### 2024-05-16 msj.tex
yufei

\documentclass[sigconf]{acmart}
% \documentclass{...}
%\input{./def/yf-formatting}
\input{./def/yf-def}

%=================================
%Yufei's stuff
%\usepackage{amsmath}
\usepackage{balance}
%\usepackage{times}
\usepackage{microtype}

\def\vgap{\vspace{0mm}}
\def\extraspacing{\vspace{1mm} \noindent}
\def\figcapup{\vspace{-2mm}}
\def\figcapdown{\vspace{-4mm}}

\def\A{\mathcal{A}}
\def\B{\mathcal{B}}
\def\C{\mathcal{C}}
\def\E{\mathcal{E}}
\def\G{\mathcal{G}}
\def\I{\mathcal{I}}
\def\II{\mathscr{I}}
\def\L{\mathcal{L}}
\def\Q{\mathcal{Q}}
\def\R{\mathcal{R}}
\def\T{\mathcal{T}}
\def\U{\mathcal{U}}
\def\V{\mathcal{V}}
\def\X{\mathcal{X}}
\def\XX{\mathscr{X}}
\def\Y{\mathcal{Y}}
\def\YY{\mathscr{Y}}
\def\Z{\mathcal{Z}}

\def\agm{\mathrm{AGM}}
\def\att{\mathbf{att}}
\def\actdom{\mathbf{actdom}}
\def\dom{\mathbf{dom}}
\def\gat{\mit{gat}}
%\def\id{\mathit{id}}
%\def\In{\mathrm{IN}}
\def\heavy{\mit{heavy}}
\def\join{\mathit{Join}}
\def\key{\mathit{key}}
%\def\light{\mathit{light}}
\def\light{\ell}
\def\load{\mathit{load}}
%\def\no{\mathit{no}}
\def\occ{\mathrm{OCC}}
\def\Out{\mathrm{OUT}}
\def\rhostar{{\rho^*}}
\def\schema{\mathit{schema}}
\def\var{\mathit{var}}

\newcommand*{\ldb}{\{\mskip-5mu\{}
\newcommand*{\rdb}{\}\mskip-5mu\}}

\allowdisplaybreaks
%=================================

%====== from acm ======
\acmDOI{}
\acmISBN{}

\acmConference[]{...}{...}{...}
\acmYear{...}
\copyrightyear{}
\acmArticle{}
\acmPrice{}
%======================

\begin{document}
%\begin{sloppy}

\title{Parallel Communication Obliviousness: One Round and Beyond}

% \author{}
% \affiliation{
% 	\institution{Chinese University of Hong Kong}
% 	\city{Hong Kong}
% 	\country{China}}

\author{}

\begin{abstract}
    This paper studies {\em communication-oblivious} algorithms under the massively parallel computation (MPC) model whose communication patterns follow a distribution that depends only on the number $p$ of machines and the input size $N$ of the underlying problem, regardless of the specific input elements. Our objective is to understand when obliviousness necessitates --- or does not necessitate --- heavier communication compared to the traditional MPC model that does not enforce such a requirement.

    \vgap

    The first part of our investigation focuses on single-round algorithms. We prove that {\em skew-free hashing}, a fundamental problem solvable with load $\tO(N/p)$ with high probability (w.h.p.) under the traditional model, demands a load of nearly $\Omega(N)$ under communication obliviousness. Intriguingly, we show that hashing can still be applied in an oblivious manner to process any natural join in one round w.h.p.\ with a load complexity matching that of the best traditional MPC algorithm. The second part of our investigation studies {\em compilation methods} that convert a traditional MPC algorithm $\A$ into a communication-oblivious counterpart. Given an $\A$ that operates within $\ell = \poly(p)$ rounds with a load at most $L = \Omega(p \log p)$, we can produce a communication-oblivious version running in $2\ell$ rounds with a load at most $(1 + \delta) L$, where $\delta > 0$ can be an arbitrarily small constant. Additionally, we establish hardness results indicating that the theoretical guarantees of our compilation can no longer be significantly improved.
\end{abstract}

\maketitle

\section{Introduction} \label{sec:intro}

In the digital era, cloud computing has emerged as a pivotal part of modern business infrastructure, reshaping the way organizations store, access, and manage data. A growing number of sectors --- ranging from healthcare and finance to academia and entertainment --- are harnessing the power of cloud services to enhance operational efficiency, effectuate cost savings, and fuel innovation. However, the migration of data from local systems to remote servers operated by third-party service providers has ushered in new privacy concerns. The crux of these concerns is the fact that sensitive data, once stored on these servers, is no longer fully under the user's control. Instead, it falls within the stewardship of service providers, thereby raising critical questions about data privacy.

\vgap

Two primary topics come to the forefront in the discourse on privacy control in cloud computing:
\myitems{
	\item {{\bf Confidential computing:}} This aims to mitigate two forms of privacy leakage that might occur on an {\em individual} server. First, rogue administrators could exploit their positions to inspect sensitive data within the server. Second, even if encryption hinders direct data accesses, administrators could still infer valuable information from the memory access patterns that the computation generates.

	\item {{\bf Oblivious communication:}} This aims to prevent another two forms of privacy leakage that could transpire during communication across {\em different} servers. First, malicious network managers might scrutinize the data flowing through the gateways. Second, even with encrypted network traffic, these managers could still deduce sensitive information from the statistics extracted from the communication.
}

Robust solutions to confidential computing are already in place, including systems that employ {\em fully homomorphic encryption} (e.g., IBM's HE4Cloud) or hardware-based {\em trusted execution environments} (nowadays offered by the leading cloud-computing platforms). When integrated with {\em oblivious RAM} techniques \cite{akl+23,go96,ln18,ppry18}, these systems safeguard sensitive information during the entire lifecycle of data: whether the data is stored in the disk, transiting through the memory hierarchy, or in use by the CPU.

\vgap

In this work, we will concentrate on oblivious communication, with the aim to understand how much communication is needed to enforce this requirement. To pave the foundation of our discussion, next we will formalize the concept of oblivious communication and the class of algorithms to be studied.

\subsection{Communication-Oblivious MPC} \label{sec:intro-prob}

Let us first clarify some math conventions in this paper. For an integer $x \ge 1$, the notation $[x]$ denotes the set $\{1, 2, ..., x\}$. Given integers $x > 0$ and $y \ge 0$, let $\binom{x}{y}$ be the number of ways of choosing $y$ elements from a set size $x$; specially, if $y > x$, define $\binom{x}{y}$ $= 0$. We use double-curly braces to represent multi-sets, e.g., $\ldb 1, 1, 1, 2, 2, 3\rdb$ is a multi-set with 6 elements. Every $\log(\cdot)$ has base $2$. The notation $\poly(n)$ denotes the class of functions polynomial in $n$.

\extraspacing {\bf The MPC Model.} Our analysis will be under the {\em massively parallel computation} (MPC) model, which has been extensively adopted in the database literature, as will be surveyed in Section~\ref{sec:intro-prev}.

\vgap

In this model, $p$ share-nothing machines are connected via a high-speed network. Initially, the input data is distributed across the $p$ machines. An algorithm then executes in {\em rounds}, each consisting of two phases: in the first phase, every machine performs computation on its local data, while in the second phase, the machines exchange messages via the network. Importantly, if machine $i \in [p]$ plans to send machine $j \in [p]$ a message $M[i,j]$ (in the second phase), it must prepare the message in the first phase. This restriction prevents machine $i$ from, for example, deciding what to send based on the messages received {\em during} the second phase. The {\em load of a round} is the maximum number of words communicated (sent and received combined) by one machine in that round. The algorithm's performance is measured by two metrics: (i) the total number of rounds executed, and (ii) the maximum load of all rounds, which is termed as the algorithm's (overall) {\em load}.

\vgap

Randomization is modeled by introducing a sequence of random bits, which is agreed upon by all machines. Such an agreement can be reached before receiving the input and therefore does not require communication during the algorithm's execution. In the first phase of a round, each machine can guide its local computation according to the random-bit sequence. When a random event is said to occur ``with high probability'' (w.h.p.), we require that the probability must be at least $1-1/p^c$ where $c$ can be an arbitrarily large constant chosen before seeing the input data.

\vgap

The value of $N$ is commonly assumed to be significantly larger than $p$, which typically means $N \ge p^c$ where $c$ is a constant dependent on the problem studied.

\extraspacing {\bf Communication-Oblivious Algorithms.} We will now formally define the notion of communication obliviousness, building upon a similar formulation in \cite{ccls20}.

\vgap

We consider that an input $\I$ to the underlying problem is a set of elements, each represented with a constant number of words. The {\em problem size} $N$ of $\I$ is defined as the set size $|\I|$. Given a specific value of $N \ge p$, let $\II(N)$ be the collection of all possible inputs with the same problem size $N$. Fix an arbitrary input $\I \in \II(N)$. As mentioned, initially the $N$ elements of $\I$ are stored across the $p$ machines. We say that a partition $\pi = (\I_1, \I_2, ..., \I_p)$ of $\I$ --- that is, $\I_1, ..., \I_p$ are mutually disjoint and their union is $\I$ --- is {\em legal} if assigning $\I_i$ to machine $i$ for each $i \in [p]$ is permitted by the underlying problem.

\vgap

Consider an arbitrary legal partition $\pi = (\I_1, \I_2, ..., \I_p)$ of $\I$ $\in$ $\II(N)$. Suppose that, when executed on $\pi$, an MPC algorithm $\A$ finishes in $\ell$ rounds where, in the $r$-th round ($r \in [\ell]$), machine $i \in [p]$ sends message $\mb{M}_r[i,j]$ to machine $j \in [p]$ (for $i = j$, $\mb{M}_r[i,j]$ is an empty message). Define $\mb{L}_r$ as the $p \times p$ matrix where $\mb{L}_r[i, j]$ equals the length of $\mb{M}_r[i,j]$, measured in the number of words. Every message $\mb{M}_r[i,j]$ is encrypted such that an adversary, who observes the network communication, can only use the sequence %\vspace{-3mm}
\myeqn{
    \sigma = (\mb{L}_1, \mb{L}_2, ..., \mb{L}_\ell) \nn
    %\label{eqn:intro:pattern}
}
to infer about $\I$. We refer to $\sigma$ as the {\em communication pattern} of $\A$ when executed on $\pi$. For a randomized algorithm $\A$, the sequence $\sigma$ may be a random variable.

\vgap

The algorithm $\A$ is {\em communication oblivious} under a specific problem size $N$ if its communication pattern $\sigma$ follows exactly the same probabilistic distribution, regardless of the choice of $\I \in \II(N)$ and the legal partition $\pi$ of $\I$. We say that $\A$ is a {\em single-round algorithm} if $\ell = 1$ for any choice of $\I$ and $\pi$; otherwise, $\A$ is a {\em multi-round algorithm}.

\subsection{Our Contributions} \label{sec:intro:ours}

Our primary objective is to discern the inherent connections between the traditional MPC model and its communication-oblivious counterpart. The core is to understand whether obliviousness would necessitate a provably higher load than the traditional, non-oblivious, MPC model. Next, we provide an overview of our findings.

\extraspacing {\bf Single-Round Hashing.} Our first contribution is a hardness result proving that {\em hashing} --- a crucial building brick deployed extensively by MPC algorithms (in both theory and practice) --- is significantly more expensive under communication obliviousness as far as single-round algorithms are concerned.

\vgap

To pinpoint the source of hardness, we introduce a problem named {\em skew-free gathering}. Fix arbitrary integers $N$ and $p$ such that $N$ is a multiple of $p$, and $p \ge t$ where $t$ is a constant integer at least 2. The input collection $\II(N)$ has only a single input $\I$: this is a set of $N$ elements, each of which --- denoted as $e$ --- is associated with an integer $\key(e)$, referred to as the {\em key} of $e$. The set of keys fulfills two conditions:
\myitems{
    \item There is one special key that is possessed by $t \cdot (N/p)$ elements, called the {\em special elements}.

    \item Every other key is possessed by exactly one element.
}
A partition $\pi = (\I_1, ..., \I_p)$ of $\I$ is {\em legal} if (i) $|\I_i| = N/p$ for all $i \in [p]$, and (ii)
the $t \cdot N/p$ special elements are placed on $t$ distinct machines, i.e., there are $t$ different machine ids $j \in [p]$ such that $\I_j$ contains $N/p$ special elements. The goal of an algorithm is to move all the special elements to one (arbitrary) machine.

\vgap

In the conventional MPC model, the problem can be solved trivially with load $O(N/p)$: simply ask the $t$ machines where the special elements are stored initially to send their data to a common machine. In contrast, we prove that if a single-round communication-oblivious algorithm is required to succeed with probability at least $2/3$, its load must be at least $\Omega(N / p^{1/t})$ with a constant probability. As $t$ increases, the load approaches $\Omega(N)$ up to a factor sub-polynomial in $p$. More precisely, there does not exist any constant $\eps > 0$ such that one can design an algorithm promising a load of $O(N/p^\eps)$ with a constant probability for all $t = O(1)$. We also show that the lower bound $\Omega(N / p^{1/t})$ can be matched by a deterministic algorithm.

\vgap

Let us now turn to {\em skew-free hashing}, where each input in $\II(N)$ is a set $\I$ of $N$ elements, each with an integer key (computed by some hash function). It is guaranteed that every possible key is possessed by $O(N/p)$ elements of $\I$ (hence, skew ``free''). Initially, each machine receives at most $\ceil{N/p}$ elements. The goal is to move all the elements of the same key to a common machine. In the traditional MPC model, the problem can be settled in one round using load $\tO(N/p)$ w.h.p.\ \cite{bks14,bks17c}, where $\tO(\cdot)$ hides a $\polylog p$ factor. As skew-free hashing generalizes skew-free gathering\footnote{Recall that skew-free gathering has the additional constraint that all the elements except for the special ones need to have distinct keys.}, our lower bound on the latter indicates that oblivious skew-free hashing demands a load of nearly $\Omega(N)$, thus creating a huge separation from the traditional model.

%It is worth pointing out that $O(N)$ is the worst possible cost under MPC and can be trivially attained --- obliviously --- by having all machines send their elements to the first machine.

\extraspacing {\bf Single-Round Joins.} The hardness of skew-free hashing under communication obliviousness is alarming because the existing one-round join algorithms \cite{assu13,au11,bks13,bks14,bks17c,kbs16} in the (conventional) MPC model depend on a method named\footnote{Also known as the {\em hyper-cube} method.} \texttt{Share} \cite{au11,bks17c}, which uses hashing as a primitive. Because the community has come very close to resolving the one-round communication complexity of (natural) joins \cite{kbs16}, it would be a huge pity if the topic had to be re-opened in the name of obliviousness.

\vgap

Our second contribution is to show that, fortunately, joins' single-round communication complexities are not affected by obliviousness! To explain, let us first formalize the {\em join problem} under the oblivious setting. Fix an arbitrary integer $N \ge 1$ as the problem size.  Let $\att$ be a finite set, where each element is called an {\em attribute}. Consider a non-empty set $\X \subseteq \att$ of attributes. A {\em tuple} over $\X$ is a function $\bm{u}: \X \rightarrow [N]$. For any non-empty subset $\Y \subseteq \X$, we define the {\em projection} of $\bm{u}$ onto $\Y$, denoted as $\bm{u}[\Y]$, as the tuple $\bm{v}$ over $\Y$ that satisfies $\bm{v}(Y) = \bm{u}(Y)$ for every attribute $Y \in \Y$. A {\em relation} $R$ is a set of tuples over the same set $\Z$ of attributes; we refer to $\Z$ as the {\em schema} of $R$ and represent it as $\schema(R)$. A {\em join} is a set $\Q$ of at least two relations such that $\sum_{R \in \Q} |R| = N$ (i.e., $N$ tuples in total in all the input relations). Define $\schema(\Q) = \bigcup_{R \in \Q} \schema(R)$. The result of $\Q$ is a relation over $\schema(\Q)$, formalized as:
\myeqn{
\join(\Q) = \set{\textrm{tuple $\bm{u}$ over $\schema(\Q)$} \mid \forall R \in \Q: \bm{u}[\schema(R)] \in R}. \nn
}
The {\em schema graph} of $\Q$ is the hypergraph $\G = (\V, \E)$ where
\myeqn{
    \V = \schema(\Q) \text{ and }
    \E = \ldb \schema(R) \mid R \in \Q \rdb. \nn
}
Note that $\E$ is a multi-set because some relations in $\Q$ may have the same schema. Focusing on ``data complexity'', we consider only joins whose schema graphs have constant sizes.

%The quantity $\max_{e \in \E} |e|$ is defined as the {\em arity} of $\G$.

\vgap

Given a hypergraph $\G = (\V, \E)$ and an integer $N \ge 1$ that is a multiple of $p$, we define the class of {\em $(\G, N)$-joins} as the set of joins that have schema graph $\G$ and problem size $N$. In the terminology of communication obliviousness, the collection $\II(N)$ comprises all the $(\G, N)$-joins of the same $\G$. Specifically, each input $\I$ of $\II(N)$ is the set of tuples in the relations of a $(\G, N)$-join $\Q$. We define a partition $\pi = (\I_1, ..., \I_p)$ of $\I$ to be {\em legal} if each $\I_i$, $i \in [p]$, consists of $N/p$ tuples. Given $\pi$, an algorithm $\A$ must output each tuple of $\join(\Q)$ on at least one machine.

\vgap

We show that any $(\G, N)$-join $\Q$ can be settled --- obliviously --- in a single round with load $\tO(N/p^{1/\psi})$ w.h.p.\ where $\psi \ge 1$ is the {\em edge quasi-packing number} \cite{kbs16} of $\G$; see Appendix~\ref{app:hypergraph} for the definition of $\psi$. The load complexity asymptotically matches that of the state-of-the-art single-round algorithm \cite{kbs16} in the traditional MPC model. Like \cite{kbs16}, our algorithm assumes that (i) $N \ge p^c$ where $c$ is a constant dependent on $\G$, and (ii) the machines are aware of certain ``heavy'' values that appear frequently in the relations of $\Q$.

\vgap

Our algorithm applies hashing in a more refined manner compared to \cite{kbs16}. The load $\tO(N/p^{1/\psi})$ stands as an intriguing contrast to our $\Omega(N)$ lower bound (up to a factor sub-polynomial in $p$) on skew-free hashing. This serves as encouraging evidence that hashing can still be powerful in designing single-round algorithms under communication obliviousness. For the join problem, our deployment of hashing circumvents the pitfall of skew-free gathering. Specifically, the deployment ensures that, for each tuple $\bm{u}$ in the join result, the $|\Q|$ tuples in $\set{\bm{u}[\schema(R)] \mid R \in \Q}$ can be transferred to a common machine (which can differ for each $\bm{u}$). Subtly, this is different from sending all tuples having an identical value on an attribute to the same machine, as required by skew-free gathering.

\extraspacing {\bf Round-Doubling Compilation.} The above discussion has focused on single-round algorithms. Our next contribution is a {\em round compilation} method for multi-round algorithms that translates a traditional MPC algorithm $\A$ into the communication-oblivious model while preserving the number of rounds of $\A$ by a factor of 2 and its load by an arbitrarily small constant factor $\delta > 0$. Specifically, if $\A$ performs at most $\ell = \poly(p)$ rounds and, w.h.p., incurs a load of at most $L$, our method yields an oblivious algorithm $\A'$ that computes the same information as $\A$ in $2\ell$ rounds with a load $(1+\delta) L$. Our compilation succeeds w.h.p.\ as long as $L = \Omega(p \log p)$, a condition satisfied by nearly all the existing MPC algorithms we are aware of.

\vgap

We also study how much improvement can still be expected over our compilation method. First, it is clear that the blow-up factor of 2 in the round number cannot be improved in general. As discussed before, for skew-free gathering, it is easy to achieve load $O(N/p)$ in one round in the traditional MPC model, but under communication obliviousness, every single-round algorithm must entail a load of nearly $\Omega(N)$ with a constant probability. Thus, the remaining question is whether the condition $L = \Omega(p \log p)$ can be significantly relaxed. Our last contribution is to answer the question in the negative by establishing a hardness result on the following {\em token passing problem}.

\vgap

Fix arbitrary integers $N$ and $p$ such that $p \ge 2$ and $N = p \cdot \ceil{p^{1-\eps}}$, where the constant $\eps$ satisfies $0 < \eps < 1$. The input collection $\II(N)$ has only a single input $\I$, which is a set of $N$ elements among which there are $N / p = \ceil{p^{1-\eps}}$ special elements called {\em tokens}. In addition to $\I$, the problem takes another two parameters: $x$ and $y$, which are distinct integers in $[p]$. A partition $\pi = (\I_1, ..., \I_p)$ of $\I$ is {\em legal} if (i) $|I_i| = N/p$ for all $i \in [p]$, and (ii) the $N/p$ tokens are all placed on machine $x$ (i.e., $I_x$ consists of nothing but tokens). The objective of the problem is to move all the tokens to machine $y$.\footnote{Unlike skew-free gathering where all the special elements can be moved to an arbitrary machine, here all the tokens must be sent to machine $y$.}

\vgap

In the traditional model, the problem can be trivially solved in one round with load $L = O(p^{1-\eps})$: simply ask machine $x$ to send all the tokens to machine $y$. We prove that, if a two-round communication oblivious algorithm is required to succeed with probability at least 2/3, its load must be $\Omega(p^{1-\eps/2})$ with a constant probability. The polynomial gap between $\Omega(p^{1-\eps/2})$ and $O(p^{1-\eps})$ indicates that the condition $L = \Omega(p \log p)$ of our round-doubling compilation  is tight up to a factor sub-polynomial in $p$. In fact, this remains true even if the blow-up factor $1+\delta$ in load preservation is increased to $\polylog p$.

\subsection{Previous Results and Relevance to Ours} \label{sec:intro-prev}

Closely related to our work is a compilation method by Chan et al.\ \cite{ccls20}. Our compilation draws inspiration from theirs but makes several improvements. First, the approach of \cite{ccls20} was designed for the ``SODA MPC'' model \cite{ksv10}, which resembles the model in Section~\ref{sec:intro-prob} but does away the notion of ``load''. In that model, each machine is equipped with $s = \Omega(N/p)$ words of memory, and the amount of communication is not a main concern as long as each machine receives no more than $s$ words in each round from all other machines combined. The analysis of \cite{ccls20} was carried out under the condition of $s = N^\eps$ for some constant $\eps > 0$. When translated into our scenario, the analysis fails to capture the regime where $L \ll s$. Second, our compilation is substantially simpler, restores the clarity for the approach underneath \cite{ccls20} in the scenario where $L = \Omega(p \log p)$, and explicitly determines the blow-up factors $1+\delta$ and 2 (for preserving the load and number of rounds, respectively)\footnote{In \cite{ccls20}, these factors were hidden in big-$O$.}. Finally, our discussion (through the token passing problem) on the condition that $L$ needs to satisfy to enable round-doubling load-preserving compilation is new.

\vgap

It should be further noted that the study in \cite{ccls20} does not address one-round MPC algorithms. Through our round compilation method, one can see that a primary distinction between the conventional MPC model and its oblivious counterpart lies in the realm of one-round algorithms. Exploring fundamental problems that can characterize this difference represents an interesting direction. Our work can be seen as a step in that direction.

\vgap

As mentioned, the existing single-round algorithms for join processing \cite{assu13,au11,bks13,bks14,bks17c,kbs16} rely crucially on the \texttt{Share} method \cite{au11,bks17c}. Given a join $\Q$, \texttt{Share} assigns a hash function $h_X(\cdot)$ to each attribute $X \in \schema(\Q)$. For each tuple $\bm{u}$ in a relation $R \in \Q$, the machine initially storing $\bm{u}$ transmits it to a set of machines determined by the hash values in $\set{h_X(\bm{u}(X)) \mid X \in \schema(R)}$. Beame et al.\ \cite{bks17c} presented a sharp concentration bound on the load of \texttt{Share} when certain skew-free requirements are satisfied. Their bound is useful to our analysis and will be reviewed in Section~\ref{sec:join}. Koutris et al.\ \cite{kbs16} proved that any one-round MPC algorithm (oblivious or not) solving the join problem must incur a load of $\Omega(N/p^{1/\psi})$ with a constant probability, if each machine is aware of only the tuples in its local storage initially.

\vgap

There is an extensive body of literature on multi-round MPC algorithms; see \cite{ajr+17,aksu15,akz19,hyt19,kst22,hcww21,hk20,h21,hy19,qt21,t18b} and the references therein. None of those algorithms was designed to achieve communication obliviousness, but all of them satisfy the condition $L = \Omega(p \log p)$ and, therefore, can be made oblivious by our compilation method. It is worth noting that our compilation also works on the one-round algorithms in \cite{assu13,au11,bks13,bks14,bks17c,kbs16}, but the algorithms after compilation will require two rounds.

\section{Hardness of Skew-Free Gathering} \label{sec:gather}

This section will discuss the skew-free gathering problem defined in Section~\ref{sec:intro-prob} and serves as a proof of the paper's first main result:

\begin{theorem} \label{thm:gathering}
    For the skew-free gathering problem parameterized by $t$, any communication-oblivious one-round algorithm succeeding with probability at least $2/3$ must entail a load of $\Omega(N/p^{1/t})$ with at least a constant probability.
    %Furthermore, this is tight: there is a communication oblivious algorithm that always succeeds and guarantees a load of $O(N/p^{1/t})$ deterministically.
\end{theorem}

In Appendix~\ref{app:sf-gathering-alg}, we show that the above lower bound $\Omega(N/p^{1/t})$ can be matched deterministically.

\vgap

To prove Theorem~\ref{thm:gathering}, let us start by identifying a set $\Pi$ of $\binom{p}{t}$ legal partitions of the (sole) input $\I \in \II(N)$ as follows. First, choose a set $Z$ of distinct integers $z_1, z_2, ..., z_t$ in $[p]$; we will refer to $Z$ as the {\em seed machine set}. Second, place $N/p$ special elements on machine $z_j$ for each $j \in [t]$, and place $N/p$ innocuous elements on each machine $i \in [p] \setminus Z$, ensuring that every  element is on exactly one machine. This defines a legal partition in $\Pi$. As there are $\binom{p}{t}$ ways to choose $Z$, the size of $\Pi$ is $\binom{p}{t}$. We require a one-round algorithm $\A$ to be oblivious only on $\Pi$, namely, its communication pattern follows the same probabilistic distribution when it is executed on $\pi \in \Pi$. We will argue that the load of $\A$ must still be $\Omega(N/p^{1/t})$ with at least a constant probability if $\A$ succeeds on every $\pi \in \Pi$ with probability at least $2/3$. This is sufficient for validating the lower bound in Theorem~\ref{thm:gathering}.

\extraspacing {\bf Conditional Expected Load.} As $\A$ executes in one-round, its communication pattern can be fully characterized by a single $p \times p$ matrix $\mb{L}$ where $\mb{L}[i,j]$ ($i, j \in [p]$) is the length of the message $M[i,j]$ that machine $i$ sends to machine $j$. The load of $\A$ can now be calculated as
\myeqn{
    \load(\mb{L}) &=&
    \max_{i=1}^p
    \Big(
    \sum_{j=1}^p \mb{L}[i, j] + \mb{L}[j, i]
    \Big).
    \label{eqn:gathering:load-matrix}
}
When $\A$ is randomized, the $M[i,j]$ of all $i, j \in [p]$ are random variables, and hence so are $\mb{L}$ and
$\load(\mb{L})$. Our objective is to prove that $\Pr[\load(\mb{L}) = \Omega(N/p^{1/t})]$ is at least a constant.

\vgap

By communication obliviousness, the distribution of $\mb{L}$ is the same for all the legal partitions in $\Pi$. Define $\L$ as the set of $p \times p$ matrices $\mb{\Lambda}$ satisfying
\myitems{
    \item $\Pr[\mb{L} = \mb{\Lambda}] > 0$, namely, $\A$ exhibits the patten $\mb{L} = \mb{\Lambda}$ with a non-zero probability;
    \item $\max_{i=1}^p \sum_{j=1}^p \mb{\Lambda}[i, j] \le N/p^{1/t}$;
    \item $\max_{i=1}^p \sum_{j=1}^p \mb{\Lambda}[j, i] \le N/p^{1/t}$.
}
Note that when $\mb{L} = \mb{\Lambda}$, {\em every} machine sends/receives at most $N/p^{1/t}$ words. The subsequent discussion will assume
\myeqn{
    \Pr[\mb{L} \in \L] \ge 1/2. \label{eqn:gathering:small-load-1/2}
}
Note that whenever $\mb{L} \notin \L$, some machine communicates at least $N/p^{1/t}$ words and hence $\load(\mb{L}) \ge N/p^{1/t}$. Thus, $\Pr[\mb{L} \notin \L] \ge 1/2$ immediately gives $\Pr[\load(\mb{L}) \ge N/p^{1/t}] \ge 1/2$, thereby verifying the claim in Theorem~\ref{thm:gathering}.

\vgap

We will aim to validate the following lower bound on the expectation of $\load(\mb{L})$ subject to the condition $\mb{L} \in \L$:
\myeqn{
    \expt[\load(\mb{L}) \mid \mb{L} \in \L] &\ge& c \cdot N/p^{1/t}
    \label{eqn:gathering:goal}
}
for some constant $0 < c < 1$. The above implies\footnote{For every $\mb{L} \in \L$, it holds that $\load(\mb{L}) \le 2 N/p^{1/t}$. If \eqref{eqn:gathering:goal-implication} incorrect, then $\expt[\load(\mb{L}) \mid \mb{L} \in \L] <  \fr{2N}{p^{1/t}} \cdot \fr{c}{4} + \fr{c}{2} \fr{N}{p^{1/t}} (1 - \fr{c}{4}) < c N / p^{1/t}$, contradicting \eqref{eqn:gathering:goal}.}
\myeqn{
    \Pr[\load(\mb{L}) \ge (c/2) \cdot N/p^{1/t} \mid \mb{L} \in \L] \ge c/4.
    \label{eqn:gathering:goal-implication}
}
Combining \eqref{eqn:gathering:small-load-1/2} and \eqref{eqn:gathering:goal-implication} yields $\Pr[\load(\mb{L}) \ge \fr{c}{2} N/p^{1/t}] \ge \fr{c}{4} \cdot \fr{1}{2} = c/8$, thus establishing Theorem~\ref{thm:gathering}.

\extraspacing {\bf Heavy Senders.} Suppose that the algorithm $\A$ exhibits a pattern $\mb{L} = \mb{\Lambda} \in \L$. Given a machine $i \in [p]$, we call machine $j \in [p] \setminus \set{i}$ a {\em heavy sender} for machine $i$ under $\mb{\Lambda}$ if $\mb{\Lambda}[j,i] \ge N/p$, namely, machine $j$ sends at least $N/p$ words to machine $i$. Define
\myeqn{
    H(\mb{\Lambda}, i) = \text{number of heavy senders for machine $i$ under $\mb{\Lambda}$}. \label{eqn:gathering:H}
}
It is easy to see that
\myeqn{
    H(\mb{\Lambda}, i) \le p^{1-1/t}
    \label{eqn:gathering:bound-H}
}
because otherwise machine $i$ would receive more than $(N/p) \cdot p^{1-1/t} = N/p^{1/t}$ words, contradicting $\mb{\Lambda} \in \L$.

\vgap

Our analysis will revolve around heavy senders. We will argue that at least one machine $i \in [p]$ satisfies
\myeqn{
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot H(\mb{\Lambda}, i)
    &=&
    \Omega(p^{1-1/t}).
    \label{eqn:gathering:goal2}
}
As $\load(\mb{\Lambda}) \ge (N/p) \cdot H(\mb{\Lambda}, i)$, the above leads to
\myeqn{
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \load(\mb{\Lambda})
    =
    \Omega\Big(\fr{N}{p} \cdot p^{1-1/t} \Big)
    =
    \Omega(N /p^{1/t}).
    \label{eqn:gathering:goal2-implication}
}
This will prove \eqref{eqn:gathering:goal} because
\myeqn{
    \expt[\load(\mb{L}) \mid \mb{L} \in \L]
    &=&
    \fr{1}{\Pr[\mb{L} \in \L]} \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \load(\mb{\Lambda})
    \nn
}
which is at least \eqref{eqn:gathering:goal2-implication}.

%as long as $H(\mb{\Lambda}, i) \ge 1$

\extraspacing {\bf Two Types of Legal Partitions.} The rest of the section will focus on proving the correctness of \eqref{eqn:gathering:goal2}. Earlier we have identified a set $\Pi$ of $\binom{p}{t}$ legal partitions, each characterized by a seed machine set $Z = \set{z_1, ..., z_t}$; for convenience, we will use $Z$ to denote the corresponding legal partition when no ambiguity can arise. Now, fix any legal partition $Z$. The algorithm $\A$ succeeds on $Z$ if and only if at least one of the following occurs:
\myitems{
    \item machines $z_1, z_2, ..., z_t$ all send their special elements to a machine $k \in [p] \setminus Z$;
    \item there is some $j \in [t]$ such that machine $z_j$ receives all the special elements from every other machine in $Z$.
}
Echoing the above, given any $Z$ and any matrix $\mb{\Lambda} \in \L$, we define
\myitems{
    \item $X(Z, \mb{\Lambda})$ as an indicator variable, which equals 1 if there exists some $k \in [p] \setminus Z$ such that $\mb{\Lambda}[z_j, k] \ge N/p$ for every $j \in [t]$, or 0 otherwise;

    \item $Y(Z, \mb{\Lambda})$ as an indicator variable, which equals 1 if there exists some $j \in [t]$ such that $\mb{\Lambda}[z_k, z_j] \ge N/p$ for every $k \in [t] \setminus \set{j}$, or 0 otherwise.
}

%

When exhibiting the pattern $\mb{L} = \bm{\Lambda}$, the algorithm $\A$ succeeds on $Z$ only if
\myeqn{
    X(Z, \mb{\Lambda}) + Y(Z, \mb{\Lambda}) \ge 1. \nn
}
We declare:
\myeqn{
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
    (X(Z, \mb{\Lambda}) + Y(Z, \mb{\Lambda}))
    &\ge&
    1/6.
    \label{eqn:gathering:1/6}
}
To see why, notice that if the above did not hold, then $\A$ would have probability less than $1/6$ succeeding on $Z$ when $\mb{L} \in \L$. Even if $\A$ always succeeds on $Z$ when $\mb{L} \notin \L$, the overall success probability on $Z$ would still be less than $1/6 + \Pr[\mb{L} \notin \L]$, which is less than 2/3 because of \eqref{eqn:gathering:small-load-1/2}. This violates the requirement that $\A$ must succeed with probability at least $2/3$ on $Z$.

\vgap

We will refer to a legal partition $Z$ as
\myitems{
    \item an {\em X-partition} if $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
    X(Z, \mb{\Lambda}) \ge 1/12$, or
    \item an {\em Y-partition} if $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
    Y(Z, \mb{\Lambda}) \ge 1/12$.
}
Note that $Z$ must be classified as at least one of the two types because of \eqref{eqn:gathering:1/6}. Thus, if $\XX$ (resp., $\YY$) represents the set of all X- (resp., Y-) partitions, we have $|\XX| + |\YY| \ge |\Pi| = \binom{p}{t}$. The next lemma relates each type of partitions to heavy-sender machines.

\begin{lemma} \label{lmm:gathering:total-X-Y}
    For any $\mb{\Lambda} \in \L$, it holds that
    \myeqn{
        \sum_{Z \in \XX} X(Z, \mb{\Lambda}) &\le& \sum_{i \in [p]} \binom{H(\mb{\Lambda}, i)}{t} \label{eqn:gathering:total-X-Y:help1} \\
        \sum_{Z \in \YY} Y(Z, \mb{\Lambda}) &\le& \sum_{i \in [p]} \binom{H(\mb{\Lambda}, i)}{t-1}. \label{eqn:gathering:total-X-Y:help2}
    }
\end{lemma}

The reader is reminded that $\binom{x}{y}$ equals 0 if $x < y$. We will prove the first inequality here and the second one in Appendix~\ref{app:proof:thm:gathering}.

\begin{proof}[Proof of \eqref{eqn:gathering:total-X-Y:help1}]
    Our proof adopts a counting argument. Initially, create an empty set $S_i$ for each $i \in [p]$. Process each $Z = \set{z_1, ..., z_t} \in \XX$ as follows. If $X(Z, \mb{\Lambda}) = 0$, do nothing. Otherwise, under the communication pattern $\mb{L} = \bm{\Lambda}$, there is at least one machine $k \in [p] \setminus Z$ that receives $N/p$ special elements from machine $z_j$ for all $j \in [t]$. Thus, machines $z_1, z_2, ..., z_t$ are heavy senders for machine $k$ under $\mb{\Lambda}$. Hence, $Z$ represents a possible way to choose $t$ machines from the $H(\mb{\Lambda}, k)$ heavy senders for machine $k$ under $\mb{\Lambda}$. Accordingly, we add $Z$ to $S_k$. After all the sets $Z \in \XX$ have been processed, the left hand side of \eqref{eqn:gathering:total-X-Y:help1} is bounded by $\sum_{i \in [p]} |S_i|$. On the other hand, for each $i \in [p]$, the size $|S_i|$ cannot exceed the total number of ways of choosing $t$ machines from the $H(\mb{\Lambda}, i)$ heavy senders for machine $i$ under $\mb{\Lambda}$. Inequality \eqref{eqn:gathering:total-X-Y:help1} now follows.
\end{proof}

\vgap

The remainder of the argument proceeds differently depending on the number of $X$-partitions.

\extraspacing {\bf At Least $\fr{1}{2} \binom{p}{t}$ X-Partitions.} That is, $|\XX| \ge \fr{1}{2} \binom{p}{t}$, which, together with the definition of X-partition, yields
\myeqn{
    \sum_{Z \in \XX} \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
    X(Z, \mb{\Lambda}) \ge \fr{1}{24} \binom{p}{t}.
    \label{eqn:gathering:X-help1}
}
We can now derive
\myeqn{
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} H(\mb{\Lambda},i)^t
    &\ge&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \binom{H(\mb{\Lambda},i)}{t}
    \nn \\
    \explain{by \eqref{eqn:gathering:total-X-Y:help1}}
    &\ge&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{Z \in \XX} X(Z, \mb{\Lambda})
    \nn \\
    &=&
    \sum_{\mb{\Lambda} \in \L} \sum_{Z \in \XX} \Pr[\mb{L} = \mb{\Lambda}] \cdot  X(Z, \mb{\Lambda})
    \nn \\
    \explain{by \eqref{eqn:gathering:X-help1}}
    &\ge&
    \fr{1}{24} \binom{p}{t}.
    \label{eqn:gathering:X-help2}
}
This gives:
\myeqn{
    %\text{LHS of \eqref{eqn:gathering:goal2}}
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \sum_{i \in [p]} H(\mb{\Lambda},i)
    &=&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^t}{H(\mb{\Lambda},i)^{t-1}}
    \nn \\
    \explain{by \eqref{eqn:gathering:bound-H}}
    &\ge&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^t}{p^{\fr{(t-1)^2}{t}}}
    \nn \\
    \explain{by \eqref{eqn:gathering:X-help2}}
    &\ge&
    \fr{1}{24} \binom{p}{t} / p^{\fr{(t-1)^2}{t}}
    \nn \\
    &=&
    \Omega\Big(
        \fr{p^t}{p^{(t-1)^2/t}}
    \Big)
    =
    \Omega\Big(p^{2-1/t} \Big).
    \nn
}
Therefore, at least one $i \in [p]$ satisfies $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot H(\mb{\Lambda},i) = \Omega(p^{2-1/t} / p ) = \Omega (p^{1-1/t})$, as claimed in \eqref{eqn:gathering:goal2}.

\extraspacing {\bf At most $\fr{1}{2} \binom{p}{t}$ X-Partitions.} Because $|\XX| + |\YY| \ge \binom{p}{t}$, the fact $|\XX| \le \binom{p}{t}/2$ indicates $|\YY| \ge \binom{p}{t}/2$. Following a derivation similar to the previous case, we can show that at least one $i \in [p]$ satisfies \eqref{eqn:gathering:goal2}. With the details presented in Appendix~\ref{app:proof:thm:gathering}, we now conclude the proof of Theorem~\ref{thm:gathering}.

\section{Single-Round Oblivious Joins} \label{sec:join}

We now turn our attention to the join problem defined in Section~\ref{sec:intro:ours}. Let $\Q$ be the input $(\G, N)$-join. Given a value $x \in [N]$, we call it
\myitems{
    \item {\em heavy} if at least $N / p^2$ tuples of a relation have value $x$ under an attribute, or formally, $\exists R \in \Q, X \in \schema(R)$ s.t. $R$ has at least $N / p^2$ tuples $\bm{u}$ with $\bm{u}(X) = x$;

    \item or {\em light}, otherwise.
} 
There can be $O(p^2)$ heavy values in total. We make the {\em heavy-statistics assumption} that each machine knows all the heavy values. For this purpose, each legal partition (of the tuples in the relations of $\Q$; see Section~\ref{sec:intro:ours}) is augmented with an ``information field'' for each machine, which contains all the heavy values. To prevent an adversary from distinguishing the $(\G, N)$ joins by inspecting the information field's length, the field is always padded with dummy data to the same length of $O(p^2)$ words.

\vgap

We will establish our second main result:

\begin{theorem} \label{thm:join}
    Consider any constant-size hypergraph $\G = (\V, \E)$. Let $N$ be a multiple of $p$ satisfying $N \ge p^{1 + 2\alpha + 1/\psi}$ where $\alpha = \max_{e \in \E} |e|$ (called the arity of $\G$) and $\psi$ is the edge quasi-packing number of $\G$. Under the heavy-statistics assumption, there is a communication oblivious algorithm that,  given any $(\G, N)$-join $\Q$, operates in a single round with load $\tO(N / p^{1/\psi})$ and computes $\join(\Q)$ w.h.p..
\end{theorem}

Our result can be compared to an algorithm of \cite{kbs16}, which has load $\tO(N / p^{1/\psi})$ w.h.p.\ but is not communication oblivious. That algorithm also makes the heavy-statistics assumption although their frequency threshold of a heavy value is $N/p$, rather than $N/p^2$ in our context.

%$\alpha = \max_{e \in \E} |e|$ (this is the arity of $\G$), and

\vgap

The rest of the section serves as a proof of Theorem~\ref{thm:join}. We will first review the \ttt{Share} method \cite{au11,bks17c} in Section~\ref{sec:join-share}. After that, we will present our join algorithm in Sections~\ref{sec:join-alg1} and \ref{sec:join-alg2} to attain the performance guarantees claimed.

%and analyzing its behavior in Section~\ref{sec:join-analysis}.

\subsection{The Share Method} \label{sec:join-share}

Let $R^*$ be a relation with $\schema(R^*) = \set{X_1, X_2, ..., X_r}$, and $M$ be an integer satisfying $|R^*| \le M \le N$. Recall that, as defined in Section~\ref{sec:intro:ours}, each tuple $\bm{u} \in R^*$ is a function from $\schema(R^*)$ to $[N]$. For any $\Y \subseteq \schema(R^*)$, define:
%{\em the degree of $R^*$ on $\Y$} as
\myeqn{
    \deg_\Y(R^*) &=& \max_{\text{tuple $\bm{u}$ over $\Y$}} \big|\set{\bm{v} \in R^* \mid \bm{v}[\Y] = \bm{u}} \big| \label{eqn:join-degree}
}
that is, the maximum number of tuples in $R^*$ having the same projection on $\Y$. Suppose that we assign to each attribute $X_i$, $i \in [r]$, a positive integer $s_i$ --- called the {\em share} of $X_i$ --- satisfying the following {\em skew-free condition}:
\myeqn{
    \text{for any non-empty $\Y \subseteq \schema(R^*)$, $\deg_\Y(R^*) \le \fr{M}{{\prod_{i \in [r]: X_i \in \Y} s_i}}$}.
    \label{eqn:share-skewfree}
}
Fix any constant $\alpha \ge r$. For each $i \in [r]$, choose independently a {\em perfectly random hash function} $h_i: [\alpha N] \rightarrow [s_i]$, that is, $h_i$ is uniformly sampled from the $(s_i)^{\alpha N}$ possible functions from $[\alpha N]$ to $[s_i]$. Given $\bm{b} = (b_1, b_2, ..., b_r)$ $\in [s_1] \times [s_2] \times ... \times [s_r]$, define
\myeqn{
    \text{\em bin $\bm{b}$}
    &=&
    \set{\bm{u} \in R^* \mid \forall i \in [r], h_i(\bm{u}(X_i)) = b_i}.
    \label{eqn:join-share:bin}
}
Note that there are in total $\prod_{i=1}^r s_i$ bins. The following concentration bound is due to Beame et al.\ \cite{bks17c}:

\begin{lemma} \label{lmm:share}
    %[\cite{bks17c}]
    For any constant $c > 0$, it holds with probability at least $1 - (1/p^*)^c$ that the sizes of all the bins are bounded by $O((\log p^* / \log\log p^*)^r \cdot M / p^*)$, where $p^* = \prod_{i=1}^r s_i$. The constant in the big-$O$ depends on $c$.
\end{lemma}

Strictly speaking, Beame et al.\ \cite{bks17c} proved the lemma only for the case where $M = |R^*|$, which is not enough for our purposes. Fortunately, it is not difficult to extend their result to any $M \in [|R^*|, N]$, as shown in Appendix~\ref{app:proof:thm:join}.

\subsection{A Non-Oblivious Join Algorithm} \label{sec:join-alg1}

This subsection will present an algorithm, which is {\em not} communication oblivious, to compute the result of a $(\G, N)$-join $\Q$ with load $\tO(N / p^{1/\psi})$ w.h.p.. The reason for describing this non-oblivious version is to allow the reader to draw a direct comparison with the solution of \cite{kbs16}. As will be clear, a key new idea is to apply the concentration bound of Lemma~\ref{lmm:share} locally on each machine, rather than globally on the entire join (as was done in \cite{kbs16}). This in turn requires us to decrease the ``heavy threshold'' from $N/p$ to $N/p^2$. Our algorithm ensures a new property (to be given in Lemma~\ref{lmm:join-comm}), which makes it easy to make the algorithm communication oblivious, as we do in the next subsection.

\vgap

Given a relation $R \in \Q$ and machine id $i \in [p]$, let $R^{(i)}$ be the set of tuples of $R$ that are initially stored on machine $i$. Define $R^{(i,\heavy)}$ as the set of tuples $\bm{u} \in R^{(i)}$ using {\em only} heavy values. We know
\myeqn{
    |R^{(i,\heavy)}| &=& O(p^{2\alpha}) \nn
    %\label{eqn:join-size-R-i-heavy}
}
because $R^{(i)}$ has at most $\alpha$ attributes (recall that $\alpha$ is the arity of the hypergraph $\G$), and there can be $O(p^2)$ heavy values.

\vgap

Before proceeding, the reader should be familiarized with the content of Appendix~\ref{app:hypergraph}, in particular, the notions of ``residual graph'' and ``fractional vertex cover''. Next, let us take an arbitrary proper subset $\Z \subset \V$ and identify an optimal fractional vertex cover $w_\Z:$ $\V$ $\setminus$ $\Z \rightarrow [0, 1]$ of the residual graph $\G_\Z = (\V_\Z, \E_\Z)$. Define
\myeqn{
    \tau(\G_\Z) &=& \sum_{X \in \V \setminus \Z} w_\Z(X).
    \label{eqn:join-tau_Z}
}
To each attribute $X \in \V$, we assign a positive integer $s_X(\Z)$ --- the share of $X$ for $G_\Z$ --- as follows:
\myitems{
    \item if $X \in \Z$, then $s_X(\Z) = 1$;
    \item otherwise, $s_X(\Z) = \lf p^{w_\Z(X) / \tau(\G_\Z)} \rf$.
}
For any non-empty subset $\Y \subseteq \V$, we show in Appendix~\ref{app:proof:thm:join}:
\myeqn{
    \prod_{X \in \Y} s_X(\Z)
    \le
    p.
    \label{eqn:join-help1}
}
Furthermore, for any relation $R \in \Q$, we show again in Appendix~\ref{app:proof:thm:join}:
\myeqn{
    \prod_{X \in \schema(R)} s_X(\Z) &=& \Omega\Big(p^{1/\tau(G_\Z)} \Big).
    \label{eqn:join-help2}
}

Henceforth, set
\myeqn{
    M = N/p. \label{eqn:join-M}
}
For each $i \in [p]$ and each relation $R \in \Q$ satisfying $\schema(R) \setminus \Z \ne \emptyset$, we define $R^{(i,\Z)}$ as the set of tuples $\bm{u} \in R^{(i)}$ such that
\myitems{
    \item $\bm{u}(X)$ is heavy for every attribute $X \in \schema(R) \cap \Z$;
    \item $\bm{u}(X)$ is light for every attribute $X \notin \schema(R) \cap \Z$.
}
For any non-empty subset $\Y \subseteq \schema(R)$, let us observe
\myeqn{
    \deg_\Y (R^{(i,\Z)})
    \le \fr{M}{\prod_{X \in \Y} s_X(\Z)}.
    \label{eqn:join-alg-skewfree}
}
Indeed, if $\Y \subseteq \Z$, then $\prod_{X \in \Y} s_X(\Z) = 1$ in which case \eqref{eqn:join-alg-skewfree} holds because $\deg_\Y (R^{(i,\Z)}) \le |R^{(i)}| \le N/p = M$. Otherwise, identify an arbitrary attribute $X \in \Y \setminus \Z$. Then, we can derive
\myeqn{
    \deg_\Y (R^{(i,\Z)}) \le \deg_{\set{X}} (R^{(i,\Z)}) \le \deg_{\set{X}} (R^{(i)}) \le N/p^2 \nn
}
where the last inequality is due to the definition of light value. By \eqref{eqn:join-help1}, $N/p^2 \le M / \prod_{X \in \Y} s_X(\Z)$, from which the correctness of \eqref{eqn:join-alg-skewfree} follows. We can therefore conclude that $R^{(i,\Z)}$ fulfills the skew-free condition prescribed in \eqref{eqn:share-skewfree}.

\vgap

We now elaborate on the join algorithm. Before receiving the $(\G,N)$-join $\Q$, we perform some preprocessing for each proper subset $\Z \subset \V$. First, obtain $s_X(\Z)$ --- the share of $X$ for $G_\Z$ --- for every attribute $X \in \V$ as explained earlier. For each $X \in \V$, independently pick a perfectly random hash function $h_{\Z, X}: [\alpha N] \rightarrow [s_X(\Z)]$ (recall that $\alpha$ is the arity of $\G$). Store a copy of all these functions on every machine. The cartesian product of $\bigtimes_{X \in \V} [s_X(\Z)]$ has a size of $\prod_{X \in \V} s_X(\Z)$, which is at most $p$ (see \eqref{eqn:join-help1}). Each element in the cartesian product can be regarded as a tuple $\bm{b}_\Z$, where $\bm{b}_\Z(X)$ is an integer in $[s_X(\Z)]$ for each $X \in \V$. We assign a distinct machine to each possible $\bm{b}_\Z$ lexicographically: machine $i \in [p]$ assigned to the $\bm{b}_\Z$ that ranks, in lexicographic order, the $i$-th among all the elements of $\bigtimes_{X \in \V} [s_X(\Z)]$. Note that some machines (precisely, $p - \prod_{X \in \V} s_X(\Z)$ machines) are not assigned to any cartesian product elements.

\vgap

After the $(\G,N)$-join $\Q$ has been given, each machine --- say the one with id $i \in [p]$ --- carries out the following steps in parallel:

\myitems{
    \item {{\bf S1:}} Broadcast $R^{(i,\heavy)}$.

    \vgap

    \item {{\bf S2:}} For each proper subset $\Z \subset \V$ and every relation $R \in \Q$ satisfying $\schema(R) \setminus \Z \ne \emptyset$, the machine examines every tuple $\bm{u} \in R^{(i,\Z)}$. For each attribute $X \in \schema(R)$, the machine computes the hash value $h_{\Z,X}(\bm{u}(X))$. Then, it sends $\bm{u}$ to every machine $\bm{b}_\Z \in \bigtimes_{X \in \V} [s_X(\Z)]$ satisfying $\bm{b}_\Z(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \schema(R)$.
    %The number of such machines is $\prod_{X \in \V \setminus \schema(R)} s_X(\Z)$, which is $O(p^{1-1/\tau(\G_\Z)})$ by \eqref{eqn:join-help3}.
}

The above steps require one round of communication. Then, every machine joins all the tuples received. To prove  correctness, consider any tuple $\bm{u} \in \join(\Q)$. If $\bm{u}$ uses a heavy value on every attribute, then it is produced by all machines due to Step {\bf S1}. Otherwise, suppose that $\bm{u}$ uses heavy values only on those attributes of $\Z$ for some $\Z \subset \V$. Then, due to Step {\bf S2}, it must be produced at the machine corresponding to the element $\bm{b}_\Z \in \bigtimes_{X \in \V} [s_X(\Z)]$ where $\bm{b}_\Z(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \V$.

\vgap

We will prove that the load incurred is $O(p^{2\alpha + 1}) + \tO(N/p^{1/\psi})$ w.h.p., which is $\tO(N/p^{1/\psi})$ for $N \ge p^{1 + 2\alpha + 1/\psi}$. Clearly, Step {\bf S1} entails load $O(p^{2\alpha + 1})$. Next, we will show that, for any particular $\Z \subset \V$, {\bf S2} generates a load of $\tO(N / p^{1/\tau(\G_\Z)})$ w.h.p., where $\tau(\G_\Z)$ is given in \eqref{eqn:join-tau_Z}. It will then follow that the overall load of {\bf S2} is $\tO(N/p^{1/\psi})$ because $\psi$ equals the maximum $\tau(\G_\Z)$ of all $\Z \subset \V$ (see \eqref{eqn:hypergraph:psi}) Appendix~\ref{app:hypergraph}) and $\V$ has a constant number of subsets.

\vgap

The lemma below characterizes the behavior of our algorithm.

\begin{lemma} \label{lmm:join-comm}
    Consider an arbitrary proper subset $\Z \subset \V$. The following statement holds w.h.p.: for any relation $R \in \Q$ and distinct machine ids $i, j \in [p]$, machine $i$ sends $\tO(M / p^{1/\tau(\G_\Z)})$ tuples in $R^{(i,\Z)}$ to machine $j$.
\end{lemma}

\begin{proof}
     As explained previously, relation $R^{(i,\Z)}$ satisfies the skew-free condition \eqref{eqn:join-alg-skewfree} with respect to the shares $\set{s_X(\Z) \mid X \in \V}$. Set $p^* = \prod_{X \in \schema(R)} s_X(\Z)$. As $|R^{(i,\Z)}| \le M \le N$, Lemma~\ref{lmm:share} asserts that the following event occurs w.h.p.:

     \minipg{0.9\linewidth}{
        $\tO(\fr{M}{p^*})$ tuples $\bm{u} \in R^{(i,\Z)}$ are hashed to each element (a.k.a., bin) $\bm{b}$ in the cartesian product $\bigtimes_{X \in \schema(R)} [s_X(\Z)]$, namely, $\bm{b}(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \schema(R)$.
     }

     \noindent Under the above event, $\tO(M/p^*)$ tuples $\bm{u} \in R^{(i,\Z)}$ are hashed to each element $\bm{b}_\Z \in \bigtimes_{X \in \V} [s_X(\Z)]$, i.e., such a tuple $\bm{u}$ satisfies $\bm{b}_\Z(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \schema(R)$. Hence, if the machine $j$ stated in the lemma is assigned to an element in  $\bigtimes_{X \in \V} [s_X(\Z)]$, then machine $i$ sends $\tO(M/p^*)$ tuples of $R^{(i,\Z)}$ to machine $j$; otherwise, machine $i$ sends no tuple of $R^{(i,\Z)}$ to machine $j$.

     \vgap

     The lemma now follows from $p^* = \Omega(p^{1/\tau(\G_\Z)})$ (see \eqref{eqn:join-help2}).
\end{proof}

As $\Q$ has $O(1)$ relations, Lemma~\ref{lmm:join-comm} indicates that every machine sends and receives $\tO(p \cdot M / p^{1/\tau(\G_\Z)}) = \tO(N / p^{1/{\tau(\G_\Z)}})$ words in Step {\bf S2} w.h.p.\ in processing the subset $\Z$.

\extraspacing {\bf Remark.} Lemma~\ref{lmm:join-comm} is a feature of our algorithm that is not shared by the algorithm of \cite{kbs16}, but is crucial for turning the algorithm into an oblivious counterpart, as discussed next.

\subsection{Making the Algorithm Oblivious} \label{sec:join-alg2}

Our algorithm in Section~\ref{sec:join-alg1} enjoys a simple communication pattern. It can be modified into a communication-oblivious version by padding enough dummy data to make the length of each message consistent with the ``worst'' case.

\vgap

Let us start with Step {\bf S1}, where machine $i \in [p]$ transmits the same message, which contains $R^{(i,\heavy)}$ for each $R \in \Q$, to every other machine. This message can have a maximum length of $\mit{len}_\mit{heavy}$ $= O(p^{2\alpha})$ words. Whenever this message is shorter than $\mit{len}_\mit{heavy}$, expand it with dummy data into length $\mit{len}_\mit{heavy}$.

\vgap

In Step {\bf S2}, for each proper subset $Z \subset \V$, machine $i \in [p]$ transmits a possibly different message to each other machine $j \in [p]$, whose length is bounded by an integer $\mit{len}_\Z = \tO(M/p^{1/\tau(\G_\Z)})$ w.h.p.\ (Lemma~\ref{lmm:join-comm}). Whenever this message is shorter than $\mit{len}_\Z$ --- including the case of an empty message --- expand it with dummy data into length $\mit{len}_\Z$. However, it is possible for the message to be actually longer than $\mit{len}_\Z$ (this happens when the high-probability event in Lemma~\ref{lmm:join-comm} does not occur). In that scenario, machine $i$ (arbitrarily) trims the message at the length of $\mit{len}_\Z$ and sends the trimmed message to machine $j$ anyway; when this happens, we say that the algorithm {\em errs}.

\vgap

In the above modified algorithm, every machine sends precisely $\mit{len}_\mit{heavy} + \sum_{\Z \subset \V} \mit{len}_\Z$ words to every other machine. It therefore exhibits a deterministic communication pattern for all $(\G,N)$-joins --- this is true regardless of whether the algorithm errs. The load is always $p \cdot (\mit{len}_\mit{heavy} + \sum_{\Z \subset \V} \mit{len}_\Z)$, which is $\tO(N/p^{1/\psi})$ as analyzed in the previous subsection. The algorithm computes the join result correctly if it does not err (an adversary cannot tell whether the algorithm has erred because everything is encrypted). Due to Lemma~\ref{lmm:join-comm}, by choosing constants appropriately, we can reduce the probability that the algorithm errs to at most $1/p^c$ for an arbitrarily large constant $c$. This completes the proof of Theorem~\ref{thm:join}.

\section{Two-Round Compilation} \label{sec:compile}

This section will concentrate on multi-round MPC algorithms, for which our main contribution is a compilation method that establishes the following theorem:

\begin{theorem} \label{thm:compile}
    Fix an arbitrary constant $\delta$ satisfying $0 < \delta < 1$. Let $\A$ be an algorithm under the traditional MPC model that operates within $\ell = \poly(p)$ rounds and entails a load at most $L$ w.h.p.. If $L  = \Omega(p \log p)$ where the hidden constant depends on $\delta$, there is a communication-oblivious algorithm that performs $2\ell$ rounds, requires a load at most $(1 + \delta) L$, and computes the same information as $\A$ on every machine w.h.p..
\end{theorem}

As explained in Section~\ref{sec:intro:ours}, the round blow-up factor 2 in the above theorem is the best possible, regardless of the constant $\delta$. The next theorem indicates that the condition $L = \Omega(p \log p)$ can no longer be relaxed significantly, again as explained in Section~\ref{sec:intro:ours}.

\begin{theorem} \label{thm:token}
    Fix an arbitrary constant $\eps$ satisfying $0 < \eps < 1$. For the token-passing problem (defined in Section~\ref{sec:intro:ours}) parameterized by $\eps$, any communication-oblivious two-round algorithm succeeding with probability at least $2/3$ must demand a load of $\Omega(p^{1-\eps/2})$ with at least a constant probability.
\end{theorem}

We will present the algorithmic procedure of our compilation method in the rest of the section, but defer its analysis to Appendix~\ref{app:compile}. The proof of Theorem~\ref{thm:token} is provided in Appendix~\ref{app:thm:token}.

\extraspacing {\bf Message Routing.} To prove Theorem~\ref{thm:compile}, we will tackle the {\em message routing problem} defined as follows. Suppose that, for each pair of $i, j \in [p]$, machine $i$ needs to send a message $\mb{M}[i,j]$ to machine $j$ (if $i = j$, then $\mb{M}[i,j]$ is an empty message). Denote by $\mb{L}[i,j]$ the length of $\mb{M}[i,j]$ in the number of words. For each $i \in [p]$, we have
\myeqn{
    \sum_{j \in [p]} \mb{L}[i,j] + \mb{L}[j,i] \le L \label{eqn:compile-len-constraint}
}
namely, every machine sends and receives no more than $L$ words in total. We want to design a communication-oblivious algorithm $\A_\mit{route}$ with all the requirements below:
\myitems{
    \item It runs in two rounds.
    \item For distinct machines $i, j \in [p]$, machine $i$ sends (precisely) $L'$ words to machine $j$ in each round, where $L' \le (1+\delta) L/p$.
    \item W.h.p., all the messages are successfully delivered at the end of the second round.
}

Equipped with $\A_\mit{route}$, we can prove Theorem~\ref{thm:compile} as follows. Recall (from Section~\ref{sec:intro-prob}) that each round of $\A$ runs in two phases where in the first phase each machine prepares the messages to be sent out in the second phase. We treat the second phase as an instance of the message routing problem and apply $\A_\mit{route}$ to deliver the messages. W.h.p., at the end of $\A_\mit{route}$, every machine acquires all the information that it would have obtained at the second phase of $\A$, and can thus perform the local computation as demanded by $\A$. Because $\A$ always terminates within $\ell = \poly(p)$ rounds, our compilation succeeds on all its rounds w.h.p..

\extraspacing {\bf Our Algorithm.} Set $\delta' = \ceil{16/\delta}$. We chop each message $\mb{M}[i,j]$ into segments, each of which has $\delta'$ words --- if the last segment is shorter than $\delta'$ words, expand it with dummy words to make its length $\delta'$. The number of segments is $\ceil{\mb{L}[i,j] / \delta'}$. Each segment, referred to as a {\em packet} henceforth, is augmented with three fields: (i) {\em sender}, i.e., the value $i$ (sender machine id), (ii) {\em recipient}, i.e., the value $j$ (recipient machine id), and (iii) {\em sequence number}, i.e., the $t$-th packet of $\mb{M}[i,j]$ has sequence number $t$. Packets may arrive at machine $j$ in an arbitrary order; the sender and sequence-number fields allow the packets to be put back into the original sequence in $\mb{M}[i,j]$. Every packet is thus $\delta' + 3$ words in length.

\vgap

Denote by $P_i$ the set of packets obtained from $\mb{M}[i,1]$, $\mb{M}[i,2]$, ..., $\mb{M}[i,p]$ combined. Let $|P_i|$ be the number of those packets. Then
\myeqn{
    |P_i| \le \sum_{j=1}^p \Big(1 + \fr{\mb{L}[i,j]}{\delta'} \Big) \le p + \fr{L}{\delta'}.
    \label{eqn:compile-P_i-size}
}
where the last inequality used \eqref{eqn:compile-len-constraint}. Set
\myeqn{
    \lambda = \lc \left(1 + \fr{L} {\delta' p}\right)\left(1 + \fr{\delta}{2}\right) \rc.
    \label{eqn:compile-lambda}
}

In the first round, each machine --- say machine $i \in [p]$ --- in parallel carries out the steps below:
\myitems{
    \item To each packet in $P_i$, assign a {\em relay machine} chosen from $[p]$ uniformly at random. Let $P_i(j)$ be the set of packets in $P_i$ assigned to the same relay machine $j \in [p]$. We use $|P_i(j)|$ to represent the number of packets in $P_i(j)$.

    \vgap

    \item For each $j \in [p]$, if $|P_i(j)| < \lambda$, add enough dummy packets to $P_i(j)$ to make the number of packets therein exactly $\lambda$. If $|P_i(j)| > \lambda$, discard (arbitrarily) some packets from $P_i(j)$ to shrink $|P_i(j)|$ to $\lambda$; however, in this case, we say that the algorithm {\em errs}.

    \vgap

    \item For each $j \in [p]$, send $P_i(j)$ to machine $j$ (of course, for $j = i$, ``sending'' $P_i(j)$ requires no communication).
}

For each $i \in [p]$, define
\myeqn{
    P^*_i = \bigcup_{j \in [p]} P_j(i) \nn
}
namely, the set of all packets to be ``relayed'' by machine $i$. These packets are transmitted to machine $i$ in the first round. Machine $i$ sorts the packets by recipient field. For each $j \in [p]$, let $P^*_i(j)$ be the set of packets in $P^*_i$ whose recipient fields are $j$. In the second round, machine $i$, in parallel to others, proceeds as follows:
\myitems{
    \item For each $j \in [p]$, if $|P^*_i(j)| < \lambda$, add dummy packets to $P^*_i(j)$ to increase $|P^*_i(j)|$ to $\lambda$. If $|P^*_i(j)| > \lambda$, discard (arbitrarily) some packets from $P^*_i(j)$ to shrink $|P^*_i(j)|$ to $\lambda$; in this case, we say that the algorithm {\em errs}.

    \vgap

    \item For each $j \in [p]$, send $P^*_i(j)$ to machine $j$.
}
This completes our 2-round algorithm for message routing.

\vgap

In each round, for any distinct $i, j \in [p]$, machine $i$ sends to machine $j$ precisely
\myeqn{
    L' = \lambda \cdot (\delta' + 3) \label{eqn:compile-L'}
}
words. It is rudimentary to verify that $L' \le (1+\delta) L /p$ when $p$ is greater than a constant. Therefore, our 2-round algorithm exhibits a deterministic communication pattern and has a load of $L' (p-1) < (1+\delta) L$. 
If the algorithm does not err, every message $\mb{M}[i,j]$ is successfully delivered to machine $j \in [p]$. In Appendix~\ref{app:compile}, we prove that the algorithm errs with probability at most $1/p^c$ where $c$ can be an arbitrarily large constant chosen before running the algorithm. This completes the proof of Theorem~\ref{thm:compile}.

\bibliographystyle{plainurl}% the mandatory bibstyle
\bibliography{ref}

\appendix

\section*{Appendix}

\section{Basic Concepts of Hypergraphs} \label{app:hypergraph}

Let $\G = (\V, \E)$ be a hypergraph with $\E$ being a multi-set where each element, called an {\em edge}, is a non-empty subset of $\V$ (the set of {\em vertices}). A {\em fractional edge packing} is a function $w: \E \rightarrow [0, 1]$ (mapping each edge $e \in \E$ to a weight $w(e)$ between 0 and 1) such that, for any attribute $X \in \V$, we have $\sum_{e \in \E: X \in e} w(e) \le 1$ (i.e., the weight sum of all the edges containing $X$ is at most 1). The {\em total weight} of $w$ is $\sum_{e \in \E} w(e)$, i.e., the weight sum of all the edges in $\E$. The {\em fractional edge packing number} of $\G$, denoted as $\tau(\G)$, is defined as the maximum total weight of all the fractional edge packings of $\G$.

\vgap

Dual to fractional edge packing is the notion of {\em fractional vertex cover}, which is a function $w: \V \rightarrow [0, 1]$ (mapping each vertex $X \in \V$ to a weight $w(X)$ between 0 and 1) such that, for any edge $e \in \E$, we have $\sum_{X \in \V: X \in e} w(X) \ge 1$ (i.e., the weight sum of all the vertices in $e$ is at least 1). The {\em total weight} of $w$ is $\sum_{X \in \V} w(X)$, i.e., the weight sum of all the vertices in $\V$. The {\em fractional vertex cover number} of $\G$ is defined as the minimum total weight of all the fractional vertex covers of $\G$. It can be shown \cite{su97} that the fractional vertex cover number of $\G$ is always identical to the fractional edge packing number $\tau(\G)$.

\vgap

Given a proper subset $\Z \subset \V$, we define $\G_\Z$ as the hypergraph obtained by ``removing'' the attributes of $\Z$ from $\G$. Specifically:
\myitems{
    \item The vertex set of $\G_\Z$ is $\V \setminus \Z$.
    \item The edges of $\G_\Z$ are decided as follows: for each edge $e \in \G$, we add the edge $e \setminus \Z$ to $\G_\Z$ if $e \setminus \Z \ne \emptyset$.
}
We will refer to $\G_\Z$ as a {\em residual graph} of $\G$. Specially, if $\Z = \emptyset$, then $\G_\Z = \G$.

\vgap

For each proper subset $\Z \subset \V$, the residual graph $\G_\Z$ has its own fractional edge packing number $\tau(\G_\Z)$. The maximum such number of all possible $\Z$ is the {\em edge quasi-packing number} $\psi(\G)$ of $\G$, or formally:
\myeqn{
    \psi(\G) &=&
    \max_{\Z \subset \V} \tau(\G_\Z).
    \label{eqn:hypergraph:psi}
}
Due to the equivalence between fractional edge packing number and fractional vertex cover number, the value $\psi(\G)$ can alternatively be defined as the maximum fractional vertex cover number of all residual graphs $G_\Z$ of $\G$.

\section{A Skew-Free Gathering Algorithm} \label{app:sf-gathering-alg}

In this section, we present a deterministic communication-oblivious algorithm to solve the skew-free gathering problem (parameterized by integer $t$) with load $O(N / p^{1/t})$, matching the lower bound in Theorem~\ref{thm:gathering}.

\vgap

Our algorithm is in essence an adaptation of the ``hyper-cube algorithm'' in \cite{au11}. Define $h = \floor{p^{1/t}}$, and order the points in the $t$-dimensional space $[h]^t$ lexicographically. For each $i \in [h^t]$, assign machine $i$ to the $i$-th point in $[h]^t$ under the lexicographic order. Note that some machines (precisely, $p - h^t$ machines) are not assigned to any point.

\vgap

Divide domain $[p]$ into $h$ disjoint intervals $I_1, I_2, ..., I_h$, ordered in ascending order of their starting points, such that:
\myitems{
    \item each of $I_1, ..., I_{h-1}$ covers $\floor{p/h}$ integers;
    \item $I_h$ covers $p - (h-1) \floor{p/h}$ integers.
}
For each interval $I_c$ with $c \in [h]$, we use $|I_c|$ to denote the number of integers covered by $I_c$. Consider any point $(c_1, c_2, ..., c_t) \in [h]^t$. Remember that the point corresponds to a machine. We associate the machine with a $t$-dimensional box $I_{c_1} \times I_{c_2} \times ... \times I_{c_t}$ and will refer to $I_{c_j}$ the machine's {\em projection} on dimension $j \in [t]$.

\vgap

Given a legal partition of the skew-free gathering problem, each machine --- say the one with id $i \in [p]$ --- sends the $N/p$ elements in its local storage to every machine whose projection on {\em at least} one dimension covers the value $i$ (there are less than $t \cdot h^{t-1}$ such machines). This completes the description of our algorithm.

\vgap

To see the algorithm's correctness, suppose that machines where the special elements are initially placed have ids $z_1, z_2, ..., z_t \in [p]$. For each $j \in [t]$, let $c_j$ be the integer in $[h]$ such that the interval $I_{c_j}$ covers $z_j$. Our algorithm ensures that the machine associated with the $t$-dimensional box $I_{c_1} \times I_{c_2} \times ... \times I_{c_t}$ receives all the special elements from the machines $z_1$, ..., $z_t$. The algorithm's communication obliviousness can be verified from the fact that, for each $i \in [p]$, machine $i$ sends exactly $N/p$ elements to a set of machines that is determined in a way independent of how the input elements are initially distributed.

\vgap

It remains to analyze the algorithm's load. In terms of sending, each machine delivers $N/p$ elements to less than $t \cdot h^{t-1} = O(h^{t-1}) = O(p^{(t-1)/t})$ machines and, thus, transmits $O((N/p) \cdot p^{(t-1)/t}) = O(N/p^{1/t})$ words. In terms of receiving, the machine corresponding to the $t$-dimensional box $I_{c_1} \times I_{c_2} \times ... \times I_{c_t}$ receives at most $(N/p) \cdot \sum_{j=1}^t |I_{c_j}|$ elements. Among all the intervals $I_1, ..., I_h$, the interval $I_h$ is the longest and covers
\myeqn{
    p - (h-1) \lf \fr{p}{h} \rf < p - (h-1) \left( \fr{p}{h} - 1 \right)
    < \fr{p}{h} + h = O(p^{1-1/t}) \nn
}
integers. Each machine, therefore, receives $O((N/p) \cdot p^{1-1/t}) = O(N / p^{1/t})$ words. This proves that our algorithm has a load of $O(N / p^{1/t})$.

\section{Completing the Proof of Thm.~\ref{thm:gathering}} \label{app:proof:thm:gathering}

\noindent {\bf Proof of Inequality (\ref{eqn:gathering:total-X-Y:help2}).} We adopt a counting argument similar to the one used to prove \eqref{eqn:gathering:total-X-Y:help1}. Initially, create an empty set $S_i$ for each $i \in [p]$. Then, process each $Z = (z_1, ..., z_t) \in \YY$ as follows. If $Y(Z, \mb{\Lambda}) = 0$, do nothing. Otherwise, under the communication pattern $\mb{L} = \mb{\Lambda}$, at least one machine $z_k$, for some $k \in [t]$, receives $N/p$ special elements from each machine $z_j$ with $j \in [t] \setminus \set{k}$. This means that all the machines in $Z \setminus \set{z_k}$ are heavy senders for machine $z_k$ under $\mb{\Lambda}$. Thus, $Z \setminus \set{z_k}$ represents a possible way to choose $t-1$ machines from the $H(\mb{\Lambda}, z_k)$ heavy senders for machine $z_k$ under $\mb{\Lambda}$. Accordingly, we add $Z$ to $S_k$. After all the sets $Z \in \YY$ have been processed, the left-hand side of \eqref{eqn:gathering:total-X-Y:help2} is bounded by $\sum_{i \in [p]} |S_i|$. On the other hand, for each $i \in [p]$, the size $|S_i|$ cannot exceed the total number of ways of choosing $t - 1$ machines from the $H(\mb{\Lambda}, i)$ heavy senders for machine $i$ under $\mb{\Lambda}$. Inequality \eqref{eqn:gathering:total-X-Y:help2} now follows.

\extraspacing {\bf Proof of \eqref{eqn:gathering:goal2} When $|\YY| \ge \fr{1}{2} \binom{p}{t}$.} By the definition of the Y-partition, the fact $|\YY| \ge \fr{1}{2} \binom{p}{t}$ yields
\myeqn{
    \sum_{Z \in \YY} \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
    Y(Z, \mb{\Lambda}) \ge \fr{1}{24} \binom{p}{t}.
    \label{eqn:gathering:Y-help1}
}
Then, we can obtain
\myeqn{
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} H(\mb{\Lambda},i)^{t-1}
    &\ge&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \binom{H(\mb{\Lambda},i)}{t-1}
    \nn \\
    \explain{by \eqref{eqn:gathering:total-X-Y:help2}}
    &\ge&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{Z \in \YY} Y(Z, \mb{\Lambda})
    \nn \\
    &=&
    \sum_{\mb{\Lambda} \in \L} \sum_{Z \in \YY} \Pr[\mb{L} = \mb{\Lambda}] \cdot  Y(Z, \mb{\Lambda})
    \nn \\
    \explain{by \eqref{eqn:gathering:Y-help1}}
    &\ge&
    \fr{1}{24} \binom{p}{t}.
    \label{eqn:gathering:Y-help2}
}
This leads to:
\myeqn{
    %\text{LHS of \eqref{eqn:gathering:goal2}}
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \sum_{i \in [p]} H(\mb{\Lambda},i)
    &=&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^{t-1}}{H(\mb{\Lambda},i)^{t-2}}
    \nn \\
    \explain{by \eqref{eqn:gathering:bound-H}}
    &\ge&
    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^{t-1}}{p^{\fr{(t-1)\cdot(t-2)}{t}}}
    \nn \\
    \explain{by \eqref{eqn:gathering:Y-help2}}
    &\ge&
    \fr{1}{24} \binom{p}{t} / p^{\fr{(t-1)\cdot(t-2)}{t}}
    \nn \\
    &=&
    \Omega\Big(
        \fr{p^t}{p^{\fr{(t-1)\cdot(t-2)}{t}}}
    \Big)
    =
    \Omega\Big(p^{3-2/t} \Big).
    \nn
}
Therefore, at least one $i \in [p]$ satisfies $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot H(\mb{\Lambda},i) = \Omega(p^{3-2/t} / p ) = \Omega (p^{1-1/t})$, as claimed in \eqref{eqn:gathering:goal2}.

\section{Completing the Proof of Thm.~\ref{thm:join}} \label{app:proof:thm:join}

\noindent {\bf Proof of Lemma~\ref{lmm:share}.} As mentioned, Beame et al.\ \cite{bks17c} has proved Lemma~\ref{lmm:share} for the case where $M = |R^*|$. We will utilize their result as a black box to prove the lemma under $|R^*| < M \leq N$.

\vgap

The number of distinct values taken by the tuples in $R^*$ is at most $|R^*|\cdot |schema(R^*)| \leq \alpha |R^*|$. Hence, we can find in the domain $[\alpha N]$ a set $S$ of $\alpha N - \alpha |R^*|$ values that are not taken by any tuple of $R^*$. From $|R^*| < M \leq N$, we know
\myeqn{
    |S| = \alpha N - \alpha |R^*| > \alpha N - (\alpha -1)N -|R^*| = N - |R^*| \geq M - |R^*|. \nn
}
Identify (arbitrarily) $M - |R^*|$ distinct values from $S$ and, for each such value $x$, create a tuple $\bm{u}$ over $\schema(R^*)$ with $\bm{u}(X) = x$ for each $X \in \schema(R^*)$. Let $R'$ be the relation obtained by inserting these $M - |R^*|$ tuples into $R^*$. Note that $|R'|$ has precisely $M$ tuples.

\vgap

It is easy to verify that $\deg_\Y(R^*) = \deg_\Y(R')$ for every non-empty subset $\Y \subseteq \schema(R')$. With this, the skew-free condition \eqref{eqn:share-skewfree} tells us that, for any non-empty subset $\Y \subseteq \schema(R')$, we must have:
\myeqn{
    \deg_\Y(R') \le \fr{|R'|}{{\prod_{i \in [r]: X_i \in \Y} s_i}}. \nn
}

Recall that, for each $i \in [r]$, we have chosen a perfectly random hash function $h_i: [\alpha N] \rightarrow [s_i]$. Given $\bm{b} = (b_1, b_2, ..., b_r)$ $\in [s_1] \times [s_2] \times ... \times [s_r]$, define
\myeqn{
    \text{\em bin $\bm{b'}$}
    &=&
    \set{\bm{u} \in R' \mid \forall i \in [r], h_i(\bm{u}(X_i)) = b_i}.
    \label{eqn:proof:thm:join:help1}
}
Clearly, bin $\bm{b'}$ contains all the tuples in the bin $\bm{b}$ defined in \eqref{eqn:join-share:bin}. By the result of \cite{bks17c}, we know that every bin $\bm{b'}$ in \eqref{eqn:proof:thm:join:help1} contains $O((\log p^* / \log\log p^*)^r \cdot |R'| / p^*) = O((\log p^* / \log\log p^*)^r \cdot M / p^*)$ tuples, where $p^* = \prod_{i=1}^r s_i$. It thus follows that every bin $\bm{b}$ in \eqref{eqn:join-share:bin} must have a size $O((\log p^* / \log\log p^*)^r \cdot M / p^*)$.

\extraspacing {\bf Proof of Equation~\eqref{eqn:join-help1}.} If $\Y \subseteq \Z$, then $\prod_{X \in \Y} s_X(\Z) = 1 \le p$. Otherwise:
\myeqn{
    \prod_{X \in \Y} s_X(\Z) =
    \prod_{X \in \Y \setminus \Z} s_X(\Z)
    \le
    \prod_{X \in \Y \setminus \Z} p^{w_\Z(X) / \tau_\Z}
    \le
    p.
    \nn
}

\extraspacing {\bf Proof of Equation~\eqref{eqn:join-help2}.} Let $e_R(\Z) = \schema(R)\setminus \Z$. Note that $e_R(\Z)$ is an edge in the residual graph $\G_\Z$. We have
\myeqn{
    \prod_{X \in \schema(R)} s_X(\Z)
    &=&
    \prod_{X \in e_R(\Z)} s_X(\Z) \cdot \prod_{X \in \Z} s_X(\Z) \nn\\
    &=& \prod_{X \in e_R(\Z)} s_X(\Z) \nn \\
    &=& \prod_{X \in e_R(\Z)} \lf p^{w_\Z(X) / \tau(\G_\Z)} \rf \nn \\
    \explain{as $p^{w_\Z(X) / \tau(\G_\Z)} \geq 1$}&\ge& \prod_{X \in e_R(\Z)} \fr{p^{w_\Z(X) / \tau(\G_\Z)}}{2} \nn \\
   \explain{as $|e_R(\Z)|\leq |\schema(R)|\leq \alpha $} &\ge&
   \fr{1}{2^\alpha} \prod_{X \in e_R(\Z)} p^{w_\Z(X) / \tau(\G_\Z)} \nn \\
    &\ge& p^{1/\tau(\G_\Z)} / 2^{\alpha} \nn
}
where the last inequality used the fact that $w_\Z$ is a fractional vertex cover of $G_\Z$ (and, hence, $\sum_{X \in {e_R(\Z)}} w_\Z(X) \ge 1$).

\section{Completing the Proof of Thm.~\ref{thm:compile}} \label{app:compile}

\noindent {\bf Chernoff Bounds.} The following Chernoff bounds (proved in \cite{stl12}) will be useful:

\begin{lemma} \label{lmm::chernoff}
    Let $X_1, X_2, ..., X_f$ be independent Bernoulli random variables that are identically distributed. Define $X = \sum_{i \in [f]} X_i$. For any $\gamma \geq 2$, it holds that
    \myeqn{
        \Pr[X \geq \gamma \cdot \expt[X]] &\leq&  \exp(-\gamma \cdot \expt[X] / 6). \label{eqn:compile:chernoff1}
    }
    For any $0 < \gamma < 1$, it holds that
    \myeqn{
        \Pr[X \geq (1 + \gamma) \expt[X]] &\leq&  \exp(- \gamma^2 \cdot \expt[X] / 3). \label{eqn:compile:chernoff2}
    }
\end{lemma}

Fix an arbitrary real value $\gamma$ satisfying $0 < \gamma < 1$ and any positive real values $p$ and $\beta$. Let us specialize the setup of random variables in the above lemma as follows:
\myitems{
    \item $\Pr[X_i = 1] = 1/p$ for each $i \in [f]$;
    \item $f \le F$ where $F \ge \beta \cdot p \log p$.
}
We claim:
\myeqn{
    \Pr[X \geq (1 + \gamma) F/p] &\leq&  \exp(- \Omega(\gamma^2 \beta \cdot \log p)) \label{eqn:compile:chernoff3}
}
where the hidden constant in the big-$\Omega$ does not depend on $\gamma$, $p$, or $\beta$. Next, we will prove the claim by distinguishing two cases.

\vgap

{\em Case 1: $\expt[X] \ge F / (2p)$.} We have:
\myeqn{
    \Pr[X \geq (1 + \gamma) F/p] &\le& \Pr[X \geq (1 + \gamma) \expt[X]] \nn \\
    \explain{by \eqref{eqn:compile:chernoff2}} &\leq&  \exp(- \gamma^2 \cdot \expt[X] / 3) \nn \\
    &\leq&  \exp(- \gamma^2 \cdot F / (6p)) \nn \\
    &\leq&  \exp(- (\gamma^2 \beta / 6) \cdot \log p) \nn
}
as claimed.

{\em Case 2: $\expt[X] < F / (2p)$.} We can derive:
\myeqn{
    \Pr\left[X \geq (1 + \gamma) F/p\right] &=&
    \Pr\left[X \geq \fr{(1 + \gamma)}{\expt[X] / (F/p)} \cdot \expt[X] \right] \nn \\
    \explain{by \eqref{eqn:compile:chernoff1}} &\leq&  \exp(- (1 + \gamma) \cdot F/(6p)) \nn \\
    &\leq&  \exp(- (\beta / 6) \cdot \log p) \nn
}
as claimed.

\extraspacing {\bf Completing the Analysis of Our Message-Routing Algorithm.} As discussed in Section~\ref{sec:compile}, it remains to prove that our algorithm errs with probability at most $1/p^c$ where $c$ can be an arbitrarily large constant chosen before running the algorithm.

\vgap

For the algorithm to err in the first round, $|P_i(j)|$ must exceed $\lambda$ (see its value in \eqref{eqn:compile-lambda}) for some $i, j \in [p]$. Set $f = |P_i|$ and $F = p + L / \delta'$; we thus have $f \le F$ by \eqref{eqn:compile-P_i-size}. For each $k \in [f]$, introduce a Bernoulli random variable $X_k$ that equals 1 if the $k$-th element of $P_i$ is relayed to machine $j$, and 0 otherwise. Thus, $\Pr[X_k = 1] = 1/p$, and $|P_i(j)| = \sum_{k \in [f]} X_k$.

\vgap

Suppose that $L \ge \beta \cdot p \log p$ where $\beta$ is a constant to be decided later. Note that this means $F > L / \delta' \ge (\beta/\delta') p \log p$.
We can derive:
\myeqn{
    \Pr[|P_i(j)| \ge \lambda]
    &\le&
    \Pr \left[|P_i(j)| \ge \left(1 + \fr{L}{\delta' p} \right) (1+\delta/2) \right] \nn \\
    &=&
    \Pr \left[|P_i(j)| \ge (1+\delta/2) \cdot F/p \right] \nn \\
    \explain{by \eqref{eqn:compile:chernoff3}}
    &\le&
    \exp(- \Omega(\delta^2 \cdot (\beta/\delta') \cdot \log p))  \nn \\
    &=&
    \exp(- \Omega(\delta^3 \cdot \beta \cdot \log p))
    \label{eqn:compile:chernoff4}
}
which can be made less than $1/p^{c_0}$ for an arbitrarily large constant $c_0$ by setting $\beta$ sufficiently large (recall that $\delta$ is a constant).

\vgap

How likely the algorithm errs in the second round is somewhat less obvious. For each $j \in [p]$, denote by $P^+_j$ the set of packets --- produced from $\mb{M}[1,j]$, $\mb{M}[2,j]$, ..., $\mb{M}[p,j]$ combined --- destined for machine $j$ (i.e., those packets have $j$ as the recipient). Because of \eqref{eqn:compile-len-constraint}, $|P^+_j|$, the number of packets in $P^+_j$, is at most $p + L/\delta'$, following a derivation similar to that in \eqref{eqn:compile-P_i-size}. The algorithm errs in the second round only if $|P^*_i(j)| > \lambda$ for some $i, j \in [p]$. Observe that $|P^*_i(j)|$ is precisely the number of packets in $P^+_j$ that have $i$ as their relay machines. Now, set $f = |P^+_j|$ and $F = p + L / \delta'$; we thus have $f \le F$ as just explained. For each $k \in [f]$, introduce a Bernoulli random variable $X_k$ that equals 1 if the $k$-th element of $P^+_j$ is relayed to machine $i$, and 0 otherwise. Thus, $\Pr[X_k = 1] = 1/p$, and $|P^*_i(j)| = \sum_{k \in [f]} X_k$. Following a derivation similar to the one leading to \eqref{eqn:compile:chernoff4}, we obtain
\myeqn{
    \Pr[|P^*_i(j)| \ge \lambda]
    &\le&
    \exp(- \Omega(\delta^3 \cdot \beta \cdot \log p))   \label{eqn:compile:chernoff5}
}

We can now conclude from \eqref{eqn:compile:chernoff4} and \eqref{eqn:compile:chernoff5} that, when $L \ge \beta \cdot p \log p$ for a sufficiently large $\beta$, our algorithm errs with a probability at most $1 / p^c$.

\section{Proof of Theorem \ref{thm:token}} \label{app:thm:token}

Let us start by constructing a set of $p(p-1)$ parameterized inputs. Denote by $\Pi$ the set of $p(p-1)$ ordered pairs $(x, y) \in [p] \times [p]$ satisfying $x \ne y$. For every $x \in [p]$, build a legal partition $\pi_x$ for token-passing by placing $\ceil{p^{1-\eps}}$ tokens on machine $x$ and placing $\ceil{p^{1-\eps}}$ innocuous elements on every other machine. Each pair $(x, y) \in \Pi$ then specifies a distinct parameterized input --- with legal partition $\pi_x$ and parameters $x, y$ --- for which the objective is to move all the tokens from machine $x$ to machine $y$.

\vgap

Let $\A$ be a two-round communication oblivious algorithm that succeeds with probability at least $2/3$ on each of the $p(p-1)$ parameterized inputs defined by $\Pi$. We will prove that $\A$ must entail a load of $\Omega(p^{1-\eps/2})$ with probability at least $1/2$.

\vgap

As $\A$ performs two rounds, its communication pattern can be represented as two $p \times p$ matrices $\mb{L}_1$ and $\mb{L}_2$. Specifically, for each $r = 1$ and $2$, the value $\mb{L}_r[i, j]$ (where $i, j \in [p]$) is the number of words that machine $i$ sends to machine $j$ in round $r$. If $\A$ is randomized, then $\mb{L}_1$ and $\mb{L}_2$ are random variables. For each $r =1$ and 2, we define a $p \times p$ matrix $\mb{L}^*_r$ from $\mb{L}_r$ as follows:
\myitems{
    \item $\mb{L}^*_r[i, j] = \mb{L}_r[i, j]$ for any $i, j \in [p]$ satisfying $i \ne j$.
    \item $\mb{L}^*_r[i, i] = \ceil{p^{1-\eps}}$ for each $i \in [p]$.
}
We remind the reader that $\mb{L}_r[i,i] = 0$ for all $i \in [p]$, by definition of communication pattern. The matrix $\mb{L}^*_r$ thus defined is a random variable decided by $\mb{L}_r$.

\vgap

By communication obliviousness, the distribution of $(\mb{L}_1, \mb{L}_2)$ is the same for all the parameterized inputs specified by the pairs in $\Pi$. Let $\L$ be the set of ordered pairs $(\mb{\Lambda}_1$, $\mb{\Lambda}_2)$ satisfying
\myitems{
    \item each of $\mb{\Lambda}_1$ and $\mb{\Lambda}_2$ is a $p \times p$ matrix;

    \item $\Pr[(\mb{L}^*_1$, $\mb{L}^*_2) = (\mb{\Lambda}_1$, $\mb{\Lambda}_2)] > 0$, namely, $\A$ exhibits with a non-zero probability a pattern $(\mb{L}_1$, $\mb{L}_2)$ such that the corresponding $(\mb{L}^*_1$, $\mb{L}^*_2)$ equals $(\mb{\Lambda}_1$, $\mb{\Lambda}_2)$;

    \item the following inequality holds for all $k \in [p]$:
    \myeqn{
        \Big(\sum_{i \in [p]} \mb{\Lambda}_1 [i,k] \Big) \cdot \Big(\sum_{j \in [p]} \mb{\Lambda}_2 [k,j] \Big) \le p^{2-\eps}/24. \label{eqn:thm:token:help-1}
    }
}
The core of our argument is to prove
\myeqn{
    \Pr[(\mb{L}^*_1, \mb{L}^*_2) \notin \L] \ge 1/2.
    \label{eqn:thm:token:help-2}
}
Note that whenever $(\mb{L}^*_1, \mb{L}^*_2) \notin \L$, it follows from the definition in \eqref{eqn:thm:token:help-1} that there is at least one $k \in [p]$ such that
\myeqn{
    \Big(\sum_{i \in [p]} \mb{L}^*_1 [i,k] \Big) \cdot \Big(\sum_{j \in [p]} \mb{L}^*_2 [k,j] \Big) > p^{2-\eps}/24 \nn
}
under which one of the following inequalities must be true:
\myitems{
    \item $\sum_{i \in [p]} \mb{L}^*_1 [i,k] > p^{1-\eps/2} / \sqrt{24}$, or
    \item $\sum_{j \in [p]} \mb{L}^*_2 [k,j] > p^{1-\eps/2} / \sqrt{24}$.
}
In the former case, we have
\myeqn{
    \sum_{i \in [p]} \mb{L}_1 [i,k] &=& \Big(\sum_{i \in [p]} \mb{L}^*_1 [i,k] \Big) - \mb{L}^*_1 [k,k] \nn \\
    &=& \Big(\sum_{i \in [p]} \mb{L}^*_1 [i,k] \Big) - \ceil{p^{1-\eps}}
    = \Omega \Big(p^{1-\eps/2}\Big). \nn
}
This means that the load of $\A$ is $\Omega(p^{1-\eps/2})$ as $\sum_{i \in [p]} \mb{L}_1 [i,k]$ is the number of words received by machine $k$ in the first round. In the latter case, following a similar derivation one can obtain $\sum_{j \in [p]} \mb{L}_2 [k,j] = \Omega(p^{1-\eps/2})$. This also implies that $\A$ has load $\Omega(p^{1-\eps/2})$ because $\sum_{j \in [p]} \mb{L}_2 [k,j]$ is the number of words sent by machine $k$ in the second round. Therefore, with probability at least $1/2$, algorithm $\A$ demands a load of $\Omega(p^{1-\eps/2})$, which completes the proof of Theorem~\ref{thm:token}.

\extraspacing {\bf Proof of Inequality \eqref{eqn:thm:token:help-2}.} Assume for contradiction purposes that \eqref{eqn:thm:token:help-2} does not hold, which means $\Pr[(\mb{L}^*_1, \mb{L}^*_2) \in \L] > 1/2$.

\vgap

For every parameterized input specified by $(x, y) \in \Pi$, we know
\myeqn{
    \Pr[\text{$\A$ succeeds on $(x, y)$} \mid (\mb{L}^*_1, \mb{L}^*_2) \in \L] \ge 1/6.
    \label{eqn:thm:token:help-4}
}
Otherwise, $\Pr[\text{$\A$ succeeds on $(x, y)$}] < 1/6 + \Pr[(\mb{L}^*_1, \mb{L}^*_2) \notin \L] < 2/3$, contradicting the fact that $\A$ succeeds on any input with probability at least $2/3$.

\begin{lemma} \label{lmm:thm:token:help-1}
    If $\A$ succeeds on the parameterized input specified by $(x, y) \in \Pi$ under the condition $(\mb{L}^*_1, \mb{L}^*_2)  = (\mb{\Lambda}_1, \mb{\Lambda}_2)$, then
    \myeqn{
        \sum_{k \in [p]} \min\big\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \big\} \ge p^{1-\eps}.
        \label{eqn:thm:token:help-3}
    }
\end{lemma}

\begin{proof}
    Let us rewrite the left hand side of \eqref{eqn:thm:token:help-3} as
    \myeqn{
        \hspace{-5mm}
        && \sum_{k \in [p]} \min\big\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \big\} \nn \\
        \hspace{-5mm}& = &
        \Big( \sum_{k \in [p] \setminus \set{x,y}} \min\big\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \big\} \Big)
        \nn \\
        \hspace{-5mm}&& + \min\big\{\mb{\Lambda}_1[x, y], \ceil{p^{1-\eps}} \big\}
        + \min\big\{\ceil{p^{1-\eps}}, \mb{\Lambda}_2[x, y] \big\}.
        \label{eqn:thm:token:help-8}
    }

    If a machine $k \in [p] \setminus \set{x, y}$ receives a token from machine $x$ in the first round and sends the token to machine $y$ in the second round, we say that the token is {\em relayed} by machine $k$. Under the condition $(\mb{L}^*_1, \mb{L}^*_2)  = (\mb{\Lambda}_1, \mb{\Lambda}_2)$, machine $k$ can relay no more than $\min\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \}$ tokens. On the other hand, if a token is not relayed by any machines, it needs to be sent from machine $x$ to machine $y$ directly (in either the first or the second round). The number of tokens that can be sent this way is at most
    \myeqn{
        \min\big\{\mb{\Lambda}_1[x, y], \ceil{p^{1-\eps}} \big\}
        + \min\big\{\ceil{p^{1-\eps}}, \mb{\Lambda}_2[x, y] \big\}. \nn
    }

    Thus, \eqref{eqn:thm:token:help-8} gives an upper bound on the total number of tokens that $\A$ can move to machine $y$ at the end of the second round. The upper bound must be at least $\ceil{p^{1-\eps}}$ for $\A$ to succeed on $(x, y)$.
\end{proof}

By putting together \eqref{eqn:thm:token:help-4} and \eqref{eqn:thm:token:help-3}, we obtain the following for every parameterized input specified by $(x, y) \in \Pi$:
\myeqn{
    \expt\Big[\sum_{k \in [p]} \min\big\{\mb{L}^*_1[x, k], \mb{L}^*_2[k, y] \big\} \bigmid (\mb{L}^*_1, \mb{L}^*_2) \in \L \Big]
    \ge
    \fr{p^{1-\eps}}{6}
    \label{eqn:thm:token:help-5}
}

For any $(\mb{\Lambda}_1, \mb{\Lambda}_2) \in \L$ and any $k \in [p]$, we have: 
\myeqn{
    && \Big( \sum_{x \in [p]} \mb{\Lambda}_1[x,k] \Big)
    \cdot \Big( \sum_{y \in [p]} \mb{\Lambda}_2[k,y] \Big) \nn \\
    &=&
    \sum_{x \in [p]} \sum_{y \in [p]}
     \mb{\Lambda}_1[x,k] \cdot \mb{\Lambda}_2[k,y]  \nn \\
    &\ge& 
    \sum_{x \in [p]} \sum_{y \in [p]}
    \min\big\{\mb{\Lambda}_1[x,k], \mb{\Lambda}_2[k,y]\big\}  \nn \\ 
    &\ge& 
    \sum_{(x,y) \in \Pi} 
    \min\big\{\mb{\Lambda}_1[x,k], \mb{\Lambda}_2[k,y]\big\}  \label{eqn:thm:token:help-6}
}
where the second inequality used the fact that $ab \ge \min\{a, b\}$ holds for any non-negative integers $a$ and $b$.

\vgap

We can now derive:
\myeqn{
    && \sum_{k \in [p]} \expt \Big[ \Big(\sum_{x \in [p]} \mb{L}^*_1[i, k] \Big) \cdot \Big( \sum_{y \in [p]} \mb{L}^*_2[k, j] \Big)
    \bigmid
    (\mb{L}^*_1, \mb{L}^*_2) \in \L
    \Big] \nn \\
    &\ge&
    \expt \Big[\sum_{k \in [p]} \sum_{(x,y) \in \Pi} \min\big\{\mb{L}^*_1[x,k], \mb{L}^*_2[k,y]\big\}
    \bigmid
    (\mb{L}^*_1, \mb{L}^*_2) \in \L
    \Big] \nn \\
    && \explain{by linearity of expectation and \eqref{eqn:thm:token:help-6}} \nn \\
    &=&
    \expt \Big[\sum_{(x,y) \in \Pi} \sum_{k \in [p]} \min\big\{\mb{L}^*_1[x,k], \mb{L}^*_2[k,y]\big\}
    \bigmid
    (\mb{L}^*_1, \mb{L}^*_2) \in \L
    \Big] \nn \\
    &=&
    \sum_{(x,y) \in \Pi}
    \expt \Big[\sum_{k \in [p]} \min\big\{\mb{L}^*_1[x,k], \mb{L}^*_2[k,y]\big\}
    \bigmid
    (\mb{L}^*_1, \mb{L}^*_2) \in \L
    \Big] \nn \\
    && \explain{linearity of expectation} \nn \\ 
    &\ge& 
    \fr{p^{1-\eps} \cdot p (p-1)}{6}. \hspace{5mm} \explain{by \eqref{eqn:thm:token:help-5}} \nn 
}

Therefore, there is at least one $k^* \in [p]$ such that 
\myeqn{
    && \expt \Big[ \Big(\sum_{x \in [p]} \mb{L}^*_1[i, k^*] \Big) \cdot \Big( \sum_{y \in [p]} \mb{L}^*_2[k^*, j] \Big)
    \bigmid
    (\mb{L}^*_1, \mb{L}^*_2) \in \L
    \Big] \nn \\ 
    &\ge& 
    \fr{p^{1-\eps}(p-1)}{6} \ge p^{2-\eps}/12
    \label{eqn:thm:token:help-7}
}
where the last inequality used $p \ge 2$.

\vgap

However, by the definition of $\L$ (in particular, the inequality \eqref{eqn:thm:token:help-1}), when $ (\mb{L}^*_1, \mb{L}^*_2) \in \L $, the product $\sum_{x \in [p]} \mb{L}^*_1[i, k]  \cdot \sum_{y \in [p]} \mb{L}^*_2[k, j]$ can be at most $p^{2-\eps}/24$ for all $k \in [p]$. This contradicts the existence of $k^*$. Therefore, our assumption
$\Pr[(\mb{L}^*_1, \mb{L}^*_2) \in \L] > 1/2$ must be a false one.

\balance

%\end{sloppy}
\end{document}

### 2024-05-16 msj.tex
yufei

@@
 
-\def\agm{\mathrm{AGM}}
-\def\att{\mathbf{att}}
-\def\actdom{\mathbf{actdom}}
-\def\dom{\mathbf{dom}}
-\def\gat{\mit{gat}}
-%\def\id{\mathit{id}}
-%\def\In{\mathrm{IN}}
-\def\heavy{\mit{heavy}}
-\def\join{\mathit{Join}}
-\def\key{\mathit{key}}
-%\def\light{\mathit{light}}
-\def\light{\ell}
-\def\load{\mathit{load}}
-%\def\no{\mathit{no}}
-\def\occ{\mathrm{OCC}}
-\def\Out{\mathrm{OUT}}
-\def\rhostar{{\rho^*}}
-\def\schema{\mathit{schema}}
-\def\var{\mathit{var}}
-
-\newcommand*{\ldb}{\{\mskip-5mu\{}
-\newcommand*{\rdb}{\}\mskip-5mu\}}
-
 \allowdisplaybreaks
@@
 
-\title{Parallel Communication Obliviousness: One Round and Beyond}
+\title{Optimal Multiway Spatial Joins}
 
@@
 \begin{abstract}
-    This paper studies {\em communication-oblivious} algorithms under the massively parallel computation (MPC) model whose communication patterns follow a distribution that depends only on the number $p$ of machines and the input size $N$ of the underlying problem, regardless of the specific input elements.
-Our objective is to understand when obliviousness necessitates --- or does not necessitate --- heavier communication compared to the traditional MPC model that does not enforce such a requirement.
-
-    \vgap
-
-    The first part of our investigation focuses on single-round algorithms.
-We prove that {\em skew-free hashing}, a fundamental problem solvable with load $\tO(N/p)$ with high probability (w.h.p.) under the traditional model, demands a load of nearly $\Omega(N)$ under communication obliviousness.
-Intriguingly, we show that hashing can still be applied in an oblivious manner to process any natural join in one round w.h.p.\ with a load complexity matching that of the best traditional MPC algorithm.
-The second part of our investigation studies {\em compilation methods} that convert a traditional MPC algorithm $\A$ into a communication-oblivious counterpart.
-Given an $\A$ that operates within $\ell = \poly(p)$ rounds with a load at most $L = \Omega(p \log p)$, we can produce a communication-oblivious version running in $2\ell$ rounds with a load at most $(1 + \delta) L$, where $\delta > 0$ can be an arbitrarily small constant.
-Additionally, we establish hardness results indicating that the theoretical guarantees of our compilation can no longer be significantly improved.
+    Later.
 \end{abstract}
@@
 
-In the digital era, cloud computing has emerged as a pivotal part of modern business infrastructure, reshaping the way organizations store, access, and manage data.
-A growing number of sectors --- ranging from healthcare and finance to academia and entertainment --- are harnessing the power of cloud services to enhance operational efficiency, effectuate cost savings, and fuel innovation.
-However, the migration of data from local systems to remote servers operated by third-party service providers has ushered in new privacy concerns.
-The crux of these concerns is the fact that sensitive data, once stored on these servers, is no longer fully under the user's control.
-Instead, it falls within the stewardship of service providers, thereby raising critical questions about data privacy.
-
-\vgap
-
-Two primary topics come to the forefront in the discourse on privacy control in cloud computing:
-\myitems{
-	\item {{\bf Confidential computing:}} This aims to mitigate two forms of privacy leakage that might occur on an {\em individual} server.
-First, rogue administrators could exploit their positions to inspect sensitive data within the server.
-Second, even if encryption hinders direct data accesses, administrators could still infer valuable information from the memory access patterns that the computation generates.
-
-	\item {{\bf Oblivious communication:}} This aims to prevent another two forms of privacy leakage that could transpire during communication across {\em different} servers.
-First, malicious network managers might scrutinize the data flowing through the gateways.
-Second, even with encrypted network traffic, these managers could still deduce sensitive information from the statistics extracted from the communication.
-}
-
-Robust solutions to confidential computing are already in place, including systems that employ {\em fully homomorphic encryption} (e.g., IBM's HE4Cloud) or hardware-based {\em trusted execution environments} (nowadays offered by the leading cloud-computing platforms).
-When integrated with {\em oblivious RAM} techniques \cite{akl+23,go96,ln18,ppry18}, these systems safeguard sensitive information during the entire lifecycle of data: whether the data is stored in the disk, transiting through the memory hierarchy, or in use by the CPU.
-
-\vgap
-
-In this work, we will concentrate on oblivious communication, with the aim to understand how much communication is needed to enforce this requirement.
-To pave the foundation of our discussion, next we will formalize the concept of oblivious communication and the class of algorithms to be studied.
-
-\subsection{Communication-Oblivious MPC} \label{sec:intro-prob}
-
-Let us first clarify some math conventions in this paper.
-For an integer $x \ge 1$, the notation $[x]$ denotes the set $\{1, 2, ..., x\}$.
-Given integers $x > 0$ and $y \ge 0$, let $\binom{x}{y}$ be the number of ways of choosing $y$ elements from a set size $x$; specially, if $y > x$, define $\binom{x}{y}$ $= 0$.
-We use double-curly braces to represent multi-sets, e.g., $\ldb 1, 1, 1, 2, 2, 3\rdb$ is a multi-set with 6 elements.
-Every $\log(\cdot)$ has base $2$.
-The notation $\poly(n)$ denotes the class of functions polynomial in $n$.
-
-\extraspacing {\bf The MPC Model.} Our analysis will be under the {\em massively parallel computation} (MPC) model, which has been extensively adopted in the database literature, as will be surveyed in Section~\ref{sec:intro-prev}.
-
-\vgap
-
-In this model, $p$ share-nothing machines are connected via a high-speed network.
-Initially, the input data is distributed across the $p$ machines.
-An algorithm then executes in {\em rounds}, each consisting of two phases: in the first phase, every machine performs computation on its local data, while in the second phase, the machines exchange messages via the network.
-Importantly, if machine $i \in [p]$ plans to send machine $j \in [p]$ a message $M[i,j]$ (in the second phase), it must prepare the message in the first phase.
-This restriction prevents machine $i$ from, for example, deciding what to send based on the messages received {\em during} the second phase.
-The {\em load of a round} is the maximum number of words communicated (sent and received combined) by one machine in that round.
-The algorithm's performance is measured by two metrics: (i) the total number of rounds executed, and (ii) the maximum load of all rounds, which is termed as the algorithm's (overall) {\em load}.
-
-\vgap
-
-Randomization is modeled by introducing a sequence of random bits, which is agreed upon by all machines.
-Such an agreement can be reached before receiving the input and therefore does not require communication during the algorithm's execution.
-In the first phase of a round, each machine can guide its local computation according to the random-bit sequence.
-When a random event is said to occur ``with high probability'' (w.h.p.), we require that the probability must be at least $1-1/p^c$ where $c$ can be an arbitrarily large constant chosen before seeing the input data.
-
-\vgap
-
-The value of $N$ is commonly assumed to be significantly larger than $p$, which typically means $N \ge p^c$ where $c$ is a constant dependent on the problem studied.
-
-\extraspacing {\bf Communication-Oblivious Algorithms.} We will now formally define the notion of communication obliviousness, building upon a similar formulation in \cite{ccls20}.
-
-\vgap
-
-We consider that an input $\I$ to the underlying problem is a set of elements, each represented with a constant number of words.
-The {\em problem size} $N$ of $\I$ is defined as the set size $|\I|$.
-Given a specific value of $N \ge p$, let $\II(N)$ be the collection of all possible inputs with the same problem size $N$.
-Fix an arbitrary input $\I \in \II(N)$.
-As mentioned, initially the $N$ elements of $\I$ are stored across the $p$ machines.
-We say that a partition $\pi = (\I_1, \I_2, ..., \I_p)$ of $\I$ --- that is, $\I_1, ..., \I_p$ are mutually disjoint and their union is $\I$ --- is {\em legal} if assigning $\I_i$ to machine $i$ for each $i \in [p]$ is permitted by the underlying problem.
-
-\vgap
-
-Consider an arbitrary legal partition $\pi = (\I_1, \I_2, ..., \I_p)$ of $\I$ $\in$ $\II(N)$.
-Suppose that, when executed on $\pi$, an MPC algorithm $\A$ finishes in $\ell$ rounds where, in the $r$-th round ($r \in [\ell]$), machine $i \in [p]$ sends message $\mb{M}_r[i,j]$ to machine $j \in [p]$ (for $i = j$, $\mb{M}_r[i,j]$ is an empty message).
-Define $\mb{L}_r$ as the $p \times p$ matrix where $\mb{L}_r[i, j]$ equals the length of $\mb{M}_r[i,j]$, measured in the number of words.
-Every message $\mb{M}_r[i,j]$ is encrypted such that an adversary, who observes the network communication, can only use the sequence %\vspace{-3mm}
-\myeqn{
-    \sigma = (\mb{L}_1, \mb{L}_2, ..., \mb{L}_\ell) \nn
-    %\label{eqn:intro:pattern}
-}
-to infer about $\I$. We refer to $\sigma$ as the {\em communication pattern} of $\A$ when executed on $\pi$. For a randomized algorithm $\A$, the sequence $\sigma$ may be a random variable.
-
-\vgap
-
-The algorithm $\A$ is {\em communication oblivious} under a specific problem size $N$ if its communication pattern $\sigma$ follows exactly the same probabilistic distribution, regardless of the choice of $\I \in \II(N)$ and the legal partition $\pi$ of $\I$.
-We say that $\A$ is a {\em single-round algorithm} if $\ell = 1$ for any choice of $\I$ and $\pi$; otherwise, $\A$ is a {\em multi-round algorithm}.
-
-\subsection{Our Contributions} \label{sec:intro:ours}
-
-Our primary objective is to discern the inherent connections between the traditional MPC model and its communication-oblivious counterpart.
-The core is to understand whether obliviousness would necessitate a provably higher load than the traditional, non-oblivious, MPC model.
-Next, we provide an overview of our findings.
-
-\extraspacing {\bf Single-Round Hashing.} Our first contribution is a hardness result proving that {\em hashing} --- a crucial building brick deployed extensively by MPC algorithms (in both theory and practice) --- is significantly more expensive under communication obliviousness as far as single-round algorithms are concerned.
-
-\vgap
-
-To pinpoint the source of hardness, we introduce a problem named {\em skew-free gathering}.
-Fix arbitrary integers $N$ and $p$ such that $N$ is a multiple of $p$, and $p \ge t$ where $t$ is a constant integer at least 2.
-The input collection $\II(N)$ has only a single input $\I$: this is a set of $N$ elements, each of which --- denoted as $e$ --- is associated with an integer $\key(e)$, referred to as the {\em key} of $e$.
-The set of keys fulfills two conditions:
-\myitems{
-    \item There is one special key that is possessed by $t \cdot (N/p)$ elements, called the {\em special elements}.
-
-    \item Every other key is possessed by exactly one element.
-}
-A partition $\pi = (\I_1, ..., \I_p)$ of $\I$ is {\em legal} if (i) $|\I_i| = N/p$ for all $i \in [p]$, and (ii)
-the $t \cdot N/p$ special elements are placed on $t$ distinct machines, i.e., there are $t$ different machine ids $j \in [p]$ such that $\I_j$ contains $N/p$ special elements.
-The goal of an algorithm is to move all the special elements to one (arbitrary) machine.
-
-\vgap
-
-In the conventional MPC model, the problem can be solved trivially with load $O(N/p)$: simply ask the $t$ machines where the special elements are stored initially to send their data to a common machine.
-In contrast, we prove that if a single-round communication-oblivious algorithm is required to succeed with probability at least $2/3$, its load must be at least $\Omega(N / p^{1/t})$ with a constant probability.
-As $t$ increases, the load approaches $\Omega(N)$ up to a factor sub-polynomial in $p$.
-More precisely, there does not exist any constant $\eps > 0$ such that one can design an algorithm promising a load of $O(N/p^\eps)$ with a constant probability for all $t = O(1)$.
-We also show that the lower bound $\Omega(N / p^{1/t})$ can be matched by a deterministic algorithm.
-
-\vgap
-
-Let us now turn to {\em skew-free hashing}, where each input in $\II(N)$ is a set $\I$ of $N$ elements, each with an integer key (computed by some hash function).
-It is guaranteed that every possible key is possessed by $O(N/p)$ elements of $\I$ (hence, skew ``free'').
-Initially, each machine receives at most $\ceil{N/p}$ elements.
-The goal is to move all the elements of the same key to a common machine.
-In the traditional MPC model, the problem can be settled in one round using load $\tO(N/p)$ w.h.p.\ \cite{bks14,bks17c}, where $\tO(\cdot)$ hides a $\polylog p$ factor.
-As skew-free hashing generalizes skew-free gathering\footnote{Recall that skew-free gathering has the additional constraint that all the elements except for the special ones need to have distinct keys.}, our lower bound on the latter indicates that oblivious skew-free hashing demands a load of nearly $\Omega(N)$, thus creating a huge separation from the traditional model.
-
-%It is worth pointing out that $O(N)$ is the worst possible cost under MPC and can be trivially attained --- obliviously --- by having all machines send their elements to the first machine.
-
-\extraspacing {\bf Single-Round Joins.} The hardness of skew-free hashing under communication obliviousness is alarming because the existing one-round join algorithms \cite{assu13,au11,bks13,bks14,bks17c,kbs16} in the (conventional) MPC model depend on a method named\footnote{Also known as the {\em hyper-cube} method.} \texttt{Share} \cite{au11,bks17c}, which uses hashing as a primitive.
-Because the community has come very close to resolving the one-round communication complexity of (natural) joins \cite{kbs16}, it would be a huge pity if the topic had to be re-opened in the name of obliviousness.
-
-\vgap
-
-Our second contribution is to show that, fortunately, joins' single-round communication complexities are not affected by obliviousness!
-To explain, let us first formalize the {\em join problem} under the oblivious setting.
-Fix an arbitrary integer $N \ge 1$ as the problem size.
-Let $\att$ be a finite set, where each element is called an {\em attribute}.
-Consider a non-empty set $\X \subseteq \att$ of attributes.
-A {\em tuple} over $\X$ is a function $\bm{u}: \X \rightarrow [N]$.
-For any non-empty subset $\Y \subseteq \X$, we define the {\em projection} of $\bm{u}$ onto $\Y$, denoted as $\bm{u}[\Y]$, as the tuple $\bm{v}$ over $\Y$ that satisfies $\bm{v}(Y) = \bm{u}(Y)$ for every attribute $Y \in \Y$.
-A {\em relation} $R$ is a set of tuples over the same set $\Z$ of attributes; we refer to $\Z$ as the {\em schema} of $R$ and represent it as $\schema(R)$.
-A {\em join} is a set $\Q$ of at least two relations such that $\sum_{R \in \Q} |R| = N$ (i.e., $N$ tuples in total in all the input relations).
-Define $\schema(\Q) = \bigcup_{R \in \Q} \schema(R)$.
-The result of $\Q$ is a relation over $\schema(\Q)$, formalized as:
-\myeqn{
-\join(\Q) = \set{\textrm{tuple $\bm{u}$ over $\schema(\Q)$} \mid \forall R \in \Q: \bm{u}[\schema(R)] \in R}. \nn
-}
-The {\em schema graph} of $\Q$ is the hypergraph $\G = (\V, \E)$ where
-\myeqn{
-    \V = \schema(\Q) \text{ and }
-    \E = \ldb \schema(R) \mid R \in \Q \rdb. \nn
-}
-Note that $\E$ is a multi-set because some relations in $\Q$ may have the same schema. Focusing on ``data complexity'', we consider only joins whose schema graphs have constant sizes.
-
-%The quantity $\max_{e \in \E} |e|$ is defined as the {\em arity} of $\G$.
-
-\vgap
-
-Given a hypergraph $\G = (\V, \E)$ and an integer $N \ge 1$ that is a multiple of $p$, we define the class of {\em $(\G, N)$-joins} as the set of joins that have schema graph $\G$ and problem size $N$.
-In the terminology of communication obliviousness, the collection $\II(N)$ comprises all the $(\G, N)$-joins of the same $\G$.
-Specifically, each input $\I$ of $\II(N)$ is the set of tuples in the relations of a $(\G, N)$-join $\Q$.
-We define a partition $\pi = (\I_1, ..., \I_p)$ of $\I$ to be {\em legal} if each $\I_i$, $i \in [p]$, consists of $N/p$ tuples.
-Given $\pi$, an algorithm $\A$ must output each tuple of $\join(\Q)$ on at least one machine.
-
-\vgap
-
-We show that any $(\G, N)$-join $\Q$ can be settled --- obliviously --- in a single round with load $\tO(N/p^{1/\psi})$ w.h.p.\ where $\psi \ge 1$ is the {\em edge quasi-packing number} \cite{kbs16} of $\G$; see Appendix~\ref{app:hypergraph} for the definition of $\psi$.
-The load complexity asymptotically matches that of the state-of-the-art single-round algorithm \cite{kbs16} in the traditional MPC model.
-Like \cite{kbs16}, our algorithm assumes that (i) $N \ge p^c$ where $c$ is a constant dependent on $\G$, and (ii) the machines are aware of certain ``heavy'' values that appear frequently in the relations of $\Q$.
-
-\vgap
-
-Our algorithm applies hashing in a more refined manner compared to \cite{kbs16}.
-The load $\tO(N/p^{1/\psi})$ stands as an intriguing contrast to our $\Omega(N)$ lower bound (up to a factor sub-polynomial in $p$) on skew-free hashing.
-This serves as encouraging evidence that hashing can still be powerful in designing single-round algorithms under communication obliviousness.
-For the join problem, our deployment of hashing circumvents the pitfall of skew-free gathering.
-Specifically, the deployment ensures that, for each tuple $\bm{u}$ in the join result, the $|\Q|$ tuples in $\set{\bm{u}[\schema(R)] \mid R \in \Q}$ can be transferred to a common machine (which can differ for each $\bm{u}$).
-Subtly, this is different from sending all tuples having an identical value on an attribute to the same machine, as required by skew-free gathering.
-
-\extraspacing {\bf Round-Doubling Compilation.} The above discussion has focused on single-round algorithms.
-Our next contribution is a {\em round compilation} method for multi-round algorithms that translates a traditional MPC algorithm $\A$ into the communication-oblivious model while preserving the number of rounds of $\A$ by a factor of 2 and its load by an arbitrarily small constant factor $\delta > 0$.
-Specifically, if $\A$ performs at most $\ell = \poly(p)$ rounds and, w.h.p., incurs a load of at most $L$, our method yields an oblivious algorithm $\A'$ that computes the same information as $\A$ in $2\ell$ rounds with a load $(1+\delta) L$.
-Our compilation succeeds w.h.p.\ as long as $L = \Omega(p \log p)$, a condition satisfied by nearly all the existing MPC algorithms we are aware of.
-
-\vgap
-
-We also study how much improvement can still be expected over our compilation method.
-First, it is clear that the blow-up factor of 2 in the round number cannot be improved in general.
-As discussed before, for skew-free gathering, it is easy to achieve load $O(N/p)$ in one round in the traditional MPC model, but under communication obliviousness, every single-round algorithm must entail a load of nearly $\Omega(N)$ with a constant probability.
-Thus, the remaining question is whether the condition $L = \Omega(p \log p)$ can be significantly relaxed.
-Our last contribution is to answer the question in the negative by establishing a hardness result on the following {\em token passing problem}.
-
-\vgap
-
-Fix arbitrary integers $N$ and $p$ such that $p \ge 2$ and $N = p \cdot \ceil{p^{1-\eps}}$, where the constant $\eps$ satisfies $0 < \eps < 1$.
-The input collection $\II(N)$ has only a single input $\I$, which is a set of $N$ elements among which there are $N / p = \ceil{p^{1-\eps}}$ special elements called {\em tokens}.
-In addition to $\I$, the problem takes another two parameters: $x$ and $y$, which are distinct integers in $[p]$.
-A partition $\pi = (\I_1, ..., \I_p)$ of $\I$ is {\em legal} if (i) $|I_i| = N/p$ for all $i \in [p]$, and (ii) the $N/p$ tokens are all placed on machine $x$ (i.e., $I_x$ consists of nothing but tokens).
-The objective of the problem is to move all the tokens to machine $y$.\footnote{Unlike skew-free gathering where all the special elements can be moved to an arbitrary machine, here all the tokens must be sent to machine $y$.}
-
-\vgap
-
-In the traditional model, the problem can be trivially solved in one round with load $L = O(p^{1-\eps})$: simply ask machine $x$ to send all the tokens to machine $y$.
-We prove that, if a two-round communication oblivious algorithm is required to succeed with probability at least 2/3, its load must be $\Omega(p^{1-\eps/2})$ with a constant probability.
-The polynomial gap between $\Omega(p^{1-\eps/2})$ and $O(p^{1-\eps})$ indicates that the condition $L = \Omega(p \log p)$ of our round-doubling compilation  is tight up to a factor sub-polynomial in $p$.
-In fact, this remains true even if the blow-up factor $1+\delta$ in load preservation is increased to $\polylog p$.
-
-\subsection{Previous Results and Relevance to Ours} \label{sec:intro-prev}
-
-Closely related to our work is a compilation method by Chan et al.\ \cite{ccls20}.
-Our compilation draws inspiration from theirs but makes several improvements.
-First, the approach of \cite{ccls20} was designed for the ``SODA MPC'' model \cite{ksv10}, which resembles the model in Section~\ref{sec:intro-prob} but does away the notion of ``load''.
-In that model, each machine is equipped with $s = \Omega(N/p)$ words of memory, and the amount of communication is not a main concern as long as each machine receives no more than $s$ words in each round from all other machines combined.
-The analysis of \cite{ccls20} was carried out under the condition of $s = N^\eps$ for some constant $\eps > 0$.
-When translated into our scenario, the analysis fails to capture the regime where $L \ll s$.
-Second, our compilation is substantially simpler, restores the clarity for the approach underneath \cite{ccls20} in the scenario where $L = \Omega(p \log p)$, and explicitly determines the blow-up factors $1+\delta$ and 2 (for preserving the load and number of rounds, respectively)\footnote{In \cite{ccls20}, these factors were hidden in big-$O$.}.
-Finally, our discussion (through the token passing problem) on the condition that $L$ needs to satisfy to enable round-doubling load-preserving compilation is new.
-
-\vgap
-
-It should be further noted that the study in \cite{ccls20} does not address one-round MPC algorithms.
-Through our round compilation method, one can see that a primary distinction between the conventional MPC model and its oblivious counterpart lies in the realm of one-round algorithms.
-Exploring fundamental problems that can characterize this difference represents an interesting direction.
-Our work can be seen as a step in that direction.
-
-\vgap
-
-As mentioned, the existing single-round algorithms for join processing \cite{assu13,au11,bks13,bks14,bks17c,kbs16} rely crucially on the \texttt{Share} method \cite{au11,bks17c}.
-Given a join $\Q$, \texttt{Share} assigns a hash function $h_X(\cdot)$ to each attribute $X \in \schema(\Q)$.
-For each tuple $\bm{u}$ in a relation $R \in \Q$, the machine initially storing $\bm{u}$ transmits it to a set of machines determined by the hash values in $\set{h_X(\bm{u}(X)) \mid X \in \schema(R)}$.
-Beame et al.\ \cite{bks17c} presented a sharp concentration bound on the load of \texttt{Share} when certain skew-free requirements are satisfied.
-Their bound is useful to our analysis and will be reviewed in Section~\ref{sec:join}.
-Koutris et al.\ \cite{kbs16} proved that any one-round MPC algorithm (oblivious or not) solving the join problem must incur a load of $\Omega(N/p^{1/\psi})$ with a constant probability, if each machine is aware of only the tuples in its local storage initially.
-
-\vgap
-
-There is an extensive body of literature on multi-round MPC algorithms; see \cite{ajr+17,aksu15,akz19,hyt19,kst22,hcww21,hk20,h21,hy19,qt21,t18b} and the references therein.
-None of those algorithms was designed to achieve communication obliviousness, but all of them satisfy the condition $L = \Omega(p \log p)$ and, therefore, can be made oblivious by our compilation method.
-It is worth noting that our compilation also works on the one-round algorithms in \cite{assu13,au11,bks13,bks14,bks17c,kbs16}, but the algorithms after compilation will require two rounds.
-
-\section{Hardness of Skew-Free Gathering} \label{sec:gather}
-
-This section will discuss the skew-free gathering problem defined in Section~\ref{sec:intro-prob} and serves as a proof of the paper's first main result:
-
-\begin{theorem} \label{thm:gathering}
-    For the skew-free gathering problem parameterized by $t$, any communication-oblivious one-round algorithm succeeding with probability at least $2/3$ must entail a load of $\Omega(N/p^{1/t})$ with at least a constant probability.
-    %Furthermore, this is tight: there is a communication oblivious algorithm that always succeeds and guarantees a load of $O(N/p^{1/t})$ deterministically.
-\end{theorem}
-
-In Appendix~\ref{app:sf-gathering-alg}, we show that the above lower bound $\Omega(N/p^{1/t})$ can be matched deterministically.
-
-\vgap
-
-To prove Theorem~\ref{thm:gathering}, let us start by identifying a set $\Pi$ of $\binom{p}{t}$ legal partitions of the (sole) input $\I \in \II(N)$ as follows.
-First, choose a set $Z$ of distinct integers $z_1, z_2, ..., z_t$ in $[p]$; we will refer to $Z$ as the {\em seed machine set}.
-Second, place $N/p$ special elements on machine $z_j$ for each $j \in [t]$, and place $N/p$ innocuous elements on each machine $i \in [p] \setminus Z$, ensuring that every  element is on exactly one machine.
-This defines a legal partition in $\Pi$.
-As there are $\binom{p}{t}$ ways to choose $Z$, the size of $\Pi$ is $\binom{p}{t}$.
-We require a one-round algorithm $\A$ to be oblivious only on $\Pi$, namely, its communication pattern follows the same probabilistic distribution when it is executed on $\pi \in \Pi$.
-We will argue that the load of $\A$ must still be $\Omega(N/p^{1/t})$ with at least a constant probability if $\A$ succeeds on every $\pi \in \Pi$ with probability at least $2/3$.
-This is sufficient for validating the lower bound in Theorem~\ref{thm:gathering}.
-
-\extraspacing {\bf Conditional Expected Load.} As $\A$ executes in one-round, its communication pattern can be fully characterized by a single $p \times p$ matrix $\mb{L}$ where $\mb{L}[i,j]$ ($i, j \in [p]$) is the length of the message $M[i,j]$ that machine $i$ sends to machine $j$.
-The load of $\A$ can now be calculated as
-\myeqn{
-    \load(\mb{L}) &=&
-    \max_{i=1}^p
-    \Big(
-    \sum_{j=1}^p \mb{L}[i, j] + \mb{L}[j, i]
-    \Big).
-    \label{eqn:gathering:load-matrix}
-}
-When $\A$ is randomized, the $M[i,j]$ of all $i, j \in [p]$ are random variables, and hence so are $\mb{L}$ and
-$\load(\mb{L})$. Our objective is to prove that $\Pr[\load(\mb{L}) = \Omega(N/p^{1/t})]$ is at least a constant.
-
-\vgap
-
-By communication obliviousness, the distribution of $\mb{L}$ is the same for all the legal partitions in $\Pi$. Define $\L$ as the set of $p \times p$ matrices $\mb{\Lambda}$ satisfying
-\myitems{
-    \item $\Pr[\mb{L} = \mb{\Lambda}] > 0$, namely, $\A$ exhibits the patten $\mb{L} = \mb{\Lambda}$ with a non-zero probability;
-    \item $\max_{i=1}^p \sum_{j=1}^p \mb{\Lambda}[i, j] \le N/p^{1/t}$;
-    \item $\max_{i=1}^p \sum_{j=1}^p \mb{\Lambda}[j, i] \le N/p^{1/t}$.
-}
-Note that when $\mb{L} = \mb{\Lambda}$, {\em every} machine sends/receives at most $N/p^{1/t}$ words. The subsequent discussion will assume
-\myeqn{
-    \Pr[\mb{L} \in \L] \ge 1/2. \label{eqn:gathering:small-load-1/2}
-}
-Note that whenever $\mb{L} \notin \L$, some machine communicates at least $N/p^{1/t}$ words and hence $\load(\mb{L}) \ge N/p^{1/t}$.
-Thus, $\Pr[\mb{L} \notin \L] \ge 1/2$ immediately gives $\Pr[\load(\mb{L}) \ge N/p^{1/t}] \ge 1/2$, thereby verifying the claim in Theorem~\ref{thm:gathering}.
-
-\vgap
-
-We will aim to validate the following lower bound on the expectation of $\load(\mb{L})$ subject to the condition $\mb{L} \in \L$:
-\myeqn{
-    \expt[\load(\mb{L}) \mid \mb{L} \in \L] &\ge& c \cdot N/p^{1/t}
-    \label{eqn:gathering:goal}
-}
-for some constant $0 < c < 1$.
-The above implies\footnote{For every $\mb{L} \in \L$, it holds that $\load(\mb{L}) \le 2 N/p^{1/t}$.
-If \eqref{eqn:gathering:goal-implication} incorrect, then $\expt[\load(\mb{L}) \mid \mb{L} \in \L] <  \fr{2N}{p^{1/t}} \cdot \fr{c}{4} + \fr{c}{2} \fr{N}{p^{1/t}} (1 - \fr{c}{4}) < c N / p^{1/t}$, contradicting \eqref{eqn:gathering:goal}.}
-\myeqn{
-    \Pr[\load(\mb{L}) \ge (c/2) \cdot N/p^{1/t} \mid \mb{L} \in \L] \ge c/4.
-    \label{eqn:gathering:goal-implication}
-}
-Combining \eqref{eqn:gathering:small-load-1/2} and \eqref{eqn:gathering:goal-implication} yields $\Pr[\load(\mb{L}) \ge \fr{c}{2} N/p^{1/t}] \ge \fr{c}{4} \cdot \fr{1}{2} = c/8$, thus establishing Theorem~\ref{thm:gathering}.
-
-\extraspacing {\bf Heavy Senders.} Suppose that the algorithm $\A$ exhibits a pattern $\mb{L} = \mb{\Lambda} \in \L$.
-Given a machine $i \in [p]$, we call machine $j \in [p] \setminus \set{i}$ a {\em heavy sender} for machine $i$ under $\mb{\Lambda}$ if $\mb{\Lambda}[j,i] \ge N/p$, namely, machine $j$ sends at least $N/p$ words to machine $i$.
-Define
-\myeqn{
-    H(\mb{\Lambda}, i) = \text{number of heavy senders for machine $i$ under $\mb{\Lambda}$}. \label{eqn:gathering:H}
-}
-It is easy to see that
-\myeqn{
-    H(\mb{\Lambda}, i) \le p^{1-1/t}
-    \label{eqn:gathering:bound-H}
-}
-because otherwise machine $i$ would receive more than $(N/p) \cdot p^{1-1/t} = N/p^{1/t}$ words, contradicting $\mb{\Lambda} \in \L$.
-
-\vgap
-
-Our analysis will revolve around heavy senders. We will argue that at least one machine $i \in [p]$ satisfies
-\myeqn{
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot H(\mb{\Lambda}, i)
-    &=&
-    \Omega(p^{1-1/t}).
-    \label{eqn:gathering:goal2}
-}
-As $\load(\mb{\Lambda}) \ge (N/p) \cdot H(\mb{\Lambda}, i)$, the above leads to
-\myeqn{
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \load(\mb{\Lambda})
-    =
-    \Omega\Big(\fr{N}{p} \cdot p^{1-1/t} \Big)
-    =
-    \Omega(N /p^{1/t}).
-    \label{eqn:gathering:goal2-implication}
-}
-This will prove \eqref{eqn:gathering:goal} because
-\myeqn{
-    \expt[\load(\mb{L}) \mid \mb{L} \in \L]
-    &=&
-    \fr{1}{\Pr[\mb{L} \in \L]} \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \load(\mb{\Lambda})
-    \nn
-}
-which is at least \eqref{eqn:gathering:goal2-implication}.
-
-%as long as $H(\mb{\Lambda}, i) \ge 1$
-
-\extraspacing {\bf Two Types of Legal Partitions.} The rest of the section will focus on proving the correctness of \eqref{eqn:gathering:goal2}.
-Earlier we have identified a set $\Pi$ of $\binom{p}{t}$ legal partitions, each characterized by a seed machine set $Z = \set{z_1, ..., z_t}$; for convenience, we will use $Z$ to denote the corresponding legal partition when no ambiguity can arise.
-Now, fix any legal partition $Z$.
-The algorithm $\A$ succeeds on $Z$ if and only if at least one of the following occurs:
-\myitems{
-    \item machines $z_1, z_2, ..., z_t$ all send their special elements to a machine $k \in [p] \setminus Z$;
-    \item there is some $j \in [t]$ such that machine $z_j$ receives all the special elements from every other machine in $Z$.
-}
-Echoing the above, given any $Z$ and any matrix $\mb{\Lambda} \in \L$, we define
-\myitems{
-    \item $X(Z, \mb{\Lambda})$ as an indicator variable, which equals 1 if there exists some $k \in [p] \setminus Z$ such that $\mb{\Lambda}[z_j, k] \ge N/p$ for every $j \in [t]$, or 0 otherwise;
-
-    \item $Y(Z, \mb{\Lambda})$ as an indicator variable, which equals 1 if there exists some $j \in [t]$ such that $\mb{\Lambda}[z_k, z_j] \ge N/p$ for every $k \in [t] \setminus \set{j}$, or 0 otherwise.
-}
-
-%
-
-When exhibiting the pattern $\mb{L} = \bm{\Lambda}$, the algorithm $\A$ succeeds on $Z$ only if
-\myeqn{
-    X(Z, \mb{\Lambda}) + Y(Z, \mb{\Lambda}) \ge 1. \nn
-}
-We declare:
-\myeqn{
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
-    (X(Z, \mb{\Lambda}) + Y(Z, \mb{\Lambda}))
-    &\ge&
-    1/6.
-    \label{eqn:gathering:1/6}
-}
-To see why, notice that if the above did not hold, then $\A$ would have probability less than $1/6$ succeeding on $Z$ when $\mb{L} \in \L$.
-Even if $\A$ always succeeds on $Z$ when $\mb{L} \notin \L$, the overall success probability on $Z$ would still be less than $1/6 + \Pr[\mb{L} \notin \L]$, which is less than 2/3 because of \eqref{eqn:gathering:small-load-1/2}.
-This violates the requirement that $\A$ must succeed with probability at least $2/3$ on $Z$.
-
-\vgap
-
-We will refer to a legal partition $Z$ as
-\myitems{
-    \item an {\em X-partition} if $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
-    X(Z, \mb{\Lambda}) \ge 1/12$, or
-    \item an {\em Y-partition} if $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
-    Y(Z, \mb{\Lambda}) \ge 1/12$.
-}
-Note that $Z$ must be classified as at least one of the two types because of \eqref{eqn:gathering:1/6}.
-Thus, if $\XX$ (resp., $\YY$) represents the set of all X- (resp., Y-) partitions, we have $|\XX| + |\YY| \ge |\Pi| = \binom{p}{t}$.
-The next lemma relates each type of partitions to heavy-sender machines.
-
-\begin{lemma} \label{lmm:gathering:total-X-Y}
-    For any $\mb{\Lambda} \in \L$, it holds that
-    \myeqn{
-        \sum_{Z \in \XX} X(Z, \mb{\Lambda}) &\le& \sum_{i \in [p]} \binom{H(\mb{\Lambda}, i)}{t} \label{eqn:gathering:total-X-Y:help1} \\
-        \sum_{Z \in \YY} Y(Z, \mb{\Lambda}) &\le& \sum_{i \in [p]} \binom{H(\mb{\Lambda}, i)}{t-1}. \label{eqn:gathering:total-X-Y:help2}
-    }
-\end{lemma}
-
-The reader is reminded that $\binom{x}{y}$ equals 0 if $x < y$. We will prove the first inequality here and the second one in Appendix~\ref{app:proof:thm:gathering}.
-
-\begin{proof}[Proof of \eqref{eqn:gathering:total-X-Y:help1}]
-    Our proof adopts a counting argument.
-Initially, create an empty set $S_i$ for each $i \in [p]$.
-Process each $Z = \set{z_1, ..., z_t} \in \XX$ as follows.
-If $X(Z, \mb{\Lambda}) = 0$, do nothing.
-Otherwise, under the communication pattern $\mb{L} = \bm{\Lambda}$, there is at least one machine $k \in [p] \setminus Z$ that receives $N/p$ special elements from machine $z_j$ for all $j \in [t]$.
-Thus, machines $z_1, z_2, ..., z_t$ are heavy senders for machine $k$ under $\mb{\Lambda}$.
-Hence, $Z$ represents a possible way to choose $t$ machines from the $H(\mb{\Lambda}, k)$ heavy senders for machine $k$ under $\mb{\Lambda}$.
-Accordingly, we add $Z$ to $S_k$.
-After all the sets $Z \in \XX$ have been processed, the left hand side of \eqref{eqn:gathering:total-X-Y:help1} is bounded by $\sum_{i \in [p]} |S_i|$.
-On the other hand, for each $i \in [p]$, the size $|S_i|$ cannot exceed the total number of ways of choosing $t$ machines from the $H(\mb{\Lambda}, i)$ heavy senders for machine $i$ under $\mb{\Lambda}$.
-Inequality \eqref{eqn:gathering:total-X-Y:help1} now follows.
-\end{proof}
-
-\vgap
-
-The remainder of the argument proceeds differently depending on the number of $X$-partitions.
-
-\extraspacing {\bf At Least $\fr{1}{2} \binom{p}{t}$ X-Partitions.} That is, $|\XX| \ge \fr{1}{2} \binom{p}{t}$, which, together with the definition of X-partition, yields
-\myeqn{
-    \sum_{Z \in \XX} \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
-    X(Z, \mb{\Lambda}) \ge \fr{1}{24} \binom{p}{t}.
-    \label{eqn:gathering:X-help1}
-}
-We can now derive
-\myeqn{
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} H(\mb{\Lambda},i)^t
-    &\ge&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \binom{H(\mb{\Lambda},i)}{t}
-    \nn \\
-    \explain{by \eqref{eqn:gathering:total-X-Y:help1}}
-    &\ge&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{Z \in \XX} X(Z, \mb{\Lambda})
-    \nn \\
-    &=&
-    \sum_{\mb{\Lambda} \in \L} \sum_{Z \in \XX} \Pr[\mb{L} = \mb{\Lambda}] \cdot  X(Z, \mb{\Lambda})
-    \nn \\
-    \explain{by \eqref{eqn:gathering:X-help1}}
-    &\ge&
-    \fr{1}{24} \binom{p}{t}.
-    \label{eqn:gathering:X-help2}
-}
-This gives:
-\myeqn{
-    %\text{LHS of \eqref{eqn:gathering:goal2}}
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \sum_{i \in [p]} H(\mb{\Lambda},i)
-    &=&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^t}{H(\mb{\Lambda},i)^{t-1}}
-    \nn \\
-    \explain{by \eqref{eqn:gathering:bound-H}}
-    &\ge&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^t}{p^{\fr{(t-1)^2}{t}}}
-    \nn \\
-    \explain{by \eqref{eqn:gathering:X-help2}}
-    &\ge&
-    \fr{1}{24} \binom{p}{t} / p^{\fr{(t-1)^2}{t}}
-    \nn \\
-    &=&
-    \Omega\Big(
-        \fr{p^t}{p^{(t-1)^2/t}}
-    \Big)
-    =
-    \Omega\Big(p^{2-1/t} \Big).
-    \nn
-}
-Therefore, at least one $i \in [p]$ satisfies $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot H(\mb{\Lambda},i) = \Omega(p^{2-1/t} / p ) = \Omega (p^{1-1/t})$, as claimed in \eqref{eqn:gathering:goal2}.
-
-\extraspacing {\bf At most $\fr{1}{2} \binom{p}{t}$ X-Partitions.} Because $|\XX| + |\YY| \ge \binom{p}{t}$, the fact $|\XX| \le \binom{p}{t}/2$ indicates $|\YY| \ge \binom{p}{t}/2$.
-Following a derivation similar to the previous case, we can show that at least one $i \in [p]$ satisfies \eqref{eqn:gathering:goal2}.
-With the details presented in Appendix~\ref{app:proof:thm:gathering}, we now conclude the proof of Theorem~\ref{thm:gathering}.
-
-\section{Single-Round Oblivious Joins} \label{sec:join}
-
-We now turn our attention to the join problem defined in Section~\ref{sec:intro:ours}. Let $\Q$ be the input $(\G, N)$-join. Given a value $x \in [N]$, we call it
-\myitems{
-    \item {\em heavy} if at least $N / p^2$ tuples of a relation have value $x$ under an attribute, or formally, $\exists R \in \Q, X \in \schema(R)$ s.t.
-$R$ has at least $N / p^2$ tuples $\bm{u}$ with $\bm{u}(X) = x$;
-
-    \item or {\em light}, otherwise.
-} 
-There can be $O(p^2)$ heavy values in total.
-We make the {\em heavy-statistics assumption} that each machine knows all the heavy values.
-For this purpose, each legal partition (of the tuples in the relations of $\Q$; see Section~\ref{sec:intro:ours}) is augmented with an ``information field'' for each machine, which contains all the heavy values.
-To prevent an adversary from distinguishing the $(\G, N)$ joins by inspecting the information field's length, the field is always padded with dummy data to the same length of $O(p^2)$ words.
-
-\vgap
-
-We will establish our second main result:
-
-\begin{theorem} \label{thm:join}
-    Consider any constant-size hypergraph $\G = (\V, \E)$.
-Let $N$ be a multiple of $p$ satisfying $N \ge p^{1 + 2\alpha + 1/\psi}$ where $\alpha = \max_{e \in \E} |e|$ (called the arity of $\G$) and $\psi$ is the edge quasi-packing number of $\G$.
-Under the heavy-statistics assumption, there is a communication oblivious algorithm that,  given any $(\G, N)$-join $\Q$, operates in a single round with load $\tO(N / p^{1/\psi})$ and computes $\join(\Q)$ w.h.p..
-\end{theorem}
-
-Our result can be compared to an algorithm of \cite{kbs16}, which has load $\tO(N / p^{1/\psi})$ w.h.p.\ but is not communication oblivious.
-That algorithm also makes the heavy-statistics assumption although their frequency threshold of a heavy value is $N/p$, rather than $N/p^2$ in our context.
-
-%$\alpha = \max_{e \in \E} |e|$ (this is the arity of $\G$), and
-
-\vgap
-
-The rest of the section serves as a proof of Theorem~\ref{thm:join}.
-We will first review the \ttt{Share} method \cite{au11,bks17c} in Section~\ref{sec:join-share}.
-After that, we will present our join algorithm in Sections~\ref{sec:join-alg1} and \ref{sec:join-alg2} to attain the performance guarantees claimed.
-
-%and analyzing its behavior in Section~\ref{sec:join-analysis}.
-
-\subsection{The Share Method} \label{sec:join-share}
-
-Let $R^*$ be a relation with $\schema(R^*) = \set{X_1, X_2, ..., X_r}$, and $M$ be an integer satisfying $|R^*| \le M \le N$.
-Recall that, as defined in Section~\ref{sec:intro:ours}, each tuple $\bm{u} \in R^*$ is a function from $\schema(R^*)$ to $[N]$.
-For any $\Y \subseteq \schema(R^*)$, define:
-%{\em the degree of $R^*$ on $\Y$} as
-\myeqn{
-    \deg_\Y(R^*) &=& \max_{\text{tuple $\bm{u}$ over $\Y$}} \big|\set{\bm{v} \in R^* \mid \bm{v}[\Y] = \bm{u}} \big| \label{eqn:join-degree}
-}
-that is, the maximum number of tuples in $R^*$ having the same projection on $\Y$.
-Suppose that we assign to each attribute $X_i$, $i \in [r]$, a positive integer $s_i$ --- called the {\em share} of $X_i$ --- satisfying the following {\em skew-free condition}:
-\myeqn{
-    \text{for any non-empty $\Y \subseteq \schema(R^*)$, $\deg_\Y(R^*) \le \fr{M}{{\prod_{i \in [r]: X_i \in \Y} s_i}}$}.
-    \label{eqn:share-skewfree}
-}
-Fix any constant $\alpha \ge r$.
-For each $i \in [r]$, choose independently a {\em perfectly random hash function} $h_i: [\alpha N] \rightarrow [s_i]$, that is, $h_i$ is uniformly sampled from the $(s_i)^{\alpha N}$ possible functions from $[\alpha N]$ to $[s_i]$.
-Given $\bm{b} = (b_1, b_2, ..., b_r)$ $\in [s_1] \times [s_2] \times ...
-\times [s_r]$, define
-\myeqn{
-    \text{\em bin $\bm{b}$}
-    &=&
-    \set{\bm{u} \in R^* \mid \forall i \in [r], h_i(\bm{u}(X_i)) = b_i}.
-    \label{eqn:join-share:bin}
-}
-Note that there are in total $\prod_{i=1}^r s_i$ bins. The following concentration bound is due to Beame et al.\ \cite{bks17c}:
-
-\begin{lemma} \label{lmm:share}
-    %[\cite{bks17c}]
-    For any constant $c > 0$, it holds with probability at least $1 - (1/p^*)^c$ that the sizes of all the bins are bounded by $O((\log p^* / \log\log p^*)^r \cdot M / p^*)$, where $p^* = \prod_{i=1}^r s_i$.
-The constant in the big-$O$ depends on $c$.
-\end{lemma}
-
-Strictly speaking, Beame et al.\ \cite{bks17c} proved the lemma only for the case where $M = |R^*|$, which is not enough for our purposes.
-Fortunately, it is not difficult to extend their result to any $M \in [|R^*|, N]$, as shown in Appendix~\ref{app:proof:thm:join}.
-
-\subsection{A Non-Oblivious Join Algorithm} \label{sec:join-alg1}
-
-This subsection will present an algorithm, which is {\em not} communication oblivious, to compute the result of a $(\G, N)$-join $\Q$ with load $\tO(N / p^{1/\psi})$ w.h.p..
-The reason for describing this non-oblivious version is to allow the reader to draw a direct comparison with the solution of \cite{kbs16}.
-As will be clear, a key new idea is to apply the concentration bound of Lemma~\ref{lmm:share} locally on each machine, rather than globally on the entire join (as was done in \cite{kbs16}).
-This in turn requires us to decrease the ``heavy threshold'' from $N/p$ to $N/p^2$.
-Our algorithm ensures a new property (to be given in Lemma~\ref{lmm:join-comm}), which makes it easy to make the algorithm communication oblivious, as we do in the next subsection.
-
-\vgap
-
-Given a relation $R \in \Q$ and machine id $i \in [p]$, let $R^{(i)}$ be the set of tuples of $R$ that are initially stored on machine $i$.
-Define $R^{(i,\heavy)}$ as the set of tuples $\bm{u} \in R^{(i)}$ using {\em only} heavy values.
-We know
-\myeqn{
-    |R^{(i,\heavy)}| &=& O(p^{2\alpha}) \nn
-    %\label{eqn:join-size-R-i-heavy}
-}
-because $R^{(i)}$ has at most $\alpha$ attributes (recall that $\alpha$ is the arity of the hypergraph $\G$), and there can be $O(p^2)$ heavy values.
-
-\vgap
-
-Before proceeding, the reader should be familiarized with the content of Appendix~\ref{app:hypergraph}, in particular, the notions of ``residual graph'' and ``fractional vertex cover''.
-Next, let us take an arbitrary proper subset $\Z \subset \V$ and identify an optimal fractional vertex cover $w_\Z:$ $\V$ $\setminus$ $\Z \rightarrow [0, 1]$ of the residual graph $\G_\Z = (\V_\Z, \E_\Z)$.
-Define
-\myeqn{
-    \tau(\G_\Z) &=& \sum_{X \in \V \setminus \Z} w_\Z(X).
-    \label{eqn:join-tau_Z}
-}
-To each attribute $X \in \V$, we assign a positive integer $s_X(\Z)$ --- the share of $X$ for $G_\Z$ --- as follows:
-\myitems{
-    \item if $X \in \Z$, then $s_X(\Z) = 1$;
-    \item otherwise, $s_X(\Z) = \lf p^{w_\Z(X) / \tau(\G_\Z)} \rf$.
-}
-For any non-empty subset $\Y \subseteq \V$, we show in Appendix~\ref{app:proof:thm:join}:
-\myeqn{
-    \prod_{X \in \Y} s_X(\Z)
-    \le
-    p.
-    \label{eqn:join-help1}
-}
-Furthermore, for any relation $R \in \Q$, we show again in Appendix~\ref{app:proof:thm:join}:
-\myeqn{
-    \prod_{X \in \schema(R)} s_X(\Z) &=& \Omega\Big(p^{1/\tau(G_\Z)} \Big).
-    \label{eqn:join-help2}
-}
-
-Henceforth, set
-\myeqn{
-    M = N/p. \label{eqn:join-M}
-}
-For each $i \in [p]$ and each relation $R \in \Q$ satisfying $\schema(R) \setminus \Z \ne \emptyset$, we define $R^{(i,\Z)}$ as the set of tuples $\bm{u} \in R^{(i)}$ such that
-\myitems{
-    \item $\bm{u}(X)$ is heavy for every attribute $X \in \schema(R) \cap \Z$;
-    \item $\bm{u}(X)$ is light for every attribute $X \notin \schema(R) \cap \Z$.
-}
-For any non-empty subset $\Y \subseteq \schema(R)$, let us observe
-\myeqn{
-    \deg_\Y (R^{(i,\Z)})
-    \le \fr{M}{\prod_{X \in \Y} s_X(\Z)}.
-    \label{eqn:join-alg-skewfree}
-}
-Indeed, if $\Y \subseteq \Z$, then $\prod_{X \in \Y} s_X(\Z) = 1$ in which case \eqref{eqn:join-alg-skewfree} holds because $\deg_\Y (R^{(i,\Z)}) \le |R^{(i)}| \le N/p = M$.
-Otherwise, identify an arbitrary attribute $X \in \Y \setminus \Z$.
-Then, we can derive
-\myeqn{
-    \deg_\Y (R^{(i,\Z)}) \le \deg_{\set{X}} (R^{(i,\Z)}) \le \deg_{\set{X}} (R^{(i)}) \le N/p^2 \nn
-}
-where the last inequality is due to the definition of light value.
-By \eqref{eqn:join-help1}, $N/p^2 \le M / \prod_{X \in \Y} s_X(\Z)$, from which the correctness of \eqref{eqn:join-alg-skewfree} follows.
-We can therefore conclude that $R^{(i,\Z)}$ fulfills the skew-free condition prescribed in \eqref{eqn:share-skewfree}.
-
-\vgap
-
-We now elaborate on the join algorithm.
-Before receiving the $(\G,N)$-join $\Q$, we perform some preprocessing for each proper subset $\Z \subset \V$.
-First, obtain $s_X(\Z)$ --- the share of $X$ for $G_\Z$ --- for every attribute $X \in \V$ as explained earlier.
-For each $X \in \V$, independently pick a perfectly random hash function $h_{\Z, X}: [\alpha N] \rightarrow [s_X(\Z)]$ (recall that $\alpha$ is the arity of $\G$).
-Store a copy of all these functions on every machine.
-The cartesian product of $\bigtimes_{X \in \V} [s_X(\Z)]$ has a size of $\prod_{X \in \V} s_X(\Z)$, which is at most $p$ (see \eqref{eqn:join-help1}).
-Each element in the cartesian product can be regarded as a tuple $\bm{b}_\Z$, where $\bm{b}_\Z(X)$ is an integer in $[s_X(\Z)]$ for each $X \in \V$.
-We assign a distinct machine to each possible $\bm{b}_\Z$ lexicographically: machine $i \in [p]$ assigned to the $\bm{b}_\Z$ that ranks, in lexicographic order, the $i$-th among all the elements of $\bigtimes_{X \in \V} [s_X(\Z)]$.
-Note that some machines (precisely, $p - \prod_{X \in \V} s_X(\Z)$ machines) are not assigned to any cartesian product elements.
-
-\vgap
-
-After the $(\G,N)$-join $\Q$ has been given, each machine --- say the one with id $i \in [p]$ --- carries out the following steps in parallel:
-
-\myitems{
-    \item {{\bf S1:}} Broadcast $R^{(i,\heavy)}$.
-
-    \vgap
-
-    \item {{\bf S2:}} For each proper subset $\Z \subset \V$ and every relation $R \in \Q$ satisfying $\schema(R) \setminus \Z \ne \emptyset$, the machine examines every tuple $\bm{u} \in R^{(i,\Z)}$.
-For each attribute $X \in \schema(R)$, the machine computes the hash value $h_{\Z,X}(\bm{u}(X))$.
-Then, it sends $\bm{u}$ to every machine $\bm{b}_\Z \in \bigtimes_{X \in \V} [s_X(\Z)]$ satisfying $\bm{b}_\Z(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \schema(R)$.
-    %The number of such machines is $\prod_{X \in \V \setminus \schema(R)} s_X(\Z)$, which is $O(p^{1-1/\tau(\G_\Z)})$ by \eqref{eqn:join-help3}.
-}
-
-The above steps require one round of communication.
-Then, every machine joins all the tuples received.
-To prove  correctness, consider any tuple $\bm{u} \in \join(\Q)$.
-If $\bm{u}$ uses a heavy value on every attribute, then it is produced by all machines due to Step {\bf S1}.
-Otherwise, suppose that $\bm{u}$ uses heavy values only on those attributes of $\Z$ for some $\Z \subset \V$.
-Then, due to Step {\bf S2}, it must be produced at the machine corresponding to the element $\bm{b}_\Z \in \bigtimes_{X \in \V} [s_X(\Z)]$ where $\bm{b}_\Z(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \V$.
-
-\vgap
-
-We will prove that the load incurred is $O(p^{2\alpha + 1}) + \tO(N/p^{1/\psi})$ w.h.p., which is $\tO(N/p^{1/\psi})$ for $N \ge p^{1 + 2\alpha + 1/\psi}$.
-Clearly, Step {\bf S1} entails load $O(p^{2\alpha + 1})$.
-Next, we will show that, for any particular $\Z \subset \V$, {\bf S2} generates a load of $\tO(N / p^{1/\tau(\G_\Z)})$ w.h.p., where $\tau(\G_\Z)$ is given in \eqref{eqn:join-tau_Z}.
-It will then follow that the overall load of {\bf S2} is $\tO(N/p^{1/\psi})$ because $\psi$ equals the maximum $\tau(\G_\Z)$ of all $\Z \subset \V$ (see \eqref{eqn:hypergraph:psi}) Appendix~\ref{app:hypergraph}) and $\V$ has a constant number of subsets.
-
-\vgap
-
-The lemma below characterizes the behavior of our algorithm.
-
-\begin{lemma} \label{lmm:join-comm}
-    Consider an arbitrary proper subset $\Z \subset \V$.
-The following statement holds w.h.p.: for any relation $R \in \Q$ and distinct machine ids $i, j \in [p]$, machine $i$ sends $\tO(M / p^{1/\tau(\G_\Z)})$ tuples in $R^{(i,\Z)}$ to machine $j$.
-\end{lemma}
-
-\begin{proof}
-     As explained previously, relation $R^{(i,\Z)}$ satisfies the skew-free condition \eqref{eqn:join-alg-skewfree} with respect to the shares $\set{s_X(\Z) \mid X \in \V}$.
-Set $p^* = \prod_{X \in \schema(R)} s_X(\Z)$.
-As $|R^{(i,\Z)}| \le M \le N$, Lemma~\ref{lmm:share} asserts that the following event occurs w.h.p.:
-
-     \minipg{0.9\linewidth}{
-        $\tO(\fr{M}{p^*})$ tuples $\bm{u} \in R^{(i,\Z)}$ are hashed to each element (a.k.a., bin) $\bm{b}$ in the cartesian product $\bigtimes_{X \in \schema(R)} [s_X(\Z)]$, namely, $\bm{b}(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \schema(R)$.
-     }
-
-     \noindent Under the above event, $\tO(M/p^*)$ tuples $\bm{u} \in R^{(i,\Z)}$ are hashed to each element $\bm{b}_\Z \in \bigtimes_{X \in \V} [s_X(\Z)]$, i.e., such a tuple $\bm{u}$ satisfies $\bm{b}_\Z(X) = h_{\Z,X}(\bm{u}(X))$ for all $X \in \schema(R)$.
-Hence, if the machine $j$ stated in the lemma is assigned to an element in  $\bigtimes_{X \in \V} [s_X(\Z)]$, then machine $i$ sends $\tO(M/p^*)$ tuples of $R^{(i,\Z)}$ to machine $j$; otherwise, machine $i$ sends no tuple of $R^{(i,\Z)}$ to machine $j$.
-
-     \vgap
-
-     The lemma now follows from $p^* = \Omega(p^{1/\tau(\G_\Z)})$ (see \eqref{eqn:join-help2}).
-\end{proof}
-
-As $\Q$ has $O(1)$ relations, Lemma~\ref{lmm:join-comm} indicates that every machine sends and receives $\tO(p \cdot M / p^{1/\tau(\G_\Z)}) = \tO(N / p^{1/{\tau(\G_\Z)}})$ words in Step {\bf S2} w.h.p.\ in processing the subset $\Z$.
-
-\extraspacing {\bf Remark.} Lemma~\ref{lmm:join-comm} is a feature of our algorithm that is not shared by the algorithm of \cite{kbs16}, but is crucial for turning the algorithm into an oblivious counterpart, as discussed next.
-
-\subsection{Making the Algorithm Oblivious} \label{sec:join-alg2}
-
-Our algorithm in Section~\ref{sec:join-alg1} enjoys a simple communication pattern.
-It can be modified into a communication-oblivious version by padding enough dummy data to make the length of each message consistent with the ``worst'' case.
-
-\vgap
-
-Let us start with Step {\bf S1}, where machine $i \in [p]$ transmits the same message, which contains $R^{(i,\heavy)}$ for each $R \in \Q$, to every other machine.
-This message can have a maximum length of $\mit{len}_\mit{heavy}$ $= O(p^{2\alpha})$ words.
-Whenever this message is shorter than $\mit{len}_\mit{heavy}$, expand it with dummy data into length $\mit{len}_\mit{heavy}$.
-
-\vgap
-
-In Step {\bf S2}, for each proper subset $Z \subset \V$, machine $i \in [p]$ transmits a possibly different message to each other machine $j \in [p]$, whose length is bounded by an integer $\mit{len}_\Z = \tO(M/p^{1/\tau(\G_\Z)})$ w.h.p.\ (Lemma~\ref{lmm:join-comm}).
-Whenever this message is shorter than $\mit{len}_\Z$ --- including the case of an empty message --- expand it with dummy data into length $\mit{len}_\Z$.
-However, it is possible for the message to be actually longer than $\mit{len}_\Z$ (this happens when the high-probability event in Lemma~\ref{lmm:join-comm} does not occur).
-In that scenario, machine $i$ (arbitrarily) trims the message at the length of $\mit{len}_\Z$ and sends the trimmed message to machine $j$ anyway; when this happens, we say that the algorithm {\em errs}.
-
-\vgap
-
-In the above modified algorithm, every machine sends precisely $\mit{len}_\mit{heavy} + \sum_{\Z \subset \V} \mit{len}_\Z$ words to every other machine.
-It therefore exhibits a deterministic communication pattern for all $(\G,N)$-joins --- this is true regardless of whether the algorithm errs.
-The load is always $p \cdot (\mit{len}_\mit{heavy} + \sum_{\Z \subset \V} \mit{len}_\Z)$, which is $\tO(N/p^{1/\psi})$ as analyzed in the previous subsection.
-The algorithm computes the join result correctly if it does not err (an adversary cannot tell whether the algorithm has erred because everything is encrypted).
-Due to Lemma~\ref{lmm:join-comm}, by choosing constants appropriately, we can reduce the probability that the algorithm errs to at most $1/p^c$ for an arbitrarily large constant $c$.
-This completes the proof of Theorem~\ref{thm:join}.
-
-\section{Two-Round Compilation} \label{sec:compile}
-
-This section will concentrate on multi-round MPC algorithms, for which our main contribution is a compilation method that establishes the following theorem:
-
-\begin{theorem} \label{thm:compile}
-    Fix an arbitrary constant $\delta$ satisfying $0 < \delta < 1$.
-Let $\A$ be an algorithm under the traditional MPC model that operates within $\ell = \poly(p)$ rounds and entails a load at most $L$ w.h.p..
-If $L  = \Omega(p \log p)$ where the hidden constant depends on $\delta$, there is a communication-oblivious algorithm that performs $2\ell$ rounds, requires a load at most $(1 + \delta) L$, and computes the same information as $\A$ on every machine w.h.p..
-\end{theorem}
-
-As explained in Section~\ref{sec:intro:ours}, the round blow-up factor 2 in the above theorem is the best possible, regardless of the constant $\delta$.
-The next theorem indicates that the condition $L = \Omega(p \log p)$ can no longer be relaxed significantly, again as explained in Section~\ref{sec:intro:ours}.
-
-\begin{theorem} \label{thm:token}
-    Fix an arbitrary constant $\eps$ satisfying $0 < \eps < 1$.
-For the token-passing problem (defined in Section~\ref{sec:intro:ours}) parameterized by $\eps$, any communication-oblivious two-round algorithm succeeding with probability at least $2/3$ must demand a load of $\Omega(p^{1-\eps/2})$ with at least a constant probability.
-\end{theorem}
-
-We will present the algorithmic procedure of our compilation method in the rest of the section, but defer its analysis to Appendix~\ref{app:compile}.
-The proof of Theorem~\ref{thm:token} is provided in Appendix~\ref{app:thm:token}.
-
-\extraspacing {\bf Message Routing.} To prove Theorem~\ref{thm:compile}, we will tackle the {\em message routing problem} defined as follows.
-Suppose that, for each pair of $i, j \in [p]$, machine $i$ needs to send a message $\mb{M}[i,j]$ to machine $j$ (if $i = j$, then $\mb{M}[i,j]$ is an empty message).
-Denote by $\mb{L}[i,j]$ the length of $\mb{M}[i,j]$ in the number of words.
-For each $i \in [p]$, we have
-\myeqn{
-    \sum_{j \in [p]} \mb{L}[i,j] + \mb{L}[j,i] \le L \label{eqn:compile-len-constraint}
-}
-namely, every machine sends and receives no more than $L$ words in total. We want to design a communication-oblivious algorithm $\A_\mit{route}$ with all the requirements below:
-\myitems{
-    \item It runs in two rounds.
-    \item For distinct machines $i, j \in [p]$, machine $i$ sends (precisely) $L'$ words to machine $j$ in each round, where $L' \le (1+\delta) L/p$.
-    \item W.h.p., all the messages are successfully delivered at the end of the second round.
-}
-
-Equipped with $\A_\mit{route}$, we can prove Theorem~\ref{thm:compile} as follows.
-Recall (from Section~\ref{sec:intro-prob}) that each round of $\A$ runs in two phases where in the first phase each machine prepares the messages to be sent out in the second phase.
-We treat the second phase as an instance of the message routing problem and apply $\A_\mit{route}$ to deliver the messages.
-W.h.p., at the end of $\A_\mit{route}$, every machine acquires all the information that it would have obtained at the second phase of $\A$, and can thus perform the local computation as demanded by $\A$.
-Because $\A$ always terminates within $\ell = \poly(p)$ rounds, our compilation succeeds on all its rounds w.h.p..
-
-\extraspacing {\bf Our Algorithm.} Set $\delta' = \ceil{16/\delta}$.
-We chop each message $\mb{M}[i,j]$ into segments, each of which has $\delta'$ words --- if the last segment is shorter than $\delta'$ words, expand it with dummy words to make its length $\delta'$.
-The number of segments is $\ceil{\mb{L}[i,j] / \delta'}$.
-Each segment, referred to as a {\em packet} henceforth, is augmented with three fields: (i) {\em sender}, i.e., the value $i$ (sender machine id), (ii) {\em recipient}, i.e., the value $j$ (recipient machine id), and (iii) {\em sequence number}, i.e., the $t$-th packet of $\mb{M}[i,j]$ has sequence number $t$.
-Packets may arrive at machine $j$ in an arbitrary order; the sender and sequence-number fields allow the packets to be put back into the original sequence in $\mb{M}[i,j]$.
-Every packet is thus $\delta' + 3$ words in length.
-
-\vgap
-
-Denote by $P_i$ the set of packets obtained from $\mb{M}[i,1]$, $\mb{M}[i,2]$, ..., $\mb{M}[i,p]$ combined. Let $|P_i|$ be the number of those packets. Then
-\myeqn{
-    |P_i| \le \sum_{j=1}^p \Big(1 + \fr{\mb{L}[i,j]}{\delta'} \Big) \le p + \fr{L}{\delta'}.
-    \label{eqn:compile-P_i-size}
-}
-where the last inequality used \eqref{eqn:compile-len-constraint}. Set
-\myeqn{
-    \lambda = \lc \left(1 + \fr{L} {\delta' p}\right)\left(1 + \fr{\delta}{2}\right) \rc.
-    \label{eqn:compile-lambda}
-}
-
-In the first round, each machine --- say machine $i \in [p]$ --- in parallel carries out the steps below:
-\myitems{
-    \item To each packet in $P_i$, assign a {\em relay machine} chosen from $[p]$ uniformly at random.
-Let $P_i(j)$ be the set of packets in $P_i$ assigned to the same relay machine $j \in [p]$.
-We use $|P_i(j)|$ to represent the number of packets in $P_i(j)$.
-
-    \vgap
-
-    \item For each $j \in [p]$, if $|P_i(j)| < \lambda$, add enough dummy packets to $P_i(j)$ to make the number of packets therein exactly $\lambda$.
-If $|P_i(j)| > \lambda$, discard (arbitrarily) some packets from $P_i(j)$ to shrink $|P_i(j)|$ to $\lambda$; however, in this case, we say that the algorithm {\em errs}.
-
-    \vgap
-
-    \item For each $j \in [p]$, send $P_i(j)$ to machine $j$ (of course, for $j = i$, ``sending'' $P_i(j)$ requires no communication).
-}
-
-For each $i \in [p]$, define
-\myeqn{
-    P^*_i = \bigcup_{j \in [p]} P_j(i) \nn
-}
-namely, the set of all packets to be ``relayed'' by machine $i$.
-These packets are transmitted to machine $i$ in the first round.
-Machine $i$ sorts the packets by recipient field.
-For each $j \in [p]$, let $P^*_i(j)$ be the set of packets in $P^*_i$ whose recipient fields are $j$.
-In the second round, machine $i$, in parallel to others, proceeds as follows:
-\myitems{
-    \item For each $j \in [p]$, if $|P^*_i(j)| < \lambda$, add dummy packets to $P^*_i(j)$ to increase $|P^*_i(j)|$ to $\lambda$.
-If $|P^*_i(j)| > \lambda$, discard (arbitrarily) some packets from $P^*_i(j)$ to shrink $|P^*_i(j)|$ to $\lambda$; in this case, we say that the algorithm {\em errs}.
-
-    \vgap
-
-    \item For each $j \in [p]$, send $P^*_i(j)$ to machine $j$.
-}
-This completes our 2-round algorithm for message routing.
-
-\vgap
-
-In each round, for any distinct $i, j \in [p]$, machine $i$ sends to machine $j$ precisely
-\myeqn{
-    L' = \lambda \cdot (\delta' + 3) \label{eqn:compile-L'}
-}
-words.
-It is rudimentary to verify that $L' \le (1+\delta) L /p$ when $p$ is greater than a constant.
-Therefore, our 2-round algorithm exhibits a deterministic communication pattern and has a load of $L' (p-1) < (1+\delta) L$.
-If the algorithm does not err, every message $\mb{M}[i,j]$ is successfully delivered to machine $j \in [p]$.
-In Appendix~\ref{app:compile}, we prove that the algorithm errs with probability at most $1/p^c$ where $c$ can be an arbitrarily large constant chosen before running the algorithm.
-This completes the proof of Theorem~\ref{thm:compile}.
-
 \bibliographystyle{plainurl}% the mandatory bibstyle
 \bibliography{ref}
-
-\appendix
-
-\section*{Appendix}
-
-\section{Basic Concepts of Hypergraphs} \label{app:hypergraph}
-
-Let $\G = (\V, \E)$ be a hypergraph with $\E$ being a multi-set where each element, called an {\em edge}, is a non-empty subset of $\V$ (the set of {\em vertices}).
-A {\em fractional edge packing} is a function $w: \E \rightarrow [0, 1]$ (mapping each edge $e \in \E$ to a weight $w(e)$ between 0 and 1) such that, for any attribute $X \in \V$, we have $\sum_{e \in \E: X \in e} w(e) \le 1$ (i.e., the weight sum of all the edges containing $X$ is at most 1).
-The {\em total weight} of $w$ is $\sum_{e \in \E} w(e)$, i.e., the weight sum of all the edges in $\E$.
-The {\em fractional edge packing number} of $\G$, denoted as $\tau(\G)$, is defined as the maximum total weight of all the fractional edge packings of $\G$.
-
-\vgap
-
-Dual to fractional edge packing is the notion of {\em fractional vertex cover}, which is a function $w: \V \rightarrow [0, 1]$ (mapping each vertex $X \in \V$ to a weight $w(X)$ between 0 and 1) such that, for any edge $e \in \E$, we have $\sum_{X \in \V: X \in e} w(X) \ge 1$ (i.e., the weight sum of all the vertices in $e$ is at least 1).
-The {\em total weight} of $w$ is $\sum_{X \in \V} w(X)$, i.e., the weight sum of all the vertices in $\V$.
-The {\em fractional vertex cover number} of $\G$ is defined as the minimum total weight of all the fractional vertex covers of $\G$.
-It can be shown \cite{su97} that the fractional vertex cover number of $\G$ is always identical to the fractional edge packing number $\tau(\G)$.
-
-\vgap
-
-Given a proper subset $\Z \subset \V$, we define $\G_\Z$ as the hypergraph obtained by ``removing'' the attributes of $\Z$ from $\G$. Specifically:
-\myitems{
-    \item The vertex set of $\G_\Z$ is $\V \setminus \Z$.
-    \item The edges of $\G_\Z$ are decided as follows: for each edge $e \in \G$, we add the edge $e \setminus \Z$ to $\G_\Z$ if $e \setminus \Z \ne \emptyset$.
-}
-We will refer to $\G_\Z$ as a {\em residual graph} of $\G$. Specially, if $\Z = \emptyset$, then $\G_\Z = \G$.
-
-\vgap
-
-For each proper subset $\Z \subset \V$, the residual graph $\G_\Z$ has its own fractional edge packing number $\tau(\G_\Z)$.
-The maximum such number of all possible $\Z$ is the {\em edge quasi-packing number} $\psi(\G)$ of $\G$, or formally:
-\myeqn{
-    \psi(\G) &=&
-    \max_{\Z \subset \V} \tau(\G_\Z).
-    \label{eqn:hypergraph:psi}
-}
-Due to the equivalence between fractional edge packing number and fractional vertex cover number, the value $\psi(\G)$ can alternatively be defined as the maximum fractional vertex cover number of all residual graphs $G_\Z$ of $\G$.
-
-\section{A Skew-Free Gathering Algorithm} \label{app:sf-gathering-alg}
-
-In this section, we present a deterministic communication-oblivious algorithm to solve the skew-free gathering problem (parameterized by integer $t$) with load $O(N / p^{1/t})$, matching the lower bound in Theorem~\ref{thm:gathering}.
-
-\vgap
-
-Our algorithm is in essence an adaptation of the ``hyper-cube algorithm'' in \cite{au11}.
-Define $h = \floor{p^{1/t}}$, and order the points in the $t$-dimensional space $[h]^t$ lexicographically.
-For each $i \in [h^t]$, assign machine $i$ to the $i$-th point in $[h]^t$ under the lexicographic order.
-Note that some machines (precisely, $p - h^t$ machines) are not assigned to any point.
-
-\vgap
-
-Divide domain $[p]$ into $h$ disjoint intervals $I_1, I_2, ..., I_h$, ordered in ascending order of their starting points, such that:
-\myitems{
-    \item each of $I_1, ..., I_{h-1}$ covers $\floor{p/h}$ integers;
-    \item $I_h$ covers $p - (h-1) \floor{p/h}$ integers.
-}
-For each interval $I_c$ with $c \in [h]$, we use $|I_c|$ to denote the number of integers covered by $I_c$.
-Consider any point $(c_1, c_2, ..., c_t) \in [h]^t$.
-Remember that the point corresponds to a machine.
-We associate the machine with a $t$-dimensional box $I_{c_1} \times I_{c_2} \times ...
-\times I_{c_t}$ and will refer to $I_{c_j}$ the machine's {\em projection} on dimension $j \in [t]$.
-
-\vgap
-
-Given a legal partition of the skew-free gathering problem, each machine --- say the one with id $i \in [p]$ --- sends the $N/p$ elements in its local storage to every machine whose projection on {\em at least} one dimension covers the value $i$ (there are less than $t \cdot h^{t-1}$ such machines).
-This completes the description of our algorithm.
-
-\vgap
-
-To see the algorithm's correctness, suppose that machines where the special elements are initially placed have ids $z_1, z_2, ..., z_t \in [p]$.
-For each $j \in [t]$, let $c_j$ be the integer in $[h]$ such that the interval $I_{c_j}$ covers $z_j$.
-Our algorithm ensures that the machine associated with the $t$-dimensional box $I_{c_1} \times I_{c_2} \times ...
-\times I_{c_t}$ receives all the special elements from the machines $z_1$, ..., $z_t$.
-The algorithm's communication obliviousness can be verified from the fact that, for each $i \in [p]$, machine $i$ sends exactly $N/p$ elements to a set of machines that is determined in a way independent of how the input elements are initially distributed.
-
-\vgap
-
-It remains to analyze the algorithm's load.
-In terms of sending, each machine delivers $N/p$ elements to less than $t \cdot h^{t-1} = O(h^{t-1}) = O(p^{(t-1)/t})$ machines and, thus, transmits $O((N/p) \cdot p^{(t-1)/t}) = O(N/p^{1/t})$ words.
-In terms of receiving, the machine corresponding to the $t$-dimensional box $I_{c_1} \times I_{c_2} \times ...
-\times I_{c_t}$ receives at most $(N/p) \cdot \sum_{j=1}^t |I_{c_j}|$ elements.
-Among all the intervals $I_1, ..., I_h$, the interval $I_h$ is the longest and covers
-\myeqn{
-    p - (h-1) \lf \fr{p}{h} \rf < p - (h-1) \left( \fr{p}{h} - 1 \right)
-    < \fr{p}{h} + h = O(p^{1-1/t}) \nn
-}
-integers. Each machine, therefore, receives $O((N/p) \cdot p^{1-1/t}) = O(N / p^{1/t})$ words. This proves that our algorithm has a load of $O(N / p^{1/t})$.
-
-\section{Completing the Proof of Thm.~\ref{thm:gathering}} \label{app:proof:thm:gathering}
-
-\noindent {\bf Proof of Inequality (\ref{eqn:gathering:total-X-Y:help2}).} We adopt a counting argument similar to the one used to prove \eqref{eqn:gathering:total-X-Y:help1}.
-Initially, create an empty set $S_i$ for each $i \in [p]$.
-Then, process each $Z = (z_1, ..., z_t) \in \YY$ as follows.
-If $Y(Z, \mb{\Lambda}) = 0$, do nothing.
-Otherwise, under the communication pattern $\mb{L} = \mb{\Lambda}$, at least one machine $z_k$, for some $k \in [t]$, receives $N/p$ special elements from each machine $z_j$ with $j \in [t] \setminus \set{k}$.
-This means that all the machines in $Z \setminus \set{z_k}$ are heavy senders for machine $z_k$ under $\mb{\Lambda}$.
-Thus, $Z \setminus \set{z_k}$ represents a possible way to choose $t-1$ machines from the $H(\mb{\Lambda}, z_k)$ heavy senders for machine $z_k$ under $\mb{\Lambda}$.
-Accordingly, we add $Z$ to $S_k$.
-After all the sets $Z \in \YY$ have been processed, the left-hand side of \eqref{eqn:gathering:total-X-Y:help2} is bounded by $\sum_{i \in [p]} |S_i|$.
-On the other hand, for each $i \in [p]$, the size $|S_i|$ cannot exceed the total number of ways of choosing $t - 1$ machines from the $H(\mb{\Lambda}, i)$ heavy senders for machine $i$ under $\mb{\Lambda}$.
-Inequality \eqref{eqn:gathering:total-X-Y:help2} now follows.
-
-\extraspacing {\bf Proof of \eqref{eqn:gathering:goal2} When $|\YY| \ge \fr{1}{2} \binom{p}{t}$.} By the definition of the Y-partition, the fact $|\YY| \ge \fr{1}{2} \binom{p}{t}$ yields
-\myeqn{
-    \sum_{Z \in \YY} \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot
-    Y(Z, \mb{\Lambda}) \ge \fr{1}{24} \binom{p}{t}.
-    \label{eqn:gathering:Y-help1}
-}
-Then, we can obtain
-\myeqn{
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} H(\mb{\Lambda},i)^{t-1}
-    &\ge&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \binom{H(\mb{\Lambda},i)}{t-1}
-    \nn \\
-    \explain{by \eqref{eqn:gathering:total-X-Y:help2}}
-    &\ge&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{Z \in \YY} Y(Z, \mb{\Lambda})
-    \nn \\
-    &=&
-    \sum_{\mb{\Lambda} \in \L} \sum_{Z \in \YY} \Pr[\mb{L} = \mb{\Lambda}] \cdot  Y(Z, \mb{\Lambda})
-    \nn \\
-    \explain{by \eqref{eqn:gathering:Y-help1}}
-    &\ge&
-    \fr{1}{24} \binom{p}{t}.
-    \label{eqn:gathering:Y-help2}
-}
-This leads to:
-\myeqn{
-    %\text{LHS of \eqref{eqn:gathering:goal2}}
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \sum_{i \in [p]} H(\mb{\Lambda},i)
-    &=&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^{t-1}}{H(\mb{\Lambda},i)^{t-2}}
-    \nn \\
-    \explain{by \eqref{eqn:gathering:bound-H}}
-    &\ge&
-    \sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot \sum_{i \in [p]} \fr{H(\mb{\Lambda},i)^{t-1}}{p^{\fr{(t-1)\cdot(t-2)}{t}}}
-    \nn \\
-    \explain{by \eqref{eqn:gathering:Y-help2}}
-    &\ge&
-    \fr{1}{24} \binom{p}{t} / p^{\fr{(t-1)\cdot(t-2)}{t}}
-    \nn \\
-    &=&
-    \Omega\Big(
-        \fr{p^t}{p^{\fr{(t-1)\cdot(t-2)}{t}}}
-    \Big)
-    =
-    \Omega\Big(p^{3-2/t} \Big).
-    \nn
-}
-Therefore, at least one $i \in [p]$ satisfies $\sum_{\mb{\Lambda} \in \L} \Pr[\mb{L} = \mb{\Lambda}] \cdot H(\mb{\Lambda},i) = \Omega(p^{3-2/t} / p ) = \Omega (p^{1-1/t})$, as claimed in \eqref{eqn:gathering:goal2}.
-
-\section{Completing the Proof of Thm.~\ref{thm:join}} \label{app:proof:thm:join}
-
-\noindent {\bf Proof of Lemma~\ref{lmm:share}.} As mentioned, Beame et al.\ \cite{bks17c} has proved Lemma~\ref{lmm:share} for the case where $M = |R^*|$.
-We will utilize their result as a black box to prove the lemma under $|R^*| < M \leq N$.
-
-\vgap
-
-The number of distinct values taken by the tuples in $R^*$ is at most $|R^*|\cdot |schema(R^*)| \leq \alpha |R^*|$.
-Hence, we can find in the domain $[\alpha N]$ a set $S$ of $\alpha N - \alpha |R^*|$ values that are not taken by any tuple of $R^*$.
-From $|R^*| < M \leq N$, we know
-\myeqn{
-    |S| = \alpha N - \alpha |R^*| > \alpha N - (\alpha -1)N -|R^*| = N - |R^*| \geq M - |R^*|. \nn
-}
-Identify (arbitrarily) $M - |R^*|$ distinct values from $S$ and, for each such value $x$, create a tuple $\bm{u}$ over $\schema(R^*)$ with $\bm{u}(X) = x$ for each $X \in \schema(R^*)$.
-Let $R'$ be the relation obtained by inserting these $M - |R^*|$ tuples into $R^*$.
-Note that $|R'|$ has precisely $M$ tuples.
-
-\vgap
-
-It is easy to verify that $\deg_\Y(R^*) = \deg_\Y(R')$ for every non-empty subset $\Y \subseteq \schema(R')$.
-With this, the skew-free condition \eqref{eqn:share-skewfree} tells us that, for any non-empty subset $\Y \subseteq \schema(R')$, we must have:
-\myeqn{
-    \deg_\Y(R') \le \fr{|R'|}{{\prod_{i \in [r]: X_i \in \Y} s_i}}. \nn
-}
-
-Recall that, for each $i \in [r]$, we have chosen a perfectly random hash function $h_i: [\alpha N] \rightarrow [s_i]$.
-Given $\bm{b} = (b_1, b_2, ..., b_r)$ $\in [s_1] \times [s_2] \times ...
-\times [s_r]$, define
-\myeqn{
-    \text{\em bin $\bm{b'}$}
-    &=&
-    \set{\bm{u} \in R' \mid \forall i \in [r], h_i(\bm{u}(X_i)) = b_i}.
-    \label{eqn:proof:thm:join:help1}
-}
-Clearly, bin $\bm{b'}$ contains all the tuples in the bin $\bm{b}$ defined in \eqref{eqn:join-share:bin}.
-By the result of \cite{bks17c}, we know that every bin $\bm{b'}$ in \eqref{eqn:proof:thm:join:help1} contains $O((\log p^* / \log\log p^*)^r \cdot |R'| / p^*) = O((\log p^* / \log\log p^*)^r \cdot M / p^*)$ tuples, where $p^* = \prod_{i=1}^r s_i$.
-It thus follows that every bin $\bm{b}$ in \eqref{eqn:join-share:bin} must have a size $O((\log p^* / \log\log p^*)^r \cdot M / p^*)$.
-
-\extraspacing {\bf Proof of Equation~\eqref{eqn:join-help1}.} If $\Y \subseteq \Z$, then $\prod_{X \in \Y} s_X(\Z) = 1 \le p$. Otherwise:
-\myeqn{
-    \prod_{X \in \Y} s_X(\Z) =
-    \prod_{X \in \Y \setminus \Z} s_X(\Z)
-    \le
-    \prod_{X \in \Y \setminus \Z} p^{w_\Z(X) / \tau_\Z}
-    \le
-    p.
-    \nn
-}
-
-\extraspacing {\bf Proof of Equation~\eqref{eqn:join-help2}.} Let $e_R(\Z) = \schema(R)\setminus \Z$. Note that $e_R(\Z)$ is an edge in the residual graph $\G_\Z$. We have
-\myeqn{
-    \prod_{X \in \schema(R)} s_X(\Z)
-    &=&
-    \prod_{X \in e_R(\Z)} s_X(\Z) \cdot \prod_{X \in \Z} s_X(\Z) \nn\\
-    &=& \prod_{X \in e_R(\Z)} s_X(\Z) \nn \\
-    &=& \prod_{X \in e_R(\Z)} \lf p^{w_\Z(X) / \tau(\G_\Z)} \rf \nn \\
-    \explain{as $p^{w_\Z(X) / \tau(\G_\Z)} \geq 1$}&\ge& \prod_{X \in e_R(\Z)} \fr{p^{w_\Z(X) / \tau(\G_\Z)}}{2} \nn \\
-   \explain{as $|e_R(\Z)|\leq |\schema(R)|\leq \alpha $} &\ge&
-   \fr{1}{2^\alpha} \prod_{X \in e_R(\Z)} p^{w_\Z(X) / \tau(\G_\Z)} \nn \\
-    &\ge& p^{1/\tau(\G_\Z)} / 2^{\alpha} \nn
-}
-where the last inequality used the fact that $w_\Z$ is a fractional vertex cover of $G_\Z$ (and, hence, $\sum_{X \in {e_R(\Z)}} w_\Z(X) \ge 1$).
-
-\section{Completing the Proof of Thm.~\ref{thm:compile}} \label{app:compile}
-
-\noindent {\bf Chernoff Bounds.} The following Chernoff bounds (proved in \cite{stl12}) will be useful:
-
-\begin{lemma} \label{lmm::chernoff}
-    Let $X_1, X_2, ..., X_f$ be independent Bernoulli random variables that are identically distributed. Define $X = \sum_{i \in [f]} X_i$. For any $\gamma \geq 2$, it holds that
-    \myeqn{
-        \Pr[X \geq \gamma \cdot \expt[X]] &\leq&  \exp(-\gamma \cdot \expt[X] / 6). \label{eqn:compile:chernoff1}
-    }
-    For any $0 < \gamma < 1$, it holds that
-    \myeqn{
-        \Pr[X \geq (1 + \gamma) \expt[X]] &\leq&  \exp(- \gamma^2 \cdot \expt[X] / 3). \label{eqn:compile:chernoff2}
-    }
-\end{lemma}
-
-Fix an arbitrary real value $\gamma$ satisfying $0 < \gamma < 1$ and any positive real values $p$ and $\beta$. Let us specialize the setup of random variables in the above lemma as follows:
-\myitems{
-    \item $\Pr[X_i = 1] = 1/p$ for each $i \in [f]$;
-    \item $f \le F$ where $F \ge \beta \cdot p \log p$.
-}
-We claim:
-\myeqn{
-    \Pr[X \geq (1 + \gamma) F/p] &\leq&  \exp(- \Omega(\gamma^2 \beta \cdot \log p)) \label{eqn:compile:chernoff3}
-}
-where the hidden constant in the big-$\Omega$ does not depend on $\gamma$, $p$, or $\beta$. Next, we will prove the claim by distinguishing two cases.
-
-\vgap
-
-{\em Case 1: $\expt[X] \ge F / (2p)$.} We have:
-\myeqn{
-    \Pr[X \geq (1 + \gamma) F/p] &\le& \Pr[X \geq (1 + \gamma) \expt[X]] \nn \\
-    \explain{by \eqref{eqn:compile:chernoff2}} &\leq&  \exp(- \gamma^2 \cdot \expt[X] / 3) \nn \\
-    &\leq&  \exp(- \gamma^2 \cdot F / (6p)) \nn \\
-    &\leq&  \exp(- (\gamma^2 \beta / 6) \cdot \log p) \nn
-}
-as claimed.
-
-{\em Case 2: $\expt[X] < F / (2p)$.} We can derive:
-\myeqn{
-    \Pr\left[X \geq (1 + \gamma) F/p\right] &=&
-    \Pr\left[X \geq \fr{(1 + \gamma)}{\expt[X] / (F/p)} \cdot \expt[X] \right] \nn \\
-    \explain{by \eqref{eqn:compile:chernoff1}} &\leq&  \exp(- (1 + \gamma) \cdot F/(6p)) \nn \\
-    &\leq&  \exp(- (\beta / 6) \cdot \log p) \nn
-}
-as claimed.
-
-\extraspacing {\bf Completing the Analysis of Our Message-Routing Algorithm.} As discussed in Section~\ref{sec:compile}, it remains to prove that our algorithm errs with probability at most $1/p^c$ where $c$ can be an arbitrarily large constant chosen before running the algorithm.
-
-\vgap
-
-For the algorithm to err in the first round, $|P_i(j)|$ must exceed $\lambda$ (see its value in \eqref{eqn:compile-lambda}) for some $i, j \in [p]$.
-Set $f = |P_i|$ and $F = p + L / \delta'$; we thus have $f \le F$ by \eqref{eqn:compile-P_i-size}.
-For each $k \in [f]$, introduce a Bernoulli random variable $X_k$ that equals 1 if the $k$-th element of $P_i$ is relayed to machine $j$, and 0 otherwise.
-Thus, $\Pr[X_k = 1] = 1/p$, and $|P_i(j)| = \sum_{k \in [f]} X_k$.
-
-\vgap
-
-Suppose that $L \ge \beta \cdot p \log p$ where $\beta$ is a constant to be decided later. Note that this means $F > L / \delta' \ge (\beta/\delta') p \log p$.
-We can derive:
-\myeqn{
-    \Pr[|P_i(j)| \ge \lambda]
-    &\le&
-    \Pr \left[|P_i(j)| \ge \left(1 + \fr{L}{\delta' p} \right) (1+\delta/2) \right] \nn \\
-    &=&
-    \Pr \left[|P_i(j)| \ge (1+\delta/2) \cdot F/p \right] \nn \\
-    \explain{by \eqref{eqn:compile:chernoff3}}
-    &\le&
-    \exp(- \Omega(\delta^2 \cdot (\beta/\delta') \cdot \log p))  \nn \\
-    &=&
-    \exp(- \Omega(\delta^3 \cdot \beta \cdot \log p))
-    \label{eqn:compile:chernoff4}
-}
-which can be made less than $1/p^{c_0}$ for an arbitrarily large constant $c_0$ by setting $\beta$ sufficiently large (recall that $\delta$ is a constant).
-
-\vgap
-
-How likely the algorithm errs in the second round is somewhat less obvious.
-For each $j \in [p]$, denote by $P^+_j$ the set of packets --- produced from $\mb{M}[1,j]$, $\mb{M}[2,j]$, ..., $\mb{M}[p,j]$ combined --- destined for machine $j$ (i.e., those packets have $j$ as the recipient).
-Because of \eqref{eqn:compile-len-constraint}, $|P^+_j|$, the number of packets in $P^+_j$, is at most $p + L/\delta'$, following a derivation similar to that in \eqref{eqn:compile-P_i-size}.
-The algorithm errs in the second round only if $|P^*_i(j)| > \lambda$ for some $i, j \in [p]$.
-Observe that $|P^*_i(j)|$ is precisely the number of packets in $P^+_j$ that have $i$ as their relay machines.
-Now, set $f = |P^+_j|$ and $F = p + L / \delta'$; we thus have $f \le F$ as just explained.
-For each $k \in [f]$, introduce a Bernoulli random variable $X_k$ that equals 1 if the $k$-th element of $P^+_j$ is relayed to machine $i$, and 0 otherwise.
-Thus, $\Pr[X_k = 1] = 1/p$, and $|P^*_i(j)| = \sum_{k \in [f]} X_k$.
-Following a derivation similar to the one leading to \eqref{eqn:compile:chernoff4}, we obtain
-\myeqn{
-    \Pr[|P^*_i(j)| \ge \lambda]
-    &\le&
-    \exp(- \Omega(\delta^3 \cdot \beta \cdot \log p))   \label{eqn:compile:chernoff5}
-}
-
-We can now conclude from \eqref{eqn:compile:chernoff4} and \eqref{eqn:compile:chernoff5} that, when $L \ge \beta \cdot p \log p$ for a sufficiently large $\beta$, our algorithm errs with a probability at most $1 / p^c$.
-
-\section{Proof of Theorem \ref{thm:token}} \label{app:thm:token}
-
-Let us start by constructing a set of $p(p-1)$ parameterized inputs.
-Denote by $\Pi$ the set of $p(p-1)$ ordered pairs $(x, y) \in [p] \times [p]$ satisfying $x \ne y$.
-For every $x \in [p]$, build a legal partition $\pi_x$ for token-passing by placing $\ceil{p^{1-\eps}}$ tokens on machine $x$ and placing $\ceil{p^{1-\eps}}$ innocuous elements on every other machine.
-Each pair $(x, y) \in \Pi$ then specifies a distinct parameterized input --- with legal partition $\pi_x$ and parameters $x, y$ --- for which the objective is to move all the tokens from machine $x$ to machine $y$.
-
-\vgap
-
-Let $\A$ be a two-round communication oblivious algorithm that succeeds with probability at least $2/3$ on each of the $p(p-1)$ parameterized inputs defined by $\Pi$.
-We will prove that $\A$ must entail a load of $\Omega(p^{1-\eps/2})$ with probability at least $1/2$.
-
-\vgap
-
-As $\A$ performs two rounds, its communication pattern can be represented as two $p \times p$ matrices $\mb{L}_1$ and $\mb{L}_2$.
-Specifically, for each $r = 1$ and $2$, the value $\mb{L}_r[i, j]$ (where $i, j \in [p]$) is the number of words that machine $i$ sends to machine $j$ in round $r$.
-If $\A$ is randomized, then $\mb{L}_1$ and $\mb{L}_2$ are random variables.
-For each $r =1$ and 2, we define a $p \times p$ matrix $\mb{L}^*_r$ from $\mb{L}_r$ as follows:
-\myitems{
-    \item $\mb{L}^*_r[i, j] = \mb{L}_r[i, j]$ for any $i, j \in [p]$ satisfying $i \ne j$.
-    \item $\mb{L}^*_r[i, i] = \ceil{p^{1-\eps}}$ for each $i \in [p]$.
-}
-We remind the reader that $\mb{L}_r[i,i] = 0$ for all $i \in [p]$, by definition of communication pattern. The matrix $\mb{L}^*_r$ thus defined is a random variable decided by $\mb{L}_r$.
-
-\vgap
-
-By communication obliviousness, the distribution of $(\mb{L}_1, \mb{L}_2)$ is the same for all the parameterized inputs specified by the pairs in $\Pi$.
-Let $\L$ be the set of ordered pairs $(\mb{\Lambda}_1$, $\mb{\Lambda}_2)$ satisfying
-\myitems{
-    \item each of $\mb{\Lambda}_1$ and $\mb{\Lambda}_2$ is a $p \times p$ matrix;
-
-    \item $\Pr[(\mb{L}^*_1$, $\mb{L}^*_2) = (\mb{\Lambda}_1$, $\mb{\Lambda}_2)] > 0$, namely, $\A$ exhibits with a non-zero probability a pattern $(\mb{L}_1$, $\mb{L}_2)$ such that the corresponding $(\mb{L}^*_1$, $\mb{L}^*_2)$ equals $(\mb{\Lambda}_1$, $\mb{\Lambda}_2)$;
-
-    \item the following inequality holds for all $k \in [p]$:
-    \myeqn{
-        \Big(\sum_{i \in [p]} \mb{\Lambda}_1 [i,k] \Big) \cdot \Big(\sum_{j \in [p]} \mb{\Lambda}_2 [k,j] \Big) \le p^{2-\eps}/24. \label{eqn:thm:token:help-1}
-    }
-}
-The core of our argument is to prove
-\myeqn{
-    \Pr[(\mb{L}^*_1, \mb{L}^*_2) \notin \L] \ge 1/2.
-    \label{eqn:thm:token:help-2}
-}
-Note that whenever $(\mb{L}^*_1, \mb{L}^*_2) \notin \L$, it follows from the definition in \eqref{eqn:thm:token:help-1} that there is at least one $k \in [p]$ such that
-\myeqn{
-    \Big(\sum_{i \in [p]} \mb{L}^*_1 [i,k] \Big) \cdot \Big(\sum_{j \in [p]} \mb{L}^*_2 [k,j] \Big) > p^{2-\eps}/24 \nn
-}
-under which one of the following inequalities must be true:
-\myitems{
-    \item $\sum_{i \in [p]} \mb{L}^*_1 [i,k] > p^{1-\eps/2} / \sqrt{24}$, or
-    \item $\sum_{j \in [p]} \mb{L}^*_2 [k,j] > p^{1-\eps/2} / \sqrt{24}$.
-}
-In the former case, we have
-\myeqn{
-    \sum_{i \in [p]} \mb{L}_1 [i,k] &=& \Big(\sum_{i \in [p]} \mb{L}^*_1 [i,k] \Big) - \mb{L}^*_1 [k,k] \nn \\
-    &=& \Big(\sum_{i \in [p]} \mb{L}^*_1 [i,k] \Big) - \ceil{p^{1-\eps}}
-    = \Omega \Big(p^{1-\eps/2}\Big). \nn
-}
-This means that the load of $\A$ is $\Omega(p^{1-\eps/2})$ as $\sum_{i \in [p]} \mb{L}_1 [i,k]$ is the number of words received by machine $k$ in the first round.
-In the latter case, following a similar derivation one can obtain $\sum_{j \in [p]} \mb{L}_2 [k,j] = \Omega(p^{1-\eps/2})$.
-This also implies that $\A$ has load $\Omega(p^{1-\eps/2})$ because $\sum_{j \in [p]} \mb{L}_2 [k,j]$ is the number of words sent by machine $k$ in the second round.
-Therefore, with probability at least $1/2$, algorithm $\A$ demands a load of $\Omega(p^{1-\eps/2})$, which completes the proof of Theorem~\ref{thm:token}.
-
-\extraspacing {\bf Proof of Inequality \eqref{eqn:thm:token:help-2}.} Assume for contradiction purposes that \eqref{eqn:thm:token:help-2} does not hold, which means $\Pr[(\mb{L}^*_1, \mb{L}^*_2) \in \L] > 1/2$.
-
-\vgap
-
-For every parameterized input specified by $(x, y) \in \Pi$, we know
-\myeqn{
-    \Pr[\text{$\A$ succeeds on $(x, y)$} \mid (\mb{L}^*_1, \mb{L}^*_2) \in \L] \ge 1/6.
-    \label{eqn:thm:token:help-4}
-}
-Otherwise, $\Pr[\text{$\A$ succeeds on $(x, y)$}] < 1/6 + \Pr[(\mb{L}^*_1, \mb{L}^*_2) \notin \L] < 2/3$, contradicting the fact that $\A$ succeeds on any input with probability at least $2/3$.
-
-\begin{lemma} \label{lmm:thm:token:help-1}
-    If $\A$ succeeds on the parameterized input specified by $(x, y) \in \Pi$ under the condition $(\mb{L}^*_1, \mb{L}^*_2)  = (\mb{\Lambda}_1, \mb{\Lambda}_2)$, then
-    \myeqn{
-        \sum_{k \in [p]} \min\big\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \big\} \ge p^{1-\eps}.
-        \label{eqn:thm:token:help-3}
-    }
-\end{lemma}
-
-\begin{proof}
-    Let us rewrite the left hand side of \eqref{eqn:thm:token:help-3} as
-    \myeqn{
-        \hspace{-5mm}
-        && \sum_{k \in [p]} \min\big\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \big\} \nn \\
-        \hspace{-5mm}& = &
-        \Big( \sum_{k \in [p] \setminus \set{x,y}} \min\big\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \big\} \Big)
-        \nn \\
-        \hspace{-5mm}&& + \min\big\{\mb{\Lambda}_1[x, y], \ceil{p^{1-\eps}} \big\}
-        + \min\big\{\ceil{p^{1-\eps}}, \mb{\Lambda}_2[x, y] \big\}.
-        \label{eqn:thm:token:help-8}
-    }
-
-    If a machine $k \in [p] \setminus \set{x, y}$ receives a token from machine $x$ in the first round and sends the token to machine $y$ in the second round, we say that the token is {\em relayed} by machine $k$.
-Under the condition $(\mb{L}^*_1, \mb{L}^*_2)  = (\mb{\Lambda}_1, \mb{\Lambda}_2)$, machine $k$ can relay no more than $\min\{\mb{\Lambda}_1[x, k], \mb{\Lambda}_2[k, y] \}$ tokens.
-On the other hand, if a token is not relayed by any machines, it needs to be sent from machine $x$ to machine $y$ directly (in either the first or the second round).
-The number of tokens that can be sent this way is at most
-    \myeqn{
-        \min\big\{\mb{\Lambda}_1[x, y], \ceil{p^{1-\eps}} \big\}
-        + \min\big\{\ceil{p^{1-\eps}}, \mb{\Lambda}_2[x, y] \big\}. \nn
-    }
-
-    Thus, \eqref{eqn:thm:token:help-8} gives an upper bound on the total number of tokens that $\A$ can move to machine $y$ at the end of the second round.
-The upper bound must be at least $\ceil{p^{1-\eps}}$ for $\A$ to succeed on $(x, y)$.
-\end{proof}
-
-By putting together \eqref{eqn:thm:token:help-4} and \eqref{eqn:thm:token:help-3}, we obtain the following for every parameterized input specified by $(x, y) \in \Pi$:
-\myeqn{
-    \expt\Big[\sum_{k \in [p]} \min\big\{\mb{L}^*_1[x, k], \mb{L}^*_2[k, y] \big\} \bigmid (\mb{L}^*_1, \mb{L}^*_2) \in \L \Big]
-    \ge
-    \fr{p^{1-\eps}}{6}
-    \label{eqn:thm:token:help-5}
-}
-
-For any $(\mb{\Lambda}_1, \mb{\Lambda}_2) \in \L$ and any $k \in [p]$, we have: 
-\myeqn{
-    && \Big( \sum_{x \in [p]} \mb{\Lambda}_1[x,k] \Big)
-    \cdot \Big( \sum_{y \in [p]} \mb{\Lambda}_2[k,y] \Big) \nn \\
-    &=&
-    \sum_{x \in [p]} \sum_{y \in [p]}
-     \mb{\Lambda}_1[x,k] \cdot \mb{\Lambda}_2[k,y]  \nn \\
-    &\ge& 
-    \sum_{x \in [p]} \sum_{y \in [p]}
-    \min\big\{\mb{\Lambda}_1[x,k], \mb{\Lambda}_2[k,y]\big\}  \nn \\ 
-    &\ge& 
-    \sum_{(x,y) \in \Pi} 
-    \min\big\{\mb{\Lambda}_1[x,k], \mb{\Lambda}_2[k,y]\big\}  \label{eqn:thm:token:help-6}
-}
-where the second inequality used the fact that $ab \ge \min\{a, b\}$ holds for any non-negative integers $a$ and $b$.
-
-\vgap
-
-We can now derive:
-\myeqn{
-    && \sum_{k \in [p]} \expt \Big[ \Big(\sum_{x \in [p]} \mb{L}^*_1[i, k] \Big) \cdot \Big( \sum_{y \in [p]} \mb{L}^*_2[k, j] \Big)
-    \bigmid
-    (\mb{L}^*_1, \mb{L}^*_2) \in \L
-    \Big] \nn \\
-    &\ge&
-    \expt \Big[\sum_{k \in [p]} \sum_{(x,y) \in \Pi} \min\big\{\mb{L}^*_1[x,k], \mb{L}^*_2[k,y]\big\}
-    \bigmid
-    (\mb{L}^*_1, \mb{L}^*_2) \in \L
-    \Big] \nn \\
-    && \explain{by linearity of expectation and \eqref{eqn:thm:token:help-6}} \nn \\
-    &=&
-    \expt \Big[\sum_{(x,y) \in \Pi} \sum_{k \in [p]} \min\big\{\mb{L}^*_1[x,k], \mb{L}^*_2[k,y]\big\}
-    \bigmid
-    (\mb{L}^*_1, \mb{L}^*_2) \in \L
-    \Big] \nn \\
-    &=&
-    \sum_{(x,y) \in \Pi}
-    \expt \Big[\sum_{k \in [p]} \min\big\{\mb{L}^*_1[x,k], \mb{L}^*_2[k,y]\big\}
-    \bigmid
-    (\mb{L}^*_1, \mb{L}^*_2) \in \L
-    \Big] \nn \\
-    && \explain{linearity of expectation} \nn \\ 
-    &\ge& 
-    \fr{p^{1-\eps} \cdot p (p-1)}{6}. \hspace{5mm} \explain{by \eqref{eqn:thm:token:help-5}} \nn 
-}
-
-Therefore, there is at least one $k^* \in [p]$ such that 
-\myeqn{
-    && \expt \Big[ \Big(\sum_{x \in [p]} \mb{L}^*_1[i, k^*] \Big) \cdot \Big( \sum_{y \in [p]} \mb{L}^*_2[k^*, j] \Big)
-    \bigmid
-    (\mb{L}^*_1, \mb{L}^*_2) \in \L
-    \Big] \nn \\ 
-    &\ge& 
-    \fr{p^{1-\eps}(p-1)}{6} \ge p^{2-\eps}/12
-    \label{eqn:thm:token:help-7}
-}
-where the last inequality used $p \ge 2$.
-
-\vgap
-
-However, by the definition of $\L$ (in particular, the inequality \eqref{eqn:thm:token:help-1}), when $ (\mb{L}^*_1, \mb{L}^*_2) \in \L $, the product $\sum_{x \in [p]} \mb{L}^*_1[i, k]  \cdot \sum_{y \in [p]} \mb{L}^*_2[k, j]$ can be at most $p^{2-\eps}/24$ for all $k \in [p]$.
-This contradicts the existence of $k^*$.
-Therefore, our assumption
-$\Pr[(\mb{L}^*_1, \mb{L}^*_2) \in \L] > 1/2$ must be a false one.
 

### 2024-05-19 msj.tex
yufei

@@
 
+\def\out{\mathrm{OUT}}
+
 \allowdisplaybreaks
@@
 \begin{abstract}
-    Later.
+    In the {\em multiway spatial join} problem, we have a constant number $k \ge 2$ sets of axis-parallel rectangles $R_1, R_2, ..., R_k$ in 2D space and need to report all $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
+\times R_k$ such that the $k$ rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ...
+\cap r_k \neq \emptyset$.
+The problem, which plays a fundamental role in spatial databases, has been extensively studied for over two decades.
+In this paper, we show how to settle the problem --- regardless of the constant $k$ --- in $O(n \log n + \out)$ time, where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the total number of $k$-tuples in the result.
+This significantly improves the state of the art, which is an algorithm with running time $n \log^{\Theta(k)} n + O(\out)$.
+Our algorithm is comparison-based
 \end{abstract}

### 2024-05-19 msj.tex
yufei

--- 2024-05-16 msj.tex
@@
 \begin{abstract}
-    Later.
+    In a {\em multiway spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$.
+The objective is to identify and report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
+\times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ...
+\cap r_k \neq \emptyset$.
+This problem holds significant importance in spatial databases and has been extensively studied for over two decades.
+We explain how to settle the problem in $O(n \log n + \out)$ time using $O(n)$ space --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
+The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
+Our result significantly improves the state of the art, which is an algorithm with running time $n \log^{\Theta(k)} n + O(\out)$.
 \end{abstract}

### 2024-05-19 msj.tex
yufei

@@
 \def\I{\mathcal{I}}
+\def\J{\mathcal{J}}
 \def\II{\mathscr{I}}
@@
 The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
-Our result significantly improves the state of the art, which is an algorithm with running time $n \log^{\Theta(k)} n + O(\out)$.
+Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{\Theta(k)} n + \out)$.
 \end{abstract}
@@
 \section{Introduction} \label{sec:intro}
+
+This paper revisits the {\em multiway spatial join} (MSJ) problem defined as follows.
+A rectangle $r$ in 2D space $\real^2$ is said to be {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.
+Let $k$ be a constant integer at least 2.
+The input of MSJ comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles in $\real^2$.
+The objective is to report all the $k$-tuples $(r_1, r_2, ..., r_k)$ such that
+\myitems{
+    \item $r_i \in R_i$ for each $1 \le i \le k$; and
+    \item $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$, namely, the $k$ rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection.
+}
+We represent the set of such $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$ and refer to it as the {\em join result}.
+Let $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.
 

### 2024-05-19 msj.tex
yufei

@@
 This paper revisits the {\em multiway spatial join} (MSJ) problem defined as follows.
-A rectangle $r$ in 2D space $\real^2$ is said to be {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.
 Let $k$ be a constant integer at least 2.
-The input of MSJ comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles in $\real^2$.
-The objective is to report all the $k$-tuples $(r_1, r_2, ..., r_k)$ such that
+The input of MSJ comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is said to be {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
+The objective is to find all the $k$-tuples $(r_1, r_2, ..., r_k)$ such that
 \myitems{
@@
 }
-We represent the set of such $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$ and refer to it as the {\em join result}.
-Let $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.
+We represent the set of these $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$, which will be referred to as the {\em join result}.
+Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.
 

### 2024-05-19 msj.tex
yufei

@@
 Let $k$ be a constant integer at least 2.
-The input of MSJ comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is said to be {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
+The input of the {\em $k$-MSJ problem} comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is said to be {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
 The objective is to find all the $k$-tuples $(r_1, r_2, ..., r_k)$ such that
@@
 }
-We represent the set of these $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$, which will be referred to as the {\em join result}.
+We represent the set of these $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$ and refer to it as the {\em join result}.
 Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.

### 2024-05-19 msj.tex
yufei

@@
 \def\L{\mathcal{L}}
+\def\P{\mathcal{P}}
 \def\Q{\mathcal{Q}}
@@
 \begin{abstract}
-    In a {\em multiway spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$.
+    In a {\em multiway spatial join}, we are given a constant number $k \geq 3$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$.
 The objective is to identify and report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
@@
 This paper revisits the {\em multiway spatial join} (MSJ) problem defined as follows.
-Let $k$ be a constant integer at least 2.
-The input of the {\em $k$-MSJ problem} comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is said to be {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
+Let $k$ be a constant integer at least 3.
+The input of {\em $k$-MSJ} comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
 The objective is to find all the $k$-tuples $(r_1, r_2, ..., r_k)$ such that
@@
 
+\vgap
+
+MSJ is a fundamental operation in spatial databases, which manage geometric entities such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
+It plays a crucial role in implementing the {\em filter-refinement mechanism} \cite{???}, the dominant approach in spatial databases for computing the overlay information of different entities.
+Specifically, a geometric entity, at the lowest level, is typically modeled as a polygon.
+Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be very time consuming when the two polygons have complex boundaries.
+To alleviate the computation burden, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $\P$, which is the smallest axis-parallel rectangle containing $\P$.
+Two polygons $\P_1$ and $\P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ have a non-empty intersection.
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-19 msj.tex
yufei

--- 2024-05-19 msj.tex
@@
 
-\title{Optimal Multiway Spatial Joins}
+\title{Optimal (Multiway) Spatial Joins}
 
@@
 \begin{abstract}
-    In a {\em multiway spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$.
+    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$.
 The objective is to identify and report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
@@
 
-This paper revisits the {\em multiway spatial join} (MSJ) problem defined as follows.
+This paper revisits the {\em spatial join} (SJ) problem defined as follows.
 Let $k$ be a constant integer at least 2.
-The input of MSJ comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is said to be {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
-The objective is to find all the $k$-tuples $(r_1, r_2, ..., r_k)$ such that
+In the {\em $k$-SJ} problem, the input comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
+The goal is to find all $k$-tuples $(r_1, r_2, ..., r_k)$ where
 \myitems{
-    \item $r_i \in R_i$ for each $1 \le i \le k$; and
+    \item $r_i \in R_i$ for each $i \in [1, k]$; and
     \item $r_1 \cap r_2 \cap ... \cap r_k \neq \emptyset$, namely, the $k$ rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection.
 }
-We represent the set of these $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$, which will be referred to as the {\em join result}.
+We represent the set of these $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$, referred to as the {\em join result}.
 Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.
+
+\vgap
+
+SJ is a fundamental operation in spatial databases, which manage geometric entities such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
+It plays a crucial role in implementing the {\em filter-refinement mechanism} \cite{???}, the dominant approach in spatial databases for computing the overlay information of different entities.
+Specifically, a geometric entity, at the lowest level, is usually modeled as a polygon.
+Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be very expensive when the polygons have complex boundaries.
+To alleviate the computation burden, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $\P$, which is the smallest axis-parallel rectangle enclosing $\P$.
+Two polygons $\P_1$ and $\P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ have a non-empty intersection.
+Motivated by this, filter-refinement first invokes an SJ
 

### 2024-05-19 msj.tex
yufei

@@
 SJ is a fundamental operation in spatial databases, which manage geometric entities such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
-It plays a crucial role in implementing the {\em filter-refinement mechanism} \cite{???}, the dominant approach in spatial databases for computing the overlay information of different entities.
+It plays a crucial role in implementing the {\em filter-refinement mechanism}, the dominant approach in spatial databases for computing the overlays of different entities.
 Specifically, a geometric entity, at the lowest level, is usually modeled as a polygon.
 Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be very expensive when the polygons have complex boundaries.
-To alleviate the computation burden, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $\P$, which is the smallest axis-parallel rectangle enclosing $\P$.
+For better efficiency, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $\P$, which is the smallest axis-parallel rectangle enclosing $\P$.
 Two polygons $\P_1$ and $\P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ have a non-empty intersection.
-Motivated by this, filter-refinement first invokes an SJ
+To overlay $k$ different sets of entities, filter-refinement first performs ``filtering'' by using an SJ to fetch all $k$-tuples $(r_1, r_2, ..., r_k)$ where each $r_i$ ($1 \le i \le k$) is the MBR of a polygon $\P_i$ and a ``refines'' every such $k$-tuple by checking if $\P_1, \P_2, ..., \P_k$ indeed have a non-empty intersection.
 

### 2024-05-19 msj.tex
yufei

@@
 
+\subsection{Previous Results} \label{sec:intro:prev}
+
+SJs have been extensively studied in the database system , which has led to the development of numerous interesting methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance on real-world applications.
+We refer the interested readers to \cite{mp01,???} as entry points into the literature.
+
+\vgap
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-19 msj.tex
yufei

@@
 Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be very expensive when the polygons have complex boundaries.
-For better efficiency, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $\P$, which is the smallest axis-parallel rectangle enclosing $\P$.
-Two polygons $\P_1$ and $\P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ have a non-empty intersection.
-To overlay $k$ different sets of entities, filter-refinement first performs ``filtering'' by using an SJ to fetch all $k$-tuples $(r_1, r_2, ..., r_k)$ where each $r_i$ ($1 \le i \le k$) is the MBR of a polygon $\P_i$ and a ``refines'' every such $k$-tuple by checking if $\P_1, \P_2, ..., \P_k$ indeed have a non-empty intersection.
+For better efficiency, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $P$, which is the smallest axis-parallel rectangle enclosing $P$.
+Two polygons $P_1$ and $P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ have a non-empty intersection.
+To overlay $k$ different sets of entities, filter-refinement first performs ``filtering'' by using an SJ to fetch all $k$-tuples $(r_1, r_2, ..., r_k)$ where each $r_i$ ($1 \le i \le k$) is the MBR of a polygon $P_i$ and a ``refines'' every such $k$-tuple by checking if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection.
 
@@
 SJs have been extensively studied in the database system , which has led to the development of numerous interesting methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance on real-world applications.
-We refer the interested readers to \cite{mp01,???} as entry points into the literature.
+We refer the interested readers to \cite{apr+00,mp01,???} as entry points into the literature.
 
 \vgap
+
+From the theory perspective, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solvable in $O(n \log n + \out)$ time by the standard planesweep technique \cite{bcko08}.
 

### 2024-05-19 msj.tex
yufei

@@
 From the theory perspective, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solvable in $O(n \log n + \out)$ time by the standard planesweep technique \cite{bcko08}.
+However, the problem becomes significantly more challenging as soon as $k$ reaches 3, i.e., the {\em multiway} scenario.
+All the solutions developed (see \cite{mp01,???} and the references therein) before 2022 all suffer from a worst-case time complexity of $O(n^k)$, namely, no better than the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
+\times R_k$.
 

### 2024-05-19 msj.tex
yufei

@@
 
+\extraspacing {\bf Math Conventions.}
+
 \subsection{Previous Results} \label{sec:intro:prev}
@@
 
+\vgap
+
+Year 2022 witnessed two independent works \cite{ty22,???} that reduce the time complexity of the $k$-SJ problem to $O((n + \out) \cdot \polylog n)$.
+Interestingly, neither \cite{ty22} nor \cite{???} studied $k$-SJ explicitly.
+In \cite{ty22}, Tao and Yi studied multiple variants of interval intersection joins under updates.
+The variant most relevant to our context is the following.
+For each $i \in [k]$, let $\I_i$ be a set of (one-dimensional) intervals in $\real$.
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-19 msj.tex
yufei

--- 2024-05-19 msj.tex
@@
 
+\vgap
+
+Year 2022 witnessed two independent works \cite{ty22,???} that, although not addressing the problem $k$-SJ directly, imply algorithms capable of solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
+Specifically, in \cite{ty22}, Tao and Yi studied multiple variants of interval intersection joins under updates.
+Most relevant to our context is the variant where the input contains, for each $i \in [k]$, a set $\I_i$ of (one-dimensional) intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ...
+\times \I_k$ such that $\bigcap_{i=1}^k I_i \neq \emptyset$.
+Set $n = \sum_{i=1}^k |\I_i|$.
+Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list the new (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $\Delta$ is the number of such $k$-tuples (i.e., change in the result size).
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-19 msj.tex
yufei

--- 2024-05-19 msj.tex
@@
 
+\vgap
+
+Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling the problem $k$-SJ directly, imply algorithms for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
+Specifically, in \cite{ty22}, Tao and Yi studied multiple variants of interval intersection joins under updates.
+Most relevant to our context is the variant where the input contains, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ...
+\times \I_k$ such that $\bigcap_{i=1}^k I_i \neq \emptyset$.
+Set $n = \sum_{i=1}^k |\I_i|$.
+Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the new (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $\Delta$ is the number of such $k$-tuples (i.e., change in the result size).
+Tao and Yi \cite{ty22} presented a data structure to achieve the purpose with the space consumption of $O(n \polylog n)$.
+Combining their structure with the planesweep methodology, one can obtain an algorithm for solving the $k$-SJ problem  in $O((n + \out) \cdot \polylog n)$ time.
+
+\vgap
+
+In \cite{kcko22}, Khamis et al.\ investigated
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-19 msj.tex
yufei

@@
 
-From the theory perspective, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solvable in $O(n \log n + \out)$ time by the standard planesweep technique \cite{bcko08}.
+From the theory perspective, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem is solvable in $O(n \log n + \out)$ time by planesweep \cite{bcko08}.
 However, the problem becomes significantly more challenging as soon as $k$ reaches 3, i.e., the {\em multiway} scenario.
@@
 Tao and Yi \cite{ty22} presented a data structure to achieve the purpose with the space consumption of $O(n \polylog n)$.
-Combining their structure with the planesweep methodology, one can obtain an algorithm for solving the $k$-SJ problem  in $O((n + \out) \cdot \polylog n)$ time.
+Combining their structure with the planesweep technique, one can obtain an algorithm for solving the $k$-SJ problem  in $O((n + \out) \cdot \polylog n)$ time.
 

### 2024-05-19 msj.tex
yufei

@@
 
-SJs have been extensively studied in the database system , which has led to the development of numerous interesting methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance on real-world applications.
-We refer the interested readers to \cite{apr+00,mp01,???} as entry points into the literature.
+SJs have been extensively studied in the database system , leading to the development of numerous interesting methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance in real-world applications.
+We refer interested readers to \cite{apr+00,mp01,???} as entry points into the literature.
 
@@
 
-From the theory perspective, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem is solvable in $O(n \log n + \out)$ time by planesweep \cite{bcko08}.
+From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solved in $O(n \log n + \out)$ time by resorting to planesweep \cite{bcko08}.
 However, the problem becomes significantly more challenging as soon as $k$ reaches 3, i.e., the {\em multiway} scenario.
-All the solutions developed (see \cite{mp01,???} and the references therein) before 2022 all suffer from a worst-case time complexity of $O(n^k)$, namely, no better than the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
+All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
 \times R_k$.
@@
 
-Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling the problem $k$-SJ directly, imply algorithms for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
-Specifically, in \cite{ty22}, Tao and Yi studied multiple variants of interval intersection joins under updates.
-Most relevant to our context is the variant where the input contains, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ...
-\times \I_k$ such that $\bigcap_{i=1}^k I_i \neq \emptyset$.
-Set $n = \sum_{i=1}^k |\I_i|$.
-Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the new (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $\Delta$ is the number of such $k$-tuples (i.e., change in the result size).
-Tao and Yi \cite{ty22} presented a data structure to achieve the purpose with the space consumption of $O(n \polylog n)$.
+Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling the $k$-SJ problem directly, imply $k$-SJ algorithms with $O((n + \out) \cdot \polylog n)$ running time.
+Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates.
+Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ...
+\times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$.
+Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples.
+Tao and Yi \cite{ty22} presented a data structure of $O(n \polylog n)$ space achieving the purpose.
 Combining their structure with the planesweep technique, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.

### 2024-05-19 msj.tex
yufei

@@
 
-This paper revisits the {\em spatial join} (SJ) problem defined as follows.
-Let $k$ be a constant integer at least 2.
+This paper revisits the {\em spatial join} (SJ) problem formulated as follows.
+Let $k \ge 2$ be a constant integer.
 In the {\em $k$-SJ} problem, the input comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
@@
 }
-We represent the set of these $k$-tuples as $\J_k(R_1, R_2, ..., R_k)$, referred to as the {\em join result}.
+We represent the set of $k$-tuples described above as $\J_k(R_1, R_2, ..., R_k)$, referred to as the {\em join result}.
 Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.
@@
 
-SJ is a fundamental operation in spatial databases, which manage geometric entities such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
-It plays a crucial role in implementing the {\em filter-refinement mechanism}, the dominant approach in spatial databases for computing the overlays of different entities.
-Specifically, a geometric entity, at the lowest level, is usually modeled as a polygon.
-Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be very expensive when the polygons have complex boundaries.
-For better efficiency, a spatial database stores the {\em minimum bounding rectangle} (MBR) of each polygon $P$, which is the smallest axis-parallel rectangle enclosing $P$.
-Two polygons $P_1$ and $P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ have a non-empty intersection.
-To overlay $k$ different sets of entities, filter-refinement first performs ``filtering'' by using an SJ to fetch all $k$-tuples $(r_1, r_2, ..., r_k)$ where each $r_i$ ($1 \le i \le k$) is the MBR of a polygon $P_i$ and a ``refines'' every such $k$-tuple by checking if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection.
+SJ is a fundamental operation in spatial databases, which manage {\em geometric entities} such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
+It plays a crucial role in implementing the {\em filter-refinement mechanism}, the dominant approach for overlaying different entities in real systems.
+To understand why, first note that a geometric entity is typically modeled as a polygon.
+Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be exceedingly expensive when the polygons have complex boundaries.
+To mitigate the issue, a spatial database stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$.
+Let $G_1, G_2, ..., G_k$ be $k$ different sets of geometric entities, and $R_1, R_2, ..., R_k$ be the corresponding sets of MBRs.
+To compute the overlay information from $G_1, G_2, ..., G_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $G_i$ whose MBR is $r_i$.
+
+%Two polygons $P_1$ and $P_2$ can overlap with each other only if their MBRs $r_1$ and $r_2$ intersect.
 
@@
 
-Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling the $k$-SJ problem directly, imply $k$-SJ algorithms with $O((n + \out) \cdot \polylog n)$ running time.
+Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling $k$-SJ directly, imply $k$-SJ algorithms with $O((n + \out) \cdot \polylog n)$ running time.
 Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates.

### 2024-05-19 msj.tex
yufei

@@
 
-SJ is a fundamental operation in spatial databases, which manage {\em geometric entities} such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
-It plays a crucial role in implementing the {\em filter-refinement mechanism}, the dominant approach for overlaying different entities in real systems.
-To understand why, first note that a geometric entity is typically modeled as a polygon.
+SJ is a fundamental operation in spatial databases (SDB), which manage {\em geometric entities} such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
+It plays a crucial role in implementing the {\em filter-refinement mechanism}, which is the dominant approach for overlaying different entities in an SDB.
+To explain this mechanism, first note that a geometric entity is typically modeled as a polygon.
 Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be exceedingly expensive when the polygons have complex boundaries.
-To mitigate the issue, a spatial database stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$.
-Let $G_1, G_2, ..., G_k$ be $k$ different sets of geometric entities, and $R_1, R_2, ..., R_k$ be the corresponding sets of MBRs.
+To mitigate the issue, a spatial database stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$; this way, each set $G$ of geometric entities spawns a set $R$ of MBRs.
+Consider $k$ sets of geometric entities $G_1, G_2, ..., G_k$, and the corresponding sets of MBRs $R_1, R_2, ..., R_k$.
 To compute the overlay information from $G_1, G_2, ..., G_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $G_i$ whose MBR is $r_i$.

### 2024-05-19 msj.tex
yufei

@@
 Consider $k$ sets of geometric entities $G_1, G_2, ..., G_k$, and the corresponding sets of MBRs $R_1, R_2, ..., R_k$.
-To compute the overlay information from $G_1, G_2, ..., G_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $G_i$ whose MBR is $r_i$.
+To compute overlays from $G_1, G_2, ..., G_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $G_i$ whose MBR is $r_i$.
 
@@
 
-SJs have been extensively studied in the database system , leading to the development of numerous interesting methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance in real-world applications.
+SJs have been extensively studied in the database-system community, leading to the development of numerous methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance in real-world applications.
 We refer interested readers to \cite{apr+00,mp01,???} as entry points into the literature.
@@
 
-From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solved in $O(n \log n + \out)$ time by resorting to planesweep \cite{bcko08}.
-However, the problem becomes significantly more challenging as soon as $k$ reaches 3, i.e., the {\em multiway} scenario.
-All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
+From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solved in $O(n \log n + \out)$ time with planesweep \cite{bcko08}.
+However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
+All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
 \times R_k$.
@@
 
-Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling $k$-SJ directly, imply $k$-SJ algorithms with $O((n + \out) \cdot \polylog n)$ running time.
+Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling $k$-SJ directly, imply provably fast $k$-SJ algorithms.
 Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates.
@@
 \times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$.
-Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples.
-Tao and Yi \cite{ty22} presented a data structure of $O(n \polylog n)$ space achieving the purpose.
-Combining their structure with the planesweep technique, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
+The objective is to design a data structure, which, given the insertion (resp., deletion) of an interval in one of the $k$ sets, can identify all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples.
+Tao and Yi \cite{ty22} presented a structure of $O(n \polylog n)$ space achieving the purpose.
+Combining their structure with planesweep, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
 

### 2024-05-19 msj.tex
yufei

@@
 
-From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solved in $O(n \log n + \out)$ time with planesweep \cite{bcko08}.
+From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time with planesweep \cite{bcko08}.
 However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
@@
 Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates.
-Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ...
+Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1,$ $I_2,$ $..., I_k) \in \I_1 \times \I_2 \times ...
 \times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$.
@@
 
-In \cite{kcko22}, Khamis et al.\ investigated
+In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend equi-joins in two ways.
+First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-joins is replaced with an ``non-empty intersection'' predicate on the attributes involved.
+The $k$-SJ problem can be regarded as a join defined next under the framework of \cite{kcko22}.
+For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$.
+For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle).
+The objective is to output all $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ...
+\times R_k$ such that $\cap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\cap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$.
 

### 2024-05-19 msj.tex
yufei

@@
 The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
-Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{\Theta(k)} n + \out)$.
+Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{2k} n + \out)$.
 \end{abstract}
@@
 First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-joins is replaced with an ``non-empty intersection'' predicate on the attributes involved.
-The $k$-SJ problem can be regarded as a join defined next under the framework of \cite{kcko22}.
+The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}.
 For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$.
 For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle).
-The objective is to output all $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ...
-\times R_k$ such that $\cap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\cap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$.
+The objective is to output all $k$-tuples $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ...
+\times R_k$ satisfying $\bigcap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$.
+Khamis et al.\ \cite{kcko22} developed an algorithm that can process this join in $O(n \log^{2k} n + \out)$ time, where $n = \sum_{i=1}^k |R_i|$.
+Their algorithm requires $O(n \log n)$ space.
 

### 2024-05-19 msj.tex
yufei

@@
 
-\extraspacing {\bf Math Conventions.}
+\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$.
+Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle.
+\todo{more?}
 
@@
 
-From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time with planesweep \cite{bcko08}.
+From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time using $O(n)$ space (see, e.g., \cite{bcko08}).
 However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
@@
 \times R_k$.
+
+\begin{table*} 
+    \begin{tabular}{c|c|c|c|c} 
+        $\bm{k}$ & {\bf method} & {\bf runtime} (big-O) & {\bf space} (big-O) & {\bf remark} \\
+        \hline\hline 
+        2 & folklore & $n \log n + \out$ & $n$ & comparison-based optimal \\ 
+        \hline
+        $\ge 3$ & before 2022 & $n^k$ & - & \\ 
+        $\ge 3$ & \cite{ty22} & $(n + \out) \cdot \polylog n$ & $n \polylog n$ & \\ 
+        $\ge 3$ & \cite{kcko22} & $n \log^{2k} n + \out$ & $n \log n$ & \\ 
+        \hline
+        $\ge 3$ & ours & $n \log n + \out$ & $n$ & comparison-based optimal
+    \end{tabular}
+
+    \vspace{3mm}
+    \caption{Comparison of results on the $k$-SJ problem} 
+    \label{tab:results-com}
+\end{table*}
 

### 2024-05-19 msj.tex
yufei

@@
 Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle.
-\todo{more?}
+Every logarithm has base 2 by default.
 
@@
 SJs have been extensively studied in the database-system community, leading to the development of numerous methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance in real-world applications.
-We refer interested readers to \cite{apr+00,mp01,???} as entry points into the literature.
+We refer interested readers to \cite{apr+00,bks93,gcn+13,js07,ks97,lr94,lr96,mp98,mp01,mp03,pd96,pmt99} as entry points into the literature.
 
@@
 However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
-All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
+All the solutions developed (see \cite{gcn+13,mp98,mp01,pmt99} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
 \times R_k$.

### 2024-05-19 msj.tex
yufei

@@
 
-SJs have been extensively studied in the database-system community, leading to the development of numerous methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance in real-world applications.
+SJs have been extensively studied in the database-system community, leading to the development of numerous methods that, although lacking strong theoretical guarantees, exhibit good empirical performance in real-world applications.
 We refer interested readers to \cite{apr+00,bks93,gcn+13,js07,ks97,lr94,lr96,mp98,mp01,mp03,pd96,pmt99} as entry points into the literature.

### 2024-05-19 msj.tex
yufei

@@
 However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
-All the solutions developed (see \cite{gcn+13,mp98,mp01,pmt99} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ...
+All the solutions developed  before 2022 (see \cite{gcn+13,mp98,mp01,pmt99} and the references therein) suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that enumerates the entire cartesian product $R_1 \times R_2 \times ...
 \times R_k$.
@@
     \begin{tabular}{c|c|c|c|c} 
-        $\bm{k}$ & {\bf method} & {\bf runtime} (big-O) & {\bf space} (big-O) & {\bf remark} \\
+        $\bm{k}$ & {\bf method} & {\bf runtime} & {\bf space} & {\bf remark} \\
         \hline\hline 
-        2 & folklore & $n \log n + \out$ & $n$ & comparison-based optimal \\ 
+        2 & folklore & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal \\ 
         \hline
-        $\ge 3$ & before 2022 & $n^k$ & - & \\ 
-        $\ge 3$ & \cite{ty22} & $(n + \out) \cdot \polylog n$ & $n \polylog n$ & \\ 
-        $\ge 3$ & \cite{kcko22} & $n \log^{2k} n + \out$ & $n \log n$ & \\ 
+        $\ge 3$ & before 2022 & $O(n^k)$ & - & \\ 
+        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & \\ 
+        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & \\ 
         \hline
-        $\ge 3$ & ours & $n \log n + \out$ & $n$ & comparison-based optimal
+        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal
     \end{tabular}
@@
 
-In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend equi-joins in two ways.
-First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-joins is replaced with an ``non-empty intersection'' predicate on the attributes involved.
+In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend the conventional equi-join in two ways.
+First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with an ``non-empty intersection'' predicate on the attributes involved.
 The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}.
@@
 \times R_k$ satisfying $\bigcap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$.
-Khamis et al.\ \cite{kcko22} developed an algorithm that can process this join in $O(n \log^{2k} n + \out)$ time, where $n = \sum_{i=1}^k |R_i|$.
+It is clear that there is one-one correspondence between the result of this join and that of k-SJ.
+Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time.
 Their algorithm requires $O(n \log n)$ space.
+
+\subsection{Our Results} \label{sec:intro:ours}
 

### 2024-05-19 msj.tex
yufei

@@
 
+\vgap
+
+Finally, it is worth mentioning that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm solving the $k$-SJ problem, even for $k = 2$.
+This can be established via a simple reduction (presented in Appendix~\ref{app:lb} for self-containment) from the {\em element distinctness} (ED) problem, where we are given $n$ real values $e_1, e_2, ..., e_n$ and need to decide whether there are distinct $i, j \in [n]$ satisfying $e_i = e_j$.
+The ED problem demands $\Omega(n \log n)$ comparisons to solve \cite{dl79}.
+
 \subsection{Our Results} \label{sec:intro:ours}
+
+In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm
 
@@
 
+\appendix
+
+\section{An $\bm{\Omega(n \log n)}$ Lower Bound} \label{app:lb}
+
 %\end{sloppy}

### 2024-05-19 msj.tex
yufei

@@
 
-Finally, it is worth mentioning that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm solving the $k$-SJ problem, even for $k = 2$.
+It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm solving the $k$-SJ problem, even for $k = 2$.
 This can be established via a simple reduction (presented in Appendix~\ref{app:lb} for self-containment) from the {\em element distinctness} (ED) problem, where we are given $n$ real values $e_1, e_2, ..., e_n$ and need to decide whether there are distinct $i, j \in [n]$ satisfying $e_i = e_j$.
@@
 
-In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm
+In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time and uses  $O(n)$ space, regardless of the value of the constant $k$.
+The time complexity is asymptotically optimal and significantly improves the current state of the art \cite{kcko22}.
 

### 2024-05-19 msj.tex
yufei

@@
 
+\vgap
+
+In terms of techniques, our main discovery is a result on the mathematical structure of the $k$-SJ problem.
+To facilitate the description of our finding, let us ``define'' the $k$-SJ problem at $k = 1$ as follows: the input is a set $R$ of $n$ rectangles in $\real^2$, and the objective is to list all the rectangles in $R$.
+Clearly, this 1-SJ problem can be trivially solved in $O(n)$ time.
+Our main technical contribution can now
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-19 msj.tex
yufei

@@
 
-In terms of techniques, our main discovery is a result on the mathematical structure of the $k$-SJ problem.
+In terms of techniques, our primary contribution is a result on the mathematical structure of the $k$-SJ problem.
 To facilitate the description of our finding, let us ``define'' the $k$-SJ problem at $k = 1$ as follows: the input is a set $R$ of $n$ rectangles in $\real^2$, and the objective is to list all the rectangles in $R$.
 Clearly, this 1-SJ problem can be trivially solved in $O(n)$ time.
-Our main technical contribution can now
+Our main technical contribution can now be stated as:
+
+\boxminipg{0.9\linewidth}{
+    {\bf Main Discovery:} For $k \ge 3$, the $k$-SJ problem can be converted, in $O(n \log n)$ time, to 
+}
 

### 2024-05-19 msj.tex
yufei

--- 2024-05-19 msj.tex
@@
 
+\todo{Rahul's result?}
+
 \subsection{Our Results} \label{sec:intro:ours}
@@
 The time complexity is asymptotically optimal and significantly improves the current state of the art \cite{kcko22}.
+
+\vgap
+
+In terms of techniques, our primary contribution is a result on the mathematical structure of the $k$-SJ problem.
+Specifically, for any $n \ge 1$ and $k \ge 2$, define function $F(n, k)$ to be the lowest value such that any instance of the $k$-SJ problem with input size $n$ and output size $\out$ can be solved in $F(n, k) + O(\out)$ time.
+Specificaly, define $F(n, 1) = n$ for any $n \ge 1$.
+Our main technical contribution can now be stated as:
+
+\begin{theorem} \label{thm:main}
+    d
+\end{theorem}
+
+% To facilitate the description of our finding, let us ``define'' the $k$-SJ problem at $k = 1$ as follows: the input is a set $R$ of $n$ rectangles in $\real^2$, and the objective is to list all the rectangles in $R$.
+Clearly, this 1-SJ problem can be trivially solved in $O(n)$ time.
+Our main technical contribution can now be stated as:
+% 
+% \boxminipg{0.9\linewidth}{
+%     {\bf Main Discovery:} For $k \ge 3$, the $k$-SJ problem can be converted, in $O(n \log n)$ time, to $F_1(k)$ instances of the $(k-1)$-SJ problem and 
+% }
 

### 2024-05-19 msj.tex
yufei

@@
 
-In terms of techniques, our primary contribution is a result on the mathematical structure of the $k$-SJ problem.
-Specifically, for any $n \ge 1$ and $k \ge 2$, define function $F(n, k)$ to be the lowest value such that any instance of the $k$-SJ problem with input size $n$ and output size $\out$ can be solved in $F(n, k) + O(\out)$ time.
-Specificaly, define $F(n, 1) = n$ for any $n \ge 1$.
+In terms of techniques, our primary contribution is a result on the mathematical structure of $k$-SJ.
+Specifically, for any $n \ge 1$ and $k \ge 2$, define function $F(n, k)$ to return the lowest positive integer such that our $k$-SJ algorithm can solve any instance of the $k$-SJ problem with input size $n$ and output size $\out$ in $F(n, k) + O(\out)$ time.
 Our main technical contribution can now be stated as:
@@
 \begin{theorem} \label{thm:main}
-    d
+    For any $n \ge 1$ and $k \ge 3$, it holds that 
+    \myeqn{
+        F(n, k) = O(k^3) \cdot ( F(n, k-1) + n \log n).
+        \label{eqn:main:reccurrence}
+    }
 \end{theorem}

### 2024-05-19 msj.tex
yufei

@@
 
-In terms of techniques, our primary contribution is a result on the mathematical structure of $k$-SJ.
-Specifically, for any $n \ge 1$ and $k \ge 2$, define function $F(n, k)$ to return the lowest positive integer such that our $k$-SJ algorithm can solve any instance of the $k$-SJ problem with input size $n$ and output size $\out$ in $F(n, k) + O(\out)$ time.
-Our main technical contribution can now be stated as:
+In terms of techniques, our primary contribution is a result on the problem's mathematical structure.
+Fix any $k \ge 3$ and an arbitrary algorithm $\A$ for the $(k-1)$-SJ problem.
+For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size $n$ and output size $\out$.
+Our main technical contribution is to establish:
 
@@
     \myeqn{
-        F(n, k) = O(k^3) \cdot ( F(n, k-1) + n \log n).
+        F(n, k, \out) = O(k^3) \cdot \big( F(n, k-1, \out) + n \log n + \out \big).
         \label{eqn:main:reccurrence}
@@
 \end{theorem}
+
+The theorem implies the existence of a recursive nature of $k$-SJ.
 

### 2024-05-19 msj.tex
yufei

@@
 
-In terms of techniques, our primary contribution is a result on the problem's mathematical structure.
-Fix any $k \ge 3$ and an arbitrary algorithm $\A$ for the $(k-1)$-SJ problem.
+In terms of techniques, our primary contribution is the revelation of a new property on the problem's mathematical structure.
+Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem.
 For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size $n$ and output size $\out$.
@@
 
-\begin{theorem} \label{thm:main}
-    For any $n \ge 1$ and $k \ge 3$, it holds that 
+\begin{theorem} \label{thm:main-recur}
+    For any $n \ge 1$, the $k$-SJ problem can be solved in
     \myeqn{
-        F(n, k, \out) = O(k^3) \cdot \big( F(n, k-1, \out) + n \log n + \out \big).
+        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + \out \big)
         \label{eqn:main:reccurrence}
     }
+    time.
+    The space usage is asymptotically the same as that of $\A$.
 \end{theorem}
 
-The theorem implies the existence of a recursive nature of $k$-SJ.
+The theorem, together with the fact that $\A$ can be any solution to $(k-1)$-SJ, implies the existence of a recursive nature of $k$-SJ.
+Indeed, we will see that the $k$-SJ problem can be converted to $O(k^3)$ instances of $(k-1)$-SJ problem, all of which have input size at most $n$ and output size at most $\out$, plus an additional processing cost of $O(n \log n + \out)$.
+For 2-SJ, we can set $\A$ simply to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$ and uses linear space.
+Combining this with \eqref{eqn:main:reccurrence} yields a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ.
+Solving the recurrence yields:
+
+\begin{theorem} \label{thm:main-alg}
+    For $k \ge 3$, we can settle the $k$-SJ problem in $O( (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space.
+\end{theorem}
+
+When $k$ is a constant, the time and space complexities become $O(n \log n + \out)$ and $O(n)$, as promised.
 

### 2024-05-19 msj.tex
yufei

@@
 
-In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time and uses  $O(n)$ space, regardless of the value of the constant $k$.
+In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time and uses  $O(n)$ space, regardless of the constant $k$.
 The time complexity is asymptotically optimal and significantly improves the current state of the art \cite{kcko22}.
@@
 In terms of techniques, our primary contribution is the revelation of a new property on the problem's mathematical structure.
-Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem.
+Fix any $k \ge 3$ and an {\em arbitrary} comparison-based algorithm $\A$ for the $(k-1)$-SJ problem.
 For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size $n$ and output size $\out$.
@@
 \begin{theorem} \label{thm:main-recur}
-    For any $n \ge 1$, the $k$-SJ problem can be solved in
+    For any $n \ge 1$, the $k$-SJ problem can be solved by a comparison-based algorithm in
     \myeqn{
@@
 
-The theorem, together with the fact that $\A$ can be any solution to $(k-1)$-SJ, implies the existence of a recursive nature of $k$-SJ.
-Indeed, we will see that the $k$-SJ problem can be converted to $O(k^3)$ instances of $(k-1)$-SJ problem, all of which have input size at most $n$ and output size at most $\out$, plus an additional processing cost of $O(n \log n + \out)$.
+The theorem implies the existence of a recursive nature of $k$-SJ.
+Indeed, we will see that the $k$-SJ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all of which have input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$.
 For 2-SJ, we can set $\A$ simply to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$ and uses linear space.
@@
 \begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle the $k$-SJ problem in $O( (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space.
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space.
 \end{theorem}
 
-When $k$ is a constant, the time and space complexities become $O(n \log n + \out)$ and $O(n)$, as promised.
-
-% To facilitate the description of our finding, let us ``define'' the $k$-SJ problem at $k = 1$ as follows: the input is a set $R$ of $n$ rectangles in $\real^2$, and the objective is to list all the rectangles in $R$.
-Clearly, this 1-SJ problem can be trivially solved in $O(n)$ time.
-Our main technical contribution can now be stated as:
-% 
-% \boxminipg{0.9\linewidth}{
-%     {\bf Main Discovery:} For $k \ge 3$, the $k$-SJ problem can be converted, in $O(n \log n)$ time, to $F_1(k)$ instances of the $(k-1)$-SJ problem and 
-% }
+When $k = O(1)$, the time and space complexities become $O(n \log n + \out)$ and $O(n)$, as promised.
 

### 2024-05-19 msj.tex
yufei

@@
 
-\todo{Rahul's result?}
+\todo{Rahul's result? No. Do it in later sections}
 
@@
 
+\vgap
+
+Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
+The answer, interestingly, is --- to a great extent --- yes as far as $k \ge 3$ is concerned.
+Specifically, if the 3D version of the 3-SJ problem could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time (which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}).
+This can be shown using an argument in \cite{kcko22}, which is included in Appendix~\ref{app:lb-cond} for self-containment.
+
 \bibliographystyle{plainurl}% the mandatory bibstyle
@@
 
-\section{An $\bm{\Omega(n \log n)}$ Lower Bound} \label{app:lb}
+\section{An $\bm{\Omega(n \log n)}$ Lower Bound for 2-SJ} \label{app:lb}
+
+\section{Hardness of 3-SJ in 3D Space} \label{app:lb-cond}
 

### 2024-05-19 msj.tex
yufei

--- 2024-05-19 msj.tex
@@
 
+\vgap
+
+Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
+The answer, interestingly, is --- subject to the absence of breakthroughs on a classical problem --- ``yes'' as far as $k \ge 3$ is concerned.
+Specifically, if the 3D version of the 3-SJ problem could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
+This can be shown using an argument in \cite{kcko22}, which we simplify and present in full in Appendix~\ref{app:lb-cond}.
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-19 msj.tex
yufei

@@
 Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
-The answer, interestingly, is --- subject to the absence of breakthroughs on a classical problem --- ``yes'' as far as $k \ge 3$ is concerned.
+The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem.
 Specifically, if the 3D version of the 3-SJ problem could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
-This can be shown using an argument in \cite{kcko22}, which we simplify and present in full in Appendix~\ref{app:lb-cond}.
+This can be shown using an argument in \cite{kcko22}, which, however, was written to prove a more generic result.
+We simply the argument for 3-SJ and present it in full in Appendix~\ref{app:lb-cond}.
 

### 2024-05-20 msj.tex
yufei

@@
 Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle.
-Every logarithm has base 2 by default.
+All logarithms have base 2 by default.
 
@@
         $\ge 3$ & before 2022 & $O(n^k)$ & - & \\ 
-        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & \\ 
-        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & \\ 
+        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & $k = O(1)$ \\ 
+        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & $k = O(1)$ \\ 
         \hline
-        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal
+        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & $k = O(1)$, comparison-based optimal
     \end{tabular}
@@
 
-In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend the conventional equi-join in two ways.
-First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with an ``non-empty intersection'' predicate on the attributes involved.
+In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extends the conventional equi-join in two ways.
+First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with a ``non-empty intersection'' predicate on the attributes involved.
 The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}.
@@
 \begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space.
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space, where $c > 1$ is a positive constant.
 \end{theorem}

### 2024-05-20 msj.tex
yufei

@@
     In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$.
-The objective is to identify and report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
+The objective is to report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
 \times R_k$ where the rectangles $r_1, r_2, ..., r_k$ have a non-empty intersection, i.e., $r_1 \cap r_2 \cap ...
@@
 SJ is a fundamental operation in spatial databases (SDB), which manage {\em geometric entities} such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
-It plays a crucial role in implementing the {\em filter-refinement mechanism}, which is the dominant approach for overlaying different entities in an SDB.
+It plays a crucial role in implementing the {\em filter-refinement mechanism}, which is the dominant approach for computing overlay information in an SDB.
 To explain this mechanism, first note that a geometric entity is typically modeled as a polygon.
 Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be exceedingly expensive when the polygons have complex boundaries.
-To mitigate the issue, a spatial database stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$; this way, each set $G$ of geometric entities spawns a set $R$ of MBRs.
+To mitigate the issue, an SDB stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$; this way, each set $G$ of geometric entities spawns a set $R$ of MBRs.
 Consider $k$ sets of geometric entities $G_1, G_2, ..., G_k$, and the corresponding sets of MBRs $R_1, R_2, ..., R_k$.
@@
 
-From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time using $O(n)$ space (see, e.g., \cite{bcko08}).
+From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time using $O(n)$ space (e.g., by planesweep \cite{bcko08}).
 However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
@@
         $\ge 3$ & before 2022 & $O(n^k)$ & - & \\
-        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & $k = O(1)$ \\ 
-        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & $k = O(1)$ \\ 
+        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & \\
+        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & \\
         \hline
-        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & $k = O(1)$, comparison-based optimal
+        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal
     \end{tabular}
@@
     \vspace{3mm}
-    \caption{Comparison of results on the $k$-SJ problem} 
+    \caption{Comparison of results on the $k$-SJ problem for a constant $k$}
     \label{tab:results-com}
@@
 It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm solving the $k$-SJ problem, even for $k = 2$.
-This can be established via a simple reduction (presented in Appendix~\ref{app:lb} for self-containment) from the {\em element distinctness} (ED) problem, where we are given $n$ real values $e_1, e_2, ..., e_n$ and need to decide whether there are distinct $i, j \in [n]$ satisfying $e_i = e_j$.
+This can be established via a reduction (presented in Appendix~\ref{app:lb} for self-containment) from the {\em element distinctness} (ED) problem, where we are given $n$ real values $e_1, e_2, ..., e_n$ and need to decide whether there are distinct $i, j \in [n]$ satisfying $e_i = e_j$.
 The ED problem demands $\Omega(n \log n)$ comparisons to solve \cite{dl79}.

### 2024-05-20 msj.tex
yufei

@@
 The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem.
-Specifically, if the 3D version of the 3-SJ problem could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
-This can be shown using an argument in \cite{kcko22}, which, however, was written to prove a more generic result.
-We simply the argument for 3-SJ and present it in full in Appendix~\ref{app:lb-cond}.
+Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
+This reduction can be inferred from an argument in \cite{kcko22} used to prove a more generic result.
+We simplify the argument for 3D 3-SJ and present the full reduction in Appendix~\ref{app:lb-cond}.
 

### 2024-05-20 msj.tex
yufei

@@
 Fix any $k \ge 3$ and an {\em arbitrary} comparison-based algorithm $\A$ for the $(k-1)$-SJ problem.
-For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size $n$ and output size $\out$.
+For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$.
 Our main technical contribution is to establish:

### 2024-05-20 msj.tex
yufei

@@
 
-Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
+Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result for $k = O(1)$ in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
 The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem.

### 2024-05-20 msj.tex
yufei

@@
 In terms of techniques, our primary contribution is the revelation of a new property on the problem's mathematical structure.
-Fix any $k \ge 3$ and an {\em arbitrary} comparison-based algorithm $\A$ for the $(k-1)$-SJ problem.
+Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem.
 For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$.
@@
 \begin{theorem} \label{thm:main-recur}
-    For any $n \ge 1$, the $k$-SJ problem can be solved by a comparison-based algorithm in
+    For any $n \ge 1$, equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in
     \myeqn{
@@
     }
-    time.
-    The space usage is asymptotically the same as that of $\A$.
+    time using asymptotically the same space as $\A$.
 \end{theorem}
@@
 Indeed, we will see that the $k$-SJ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all of which have input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$.
+Furthermore, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues.
 For 2-SJ, we can set $\A$ simply to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$ and uses linear space.
@@
 Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result for $k = O(1)$ in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
-The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem.
+The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory.
 Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.

### 2024-05-20 msj.tex
yufei

@@
 This problem holds significant importance in spatial databases and has been extensively studied for over two decades.
-We explain how to settle the problem in $O(n \log n + \out)$ time using $O(n)$ space --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
+We explain how to settle the problem in $O(n \log n + \out)$ time using $O(n + \out)$ space --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
 The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
@@
         $\ge 3$ & before 2022 & $O(n^k)$ & - & \\
-        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & \\
-        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & \\
+        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n + \out)$ & \\
+        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n + \out)$ & \\
         \hline
-        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal
+        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n + \out)$ & comparison-based optimal
     \end{tabular}
@@
 Tao and Yi \cite{ty22} presented a structure of $O(n \polylog n)$ space achieving the purpose.
-Combining their structure with planesweep, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
+Combining their structure with planesweep, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time using $O(n \polylog n + \out)$ space.
 
@@
 Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time.
-Their algorithm requires $O(n \log n)$ space.
+Their algorithm requires $O(n \log n + \out)$ space.
 
@@
 
-In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time and uses  $O(n)$ space, regardless of the constant $k$.
-The time complexity is asymptotically optimal and significantly improves the current state of the art \cite{kcko22}.
+In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time and uses  $O(n + \out)$ space, regardless of the constant $k$.
+The time complexity is asymptotically optimal.
 
@@
     }
-    time using asymptotically the same space as $\A$.
+    time using $O(n + \out)$ space, plus the space cost of $\A$.
 \end{theorem}
@@
 \begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space, where $c > 1$ is a positive constant.
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n + \out)$ space, where $c > 1$ is a positive constant.
 \end{theorem}
 
-When $k = O(1)$, the time and space complexities become $O(n \log n + \out)$ and $O(n)$, as promised.
+When $k = O(1)$, the time and space complexities become $O(n \log n$ $+ \out)$ and $O(n + \out)$, as promised.
 

### 2024-05-22 msj.tex
yufei

@@
 
+\section{Geometric Building Bricks} \label{sec:bricks}
+
+This section will introduce four computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm. Let us start with some terminology:
+\myitems{
+    \item For a horizontal segment $h = [x_1, x_2] \times y$ and a rectangle $r$, we call $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left end point of $h$ (i.e., $(x_1, y) \in r$).
+}
+
+\begin{figure*}
+    \begin{tabular}{cccc}
+        \includegraphics[height=30mm]{./artwork/prob-a} &
+        \includegraphics[height=30mm]{./artwork/prob-a} &
+        \includegraphics[height=30mm]{./artwork/prob-a} &
+        \includegraphics[height=30mm]{./artwork/prob-a}
+    \end{tabular}
+
+    \figcapup
+    \caption{Four geometric building brick problems}
+    \label{fig:probs}
+    \figcapdown
+\end{figure*}
+
+\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that some rectangle $r \in R$, which covers the left end point of $h$, contains the point $(x, y)$.
+If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
+In Figure~\ref{fig:probs}a
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-22 msj.tex
yufei

--- 2024-05-20 msj.tex
@@
 
+\section{Geometric Building Bricks} \label{sec:bricks}
+
+This section will define some notions useful for our discussion. Then, we will introduce four computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
+
+\extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$.
+Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
+Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$.
+We say that a horizontal/vertical segment $s$ {\em crosses} $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+
+\begin{figure*}
+    \begin{tabular}{cccc}
+        \includegraphics[height=30mm]{./artwork/prob-a} &
+        \includegraphics[height=30mm]{./artwork/prob-a} &
+        \includegraphics[height=30mm]{./artwork/prob-a} &
+        \includegraphics[height=30mm]{./artwork/prob-a}
+    \end{tabular}
+
+    \figcapup
+    \caption{Four geometric building brick problems}
+    \label{fig:probs}
+    \figcapdown
+\end{figure*}
+
+\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering the left end point of $h$ and the point $(x, y)$.
+If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
+In Figure~\ref{fig:probs}a, suppose $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+
+\vgap
+
+In Appendix~\ref{app:prob-a}, we explain how the problem can be solved in $O(n \log n)$ time where $n = |H| + |R|$.
+
 \bibliographystyle{plainurl}% the mandatory bibstyle
@@
 
+\section{Algorithm for Problem $\bm{\EuScript{A}}$} \label{app:prob-a}
+
 %\end{sloppy}

### 2024-05-22 msj.tex
yufei

@@
 Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$.
-We say that a horizontal/vertical segment $s$ {\em crosses} $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+We say that a horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
 
@@
         \includegraphics[height=30mm]{./artwork/prob-a} &
-        \includegraphics[height=30mm]{./artwork/prob-a} &
+        \includegraphics[height=30mm]{./artwork/prob-b} &
         \includegraphics[height=30mm]{./artwork/prob-a} &
@@
 If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
-In Figure~\ref{fig:probs}a, suppose $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+In Figure~\ref{fig:probs}a, suppose that $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
 
@@
 In Appendix~\ref{app:prob-a}, we explain how the problem can be solved in $O(n \log n)$ time where $n = |H| + |R|$.
+
+\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+The goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$.
+If such an $h$ does not exist, output nothing for $r$.
+In Figure~\ref{fig:probs}b, suppose that  $H = \set{h_1, h_2, h_3}$ and $R = \set{r}$; the output is $\set{(r, h_2)}$.
 

### 2024-05-22 msj.tex
yufei

@@
 \def\vgap{\vspace{0mm}}
-\def\extraspacing{\vspace{1mm} \noindent}
+\def\extraspacing{\vspace{2mm} \noindent}
 \def\figcapup{\vspace{-2mm}}
@@
     \label{tab:results-com}
+    \figcapdown
 \end{table*}
@@
 We say that a horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+We say that a rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
 
@@
         \includegraphics[height=30mm]{./artwork/prob-b} &
-        \includegraphics[height=30mm]{./artwork/prob-a} &
-        \includegraphics[height=30mm]{./artwork/prob-a}
+        \includegraphics[height=30mm]{./artwork/prob-c} &
+        \includegraphics[height=30mm]{./artwork/prob-a} \\
+        (a) Problem $\EuScript{A}$ &
+        (b) Problem $\EuScript{B}$ &
+        (c) Problem $\EuScript{C}$ &
+        (d) Problem $\EuScript{D}$
     \end{tabular}
@@
 If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
-In Figure~\ref{fig:probs}a, suppose that $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
 
@@
 
-In Appendix~\ref{app:prob-a}, we explain how the problem can be solved in $O(n \log n)$ time where $n = |H| + |R|$.
+In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
 
@@
 If such an $h$ does not exist, output nothing for $r$.
-In Figure~\ref{fig:probs}b, suppose that  $H = \set{h_1, h_2, h_3}$ and $R = \set{r}$; the output is $\set{(r, h_2)}$.
+Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r}$; the output is $\set{(r, h_2)}$.
+
+\vgap
+
+In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
+
+\extraspacing {\bf Problem $\bm{\EuScript{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+The goal is to report, for each segment $h \in H$, all the rectangles $r \in R$ containing $h$.
+Furthermore, those rectangles need to be sorted by the x-coordinates of their right boundaries.
+Formally, if $r_1, r_2, ..., r_t$ (for some $t \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_t)$, provided that $x_i > x_{i-1}$ for each $i \in [2, t]$ where $x_i$ (resp., $x_{i-1}$) is the x-coordinate of the right boundary of $r_i$ (resp., $r_{i-1}$).
+Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
+
+\vgap
+
+In Appendix~\ref{app:prob-c}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
 
@@
 
+\section{Algorithm for Problem $\bm{\EuScript{A}}$} \label{app:prob-a}
+
+\section{Algorithm for Problem $\bm{\EuScript{B}}$} \label{app:prob-b}
+
+\section{Algorithm for Problem $\bm{\EuScript{C}}$} \label{app:prob-c}
+
 \section{An $\bm{\Omega(n \log n)}$ Lower Bound for 2-SJ} \label{app:lb}
@@
 
-\section{Algorithm for Problem $\bm{\EuScript{A}}$} \label{app:prob-a}
-
 %\end{sloppy}

### 2024-05-22 msj.tex
yufei

@@
 
-\def\vgap{\vspace{0mm}}
+\def\vgap{\vspace{1mm}}
 \def\extraspacing{\vspace{2mm} \noindent}
@@
         \includegraphics[height=30mm]{./artwork/prob-a} & 
+        \hspace{3mm}
         \includegraphics[height=30mm]{./artwork/prob-b} &
+        \hspace{3mm}
         \includegraphics[height=30mm]{./artwork/prob-c} &
-        \includegraphics[height=30mm]{./artwork/prob-a} \\
+        \hspace{3mm}
+        \includegraphics[height=30mm]{./artwork/prob-d} \\
         (a) Problem $\EuScript{A}$ &
+        \hspace{3mm}
         (b) Problem $\EuScript{B}$ &
+        \hspace{3mm}
         (c) Problem $\EuScript{C}$ &
+        \hspace{3mm}
         (d) Problem $\EuScript{D}$
@@
 
+\extraspacing {\bf Problem $\bm{\EuScript{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+The goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$.
+Furthermore, those segments need to be sorted by their y-coordinates.
+Formally, if $h_1, h_2, ..., h_t$ (for some $t \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_t)$, provided that $y_i > y_{i-1}$ for each $i \in [2, t]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
+Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.
+
+\vgap
+
+In Appendix~\ref{app:prob-d}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.
+
 \bibliographystyle{plainurl}% the mandatory bibstyle
@@
 
+\section{Algorithm for Problem $\bm{\EuScript{D}}$} \label{app:prob-d}
+
 \section{An $\bm{\Omega(n \log n)}$ Lower Bound for 2-SJ} \label{app:lb}

### 2024-05-22 msj.tex
yufei

@@
 
-This section will define some notions useful for our discussion. Then, we will introduce four computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
+This section will first define some notions useful for our discussion and then introduce four computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
 

### 2024-05-22 msj.tex
yufei

@@
 \extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering the left end point of $h$ and the point $(x, y)$.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$.
 If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
@@
 If such an $h$ does not exist, output nothing for $r$.
-Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r}$; the output is $\set{(r, h_2)}$.
+Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
 

### 2024-05-22 msj.tex
yufei

@@
 \def\extraspacing{\vspace{2mm} \noindent}
-\def\figcapup{\vspace{-2mm}}
-\def\figcapdown{\vspace{-4mm}}
+\def\figcapup{\vspace{-0mm}}
+\def\figcapdown{\vspace{-0mm}}
 
@@
 
-\todo{Rahul's result? No. Do it in later sections}
-
 \subsection{Our Results} \label{sec:intro:ours}
@@
 
+\section{H-V Multiway Spatial Joins} \label{sec:hv}
+
+This section deals with a special version of the $k$-SJ problem.
+Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ..., R_k$.
+In the special version --- which we name the {\em H-V $k$-SJ} problem ---  we impose the constraints that (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
+For better clarity, we will represent the input sets as $R_1, R_2, ..., R_{k-2}, H$ ($=R_{k-1}$), and $V$ ($=R_k$).
+Remember that each $R_i$ ($i \in [k-2]$) is a set of rectangles.
+As we will see in the next section, solving the H-V $k$-SJ problem is an imperative step for settling the original $k$-SJ problem.
+
+\vgap
+
+Our objective is to prove that the H-V $k$-SJ problem can be reduced to the $(k-1)$-SJ problem --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
+For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
+Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ when it is executed on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
+We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \in [2, k-2]$, that is, its efficiency on $\kappa$-SJ should not be worse than that on $(\kappa+1)$-SJ.
+
+\vgap
+
+\todo{Remark on Rahul's result}
+
 \bibliographystyle{plainurl}% the mandatory bibstyle

### 2024-05-22 msj.tex
yufei

@@
 Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be exceedingly expensive when the polygons have complex boundaries.
-To mitigate the issue, an SDB stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$; this way, each set $G$ of geometric entities spawns a set $R$ of MBRs.
-Consider $k$ sets of geometric entities $G_1, G_2, ..., G_k$, and the corresponding sets of MBRs $R_1, R_2, ..., R_k$.
-To compute overlays from $G_1, G_2, ..., G_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $G_i$ whose MBR is $r_i$.
+To mitigate the issue, an SDB stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$; this way, each set $\Gamma$ of geometric entities spawns a set $R$ of MBRs.
+Consider $k$ sets of geometric entities $\Gamma_1, \Gamma_2, ..., \Gamma_k$, and the corresponding sets of MBRs $R_1, R_2, ..., R_k$.
+To compute overlays from $\Gamma_1, \Gamma_2, ..., \Gamma_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $\Gamma_i$ whose MBR is $r_i$.
 
@@
     }
-    time using $O(n + \out)$ space, plus the space cost of $\A$.
+    time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
 \end{theorem}
@@
 For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
-Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ when it is executed on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
-We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \in [2, k-2]$, that is, its efficiency on $\kappa$-SJ should not be worse than that on $(\kappa+1)$-SJ.
+Denote by
+\myitems{
+    \item $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$;
+    \item $G_\kappa(n, \out)$ the worst-case space consumed by $\A$ on any instance mentioned above.
+}
+We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ and $G_\kappa(n, \out) \le G_{\kappa + 1}(n, \out)$ for any $\kappa \in [2, k-2]$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
 
 \vgap
+
+The rest of the section serves as a proof for:
+
+\begin{lemma} \label{lmm:hv}
+    The H-V $k$-SJ problem can be solved in 
+    \myeqn{
+        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + \out \big) \nn 
+    }
+    time using $O(n + \out + G_{k-1}(n, \out))$ space.
+\end{lemma}
 

### 2024-05-22 msj.tex
yufei

@@
 \begin{theorem} \label{thm:main-recur}
-    For any $n \ge 1$, equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in
+    For any $n \ge 1$, equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in time
     \myeqn{
-        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + \out \big)
+        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + \out \big).
         \label{eqn:main:reccurrence}
     }
-    time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
+%     time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
 \end{theorem}
@@
 \begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n + \out)$ space, where $c > 1$ is a positive constant.
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + \out))$ time.
+    %and $O(k \cdot n + \out)$ space, where $c > 1$ is a positive constant.
 \end{theorem}
 
-When $k = O(1)$, the time and space complexities become $O(n \log n$ $+ \out)$ and $O(n + \out)$, as promised.
+We will also prove that the space of our algorithm is $O(k \cdot n)$. When $k = O(1)$, the time and space complexities become $O(n \log n$ $+ \out)$ and $O(n + \out)$, as promised.
 

### 2024-05-22 msj.tex
yufei

@@
 Remember that each $R_i$ ($i \in [k-2]$) is a set of rectangles.
-As we will see in the next section, solving the H-V $k$-SJ problem is an imperative step for settling the original $k$-SJ problem.
+As shown in the next section, solving the H-V $k$-SJ problem is an imperative step for settling the original $k$-SJ problem.
 
@@
     }
-    time using $O(n + \out + G_{k-1}(n, \out))$ space.
+    time using $O(n + \out) + G_{k-1}(n, \out)$ space.
 \end{lemma}
+
+\subsection{Categorizing the Result Tuples} \label{sec:hv:result-part}
 

### 2024-05-22 msj.tex
yufei

@@
 
+Let us represent the result of the H-V $k$-SJ problem as $\J_k(R_1, ..., R_{k-2},$ $H, V)$.
+Recall that $\J_k(R_1, ..., R_{k-2},$ $H, V)$ contains all such $k$-tuples $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
+\times R_{k-2} \times H \times V$ satisfying the condition that $h \cap v \bigcap_{i=1}^{k-2} r_i$ is not empty.
+The condition is equivalent to saying that the intersection point of $h$ and $v$ falls in every $r_i$ of $i \in [k-2]$.
+
 \todo{Remark on Rahul's result}

### 2024-05-22 msj.tex
yufei

@@
 
+\vgap
+
+Let us now focus on any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in  \J_k(R_1, ..., R_{k-2},$ $H, V)$. We classify the tuple into one of the two categories below:
+
+\myitems{
+    \item {{\bf Cat.\ 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$, and $v$  crosses all of $r_1, ..., r_{k-2}$;
+
+    \item {{\bf Cat.\ 2:}} $h$ or $v$ fails to cross some rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. 
+}
+
 \todo{Remark on Rahul's result}

### 2024-05-22 msj.tex
yufei

--- 2024-05-22 msj.tex
@@
 
+\vgap
+
+Let us now focus on any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in  \J_k(R_1, ..., R_{k-2},$ $H, V)$. We classify the tuple into one of the two categories below:
+
+\myitems{
+    \item {{\bf Type 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$, and $v$  crosses all of $r_1, ..., r_{k-2}$;
+
+    \item {{\bf Type 2:}} $h$ or $v$ fails to cross some rectangle in $\set{r_1, r_2, ..., r_{k-2}}$.
+This means the existence of at least one rectangle $r_i$, for some $i \in [k-2]$, covers either an endpoint of $h$, or an endpoint of $v$, or both.
+}
+Figure~\ref{fig:hv:cats} illustrates a result tuple of each type, assuming $k = 4$.
+
+\begin{figure} 
+    \begin{tabular}{cc} 
+        \includegraphics[height=27mm]{./artwork/cat1} & 
+        \hspace{5mm} \includegraphics[height=27mm]{./artwork/cat2} \\
+        (a) Cat.\ 1 result tuple & 
+        (b) Cat.\ 2 result tuple
+    \end{tabular}
+
+    \figcapup 
+    \caption{Two categories of H-V $k$-SJ result tuples ($k$ = 4)} 
+    \label{fig:hv:types}
+    \figcapdown
+\end{figure}
+
 \todo{Remark on Rahul's result}

### 2024-05-22 msj.tex
yufei

@@
     \item {{\bf Type 2:}} $h$ or $v$ fails to cross some rectangle in $\set{r_1, r_2, ..., r_{k-2}}$.
-This means the existence of at least one rectangle $r_i$, for some $i \in [k-2]$, covers either an endpoint of $h$, or an endpoint of $v$, or both.
+This means the existence of at least one rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of either $h$ or $v$ or both.
 }
-Figure~\ref{fig:hv:cats} illustrates a result tuple of each type, assuming $k = 4$.
+Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.
 
@@
     \begin{tabular}{cc} 
-        \includegraphics[height=27mm]{./artwork/cat1} & 
-        \hspace{5mm} \includegraphics[height=27mm]{./artwork/cat2} \\
+        \includegraphics[height=27mm]{./artwork/type1} & 
+        \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
         (a) Cat.\ 1 result tuple & 
@@
 \end{figure}
+
+\vgap
+
+We will find result tuples of each type separately.
+Before proceeding, however, let us observe a sense of symmetry for Type 2.
+As mentioned, for a result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
+As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, etc.
+It is possible for the result tuple to belong to multiple sub-types simultaneously.
+
+\vgap
+
+Our subsequent discussion will proceed as follows. In Section~\ref{sec:hv:type1}, we
+
+\subsection{Finding Result Tuples of Type 1} \label{sec:hv:type1}
+
+\subsection{Finding Result Tuples of Type 2} \label{sec:hv:type2}
 

### 2024-05-22 msj.tex
yufei

@@
 
-Our subsequent discussion will proceed as follows. In Section~\ref{sec:hv:type1}, we
+Our subsequent discussion will proceed as follows.
+Section~\ref{sec:hv:type1} will explain how to produce all the result tuples of Type 1.
+Section~\ref{sec:hv:type2} will explain how to produce the result tuples $(r_1, ..., r_{k-2}, h, v)$ of a particular sub-type of Type 2: those tuples where $r_1$ covers the left endpoint of $h$.
+Other sub-types can be computed using the same algorithm, by symmetry.
+
+\vgap
+
+A remark is in order about duplicate removal.
+A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types.
+Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm.
+However, this does {\em not} mean that the tuple need to be {\em reported} multiple times.
+Whenever a Type-2 result tuple is found, we can immediately decide all the sub-types it belongs to.
+To avoid outputting the tuple more than once, it suffices to
 

### 2024-05-22 msj.tex
yufei

@@
 Whenever a Type-2 result tuple is found, we can immediately decide all the sub-types it belongs to.
-To avoid outputting the tuple more than once, it suffices to
+To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting.
+One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.
 

### 2024-05-22 msj.tex
yufei

@@
 Furthermore, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues.
-For 2-SJ, we can set $\A$ simply to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$ and uses linear space.
+For 2-SJ, we can set $\A$ simply to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.
 Combining this with \eqref{eqn:main:reccurrence} yields a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ.
@@
 
-Let us now focus on any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in  \J_k(R_1, ..., R_{k-2},$ $H, V)$. We classify the tuple into one of the two categories below:
+Let us focus on any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in  \J_k(R_1, ..., R_{k-2},$ $H, V)$. We classify the tuple into one of the two categories below:
 
@@
     \item {{\bf Type 2:}} $h$ or $v$ fails to cross some rectangle in $\set{r_1, r_2, ..., r_{k-2}}$.
-This means the existence of at least one rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of either $h$ or $v$ or both.
+This means the existence of at least one rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
 }
@@
 As mentioned, for a result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
-As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, etc.
+As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc.
 It is possible for the result tuple to belong to multiple sub-types simultaneously.

### 2024-05-22 msj.tex
yufei

@@
 Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm.
-However, this does {\em not} mean that the tuple need to be {\em reported} multiple times.
-Whenever a Type-2 result tuple is found, we can immediately decide all the sub-types it belongs to.
+However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times.
+Whenever a Type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to.
 To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting.

### 2024-05-22 msj.tex
yufei

@@
     \myeqn{
-        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + \out \big).
-        \label{eqn:main:reccurrence}
+        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big).
+        \nn
+        %\label{eqn:main:reccurrence}
     }
@@
 \begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + \out))$ time.
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + k \cdot \out))$ time.
     %and $O(k \cdot n + \out)$ space, where $c > 1$ is a positive constant.
@@
 
-We will also prove that the space of our algorithm is $O(k \cdot n)$. When $k = O(1)$, the time and space complexities become $O(n \log n$ $+ \out)$ and $O(n + \out)$, as promised.
+We will also prove that the space of our algorithm is $O(k \cdot n + k \cdot \out)$. When $k = O(1)$, the time and space complexities become $O(n \log n$ $+ \out)$ and $O(n + \out)$, as promised.
 
@@
     \myeqn{
-        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + \out \big) \nn 
+        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn 
     }

### 2024-05-23 msj.tex
yufei

@@
 }
-We represent the set of $k$-tuples described above as $\J_k(R_1, R_2, ..., R_k)$, referred to as the {\em join result}.
-Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J_k(R_1, R_2, ..., R_k)|$, i.e., the output size.
+We represent the set of $k$-tuples described above as $\J(R_1, R_2, ..., R_k)$, referred to as the {\em join result}.
+Set $n = \sum_{i=1}^k |R_i|$, i.e., the input size, and $\out = |\J(R_1, R_2, ..., R_k)|$, i.e., the output size.
 
@@
 Determining whether two entities overlap amounts to deciding if two polygons intersect, which can be exceedingly expensive when the polygons have complex boundaries.
-To mitigate the issue, an SDB stores, for each polygon $g$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $g$; this way, each set $\Gamma$ of geometric entities spawns a set $R$ of MBRs.
+To mitigate the issue, an SDB stores, for each polygon $\gamma$, its {\em minimum bounding rectangle} (MBR) defined as the smallest axis-parallel rectangle enclosing $\gamma$; this way, each set $\Gamma$ of geometric entities spawns a set $R$ of MBRs.
 Consider $k$ sets of geometric entities $\Gamma_1, \Gamma_2, ..., \Gamma_k$, and the corresponding sets of MBRs $R_1, R_2, ..., R_k$.
-To compute overlays from $\Gamma_1, \Gamma_2, ..., \Gamma_k$, filter-refinement first executes (i) ``a filter step'', which performs an SJ to obtain $\J_k(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J_k(R_1, R_2, ..., R_k)$, examines if $P_1, P_2, ..., P_k$ indeed have a non-empty intersection, where $P_i$ ($i \in [1, k]$) is the entity in $\Gamma_i$ whose MBR is $r_i$.
+To compute overlays from $\Gamma_1, \Gamma_2, ..., \Gamma_k$, filter-refinement first executes (i) a ``filter step'', which performs an SJ to obtain $\J(R_1, R_2, ..., R_k)$, and (ii) a ``refinement step'', which, for each $(r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$, examines if $\gamma_1, \gamma_2, ..., \gamma_k$ indeed have a non-empty intersection, where $\gamma_i$ ($i \in [1, k]$) is the entity in $\Gamma_i$ whose MBR is $r_i$.
 
@@
 
-Our objective is to prove that the H-V $k$-SJ problem can be reduced to the $(k-1)$-SJ problem --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
-For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
+Our objective is to prove that the H-V $k$-SJ problem can be efficiently reduced to the $(k-1)$-SJ problem --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
+For this purpose, we will assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
 Denote by
@@
 
-Let us represent the result of the H-V $k$-SJ problem as $\J_k(R_1, ..., R_{k-2},$ $H, V)$.
-Recall that $\J_k(R_1, ..., R_{k-2},$ $H, V)$ contains all such $k$-tuples $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
+Let us represent the result of the H-V $k$-SJ problem as $\J(R_1, ..., R_{k-2},$ $H, V)$.
+Recall that $\J(R_1, ..., R_{k-2},$ $H, V)$ contains all such $k$-tuples $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
 \times R_{k-2} \times H \times V$ satisfying the condition that $h \cap v \bigcap_{i=1}^{k-2} r_i$ is not empty.
@@
 
-Let us focus on any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in  \J_k(R_1, ..., R_{k-2},$ $H, V)$. We classify the tuple into one of the two categories below:
+Let us focus on any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in  \J(R_1, ..., R_{k-2},$ $H, V)$. We classify the tuple into one of the two categories below:
 
@@
         \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
-        (a) Cat.\ 1 result tuple & 
-        (b) Cat.\ 2 result tuple
+        (a) Type 1 result tuple &
+        (b) Type 2 result tuple
     \end{tabular}
@@
 
+\subsection{Remarks} \label{sec:hv:remarks}
+
 \todo{Remark on Rahul's result}

### 2024-05-23 msj.tex
yufei

@@
 \begin{figure*}
-    \begin{tabular}{cccc}
+    \begin{tabular}{ccc}
         \includegraphics[height=30mm]{./artwork/prob-a} & 
@@
         \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-c} &
-        \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-d} \\
-        (a) Problem $\EuScript{A}$ &
+        \includegraphics[height=30mm]{./artwork/prob-c} \\
+%         &
+%         \hspace{3mm}
+%         \includegraphics[height=30mm]{./artwork/prob-d} \\
+        (a) Problem $\EuScript{A}$ and $\EuScript{D}$ &
         \hspace{3mm}
@@
         \hspace{3mm}
-        (c) Problem $\EuScript{C}$ &
-        \hspace{3mm}
-        (d) Problem $\EuScript{D}$
+        (c) Problem $\EuScript{C}$
+%         &
+%         \hspace{3mm}
+%         (d) Problem $\EuScript{D}$
     \end{tabular}
@@
 \extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$.
-If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
-Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+The goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$.
+If such an $h$ does not exist, output nothing for $r$.
+Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
 
@@
 \extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$.
-If such an $h$ does not exist, output nothing for $r$.
-Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$.
+If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
+Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
 
@@
 Formally, if $h_1, h_2, ..., h_t$ (for some $t \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_t)$, provided that $y_i > y_{i-1}$ for each $i \in [2, t]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
-Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.
+See Figure~\ref{fig:probs}a again where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.
 
@@
 \end{lemma}
+
+Recall that $n$ is the total size of $R_1, ..., R_{k-2}, H$, and $V$. Henceforth, we assume that a distinct ID from 1 to $n$ is given to each rectangle or segment in $R_1, ..., R_{k-2}, H$, and $V$.
 
@@
 However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times.
-Whenever a Type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to.
+Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to.
 To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting.
@@
 \subsection{Finding Result Tuples of Type 1} \label{sec:hv:type1}
+
+This subsection will explain how to find the $k$-tuples in $\J(R_1, ..., R_{k-2},$ $H, V)$ that belong to Type 1.
+We will represent the set of such tuples as $\J_1(R_1, ..., R_{k-2}, H, V)$ and define its ``projection'' onto $(R_1, ..., R_{k-2})$ as
+\myeqn{
+    && \Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V)) \nn \\
+    &=& \{(r_1, ..., r_{k-2}) \mid (r_1, ..., r_{k-2}, h, v) \in \J_1(R_1, ..., R_{k-2}, H, V)
+    \}. \nn
+}
+We will compute $\J_1(R_1, ..., R_{k-2}, H, V)$ in two steps:
+\myitems{
+    \item {{\bf Step 1:}} Compute $\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))$;
+    \item {{\bf Step 2:}} Generate $\J_1(R_1, ..., R_{k-2}, H, V)$ from the result of Step 1.
+}
+
+\noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ constructed as follows. For each rectangle $r_1 \in R_1$, we compute four segments:
+\myitems{
+    \item $h_1$: the lowest segment (i.e., with the smallest y-coordinate) among all the segments in $H$ that cross $r$;
+    \item $h_2$: the highest segment (i.e., with the largest y-coordinate) among all the segments in $H$ that cross $r$;
+    \item $v_1$: the leftmost segment (i.e., with the smallest x-coordinate) among all the segments in $V$ that cross $r$;
+    \item $v_2$: the rightmost segment (i.e., with the largest x-coordinate) among all the segments in $V$ that cross $r$.
+}
+Define $r_1'$ the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:four-segs}, where $r_1'$ is the shaded area.
+Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
+It is possible for $h_1$ and $h_2$ to be the same segment (same for $v_1$ and $v_2$).
+
+\begin{figure}
+    \includegraphics[height=25mm]{./artwork/four-segs}
+    \figcapup
+    \caption{Constructing $r_1'$ from $r_1$}
+    \label{fig:four-segs}
+    \figcapdown
+\end{figure}
+
+a rectangle $r_1'$ to an initially empty set $R_1'$ if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Specifically,
 
@@
 \appendix
+
+\section*{Appendix}
 

### 2024-05-23 msj.tex
yufei

@@
 
-In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
+In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time and $O(n)$ space where $n = |H| + |R|$.
 
@@
 
-In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
+In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time and $O(n)$ space where $n = |H| + |R|$.
 
@@
 
-In Appendix~\ref{app:prob-c}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
+In Appendix~\ref{app:prob-c}, we explain how to solve the problem in $O(n \log n + \out)$ time and $O(n + \out)$ space where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
 
@@
 
-In Appendix~\ref{app:prob-d}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.
+In Appendix~\ref{app:prob-d}, we explain how to solve the problem in $O(n \log n + \out)$ time and $O(n + \out)$ space where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.
 
@@
 
+\begin{figure}
+    \includegraphics[height=25mm]{./artwork/four-segs}
+    \figcapup
+    \caption{Constructing $r_1'$ from $r_1$}
+    \label{fig:four-segs}
+    \figcapdown
+\end{figure}
+
 \noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ constructed as follows. For each rectangle $r_1 \in R_1$, we compute four segments:
@@
 Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
-It is possible for $h_1$ and $h_2$ to be the same segment (same for $v_1$ and $v_2$).
-
-\begin{figure}
-    \includegraphics[height=25mm]{./artwork/four-segs}
-    \figcapup
-    \caption{Constructing $r_1'$ from $r_1$}
-    \label{fig:four-segs}
-    \figcapdown
-\end{figure}
-
-a rectangle $r_1'$ to an initially empty set $R_1'$ if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Specifically,
+Moreover, it is possible for $h_1$ and $h_2$ to be the same segment (same for $v_1$ and $v_2$).
+Define
+\myeqn{
+    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}}. \nn
+}
+Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\EuScript{A}$.
+By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
+It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time and $O(n)$ space.
 

### 2024-05-23 msj.tex
yufei

@@
 }
-Define $r_1'$ the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:four-segs}, where $r_1'$ is the shaded area.
+Define $r_1'$ as the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:four-segs}, where $r_1'$ is the shaded area.
 Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
-Moreover, it is possible for $h_1$ and $h_2$ to be the same segment (same for $v_1$ and $v_2$).
+Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$.
 Define
@@
 It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time and $O(n)$ space.
+
+\vgap
+
+We now solve the $(k-2)$-SJ problem on $R_1', R_2, ..., R_{k-2}$ using the algorithm $\A$ supplied.
+Let us represent the result of this $(k-2)$-SJ as $\J(R_1', R_2, ..., R_{k-2})$.
+The following observation is crucial.
+
+\begin{lemma} \label{sec:hv:type1:proj}
+    A $(k-2)$-tuple $(r_1', r_2, ..., r_k)$ exists in $\J(R_1', R_2, ..., R_{k-2})$ if and only if the $(k-2)$-tuple $(r_1, r_2, ..., r_k)$ exists in $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$, where $r_1$ is the rectangle in $R_1$ defining $r_1'$.
+\end{lemma}
 

### 2024-05-23 msj.tex
yufei

@@
 %         \includegraphics[height=30mm]{./artwork/prob-d} \\
-        (a) Problem $\EuScript{A}$ and $\EuScript{D}$ &
+        (a) Problems $\EuScript{A}$ and $\EuScript{D}$ &
         \hspace{3mm}
@@
 
-\begin{lemma} \label{sec:hv:type1:proj}
+\begin{lemma} \label{lmm:hv:type1:proj}
     A $(k-2)$-tuple $(r_1', r_2, ..., r_k)$ exists in $\J(R_1', R_2, ..., R_{k-2})$ if and only if the $(k-2)$-tuple $(r_1, r_2, ..., r_k)$ exists in $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$, where $r_1$ is the rectangle in $R_1$ defining $r_1'$.
 \end{lemma}
+
+The proof is presented in Appendix~\ref{app:lmm:hv:type1:proj}.
+Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2}, H, V)$, which in turn is bounded by $\out$.
+Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time using $G_{k-2}(n, \out)$ space.
 
@@
 
+\section{Proof of Lemma~\ref{lmm:hv:type1:proj}} \label{app:lmm:hv:type1:proj}
+
 \section{An $\bm{\Omega(n \log n)}$ Lower Bound for 2-SJ} \label{app:lb}

### 2024-05-23 msj.tex
yufei

@@
 
-This subsection will explain how to find the $k$-tuples in $\J(R_1, ..., R_{k-2},$ $H, V)$ that belong to Type 1.
-We will represent the set of such tuples as $\J_1(R_1, ..., R_{k-2}, H, V)$ and define its ``projection'' onto $(R_1, ..., R_{k-2})$ as
+This subsection will explain how to find the $k$-tuples of Type 1.
+Let us represent the set of such tuples as $\J_1(R_1, ..., R_{k-2}, H, V)$ and define its ``projection'' onto $(R_1, ..., R_{k-2})$ as
 \myeqn{
@@
 
-\noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ constructed as follows. For each rectangle $r_1 \in R_1$, we compute four segments:
+\noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ. To construct the instance, for each rectangle $r_1 \in R_1$, we compute four segments:
 \myitems{
@@
 Let us represent the result of this $(k-2)$-SJ as $\J(R_1', R_2, ..., R_{k-2})$.
-The following observation is crucial.
+The lemma below shows that the result is essentially what we aim to compute in Step 1.
 
 \begin{lemma} \label{lmm:hv:type1:proj}
-    A $(k-2)$-tuple $(r_1', r_2, ..., r_k)$ exists in $\J(R_1', R_2, ..., R_{k-2})$ if and only if the $(k-2)$-tuple $(r_1, r_2, ..., r_k)$ exists in $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$, where $r_1$ is the rectangle in $R_1$ defining $r_1'$.
+    A $(k-2)$-tuple $(r_1', r_2, ..., r_{k-2})$ exists in $\J(R_1', R_2, ...,$ $R_{k-2})$ if and only if the $(k-2)$-tuple $(r_1, r_2, ..., r_{k-2})$ is in $\Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$, where $r_1$ is the rectangle in $R_1$ defining $r_1'$.
 \end{lemma}
@@
 The proof is presented in Appendix~\ref{app:lmm:hv:type1:proj}.
-Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2}, H, V)$, which in turn is bounded by $\out$.
+Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2},$ $H, V)$, which in turn is bounded by $\out$.
 Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time using $G_{k-2}(n, \out)$ space.
+
+\extraspacing {\bf Step 2.}
 
@@
 
-\subsection{Remarks} \label{sec:hv:remarks}
+\subsection{Putting Things Together} \label{sec:hv:analysis}
 

### 2024-05-23 msj.tex
yufei

@@
 
+\def\cross{\mathrm{cross}}
 \def\out{\mathrm{OUT}}
@@
 
-\begin{figure}
-    \includegraphics[height=25mm]{./artwork/four-segs}
-    \figcapup
-    \caption{Constructing $r_1'$ from $r_1$}
-    \label{fig:four-segs}
-    \figcapdown
-\end{figure}
-
 \noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ. To construct the instance, for each rectangle $r_1 \in R_1$, we compute four segments:
@@
 }
-Define $r_1'$ as the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:four-segs}, where $r_1'$ is the shaded area.
+Define $r_1'$ as the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:type1}(a), where $r_1'$ is the shaded area.
 Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
@@
 
-\extraspacing {\bf Step 2.}
+\extraspacing {\bf Step 2.} We are ready to explain how to produce $\J_1(R_1, ..., R_{k-2}, H, V)$ from $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$.
+Our algorithm for this purpose will find a set $S$ satisfying
+\myeqn{
+    \J_1(R_1, ..., R_{k-2}, H, V) 
+    \subseteq 
+    S
+    \subseteq 
+    \J(R_1, ..., R_{k-2}, H, V). \nn
+}
+Recall that $\J(R_1, ..., R_{k-2}, H, V)$ is the result of the original H-V $k$-SJ problem, which has a size of $\out$.
+Only the $k$-tuples in $S$ that are of Type 1 will be reported.
+We will achieve the purpose in time $O(n \log n + |S|)$ and space $O(n + |S|)$.
+
+\vgap
+
+Let us now fix an arbitrary $(k-2)$-tuple in $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$. Denote by 
+\myeqn{
+    B = \cap_{i=1}^{k-2} r_i. \nn
+}
+Note that $B$ is a rectangle itself. We say that $B$ is 
+\myitems{
+    \item {\em left-bounded} by $r_i$, if the left edge of $B$ is a part of the left boundary of $r_i$ for some $i \in [k-2]$;
+    \item {\em bottom-bounded} by $r_j$, if the bottom edge of $B$ is a part of the bottom boundary of $r_j$ for some $j \in [k-2]$.
+}
+Figure~\ref{fig:type1}(b) shows an example with rectangles $r_1, r_2$, and $r_3$ (here $k = 5$).
+Their intersection is the gray rectangle $B$, which is left-bounded by $r_3$ and bottom-bounded by $r_2$.
+
+\vgap
+
+We will compute:
+\myitems{
+    \item $\cross(B, H)$, the set of segments in $H$ that cross $B$; 
+    \item $\cross(B, V)$, the set of segments in $V$ that cross $B$.
+}
+
+\begin{figure} 
+    \begin{tabular}{cc}
+            \includegraphics[height=25mm]{./artwork/four-segs} &
+            \includegraphics[height=28mm]{./artwork/oracle} \\ 
+            (a) Constructing $r_1'$ from $r_1$ & 
+            (b) Crossing oracles
+    \end{tabular}
+    \figcapup
+    \caption{Illustrations for Section~\ref{sec:hv:type1}}
+    \label{fig:type1}
+    \figcapdown
+\end{figure}
+
+\vgap
+
+We say that a $k$-tuple in $\J(R_1, ..., R_{k-2}, H, V)$ is {\em supported} by $(r_1, ..., r_{k-2})$ if it has the form $(r_1, ..., r_{k-2}, h, v)$, where $h \in H$ and $v \in V$.
+This means that the intersection of
+
+Let us discuss how to generate all the $k$-tuples in
 

### 2024-05-23 msj.tex
yufei

--- 2024-05-23 msj.tex
@@
 
-\begin{figure}
-    \includegraphics[height=25mm]{./artwork/four-segs}
-    \figcapup
-    \caption{Constructing $r_1'$ from $r_1$}
-    \label{fig:four-segs}
-    \figcapdown
-\end{figure}
-
 \noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ. To construct the instance, for each rectangle $r_1 \in R_1$, we compute four segments:
 \myitems{
-    \item $h_1$: the lowest segment (i.e., with the smallest y-coordinate) among all the segments in $H$ that cross $r$;
-    \item $h_2$: the highest segment (i.e., with the largest y-coordinate) among all the segments in $H$ that cross $r$;
-    \item $v_1$: the leftmost segment (i.e., with the smallest x-coordinate) among all the segments in $V$ that cross $r$;
-    \item $v_2$: the rightmost segment (i.e., with the largest x-coordinate) among all the segments in $V$ that cross $r$.
+    \item $h_1$ (resp., $h_2$): the lowest (resp., highest) segment --- i.e., with the smallest (resp., largest) y-coordinate --- among all the segments in $H$ that cross $r$;
+
+    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment --- i.e., with the smallest (resp., largest) x-coordinate --- among all the segments in $V$ that cross $r$.
 }
-Define $r_1'$ as the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:four-segs}, where $r_1'$ is the shaded area.
+Define $r_1'$ as the rectangle defined by the above four segments in the way shown in Figure~\ref{fig:type1}(a) ($r_1'$ is the shaded area).
 Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
@@
 
-\extraspacing {\bf Step 2.}
+\extraspacing {\bf Step 2.} We are ready to explain how to produce $\J_1(R_1, ..., R_{k-2}, H, V)$ from $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$.
+Our algorithm for this purpose will find a set $S$ satisfying
+\myeqn{
+    \J_1(R_1, ..., R_{k-2}, H, V) 
+    \subseteq 
+    S
+    \subseteq 
+    \J(R_1, ..., R_{k-2}, H, V). \nn
+}
+Recall that $\J(R_1, ..., R_{k-2}, H, V)$ is the result of the original H-V $k$-SJ problem, which has a size of $\out$.
+Only the $k$-tuples in $S$ that are of Type 1 will be reported.
+We will achieve the purpose in time $O(n \log n + |S|)$ and space $O(n + |S|)$.
+
+\vgap
+
+Now, take an arbitrary $(k-2)$-tuple in $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$. Denote by 
+\myeqn{
+    B = \cap_{i=1}^{k-2} r_i. \nn
+}
+Note that $B$ is a rectangle itself. We say that $B$ is 
+\myitems{
+    \item {\em bottom-bounded} by $r_a$, if the bottom edge of $B$ is a part of the bottom boundary of $r_a$ for some $a \in [k-2]$;
+
+    \item {\em left-bounded} by $r_b$, if the left edge of $B$ is a part of the left boundary of $r_b$ for some $b \in [k-2]$.
+}
+Figure~\ref{fig:type1}(b) shows an example with rectangles $r_1, r_2$, and $r_3$ (here $k = 5$).
+Their intersection is the gray rectangle $B$, which is bottom-bounded by $r_2$ and left-bounded by $r_3$.
+
+\vgap
+
+We aim to compute 
+\myitems{
+    \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$; 
+    \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
+}
+In Figure~\ref{fig:type1}(b), $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2, v_3}$.
+As mentioned before, $r_a = r_2$ and $r_b = r_3$.
+We have $\cross(r_a, B; H) = \set{h_1, h_2}$ and $\cross(r_b, B; V) = \set{v_1, v_3}$.
+Note that $h_3$ is not in $\cross(r_a, B; H)$ because $h_3$ does not cross $B$, and $v_2$ is not in $\cross(r_b, B; V)$ because $v_2$ does not cross $r_3$.
+
+\vgap
+
+Equipped with the above, we can generate the target $S$ as follows:
+
+\mytab{
+    \> {\bf algorithm generate-$S$} \\
+    \> 1.\> $S = \emptyset$ \\ 
+    \> 2.\> {\bf for} each $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$ {\bf do} \\
+    \> 3.\>\> define $B, r_a$, and $r_b$ as above \\
+    \> 4.\>\> {\bf for} each $(h, v) \in \cross(r_a, B; H) \times cross(r_b, B; V)$ {\bf do} \\
+}
+
+\vgap
+
+\begin{figure} 
+    \begin{tabular}{cc}
+            \hspace{-3mm}
+            \includegraphics[height=30mm]{./artwork/four-segs} &
+            \hspace{-5mm}
+            \includegraphics[height=37mm]{./artwork/oracle} \\ 
+            \hspace{-3mm} (a) Constructing $r_1'$ from $r_1$ & 
+            \hspace{-5mm} (b) Crossing oracles
+    \end{tabular}
+    \figcapup
+    \caption{Illustrations for Section~\ref{sec:hv:type1}}
+    \label{fig:type1}
+    \figcapdown
+\end{figure}
 

### 2024-05-23 msj.tex
yufei

@@
 }
-Define $r_1'$ as the rectangle defined by the above four segments in the way shown in Figure~\ref{fig:type1}(a) ($r_1'$ is the shaded area).
+Define $r_1'$ as the rectangle defined by the above four segments in the way shown in Figure~\ref{fig:hv:type1}(a) ($r_1'$ is the shaded area).
 Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
@@
     \subseteq 
-    \J(R_1, ..., R_{k-2}, H, V). \nn
+    \J(R_1, ..., R_{k-2}, H, V). \label{eqn:hv:type1:S}
 }
@@
 }
-Figure~\ref{fig:type1}(b) shows an example with rectangles $r_1, r_2$, and $r_3$ (here $k = 5$).
+Figure~\ref{fig:hv:type1}(b) shows an example with rectangles $r_1, r_2$, and $r_3$ (here $k = 5$).
 Their intersection is the gray rectangle $B$, which is bottom-bounded by $r_2$ and left-bounded by $r_3$.
@@
 }
-In Figure~\ref{fig:type1}(b), $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2, v_3}$.
+In Figure~\ref{fig:hv:type1}(b), $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2, v_3}$.
 As mentioned before, $r_a = r_2$ and $r_b = r_3$.
@@
     \> 3.\>\> define $B, r_a$, and $r_b$ as above \\
-    \> 4.\>\> {\bf for} each $(h, v) \in \cross(r_a, B; H) \times cross(r_b, B; V)$ {\bf do} \\
+    \> 4.\>\> {\bf for} each $(h, v) \in \cross(r_a, B; H) \times \cross(r_b, B; V)$ {\bf do} \\
+    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $S$
 }
+
+\noindent In Appendix~\ref{app:hv:type1:gen-S}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired.
 
@@
     \caption{Illustrations for Section~\ref{sec:hv:type1}}
-    \label{fig:type1}
+    \label{fig:hv:type1}
     \figcapdown
@@
 
+\section{Correctness of Algorithm generate-S} \label{app:hv:type1:gen-S}
+
 \section{An $\bm{\Omega(n \log n)}$ Lower Bound for 2-SJ} \label{app:lb}

### 2024-05-23 msj.tex
yufei

@@
 This problem holds significant importance in spatial databases and has been extensively studied for over two decades.
-We explain how to settle the problem in $O(n \log n + \out)$ time using $O(n + \out)$ space --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
+We explain how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
 The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
@@
 
-From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time using $O(n)$ space (e.g., by planesweep \cite{bcko08}).
+From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time (e.g., by planesweep \cite{bcko08}).
 However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
@@
 \times R_k$.
-
-\begin{table*} 
-    \begin{tabular}{c|c|c|c|c} 
-        $\bm{k}$ & {\bf method} & {\bf runtime} & {\bf space} & {\bf remark} \\
-        \hline\hline 
-        2 & folklore & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal \\ 
-        \hline
-        $\ge 3$ & before 2022 & $O(n^k)$ & - & \\
-        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n + \out)$ & \\
-        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n + \out)$ & \\
-        \hline
-        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n + \out)$ & comparison-based optimal
-    \end{tabular}
-
-    \vspace{3mm}
-    \caption{Comparison of results on the $k$-SJ problem for a constant $k$}
-    \label{tab:results-com}
-    \figcapdown
-\end{table*}
 
@@
 Tao and Yi \cite{ty22} presented a structure of $O(n \polylog n)$ space achieving the purpose.
-Combining their structure with planesweep, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time using $O(n \polylog n + \out)$ space.
+Combining their structure with planesweep, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
 
@@
 Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time.
-Their algorithm requires $O(n \log n + \out)$ space.
 
@@
 
+\begin{table} 
+    \begin{tabular}{c|c|c|c} 
+        $\bm{k}$ & {\bf method} & {\bf runtime} & {\bf remark} \\
+        \hline\hline 
+        2 & folklore & $O(n \log n + \out)$ & optimal \\ 
+        \hline
+        $\ge 3$ & before 2022 & $O(n^k)$ &  \\
+        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & \\
+        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & \\
+        \hline
+        $\ge 3$ & ours & $O(n \log n + \out)$ & optimal
+    \end{tabular}
+
+    \vspace{3mm}
+    \figcapup 
+    \caption{Result comparison on $k$-SJ problem for a constant $k$}
+    \label{tab:results-com}
+    \figcapdown \vspace{-5mm}
+\end{table}
+
 \subsection{Our Results} \label{sec:intro:ours}
 
-In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time and uses  $O(n + \out)$ space, regardless of the constant $k$.
+In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time, regardless of the constant $k$.
 The time complexity is asymptotically optimal.
@@
 
-We will also prove that the space of our algorithm is $O(k \cdot n + k \cdot \out)$. When $k = O(1)$, the time and space complexities become $O(n \log n$ $+ \out)$ and $O(n + \out)$, as promised.
+When $k = O(1)$, the time complexity becomes $O(n \log n$ $+ \out)$, as promised.
 
@@
 
-In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time and $O(n)$ space where $n = |H| + |R|$.
+In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
 
@@
 
-In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time and $O(n)$ space where $n = |H| + |R|$.
+In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
 
@@
 
-In Appendix~\ref{app:prob-c}, we explain how to solve the problem in $O(n \log n + \out)$ time and $O(n + \out)$ space where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
+In Appendix~\ref{app:prob-c}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
 
@@
 
-In Appendix~\ref{app:prob-d}, we explain how to solve the problem in $O(n \log n + \out)$ time and $O(n + \out)$ space where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.
+In Appendix~\ref{app:prob-d}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.
 
@@
 For this purpose, we will assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
-Denote by
-\myitems{
-    \item $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$;
-    \item $G_\kappa(n, \out)$ the worst-case space consumed by $\A$ on any instance mentioned above.
-}
-We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ and $G_\kappa(n, \out) \le G_{\kappa + 1}(n, \out)$ for any $\kappa \in [2, k-2]$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
+Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
+We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
 
@@
 \begin{lemma} \label{lmm:hv}
-    The H-V $k$-SJ problem can be solved in 
+    The H-V $k$-SJ problem can be solved in time
     \myeqn{
-        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn 
+        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big). \nn 
     }
-    time using $O(n + \out) + G_{k-1}(n, \out)$ space.
 \end{lemma}
@@
 Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.
-
-\begin{figure} 
-    \begin{tabular}{cc} 
-        \includegraphics[height=27mm]{./artwork/type1} & 
-        \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
-        (a) Type 1 result tuple &
-        (b) Type 2 result tuple
-    \end{tabular}
-
-    \figcapup 
-    \caption{Two categories of H-V $k$-SJ result tuples ($k$ = 4)} 
-    \label{fig:hv:types}
-    \figcapdown
-\end{figure}
 
@@
 \subsection{Finding Result Tuples of Type 1} \label{sec:hv:type1}
+
+\begin{figure} 
+    \begin{tabular}{cc} 
+        \includegraphics[height=27mm]{./artwork/type1} & 
+        \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
+        (a) Type 1 result tuple &
+        (b) Type 2 result tuple
+    \end{tabular}
+
+    \figcapup 
+    \caption{Two categories of H-V $k$-SJ result tuples ($k$ = 4)} 
+    \label{fig:hv:types}
+    \figcapdown
+\end{figure}
 
@@
 
-\noindent In Appendix~\ref{app:hv:type1:gen-S}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired.
+\noindent In Appendix~\ref{app:hv:type1:gen-S}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired. The running time of the above algorithm is
+\myeqn{
+    O(1 + |\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))| + |S|) = O(1 + \out) \nn
+}
+and its space is bounded by $O(\out)$.
 

### 2024-05-23 msj.tex
yufei

@@
 By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
-It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time and $O(n)$ space.
+It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time.
 
@@
 Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2},$ $H, V)$, which in turn is bounded by $\out$.
-Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time using $G_{k-2}(n, \out)$ space.
+Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time.
 
@@
 Only the $k$-tuples in $S$ that are of Type 1 will be reported.
-We will achieve the purpose in time $O(n \log n + |S|)$ and space $O(n + |S|)$.
+We will achieve the purpose in $O(n \log n + |S|)$ time.
+
+\begin{figure} 
+    \begin{tabular}{cc}
+            \hspace{-3mm}
+            \includegraphics[height=30mm]{./artwork/four-segs} &
+            \hspace{-5mm}
+            \includegraphics[height=37mm]{./artwork/oracle} \\ 
+            \hspace{-3mm} (a) Constructing $r_1'$ from $r_1$ & 
+            \hspace{-5mm} (b) Crossing oracles
+    \end{tabular}
+    \figcapup
+    \caption{Illustrations for Section~\ref{sec:hv:type1}}
+    \label{fig:hv:type1}
+    \figcapdown
+\end{figure}
 
@@
 
-\noindent In Appendix~\ref{app:hv:type1:gen-S}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired. The running time of the above algorithm is
+\noindent In Appendix~\ref{app:hv:type1:gen-S}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired.
+If we exclude the time to compute all the $\cross(r_a, B; H)$ and $\cross(r_b, B; V)$, the running time of the above algorithm is
 \myeqn{
-    O(1 + |\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))| + |S|) = O(1 + \out) \nn
+    O(1 + |\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))| + |S|) \nn
 }
-and its space is bounded by $O(\out)$.
-
-\vgap
-
-\begin{figure} 
-    \begin{tabular}{cc}
-            \hspace{-3mm}
-            \includegraphics[height=30mm]{./artwork/four-segs} &
-            \hspace{-5mm}
-            \includegraphics[height=37mm]{./artwork/oracle} \\ 
-            \hspace{-3mm} (a) Constructing $r_1'$ from $r_1$ & 
-            \hspace{-5mm} (b) Crossing oracles
-    \end{tabular}
-    \figcapup
-    \caption{Illustrations for Section~\ref{sec:hv:type1}}
-    \label{fig:hv:type1}
-    \figcapdown
-\end{figure}
+which is $O(1 + \out)$ because of \eqref{eqn:hv:type1:S}.
+In what follows, we will explain how to compute all the $\cross(r_a, B; H)$ in $O(n \log n + |S|)$ time.
+A symmetric algorithm can be used to compute
 

### 2024-05-23 msj.tex
yufei

@@
 Only the $k$-tuples in $S$ that are of Type 1 will be reported.
-We will achieve the purpose in $O(n \log n + |S|)$ time.
+We will achieve the purpose in $O(n \log n + |S|) = O(n \log n + \out)$ time.
 
@@
 In what follows, we will explain how to compute all the $\cross(r_a, B; H)$ in $O(n \log n + |S|)$ time.
-A symmetric algorithm can be used to compute
+A symmetric algorithm can be used to compute all the $\cross(r_b, B; V)$ in the same complexity.
+This will then allow us to conclude that Step 2 can be done in $O(n \log n + |S|)$ time.
 

### 2024-05-23 msj.tex
yufei

@@
 \cap r_k \neq \emptyset$.
-This problem holds significant importance in spatial databases and has been extensively studied for over two decades.
+This problem holds significant importance in spatial databases and has been extensively studied in the database community.
 We explain how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
@@
 
-The proof is presented in Appendix~\ref{app:lmm:hv:type1:proj}.
+The proof can be found in Appendix~\ref{app:hv}.
 Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2},$ $H, V)$, which in turn is bounded by $\out$.
@@
 
+Appendix~\ref{app:hv} proves the next property:
+
+\begin{lemma} \label{lmm:hv:type1:cross-non-empty}
+    Neither $\cross(r_a, B; H)$ nor $\cross(r_b, B; H)$ can be empty.
+\end{lemma}
+
+\vgap
+
 Equipped with the above, we can generate the target $S$ as follows:
@@
 \mytab{
-    \> {\bf algorithm generate-$S$} \\
+    \> {\bf Algorithm Generate-$S$} \\
     \> 1.\> $S = \emptyset$ \\ 
@@
 
-\noindent In Appendix~\ref{app:hv:type1:gen-S}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired.
+\noindent In Appendix~\ref{app:hv}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired.
 If we exclude the time to compute all the $\cross(r_a, B; H)$ and $\cross(r_b, B; V)$, the running time of the above algorithm is
@@
 
-\section{Proof of Lemma~\ref{lmm:hv:type1:proj}} \label{app:lmm:hv:type1:proj}
+\section{Completing the Proof of Lemma~\ref{lmm:hv}} \label{app:hv}
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:proj}.}
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-non-empty}.}
+
+\extraspacing {\bf Correctness of Algorithm Generate-$S$.}
 

### 2024-05-23 msj.tex
yufei

@@
 \begin{lemma} \label{lmm:hv:type1:cross-non-empty}
-    Neither $\cross(r_a, B; H)$ nor $\cross(r_b, B; H)$ can be empty.
+    $\cross(r_a, B; H) \ne \emptyset$ and $\cross(r_b, B; H) \ne \emptyset$.
 \end{lemma}
@@
 \noindent In Appendix~\ref{app:hv}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired.
-If we exclude the time to compute all the $\cross(r_a, B; H)$ and $\cross(r_b, B; V)$, the running time of the above algorithm is
-\myeqn{
-    O(1 + |\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))| + |S|) \nn
-}
-which is $O(1 + \out)$ because of \eqref{eqn:hv:type1:S}.
+If we exclude the time to compute all the $\cross(r_a, B; H)$ and $\cross(r_b, B; V)$, the running time of the above algorithm is $O(1 + |S|)$, which is $O(1 + \out)$ because of \eqref{eqn:hv:type1:S}.
 In what follows, we will explain how to compute all the $\cross(r_a, B; H)$ in $O(n \log n + |S|)$ time.

### 2024-05-24 msj.tex
yufei

@@
 
+\def\xleft{\mathrm{left}}
+\def\xright{\mathrm{right}}
+\def\ybot{\mathrm{bot}}
+\def\ytop{\mathrm{top}}
+\def\gleft{\mathrm{left\text{-}guard}}
+\def\gbot{\mathrm{bot\text{-}guard}}
 \def\cross{\mathrm{cross}}
@@
 \cap r_k \neq \emptyset$.
-This problem holds significant importance in spatial databases and has been extensively studied in the database community.
-We explain how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
-The time complexity is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
+The problem holds significant importance in spatial databases and has been extensively studied in the database community.
+In this paper, we show how to settle the problem in $O(n \log n + \out)$ time --- regardless of the constant $k$ --- where $n = \sum_{i=1}^k |R_i|$ and $\out$ is the result size (i.e., the total number of $k$-tuples reported).
+The runtime  is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
 Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{2k} n + \out)$.
@@
 
-This paper revisits the {\em spatial join} (SJ) problem formulated as follows.
+This paper studies the {\em spatial join} (SJ) problem formulated as follows.
 Let $k \ge 2$ be a constant integer.
-In the {\em $k$-SJ} problem, the input comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle $r$ in 2D space is {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
+In the {\em $k$-SJ} problem, the input comprises $k$ sets --- denoted as $R_1, R_2, ..., R_k$ --- of axis-parallel rectangles\footnote{A rectangle is {\em axis-parallel} if it has the form $r = [x_1, x_2] \times [y_1, y_2]$.} in $\real^2$.
 The goal is to find all $k$-tuples $(r_1, r_2, ..., r_k)$ where
@@
 
-SJ is a fundamental operation in spatial databases (SDB), which manage {\em geometric entities} such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
-It plays a crucial role in implementing the {\em filter-refinement mechanism}, which is the dominant approach for computing overlay information in an SDB.
+SJ is a fundamental operation in spatial databases (SDB), which manage geometric entities such as land parcels, service areas, habitat zones, commercial districts, administrative boundaries, etc.
+The operation plays a crucial role in implementing the {\em filter-refinement mechanism}, which is the dominant approach for computing overlay information in an SDB.
 To explain this mechanism, first note that a geometric entity is typically modeled as a polygon.
@@
 \extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$.
-Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle.
-All logarithms have base 2 by default.
+Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $\bm{t} = (e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ...
+\times S_k$ as a $k$-dimensional vector such that $\bm{t}[i] = e_i$ for each $i \in [k]$.
+
+\vgap
+
+Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle. All logarithms have base 2 by default.
 
@@
 From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time (e.g., by planesweep \cite{bcko08}).
-However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario.
+However, the problem becomes much more challenging for $k \ge 3$, known as the {\em multiway} scenario.
 All the solutions developed  before 2022 (see \cite{gcn+13,mp98,mp01,pmt99} and the references therein) suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that enumerates the entire cartesian product $R_1 \times R_2 \times ...
@@
 For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$.
-For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle).
-The objective is to output all $k$-tuples $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ...
-\times R_k$ satisfying $\bigcap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$.
+For each tuple $\bm{t} \in R_i$, its values $\bm{t}(X)$ and $\bm{t}(Y)$ on the two attributes are both intervals (effectively defining a rectangle).
+The objective is to output all $k$-tuples $(\bm{t}_1, \bm{t}_2, ..., \bm{t}_k) \in R_1 \times R_2 \times ...
+\times R_k$ satisfying $\bigcap_{i=1}^k \bm{t}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{t}_i(Y) \ne \emptyset$.
 It is clear that there is one-one correspondence between the result of this join and that of k-SJ.
@@
 It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm solving the $k$-SJ problem, even for $k = 2$.
-This can be established via a reduction (presented in Appendix~\ref{app:lb} for self-containment) from the {\em element distinctness} (ED) problem, where we are given $n$ real values $e_1, e_2, ..., e_n$ and need to decide whether there are distinct $i, j \in [n]$ satisfying $e_i = e_j$.
+This can be established via a reduction from the {\em element distinctness} (ED) problem; see \cite{dl79}.
+
+%where we are given $n$ real values $e_1, e_2, ..., e_n$ and need to decide whether there are distinct $i, j \in [n]$ satisfying $e_i = e_j$.
 The ED problem demands $\Omega(n \log n)$ comparisons to solve \cite{dl79}.
@@
         O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big).
-        \nn
-        %\label{eqn:main:reccurrence}
+        %\nn
+        \label{eqn:main:reccurrence}
     }
@@
 The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory.
-Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly surprising breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
+Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
 This reduction can be inferred from an argument in \cite{kcko22} used to prove a more generic result.
@@
 
+\begin{figure}
+    \includegraphics[height=23mm]{./artwork/guard}
+
+    \figcapup
+    \caption{Suppose $\bm{t} = \set{r_1, r_2, r_3}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
+    \label{fig:guard}
+    \figcapdown
+\end{figure}
+
 \section{Geometric Building Bricks} \label{sec:bricks}
 
-This section will first define some notions useful for our discussion and then introduce four computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
+This section will first define some notions frequently used in our presentation and then introduce several computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
 
@@
 We say that a rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
+
+\vgap
+
+Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$.
+Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $r_i$ ($i \le [k]$) is a rectangle.
+We define $B_\bm{t} = \bigcap_{i=1}^t r_i$, namely, the intersection of the rectangles in $\bm{t}$.
+Note that $B_\bm{t}$ is a rectangle itself.
+We define:
+\myitems{
+    \item $\gleft(\bm{t})$ to be the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$.
+In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
+
+    \item $\gbot(\bm{t})$ to be the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$.
+In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
+}
+See Figure~\ref{fig:guard} for
 
@@
             \hspace{-3mm} (a) Constructing $r_1'$ from $r_1$ & 
-            \hspace{-5mm} (b) Crossing oracles
+            \hspace{-5mm} (b) Generating the set $S$ 
     \end{tabular}

### 2024-05-24 msj.tex
yufei

@@
     \figcapup
-    \caption{Suppose $\bm{t} = \set{r_1, r_2, r_3}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
+    \caption{Define 3-tuple $\bm{t} = \set{r_1, r_2, r_3}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
     \label{fig:guard}
@@
 
-\section{Geometric Building Bricks} \label{sec:bricks}
+\section{Preliminaries in Geometry} \label{sec:bricks}
 
-This section will first define some notions frequently used in our presentation and then introduce several computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
+This section will first define some notions frequently used in our presentation and then introduce several  geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
 
@@
 Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$.
-We say that a horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+\yf{need this?} We say that a horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
 We say that a rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
@@
 }
-See Figure~\ref{fig:guard} for
+See Figure~\ref{fig:guard} for an illustration.
 
 \begin{figure*}
-    \begin{tabular}{ccc}
+    \begin{tabular}{cccc}
         \includegraphics[height=30mm]{./artwork/prob-a} & 
         \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-b} &
+        \includegraphics[height=27mm]{./artwork/prob-b} &
         \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-c} \\
-%         &
-%         \hspace{3mm}
-%         \includegraphics[height=30mm]{./artwork/prob-d} \\
-        (a) Problems $\EuScript{A}$ and $\EuScript{D}$ &
+        \includegraphics[height=30mm]{./artwork/prob-c} &
+        \hspace{3mm}
+        \includegraphics[height=30mm]{./artwork/prob-d} \\[2mm]
+        (a) Problems $\EuScript{A}$ &
         \hspace{3mm}
@@
         (c) Problem $\EuScript{C}$
-%         &
-%         \hspace{3mm}
-%         (d) Problem $\EuScript{D}$
+        &
+        \hspace{3mm}
+        (d) Problem $\EuScript{D}$
     \end{tabular}
@@
 \end{figure*}
+
+\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $P$ of points and set $R$ of rectangles, all in 2D space.
+In the {\em detection version} of Problem $\EuScript{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$.
+Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$.
+In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time, where $n = |P| + |R|$.
+
+\vgap
+
+In the {\em reporting version} of the problem, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$, provided that at least one such $r$ exists.
+For example, in Figure~\ref{fig:guard}a, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$.
+In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
+
+\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space.
+The goal is to report, for each segment $h \in R$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$.
+If $h$ does not intersect with any segment in $V$, report nothing for $h$.
+Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |V|$.
+
+\extraspacing {\bf Problem $\bm{\EuScript{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$.
+If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
+Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+
+\vgap
+
+In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
 
@@
 In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
-
-\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$.
-If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
-Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-
-\vgap
-
-In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
 
@@
 
-\section{Algorithm for Problem $\bm{\EuScript{A}}$} \label{app:prob-a}
+\section{Building Brick Geometry Algorithms} \label{app:bricks}
 
-\section{Algorithm for Problem $\bm{\EuScript{B}}$} \label{app:prob-b}
+\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{A}}$}
 
-\section{Algorithm for Problem $\bm{\EuScript{C}}$} \label{app:prob-c}
+\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{B}}$}
 
-\section{Algorithm for Problem $\bm{\EuScript{D}}$} \label{app:prob-d}
+\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{C}}$}
+
+\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{D}}$}
 

### 2024-05-24 msj.tex
yufei

@@
 \begin{figure*}
-    \begin{tabular}{cccc}
+    \begin{tabular}{ccccc}
         \includegraphics[height=30mm]{./artwork/prob-a} & 
@@
         \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-d} \\[2mm]
-        (a) Problems $\EuScript{A}$ &
+        \includegraphics[height=30mm]{./artwork/prob-d} &
         \hspace{3mm}
-        (b) Problem $\EuScript{B}$ &
+        \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
+        (a) Problems $\mathscr{A}$ &
         \hspace{3mm}
-        (c) Problem $\EuScript{C}$
-        &
+        (b) Problem $\mathscr{B}$ &
         \hspace{3mm}
-        (d) Problem $\EuScript{D}$
+        (c) Problem $\mathscr{C}$ &
+        \hspace{3mm}
+        (d) Problem $\mathscr{D}$ &
+        \hspace{3mm}
+        (3) Problem $\mathscr{E}$
     \end{tabular}
@@
 
-\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $P$ of points and set $R$ of rectangles, all in 2D space.
-In the {\em detection version} of Problem $\EuScript{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$.
+\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of points and set $R$ of rectangles, all in 2D space.
+In the {\em detection version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$.
 Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$.
@@
 
-\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space.
+\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space.
 The goal is to report, for each segment $h \in R$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$.
@@
 
-\extraspacing {\bf Problem $\bm{\EuScript{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$.
+\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the point $(x_1, y)$ and the point $(x, y)$.
 If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
-Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
+
+\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+In the {\em first-crossing} version of the problem, the goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$.
+If such an $h$ does not exist, output nothing for $r$.
+Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
+In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
 
@@
 
-In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
+In the {\em all-crossing} version of the problem, the goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$.
+Furthermore, those segments need to be sorted by their y-coordinates.
+Formally, if $h_1, h_2, ..., h_z$ (for some $z \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
+In the example of Figure~\ref{fig:probs}d, the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.
+In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.
 
-\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$.
-If such an $h$ does not exist, output nothing for $r$.
-Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
-
-\vgap
-
-In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
-
-\extraspacing {\bf Problem $\bm{\EuScript{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
 The goal is to report, for each segment $h \in H$, all the rectangles $r \in R$ containing $h$.
 Furthermore, those rectangles need to be sorted by the x-coordinates of their right boundaries.
-Formally, if $r_1, r_2, ..., r_t$ (for some $t \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_t)$, provided that $x_i > x_{i-1}$ for each $i \in [2, t]$ where $x_i$ (resp., $x_{i-1}$) is the x-coordinate of the right boundary of $r_i$ (resp., $r_{i-1}$).
+Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
 Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
-
-\vgap
-
-In Appendix~\ref{app:prob-c}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
-
-\extraspacing {\bf Problem $\bm{\EuScript{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$.
-Furthermore, those segments need to be sorted by their y-coordinates.
-Formally, if $h_1, h_2, ..., h_t$ (for some $t \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_t)$, provided that $y_i > y_{i-1}$ for each $i \in [2, t]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
-See Figure~\ref{fig:probs}a again where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.
-
-\vgap
-
-In Appendix~\ref{app:prob-d}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.
+In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
 
@@
 }
-Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\EuScript{A}$.
+Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{A}$.
 By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{A}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{B}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{C}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{D}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$}
+
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$}
 

### 2024-05-24 msj.tex
yufei

@@
 
-\section{H-V Multiway Spatial Joins} \label{sec:hv}
+\section{The Core: H-V Multiway Spatial Joins} \label{sec:hv}
 
-This section deals with a special version of the $k$-SJ problem.
-Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ..., R_k$.
-In the special version --- which we name the {\em H-V $k$-SJ} problem ---  we impose the constraints that (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
+Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ...,$ $R_k$.
+We now formulate a special version of $k$-SJ, named the {\em H-V $k$-SJ} problem.
+The special nature is reflected in the introduction of three constraints: (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
 For better clarity, we will represent the input sets as $R_1, R_2, ..., R_{k-2}, H$ ($=R_{k-1}$), and $V$ ($=R_k$).
-Remember that each $R_i$ ($i \in [k-2]$) is a set of rectangles.
-As shown in the next section, solving the H-V $k$-SJ problem is an imperative step for settling the original $k$-SJ problem.
+We want to output the join result $\J(R_1, ..., R_{k-2},$ $H, V)$, including every $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
+\times R_{k-2} \times H \times V$ such that $h \cap v \cap \bigcap_{i=1}^{k-2} r_i$ is not empty.
+As we will see later, conquering this sepcial version is the most challenging step in settling the original $k$-SJ problem optimally.
 
@@
 
-Our objective is to prove that the H-V $k$-SJ problem can be efficiently reduced to the $(k-1)$-SJ problem --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
-For this purpose, we will assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
+Our objective is to prove H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
+For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
 Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
 We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
-
-\vgap
-
-The rest of the section serves as a proof for:
+We will prove:
 
@@
 
-Recall that $n$ is the total size of $R_1, ..., R_{k-2}, H$, and $V$. Henceforth, we assume that a distinct ID from 1 to $n$ is given to each rectangle or segment in $R_1, ..., R_{k-2}, H$, and $V$.
+%The condition is equivalent to saying that the intersection point of $h$ and $v$ falls in every $r_i$ of $i \in [k-2]$.
 
-\subsection{Categorizing the Result Tuples} \label{sec:hv:result-part}
-
-Let us represent the result of the H-V $k$-SJ problem as $\J(R_1, ..., R_{k-2},$ $H, V)$.
-Recall that $\J(R_1, ..., R_{k-2},$ $H, V)$ contains all such $k$-tuples $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
-\times R_{k-2} \times H \times V$ satisfying the condition that $h \cap v \bigcap_{i=1}^{k-2} r_i$ is not empty.
-The condition is equivalent to saying that the intersection point of $h$ and $v$ falls in every $r_i$ of $i \in [k-2]$.
-
-\vgap
-
-Let us focus on any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in  \J(R_1, ..., R_{k-2},$ $H, V)$. We classify the tuple into one of the two categories below:
+To get started, consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ in the join result $\J(R_1, ..., R_{k-2}, H, V)$. We classify the tuple into one of the two types below:
 
@@
 
-    \item {{\bf Type 2:}} $h$ or $v$ fails to cross some rectangle in $\set{r_1, r_2, ..., r_{k-2}}$.
-This means the existence of at least one rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
+    \item {{\bf Type 2:}} $h$ or $v$ fails to cross some rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
+    there exists at least one rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
 }
@@
 Before proceeding, however, let us observe a sense of symmetry for Type 2.
-As mentioned, for a result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
+As mentioned, for each result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
 As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc.
@@
         \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
-        (a) Type 1 result tuple &
-        (b) Type 2 result tuple
+        (a) Type 1 &
+        (b) Type 2
     \end{tabular}
@@
     \figcapup 
-    \caption{Two categories of H-V $k$-SJ result tuples ($k$ = 4)} 
+    \caption{Classifying H-V $k$-SJ result tuples ($k$ = 4)}
     \label{fig:hv:types}

### 2024-05-24 msj.tex
yufei

@@
 \begin{theorem} \label{thm:main-recur}
-    For any $n \ge 1$, equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in time
+    Equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in time
     \myeqn{
-        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big).
+        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big)
         %\nn
@@
     }
+    where $n$ (resp., $\out$) is the input (resp., output) size of the problem.
 %     time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
@@
 We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
-We will prove:
+
+\vgap
+
+We will establish:
 
@@
     \myeqn{
-        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big). \nn 
+        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
     }
+    where $n$ (resp., $\out$) is the input (resp., output) size of the problem.
 \end{lemma}
@@
 
-We will find result tuples of each type separately.
+In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.
+
+%We will find result tuples of each type separately.
 Before proceeding, however, let us observe a sense of symmetry for Type 2.
@@
 \vgap 
-
-Our subsequent discussion will proceed as follows.
+%
+% Our subsequent discussion will proceed as follows.
 Section~\ref{sec:hv:type1} will explain how to produce all the result tuples of Type 1.
@@
 Other sub-types can be computed using the same algorithm, by symmetry.
-
-\vgap
-
-A remark is in order about duplicate removal.
+%
+% \vgap
+%
+% A remark is in order about duplicate removal.
 A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types.
@@
 
-\subsection{Finding Result Tuples of Type 1} \label{sec:hv:type1}
+\section{H-V $k$-SJ: Result Tuples of Type 1} \label{sec:hv:type1}
 
@@
 
-\subsection{Finding Result Tuples of Type 2} \label{sec:hv:type2}
+\section{Finding Result Tuples of Type 2} \label{sec:hv:type2}
 

### 2024-05-24 msj.tex
yufei

@@
 
-\section{H-V $k$-SJ: Result Tuples of Type 1} \label{sec:hv:type1}
-
 \begin{figure}
@@
 
-This subsection will explain how to find the $k$-tuples of Type 1.
-Let us represent the set of such tuples as $\J_1(R_1, ..., R_{k-2}, H, V)$ and define its ``projection'' onto $(R_1, ..., R_{k-2})$ as
+\section{H-V $k$-SJ: Result Tuples of Type 1} \label{sec:hv:type1}
+
+As before, let $R_1, ..., R_{k-2}, H$, and $V$ be the input sets of the H-V $k$-SJ problem.
+Denote by $\J_1$ the set of type-1 result tuples defined in Section~\ref{sec:hv}.
+In this section, we aim to compute a set $\J^*$ satisfying
 \myeqn{
-    && \Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V)) \nn \\
-    &=& \{(r_1, ..., r_{k-2}) \mid (r_1, ..., r_{k-2}, h, v) \in \J_1(R_1, ..., R_{k-2}, H, V)
-    \}. \nn
+    \J_1
+    \subseteq
+    \J^*
+    \subseteq
+    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:S}
 }
-We will compute $\J_1(R_1, ..., R_{k-2}, H, V)$ in two steps:
-\myitems{
-    \item {{\bf Step 1:}} Compute $\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))$;
-    \item {{\bf Step 2:}} Generate $\J_1(R_1, ..., R_{k-2}, H, V)$ from the result of Step 1.
-}
+where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ. From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.
 
-\noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ. To construct the instance, for each rectangle $r_1 \in R_1$, we compute four segments:
+\extraspacing {\bf The Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} For each rectangle $r_1 \in R_1$, we compute four segments:
 \myitems{
@@
 }
-Define $r_1'$ as the rectangle defined by the above four segments in the way shown in Figure~\ref{fig:hv:type1}(a) ($r_1'$ is the shaded area).
+Define $r_1'$ as the rectangle bounded by the above four segments as shown in Figure~\ref{fig:hv:type1}(a) ($r_1'$ is the shaded area).
 Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
@@
 \myeqn{
-    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}}. \nn
+    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}} \nn
 }
-Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{A}$.
+which we refer to as the {\em trimmed} version of $R_1$.
+Computing the ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{D}$ (the first-crossing version).
 By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.

### 2024-05-24 msj.tex
yufei

@@
 \begin{lemma} \label{lmm:hv}
-    The H-V $k$-SJ problem can be solved in time
+    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
     \myeqn{
@@
 \vgap
+
+In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, n]$, spending $O(k n \log n)$ time in total.
 

### 2024-05-24 msj.tex
yufei

@@
 
+\yf{Rahul's result here}
+
 %We will find result tuples of each type separately.
@@
 }
-where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ. From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.
+where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ.
+From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.
+Remember that $\out$ is the size of $\J(R_1, ..., R_{k-2}, H, V)$.
 
@@
 }
-Define $r_1'$ as the rectangle bounded by the above four segments as shown in Figure~\ref{fig:hv:type1}(a) ($r_1'$ is the shaded area).
+Define $r_1'$ as the rectangle bounded by the above four segments as shown in Figure~\ref{fig:hv:type1:trim} ($r_1'$ is the shaded area).
+We say that $r_1'$ is the {\em trimmed rectangle} of $r_1$, and conversely, $r_1$ is the {\em full rectangle} of $r_1'$.
 Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
@@
 In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, n]$, spending $O(k n \log n)$ time in total.
+We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R_1', R_2, ..., R_{k-2})$.
+We prove in Appendix~\ref{app:hv}:
 
-We now solve the $(k-2)$-SJ problem on $R_1', R_2, ..., R_{k-2}$ using the algorithm $\A$ supplied.
-Let us represent the result of this $(k-2)$-SJ as $\J(R_1', R_2, ..., R_{k-2})$.
-The lemma below shows that the result is essentially what we aim to compute in Step 1.
+%The lemma below shows that the result is essentially what we aim to compute in Step 1.
 
 \begin{lemma} \label{lmm:hv:type1:proj}
-    A $(k-2)$-tuple $(r_1', r_2, ..., r_{k-2})$ exists in $\J(R_1', R_2, ...,$ $R_{k-2})$ if and only if the $(k-2)$-tuple $(r_1, r_2, ..., r_{k-2})$ is in $\Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$, where $r_1$ is the rectangle in $R_1$ defining $r_1'$.
+    $|\J(R_1', R'_2, ..., R'_{k-2})| \le \out$.
 \end{lemma}
 
-The proof can be found in Appendix~\ref{app:hv}.
-Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2},$ $H, V)$, which in turn is bounded by $\out$.
-Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time.
+We reminder the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
 
-\extraspacing {\bf Step 2.} We are ready to explain how to produce $\J_1(R_1, ..., R_{k-2}, H, V)$ from $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$.
-Our algorithm for this purpose will find a set $S$ satisfying
-\myeqn{
-    \J_1(R_1, ..., R_{k-2}, H, V) 
-    \subseteq 
-    S
-    \subseteq 
-    \J(R_1, ..., R_{k-2}, H, V). \label{eqn:hv:type1:S}
-}
-Recall that $\J(R_1, ..., R_{k-2}, H, V)$ is the result of the original H-V $k$-SJ problem, which has a size of $\out$.
-Only the $k$-tuples in $S$ that are of Type 1 will be reported.
-We will achieve the purpose in $O(n \log n + |S|) = O(n \log n + \out)$ time.
+\yf{example}
 
 \begin{figure}
-    \begin{tabular}{cc}
-            \hspace{-3mm}
-            \includegraphics[height=30mm]{./artwork/four-segs} &
-            \hspace{-5mm}
-            \includegraphics[height=37mm]{./artwork/oracle} \\ 
-            \hspace{-3mm} (a) Constructing $r_1'$ from $r_1$ & 
-            \hspace{-5mm} (b) Generating the set $S$ 
-    \end{tabular}
+    \includegraphics[height=30mm]{./artwork/four-segs}
     \figcapup
-    \caption{Illustrations for Section~\ref{sec:hv:type1}}
-    \label{fig:hv:type1}
+    \caption{Trimming}
+    \label{fig:hv:type1:trim}
     \figcapdown
@@
 
-\vgap
-
-Now, take an arbitrary $(k-2)$-tuple in $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$. Denote by 
-\myeqn{
-    B = \cap_{i=1}^{k-2} r_i. \nn
-}
-Note that $B$ is a rectangle itself. We say that $B$ is 
-\myitems{
-    \item {\em bottom-bounded} by $r_a$, if the bottom edge of $B$ is a part of the bottom boundary of $r_a$ for some $a \in [k-2]$;
-
-    \item {\em left-bounded} by $r_b$, if the left edge of $B$ is a part of the left boundary of $r_b$ for some $b \in [k-2]$.
-}
-Figure~\ref{fig:hv:type1}(b) shows an example with rectangles $r_1, r_2$, and $r_3$ (here $k = 5$).
-Their intersection is the gray rectangle $B$, which is bottom-bounded by $r_2$ and left-bounded by $r_3$.
+\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple in $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$.
+The reader should recall from Section~\ref{sec:bricks} that $B_\bm{t}$ is the intersection of of $r'_1, r'_2, ..., r'_{k-2}$, $\gleft(\bm{t})$ is the $r_i$ ($i \in [k-2]$) satisfying $\xleft(r'_i) = \xleft(B_\bm{t})$, and $\gbot(\bm{t})$ is the $r_i$ ($i \in [k-2]$) satisfying $\ybot(r'_i) = \ybot(B_\bm{t})$.
 
@@
 
-We aim to compute 
-\myitems{
-    \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$; 
-    \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
+We now introduce:
+% \myitems{
+%     \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$;
+%     \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
+% }
+\myeqn{
+    \cross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
+    \cross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
 }
-In Figure~\ref{fig:hv:type1}(b), $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2, v_3}$.
-As mentioned before, $r_a = r_2$ and $r_b = r_3$.
-We have $\cross(r_a, B; H) = \set{h_1, h_2}$ and $\cross(r_b, B; V) = \set{v_1, v_3}$.
-Note that $h_3$ is not in $\cross(r_a, B; H)$ because $h_3$ does not cross $B$, and $v_2$ is not in $\cross(r_b, B; V)$ because $v_2$ does not cross $r_3$.
-
-\vgap
-
-Appendix~\ref{app:hv} proves the next property:
+These sets have important properties, as proved in Appendix~\ref{app:hv}:
 
 \begin{lemma} \label{lmm:hv:type1:cross-non-empty}
-    $\cross(r_a, B; H) \ne \emptyset$ and $\cross(r_b, B; H) \ne \emptyset$.
+    Suppose that $r_1, r_2, ..., r_{k-2}$ are the full rectangles of $r_1', r_2', ..., r_{k-2}'$, respectively. All the following statements are true:
+    \myenums{
+        \item $\cross(\bm{t}, H) \ne \emptyset$ and $\cross(\bm{t}, V) \ne \emptyset$.
+    }
 \end{lemma}
+
+\yf{example}
 
@@
 \section{Finding Result Tuples of Type 2} \label{sec:hv:type2}
-
-\subsection{Putting Things Together} \label{sec:hv:analysis}
-
-\todo{Remark on Rahul's result}
 

### 2024-05-24 msj.tex
yufei

@@
 \begin{lemma} \label{lmm:hv:type1:cross-non-empty}
-    Suppose that $r_1, r_2, ..., r_{k-2}$ are the full rectangles of $r_1', r_2', ..., r_{k-2}'$, respectively. All the following statements are true:
+    Let $r_i$ be the full rectangle of $r_i'$ for each $i \in [k-2]$. All the following statements are true:
     \myenums{
         \item $\cross(\bm{t}, H) \ne \emptyset$ and $\cross(\bm{t}, V) \ne \emptyset$.
+        \item For any $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+        \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$.
     }

### 2024-05-24 msj.tex
yufei

@@
 
-\begin{lemma} \label{lmm:hv:type1:cross-non-empty}
+\begin{lemma} \label{lmm:hv:type1:properties}
     Let $r_i$ be the full rectangle of $r_i'$ for each $i \in [k-2]$. All the following statements are true:
@@
         \item $\cross(\bm{t}, H) \ne \emptyset$ and $\cross(\bm{t}, V) \ne \emptyset$.
-        \item For any $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+        \item For any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
         \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$.
@@
 
-Equipped with the above, we can generate the target $S$ as follows:
+Equipped Lemma~\ref{lmm:hv:type1:properties}, we can generate the target $S$ as follows:
 
 \mytab{
-    \> {\bf Algorithm Generate-$S$} \\
+    \> {\bf Algorithm Generate-$\J^*$} \\
     \> 1.\> $S = \emptyset$ \\ 

### 2024-05-24 msj.tex
yufei

@@
 }
-These sets have important properties, as proved in Appendix~\ref{app:hv}:
+These sets have important properties, as proved in Appendix~\ref{app:hv} (recall the notation $\bm{t}[i] = r_i'$ for each $i \in [k-2]$):
 
 \begin{lemma} \label{lmm:hv:type1:properties}
-    Let $r_i$ be the full rectangle of $r_i'$ for each $i \in [k-2]$. All the following statements are true:
+    %Let $r_i$ be the full rectangle of $r_i'$ for each $i \in [k-2]$.
+    All the following statements are true:
+    \myenums{
+        \item For each $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, we have:
+
+        \vspace{2mm}
+
         \myenums{
             \item $\cross(\bm{t}, H) \ne \emptyset$ and $\cross(\bm{t}, V) \ne \emptyset$.
-        \item For any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+
+            \item Let $r_i$ ($i \in [k-2]$) be the full rectangle of $\bm{t}[i']$. Then,
+            for any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+
             \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$.
+        }
+
+        \vspace{2mm}
+
+        \item $\sum_{\bm{t}} |\cross(\bm{t}, H)| \le \out$ and $\sum_{\bm{t}} |\cross(\bm{t}, V)| \le \out$.
     }
@@
     \> {\bf Algorithm Generate-$\J^*$} \\
-    \> 1.\> $S = \emptyset$ \\ 
-    \> 2.\> {\bf for} each $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$ {\bf do} \\
-    \> 3.\>\> define $B, r_a$, and $r_b$ as above \\
-    \> 4.\>\> {\bf for} each $(h, v) \in \cross(r_a, B; H) \times \cross(r_b, B; V)$ {\bf do} \\
-    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $S$
+    \> 1.\> $\J^* = \emptyset$ \\
+    \> 2.\> {\bf for} each $(k-2)$-tuple $\bm{t} \in \J(R_1,$ $...,$ $R_{k-2}))$ {\bf do} \\
+    \> 3.\>\> $r_i \leftarrow$ the full rectangle of $\bm{t}[i]$, for each $i \in [k-2]$ \\
+    \>\>\> /* recall that $\bm{t}[i]$ is the $i$-th rectangle in $\bm{t}$ */ \\
+    \> 4.\>\> {\bf for} each $(h, v) \in \cross(\bm{t}, H) \times \cross(\bm{t}, V)$ {\bf do} \\
+    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $\J^*$
 }

### 2024-05-24 msj.tex
yufei

@@
     \subseteq
-    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:S}
+    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:J*}
 }
@@
 
-Equipped Lemma~\ref{lmm:hv:type1:properties}, we can generate the target $S$ as follows:
+Equipped Lemma~\ref{lmm:hv:type1:properties}, we can generate our target $\J^*$ as follows:
 
@@
     \> 3.\>\> $r_i \leftarrow$ the full rectangle of $\bm{t}[i]$, for each $i \in [k-2]$ \\
-    \>\>\> /* recall that $\bm{t}[i]$ is the $i$-th rectangle in $\bm{t}$ */ \\
     \> 4.\>\> {\bf for} each $(h, v) \in \cross(\bm{t}, H) \times \cross(\bm{t}, V)$ {\bf do} \\
@@
 
-\noindent In Appendix~\ref{app:hv}, we prove that the set $S$ computed satisfies \eqref{eqn:hv:type1:S}, as desired.
-If we exclude the time to compute all the $\cross(r_a, B; H)$ and $\cross(r_b, B; V)$, the running time of the above algorithm is $O(1 + |S|)$, which is $O(1 + \out)$ because of \eqref{eqn:hv:type1:S}.
-In what follows, we will explain how to compute all the $\cross(r_a, B; H)$ in $O(n \log n + |S|)$ time.
-A symmetric algorithm can be used to compute all the $\cross(r_b, B; V)$ in the same complexity.
-This will then allow us to conclude that Step 2 can be done in $O(n \log n + |S|)$ time.
+\noindent Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
+Furthermore, if we are given $\cross(\bm{t}, H)$ and $\cross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + |\J^*|) = O(1 + \out)$ time.
 

### 2024-05-24 msj.tex
yufei

@@
 
-\section{Building Brick Geometry Algorithms} \label{app:bricks}
+\section{Building Brick Algorithms} \label{app:bricks}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.}
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.}
 
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-non-empty}.}
-
-\extraspacing {\bf Correctness of Algorithm Generate-$S$.}
-
-\section{Correctness of Algorithm generate-S} \label{app:hv:type1:gen-S}
-
-\section{An $\bm{\Omega(n \log n)}$ Lower Bound for 2-SJ} \label{app:lb}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
 

### 2024-05-24 msj.tex
yufei

@@
 \times S_k$ as a $k$-dimensional vector such that $\bm{t}[i] = e_i$ for each $i \in [k]$.
-
-\vgap
-
-Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle. All logarithms have base 2 by default.
+Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle.
+All logarithms have base 2 by default.
 
@@
 
-In this paper, we show that the $k$-SJ problem can be settled by a comparison-based algorithm that runs in $O(n \log n + \out)$ time, regardless of the constant $k$.
-The time complexity is asymptotically optimal.
+In this paper, we solve the $k$-SJ problem with a comparison-based algorithm that runs in $O(n \log n + \out)$ time regardless of the constant $k$. The time complexity is asymptotically optimal.
 
@@
 For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$.
-Our main technical contribution is to establish:
+Our core technical contribution is to establish:
 
@@
 When $k = O(1)$, the time complexity becomes $O(n \log n$ $+ \out)$, as promised.
-
-\vgap
-
 Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result for $k = O(1)$ in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
-The answer, interestingly, is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory.
+Interestingly, the answer is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory.
 Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
@@
 \extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$.
+We say that a horizontal segment $h_1$ is {\em lower} (resp., {\em higher}) than another horizontal segment $h_2$ if the y-coordinate of $h_1$ is smaller (resp., larger) than that of $h_2$.
+Similarly, a vertical segment $v_1$ is {\em to the left} (resp., {\em right}) {\em of} another vertical segment $v_2$ if the x-coordinate of $v_1$ is smaller (resp., larger) than that of $v_2$.
+
 Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
 Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$.
-\yf{need this?} We say that a horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
-We say that a rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
+\yf{need this?} A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
 
@@
 Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$.
-Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $r_i$ ($i \le [k]$) is a rectangle.
-We define $B_\bm{t} = \bigcap_{i=1}^t r_i$, namely, the intersection of the rectangles in $\bm{t}$.
+Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
+We use $B_\bm{t}$ to represent $\bigcap_{i=1}^t r_i$, namely, the intersection of the rectangles in $\bm{t}$.
 Note that $B_\bm{t}$ is a rectangle itself.
@@
 \myitems{
-    \item $\gleft(\bm{t})$ to be the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$.
+    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$.
 In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
 
-    \item $\gbot(\bm{t})$ to be the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$.
+    \item $\gbot(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$.
 In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
@@
         \hspace{3mm}
-        (3) Problem $\mathscr{E}$
+        (e) Problem $\mathscr{E}$
     \end{tabular}
@@
 \extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space.
-The goal is to report, for each segment $h \in R$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$.
+The goal is to report, for each segment $h \in H$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$.
 If $h$ does not intersect with any segment in $V$, report nothing for $h$.
@@
 \extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the point $(x_1, y)$ and the point $(x, y)$.
-If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that at least one rectangle $r \in R$ covers both the point $(x_1, y)$ and the point $(x, y)$.
+If the point $p$ exists (i.e., $h$ has at least one left-end covering rectangle in $R$), we should output a tuple $(h, p)$; otherwise, output nothing for $h$.
 Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
@@
 \extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-In the {\em first-crossing} version of the problem, the goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$.
+In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$.
 If such an $h$ does not exist, output nothing for $r$.
@@
 
-In the {\em all-crossing} version of the problem, the goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$.
-Furthermore, those segments need to be sorted by their y-coordinates.
+In the {\em find-all} version of the problem, the goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$.
+Furthermore, those segments need to be sorted by y-coordinate.
 Formally, if $h_1, h_2, ..., h_z$ (for some $z \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
@@
 \mytab{
-    \> {\bf Algorithm Generate-$\J^*$} \\
+    \> {\bf algorithm generate-$\J^*$} \\
     \> 1.\> $\J^* = \emptyset$ \\
-    \> 2.\> {\bf for} each $(k-2)$-tuple $\bm{t} \in \J(R_1,$ $...,$ $R_{k-2}))$ {\bf do} \\
+    \> 2.\> {\bf for} each $(k-2)$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ {\bf do} \\
     \> 3.\>\> $r_i \leftarrow$ the full rectangle of $\bm{t}[i]$, for each $i \in [k-2]$ \\

### 2024-05-24 msj.tex
yufei

@@
 For better clarity, we will represent the input sets as $R_1, R_2, ..., R_{k-2}, H$ ($=R_{k-1}$), and $V$ ($=R_k$).
-We want to output the join result $\J(R_1, ..., R_{k-2},$ $H, V)$, including every $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
+The goal is to output the join result $\J(R_1, ..., R_{k-2},$ $H, V)$, including every $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
 \times R_{k-2} \times H \times V$ such that $h \cap v \cap \bigcap_{i=1}^{k-2} r_i$ is not empty.
-As we will see later, conquering this sepcial version is the most challenging step in settling the original $k$-SJ problem optimally.
+As we will see, conquering this sepcial version is the most challenging step in settling the original $k$-SJ problem optimally.
 
@@
 
-Our objective is to prove H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
+Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
 For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
@@
 
-To get started, consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ in the join result $\J(R_1, ..., R_{k-2}, H, V)$. We classify the tuple into one of the two types below:
+Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ in the join result $\J(R_1, ...,$ $R_{k-2}, H, V)$. We classify the tuple into one of the two types below:
 
 \myitems{
-    \item {{\bf Type 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$, and $v$  crosses all of $r_1, ..., r_{k-2}$;
+    \item {{\bf Type 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$ and, at the same time, $v$ crosses all of $r_1, ..., r_{k-2}$;
 
-    \item {{\bf Type 2:}} $h$ or $v$ fails to cross some rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
-    there exists at least one rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
+    \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
+    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
 }

### 2024-05-24 msj.tex
yufei

@@
 
-\yf{Rahul's result here}
+\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently.
+They developed a structure of $O(n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
+Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
+We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.
+Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas.
 

### 2024-05-24 msj.tex
yufei

@@
 where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ.
+Remember that the output size $\out$ is defined to $|\J(R_1, ..., R_{k-2}, H, V)|$.
 From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.
-Remember that $\out$ is the size of $\J(R_1, ..., R_{k-2}, H, V)$.
 
-\extraspacing {\bf The Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} For each rectangle $r_1 \in R_1$, we compute four segments:
+\extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} For each rectangle $r_1 \in R_1$, we compute four segments:
 \myitems{
-    \item $h_1$ (resp., $h_2$): the lowest (resp., highest) segment --- i.e., with the smallest (resp., largest) y-coordinate --- among all the segments in $H$ that cross $r$;
+    \item $h_1$ (resp., $h_2$): the lowest (resp., highest) segment in $H$ that crosses $r$;
 
-    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment --- i.e., with the smallest (resp., largest) x-coordinate --- among all the segments in $V$ that cross $r$.
+    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment in $V$ that crosses $r$.
 }
-Define $r_1'$ as the rectangle bounded by the above four segments as shown in Figure~\ref{fig:hv:type1:trim} ($r_1'$ is the shaded area).
+Define $r_1' = [x_1, x_2] \times [y_1, y_2]$, where $x_1$ (resp., $x_2$) is x-coordinate of $v_1$ (resp., $v_2$), and $y_1$ (resp., $y_2$) is y-coordinate of $h_1$ (resp., $h_2$).
 We say that $r_1'$ is the {\em trimmed rectangle} of $r_1$, and conversely, $r_1$ is the {\em full rectangle} of $r_1'$.
@@
 }
-which we refer to as the {\em trimmed} version of $R_1$.
-Computing the ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{D}$ (the first-crossing version).
+which we call the {\em trimmed} version of $R_1$.
+Computing the ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{D}$ (the find-lowest version).
 By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
@@
 
-\begin{figure}
-    \includegraphics[height=30mm]{./artwork/four-segs}
-    \figcapup
-    \caption{Trimming}
-    \label{fig:hv:type1:trim}
-    \figcapdown
-\end{figure}
+% \begin{figure}
+%     \includegraphics[height=30mm]{./artwork/four-segs}
+%     \figcapup
+%     \caption{Trimming}
+%     \label{fig:hv:type1:trim}
+%     \figcapdown
+% \end{figure}
 

### 2024-05-24 msj.tex
yufei

@@
 
-\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple in $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$.
-The reader should recall from Section~\ref{sec:bricks} that $B_\bm{t}$ is the intersection of of $r'_1, r'_2, ..., r'_{k-2}$, $\gleft(\bm{t})$ is the $r_i$ ($i \in [k-2]$) satisfying $\xleft(r'_i) = \xleft(B_\bm{t})$, and $\gbot(\bm{t})$ is the $r_i$ ($i \in [k-2]$) satisfying $\ybot(r'_i) = \ybot(B_\bm{t})$.
-
-\vgap
-
+\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$.
+The reader should recall from Section~\ref{sec:bricks} that
+\myitems{
+    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i]$, i.e., the intersection of $r'_1, r'_2, ..., r'_{k-2}$;
+    \item $\gleft(\bm{t})$ is the $r_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
+    \item $\gbot(\bm{t})$ is the $r_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
+}
 We now introduce:
@@
 }
-These sets have important properties, as proved in Appendix~\ref{app:hv} (recall the notation $\bm{t}[i] = r_i'$ for each $i \in [k-2]$):
+These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:
 
@@
 
-            \item Let $r_i$ ($i \in [k-2]$) be the full rectangle of $\bm{t}[i']$. Then,
-            for any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+            \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i']$. 
+            For any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
 
@@
 
-Equipped Lemma~\ref{lmm:hv:type1:properties}, we can generate our target $\J^*$ as follows:
+Equipped with Lemma~\ref{lmm:hv:type1:properties}, we generate our target $\J^*$ as follows:
 
@@
 
-\noindent Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
+Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
 Furthermore, if we are given $\cross(\bm{t}, H)$ and $\cross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + |\J^*|) = O(1 + \out)$ time.

### 2024-05-24 msj.tex
yufei

@@
 If $h$ does not intersect with any segment in $V$, report nothing for $h$.
-Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |V|$.
@@
 Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
-Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
+Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
@@
 
-In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, n]$, spending $O(k n \log n)$ time in total.
+In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, k]$, spending $O(k n \log n)$ time in total.
 We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
@@
 
-We reminder the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
+We remind the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
 
@@
 
-            \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i']$. 
+            \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i]$. 
             For any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
 
-            \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$.
+            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$.
         }

### 2024-05-24 msj.tex
yufei

@@
 Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
-Furthermore, if we are given $\cross(\bm{t}, H)$ and $\cross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + |\J^*|) = O(1 + \out)$ time.
+Furthermore, if we are given $\cross(\bm{t}, H)$ and $\cross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
+
+\vgap
+
+The rest of the section will focus on explaining how to prepare the sets $\cross(\bm{t}, H)$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
+An analogous algorithm can be used to compute the sets $\cross(\bm{t}, V)$ of all $\bm{t}$ in the same time complexity.
 

### 2024-05-24 msj.tex
yufei

@@
 
-When $k = O(1)$, the time complexity becomes $O(n \log n$ $+ \out)$, as promised.
+When $k = O(1)$, the time complexity becomes $O(n \log n$ $+ \out)$, as promised; the space consumption of our algorithm is $O(n + \out)$.
 Now that Theorem~\ref{thm:main-alg} offers a satisfactory $k$-SJ result for $k = O(1)$ in 2D space, it is natural to wonder whether the constraint on dimensionality 2 is necessary.
@@
 Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
-For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$.
+To facilitate our analysis, let us define the 1-SJ problem as follows: the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$.
+The 1-SJ problem can be trivially ``solved'' in $O(n)$ time.
+We assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [1, k-1]$.
 Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
-We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
+We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \ge 1$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
 
@@
 \extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently.
-They developed a structure of $O(n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
+They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
 Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
@@
 
+\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}
+
 \section{Finding Result Tuples of Type 2} \label{sec:hv:type2}

### 2024-05-24 msj.tex
yufei

@@
 
-\def\vgap{\vspace{1mm}}
+\def\vgap{\vspace{0mm}}
 \def\extraspacing{\vspace{2mm} \noindent}
@@
 \def\ytop{\mathrm{top}}
+\def\maxtop{\mathrm{maxtop}}
 \def\gleft{\mathrm{left\text{-}guard}}
@@
 
-\extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} For each rectangle $r_1 \in R_1$, we compute four segments:
+\extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r \in R_i$, we compute four segments:
 \myitems{
@@
 }
-Define $r_1' = [x_1, x_2] \times [y_1, y_2]$, where $x_1$ (resp., $x_2$) is x-coordinate of $v_1$ (resp., $v_2$), and $y_1$ (resp., $y_2$) is y-coordinate of $h_1$ (resp., $h_2$).
-We say that $r_1'$ is the {\em trimmed rectangle} of $r_1$, and conversely, $r_1$ is the {\em full rectangle} of $r_1'$.
-Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
+Define $r' = [x_1, x_2] \times [y_1, y_2]$, where $x_1$ (resp., $x_2$) is x-coordinate of $v_1$ (resp., $v_2$), and $y_1$ (resp., $y_2$) is y-coordinate of $h_1$ (resp., $h_2$).
+We say that $r'$ is the {\em trimmed rectangle} of $r$, and conversely, $r$ is the {\em full rectangle} of $r'$.
+Note that $r'$ exists if and only if $r$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
 Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$.
-Define
-\myeqn{
-    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}} \nn
-}
-which we call the {\em trimmed} version of $R_1$.
-Computing the ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{D}$ (the find-lowest version).
-By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
-It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time.
 
@@
 
-In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, k]$, spending $O(k n \log n)$ time in total.
+Construct
+\myeqn{
+    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \nn
+}
+Computing the ``segment $h_1$'' for each $r \in R_1$ is an instance of Problem $\mathscr{D}$ (the find-lowest version).
+By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
+It thus follows from Section~\ref{sec:bricks} that $R_i'$ can be produced in $O(n \log n)$ time.
+
+\vgap
+
 We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
@@
     \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i]$, i.e., the intersection of $r'_1, r'_2, ..., r'_{k-2}$;
-    \item $\gleft(\bm{t})$ is the $r_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
-    \item $\gbot(\bm{t})$ is the $r_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
+    \item $\gleft(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
+    \item $\gbot(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
 }
@@
 
-\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}
+\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r' \in R_i'$, we define
+\myeqn{
+    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): r' = \gbot(\bm{t})} \xright(B_\bm{t}). 
+}
+The value $\maxtop(r')$ is set to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
+When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
+\myeqn{
+    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
+}
+We refer to $r^*$ as the {\em top-sliced rectangle} of $r'$.
+
+\yf{example}
+
+\vgap
+
+Now construct from $R_i'$ a new set of rectangles: 
+\myeqn{
+    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \nn
+}
 

### 2024-05-24 msj.tex
yufei

@@
 \def\cross{\mathrm{cross}}
+\def\dcross{\mathrm{d\text{-}cross}}
 \def\out{\mathrm{OUT}}
@@
 \myeqn{
-    \cross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
-    \cross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
+    \dcross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
+    \dcross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
 }
-These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:
+The prefix ``d-'' stands for double. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:
 
@@
         \myenums{
-            \item $\cross(\bm{t}, H) \ne \emptyset$ and $\cross(\bm{t}, V) \ne \emptyset$.
+            \item $\dcross(\bm{t}, H) \ne \emptyset$ and $\dcross(\bm{t}, V) \ne \emptyset$.
 
             \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i]$. 
-            For any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+            For any $h \in \dcross(\bm{t}, H)$ and any $v \in \dcross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
 
-            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$.
+            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \dcross(\bm{t}, H)$ and $v \in \dcross(\bm{t}, V)$.
         }
@@
 
-        \item $\sum_{\bm{t}} |\cross(\bm{t}, H)| \le \out$ and $\sum_{\bm{t}} |\cross(\bm{t}, V)| \le \out$.
+        \item $\sum_{\bm{t}} |\dcross(\bm{t}, H)| \le \out$ and $\sum_{\bm{t}} |\dcross(\bm{t}, V)| \le \out$.
     }
@@
     \> 3.\>\> $r_i \leftarrow$ the full rectangle of $\bm{t}[i]$, for each $i \in [k-2]$ \\
-    \> 4.\>\> {\bf for} each $(h, v) \in \cross(\bm{t}, H) \times \cross(\bm{t}, V)$ {\bf do} \\
+    \> 4.\>\> {\bf for} each $(h, v) \in \dcross(\bm{t}, H) \times \dcross(\bm{t}, V)$ {\bf do} \\
     \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $\J^*$
@@
 Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
-Furthermore, if we are given $\cross(\bm{t}, H)$ and $\cross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
+Furthermore, if we are given $\dcross(\bm{t}, H)$ and $\dcross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
 
@@
 
-The rest of the section will focus on explaining how to prepare the sets $\cross(\bm{t}, H)$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
-An analogous algorithm can be used to compute the sets $\cross(\bm{t}, V)$ of all $\bm{t}$ in the same time complexity.
+The rest of the section will focus on explaining how to prepare the sets $\dcross(\bm{t}, H)$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
+An analogous algorithm can be used to compute the sets $\dcross(\bm{t}, V)$ of all $\bm{t}$ in the same time complexity.
 
@@
 \myeqn{
-    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \nn
+    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
 }
+In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
 
@@
 
+\extraspacing {\bf Computing the Top-Sliced Rectangles.}
+
 \section{Hardness of 3-SJ in 3D Space} \label{app:lb-cond}

### 2024-05-24 msj.tex
yufei

@@
 
-\section{Preliminaries in Geometry} \label{sec:bricks}
-
-This section will first define some notions frequently used in our presentation and then introduce several  geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
-
-\extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$.
-We say that a horizontal segment $h_1$ is {\em lower} (resp., {\em higher}) than another horizontal segment $h_2$ if the y-coordinate of $h_1$ is smaller (resp., larger) than that of $h_2$.
-Similarly, a vertical segment $v_1$ is {\em to the left} (resp., {\em right}) {\em of} another vertical segment $v_2$ if the x-coordinate of $v_1$ is smaller (resp., larger) than that of $v_2$.
-
-Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
-Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$.
-\yf{need this?} A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
-A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
-
-\vgap
-
-Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$.
-Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
-We use $B_\bm{t}$ to represent $\bigcap_{i=1}^t r_i$, namely, the intersection of the rectangles in $\bm{t}$.
-Note that $B_\bm{t}$ is a rectangle itself.
-We define:
-\myitems{
-    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$.
-In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
-
-    \item $\gbot(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$.
-In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
-}
-See Figure~\ref{fig:guard} for an illustration.
-
 \begin{figure*}
@@
 
-\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of points and set $R$ of rectangles, all in 2D space.
+\section{Preliminaries in Geometry} \label{sec:bricks}
+
+This section will first define some notions frequently used in our presentation and then introduce several  geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
+
+\extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$.
+We say that a horizontal segment $h_1$ is {\em lower} (resp., {\em higher}) than another horizontal segment $h_2$ if the y-coordinate of $h_1$ is smaller (resp., larger) than that of $h_2$.
+Similarly, a vertical segment $v_1$ is {\em to the left} (resp., {\em right}) {\em of} another vertical segment $v_2$ if the x-coordinate of $v_1$ is smaller (resp., larger) than that of $v_2$.
+
+\vgap
+
+Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
+Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$.
+\yf{need this?} A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
+
+\vgap
+
+Let $S$ be a set of segments all of which are either horizontal or vertical. Given a rectangle $r$, we define 
+\myeqn{
+    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}. 
+    \label{eqn:cross}
+}
+
+\vgap
+
+Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$.
+Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
+We use $B_\bm{t}$ to represent $\bigcap_{i=1}^t r_i$, namely, the intersection of the rectangles in $\bm{t}$.
+Note that $B_\bm{t}$ is a rectangle itself.
+We define:
+\myitems{
+    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$.
+In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
+
+    \item $\gbot(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$.
+In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
+}
+See Figure~\ref{fig:guard} for an illustration.
+
+\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of 2D points and set $R$ of rectangles.
 In the {\em detection version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$.
@@
 \extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$.
-If such an $h$ does not exist, output nothing for $r$.
+In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see \eqref{eqn:cross}).
+If no segment in $H$ crosses $r$, output nothing for $r$.
 Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
@@
 
-In the {\em find-all} version of the problem, the goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$.
-Furthermore, those segments need to be sorted by y-coordinate.
-Formally, if $h_1, h_2, ..., h_z$ (for some $z \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
+In the {\em find-all} version of the problem, the goal is to report, for each rectangle $r \in R$, the entire $\cross_H(r)$ sorted by y-coordinate.
+Formally, if $\cross_H(r) = \set{h_1, h_2, ..., h_z}$ (for some $z \ge 1$), we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
 In the example of Figure~\ref{fig:probs}d, the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.

### 2024-05-24 msj.tex
yufei

@@
 \myeqn{
-    \dcross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
-    \dcross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
+    \dcross_H(\bm{t}) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \label{eqn:dcross-H} \\
+    \dcross_V(\bm{t}) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \label{eqn:dcross-H} 
 }
@@
         \myenums{
-            \item $\dcross(\bm{t}, H) \ne \emptyset$ and $\dcross(\bm{t}, V) \ne \emptyset$.
+            \item $\dcross_H(\bm{t}) \ne \emptyset$ and $\dcross_V(\bm{t}) \ne \emptyset$.
 
             \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i]$. 
-            For any $h \in \dcross(\bm{t}, H)$ and any $v \in \dcross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+            For any $h \in \dcross_H(\bm{t})$ and any $v \in \dcross_V(\bm{t})$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
 
-            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \dcross(\bm{t}, H)$ and $v \in \dcross(\bm{t}, V)$.
+            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
         }
@@
 
-        \item $\sum_{\bm{t}} |\dcross(\bm{t}, H)| \le \out$ and $\sum_{\bm{t}} |\dcross(\bm{t}, V)| \le \out$.
+        \item $\sum_{\bm{t}} |\dcross_H(\bm{t})| \le \out$ and $\sum_{\bm{t}} |\dcross_V(\bm{t})| \le \out$.
     }
@@
     \> 3.\>\> $r_i \leftarrow$ the full rectangle of $\bm{t}[i]$, for each $i \in [k-2]$ \\
-    \> 4.\>\> {\bf for} each $(h, v) \in \dcross(\bm{t}, H) \times \dcross(\bm{t}, V)$ {\bf do} \\
+    \> 4.\>\> {\bf for} each $(h, v) \in \dcross_H(\bm{t}) \times \dcross_V(\bm{t})$ {\bf do} \\
     \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $\J^*$
@@
 Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
-Furthermore, if we are given $\dcross(\bm{t}, H)$ and $\dcross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
+Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
 
@@
 
-The rest of the section will focus on explaining how to prepare the sets $\dcross(\bm{t}, H)$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
-An analogous algorithm can be used to compute the sets $\dcross(\bm{t}, V)$ of all $\bm{t}$ in the same time complexity.
+The rest of the section will focus on explaining how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
+An analogous algorithm can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ in the same time complexity.
 
-\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r' \in R_i'$, we define
+\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r' \in R_i'$, define
 \myeqn{
-    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): r' = \gbot(\bm{t})} \xright(B_\bm{t}). 
+    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): r' = \gbot(\bm{t})} \ytop(B_\bm{t}). 
 }
@@
 }
-We refer to $r^*$ as the {\em top-sliced rectangle} of $r'$.
+and call it the {\em top-sliced rectangle} of $r'$.
+
+\vgap
 

### 2024-05-24 msj.tex
yufei

@@
     \dcross_H(\bm{t}) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \label{eqn:dcross-H} \\
-    \dcross_V(\bm{t}) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \label{eqn:dcross-H} 
+    \dcross_V(\bm{t}) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
+    %\label{eqn:dcross-H} 
 }
@@
 
-The rest of the section will focus on explaining how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
-An analogous algorithm can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ in the same time complexity.
+The rest of the section will focus on how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
+An analogous method can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ in the same time complexity.
 
@@
 
-Now construct from $R_i'$ a new set of rectangles: 
+Next, we construct from $R_i'$ a new set of rectangles: 
 \myeqn{
@@
 In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
+
+\vgap
+
+Now, we compute $\cross_H(r^*)$ for each $r^* \in R^*_i$
+
+\extraspacing {\bf Computing the ``d-cross'' Sets.}
 

### 2024-05-24 msj.tex
yufei

@@
 
-In the {\em find-all} version of the problem, the goal is to report, for each rectangle $r \in R$, the entire $\cross_H(r)$ sorted by y-coordinate.
+In the {\em find-all-sorted} version of the problem, the goal is to report, for each rectangle $r \in R$, the entire $\cross_H(r)$ sorted by y-coordinate.
 Formally, if $\cross_H(r) = \set{h_1, h_2, ..., h_z}$ (for some $z \ge 1$), we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
@@
 
-Now, we compute $\cross_H(r^*)$ for each $r^* \in R^*_i$
+Now, we compute $\cross_H(r^*)$ for each $r^* \in R^*_i$ --- where $\cross_H(r^*)$, defined in \eqref{eqn:cross}, is the set of segments in $H$ crossing $r^*$ --- which is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
 

### 2024-05-24 msj.tex
yufei

@@
 Now, we compute $\cross_H(r^*)$ for each $r^* \in R^*_i$ --- where $\cross_H(r^*)$, defined in \eqref{eqn:cross}, is the set of segments in $H$ crossing $r^*$ --- which is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
+The following lemma, proven in Appendix~\ref{app:hv}, presents some crucial properties for these ``cross'' sets.
+
+\begin{lemma} \label{lmm:hv:type1:cross-r*}
+    Both statements below are true: 
+    \myenums{
+        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
+
+        \item Consider any $k$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
+Then, it must hold that $\dcross_H(\bm{t}) \subseteq \cross_H(r^*)$.
+Furthermore, if the segments of $\cross_H(r^*)$ have been sorted in ascending order, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+    }
+\end{lemma}
+
+\yf{example}
 
@@
 
-\section{Finding Result Tuples of Type 2} \label{sec:hv:type2}
+\section{H-V $k$-SJ: Result Tuples of Type 2} \label{sec:hv:type2}
 

### 2024-05-24 msj.tex
yufei

@@
 
-Now, we compute $\cross_H(r^*)$ for each $r^* \in R^*_i$ --- where $\cross_H(r^*)$, defined in \eqref{eqn:cross}, is the set of segments in $H$ crossing $r^*$ --- which is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
-The following lemma, proven in Appendix~\ref{app:hv}, presents some crucial properties for these ``cross'' sets.
+We pay particular attention to the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$, defined in \eqref{eqn:cross}, is the set of segments in $H$ crossing $r^*$.
+The following lemma, proven in Appendix~\ref{app:hv}, presents some crucial properties for these sets.
 
@@
 \end{lemma}
+
+Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$. Statement (1) of the above lemma assures us that
 

### 2024-05-24 msj.tex
yufei

@@
 Then, it must hold that $\dcross_H(\bm{t}) \subseteq \cross_H(r^*)$.
-Furthermore, if the segments of $\cross_H(r^*)$ have been sorted in ascending order, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ have been sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
     }
@@
 
-Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$. Statement (1) of the above lemma assures us that
+Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
+Statement (1) of the above lemma assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is $O(n \log n + \out)$.
+Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.
 

### 2024-05-24 msj.tex
yufei

@@
 
+\yf{example}
+
 Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
-Statement (1) of the above lemma assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is $O(n \log n + \out)$.
+Statement (1) of the above lemma assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
 Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.
 
-\yf{example}
-
-\extraspacing {\bf Computing the ``d-cross'' Sets.}
+\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready explain how to compute $\dcross_H(\bm{t})$ for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+By Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*},
 

### 2024-05-24 msj.tex
yufei

@@
 
-\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready explain how to compute $\dcross_H(\bm{t})$ for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
-By Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*},
+\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready explain how to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+This, in fact, has been made easy by Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
+First, identify the rectangle $r' = \gbot(\bm{t})$ and fetch its (already computed) top-sliced rectangle $r^*$.
+Second, compute $B_\bm{t}$ and $\gleft(\bm{t})$.
+These two steps require $O(k)$ time.
+Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
+For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely,  whether $h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$; this can be done in constant time.
+Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
+This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|)$ time.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k |\J| + |\J^*|) = O(k \cdot \out)$ time.
 

### 2024-05-24 msj.tex
yufei

@@
 By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
-It thus follows from Section~\ref{sec:bricks} that $R_i'$ can be produced in $O(n \log n)$ time.
+It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(n \log n)$ total time.
 
@@
 We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
-This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R_1', R_2, ..., R_{k-2})$.
+This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$.
 We prove in Appendix~\ref{app:hv}:
@@
 
-\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready explain how to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
-This, in fact, has been made easy by Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
+\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
 First, identify the rectangle $r' = \gbot(\bm{t})$ and fetch its (already computed) top-sliced rectangle $r^*$.
-Second, compute $B_\bm{t}$ and $\gleft(\bm{t})$.
+Second, obtain $B_\bm{t}$ and $\gleft(\bm{t})$.
 These two steps require $O(k)$ time.
@@
 This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|)$ time.
-Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k |\J| + |\J^*|) = O(k \cdot \out)$ time.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + |\J^*|) = O(k \cdot \out)$ time.
+
+\vgap
+
+We now conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
 

### 2024-05-24 msj.tex
yufei

@@
 Note that $r'$ exists if and only if $r$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
-Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$.
+
+%Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$.
 
@@
 }
-Computing the ``segment $h_1$'' for each $r \in R_1$ is an instance of Problem $\mathscr{D}$ (the find-lowest version).
-By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$.
+Computing the ``segment $h_1$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version).
+By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r \in R_i$.
 It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(n \log n)$ total time.

### 2024-05-24 msj.tex
yufei

@@
 
-\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$.
+\extraspacing {\bf Generating $\bm{\J^*}$.} Take any $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$.
 The reader should recall from Section~\ref{sec:bricks} that
 \myitems{
-    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i]$, i.e., the intersection of $r'_1, r'_2, ..., r'_{k-2}$;
+    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i] = \bigcap_{i=1}^{k-2} r'_i$;
     \item $\gleft(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
@@
 }
-The prefix ``d-'' stands for double. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:
+The prefix ``d-'' stands for ``double''. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:
 
@@
 
-Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
-Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
+By statements (b) and (c) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
+Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, statement (a) of Lemma~\ref{lmm:hv:type1:properties} assures us that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
 

### 2024-05-24 msj.tex
yufei

@@
 
-\begin{lemma} \label{lmm:hv:type1:proj}
+\begin{lemma} \label{lmm:hv:type1:recur-output}
     $|\J(R_1', R'_2, ..., R'_{k-2})| \le \out$.
@@
 
-We remind the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
+We remind the reader that $\out$ is the output size of the original H-V $k$-SJ problem.
+As a corollary of Lemma~\ref{lmm:hv:type1:recur-output}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
 
@@
 \myeqn{
-    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): r' = \gbot(\bm{t})} \ytop(B_\bm{t}). 
+    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
 }
-The value $\maxtop(r')$ is set to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
+We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
 When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
@@
 
-We pay particular attention to the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$, defined in \eqref{eqn:cross}, is the set of segments in $H$ crossing $r^*$.
-The following lemma, proven in Appendix~\ref{app:hv}, presents some crucial properties for these sets.
+We are particularly concerned with the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
+The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties for these sets.
 
@@
 Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
-Then, it must hold that $\dcross_H(\bm{t}) \subseteq \cross_H(r^*)$.
-Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ have been sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
+Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
     }
@@
 Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
-Statement (1) of the above lemma assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
+Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*} assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
 Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.
@@
 This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|)$ time.
-Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + |\J^*|) = O(k \cdot \out)$ time.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)| = O(k \cdot \out)$ time, where the last step used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.
 
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:proj}.}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}
 
@@
 
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
+
 \section{Hardness of 3-SJ in 3D Space} \label{app:lb-cond}

### 2024-05-24 msj.tex
yufei

@@
 \times R_{k-2} \times H \times V$ such that $h \cap v \cap \bigcap_{i=1}^{k-2} r_i$ is not empty.
-As we will see, conquering this sepcial version is the most challenging step in settling the original $k$-SJ problem optimally.
+
+%As we will see, conquering this sepcial version is the most challenging step in settling the original $k$-SJ problem optimally.
 
@@
 \end{lemma}
+
+The part of the paper from now until the appendix will be devoted to proving the above lemma. This is the most challenging step in settling the general $k$-SJ problem optimally.
+
+\vgap
 

### 2024-05-24 msj.tex
yufei

@@
 
-The part of the paper from now until the appendix will be devoted to proving the above lemma. This is the most challenging step in settling the general $k$-SJ problem optimally.
+The part of the paper from this point onwards until the appendix will be devoted to proving the above lemma.
+This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Appendix~\ref{app:kSJ}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
 
@@
 
+\section{Solving the $k$-SJ Problem} \label{app:ksj}
+
 \section{Hardness of 3-SJ in 3D Space} \label{app:lb-cond}

### 2024-05-24 msj.tex
yufei

@@
 The part of the paper from this point onwards until the appendix will be devoted to proving the above lemma.
-This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Appendix~\ref{app:kSJ}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
+This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Appendix~\ref{app:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
 

### 2024-05-24 msj.tex
yufei

@@
 Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
-This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|)$ time.
-Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)| = O(k \cdot \out)$ time, where the last step used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.
+This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|))$ time.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.
 
@@
 
-We now conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+We 
+conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
 

### 2024-05-24 msj.tex
yufei

@@
 
-Let $S$ be a set of segments all of which are either horizontal or vertical. Given a rectangle $r$, we define 
+Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
 \myeqn{
@@
     \dcross_H(\bm{t}) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \label{eqn:dcross-H} \\
-    \dcross_V(\bm{t}) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
+    \dcross_V(\bm{t}) = \set{v \in V \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
     %\label{eqn:dcross-H} 
@@
 Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
-Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*} assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
+Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
 Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.

### 2024-05-24 msj.tex
yufei

@@
 \myeqn{
-    \dcross_H(\bm{t}) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \label{eqn:dcross-H} \\
-    \dcross_V(\bm{t}) = \set{v \in V \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
+    \dcross_H(\bm{t}) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}} \label{eqn:dcross-H} \\
+    \dcross_V(\bm{t}) = \set{v \in V \mid \text{$v$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}}. \nn
     %\label{eqn:dcross-H} 
@@
 \extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
-First, identify the rectangle $r' = \gbot(\bm{t})$ and fetch its (already computed) top-sliced rectangle $r^*$.
-Second, obtain $B_\bm{t}$ and $\gleft(\bm{t})$.
-These two steps require $O(k)$ time.
+First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time.
 Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
-For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely,  whether $h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$; this can be done in constant time.
+For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$; this can be done in constant time.
 Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.

### 2024-05-24 msj.tex
yufei

@@
 Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
-For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$; this can be done in constant time.
+For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time.
 Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.

### 2024-05-25 msj.tex
yufei

@@
 
-%We will find result tuples of each type separately.
-Before proceeding, however, let us observe a sense of symmetry for Type 2.
-As mentioned, for each result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
-As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc.
-It is possible for the result tuple to belong to multiple sub-types simultaneously.
-
-\vgap 
-%
-% Our subsequent discussion will proceed as follows.
-Section~\ref{sec:hv:type1} will explain how to produce all the result tuples of Type 1.
-Section~\ref{sec:hv:type2} will explain how to produce the result tuples $(r_1, ..., r_{k-2}, h, v)$ of a particular sub-type of Type 2: those tuples where $r_1$ covers the left endpoint of $h$.
-Other sub-types can be computed using the same algorithm, by symmetry.
-%
-% \vgap
-%
-% A remark is in order about duplicate removal.
-A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types.
-Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm.
-However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times.
-Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to.
-To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting.
-One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.
-
 \begin{figure}
@@
 
-        \item Consider any $k$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
 Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
@@
 Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
-This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|))$ time.
+This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
 Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.
@@
 \section{H-V $k$-SJ: Result Tuples of Type 2} \label{sec:hv:type2}
+
+Still, denote by $R_1, ..., R_{k-2}, H$, and $V$ the input sets of the H-V $k$-SJ problem. This section will explain how to find the result tuples of Type 2 as defined in Section~\ref{sec:hv}.
+
+\vgap
+
+As mentioned before, for a $(k-2)$-tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
+As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc.
+It is possible for the result tuple to belong to multiple sub-types simultaneously.
+
+\vgap
+
+Next, we will focus on producing the result tuples of a particular sub-type: those tuples $(r_1, ..., r_{k-2}, h, v)$ where $r_1$ covers the left endpoint of $h$ --- let us denote this set of result tuples as $\J_2$.
+Other sub-types can be computed using the same algorithm, by symmetry.
+
+\vgap
+
+A remark is in order about duplicate removal.
+By finding each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm.
+However, this does not mean that the tuple needs to be {\em reported} multiple times.
+Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to.
+To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting.
+One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.
+
+\extraspacing {\bf Set $\bm{H'}$.}
 

### 2024-05-25 msj.tex
yufei

@@
 
-We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+We now solve a $(k-2)$-SJ problem with $R_1', R_2', ..., R_{k-2}'$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
 This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$.
 We prove in Appendix~\ref{app:hv}:
-
-%The lemma below shows that the result is essentially what we aim to compute in Step 1.
 
@@
 
-% \begin{figure}
-%     \includegraphics[height=30mm]{./artwork/four-segs}
-%     \figcapup
-%     \caption{Trimming}
-%     \label{fig:hv:type1:trim}
-%     \figcapdown
-% \end{figure}
-
 \extraspacing {\bf Generating $\bm{\J^*}$.} Take any $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$.
@@
 We now introduce:
-% \myitems{
-%     \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$;
-%     \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
-% }
 \myeqn{
@@
 
-\extraspacing {\bf Set $\bm{H'}$.}
+\extraspacing {\bf Set $\bm{H'}$.} For each horizontal segment $h = [x_1, x_2] \times y \in H$, we want to find the rightmost point $p$ on $h$, such that $p$ is covered by at least one rectangle $r_{k-2} \in R_{k-2}$.
+This point $p$ exists if and only if $h$ has at least one left-end covering rectangle in $R_{k-2}$.
+If $p$ exists and has coordinates $(x, y)$, we construct a segment $h' = [x_1, x] \times y$,  referred to as the {\em trimmed segment} of $h$; conversely, we call $h$ the {\em full segment} of $h'$.
+
+\vgap
+
+Construct
+\myeqn{
+    H' &=& \{h' \mid h \in H \text{ and its trimmed segment $h'$ exists} \}. \nn
+}
+The construction requires solving an instance of Problem $\mathscr{C}$, which takes $O(n \log n)$ time based on the discussion in Section~\ref{sec:bricks}.
+
+\vgap
+
+We now solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-1}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, R_2, ..., R_{k-1}, H', V)$.
+We prove in Appendix~\ref{app:hv}:
+
+\begin{lemma} \label{lmm:hv:type2:recur-output}
+    $|\J(R_1, ..., R_{k-1}, H', V)| \le \out$.
+\end{lemma}
 
@@
 
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
+
 \section{Solving the $k$-SJ Problem} \label{app:ksj}

### 2024-05-25 msj.tex
yufei

@@
 \def\dcross{\mathrm{d\text{-}cross}}
+\def\contained{\mathrm{contained}}
 \def\out{\mathrm{OUT}}
@@
 
+%Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. \yf{need this?}
+
 Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
-Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$.
-\yf{need this?} A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
 A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
@@
 }
-
-\vgap
+Let $R$ be a set of rectangles. Given a horizontal segment, we define 
+\myeqn{
+    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}. 
+    \label{eqn:contained}
+}
 
@@
 \extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, all the rectangles $r \in R$ containing $h$.
-Furthermore, those rectangles need to be sorted by the x-coordinates of their right boundaries.
-Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
+The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ (defined in \eqref{eqn:contained}), where the rectangles are sorted by their right boundaries.
+Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
 Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
@@
 
-We 
-conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+We conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.
 
@@
 We now solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-1}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
-This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, R_2, ..., R_{k-1}, H', V)$.
+This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, ..., R_{k-1}, H', V)$.
 We prove in Appendix~\ref{app:hv}:
@@
 \end{lemma}
+
+The $(k-1)$-SJ can be settled in $F_{k-1}(n, \out)$ time.
+
+\yf{example}
+
+\extraspacing {\bf Generating $\bm{\J^*}$.}
 

### 2024-05-25 msj.tex
yufei

@@
 
-We now solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-1}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
-This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, ..., R_{k-1}, H', V)$.
+We now solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-3}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, ..., R_{k-3}, H', V)$.
 We prove in Appendix~\ref{app:hv}:
@@
 \begin{lemma} \label{lmm:hv:type2:recur-output}
-    $|\J(R_1, ..., R_{k-1}, H', V)| \le \out$.
+    $|\J(R_1, ..., R_{k-3}, H', V)| \le \out$.
 \end{lemma}
@@
 
-\extraspacing {\bf Generating $\bm{\J^*}$.}
+\extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h, v)$.
 

### 2024-05-25 msj.tex
yufei

@@
 
-\extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h, v)$.
+\extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$.
+Note that $\bm{t}[k-1] = h'$ must intersect with $v$ (because $\bm{t}$ is a result tuple of the $(k-1)$-SJ).
+If $p$ is the intersection point between $h$ and $v$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment from the left endpoint of $h'$ to $p$.
+\myeqn{
+    \contained_{R_{k-2}}(\bm{t}) &=& \{r_{k-2} \in R_{k-2} \mid \text{ $r_{k-2}$ contains the effective} \nn \\
+    && \text{horizontal segment of $\bm{t}$}   \}
+    \nn
+}
+The above definition should not be confused with \eqref{eqn:contained}, where the $\contained$ function takes a horizontal segment as the input.
+
+\begin{lemma} \label{lmm:hv:type2:properties}
+    For each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, we have:
+        \myenums{
+            \item $\contained_{R_{k-2}}(\bm{t}) \ne \emptyset$.
+
+            \item Denote by $h$ the full segment of $\bm{t}[k-1]$. 
+            For any $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r_{k-2}, h, \bm{t}[k])$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+
+            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+        }
+\end{lemma}
 

### 2024-05-25 msj.tex
yufei

@@
             For any $h \in \dcross_H(\bm{t})$ and any $v \in \dcross_V(\bm{t})$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
-
-            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
         }
@@
         \vspace{2mm}
+
+        \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $\bm{t} = (r_1, r_2, ...,$ $r_{k-2})$ belongs to $\J(R_1', R'_2, ..., R'_{k-2})$, and       
+        $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
 
@@
 
-By statements (b) and (c) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
+By statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
 Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, statement (a) of Lemma~\ref{lmm:hv:type1:properties} assures us that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.

### 2024-05-25 msj.tex
yufei

@@
 
-        \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $\bm{t} = (r_1, r_2, ...,$ $r_{k-2})$ belongs to $\J(R_1', R'_2, ..., R'_{k-2})$, and       
-        $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+        \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2})$ belongs to $\J(R_1', R'_2, ..., R'_{k-2})$, and       
+        $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, where $r'_i$ ($i \in [k-2])$ is the trimmed rectangle of $r_i$.
 
@@
 \begin{lemma} \label{lmm:hv:type2:properties}
-    For each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, we have:
+    All the following statements are true:
+    \myenums{
+        \item For each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, we have:
+
+        \vspace{2mm}
+
         \myenums{
@@
             \item Denote by $h$ the full segment of $\bm{t}[k-1]$. 
-            For any $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r_{k-2}, h, \bm{t}[k])$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
+            For any $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r_{k-2}, h, \bm{t}[k])$ belongs to $\J_2$.
+        }
 
-            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+        \vspace{2mm}
+
+        \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_2$, then $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1, ..., R_{k-3}, H', V)$ and        
+        $h' \in \contained_{R_{k-2}}(\bm{t})$, where $h'$ is the trimmed segment of $h$.
+
+        \item d
     }

### 2024-05-25 msj.tex
yufei

@@
 \extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$.
-Note that $\bm{t}[k-1] = h'$ must intersect with $v$ (because $\bm{t}$ is a result tuple of the $(k-1)$-SJ).
+Note that $\bm{t}[k-2] = h'$ must intersect with $v$ (because $\bm{t}$ is a result tuple of the $(k-1)$-SJ).
 If $p$ is the intersection point between $h$ and $v$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment from the left endpoint of $h'$ to $p$.
@@
 The above definition should not be confused with \eqref{eqn:contained}, where the $\contained$ function takes a horizontal segment as the input.
+
+\vgap
+
+We prove the next lemma in Appendix~\ref{app:hv}.
 
@@
 
-            \item Denote by $h$ the full segment of $\bm{t}[k-1]$. 
+            \item Denote by $h$ the full segment of $\bm{t}[k-2]$. 
             For any $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r_{k-2}, h, \bm{t}[k])$ belongs to $\J_2$.
@@
 
-        \item d
+        \item $\sum_{\bm{t}} |\contained_{R_{k-2}}(\bm{t})| \le \out$.
     }
 \end{lemma}
+
+Equipped with Lemma~\ref{lmm:hv:type1:properties}, we generate our target $\J_2$ as follows:
+
+\mytab{
+    \> {\bf algorithm generate-$\J_2$} \\
+    \> 1.\> $\J_2 = \emptyset$ \\
+    \> 2.\> {\bf for} each $(k-2)$-tuple $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ {\bf do} \\
+    \> 3.\>\> $h \leftarrow$ the full segment of $\bm{t}[k-2]$ \\
+    \> 4.\>\> {\bf for} each $(h, v) \in \dcross_H(\bm{t}) \times \dcross_V(\bm{t})$ {\bf do} \\
+    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $\J^*$
+}
+
+By statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
+Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, statement (a) of Lemma~\ref{lmm:hv:type1:properties} assures us that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
 

### 2024-05-25 msj.tex
yufei

@@
 \extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$.
-Note that $\bm{t}[k-2] = h'$ must intersect with $v$ (because $\bm{t}$ is a result tuple of the $(k-1)$-SJ).
+Note that $\bm{t}[k-2] = h'$ must intersect with $v$ as $\bm{t}$ is a result tuple of the $(k-1)$-SJ.
 If $p$ is the intersection point between $h$ and $v$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment from the left endpoint of $h'$ to $p$.
@@
             \item Denote by $h$ the full segment of $\bm{t}[k-2]$. 
-            For any $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r_{k-2}, h, \bm{t}[k])$ belongs to $\J_2$.
+            For any $r \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k])$ belongs to $\J_2$.
         }
@@
     \> 3.\>\> $h \leftarrow$ the full segment of $\bm{t}[k-2]$ \\
-    \> 4.\>\> {\bf for} each $(h, v) \in \dcross_H(\bm{t}) \times \dcross_V(\bm{t})$ {\bf do} \\
-    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $\J^*$
+    \> 4.\>\> {\bf for} each $r \in \contained_{R_{k-2}}(\bm{t})$ {\bf do} \\
+    \> 5.\>\>\> add $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k])$ to $\J^*$
 }
 
-By statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
+%By statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J_2$ thus produced is indeed what we aim to compute.
 Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, statement (a) of Lemma~\ref{lmm:hv:type1:properties} assures us that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.

### 2024-05-25 msj.tex
yufei

@@
 
-        \item If a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2})$ belongs to $\J(R_1', R'_2, ..., R'_{k-2})$, and       
-        $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, where $r'_i$ ($i \in [k-2])$ is the trimmed rectangle of $r_i$.
+        \item Consider any a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v) \in \J_1$.
+Let $r'_i$ ($i \in [k-2])$ be the trimmed rectangle of $r_i$, and set $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2})$.
+Then, $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, $h \in \dcross_H(\bm{t})$, and $v \in \dcross_V(\bm{t})$.
 
@@
 
-As mentioned before, for a $(k-2)$-tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
+As mentioned before, for a result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
 As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc.
@@
 
-Next, we will focus on producing the result tuples of a particular sub-type: those tuples $(r_1, ..., r_{k-2}, h, v)$ where $r_1$ covers the left endpoint of $h$ --- let us denote this set of result tuples as $\J_2$.
+Next, we will focus on producing the result tuples of a particular sub-type: those tuples $(r_1, ..., r_{k-2}, h, v)$ where $r_{k-2}$ covers the left endpoint of $h$ --- let us denote this set of result tuples as $\J_2$.
 Other sub-types can be computed using the same algorithm, by symmetry.
@@
             \item Denote by $h$ the full segment of $\bm{t}[k-2]$. 
-            For any $r \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k])$ belongs to $\J_2$.
+            For any $r \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ belongs to $\J_2$.
         }
@@
 
-        \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_2$, then $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1, ..., R_{k-3}, H', V)$ and        
-        $h' \in \contained_{R_{k-2}}(\bm{t})$, where $h'$ is the trimmed segment of $h$.
+        \item Consider any $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v) \in \J_2$.
+Let $h'$ be the trimmed segment of $h$, and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
+Then, $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ and $h' \in \contained_{R_{k-2}}(\bm{t})$.
 
@@
     \> 4.\>\> {\bf for} each $r \in \contained_{R_{k-2}}(\bm{t})$ {\bf do} \\
-    \> 5.\>\>\> add $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k])$ to $\J^*$
+    \> 5.\>\>\> add $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ to $\J_2?
+    $
 }

### 2024-05-25 msj.tex
yufei

@@
     \> 4.\>\> {\bf for} each $r \in \contained_{R_{k-2}}(\bm{t})$ {\bf do} \\
-    \> 5.\>\>\> add $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ to $\J_2?
-    $
+    \> 5.\>\>\> add $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ to $\J_2$
 }
 
-%By statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J_2$ thus produced is indeed what we aim to compute.
-Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, statement (a) of Lemma~\ref{lmm:hv:type1:properties} assures us that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
+The correctness of the algorithm follows from statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}.
+Furthermore, if we are given $\contained_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time.
 

### 2024-05-25 msj.tex
yufei

@@
 \myeqn{
-    \contained_{R_{k-2}}(\bm{t}) &=& \{r_{k-2} \in R_{k-2} \mid \text{ $r_{k-2}$ contains the effective} \nn \\
-    && \text{horizontal segment of $\bm{t}$}   \}
+    \contained_{R_{k-2}}(\bm{t}) &=& \{r_{k-2} \in R_{k-2} \mid \text{ the effective horizontal } \nn \\
+    && \text{segment of $\bm{t}$ is contained in $r_{k-2}$}   \}
     \nn

### 2024-05-25 msj.tex
yufei

@@
     \myenums{
-        \item For each $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, we have:
-
-        \vspace{2mm}
-
-        \myenums{
-            \item $\dcross_H(\bm{t}) \ne \emptyset$ and $\dcross_V(\bm{t}) \ne \emptyset$.
-
-            \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i]$. 
-            For any $h \in \dcross_H(\bm{t})$ and any $v \in \dcross_V(\bm{t})$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.
-        }
-
-        \vspace{2mm}
+        \item Consider any $(k-2)$-tuple $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$.
+Let $r_i$ ($i$ $\in$  $[k-2]$) be the full rectangle of $\bm{t}[i]$.
+Then, for any $h \in \dcross_H(\bm{t})$ and any $v \in \dcross_V(\bm{t})$, the $k$-tuple $(r_1, r_2,$ $...,$ $r_{k-2}, h, v)$ must belong to $\J(R_1, ..., R_{k-2}, H, V)$.
 
@@
 Let $r'_i$ ($i \in [k-2])$ be the trimmed rectangle of $r_i$, and set $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2})$.
-Then, $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, $h \in \dcross_H(\bm{t})$, and $v \in \dcross_V(\bm{t})$.
+Then, we must have $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, $h \in \dcross_H(\bm{t})$, and $v \in \dcross_V(\bm{t})$.
 
@@
 
-By statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
-Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, statement (a) of Lemma~\ref{lmm:hv:type1:properties} assures us that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.
+By statements (1) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
+Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, the above algorithm runs in $O(1 + |\J(R'_1, ..., R'_{k-2})| + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time, where the derivation used \eqref{eqn:hv:type1:J*} and Lemma~\ref{lmm:hv:type1:recur-output}.
 
@@
 This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
-Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type1:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
 

### 2024-05-25 msj.tex
yufei

@@
 Let $r'_i$ ($i \in [k-2])$ be the trimmed rectangle of $r_i$, and set $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2})$.
-Then, we must have $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, $h \in \dcross_H(\bm{t})$, and $v \in \dcross_V(\bm{t})$.
+Then, we must have
+        \myitems{
+            \item $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$;
+            \item $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$. 
+        }
+
+        \vspace{2mm}
 
@@
     \myenums{
-        \item For each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, we have:
-
-        \vspace{2mm}
-
-        \myenums{
-            \item $\contained_{R_{k-2}}(\bm{t}) \ne \emptyset$.
-
-            \item Denote by $h$ the full segment of $\bm{t}[k-2]$. 
-            For any $r \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ belongs to $\J_2$.
-        }
-
-        \vspace{2mm}
+        \item Consider any $(k-1)$-tuple $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$.
+Denote by $h$ the full segment of $\bm{t}[k-2]$.
+Then, for any $r \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ must belong to $\J_2$.
 

### 2024-05-25 msj.tex
yufei

@@
 
-The correctness of the algorithm follows from statements (b) and (2) of Lemma~\ref{lmm:hv:type1:properties}.
-Furthermore, if we are given $\contained_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time.
+The correctness of the algorithm follows from statements (1) and (2) of Lemma~\ref{lmm:hv:type2:properties}.
+Furthermore, if we are given $\contained_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 

### 2024-05-25 msj.tex
yufei

@@
 \def\maxtop{\mathrm{maxtop}}
+\def\minleft{\mathrm{minleft}}
 \def\gleft{\mathrm{left\text{-}guard}}
@@
 \extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ (defined in \eqref{eqn:contained}), where the rectangles are sorted by their right boundaries.
+The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries.
 Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
@@
 
-We are particularly concerned with the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
-The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties for these sets.
+Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
+The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.
 
@@
 
+\extraspacing {\bf Set $\bm{H^*}$.} Fix any $i \in [k-2]$. For each segment $h' \in H'$, define
+\myeqn{
+    \minleft(h') &=& \min_{v \in V: h' \cap v \ne \emptyset} \text{$x$-coordinate of $v$}. 
+}
+We set $\minleft(h')$ to $\infty$, if $h'$ intersects no segment in $V$. When $\minleft(h') \ne \infty$, assuming $h' = [x_1, x_2] \times y$, we introduce a horizontal segment
+\myeqn{
+    h^* &=& [x_1, \minleft(h')] \times y. \nn
+}
+and call it the {\em minimal segment} of $r'$.
+
+\vgap
+
+\yf{example}
+
+\vgap
+
+Next, we construct from $H'$ a new set of rectangles: 
+\myeqn{
+    H^* = \set{h^* \mid h' \in H' \text{ and its minimal segment $h^*$ exists}}. \label{eqn:hv:type2:H*}
+}
+Computing $H^*$ from $H'$ and $V$ is an instance of Problem $\mathscr{B}$, and can be achieved in $O(n \log n)$ as per the discussion from Section~\ref{sec:bricks}.
+
+\vgap
+
+We will utilize the sets $\contained_{R_{k-2}}(h^*)$ of the segments $h^*$ in $H^*$, where $\contained_{R_{k-2}}(h^*)$ --- defined in \eqref{eqn:contained} --- is the set of rectangles in $R_{k-2}$ containing $h^*$.
+The next lemma is proven in Appendix~\ref{app:hv}.
+
+\begin{lemma} \label{lmm:hv:type2:contained-h*}
+    Both statements below are true: 
+    \myenums{
+        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
+
+        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
+Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
+Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+    }
+\end{lemma}
+
+\yf{example}
+
+Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
+Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
+Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.
+
+\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
+First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time.
+Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
+For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time.
+Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
+This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type1:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
+
+\vgap
+
+We conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+
 \bibliographystyle{plainurl}% the mandatory bibstyle
@@
 
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}
+
 \section{Solving the $k$-SJ Problem} \label{app:ksj}

### 2024-05-25 msj.tex
yufei

@@
     \myenums{
-        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
+        \item $\sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)| \le \out$.
 
-        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
-Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
-Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
-Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+        \item Consider any tuple $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$.
+Set $h' = \bm{t}[k-2]$ and let $h^*$ be the minimal segment of $h'$.
+Then, we have $$\contained_{R_{k-2}}(\bm{t}) \subseteq \contained_{R_{k-2}}(h^*).$$ Furthermore, if the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
     }

### 2024-05-25 msj.tex
yufei

--- 2024-05-25 msj.tex
@@
     \myenums{
-        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
+        \item $\sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)| \le \out$.
 
-        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
-Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
-Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
-Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+        \item Consider any tuple $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$.
+Set $h' = \bm{t}[k-2]$ and let $h^*$ be the minimal segment of $h'$.
+Then, $$\contained_{R_{k-2}}(\bm{t}) \subseteq \contained_{R_{k-2}}(h^*).$$ Furthermore, if the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
     }

### 2024-05-25 msj.tex
yufei

@@
 
-Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
-Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
-Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.
+Finding the $\contained_{R_{k-2}}(h^*)$ sets of all $h^* \in H^*$ is an instance of the find-all-sorted version of Problem $\mathscr{E}$ (with $H^*$ and $R_{k-2}$ as the input).
+The cost is $O(n \log n + \out)$ according statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*} and the discussion in Section~\ref{sec:bricks}.
+Note that, for each $\contained_{R_{k-2}}(h^*)$ computed, the rectangles $r$ therein have been sorted in descending order of $\xright(r)$.
 

### 2024-05-25 msj.tex
yufei

@@
     && \text{segment of $\bm{t}$ is contained in $r_{k-2}$}   \}
-    \nn
+    \label{eqn:type2:contained-t}
 }
@@
 
-\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
-First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time.
-Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
-For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time.
-Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
-This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
-Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type1:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
+\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} We are ready to produce $\contained_{R_{k-2}}(h^*)$, defined in \eqref{eqn:type2:contained-t}, for any $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, based on Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*}.
+First, fetch the (already computed) minimal segment $h^*$ of $h'$ in $O(1)$ time.
+Then, scan the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$.
+For each $r$ scanned, check whether $r \in
+\contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers the intersection point between $h$ and $v$; this takes $O(1)$ time. Abort the scan as soon as $r \notin 
+\contained_{R_{k-2}}(\bm{t})$.
+This way, we compute $\contained_{R_{k-2}}(t)$ in $O(k + |\contained_{R_{k-2}}(t)|)$ time.
+Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 

### 2024-05-25 msj.tex
yufei

@@
 
-We conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+We conclude that $\J_2$ (the set of type-2 result tuples $(r_1, ..., r_{k-2}, h, v)$ where $r_{k-2}$ covers the left endpoint of $h$) can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+Remember that, to generate the entire type-2 result, we need to repeat the algorithm $4(k-2)$ times (one for each sub-type).
+The total running time is therefore $O(k) \cdot (F_{k-1}(n, \out) + O(n \log n +  k \cdot \out))$, as claimed in Lemma~\ref{lmm:hv}.
 

### 2024-05-25 msj.tex
yufei

--- 2024-05-25 msj.tex
@@
 
-The part of the paper from this point onwards until the appendix will be devoted to proving the above lemma.
-This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Appendix~\ref{app:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
+The part of the paper from this point till the end of Section~\ref{sec:hv:type2} will be devoted to proving the above lemma.
+This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Appendix~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
 
@@
 
-We conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.
 
@@
 
-We conclude that $\J_1$ (the set of type-1 result tuples) can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+We conclude that $\J_2$ --- the set of type-2 result tuples $(r_1, ..., r_{k-2},$ $h, v)$ where $r_{k-2}$ covers the left endpoint of $h$ --- can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+Remember that, to generate the entire type-2 result, we need to repeat the algorithm $4(k-2)$ times (one for each sub-type).
+The total running time is therefore $O(k) \cdot (F_{k-1}(n, \out) + n \log n +  k \cdot \out)$, as claimed in Lemma~\ref{lmm:hv}.
+
+\section{Settling $k$-SJ} \label{sec:ksj}
 
@@
 
-\section{Solving the $k$-SJ Problem} \label{app:ksj}
-
 \section{Hardness of 3-SJ in 3D Space} \label{app:lb-cond}

### 2024-05-25 msj.tex
yufei

@@
 }
-The construction requires solving an instance of Problem $\mathscr{C}$, which takes $O(n \log n)$ time based on the discussion in Section~\ref{sec:bricks}.
+The construction requires solving an instance of Problem $\mathscr{C}$ (with $H$ and $R_{k-2}$ as the input), which takes $O(n \log n)$ time based on the discussion in Section~\ref{sec:bricks}.
 

### 2024-05-25 msj.tex
yufei

@@
 }
-Computing the ``segment $h_1$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version).
+Computing the ``segment $h_1$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input).
 By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r \in R_i$.
@@
 
-Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$.
+Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$ (with $H$ and $R^*_i$ as the input).
 Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.

### 2024-05-25 msj.tex
yufei

@@
 The part of the paper from this point till the end of Section~\ref{sec:hv:type2} will be devoted to proving the above lemma.
-This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Appendix~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
+This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Section~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
 

### 2024-05-25 msj.tex
yufei

@@
 As mentioned before, for a result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both.
-As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc.
+As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc.
 It is possible for the result tuple to belong to multiple sub-types simultaneously.
@@
 By finding each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm.
-However, this does not mean that the tuple needs to be {\em reported} multiple times.
+However, this does not mean that the tuple needs to be reported multiple times.
 Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to.
@@
 
-\extraspacing {\bf Set $\bm{H'}$.} For each horizontal segment $h = [x_1, x_2] \times y \in H$, we want to find the rightmost point $p$ on $h$, such that $p$ is covered by at least one rectangle $r_{k-2} \in R_{k-2}$.
+\extraspacing {\bf Set $\bm{H'}$.} Take any horizontal segment $h = [x_1, x_2] \times y \in H$.
+Let $p$ be the rightmost point $p$ on $h$ satisfying the condition that at least one rectangle in $R_{k-2}$ covers $p$.
 This point $p$ exists if and only if $h$ has at least one left-end covering rectangle in $R_{k-2}$.
-If $p$ exists and has coordinates $(x, y)$, we construct a segment $h' = [x_1, x] \times y$,  referred to as the {\em trimmed segment} of $h$; conversely, we call $h$ the {\em full segment} of $h'$.
+If $p$ exists and has coordinates $(x, y)$, we refer to the segment $h' = [x_1, x] \times y$ as the {\em trimmed segment} of $h$; conversely, we call $h$ the {\em full segment} of $h'$.
 
@@
 }
-The construction requires solving an instance of Problem $\mathscr{C}$ (with $H$ and $R_{k-2}$ as the input), which takes $O(n \log n)$ time based on the discussion in Section~\ref{sec:bricks}.
+The construction is an instance of Problem $\mathscr{C}$ (with $H$ and $R_{k-2}$ as the input) and finishes in $O(n \log n)$ time based on the discussion in Section~\ref{sec:bricks}.
 

### 2024-05-25 msj.tex
yufei

@@
 
-We now solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-3}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
-This $(k-1)$-SJ has an input size at most $n$, and let us represent its result as $\J(R_1, ..., R_{k-3}, H', V)$.
-We prove in Appendix~\ref{app:hv}:
+Now, solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-3}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+Let $\J(R_1, ...,$ $R_{k-3}, H', V)$ represent the result of this $(k-1)$-SJ.
+As the $(k-1)$-SJ has an input size at most $n$, given the lemma below (which is proved in Appendix~\ref{app:hv}), we assert that $\J(R_1, ..., R_{k-3}, H', V)$ can be computed in $F_{k-1}(n, \out)$ time.
 
@@
 \end{lemma}
-
-The $(k-1)$-SJ can be settled in $F_{k-1}(n, \out)$ time.
 

### 2024-05-25 msj.tex
yufei

@@
 Now, solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-3}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
-Let $\J(R_1, ...,$ $R_{k-3}, H', V)$ represent the result of this $(k-1)$-SJ.
-As the $(k-1)$-SJ has an input size at most $n$, given the lemma below (which is proved in Appendix~\ref{app:hv}), we assert that $\J(R_1, ..., R_{k-3}, H', V)$ can be computed in $F_{k-1}(n, \out)$ time.
+Let $\J(R_1, ...,$ $R_{k-3}, H', V)$ represent the result of this $(k-1)$-SJ, whose input size is at most $n$.
+Given the lemma below (which is proved in Appendix~\ref{app:hv}), we assert that $\J(R_1, ..., R_{k-3}, H', V)$ can be computed in $F_{k-1}(n, \out)$ time.
 

### 2024-05-25 msj.tex
yufei

@@
 \extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$.
-Note that $\bm{t}[k-2] = h'$ must intersect with $v$ as $\bm{t}$ is a result tuple of the $(k-1)$-SJ.
-If $p$ is the intersection point between $h$ and $v$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment from the left endpoint of $h'$ to $p$.
+Note that $\bm{t}[k-2] = h'$ must intersect with $v$, because $\bm{t}$ is a result tuple of the $(k-1)$-SJ on $R_1, ..., R_{k-3}, H', V$.
+Suppose $h = [x_1, x_2] \times y$ and $p = (x, y)$ is the intersection point between $h$ and $v$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $h' = [x_1, x] \times y$.
 \myeqn{

### 2024-05-25 msj.tex
yufei

@@
 \myeqn{
-    \contained_{R_{k-2}}(\bm{t}) &=& \{r_{k-2} \in R_{k-2} \mid \text{ the effective horizontal } \nn \\
-    && \text{segment of $\bm{t}$ is contained in $r_{k-2}$}   \}
+    \contained_{R_{k-2}}(\bm{t}) &=& \{r \in R_{k-2} \mid \text{ the effective horizontal } \nn \\
+    && \text{segment of $\bm{t}$ is contained in $r$}   \}
     \label{eqn:type2:contained-t}

### 2024-05-25 msj.tex
yufei

@@
 }
-The above definition should not be confused with \eqref{eqn:contained}, where the $\contained$ function takes a horizontal segment as the input.
+The above should not be confused with \eqref{eqn:contained}, where the ``$\contained$'' function takes a segment as the input, rather than a tuple.
 

### 2024-05-25 msj.tex
yufei

@@
 
-        \item Consider any $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v) \in \J_2$.
+        \item Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
 Let $h'$ be the trimmed segment of $h$, and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
@@
 
-\extraspacing {\bf Set $\bm{H^*}$.} Fix any $i \in [k-2]$. For each segment $h' \in H'$, define
+\extraspacing {\bf Set $\bm{H^*}$.} 
+For each segment $h' \in H'$, define
 \myeqn{
-    \minleft(h') &=& \min_{v \in V: h' \cap v \ne \emptyset} \text{$x$-coordinate of $v$}. 
+    \minleft(h') = \min_{\substack{\text{$\bm{t} \in \J(R_1, ..., R_{k-3}, H', V):$}\\ \text{$\bm{t}[k-2] = h'$}}} \text{$x$-coordinate of $\bm{t}[k-1]$}. 
+    \label{eqn:minleft}
 }
-We set $\minleft(h')$ to $\infty$, if $h'$ intersects no segment in $V$. When $\minleft(h') \ne \infty$, assuming $h' = [x_1, x_2] \times y$, we introduce a horizontal segment
+We reminder the reader that, in the above equation, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment.
+We set $\minleft(h')$ to $\infty$, if no $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ has $h'$ in its field $\bm{t}[k-2]$.
+When $\minleft(h') \ne \infty$, assuming $h' = [x_1, x_2] \times y$, we introduce a horizontal segment
 \myeqn{
@@
 }
-and call it the {\em minimal segment} of $r'$.
+and call it the {\em minimal segment} of $h'$.
 
@@
 
-Next, we construct from $H'$ a new set of rectangles: 
+Next, we construct from $H'$ a new set of horizontal segments: 
 \myeqn{

### 2024-05-25 msj.tex
yufei

@@
 
-Next, we construct from $H'$ a new set of horizontal segments: 
+Next, we construct a new set of horizontal segments: 
 \myeqn{
@@
 }
-Computing $H^*$ from $H'$ and $V$ is an instance of Problem $\mathscr{B}$, and can be achieved in $O(n \log n)$ as per the discussion from Section~\ref{sec:bricks}.
+This can be done in $O(n + k \cdot \out)$ time, as shown in Appendix~\ref{app:hv}.
+
+%Computing $H^*$ from $H'$ and $V$ is an instance of Problem $\mathscr{B}$, and can be achieved in $O(n \log n)$ as per the discussion from Section~\ref{sec:bricks}.
 
@@
 
-\extraspacing {\bf Computing the Top-Sliced Rectangles.}
+\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}
 
@@
 
+\extraspacing {\bf Computing $\bm{H^*}$.}
+
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}

### 2024-05-25 msj.tex
yufei

@@
 We will utilize the sets $\contained_{R_{k-2}}(h^*)$ of the segments $h^*$ in $H^*$, where $\contained_{R_{k-2}}(h^*)$ --- defined in \eqref{eqn:contained} --- is the set of rectangles in $R_{k-2}$ containing $h^*$.
-The next lemma is proven in Appendix~\ref{app:hv}.
+These sets have some useful properties:
 
@@
 
+The proof can be found in Appendix~\ref{app:hv}.
+
 \yf{example}
@@
 \extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} We are ready to produce $\contained_{R_{k-2}}(h^*)$, defined in \eqref{eqn:type2:contained-t}, for any $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, based on Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*}.
-First, fetch the (already computed) minimal segment $h^*$ of $h'$ in $O(1)$ time.
+First, fetch the (already computed) minimal segment $h^*$ of $\bm{t}[k-2]$ in $O(1)$ time.
 Then, scan the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$.

### 2024-05-25 msj.tex
yufei

@@
 Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
-We use $B_\bm{t}$ to represent $\bigcap_{i=1}^t r_i$, namely, the intersection of the rectangles in $\bm{t}$.
-Note that $B_\bm{t}$ is a rectangle itself.
-We define:
+We define
+\myeqn{
+    B_\bm{t} &=& \cap_{i=1}^t r_i \label{eqn:B_t}
+}
+namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Let us introduce:
 \myitems{
@@
 \extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$.
-Note that $\bm{t}[k-2] = h'$ must intersect with $v$, because $\bm{t}$ is a result tuple of the $(k-1)$-SJ on $R_1, ..., R_{k-3}, H', V$.
-Suppose $h = [x_1, x_2] \times y$ and $p = (x, y)$ is the intersection point between $h$ and $v$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $h' = [x_1, x] \times y$.
+Observe that $B_\bm{t}$ --- defined in \eqref{eqn:B_t} --- is a point (due to $h'$ and $v$).
+Suppose $h' = [x_1, x_2] \times y$ and $B_\bm{t} = (x, y)$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $[x_1, x] \times y$.
+
+\vgap
+
+Define
 \myeqn{
@@
 
-We prove the next lemma in Appendix~\ref{app:hv}.
+We prove the next lemma in Appendix~\ref{app:hv} (to avoid confusion, the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).
 
@@
 }
-We reminder the reader that, in the above equation, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment.
 We set $\minleft(h')$ to $\infty$, if no $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ has $h'$ in its field $\bm{t}[k-2]$.
@@
 First, fetch the (already computed) minimal segment $h^*$ of $\bm{t}[k-2]$ in $O(1)$ time.
-Then, scan the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$.
-For each $r$ scanned, check whether $r \in
-\contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers the intersection point between $h$ and $v$; this takes $O(1)$ time. Abort the scan as soon as $r \notin 
+Then, scan the rectangles $r$ of $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$.
+For each $r$ scanned, check whether $r \in \contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers $B_\bm{t}$ (recall that $B_\bm{t}$ is a point); the cost for such a check is $O(1)$.
+Abort the scan as soon as $r \notin
 \contained_{R_{k-2}}(\bm{t})$.
-This way, we compute $\contained_{R_{k-2}}(t)$ in $O(k + |\contained_{R_{k-2}}(t)|)$ time.
-Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+This way, $\contained_{R_{k-2}}(t)$ can be decided in $O(k + |\contained_{R_{k-2}}(t)|)$ time.
+Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3},$ $H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 

### 2024-05-25 msj.tex
yufei

@@
 
-\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} We are ready to produce $\contained_{R_{k-2}}(h^*)$, defined in \eqref{eqn:type2:contained-t}, for any $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, based on Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*}.
+\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*} allows us to produce $\contained_{R_{k-2}}(h^*)$, defined in \eqref{eqn:type2:contained-t}, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ as follows.
 First, fetch the (already computed) minimal segment $h^*$ of $\bm{t}[k-2]$ in $O(1)$ time.

### 2024-05-25 msj.tex
yufei

@@
 \contained_{R_{k-2}}(\bm{t})$.
-This way, $\contained_{R_{k-2}}(t)$ can be decided in $O(k + |\contained_{R_{k-2}}(t)|)$ time.
-Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3},$ $H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+This way, $\contained_{R_{k-2}}(t)$ can be decided in $O(k + |\contained_{R_{k-2}}(\bm{t})|)$ time.
+Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3},$ $H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 

### 2024-05-25 msj.tex
yufei

@@
 \myeqn{
-    \contained_{R_{k-2}}(\bm{t}) &=& \{r \in R_{k-2} \mid \text{ the effective horizontal } \nn \\
+    \contained_{R_{k-2}}(\bm{t}) &=& \{r \in R_{k-2} \mid \text{ the effective horizontal } \nn \\[-1mm]
     && \text{segment of $\bm{t}$ is contained in $r$}   \}

### 2024-05-26 msj.tex
yufei

@@
 
-        \item $\sum_{\bm{t}} |\dcross_H(\bm{t})| \le \out$ and $\sum_{\bm{t}} |\dcross_V(\bm{t})| \le \out$.
+        \item $\sum_{\bm{t}} |\dcross_H(\bm{t})| \le \out$ and $\sum_{\bm{t}} |\dcross_V(\bm{t})| \le \out$, where the two summations are over all $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$.
     }
@@
 
-        \item $\sum_{\bm{t}} |\contained_{R_{k-2}}(\bm{t})| \le \out$.
+        \item $\sum_{\bm{t}} |\contained_{R_{k-2}}(\bm{t})| \le \out$, where the summation is over all $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$.
     }

### 2024-05-26 msj.tex
yufei

@@
 \begin{figure}
-    \includegraphics[height=23mm]{./artwork/guard}
+    \includegraphics[height=27mm]{./artwork/guard}
 
     \figcapup
-    \caption{Define 3-tuple $\bm{t} = \set{r_1, r_2, r_3}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
+    \caption{Consider a 4-tuple $\bm{t} = \set{r_1, r_2, r_3, r_4}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
     \label{fig:guard}
@@
     \begin{tabular}{cc}
-        \includegraphics[height=27mm]{./artwork/type1} &
+        \includegraphics[height=23mm]{./artwork/type1} &
         \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
@@
 
+\begin{figure}
+    \begin{tabular}{cc}
+        \includegraphics[height=27mm]{./artwork/cat1} &
+        \hspace{5mm} \includegraphics[height=27mm]{./artwork/cat2} \\
+        (a) Category 1 &
+        (b) Category 2
+    \end{tabular}
+
+    \figcapup
+    \caption{Classifying $k$-SJ result tuples ($k$ = 4)}
+    \label{fig:ksj:cats}
+    \figcapdown
+\end{figure}
+
 \section{Settling $k$-SJ} \label{sec:ksj}
+
+This section will tackle the (generic form of the) $k$-SJ problem, where the input comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$.
+The join result, denoted by $\J(R_1, R_2, ..., R_k)$, is the set of $k$-tuples $\bm{t} = (r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
+\times R_k$ satisfying the condition that $B_\bm{t}$ --- defined as $\bigcap_{i=1}^k r_i \ne \emptyset$; see \eqref{eqn:B_t} --- is non-empty.
+
+\vgap
+
+Consider any result tuple $\bm{t} \in \J(R_1, R_2, ..., R_k)$, and let $p$ be the top-left corner of $B_\bm{t}$.
+Depending on how $p$ is determined, we classify $\bm{t}$ into one of the two categories below:
+\myitems{
+    \item {{\bf Cat.\ 1:}} $p$ is a corner of a rectangle $\bm{t}[i]$, for some $i \in [k]$.
+
+    \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $\bm{t}[1], ..., \bm{t}[k]$.
+Note that this means $p$ must be the intersection point between the left edge of some rectangle $\bm{t}[i]$ and the top edge of another rectangle $\bm{t}[j]$, where $i, j \in [k]$ and $i \ne j$.
+}
+Figure~\ref{fig:ksj:cats} illustrates a tuple $\bm{t} = (r_1, r_2, ..., r_4)$ of each category.
+
+\vgap
+
+We assume the existence of an algorithm $\A$ that can solve any $(k-1)$-SJ problem in $F_{k-1}(n, \out)$ time, where $n$ and $\out$ are the input and output sizes, respectively.
+Equipped with $\A$, we will find the result tuples of each category separately, within the time complexity stated in Theorem~\ref{thm:main-recur}.
+
+\extraspacing {\bf Category 1.}
 

### 2024-05-26 msj.tex
yufei

@@
 
-Consider any result tuple $\bm{t} \in \J(R_1, R_2, ..., R_k)$, and let $p$ be the top-left corner of $B_\bm{t}$.
+Consider any result tuple $\bm{t} = (r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$, and let $p$ be the top-left corner of $B_\bm{t}$.
 Depending on how $p$ is determined, we classify $\bm{t}$ into one of the two categories below:
 \myitems{
-    \item {{\bf Cat.\ 1:}} $p$ is a corner of a rectangle $\bm{t}[i]$, for some $i \in [k]$.
+    \item {{\bf Cat.\ 1:}} $p$ is the top-left corner of $r_i$, for some $i \in [k]$.
 
-    \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $\bm{t}[1], ..., \bm{t}[k]$.
-Note that this means $p$ must be the intersection point between the left edge of some rectangle $\bm{t}[i]$ and the top edge of another rectangle $\bm{t}[j]$, where $i, j \in [k]$ and $i \ne j$.
+    \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $r_1, ..., r_k$.
+Note that this means $p$ must be the intersection point between the left edge of some rectangle $r_i$ and the top edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
 }
-Figure~\ref{fig:ksj:cats} illustrates a tuple $\bm{t} = (r_1, r_2, ..., r_4)$ of each category.
+Figure~\ref{fig:ksj:cats} illustrates a tuple of each category, assuming $k=4$.
 
@@
 
-We assume the existence of an algorithm $\A$ that can solve any $(k-1)$-SJ problem in $F_{k-1}(n, \out)$ time, where $n$ and $\out$ are the input and output sizes, respectively.
-Equipped with $\A$, we will find the result tuples of each category separately, within the time complexity stated in Theorem~\ref{thm:main-recur}.
+As stated in Theorem~\ref{thm:main-recur}, we are given an algorithm $\A$ that can solve any $(k-1)$-SJ problem in $F_{k-1}(n, \out)$ time, where $n$ and $\out$ are the input and output sizes, respectively.
+Equipped with $\A$, we will show how to find the result tuples of each category within the time complexity of \eqref{eqn:main:reccurrence}.
 
-\extraspacing {\bf Category 1.}
+\extraspacing {\bf Category 1.} Fix an arbitrary $i \in [k]$.
+We now explain how to find all the result tuples $\bm{t} \in \J(R_1, R_2, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$.
 

### 2024-05-26 msj.tex
yufei

@@
 \def\contained{\mathrm{contained}}
+\def\cat{\mathrm{cat1}}
 \def\out{\mathrm{OUT}}
@@
 \extraspacing {\bf Category 1.} Fix an arbitrary $i \in [k]$.
-We now explain how to find all the result tuples $\bm{t} \in \J(R_1, R_2, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$.
+Define $\J^\cat_i$ as the set of result tuples $\bm{t} = (r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$.
+We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.
+
+\vgap
+
+For every $\bm{t} = (r_1, r_2, ..., r_k) \in \J^\cat_k$, the top-left corner of $r_k$ must be covered by all of $r_1, ..., r_{k-1}$.
+This observation motivates us to find the target result tuples as follows.
+First, collect the set $P$ of top-left corners of all the rectangles in $R_k$.
+Remove from $P$ every point $p$ with the property that, there exists at least one $j \in [k-1]$ such that $p$ is covered by no rectangle in $R_j$.
+This requires solving $k-1$ instances of the detection version of Problem $\mathscr{A}$ (pairing $P$ together with a different $R_j$ where $j \in [k-1]$ as the input in each instance); the cost is $O(k n \log n)$ by the discussion in Section~\ref{sec:bricks}.
+
+\vgap
+
+Let $P'$ be the set of remaining points in $P$ after the aforementioned removal.
+Next, for each $j \in [k-1]$, solve the reporting version of Problem $\mathscr{A}$ by feeding $P'$ and $R_j$ as the input.
+Doing so produces, for each point $p \in P'$, $\contained_{R_j}(p)$, which --- as defined in \eqref{eqn:contained} (noticing that $p$ is a degenerated ``horizontal segment'') --- is the set of rectangles in $R_j$ covering $p$.
+By the discussion in Section~\ref{sec:bricks}, the total cost of this step is bounded by
+\myeqn{
+    O\Big(kn \log n + \sum_{j \in [k-1]} \sum_{p \in P'} |\contained_{R_j}(p)| \Big). \label{eqn:ksj:cat1:help1}
+}
+
+We are now ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r_k \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
+\myeqn{
+    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \label{eqn:ksj:cat1:help2}
+}
 

### 2024-05-26 msj.tex
yufei

@@
 \myeqn{
-    O\Big(kn \log n + \sum_{j \in [k-1]} \sum_{p \in P'} |\contained_{R_j}(p)| \Big). \label{eqn:ksj:cat1:help1}
+    O\Big(kn \log n + \sum_{p \in P'} \sum_{j \in [k-1]} |\contained_{R_j}(p)| \Big). \label{eqn:ksj:cat1:help1}
 }
@@
 }
+we add $(r_1, ..., r_{k-1}, r_k)$ to $\J^\cat_k$.
+Performing the above for all $p \in P'$ correctly generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time.
+The way $P'$ is computed ensures that $\contained_{R_j}(p)$ is non-empty for all $j \in [k-1]$.
+Hence, $\sum_{j \in [k-1]} |\contained_{R_j}(p)|$ is bounded by $\prod_{j \in [k-1]} |\contained_{R_j}(p)|$.
+This implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.
 

### 2024-05-26 msj.tex
yufei

@@
 
-This section will tackle the (generic form of the) $k$-SJ problem, where the input comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$.
-The join result, denoted by $\J(R_1, R_2, ..., R_k)$, is the set of $k$-tuples $\bm{t} = (r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
-\times R_k$ satisfying the condition that $B_\bm{t}$ --- defined as $\bigcap_{i=1}^k r_i \ne \emptyset$; see \eqref{eqn:B_t} --- is non-empty.
+This section will tackle the $k$-SJ problem in its general form, where the input comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$.
+The join result $\J(R_1, R_2, ..., R_k)$ is the set of $k$-tuples $\bm{t} = (r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
+\times R_k$ satisfying the condition that $B_\bm{t}$ --- which is $\bigcap_{i=1}^k r_i \ne \emptyset$ (see \eqref{eqn:B_t}) --- is non-empty.
 
@@
 \myitems{
-    \item {{\bf Cat.\ 1:}} $p$ is the top-left corner of $r_i$, for some $i \in [k]$.
+    \item {{\bf Cat.\ 1:}} $p$ is the top-left corner of $r_i$ for some $i \in [k]$.
 
     \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $r_1, ..., r_k$.
-Note that this means $p$ must be the intersection point between the left edge of some rectangle $r_i$ and the top edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
+This means $p$ must be the intersection point between the left edge of some rectangle $r_i$ and the top edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
 }
@@
 
-We are now ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r_k \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
+We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r_k \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
 \myeqn{
@@
 This implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.
+
+\vgap
+
+We can now conclude that the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k^2 n \log n + k \cdot \sum_{i=1}^k |\J^\cat_i|) = O(k^2 n \log n + k \cdot \out)$.
 

### 2024-05-26 msj.tex
yufei

@@
 For every $\bm{t} = (r_1, r_2, ..., r_k) \in \J^\cat_k$, the top-left corner of $r_k$ must be covered by all of $r_1, ..., r_{k-1}$.
-This observation motivates us to find the target result tuples as follows.
+This observation motivates us to find $\J^\cat_i$ as follows.
 First, collect the set $P$ of top-left corners of all the rectangles in $R_k$.
 Remove from $P$ every point $p$ with the property that, there exists at least one $j \in [k-1]$ such that $p$ is covered by no rectangle in $R_j$.
-This requires solving $k-1$ instances of the detection version of Problem $\mathscr{A}$ (pairing $P$ together with a different $R_j$ where $j \in [k-1]$ as the input in each instance); the cost is $O(k n \log n)$ by the discussion in Section~\ref{sec:bricks}.
+This requires solving $k-1$ instances of the detection version of Problem $\mathscr{A}$ (in each instance, the input includes $P$ together with a different $R_j$, $j \in [k-1]$); the cost is $O(k n \log n)$ by the discussion in Section~\ref{sec:bricks}.
 
@@
 Next, for each $j \in [k-1]$, solve the reporting version of Problem $\mathscr{A}$ by feeding $P'$ and $R_j$ as the input.
-Doing so produces, for each point $p \in P'$, $\contained_{R_j}(p)$, which --- as defined in \eqref{eqn:contained} (noticing that $p$ is a degenerated ``horizontal segment'') --- is the set of rectangles in $R_j$ covering $p$.
+Doing so produces, for each point $p \in P'$, the set $\contained_{R_j}(p)$, which --- as defined in \eqref{eqn:contained},  noticing that $p$ is a degenerated ``horizontal segment'' --- is the set of rectangles in $R_j$ covering $p$.
 By the discussion in Section~\ref{sec:bricks}, the total cost of this step is bounded by
@@
 
-We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r_k \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
+We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
 \myeqn{
@@
 }
-we add $(r_1, ..., r_{k-1}, r_k)$ to $\J^\cat_k$.
-Performing the above for all $p \in P'$ correctly generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time.
-The way $P'$ is computed ensures that $\contained_{R_j}(p)$ is non-empty for all $j \in [k-1]$.
-Hence, $\sum_{j \in [k-1]} |\contained_{R_j}(p)|$ is bounded by $\prod_{j \in [k-1]} |\contained_{R_j}(p)|$.
-This implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.
+we add $(r_1, ..., r_{k-1}, r)$ to $\J^\cat_k$.
+Performing the above for all $p \in P'$ generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time.
+The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for all $j \in [k-1]$.
+Hence, we have $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.
 
@@
 
-We can now conclude that the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k^2 n \log n + k \cdot \sum_{i=1}^k |\J^\cat_i|) = O(k^2 n \log n + k \cdot \out)$.
+Therefore, the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k^2 n \log n + k \cdot \sum_{i=1}^k |\J^\cat_i|) = O(k^2 n \log n + k \cdot \out)$.
 

### 2024-05-26 msj.tex
yufei

@@
 \def\cat{\mathrm{cat1}}
+\def\catt{\mathrm{cat2}}
 \def\out{\mathrm{OUT}}
@@
     \item {{\bf Cat.\ 2:}} $p$ is not a corner of any of $r_1, ..., r_k$.
-This means $p$ must be the intersection point between the left edge of some rectangle $r_i$ and the top edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
+This means $p$ must be the intersection point between the top edge of some rectangle $r_i$ and the left edge of another rectangle $r_j$, where $i, j \in [k]$ and $i \ne j$.
 }
@@
 
-\extraspacing {\bf Category 1.} Fix an arbitrary $i \in [k]$.
-Define $\J^\cat_i$ as the set of result tuples $\bm{t} = (r_1, r_2, ..., r_k) \in \J(R_1, R_2, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$.
+\extraspacing {\bf Category 1.} Given an $i \in [k]$, we denote by $\J^\cat_i$ the set of $k$-tuples $\bm{t} = (r_1, ..., r_k) \in \J(R_1, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$.
 We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.
@@
 The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for all $j \in [k-1]$.
-Hence, we have $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|)$.
+Hence, we have $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|) = O(kn \log n + k \cdot \out)$ .
 
@@
 
-Therefore, the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k^2 n \log n + k \cdot \sum_{i=1}^k |\J^\cat_i|) = O(k^2 n \log n + k \cdot \out)$.
+Therefore, the total time of computing all of $\J^\cat_1, ..., \J^\cat_k$ is $O(k) \cdot (kn \log n + k \cdot \out)$.
+A category-1 result tuple $\bm{t}$ may be seen more than once (this happens if the top-left corner of $B_\bm{t}$ is the top-left corner of more than one rectangle in $\bm{t}$).
+Duplicate removal can be achieved at no extra cost asymptotically, following the ideas explained in Section~\ref{sec:hv:type2}.
+
+\extraspacing {\bf Category 2.} Given $i, j \in [k]$ with $i \ne j$, we denote by $\J^\catt_{i,j}$ the set of $k$-tuples $\bm{t} = (r_1, ..., r_k) \in \J(R_1, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the intersection between the top edge of $r_i$ and the left edge of $r_j$.
+The Category 2 of result tuples is the union of the $\J^\catt_{i,j}$ of all possible $i, j$.
+
+top-left corner of $\bm{t}[i]$. We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.
 

### 2024-05-26 msj.tex
yufei

@@
 
-top-left corner of $\bm{t}[i]$. We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.
+\vgap
+
+The computation of $\J^\catt_{i,j}$ is an instance of H-V $k$-SJ.
+Specifically, collect the top-edges of all rectangles $r_i \in R_i$ into a set $H$, and collect the left-edges of all rectangles $r_j \in R_j$ into a set $V$.
+This yields an H-V $k$-SJ problem whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$, and $H$ and $V$
 

### 2024-05-26 msj.tex
yufei

@@
 Specifically, collect the top-edges of all rectangles $r_i \in R_i$ into a set $H$, and collect the left-edges of all rectangles $r_j \in R_j$ into a set $V$.
-This yields an H-V $k$-SJ problem whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$, and $H$ and $V$
+This yields an H-V $k$-SJ problem whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$.
+Its output includes every $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, $h \in H$, and $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$.
 

### 2024-05-26 msj.tex
yufei

@@
 Its output includes every $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, $h \in H$, and $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$.
+There is one-one correspondence between the output of the H-V $k$-SJ and $\J^\catt_{i,j}$.
+Thus, by Lemma~\ref{lmm:hv}, the H-V $k$-SJ can be solved in $O(k) \cdot (F_{k-1}(n, |\J^\catt_{i,j}|) + n \log n + k \cdot |\J^\catt_{i,j}|)$ time.
+Converting the output into $\J^\catt_{i,j}$ takes another $O(|\J^\catt_{i,j}|)$ time.
+Applying the trivial bound $|\J^\catt_{i,j}| \le \out$, we can assert that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time
+
+\vgap
+
+Performing the above for all $i, j \in [k]$ with $i \ne j$ results in a total time complexity of $O(k^3) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$.
+A category-1 result tuple $\bm{t}$ may be seen more than once (this happens if the top-left corner of $B_\bm{t}$ is the top-left corner of more than one rectangle in $\bm{t}$).
+Duplicate removal can be achieved at no extra cost asymptotically, following the ideas explained in Section~\ref{sec:hv:type2}.
 

### 2024-05-26 msj.tex
yufei

@@
 
-\vgap
+\subsection{Proof of Theorem~\ref{thm:main-recur}} \label{sec:ksj:thm1}
 
@@
 Performing the above for all $i, j \in [k]$ with $i \ne j$ results in a total time complexity of $O(k^3) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$.
-A category-1 result tuple $\bm{t}$ may be seen more than once (this happens if the top-left corner of $B_\bm{t}$ is the top-left corner of more than one rectangle in $\bm{t}$).
-Duplicate removal can be achieved at no extra cost asymptotically, following the ideas explained in Section~\ref{sec:hv:type2}.
+A category-2 result tuple $\bm{t}$ may be seen more than once (this can happen if, for example, more than one rectangle in $\bm{t}$ has the same top-edge or the same left-edge).
+Again, duplicate removal can be achieved at no extra cost asymptotically.
+
+\vgap
+
+We now complete the proof of Theorem~\ref{thm:main-recur}.
 

### 2024-05-26 msj.tex
yufei

@@
 
-\subsection{Proof of Theorem~\ref{thm:main-recur}} \label{sec:ksj:thm1}
+\vgap
 
+The rest of this section serves as a proof of Theorem~\ref{thm:main-recur}.
+Theorem~\ref{thm:main-alg} is a corollary of Theorem~\ref{thm:main-recur}, as we show in Appendix~\ref{app:main-alg}.
 As stated in Theorem~\ref{thm:main-recur}, we are given an algorithm $\A$ that can solve any $(k-1)$-SJ problem in $F_{k-1}(n, \out)$ time, where $n$ and $\out$ are the input and output sizes, respectively.
@@
 
+\section{Proof of Theorem~\ref{thm:main-alg}} \label{app:main-alg}
+
 \section{Hardness of 3-SJ in 3D Space} \label{app:lb-cond}

### 2024-05-26 msj.tex
yufei

@@
 
-The rest of this section serves as a proof of Theorem~\ref{thm:main-recur}.
-Theorem~\ref{thm:main-alg} is a corollary of Theorem~\ref{thm:main-recur}, as we show in Appendix~\ref{app:main-alg}.
+The rest of this section serves as a proof of Theorem~\ref{thm:main-recur}. Theorem~\ref{thm:main-alg} is a corollary of Theorem~\ref{thm:main-recur}, as proved in Appendix~\ref{app:main-alg}.
+
+\vgap
+
 As stated in Theorem~\ref{thm:main-recur}, we are given an algorithm $\A$ that can solve any $(k-1)$-SJ problem in $F_{k-1}(n, \out)$ time, where $n$ and $\out$ are the input and output sizes, respectively.
@@
 This yields an H-V $k$-SJ problem whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$.
-Its output includes every $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, $h \in H$, and $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$.
+Each result tuple includes every $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, $h \in H$, and $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$.
 There is one-one correspondence between the output of the H-V $k$-SJ and $\J^\catt_{i,j}$.
@@
 
-Performing the above for all $i, j \in [k]$ with $i \ne j$ results in a total time complexity of $O(k^3) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$.
+Performing the above for all $i, j \in [k]$ with $i \ne j$ leads to a total time complexity of $O(k^3) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$.
 A category-2 result tuple $\bm{t}$ may be seen more than once (this can happen if, for example, more than one rectangle in $\bm{t}$ has the same top-edge or the same left-edge).

### 2024-05-26 msj.tex
yufei

@@
 The computation of $\J^\catt_{i,j}$ is an instance of H-V $k$-SJ.
-Specifically, collect the top-edges of all rectangles $r_i \in R_i$ into a set $H$, and collect the left-edges of all rectangles $r_j \in R_j$ into a set $V$.
+Specifically, collect the top-edges of all rectangles of $R_i$ into a set $H$, and collect the left-edges of all the rectangles of $R_j$ into a set $V$.
 This yields an H-V $k$-SJ problem whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$.

### 2024-05-26 msj.tex
yufei

@@
 Converting the output into $\J^\catt_{i,j}$ takes another $O(|\J^\catt_{i,j}|)$ time.
-Applying the trivial bound $|\J^\catt_{i,j}| \le \out$, we can assert that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time
+Applying $|\J^\catt_{i,j}| \le \out$, we can assert that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time
 

### 2024-05-26 msj.tex
yufei

@@
 This yields an H-V $k$-SJ problem whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$.
-Each result tuple includes every $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, $h \in H$, and $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$.
+Each result tuple consists of a rectangle $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, a horizontal segment $h \in H$, and a vertical segment $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$.
 There is one-one correspondence between the output of the H-V $k$-SJ and $\J^\catt_{i,j}$.

### 2024-05-27 msj.tex
yufei

@@
 \extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$.
-Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $\bm{t} = (e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ...
-\times S_k$ as a $k$-dimensional vector such that $\bm{t}[i] = e_i$ for each $i \in [k]$.
-Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle.
+Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ...
+\times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$.
+Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$.
 All logarithms have base 2 by default.
@@
 Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem.
-For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$.
+Define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$.
 Our core technical contribution is to establish:
@@
 
-The theorem implies the existence of a recursive nature of $k$-SJ.
-Indeed, we will see that the $k$-SJ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all of which have input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$.
-Furthermore, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues.
-For 2-SJ, we can set $\A$ simply to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.
-Combining this with \eqref{eqn:main:reccurrence} yields a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ.
+The theorem implies a recursive nature of $k$-SJ.
+Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$.
+Moreover, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues.
+For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.
+Combining this with \eqref{eqn:main:reccurrence} gives a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ.
 Solving the recurrence yields:
@@
 
+\input{thm1.2}
+
 \section{Hardness of 3-SJ in 3D Space} \label{app:lb-cond}
+
+\input{hard-3d}
 

### 2024-05-27 msj.tex
yufei

@@
 Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
-A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$.
+A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s \cap r \ne \emptyset$ but $r$ covers neither of the two endpoints of $s$.
 A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
@@
 \myeqn{
-    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}. 
+    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}
     \label{eqn:cross}
 }
+namely, the set of segments in $S$ crossing $r$.
 Let $R$ be a set of rectangles. Given a horizontal segment, we define 
 \myeqn{
-    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}. 
+    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}
     \label{eqn:contained}
 }
+namely, the set of rectangles in $R$ containing $h$.
+
+\vgap
 
@@
 }
-namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Let us introduce:
+namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, define:
 \myitems{
@@
 Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$.
-In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time, where $n = |P| + |R|$.
+This can be achieved in $O(n \log n)$ time where $n = |P| + |R|$ as shown
+in Appendix~\ref{app:bricks}.
 
@@
 For example, in Figure~\ref{fig:guard}a, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$.
-In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
+As shown in Appendix~\ref{app:bricks}, this can be achieved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
 
-\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space.
+\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments.
 The goal is to report, for each segment $h \in H$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$.
@@
 Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |V|$.
+This can be achieved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.
 
-\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
-The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that at least one rectangle $r \in R$ covers both the point $(x_1, y)$ and the point $(x, y)$.
+\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ such that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that at least one rectangle $r \in R$ covers both the point $(x_1, y)$ and the point $(x, y)$.
 If the point $p$ exists (i.e., $h$ has at least one left-end covering rectangle in $R$), we should output a tuple $(h, p)$; otherwise, output nothing for $h$.
@@
 
-\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
 In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see \eqref{eqn:cross}).
@@
 Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
-In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
+This can be achieved in $O(n \log n)$ time where $n = |H| + |R|$ as shown
+In Appendix~\ref{app:bricks}.
 
@@
 
-\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space.
+\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
 The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries.

### 2024-05-27 msj.tex
yufei

@@
 
-In the {\em reporting version} of the problem, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$, provided that at least one such $r$ exists.
-For example, in Figure~\ref{fig:guard}a, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$.
+In the {\em reporting version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$.
+In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_1, \emptyset), (p_2: r_1, r_2), (p_3: r_2)}$.
 As shown in Appendix~\ref{app:bricks}, this can be achieved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
@@
 Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-This can be achieved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.
+This problem can be solved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.
 
@@
 Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.
+This problem can be solved in in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
 
 \extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
-In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see \eqref{eqn:cross}).
+In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see its definition in \eqref{eqn:cross}).
 If no segment in $H$ crosses $r$, output nothing for $r$.
 Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
-This can be achieved in $O(n \log n)$ time where $n = |H| + |R|$ as shown
-In Appendix~\ref{app:bricks}.
+This can be achieved in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
 
@@
 In the {\em find-all-sorted} version of the problem, the goal is to report, for each rectangle $r \in R$, the entire $\cross_H(r)$ sorted by y-coordinate.
-Formally, if $\cross_H(r) = \set{h_1, h_2, ..., h_z}$ (for some $z \ge 1$), we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$).
+Formally, if $\cross_H(r) = \set{h_1, h_2, ..., h_z}$ for some $z \ge 1$, we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-coordinate of $h_i$ (resp., $h_{i-1}$).
 In the example of Figure~\ref{fig:probs}d, the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.

### 2024-05-27 msj.tex
yufei

@@
 The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries.
-Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
-Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
+Formally, if $r_1, r_2, ..., r_z$ for some $z \ge 1$ are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
+Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3), (h_3: \emptyset}$.
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

### 2024-05-27 msj.tex
yufei

@@
 \begin{theorem} \label{thm:main-recur}
-    Equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in time
+    Equipped with the algorithm $\A$ as described above, the $k$-SJ problem with $k \ge 3$ can be solved in time
     \myeqn{
@@
 Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ...,$ $R_k$.
-We now formulate a special version of $k$-SJ, named the {\em H-V $k$-SJ} problem.
+We now formulate a special version of $k$-SJ, named the {\em H-V $k$-SJ problem}.
 The special nature is reflected in the introduction of three constraints: (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
@@
 Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
-To facilitate our analysis, let us define the 1-SJ problem as follows: the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$.
-The 1-SJ problem can be trivially ``solved'' in $O(n)$ time.
-We assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [1, k-1]$.
+To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time.
+We assume the existence of an algorithm $\A$ that can settle $\kappa$-SJ for all $\kappa \in [1, k-1]$.
 Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
@@
 
-This section will tackle the $k$-SJ problem in its general form, where the input comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$.
+This section will tackle the $k$-SJ problem in its general form, where the input comprises $k \ge 3$ sets of rectangles $R_1, R_2, ..., R_k$.
 The join result $\J(R_1, R_2, ..., R_k)$ is the set of $k$-tuples $\bm{t} = (r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...

### 2024-05-27 msj.tex
yufei

@@
 The part of the paper from this point till the end of Section~\ref{sec:hv:type2} will be devoted to proving the above lemma.
-This is the most challenging step in settling the general $k$-SJ problem optimally, as we will see in Section~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
+This is the most challenging step in solving the general $k$-SJ problem optimally, as will be discussed in Section~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
 

### 2024-05-27 msj.tex
yufei

@@
 Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.
-
-\vgap
-
 In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.
@@
 We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.
-Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas.
+Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.
 
@@
     \begin{tabular}{cc}
-        \includegraphics[height=23mm]{./artwork/type1} &
-        \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
+        \includegraphics[height=27mm]{./artwork/type1} &
+        \hspace{5mm} \includegraphics[height=23mm]{./artwork/type2} \\
         (a) Type 1 &
@@
 where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ.
-Remember that the output size $\out$ is defined to $|\J(R_1, ..., R_{k-2}, H, V)|$.
+Remember that the output size $\out$ is defined as $|\J(R_1, ..., R_{k-2}, H, V)|$.
 From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.
@@
 By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r \in R_i$.
-It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(n \log n)$ total time.
+It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(kn \log n)$ total time.
 
@@
 
-We now solve a $(k-2)$-SJ problem with $R_1', R_2', ..., R_{k-2}'$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+We now solve a $(k-2)$-SJ problem on the input $\set{R_1', R_2', ..., R_{k-2}'}$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
 This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$.
@@
 
-We remind the reader that $\out$ is the output size of the original H-V $k$-SJ problem.
 As a corollary of Lemma~\ref{lmm:hv:type1:recur-output}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
@@
 
-We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(kn \log n +  k \cdot \out)$ time.
 

### 2024-05-27 msj.tex
yufei

@@
 The rest of the section will focus on how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
-An analogous method can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ in the same time complexity.
+An analogous method can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ within the same time complexity.
 
@@
 Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$ (with $H$ and $R^*_i$ as the input).
-Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$.
+Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(k n \log n + \out)$.
 Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.

### 2024-05-27 thm1.2.tex
thm1.2

Theorem~\ref{thm:main-recur} implies the existence of a recursive algorithm $\A$ for $k$-SJ. For 2-SJ, we can set $\A$ to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$. For $k \geq 3$, $\A$ recursively solves $O(k^3)$ subproblems of $k-1$-SJ---each of which has an input size of at most $n$ and an output size of at most $\out$---and then produces the output for $k$-SJ using $O(n\log n + k\cdot \out)$ additional processing
time. It is rudimentary to prove that $\A$ is comparison-based by induction.
Moreover, we have the following recurrence.
For $k \geq 3$, there is a constant $c$ such that
\myeqn{
    F_{k}(n, \out) = c \cdot k^3 \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big).\nn
}
We solve the recurrence by expanding it $k-2$ times. And we get
\myeqn{
    F_{k}(n,\out) &=& c \cdot k^3(n\log n + k\cdot \out )+ \nn \\
    && c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out)\nn\\
    &\leq& 2c^2\cdot k^3(k-1)^3(n\log n + k \cdot\out)+  \nn\\
    && c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) \nn \\
    &\leq& 3c^3\cdot k^3(k-1)^3(k-2)^3(n\log n + k \cdot\out)+  \nn\\
    && c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out)\nn\\
    &\leq& ... \nn \\
    &\leq& (k-2)c^{k-2}(k!/2)^3(n\log n + k \cdot\out) \nn \\
    && + c^{k-2}(k!/2)^3 \cdot F_{2}(n,\out) \nn \\
    &=& (k-1)c^{k-2}(k!/2)^3(n\log n + k \cdot\out) \nn \\
    &=& O((c')^k(n\log n + k \cdot\out)), \nn
}
where $c'$ is a constant satisfying that $(c')^k \geq (k-1)c^{k-2}/2^3$. 
As a result, Theorem~\ref{thm:main-alg} holds.

### 2024-05-27 msj.tex
yufei

@@
 
-From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time (e.g., by planesweep \cite{bcko08}).
+From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved by a comparison-based algorithm in $O(n \log n + \out)$ time (e.g., by planesweep \cite{bcko08}).
 However, the problem becomes much more challenging for $k \ge 3$, known as the {\em multiway} scenario.
@@
 Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(k n \log n + \out)$.
-Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.
+Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein have been sorted in ascending order of y-coordinate.
 
@@
 
-\input{thm1.2}
+As mentioned in Section~\ref{sec:intro:prev}, 2-SJ can be solved in $F_2(n, \out) = O(n \log n + \out)$ time using a comparison-based algorithm.
+By Theorem~\ref{thm:main-recur}, in general, a $(k-1)$-SJ algorithm with runtime $F_{k-1}(n, \out)$ spawns a $k$-SJ algorithm whose running time $F_k(n, \out)$ obeys \eqref{eqn:main:reccurrence}.
+More specifically, for $k \ge 3$, there is a constant $c > 0$ such that
+\myeqn{
+    && F_{k}(n, \out) \nn \\
+    &\le& c \cdot k^3 \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big) \nn \\
+    &=& c \cdot k^3(n\log n + k\cdot \out )+ \nn \\
+    && c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out)\nn\\
+    &<& (c+c^2)\cdot k^3(k-1)^3(n\log n + k \cdot\out)+  \nn\\
+    && c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) \nn \\
+    &\leq& (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3(n\log n + k \cdot\out)+  \nn\\
+    && c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out)\nn\\
+    &\leq& ... \nn \\
+    &\leq& \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!/2)^3\cdot (n\log n + k \cdot\out)+ \nn \\
+    && c^{k-2}(k!/2)^3 \cdot F_{2}(n,\out) \nn \\
+    &\leq & 2c^{k-1}(k!/2)^3 \cdot O(n\log n + k \cdot\out). \nn
+}
+As a result, Theorem~\ref{thm:main-alg} holds.
 

### 2024-05-27 msj.tex
yufei

--- 2024-05-27 msj.tex
@@
     }
-    where $n$ (resp., $\out$) is the input (resp., output) size of the problem.
+    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the $k$-SJ algorithm.
+
 %     time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
@@
 \begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + k \cdot \out))$ time.
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^{k-1} (k!)^3 \cdot (n \log n + k \cdot \out))$ time.
     %and $O(k \cdot n + \out)$ space, where $c > 1$ is a positive constant.
@@
 
-\input{thm1.2}
+As mentioned in Section~\ref{sec:intro:prev}, 2-SJ can be solved in $F_2(n, \out) = O(n \log n + \out)$ time using a comparison-based algorithm.
+By Theorem~\ref{thm:main-recur}, in general, a comparison-based $(k-1)$-SJ algorithm with runtime $F_{k-1}(n, \out)$ spawns a comparison-based $k$-SJ algorithm whose running time $F_k(n, \out)$ obeys \eqref{eqn:main:reccurrence}.
+Specifically, for $k \ge 3$, there is a constant $c > 1$ such that
+\myeqn{
+    && F_{k}(n, \out) \nn \\
+    &\le& c \cdot k^3 \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big) \nn \\
+    &=& c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out) + \nn \\
+    && c \cdot k^3(n\log n + k\cdot \out ) \nn\\
+    &<& c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) +  \nn\\
+    && (c+c^2)\cdot k^3(k-1)^3(n\log n + k \cdot\out) \nn \\
+    &\leq& c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out) + \nn\\
+    && (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3(n\log n + k \cdot\out) \nn\\
+    &\leq& ... \nn \\
+    &\leq& c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \nn \\
+    && \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3\cdot (n\log n + k \cdot\out) \nn \\
+    &\leq & 2c^{k-1} \cdot (k!)^3 \cdot O(n\log n + k \cdot\out) \nn
+}
+which completes the proof of Theorem~\ref{thm:main-alg}.
 

### 2024-05-27 msj.tex
yufei

@@
 Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$.
-Moreover, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues.
 For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.
@@
 
-%As we will see, conquering this sepcial version is the most challenging step in settling the original $k$-SJ problem optimally.
-
 \vgap
@@
     }
-    where $n$ (resp., $\out$) is the input (resp., output) size of the problem.
+    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the H-V $k$-SJ algorithm.
 \end{lemma}
@@
     &=& c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out) + \nn \\
-    && c \cdot k^3(n\log n + k\cdot \out ) \nn\\
+    && c \cdot k^3 \cdot (n\log n + k\cdot \out ) \nn\\
     &<& c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) +  \nn\\
-    && (c+c^2)\cdot k^3(k-1)^3(n\log n + k \cdot\out) \nn \\
+    && (c+c^2)\cdot k^3(k-1)^3 \cdot (n\log n + k \cdot\out) \nn \\
     &\leq& c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out) + \nn\\
-    && (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3(n\log n + k \cdot\out) \nn\\
+    && (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3 \cdot (n\log n + k \cdot\out) \nn\\
     &\leq& ... \nn \\
     &\leq& c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \nn \\
-    && \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3\cdot (n\log n + k \cdot\out) \nn \\
+    && \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3 \cdot (n\log n + k \cdot\out) \nn \\
     &\leq & 2c^{k-1} \cdot (k!)^3 \cdot O(n\log n + k \cdot\out) \nn

### 2024-05-27 msj.tex
yufei

@@
 It is possible for the result tuple to belong to multiple sub-types simultaneously.
-
-\vgap
-
-Next, we will focus on producing the result tuples of a particular sub-type: those tuples $(r_1, ..., r_{k-2}, h, v)$ where $r_{k-2}$ covers the left endpoint of $h$ --- let us denote this set of result tuples as $\J_2$.
-Other sub-types can be computed using the same algorithm, by symmetry.
+Next, we will focus on producing the result tuples of a particular sub-type:
+\myeqn{
+    \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers} \nn \\[-1mm]
+    && \text{the left endpoint of $h$}\}. \nn
+}
+Other sub-types can be found analogously.
 

### 2024-05-27 msj.tex
yufei

@@
 \extraspacing {\bf Set $\bm{H'}$.} Take any horizontal segment $h = [x_1, x_2] \times y \in H$.
-Let $p$ be the rightmost point $p$ on $h$ satisfying the condition that at least one rectangle in $R_{k-2}$ covers $p$.
-This point $p$ exists if and only if $h$ has at least one left-end covering rectangle in $R_{k-2}$.
+Let $p$ be the rightmost point on $h$ satisfying the condition that at least one rectangle in $R_{k-2}$ covers $p$.
+This $p$ exists if and only if $h$ has at least one left-end covering rectangle in $R_{k-2}$.
 If $p$ exists and has coordinates $(x, y)$, we refer to the segment $h' = [x_1, x] \times y$ as the {\em trimmed segment} of $h$; conversely, we call $h$ the {\em full segment} of $h'$.
@@
 
-Now, solve a $(k-1)$-SJ problem with $R_1, ..., R_{k-3}, H', V$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+Now, solve a $(k-1)$-SJ problem on the input $\set{R_1, ..., R_{k-3}, H', V}$ using the algorithm $\A$ supplied (by Lemma~\ref{lmm:hv}).
 Let $\J(R_1, ...,$ $R_{k-3}, H', V)$ represent the result of this $(k-1)$-SJ, whose input size is at most $n$.
@@
 \extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$.
-Observe that $B_\bm{t}$ --- defined in \eqref{eqn:B_t} --- is a point (due to $h'$ and $v$).
+Note that $B_\bm{t}$ --- defined in \eqref{eqn:B_t} --- is a point (due to $h'$ and $v$).
 Suppose $h' = [x_1, x_2] \times y$ and $B_\bm{t} = (x, y)$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $[x_1, x] \times y$.
-
-\vgap
-
-Define
+This allows us to define
 \myeqn{
@@
 }
-The above should not be confused with \eqref{eqn:contained}, where the ``$\contained$'' function takes a segment as the input, rather than a tuple.
+The above should not be confused with \eqref{eqn:contained}, where the ``$\contained$'' function takes a segment as the parameter, rather than a tuple.
 
@@
         \item Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
-Let $h'$ be the trimmed segment of $h$, and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
+Let $h'$ be the trimmed segment of $h$ and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
 Then, $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ and $h' \in \contained_{R_{k-2}}(\bm{t})$.
@@
 The correctness of the algorithm follows from statements (1) and (2) of Lemma~\ref{lmm:hv:type2:properties}.
-Furthermore, if we are given $\contained_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+Furthermore, if we are given $\contained_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of Lemma~\ref{lmm:hv:type2:properties}
+assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 

### 2024-05-27 msj.tex
yufei

@@
     \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers} \nn \\[-1mm]
-    && \text{the left endpoint of $h$}\}. \nn
+    && \text{the left endpoint of $h$}\}. \label{eqn:hv:type2:J2}
 }
@@
 
-\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*} allows us to produce $\contained_{R_{k-2}}(h^*)$, defined in \eqref{eqn:type2:contained-t}, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ as follows.
+\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*} allows us to produce $\contained_{R_{k-2}}(\bm{t})$ --- defined in \eqref{eqn:type2:contained-t} --- for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ as follows.
 First, fetch the (already computed) minimal segment $h^*$ of $\bm{t}[k-2]$ in $O(1)$ time.
 Then, scan the rectangles $r$ of $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$.
-For each $r$ scanned, check whether $r \in \contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers $B_\bm{t}$ (recall that $B_\bm{t}$ is a point); the cost for such a check is $O(1)$.
+For each $r$ scanned, check whether $r \in \contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers $B_\bm{t}$ (recall that $B_\bm{t}$ is a point); the cost of this inspection is $O(1)$.
 Abort the scan as soon as $r \notin
@@
 
-We conclude that $\J_2$ --- the set of type-2 result tuples $(r_1, ..., r_{k-2},$ $h, v)$ where $r_{k-2}$ covers the left endpoint of $h$ --- can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
+We conclude that $\J_2$ --- see \eqref{eqn:hv:type2:J2} --- can be computed in $F_{k-1}(n, \out) + O(n \log n +  k \cdot \out)$ time.
 Remember that, to generate the entire type-2 result, we need to repeat the algorithm $4(k-2)$ times (one for each sub-type).

### 2024-05-27 msj.tex
yufei

@@
 
-The rest of this section serves as a proof of Theorem~\ref{thm:main-recur}. Theorem~\ref{thm:main-alg} is a corollary of Theorem~\ref{thm:main-recur}, as proved in Appendix~\ref{app:main-alg}.
-
-\vgap
-
+The rest of this section serves as a proof of Theorem~\ref{thm:main-recur}.
+Theorem~\ref{thm:main-alg} is a corollary of Theorem~\ref{thm:main-recur}, as proved in Appendix~\ref{app:main-alg}.
 As stated in Theorem~\ref{thm:main-recur}, we are given an algorithm $\A$ that can solve any $(k-1)$-SJ problem in $F_{k-1}(n, \out)$ time, where $n$ and $\out$ are the input and output sizes, respectively.
@@
 
-\extraspacing {\bf Category 1.} Given an $i \in [k]$, we denote by $\J^\cat_i$ the set of $k$-tuples $\bm{t} = (r_1, ..., r_k) \in \J(R_1, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i]$.
-We will compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.
+\extraspacing {\bf Category 1.} Given an $i \in [k]$, we denote by $\J^\cat_i$ the set of $k$-tuples $\bm{t} = (r_1, ..., r_k) \in \J(R_1, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the top-left corner of $\bm{t}[i] = r_i$.
+We will show how to compute $\J^\cat_i$ for $i = k$; the set $\J^\cat_i$ of every other $i$ can be produced in the same manner.
 
@@
 Next, for each $j \in [k-1]$, solve the reporting version of Problem $\mathscr{A}$ by feeding $P'$ and $R_j$ as the input.
-Doing so produces, for each point $p \in P'$, the set $\contained_{R_j}(p)$, which --- as defined in \eqref{eqn:contained},  noticing that $p$ is a degenerated ``horizontal segment'' --- is the set of rectangles in $R_j$ covering $p$.
+This produces the set $\contained_{R_j}(p)$ for each point $p \in P'$, where $\contained_{R_j}(p)$ is defined in \eqref{eqn:contained} (treating $p$ as a degenerated ``horizontal segment'') and includes all rectangles of $R_j$ covering $p$.
 By the discussion in Section~\ref{sec:bricks}, the total cost of this step is bounded by

### 2024-05-27 msj.tex
yufei

@@
 
-We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, let $r \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple 
+We are ready to generate $\J^\cat_k$. Take any point $p \in P'$, and let $r \in R_k$ be the rectangle with $p$ as the top-left corner. For every $(k-1)$-tuple
 \myeqn{
-    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \label{eqn:ksj:cat1:help2}
+    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \nn
+    %\label{eqn:ksj:cat1:help2}
 }
@@
 Performing the above for all $p \in P'$ generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time.
-The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for all $j \in [k-1]$.
-Hence, we have $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|) = O(kn \log n + k \cdot \out)$ .
+The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for each $j \in [k-1]$.
+Hence, $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|) = O(kn \log n + k \cdot \out)$ .
 
@@
 A category-1 result tuple $\bm{t}$ may be seen more than once (this happens if the top-left corner of $B_\bm{t}$ is the top-left corner of more than one rectangle in $\bm{t}$).
-Duplicate removal can be achieved at no extra cost asymptotically, following the ideas explained in Section~\ref{sec:hv:type2}.
+Duplicate removal can be implemented at no extra cost asymptotically, following the ideas explained in Section~\ref{sec:hv:type2}.
 
@@
 
-The computation of $\J^\catt_{i,j}$ is an instance of H-V $k$-SJ.
+The computation of $\J^\catt_{i,j}$ is an instance of the H-V $k$-SJ problem.
 Specifically, collect the top-edges of all rectangles of $R_i$ into a set $H$, and collect the left-edges of all the rectangles of $R_j$ into a set $V$.
-This yields an H-V $k$-SJ problem whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$.
+This yields an H-V $k$-SJ instance whose input comprises all the $R_z$ with $z \in [k] \setminus \set{i,j}$,  $H$, and $V$.
 Each result tuple consists of a rectangle $r_z \in R_z$, for $z \in [k] \setminus \set{i,j}$, a horizontal segment $h \in H$, and a vertical segment $v \in V$ such that $h \cap v \cap \bigcap_{z \in [k] \setminus \set{i,j}} r_z \ne \emptyset$.
@@
 Converting the output into $\J^\catt_{i,j}$ takes another $O(|\J^\catt_{i,j}|)$ time.
-Applying $|\J^\catt_{i,j}| \le \out$, we can assert that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time
+Applying $|\J^\catt_{i,j}| \le \out$, we know that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time
 

### 2024-05-27 msj.tex
yufei

@@
 Formally, if $r_1, r_2, ..., r_z$ for some $z \ge 1$ are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
-Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3), (h_3: \emptyset}$.
+Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3), (h_3: \emptyset)}$.
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

### 2024-05-27 hard-3d.tex
hard-3d

The input to the 3-SJ problem in 3D space is three sets of axis-parallel boxes $S_1$, $S_2$ and $S_3$. 
% A box $s_i \in S_i$, for $i \in [3]$, is of the form $s_i = [x_1,x_2]\times[y_1,y_2]\times[z_1,z_2]$. 
The goal is to find the set of tuples $(s_1,s_2,s_3)$ such that $s_i \in S_i$ for $i \in [3]$ and $s_1 \cap s_2 \cap s_3 \neq \emptyset$. Denote the output set as $\J(S_1,S_2,S_3)$.
Let $n:=|S_1|+|S_2|+|S_3|$ be the input size, and let $\out = |\J(S_1,S_2,S_3)|$.

We will use the detection version of the 3-SJ problem in 3D space to solve the triangle detection problem defined as follows. Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph. We will convert this problem to an instance of the 3-SJ problem in 3D space with input size $12m=O(m)$. First, we represent each node in $G$ as a unique integer in $[|V|]$, and create three empty sets of 3D boxes $S_1$, $S_2$, $S_3$. For each edge $\set{u,v}\in E$, we add the following boxes to the sets $S_1$:
\begin{itemize}
    \item $[u,u]\times [v,v]\times(-\infty,u-1]$,
    \item $[u,u]\times [v,v]\times[u+1,+\infty)$,
    \item $[v,v]\times [u,u] \times (-\infty,v-1]$,
    \item $[v,v]\times [u,u] \times [v+1,+\infty)$;
\end{itemize}
and the following boxes to the sets $S_2$:
\begin{itemize}
    \item $(-\infty,v-1]\times[u,u]\times [v,v]$,
    \item $[v+1,+\infty)\times[u,u]\times [v,v]$,
    \item $(-\infty,u-1] \times[v,v]\times [u,u]$,
    \item $ [u+1,+\infty)\times [v,v]\times [u,u]$;
\end{itemize}
and the following boxes to the sets $S_3$:
\begin{itemize}
    \item $[u,u]\times[v+1,\infty)\times [v,v]$,
    \item $[u,u]\times(-\infty,v-1]\times [v,v]$,
    \item $[v,v]\times[u+1,+\infty)\times [u,u]$,
    \item $[v,v]\times(-\infty,u-1]\times [u,u]$.
\end{itemize}
There is a triangle in $G$ if and only if $\J(S_1,S_2,S_3) \neq \emptyset$. We prove as follows. If there is a triangle in $G$ with nodes $u$, $v$, $w$, where $u < v < w$, then the three boxes:
\begin{itemize}
    \item $[u,u]\times [v,v]\times[u+1, +\infty)\in S_1$,
    \item $(-\infty,w-1]\times[v,v]\times [w,w] \in S_2$,
    \item $[u,u]\times(-\infty, w-1]\times [w,w] \in S_3$,
\end{itemize}
will form a result tuple in $\J(S_1,S_2,S_3)$. Conversely, w.l.o.g., suppose there is a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t. $s_1 = [u,u]\times [v,v]\times[u+1, +\infty)$, $s_2 = (-\infty,w-1]\times[v,v]\times [w,w]$, and $s_3 = [u,u]\times(-\infty, w-1]\times [w,w]$. Then, we have $u < v < w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$. Therefore, if we can detect whethere there is an output in $O(n\polylog n)$ time for the 3-SJ problem in 3D space, then we can solve the triangle detection problem 
in $O(m \polylog m)$ time. However, the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.

### 2024-05-27 hard-3d.tex
hard-3d

@@
-The input to the 3-SJ problem in 3D space is three sets of axis-parallel boxes $S_1$, $S_2$ and $S_3$. 
+The input to the 3-SJ problem in 3D space comprises
+ three sets of axis-parallel boxes $S_1$, $S_2$, and $S_3$. 
 % A box $s_i \in S_i$, for $i \in [3]$, is of the form $s_i = [x_1,x_2]\times[y_1,y_2]\times[z_1,z_2]$. 
@@
 
-We will use the detection version of the 3-SJ problem in 3D space to solve the triangle detection problem defined as follows.
+The triangle detection problem is defined as follows.
 Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph.
@@
 \begin{itemize}
-    \item $(-\infty,v-1]\times[u,u]\times [v,v]$,
-    \item $[v+1,+\infty)\times[u,u]\times [v,v]$,
-    \item $(-\infty,u-1] \times[v,v]\times [u,u]$,
-    \item $ [u+1,+\infty)\times [v,v]\times [u,u]$;
+    \item $(-\infty,u-1]\times[u,u]\times [v,v]$,
+    \item $[u+1,+\infty)\times[u,u]\times [v,v]$,
+    \item $(-\infty,v-1] \times[v,v]\times [u,u]$,
+    \item $ [v+1,+\infty)\times [v,v]\times [u,u]$;
 \end{itemize}
@@
     \item $[u,u]\times [v,v]\times[u+1, +\infty)\in S_1$,
-    \item $(-\infty,w-1]\times[v,v]\times [w,w] \in S_2$,
+    \item $(-\infty,v-1]\times[v,v]\times [w,w] \in S_2$,
     \item $[u,u]\times(-\infty, w-1]\times [w,w] \in S_3$,
@@
 will form a result tuple in $\J(S_1,S_2,S_3)$.
-Conversely, w.l.o.g., suppose there is a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t.
-$s_1 = [u,u]\times [v,v]\times[u+1, +\infty)$, $s_2 = (-\infty,w-1]\times[v,v]\times [w,w]$, and $s_3 = [u,u]\times(-\infty, w-1]\times [w,w]$.
-Then, we have $u < v < w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
-Therefore, if we can detect whethere there is an output in $O(n\polylog n)$ time for the 3-SJ problem in 3D space, then we can solve the triangle detection problem
+Conversely, w.l.o.g., assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t.
+$s_1 = [v,v]\times [u,u]\times[v+1, +\infty)$, $s_2 = (-\infty,u-1]\times[u,u]\times [w,w]$, and $s_3 = [v,v]\times(-\infty, w-1]\times [w,w]$.
+We can reason about the other forms of result tuples analogously.
+From the assumption, we have $v< u< w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
+Therefore, if we can detect whether there is an output in $O(n\polylog n)$ time for the 3-SJ problem in 3D space, then we can solve the triangle detection problem
 in $O(m \polylog m)$ time. However, the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.

### 2024-05-27 hard-3d.tex
hard-3d

@@
 \end{itemize}
-There is a triangle in $G$ if and only if $\J(S_1,S_2,S_3) \neq \emptyset$. We prove as follows. If there is a triangle in $G$ with nodes $u$, $v$, $w$, where $u < v < w$, then the three boxes:
+Now we prove that there is a triangle in $G$ iff.
+ $\J(S_1,S_2,S_3) \neq \emptyset$. If there is a triangle in $G$ with nodes $u$, $v$, $w$  satisfying that $u < v < w$, then the three boxes:
 \begin{itemize}

### 2024-05-27 hard-3d.tex
hard-3d

@@
 Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph.
-We will convert this problem to an instance of the 3-SJ problem in 3D space with input size $12m=O(m)$.
+We will reduce this problem to an instance of the 3-SJ problem in 3D space with input size $12m=O(m)$.
 First, we represent each node in $G$ as a unique integer in $[|V|]$, and create three empty sets of 3D boxes $S_1$, $S_2$, $S_3$.
@@
 \begin{itemize}
-    \item $[u,u]\times[v+1,\infty)\times [v,v]$,
+    \item $[u,u]\times[v+1,+\infty)\times [v,v]$,
     \item $[u,u]\times(-\infty,v-1]\times [v,v]$,
@@
 We can reason about the other forms of result tuples analogously.
-From the assumption, we have $v< u< w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
+From the reduction, we have $v< u< w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
 Therefore, if we can detect whether there is an output in $O(n\polylog n)$ time for the 3-SJ problem in 3D space, then we can solve the triangle detection problem

### 2024-05-27 msj.tex
yufei

@@
 \extraspacing {\bf Set $\bm{H'}$.} Take any horizontal segment $h = [x_1, x_2] \times y \in H$.
-Let $p$ be the rightmost point on $h$ satisfying the condition that at least one rectangle in $R_{k-2}$ covers $p$.
+Let $p$ be the rightmost point on $h$ satisfying the condition that at least one left-end covering rectangle of $h$ in $R_{k-2}$ covers $p$.
 This $p$ exists if and only if $h$ has at least one left-end covering rectangle in $R_{k-2}$.
@@
 By Theorem~\ref{thm:main-recur}, in general, a comparison-based $(k-1)$-SJ algorithm with runtime $F_{k-1}(n, \out)$ spawns a comparison-based $k$-SJ algorithm whose running time $F_k(n, \out)$ obeys \eqref{eqn:main:reccurrence}.
-Specifically, for $k \ge 3$, there is a constant $c > 1$ such that
+Specifically, for $k \ge 3$, there is a constant $c \ge 2$ such that
 \myeqn{

### 2024-05-27 hard-3d.tex
hard-3d

@@
 From the reduction, we have $v< u< w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
-Therefore, if we can detect whether there is an output in $O(n\polylog n)$ time for the 3-SJ problem in 3D space, then we can solve the triangle detection problem
-in $O(m \polylog m)$ time. However, the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
+Hence, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem 
+in $O(m \polylog m)$ time.
+It is standard to convert an $O(n \log n + \out)$-time reporting algorthm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
+More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.
+We can simply run this algorithm for $cn\polylog n$ time and check whether there is an output.
+If the algorithm terminates and reports no output, then we declare that $\out = 0$.
+Otherwise, the algorithm either reports an output or runs for more than $cn\polylog n$ time.
+Either case indicates that there is at least one output for the 3-SJ problem in 3D space.

### 2024-05-27 hard-3d.tex
hard-3d

@@
 in $O(m \polylog m)$ time.
-It is standard to convert an $O(n \log n + \out)$-time reporting algorthm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
+Furthermore, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
 More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.
-We can simply run this algorithm for $cn\polylog n$ time and check whether there is an output.
+We can simply run this algorithm for $cn\polylog n$ time and check the output and whether it terminates.
 If the algorithm terminates and reports no output, then we declare that $\out = 0$.
 Otherwise, the algorithm either reports an output or runs for more than $cn\polylog n$ time.
-Either case indicates that there is at least one output for the 3-SJ problem in 3D space.
+Both cases indicate that there is at least one output for the 3-SJ problem in 3D space.
+The input to the 3-SJ problem in 3D space comprises
+ three sets of axis-parallel boxes $S_1$, $S_2$, and $S_3$. 
+% A box $s_i \in S_i$, for $i \in [3]$, is of the form $s_i = [x_1,x_2]\times[y_1,y_2]\times[z_1,z_2]$. 
+The goal is to find the set of tuples $(s_1,s_2,s_3)$ such that $s_i \in S_i$ for $i \in [3]$ and $s_1 \cap s_2 \cap s_3 \neq \emptyset$. Denote the output set as $\J(S_1,S_2,S_3)$.
+Let $n:=|S_1|+|S_2|+|S_3|$ be the input size, and let $\out = |\J(S_1,S_2,S_3)|$.
+
+The triangle detection problem is defined as follows.
+Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph.
+We will reduce this problem to an instance of the 3-SJ problem in 3D space with input size $12m=O(m)$.
+First, we represent each node in $G$ as a unique integer in $[|V|]$, and create three empty sets of 3D boxes $S_1$, $S_2$, $S_3$.
+For each edge $\set{u,v}\in E$, we add the following boxes to the sets $S_1$:
+\begin{itemize}
+    \item $[u,u]\times [v,v]\times(-\infty,u-1]$,
+    \item $[u,u]\times [v,v]\times[u+1,+\infty)$,
+    \item $[v,v]\times [u,u] \times (-\infty,v-1]$,
+    \item $[v,v]\times [u,u] \times [v+1,+\infty)$;
+\end{itemize}
+and the following boxes to the sets $S_2$:
+\begin{itemize}
+    \item $(-\infty,u-1]\times[u,u]\times [v,v]$,
+    \item $[u+1,+\infty)\times[u,u]\times [v,v]$,
+    \item $(-\infty,v-1] \times[v,v]\times [u,u]$,
+    \item $ [v+1,+\infty)\times [v,v]\times [u,u]$;
+\end{itemize}
+and the following boxes to the sets $S_3$:
+\begin{itemize}
+    \item $[u,u]\times[v+1,+\infty)\times [v,v]$,
+    \item $[u,u]\times(-\infty,v-1]\times [v,v]$,
+    \item $[v,v]\times[u+1,+\infty)\times [u,u]$,
+    \item $[v,v]\times(-\infty,u-1]\times [u,u]$.
+\end{itemize}
+Now we prove that there is a triangle in $G$ iff.
+ $\J(S_1,S_2,S_3) \neq \emptyset$. If there is a triangle in $G$ with nodes $u$, $v$, $w$ satisfying that $u < v < w$, then the three boxes:
+\begin{itemize}
+    \item $[u,u]\times [v,v]\times[u+1, +\infty)\in S_1$,
+    \item $(-\infty,v-1]\times[v,v]\times [w,w] \in S_2$,
+    \item $[u,u]\times(-\infty, w-1]\times [w,w] \in S_3$,
+\end{itemize}
+will form a result tuple in $\J(S_1,S_2,S_3)$.
+Conversely, w.l.o.g., assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t.
+$s_1 = [v,v]\times [u,u]\times[v+1, +\infty)$, $s_2 = (-\infty,u-1]\times[u,u]\times [w,w]$, and $s_3 = [v,v]\times(-\infty, w-1]\times [w,w]$.
+We can reason about the other forms of result tuples analogously.
+From the reduction, we have $v< u< w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
+Hence, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem 
+in $O(m \polylog m)$ time.
+Furthermore, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
+More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.
+We can simply run this algorithm for $cn\polylog n$ time and check the output and whether it terminates.
+If the algorithm terminates and reports no output, then we declare that $\out = 0$.
+Otherwise, the algorithm either reports an output or runs for more than $cn\polylog n$ time.
+Both cases indicate that there is at least one output for the 3-SJ problem in 3D space.

### 2024-05-27 msj.tex
yufei

@@
 
+%test
 %=================================
@@
 \extraspacing {\bf Set $\bm{H'}$.} Take any horizontal segment $h = [x_1, x_2] \times y \in H$.
-Let $p$ be the rightmost point on $h$ satisfying the condition that at least one left-end covering rectangle of $h$ in $R_{k-2}$ covers $p$.
+Recall from Section~\ref{sec:bricks} that a left-end covering rectangle of $h$ is a rectangle covering the left endpoint of $h$.
+Let $p$ be the rightmost point on $h$ such that at least one left-end covering rectangle of $h$ in $R_{k-2}$ covers $p$.
 This $p$ exists if and only if $h$ has at least one left-end covering rectangle in $R_{k-2}$.

### 2024-05-27 hard-3d.tex
hard-3d

@@
 will form a result tuple in $\J(S_1,S_2,S_3)$.
-Conversely, w.l.o.g., assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t.
+Conversely, assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t.
 $s_1 = [v,v]\times [u,u]\times[v+1, +\infty)$, $s_2 = (-\infty,u-1]\times[u,u]\times [w,w]$, and $s_3 = [v,v]\times(-\infty, w-1]\times [w,w]$.
@@
 We can simply run this algorithm for $cn\polylog n$ time and check the output and whether it terminates.
-If the algorithm terminates and reports no output, then we declare that $\out = 0$.
+If the algorithm terminates and reports nothing, then we declare that there is no output.
 Otherwise, the algorithm either reports an output or runs for more than $cn\polylog n$ time.
 Both cases indicate that there is at least one output for the 3-SJ problem in 3D space.
-The input to the 3-SJ problem in 3D space comprises
- three sets of axis-parallel boxes $S_1$, $S_2$, and $S_3$. 
-% A box $s_i \in S_i$, for $i \in [3]$, is of the form $s_i = [x_1,x_2]\times[y_1,y_2]\times[z_1,z_2]$. 
-The goal is to find the set of tuples $(s_1,s_2,s_3)$ such that $s_i \in S_i$ for $i \in [3]$ and $s_1 \cap s_2 \cap s_3 \neq \emptyset$. Denote the output set as $\J(S_1,S_2,S_3)$.
-Let $n:=|S_1|+|S_2|+|S_3|$ be the input size, and let $\out = |\J(S_1,S_2,S_3)|$.
-
-The triangle detection problem is defined as follows.
-Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph.
-We will reduce this problem to an instance of the 3-SJ problem in 3D space with input size $12m=O(m)$.
-First, we represent each node in $G$ as a unique integer in $[|V|]$, and create three empty sets of 3D boxes $S_1$, $S_2$, $S_3$.
-For each edge $\set{u,v}\in E$, we add the following boxes to the sets $S_1$:
-\begin{itemize}
-    \item $[u,u]\times [v,v]\times(-\infty,u-1]$,
-    \item $[u,u]\times [v,v]\times[u+1,+\infty)$,
-    \item $[v,v]\times [u,u] \times (-\infty,v-1]$,
-    \item $[v,v]\times [u,u] \times [v+1,+\infty)$;
-\end{itemize}
-and the following boxes to the sets $S_2$:
-\begin{itemize}
-    \item $(-\infty,u-1]\times[u,u]\times [v,v]$,
-    \item $[u+1,+\infty)\times[u,u]\times [v,v]$,
-    \item $(-\infty,v-1] \times[v,v]\times [u,u]$,
-    \item $ [v+1,+\infty)\times [v,v]\times [u,u]$;
-\end{itemize}
-and the following boxes to the sets $S_3$:
-\begin{itemize}
-    \item $[u,u]\times[v+1,+\infty)\times [v,v]$,
-    \item $[u,u]\times(-\infty,v-1]\times [v,v]$,
-    \item $[v,v]\times[u+1,+\infty)\times [u,u]$,
-    \item $[v,v]\times(-\infty,u-1]\times [u,u]$.
-\end{itemize}
-Now we prove that there is a triangle in $G$ iff.
- $\J(S_1,S_2,S_3) \neq \emptyset$. If there is a triangle in $G$ with nodes $u$, $v$, $w$ satisfying that $u < v < w$, then the three boxes:
-\begin{itemize}
-    \item $[u,u]\times [v,v]\times[u+1, +\infty)\in S_1$,
-    \item $(-\infty,v-1]\times[v,v]\times [w,w] \in S_2$,
-    \item $[u,u]\times(-\infty, w-1]\times [w,w] \in S_3$,
-\end{itemize}
-will form a result tuple in $\J(S_1,S_2,S_3)$.
-Conversely, w.l.o.g., assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t.
-$s_1 = [v,v]\times [u,u]\times[v+1, +\infty)$, $s_2 = (-\infty,u-1]\times[u,u]\times [w,w]$, and $s_3 = [v,v]\times(-\infty, w-1]\times [w,w]$.
-We can reason about the other forms of result tuples analogously.
-From the reduction, we have $v< u< w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
-Hence, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem 
-in $O(m \polylog m)$ time.
-Furthermore, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
-More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.
-We can simply run this algorithm for $cn\polylog n$ time and check the output and whether it terminates.
-If the algorithm terminates and reports no output, then we declare that $\out = 0$.
-Otherwise, the algorithm either reports an output or runs for more than $cn\polylog n$ time.
-Both cases indicate that there is at least one output for the 3-SJ problem in 3D space.

### 2024-05-27 msj.tex
yufei

@@
 
-%test
 %=================================
@@
 
+\begin{figure*}
+    \begin{tabular}{cccc}
+        \hspace{-5mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex1-a} &
+        \hspace{-4mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex1-b} & 
+        \hspace{-4mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex1-c} &
+        \hspace{-4mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex1-d} \\ 
+        \hspace{-5mm} (a) $R_1$, $R_2$, $H$, and $V$ &
+        \hspace{-4mm} (b) $R_1'$ and $R_2'$ &
+        \hspace{-4mm} (c) $R_1'$, $R_2'$, $H$, and $V$ &
+        \hspace{-4mm} (d) $R_1^*$, $R_2^*$ ($= \emptyset$), $H$, and $V$
+    \end{tabular}
+    \figcapup 
+    \caption{Finding H-V $k$-SJ result tuples of type 1 ($k = 4$)} 
+    \label{fig:hv:type1:ex}
+    \figcapdown 
+\end{figure*}
+
 \section{H-V $k$-SJ: Result Tuples of Type 1} \label{sec:hv:type1}
@@
 
+\extraspacing {\underline{\em Example.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
+The set $\J_1$ contains the following 4-tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
+\qed
+
 \extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r \in R_i$, we compute four segments:
 \myitems{
-    \item $h_1$ (resp., $h_2$): the lowest (resp., highest) segment in $H$ that crosses $r$;
+    \item $h_\bot$ (resp., $h_\top$): the lowest (resp., highest) segment in $H$ that crosses $r$;
 
-    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment in $V$ that crosses $r$.
+    \item $v_\vdash$ (resp., $v_\dashv$): the leftmost (resp., rightmost) segment in $V$ that crosses $r$.
 }
-Define $r' = [x_1, x_2] \times [y_1, y_2]$, where $x_1$ (resp., $x_2$) is x-coordinate of $v_1$ (resp., $v_2$), and $y_1$ (resp., $y_2$) is y-coordinate of $h_1$ (resp., $h_2$).
+Define $r' = [x_\vdash, x_\dashv] \times [y_\bot, y_\top]$, where $x_\vdash$ (resp., $x_\dashv$) is the x-coordinate of $v_\vdash$ (resp., $v_\dashv$), and $y_\bot$ (resp., $y_\top$) is the y-coordinate of $h_\bot$ (resp., $h_\top$).
 We say that $r'$ is the {\em trimmed rectangle} of $r$, and conversely, $r$ is the {\em full rectangle} of $r'$.
@@
 
-\yf{example}
+\extraspacing {\underline{\em Example.}} Figure~\ref{fig:hv:type1:ex}b shows the rectangles in $R'_1 = \set{\alpha'}$ and $R'_2$ $=$ $\{\beta'_1,$ $\beta'_2\}$.
+For instance, $\alpha'$, which is trimmed from rectangle $\alpha$, is decided by $h_\bot$ $=$ $h_2$, $h_\top$ $=$ $h_6$, $v_\vdash = v_3$, and $v_\dashv = v_5$.
+The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2, ..., R'_{k-2}) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
+\qed
+
+%\yf{example}
 
@@
 We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(kn \log n +  k \cdot \out)$ time.
+
+\begin{figure*}
+    \begin{tabular}{cccc}
+        \hspace{-3mm}
+        \includegraphics[height=43mm]{./artwork/alg-ex1} &
+        \hspace{-3mm}
+        \includegraphics[height=43mm]{./artwork/alg-ex1} & 
+        \hspace{-3mm}
+        \includegraphics[height=43mm]{./artwork/alg-ex1} &
+        \hspace{-3mm}
+        \includegraphics[height=43mm]{./artwork/alg-ex1}
+    \end{tabular}
+\end{figure*}
 

### 2024-05-27 msj.tex
yufei

@@
 \extraspacing {\underline{\em Example.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
-The set $\J_1$ contains the following 4-tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
+The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
 \qed
@@
 
-\yf{example}
+\noindent {\underline{\em Example.} Let us examine, in turn, the two 2-tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1', R_2')$.
+For $\bm{t}_1$, $B_{\bm{t}_1}$ is the rectangle \ttt{ABCD}, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$.
+Accordingly, $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_V(\bm{t}_1) = \set{v_3}$.
+For $\bm{t}_2$, $B_{\bm{t}_2}$ is the rectangle \ttt{AEFG}, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$.
+Accordingly, $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ and $\dcross_V(\bm{t}_1) = \set{v_3, v_4, v_5}$.
+\qed
 
+\vspace{1mm}
 \vgap
@@
 An analogous method can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ within the same time complexity.
+
+\noindent {\underline{\em Example.} In our running example, the $\J^*$ computed includes 10 tuples in total: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_4, v_5}$.
+All these 10 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (mentioned earlier).
+\qed
 

### 2024-05-27 msj.tex
yufei

@@
 \def\vgap{\vspace{0mm}}
-\def\extraspacing{\vspace{2mm} \noindent}
-\def\figcapup{\vspace{-0mm}}
-\def\figcapdown{\vspace{-0mm}}
+\def\extraspacing{\vspace{1mm} \noindent}
+\def\figcapup{\vspace{-2mm}}
+\def\figcapdown{\vspace{-4mm}}
 
@@
 
-\extraspacing {\underline{\em Example.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
+\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
 The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
@@
 
-\extraspacing {\underline{\em Example.}} Figure~\ref{fig:hv:type1:ex}b shows the rectangles in $R'_1 = \set{\alpha'}$ and $R'_2$ $=$ $\{\beta'_1,$ $\beta'_2\}$.
+\extraspacing {\underline{\em Example 4.2.}} Figure~\ref{fig:hv:type1:ex}b shows the rectangles in $R'_1 = \set{\alpha'}$ and $R'_2$ $=$ $\{\beta'_1,$ $\beta'_2\}$.
 For instance, $\alpha'$, which is trimmed from rectangle $\alpha$, is decided by $h_\bot$ $=$ $h_2$, $h_\top$ $=$ $h_6$, $v_\vdash = v_3$, and $v_\dashv = v_5$.
@@
 
-\noindent {\underline{\em Example.} Let us examine, in turn, the two 2-tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1', R_2')$.
+\noindent {\underline{\em Example 4.3.} Let us examine, in turn, the two 2-tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1', R_2')$.
 For $\bm{t}_1$, $B_{\bm{t}_1}$ is the rectangle \ttt{ABCD}, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$.
@@
 
-\noindent {\underline{\em Example.} In our running example, the $\J^*$ computed includes 10 tuples in total: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_4, v_5}$.
-All these 10 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (mentioned earlier).
+\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 10 tuples in total: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_4, v_5}$.
+All these 10 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
 \qed
@@
 
-\yf{example}
+\extraspacing {\underline{\em Example 4.4.}
 

### 2024-05-27 msj.tex
yufei

@@
 \noindent {\underline{\em Example 4.3.} Let us examine, in turn, the two 2-tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1', R_2')$.
-For $\bm{t}_1$, $B_{\bm{t}_1}$ is the rectangle \ttt{ABCD}, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$.
+For $\bm{t}_1$, $B_{\bm{t}_1}$ is the rectangle \ttt{ABCD} in Figure~\ref{fig:hv:type1:ex}c, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$.
 Accordingly, $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_V(\bm{t}_1) = \set{v_3}$.
@@
 
-\vspace{1mm}
 \vgap
@@
 
-\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 10 tuples in total: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_4, v_5}$.
+\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 10 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_4, v_5}$.
 All these 10 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
@@
 
-\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r' \in R_i'$, define
+\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. Define for each $r' \in R_i'$:
 \myeqn{
@@
 
-\vgap
+%\vgap
 
-\extraspacing {\underline{\em Example 4.4.}
+\extraspacing {\underline{\em Example 4.5.} The two tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1',$ $R_2')$ satisfy $\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$.
+Therefore, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}), \ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
+The top-sliced rectangle of $r'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
+Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
+\qed
 

### 2024-05-27 msj.tex
yufei

@@
         \hspace{-4mm} (c) $R_1'$, $R_2'$, $H$, and $V$ &
-        \hspace{-4mm} (d) $R_1^*$, $R_2^*$ ($= \emptyset$), $H$, and $V$
+        \hspace{-4mm} (d) $R_1^*$, $R_2^* = \emptyset$, and $H$ 
     \end{tabular}
@@
 
-\extraspacing {\underline{\em Example 4.5.} The two tuples $\bm{t}_1 = (\alpha', \beta_1')$ and $\bm{t}_2 = (\alpha', \beta_2')$ in $\J(R_1',$ $R_2')$ satisfy $\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$.
-Therefore, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}), \ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
+\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ serve as both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$.
+Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
 The top-sliced rectangle of $r'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
@@
 
-\yf{example}
+\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate.
+Recall that
+$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list.
+\qed
 

### 2024-05-27 msj.tex
yufei

@@
 
-\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ serve as both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$.
+\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ is both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$.
 Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
-The top-sliced rectangle of $r'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
+The top-sliced rectangle of $\alpha'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
 Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.

### 2024-05-28 msj.tex
yufei

@@
 }
-Computing the ``segment $h_1$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input).
-By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r \in R_i$.
+Computing the ``segment $h_\bot$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input).
+By symmetry, so is the computation of $h_\top, v_\vdash$, and $v_\dashv$ segments for each $r \in R_i$.
 It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(kn \log n)$ total time.
@@
 Accordingly, $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_V(\bm{t}_1) = \set{v_3}$.
-For $\bm{t}_2$, $B_{\bm{t}_2}$ is the rectangle \ttt{AEFG}, and $\gleft(\bm{t}_1) = \gbot(\bm{t}_1) = \alpha'$.
-Accordingly, $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ and $\dcross_V(\bm{t}_1) = \set{v_3, v_4, v_5}$.
+For $\bm{t}_2$, $B_{\bm{t}_2}$ is the rectangle \ttt{AEFG}, and $\gleft(\bm{t}_2) = \gbot(\bm{t}_2) = \alpha'$.
+Accordingly, $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ and $\dcross_V(\bm{t}_2) = \set{v_3, v_5}$.
 \qed
@@
 
-\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 10 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_4, v_5}$.
-All these 10 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
+\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 7 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_5}$.
+All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
 \qed

### 2024-05-28 hard-3d.tex
hard-3d

@@
 Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph.
-We will reduce this problem to an instance of the 3-SJ problem in 3D space with input size $12m=O(m)$.
+We will reduce this problem to an instance of the 3-SJ problem in 3D space with input size $6m=O(m)$.
 First, we represent each node in $G$ as a unique integer in $[|V|]$, and create three empty sets of 3D boxes $S_1$, $S_2$, $S_3$.
@@
 \begin{itemize}
-    \item $[u,u]\times [v,v]\times(-\infty,u-1]$,
-    \item $[u,u]\times [v,v]\times[u+1,+\infty)$,
-    \item $[v,v]\times [u,u] \times (-\infty,v-1]$,
-    \item $[v,v]\times [u,u] \times [v+1,+\infty)$;
+    \item $[u,u]\times [v,v]\times(-\infty,+\infty)$,
+    \item $[v,v]\times [u,u] \times (-\infty,+\infty)$,
 \end{itemize}
@@
 \begin{itemize}
-    \item $(-\infty,u-1]\times[u,u]\times [v,v]$,
-    \item $[u+1,+\infty)\times[u,u]\times [v,v]$,
-    \item $(-\infty,v-1] \times[v,v]\times [u,u]$,
-    \item $ [v+1,+\infty)\times [v,v]\times [u,u]$;
+    \item $(-\infty,+\infty)\times[u,u]\times [v,v]$,
+    \item $(-\infty,+\infty) \times[v,v]\times [u,u]$,
 \end{itemize}
@@
 \begin{itemize}
-    \item $[u,u]\times[v+1,+\infty)\times [v,v]$,
-    \item $[u,u]\times(-\infty,v-1]\times [v,v]$,
-    \item $[v,v]\times[u+1,+\infty)\times [u,u]$,
-    \item $[v,v]\times(-\infty,u-1]\times [u,u]$.
+    \item $[u,u]\times(-\infty,+\infty)\times [v,v]$,
+    \item $[v,v]\times(-\infty,+\infty)\times [u,u]$,
 \end{itemize}
-Now we prove that there is a triangle in $G$ iff.
- $\J(S_1,S_2,S_3) \neq \emptyset$. If there is a triangle in $G$ with nodes $u$, $v$, $w$ satisfying that $u < v < w$, then the three boxes:
+
+Now, we prove that there is a triangle in $G$ iff.
+ $\J(S_1,S_2,S_3) \neq \emptyset$. If there is a triangle in $G$ with nodes $u$, $v$, $w$, then the three boxes:
 \begin{itemize}
-    \item $[u,u]\times [v,v]\times[u+1, +\infty)\in S_1$,
-    \item $(-\infty,v-1]\times[v,v]\times [w,w] \in S_2$,
-    \item $[u,u]\times(-\infty, w-1]\times [w,w] \in S_3$,
+    \item $[u,u]\times [v,v]\times(-\infty,+\infty)\in S_1$,
+    \item $(-\infty,+\infty)\times[v,v]\times [w,w] \in S_2$,
+    \item $[u,u]\times(-\infty,+\infty)\times [w,w] \in S_3$,
 \end{itemize}
 will form a result tuple in $\J(S_1,S_2,S_3)$.
-Conversely, assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, s.t.
-$s_1 = [v,v]\times [u,u]\times[v+1, +\infty)$, $s_2 = (-\infty,u-1]\times[u,u]\times [w,w]$, and $s_3 = [v,v]\times(-\infty, w-1]\times [w,w]$.
-We can reason about the other forms of result tuples analogously.
-From the reduction, we have $v< u< w$, and the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
-Hence, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem 
+Conversely, assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, where $s_1 = [v,v]\times [u,u]\times(-\infty,+\infty)$, $s_2 = (-\infty,+\infty)\times[u,u]\times [w,w]$, and $s_3 = [v,v]\times(-\infty,+\infty)\times [w,w]$.
+From how $S_1$, $S_2$ and $S_3$ are constructed, the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
+Therefore, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem
 in $O(m \polylog m)$ time.
-Furthermore, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
+Besides, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
 More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.

### 2024-05-28 hard-3d.tex
hard-3d

@@
 % A box $s_i \in S_i$, for $i \in [3]$, is of the form $s_i = [x_1,x_2]\times[y_1,y_2]\times[z_1,z_2]$. 
-The goal is to find the set of tuples $(s_1,s_2,s_3)$ such that $s_i \in S_i$ for $i \in [3]$ and $s_1 \cap s_2 \cap s_3 \neq \emptyset$. Denote the output set as $\J(S_1,S_2,S_3)$.
+The goal is to report the set of tuples $(s_1,s_2,s_3)$ such that $s_i \in S_i$ for $i \in [3]$ and $s_1 \cap s_2 \cap s_3 \neq \emptyset$. Denote the output set as $\J(S_1,S_2,S_3)$.
 Let $n:=|S_1|+|S_2|+|S_3|$ be the input size, and let $\out = |\J(S_1,S_2,S_3)|$.
@@
 From how $S_1$, $S_2$ and $S_3$ are constructed, the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
-Therefore, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem
+Hence, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem
 in $O(m \polylog m)$ time.
-Besides, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
+Moreover, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
 More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.
@@
 Both cases indicate that there is at least one output for the 3-SJ problem in 3D space.
+Therefore, if the 3D version of the 3-SJ problem can be solved in $O(n\polylog n + \out)$ time, then the triangle detection problem can be solved in $O(m\polylog m)$ time.

### 2024-05-28 hard-3d.tex
hard-3d

@@
 The goal is to report the set of tuples $(s_1,s_2,s_3)$ such that $s_i \in S_i$ for $i \in [3]$ and $s_1 \cap s_2 \cap s_3 \neq \emptyset$. Denote the output set as $\J(S_1,S_2,S_3)$.
-Let $n:=|S_1|+|S_2|+|S_3|$ be the input size, and let $\out = |\J(S_1,S_2,S_3)|$.
+Let $n=|S_1|+|S_2|+|S_3|$ be the input size, and let $\out = |\J(S_1,S_2,S_3)|$.
 
 The triangle detection problem is defined as follows.
-Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph.
+Given a graph $G = (V,E)$, where $m=|E|$, we want to find out whether there exists a triangle in the graph.
 We will reduce this problem to an instance of the 3-SJ problem in 3D space with input size $6m=O(m)$.

### 2024-05-28 hard-3d.tex

@@
 in $O(m \polylog m)$ time.
-Moreover, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
+Moreover, it is standard to convert an $O(n \polylog n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
 More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.

### 2024-05-28 appB.tex
yufei

--- 2024-05-28 hard-3d.tex
@@
-The input to the 3-SJ problem in 3D space comprises
- three sets of axis-parallel boxes $S_1$, $S_2$, and $S_3$. 
-% A box $s_i \in S_i$, for $i \in [3]$, is of the form $s_i = [x_1,x_2]\times[y_1,y_2]\times[z_1,z_2]$. 
-The goal is to find the set of tuples $(s_1,s_2,s_3)$ such that $s_i \in S_i$ for $i \in [3]$ and $s_1 \cap s_2 \cap s_3 \neq \emptyset$. Denote the output set as $\J(S_1,S_2,S_3)$.
-Let $n:=|S_1|+|S_2|+|S_3|$ be the input size, and let $\out = |\J(S_1,S_2,S_3)|$.
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}
 
-The triangle detection problem is defined as follows.
-Given a graph $G = (V,E)$, where $m:=|E|$, we want to find out whether there exists a triangle in the graph.
-We will reduce this problem to an instance of the 3-SJ problem in 3D space with input size $6m=O(m)$.
-First, we represent each node in $G$ as a unique integer in $[|V|]$, and create three empty sets of 3D boxes $S_1$, $S_2$, $S_3$.
-For each edge $\set{u,v}\in E$, we add the following boxes to the sets $S_1$:
-\begin{itemize}
-    \item $[u,u]\times [v,v]\times(-\infty,+\infty)$,
-    \item $[v,v]\times [u,u] \times (-\infty,+\infty)$,
-\end{itemize}
-and the following boxes to the sets $S_2$:
-\begin{itemize}
-    \item $(-\infty,+\infty)\times[u,u]\times [v,v]$,
-    \item $(-\infty,+\infty) \times[v,v]\times [u,u]$,
-\end{itemize}
-and the following boxes to the sets $S_3$:
-\begin{itemize}
-    \item $[u,u]\times(-\infty,+\infty)\times [v,v]$,
-    \item $[v,v]\times(-\infty,+\infty)\times [u,u]$,
-\end{itemize}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
 
-Now, we prove that there is a triangle in $G$ iff.
- $\J(S_1,S_2,S_3) \neq \emptyset$. If there is a triangle in $G$ with nodes $u$, $v$, $w$, then the three boxes:
-\begin{itemize}
-    \item $[u,u]\times [v,v]\times(-\infty,+\infty)\in S_1$,
-    \item $(-\infty,+\infty)\times[v,v]\times [w,w] \in S_2$,
-    \item $[u,u]\times(-\infty,+\infty)\times [w,w] \in S_3$,
-\end{itemize}
-will form a result tuple in $\J(S_1,S_2,S_3)$.
-Conversely, assume the existence of a tuple $(s_1,s_2,s_3)\in \J(S_1,S_2,S_3)$, where $s_1 = [v,v]\times [u,u]\times(-\infty,+\infty)$, $s_2 = (-\infty,+\infty)\times[u,u]\times [w,w]$, and $s_3 = [v,v]\times(-\infty,+\infty)\times [w,w]$.
-From how $S_1$, $S_2$ and $S_3$ are constructed, the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ form a triangle in $G$.
-Therefore, if we can detect whether there is an output for the 3-SJ problem in 3D space in $O(n\polylog n)$ time, then we can solve the triangle detection problem
-in $O(m \polylog m)$ time.
-Besides, it is standard to convert an $O(n \log n + \out)$-time reporting algorithm for the 3-SJ problem in 3D space to an algorithm that detects whether there is an output for the same problem in $O(n \polylog n)$ time.
-More specifically, when $\out = 0$, the reporting algorithm runs in $cn\polylog n$ time for some constant $c$.
-We can simply run this algorithm for $cn\polylog n$ time and check the output and whether it terminates.
-If the algorithm terminates and reports nothing, then we declare that there is no output.
-Otherwise, the algorithm either reports an output or runs for more than $cn\polylog n$ time.
-Both cases indicate that there is at least one output for the 3-SJ problem in 3D space.
+\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}
+
+\extraspacing {\bf Computing $\bm{H^*}$.}
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}

### 2024-05-28 msj.tex
yufei

@@
         \hspace{-4mm} (c) $R_1'$, $R_2'$, $H$, and $V$ &
-        \hspace{-4mm} (d) $R_1^*$, $R_2^* = \emptyset$, and $H$ 
+        \hspace{-4mm} (d) $R_1^*$, $R_2^*$ ($= \emptyset$), and $H$
     \end{tabular}
@@
     \begin{tabular}{cccc}
-        \hspace{-3mm}
-        \includegraphics[height=43mm]{./artwork/alg-ex1} &
-        \hspace{-3mm}
-        \includegraphics[height=43mm]{./artwork/alg-ex1} & 
-        \hspace{-3mm}
-        \includegraphics[height=43mm]{./artwork/alg-ex1} &
-        \hspace{-3mm}
-        \includegraphics[height=43mm]{./artwork/alg-ex1}
+        \hspace{-5mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex2-a} &
+        \hspace{-4mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex2-b} &
+        \hspace{-4mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex2-c} &
+        \hspace{-4mm}
+        \includegraphics[height=45mm]{./artwork/alg-ex2-d} \\
+        \hspace{-5mm} (a) $R_1$, $R_2$, $H$, and $V$ &
+        \hspace{-4mm} (b) $R_1$, $H'$, and $V$ &
+        \hspace{-4mm} (c) $R_1$, $R_2$, $H'$, and $V$ &
+        \hspace{-4mm} (d) $R_2$ and $H^*$
     \end{tabular}
+    \figcapup
+    \caption{Finding H-V $k$-SJ result tuples of type 2 ($k = 4$)}
+    \label{fig:hv:type2:ex}
+    \figcapdown
 \end{figure*}
@@
 
+\extraspacing {\underline{\em Example 5.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$.
+The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_1, h_3, v_2)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_2)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$.
+\qed
+
 \extraspacing {\bf Set $\bm{H'}$.} Take any horizontal segment $h = [x_1, x_2] \times y \in H$.
@@
 
-\yf{example}
+\noindent {\underline{\em Example 5.2.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$.
+The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_1, h_3, v_2)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_2)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$.
+\qed
 
@@
 
-\begin{figure}
-    \begin{tabular}{cc}
-        \includegraphics[height=27mm]{./artwork/cat1} &
-        \hspace{5mm} \includegraphics[height=27mm]{./artwork/cat2} \\
-        (a) Category 1 &
-        (b) Category 2
-    \end{tabular}
-
-    \figcapup
-    \caption{Classifying $k$-SJ result tuples ($k$ = 4)}
-    \label{fig:ksj:cats}
-    \figcapdown
-\end{figure}
-
 \section{Settling $k$-SJ} \label{sec:ksj}
@@
 
+\begin{figure}
+    \begin{tabular}{cc}
+        \includegraphics[height=27mm]{./artwork/cat1} &
+        \hspace{5mm} \includegraphics[height=27mm]{./artwork/cat2} \\
+        (a) Category 1 &
+        (b) Category 2
+    \end{tabular}
+
+    \figcapup
+    \caption{Classifying $k$-SJ result tuples ($k$ = 4)}
+    \label{fig:ksj:cats}
+    \figcapdown
+\end{figure}
+
 \extraspacing {\bf Category 2.} Given $i, j \in [k]$ with $i \ne j$, we denote by $\J^\catt_{i,j}$ the set of $k$-tuples $\bm{t} = (r_1, ..., r_k) \in \J(R_1, ..., R_k)$ such that the top-left corner of $B_\bm{t}$ is the intersection between the top edge of $r_i$ and the left edge of $r_j$.
@@
 
+\input{appB.tex}
+
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}

### 2024-05-28 msj.tex
yufei

@@
 
-\noindent {\underline{\em Example 5.2.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$.
-The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_1, h_3, v_2)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_2)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$.
+\noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ and thus has no trimmed segment.
+Segment $h_2$ has $\beta_2$ and $\beta_3$ as its left-covering rectangles.
+As the entire $h_2$ is covered by $\beta_2$
 \qed

### 2024-05-28 appB.tex
lmm4.1

@@
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}
+ We will prove by associating each tuple $(r_1',...,r_{k-2}')\in \J(R_1',...,R_{k-2}')$ with a unique tuple in $\J(R_1,...,R_{k-2},H,V)$.
+More specifically, let $r_i$ be the full rectangle of $r_i'$ for $i\in[k-2]$.
+Let $r'_l = \gleft(\bm{t'})$ and $r'_b = \gbot(\bm{t'})$, where $l,b\in[k-2]$.
+Let $h$ be the segment $h_\bot$ for the rectangle $r_b$, and $v$ be the segment $v_\vdash$ for the rectangle $r_l$.
+The intersection of $h$ and $v$ is the bottom-left coner of the intersection area $\bigcap_{i = 1}^{k-2}r_i'$.
+Hence, $h\cap v\cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$ and the tuple $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
+The result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
+the full rectangle of $\bm{t_1'}[i]$ and $\bm{t_2'}[i]$ are different.
+Therefore, $|\J(R_1',...,R_{k-2}')| \leq |\J(R_1,...,R_{k-2})| = \out$.
 

### 2024-05-28 msj.tex
yufei

@@
 \noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ and thus has no trimmed segment.
-Segment $h_2$ has $\beta_2$ and $\beta_3$ as its left-covering rectangles.
-As the entire $h_2$ is covered by $\beta_2$
+Segment $h_2$ has one left-covering rectangle in $R_2$, which is $\beta_1$.
+As the entire $h_2$ is covered by $\beta_1$, it is equivalent to its trimmed segment $h_2'$.
+Segment $h_3$ has two left-covering rectangles in $R_2$, which are $\beta_2$ and $\beta_3$.
 \qed

### 2024-05-28 msj.tex
yufei

@@
 
-\noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ and thus has no trimmed segment.
+\noindent {\underline{\em Example 5.2.}} In Figure~\ref{fig:guard}a, segment $h_1$ has no left-covering rectangle in $R_2$ and thus has no trimmed segment.
 Segment $h_2$ has one left-covering rectangle in $R_2$, which is $\beta_1$.
-As the entire $h_2$ is covered by $\beta_1$, it is equivalent to its trimmed segment $h_2'$.
+As the entire $h_2$ is covered by $\beta_1$, it is equivalent to its trimmed segment $h_2'$; see Figure~\ref{fig:hv:type2:ex}b.
 Segment $h_3$ has two left-covering rectangles in $R_2$, which are $\beta_2$ and $\beta_3$.
+The right endpoint of its trimmed segment $h_3'$, as shown in Figure~\ref{fig:guard}b, is decided by the right edge of $\beta_3$.
+Therefore, $H' = \set{h_2', h_3'}$.
+It is clear from Figure~\ref{fig:guard}b that $\J(R_1, , H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
 \qed
@@
 The above should not be confused with \eqref{eqn:contained}, where the ``$\contained$'' function takes a segment as the parameter, rather than a tuple.
+
+\noindent {\underline{\em Example 5.3.}} For $\bm{t}_1 = (\alpha, h_2', v_1) \in \J(R_1, , H', V)$.
+Therefore, $H' = \set{h_2', h_3'}$.
+It is clear from Figure~\ref{fig:guard}b that $\J(R_1, , H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
 

### 2024-05-28 appB.tex
lmm4.1

@@
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}
- We will prove by associating each tuple $(r_1',...,r_{k-2}')\in \J(R_1',...,R_{k-2}')$ with a unique tuple in $\J(R_1,...,R_{k-2},H,V)$.
+ We will prove by associating each tuple $(r_1',...,r_{k-2}')\in \J(R_1',...,R_{k-2}')$ with a unique tuple in the H-V $k$-SJ result set $\J(R_1,...,R_{k-2},H,V)$.
 More specifically, let $r_i$ be the full rectangle of $r_i'$ for $i\in[k-2]$.
@@
 The result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
-the full rectangle of $\bm{t_1'}[i]$ and $\bm{t_2'}[i]$ are different.
+the full rectangles of $\bm{t_1'}[i]$ and $\bm{t_2'}[i]$ are different.
 Therefore, $|\J(R_1',...,R_{k-2}')| \leq |\J(R_1,...,R_{k-2})| = \out$.

### 2024-05-28 msj.tex
yufei

@@
         \hspace{-4mm} (b) $R_1$, $H'$, and $V$ &
-        \hspace{-4mm} (c) $R_1$, $R_2$, $H'$, and $V$ &
+        \hspace{-4mm} (c) $R_2$, $H'$, and $V$ &
         \hspace{-4mm} (d) $R_2$ and $H^*$
@@
 Therefore, $H' = \set{h_2', h_3'}$.
-It is clear from Figure~\ref{fig:guard}b that $\J(R_1, , H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
+It is clear from Figure~\ref{fig:guard}b that $\J(R_1, H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
 \qed
@@
 
-\noindent {\underline{\em Example 5.3.}} For $\bm{t}_1 = (\alpha, h_2', v_1) \in \J(R_1, , H', V)$.
-Therefore, $H' = \set{h_2', h_3'}$.
-It is clear from Figure~\ref{fig:guard}b that $\J(R_1, , H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
-
+\extraspacing {\underline{\em Example 5.3.}} Consider tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$ of $ \J(R_1, H', V)$ given in Example 5.2.
+For $\bm{t}_1 = (\alpha, h_2', v_1)$, its effective horizontal segment is $\texttt{DC}$ (see Figure~\ref{fig:hv:type2:ex}b.
+For $\bm{t}_2 = (\alpha, h_3', v_1)$, its effective horizontal segment is \texttt{EB}.
+For $\bm{t}_2 = (\alpha, h_3', v_2) \in \J(R_1, H', V)$, its effective horizontal segment is \texttt{EA}.
 \vgap

### 2024-05-28 4.1.tex
yufei

--- 2024-05-28 appB.tex
@@
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}
  We will prove by associating each tuple $(r_1',...,r_{k-2}')\in \J(R_1',...,R_{k-2}')$ with a unique tuple in the H-V $k$-SJ result set $\J(R_1,...,R_{k-2},H,V)$.
@@
 Therefore, $|\J(R_1',...,R_{k-2}')| \leq |\J(R_1,...,R_{k-2})| = \out$.
-
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
-
-\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}
-
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
-
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
-
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}
-
-\extraspacing {\bf Computing $\bm{H^*}$.}
-
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}

### 2024-05-28 msj.tex
yufei

@@
 
-\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 7 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_5}$.
-All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
-\qed
-
-\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. Define for each $r' \in R_i'$:
-\myeqn{
-    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
-}
-We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
-When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
-\myeqn{
-    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
-}
-and call it the {\em top-sliced rectangle} of $r'$.
-
-%\vgap
-
-\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ is both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$.
-Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
-The top-sliced rectangle of $\alpha'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
-Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
-\qed
-
-\vgap
-
-Next, we construct from $R_i'$ a new set of rectangles: 
-\myeqn{
-    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
-}
-In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
-
-\vgap
-
-Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
-The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.
-
-\begin{lemma} \label{lmm:hv:type1:cross-r*}
-    Both statements below are true: 
-    \myenums{
-        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
-
-        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
-Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
-Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
-Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
-    }
-\end{lemma}
-
-\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate.
-Recall that
-$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list.
-\qed
-
-Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$ (with $H$ and $R^*_i$ as the input).
-Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(k n \log n + \out)$.
-Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein have been sorted in ascending order of y-coordinate.
-
-\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
-First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time.
-Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
-For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time.
-Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
-This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
-Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type1:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
-
-\vgap
-
-We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(kn \log n +  k \cdot \out)$ time.
-
 \begin{figure*}
@@
 
+\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 7 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_5}$.
+All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
+\qed
+
+\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. Define for each $r' \in R_i'$:
+\myeqn{
+    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
+}
+We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
+When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
+\myeqn{
+    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
+}
+and call it the {\em top-sliced rectangle} of $r'$.
+
+%\vgap
+
+\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ is both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$.
+Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
+The top-sliced rectangle of $\alpha'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
+Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
+\qed
+
+\vgap
+
+Next, we construct from $R_i'$ a new set of rectangles: 
+\myeqn{
+    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
+}
+In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
+
+\vgap
+
+Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
+The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.
+
+\begin{lemma} \label{lmm:hv:type1:cross-r*}
+    Both statements below are true: 
+    \myenums{
+        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
+
+        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
+Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
+Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+    }
+\end{lemma}
+
+\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate.
+Recall that
+$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list.
+\qed
+
+Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$ (with $H$ and $R^*_i$ as the input).
+Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(k n \log n + \out)$.
+Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein have been sorted in ascending order of y-coordinate.
+
+\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
+First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time.
+Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
+For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time.
+Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
+This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type1:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
+
+\vgap
+
+We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(kn \log n +  k \cdot \out)$ time.
+
 \section{H-V $k$-SJ: Result Tuples of Type 2} \label{sec:hv:type2}
@@
 \extraspacing {\underline{\em Example 5.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$.
-The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_1, h_3, v_2)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_2)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$.
+The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$.
 \qed
@@
 
-\extraspacing {\underline{\em Example 5.3.}} Consider tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$ of $ \J(R_1, H', V)$ given in Example 5.2.
-For $\bm{t}_1 = (\alpha, h_2', v_1)$, its effective horizontal segment is $\texttt{DC}$ (see Figure~\ref{fig:hv:type2:ex}b.
+\extraspacing {\underline{\em Example 5.3.}} Consider the tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$ of $ \J(R_1, H', V)$ given in Example 5.2.
+For $\bm{t}_1 = (\alpha, h_2', v_1)$, its effective horizontal segment is $\texttt{DC}$ (see Figure~\ref{fig:hv:type2:ex}b).
 For $\bm{t}_2 = (\alpha, h_3', v_1)$, its effective horizontal segment is \texttt{EB}.
-For $\bm{t}_2 = (\alpha, h_3', v_2) \in \J(R_1, H', V)$, its effective horizontal segment is \texttt{EA}.
+For $\bm{t}_3 = (\alpha, h_3', v_2) \in \J(R_1, H', V)$, its effective horizontal segment is \texttt{EA}.
+Accordingly, as can be seen from Figure~\ref{fig:hv:type2:ex}c,
+$\contained_{R_{k-2}}(\bm{t}_1) = \contained_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contained_{R_2}(\bm{t}_2) = \set{\beta_3}$.
+\qed
+
 \vgap
@@
 assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+
+\extraspacing {\underline{\em Example 5.4.}} For $\bm{t}_1 = (\alpha, h_2', v_1)$, first find the full segment of $h_2'$, which is $h_2$.
+As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$.
+For $\bm{t}_2 = (\alpha, h_3', v_2)$, the full segment of $h_3'$ is $h_3$.
+As $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, Line 5 adds $(\alpha, \beta_1, h_3, v_2)$, $(\alpha, \beta_2, h_3, v_2)$, and $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
+Finally, the processing of $\tm{t}_3 = (\alpha, h_3', v_2)$ adds $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
+\qed
 
@@
 
-\input{appB.tex}
+\input{4.1.tex}
 
@@
 
+\input{4.2.tex}
+
 \extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}
 
+\input{4-comp.tex}
+
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
 
+\input{4.3.tex}
+
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
 
+\input{5.1.tex}
+
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}
 
+\input{5.2.tex}
+
 \extraspacing {\bf Computing $\bm{H^*}$.}
 
+\input{5-comp.tex}
+
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}
+
+\input{5.3.tex}
 

### 2024-05-28 4.1.tex
4.1

@@
- We will prove by associating each tuple $(r_1',...,r_{k-2}')\in \J(R_1',...,R_{k-2}')$ with a unique tuple in the H-V $k$-SJ result set $\J(R_1,...,R_{k-2},H,V)$.
-More specifically, let $r_i$ be the full rectangle of $r_i'$ for $i\in[k-2]$.
-Let $r'_l = \gleft(\bm{t'})$ and $r'_b = \gbot(\bm{t'})$, where $l,b\in[k-2]$.
+ We will prove by associating each tuple $\bm{t} \in \J(R_1',...,R_{k-2}')$ with a unique tuple in the H-V $k$-SJ result set $\J(R_1,...,R_{k-2},H,V)$.
+More specifically, let $r_i$ be the full rectangle of $\bm{t}[i]$ for $i\in[k-2]$.
+Let $\bm{t}[l] = \gleft(\bm{t})$ and $\bm{t}[b] = \gbot(\bm{t})$, where $l,b\in[k-2]$.
 Let $h$ be the segment $h_\bot$ for the rectangle $r_b$, and $v$ be the segment $v_\vdash$ for the rectangle $r_l$.
-The intersection of $h$ and $v$ is the bottom-left coner of the intersection area $\bigcap_{i = 1}^{k-2}r_i'$.
+The intersection of $h$ and $v$ is the bottom-left coner of the intersection area $\bigcap_{i = 1}^{k-2}\bm{t}[i]$.
 Hence, $h\cap v\cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$ and the tuple $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
-The result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
-the full rectangles of $\bm{t_1'}[i]$ and $\bm{t_2'}[i]$ are different.
+The result tuple is unique because for any two distinct tuples $\bm{t_1}$ and $\bm{t_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
+the full rectangles of $\bm{t_1}[i]$ and $\bm{t_2}[i]$ are different.
 Therefore, $|\J(R_1',...,R_{k-2}')| \leq |\J(R_1,...,R_{k-2})| = \out$.

### 2024-05-28 4.2.tex
4.2

First, we prove (1).

Next, we prove (2).

Finally, we prove (3).

### 2024-05-28 msj.tex
yufei

@@
 As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$.
-For $\bm{t}_2 = (\alpha, h_3', v_2)$, the full segment of $h_3'$ is $h_3$.
-As $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, Line 5 adds $(\alpha, \beta_1, h_3, v_2)$, $(\alpha, \beta_2, h_3, v_2)$, and $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
-Finally, the processing of $\tm{t}_3 = (\alpha, h_3', v_2)$ adds $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
+For $\bm{t}_2 = (\alpha, h_3', v_1)$, the full segment of $h_3'$ is $h_3$.
+As $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, Line 5 adds $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_1)$ to $\J_2$.
+Finally, the processing of $\bm{t}_3 = (\alpha, h_3', v_2)$ adds $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
 \qed
@@
 
-\vgap
-
-\yf{example}
+\extraspacing {\underline{\em Example 5.5.}} As mentioned $\J(R_1, H', V)$ has 3 tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$.
+Both $\bm{t}_2 = (\alpha, h_3', v_2)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$ have $h_3'$ as the horizontal segment.
+Therefore, $\minleft(h_3')$ equals the smaller between the x-coordinate of $v_1$ and that of $v_2$.
+The minimal segment $h_3^*$ of $h_3'$ is shown in Figure~\ref{fig:hv:type2:ex}(c).
+On the other hand, it is easy to verify that $\minleft(h_2')$ is the x-coordinate of $v_1$.
+The minimal segment $h_2^*$ of $h_2'$ is also shown in Figure~\ref{fig:hv:type2:ex}(c).
+\qed
 

### 2024-05-28 msj.tex
yufei

@@
 
+\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 7 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_5}$.
+All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
+\qed
+
+\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. Define for each $r' \in R_i'$:
+\myeqn{
+    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
+}
+We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
+When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
+\myeqn{
+    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
+}
+and call it the {\em top-sliced rectangle} of $r'$.
+
+%\vgap
+
+\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ is both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$.
+Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
+The top-sliced rectangle of $\alpha'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
+Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
+\qed
+
+\vgap
+
+Next, we construct from $R_i'$ a new set of rectangles: 
+\myeqn{
+    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
+}
+In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
+
+\vgap
+
+Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
+The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.
+
+\begin{lemma} \label{lmm:hv:type1:cross-r*}
+    Both statements below are true: 
+    \myenums{
+        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
+
+        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
+Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
+Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+    }
+\end{lemma}
+
+\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate.
+Recall that
+$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list.
+\qed
+
+Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$ (with $H$ and $R^*_i$ as the input).
+Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(k n \log n + \out)$.
+Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein have been sorted in ascending order of y-coordinate.
+
+\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
+First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time.
+Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
+For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time.
+Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
+This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
+Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type1:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
+
+\vgap
+
+We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(kn \log n +  k \cdot \out)$ time.
+
 \begin{figure*}
@@
 
-\extraspacing {\underline{\em Example 4.4.} In our running example, the $\J^*$ computed includes 7 tuples: $(\alpha, \beta_1, h_2, v_3)$, and $\{\alpha\} \times \{\beta_2\} \times \set{h_2, h_4, h_5} \times \set{v_3, v_5}$.
-All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
-\qed
-
-\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. Define for each $r' \in R_i'$:
-\myeqn{
-    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
-}
-We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
-When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
-\myeqn{
-    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
-}
-and call it the {\em top-sliced rectangle} of $r'$.
-
-%\vgap
-
-\extraspacing {\underline{\em Example 4.5.} Recall from Example 4.3 that rectangle $\alpha'$ is both $\gbot(\bm{t}_1)$ and $\gbot(\bm{t}_2)$.
-Thus, $\maxtop(\alpha') = \max\{\ytop(B_{\bm{t}_1}),$ $\ytop(B_{\bm{t}_2})\} = \max\{6, 9.5\} = 9.5$.
-The top-sliced rectangle of $\alpha'$ is the rectangle $\alpha^*$ in Figure~\ref{fig:hv:type1:ex}d.
-Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
-\qed
-
-\vgap
-
-Next, we construct from $R_i'$ a new set of rectangles: 
-\myeqn{
-    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
-}
-In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
-
-\vgap
-
-Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
-The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.
-
-\begin{lemma} \label{lmm:hv:type1:cross-r*}
-    Both statements below are true: 
-    \myenums{
-        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
-
-        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
-Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$.
-Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
-Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
-    }
-\end{lemma}
-
-\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate.
-Recall that
-$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list.
-\qed
-
-Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$ (with $H$ and $R^*_i$ as the input).
-Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*}, as well as the discussion in Section~\ref{sec:bricks}, assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(k n \log n + \out)$.
-Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein have been sorted in ascending order of y-coordinate.
-
-\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}.
-First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time.
-Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates.
-For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time.
-Abort the scan as soon as $h \notin \dcross_H(\bm{t})$.
-This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time.
-Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type1:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
-
-\vgap
-
-We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(kn \log n +  k \cdot \out)$ time.
-
 \section{H-V $k$-SJ: Result Tuples of Type 2} \label{sec:hv:type2}
@@
 Accordingly, as can be seen from Figure~\ref{fig:hv:type2:ex}c,
-$\contained_{R_{k-2}}(\bm{t}_1) = \contained_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contained_{R_2}(\bm{t}_2) = \set{\beta_3}$.
+$\contained_{R_{k-2}}(\bm{t}_1) = \contained_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contained_{R_2}(\bm{t}_3) = \set{\beta_3}$.
 \qed
@@
 
-\yf{example}
+\extraspacing {\underline{\em Example 5.6.}} It is clear from Figure~\ref{fig:hv:type2:ex}(d) that $\contained_{R_2}(h_3^*)$ has rectangles $\beta_3, \beta_2$, $\beta_1$, sorted in descending order of their right boundaries' x-coordinates.
+Consider $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$.
+Segment $h_3^*$ is the minimal segment of $h_3'$.
+As stated in Lemma~\ref{lmm:hv:type2:contained-h*}, both $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$ and $\contained_{R_2}(\bm{t}_3) = \set{\beta_1}$ are prefixes of the sorted order of $\contained_{R_2}(h_3^*)$.
+Regarding $h_2^*$, it is the minimal segment of $h_2'$, and $\contained_{R_2}(h_2^*)$ contains only  $\beta_1$.
+For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contained_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contained_{R_2}(h_2^*)$, as is consistent with the lemma.
+\qed
 

### 2024-05-28 msj.tex
yufei

@@
 Recall that
-$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$. Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list.
+$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$.
+Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list, as is consistent with Lemma~\ref{lmm:hv:type1:cross-r*}.
 \qed
@@
 
-\noindent {\underline{\em Example 5.2.}} In Figure~\ref{fig:guard}a, segment $h_1$ has no left-covering rectangle in $R_2$ and thus has no trimmed segment.
+\noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ (see Figure~\ref{fig:guard}a) and thus has no trimmed segment.
 Segment $h_2$ has one left-covering rectangle in $R_2$, which is $\beta_1$.
@@
 \extraspacing {\underline{\em Example 5.4.}} For $\bm{t}_1 = (\alpha, h_2', v_1)$, first find the full segment of $h_2'$, which is $h_2$.
-As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$.
+As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 of the algorithm adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$.
 For $\bm{t}_2 = (\alpha, h_3', v_1)$, the full segment of $h_3'$ is $h_3$.
@@
 
-\extraspacing {\underline{\em Example 5.5.}} As mentioned $\J(R_1, H', V)$ has 3 tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$.
-Both $\bm{t}_2 = (\alpha, h_3', v_2)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$ have $h_3'$ as the horizontal segment.
+\extraspacing {\underline{\em Example 5.5.}} As mentioned, $\J(R_1, H', V)$ has 3 tuples $\bm{t}_1$, $\bm{t}_2$, and $\bm{t}_3$.
+Both $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$ have $h_3'$ as the horizontal segment.
 Therefore, $\minleft(h_3')$ equals the smaller between the x-coordinate of $v_1$ and that of $v_2$.

### 2024-05-28 4.2.tex
4.2

@@
 First, we prove (1).
+By the definition of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$, $h$ and $v$ crosses $B_\bm{t}$.
+Thus, the intersection of $h$ and $v$ is a point in $B_\bm{t}$.
+Since $B_\bm{t}$ is the intersection of $\bm{t}[1],...,\bm{t}[k-2]$, the intersection of $h$ and $v$ lies in all the rectangles $r_1,...,r_{k-2}$.
+Therefore, $h\cap v\bigcap_{i = 1}^{k-2}r_i \neq \emptyset$, and  the tuple $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
 
-Next, we prove (2).
-
+Next, we prove (2). Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both $h$
+ and $v$ cross the rectangles $r_1,...,r_{k-2}$.
+Hence, for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(\xright(r_i'))]$.
 Finally, we prove (3).

### 2024-05-28 4.2.tex
4.2

@@
  and $v$ cross the rectangles $r_1,...,r_{k-2}$.
-Hence, for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(\xright(r_i'))]$.
+Hence, for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
 Finally, we prove (3).

### 2024-05-28 4.2.tex
4.2

@@
 Hence, for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
+Combining the two facts, $h$ and $v$ cross the rectangles $r_i'$, for $i\in[k-2]$.
+Hence, the point $h \cap v$ is in $B_\bm{t}$, and $h \cap v$ crosses $B_{\bm{t}}$.
+
+ The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+
 Finally, we prove (3).

### 2024-05-28 4.2.tex
4.2

@@
 Next, we prove (2). Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both $h$
- and $v$ cross the rectangles $r_1,...,r_{k-2}$.
-Hence, for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
-Combining the two facts, $h$ and $v$ cross the rectangles $r_i'$, for $i\in[k-2]$.
+ and $v$ cross the rectangles $r_1,...,r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
+Additionally, for each $i\in[k-2]$, the x-range (resp.
+y-range) of $r_i'$ is a subset of the x-range (resp.
+y-range) of $r_i$, so $h$ (resp.
+$v$) crosses the rectangles $r_i'$.
 Hence, the point $h \cap v$ is in $B_\bm{t}$, and $h \cap v$ crosses $B_{\bm{t}}$.
-
 The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.

### 2024-05-28 4.2.tex
4.2

@@
 Finally, we prove (3). 
+For each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $h \in \dcross_H(\bm{t'})$, we will identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
+Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $v$ be the segment $v_\vdash$ of $\gleft(\bm{t'})$.
+Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
+Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
+Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$. As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
+
+Similarly, for each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $v \in \dcross_V(\bm{t'})$, we can identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
+Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $h$ be the segment $h_\bot$ of $\gbot(\bm{t'})$.
+Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
+Since $v$ crosses $B_{\bm{t'}}$, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
+As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq \out$.

### 2024-05-28 msj.tex
yufei

@@
 Interestingly, the answer is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory.
-Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O(n \polylog n + \out)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
+Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O((n + \out) \cdot \polylog n)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
 This reduction can be inferred from an argument in \cite{kcko22} used to prove a more generic result.
@@
 
-\noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ (see Figure~\ref{fig:guard}a) and thus has no trimmed segment.
+\noindent {\underline{\em Example 5.2.}} Segment $h_1$ has no left-covering rectangle in $R_2$ (see Figure~\ref{fig:hv:type2:ex}a) and thus has no trimmed segment.
 Segment $h_2$ has one left-covering rectangle in $R_2$, which is $\beta_1$.
@@
 Segment $h_3$ has two left-covering rectangles in $R_2$, which are $\beta_2$ and $\beta_3$.
-The right endpoint of its trimmed segment $h_3'$, as shown in Figure~\ref{fig:guard}b, is decided by the right edge of $\beta_3$.
+The right endpoint of its trimmed segment $h_3'$, as shown in Figure~\ref{fig:hv:type2:ex}b, is decided by the right edge of $\beta_3$.
 Therefore, $H' = \set{h_2', h_3'}$.
-It is clear from Figure~\ref{fig:guard}b that $\J(R_1, H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
+It is clear from Figure~\ref{fig:hv:type2:ex}b that $\J(R_1, H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
 \qed
@@
 
-\input{hard-3d}
+An axis-parallel rectangle in 3D has the form $r = [x_1, x_2] \times [y_1, y_2] \times [z_1, z_2]$.
+We will refer to $[x_1, x_2]$ as the {\em x-projection} of $r$, and define its {\em y-} and {\em z-projections} analogously.
+
+\vgap
+
+In the 3D 3-SJ problem, the input comprises three sets of axis-parallel rectangles in $\real^3$: $R_1, R_2$, and $R_3$.
+The goal is to report all 3-tuples $(r_1, r_2, r_3) \in R_1 \times R_2 \times R_3$ such that $r_1 \cap r_2 \cap r_3 \neq \emptyset$.
+Denote by $\J(R_1, R_2, R_3)$ the set of those 3-tuples.
+Define the input size as $n = |R_1| + |R_2| + |R_3|$ and the output size as $\out = |\J(R_1, R_2, R_3)|$.
+
+\vgap
+
+In the {\em triangle detection}, we are given an undirected graph $G = (V, E)$ and need to determine whether $G$ has a triangle (a.k.a.\, 3-clique).
+Set $m = |E|$.
+We consider that $G$ has no isolated vertices (namely, vertices with degree 0), and therefore $|V| \le 2m$.
+
+\vgap
+
+Next, we will prove that if the 3D 3-SJ problem can be solved in $O((n + \out) \cdot \polylog n)$ time, then the triangle detection problem can be solved in $O(m \polylog m)$ time.
+This would be truly surprising because as mentioned in Section~\ref{sec:intro:ours} the state of the art algorithm for triangle detection runs in $O(m^{1.41})$ time \cite{ayz97}.
+Thus, in the absence of such a breakthrough, no $O((n + \out) \cdot \polylog n)$ time algorithms can exist for the 3D 3-SJ problem.
+
+\vgap
+
+Given a graph $G = (V,E)$ for triangle detection, we will construct an instance of the 3D 3-SJ problem with input size $3m$.
+W.l.o.g., let us assume that each vertex of $V$ is represented as a unique integer in $[|V|]$.
+Initialize 3 empty sets of rectangles: $R_1, R_2$, and $R_3$.
+For each edge $\set{u, v}\in E$ where $u < v$, we add rectangles
+\begin{itemize}
+    \item $(-\infty, \infty) \times [u,u] \times [v,v]$ to $R_1$;
+    \item $[v,v] \times(-\infty, \infty)\times [u,u]$ to $R_2$;
+    \item $[v,v] \times [u,u] \times(-\infty, \infty) $ to $R_3$.
+\end{itemize}
+Note that every rectangle in $R_1$ (resp., $R_2$ and $R_3$) has $(-\infty, \infty)$ as the x- (resp., y- and z-) projection.
+
+\vgap
+
+The construction has the property that $G$ has a triangle if and only if $\J(R_1, R_2, R_3) \neq \emptyset$. We can prove this with the following argument.
+\myitems{
+    \item Suppose that $G$ with triangle with vertices $u$, $v$, and $w$ such that $u < v  < w$.
+Then, by our construction, $r_1 = (-\infty, \infty) \times [u,u] \times [v,v] \in R_1$, $r_2 = [w, w] \times (-\infty, \infty) \times[v,v] \in R_2$, and $r_3 = [w,w] \times [u, u] \times (-\infty, \infty) \in r_3$.
+It is clear that $(r_1, r_2, r_3)$ is a result tuple in $\J(R_1, R_2, R_3)$.
+
+    \item Conversely, consider that $\J(R_1, R_2, R_3)$ is non-empty.
+Consider an arbitrary result tuple $(r_1, r_2, r_3) \in \J(R_1, R_2, R_3)$.
+Assume, w.l.o.g., that $r_1 = (-\infty, \infty) \times [u, u] \times [v, v]$.
+    Because the z-projection of $r_2$ must match that of $r_1$, we assert that $r_2$ must have the form $[w, w] \times (-\infty, \infty) \times [v, v]$ where $w > u$.
+Because the x-projection of $r_3$ must match that of $r_2$ and the y-projection of $r_3$ must match that of $r_1$, it follows that $r_3$ must have the form $[w, w] \times [u, u] \times (-\infty, \infty)$.
+This means that the edges $\set{u,v}$, $\set{v,w}$, and $\set{u,w}$ must all exist in $G$, and thus form a triangle.
+}
+
+Now, assume that there exists an algorithm $\A$ capable of solving the 3D 3-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
+For $\out = 0$, this algorithm must perform at most $c \cdot n \polylog n$ steps when executed on an input with input size $n$.
+We run $\A$ on the 3D 3-SJ instance $R_1, R_2, R_3$ constructed earlier in a {\em cost-monitoring manner}:
+\myitems{
+    \item If $\A$ terminates within $c \cdot (3m) \polylog (3m)$ steps, we check whether it finds any result tuple in $\J(R_1, R_2, R_3)$.
+If so, a triangle has been found in $G$; otherwise, we declare that $G$ has no triangles.
+
+    \item If $\A$ has performed $1 + c \cdot (3m) \polylog (3m)$ steps, we manually terminate the algorithm and declare that $\J(R_1, R_2, R_3) \ne \empty$, meaning that $G$ must have at least one triangle.
+}
+The above strategy thus settles the triangle detection problem in $O(m\polylog m)$ time.
 

### 2024-05-28 msj.tex
yufei

@@
 
-Next, we will prove that if the 3D 3-SJ problem can be solved in $O((n + \out) \cdot \polylog n)$ time, then the triangle detection problem can be solved in $O(m \polylog m)$ time.
-This would be truly surprising because as mentioned in Section~\ref{sec:intro:ours} the state of the art algorithm for triangle detection runs in $O(m^{1.41})$ time \cite{ayz97}.
+The subsequent discussion will show that if the 3D 3-SJ problem can be solved in $O((n + \out) \cdot \polylog n)$ time, then the triangle detection problem can be solved in $O(m \polylog m)$ time.
+This would be truly surprising because as mentioned in Section~\ref{sec:intro:ours} the state-of-the-art algorithm for triangle detection runs in $O(m^{1.41})$ time \cite{ayz97}.
 Thus, in the absence of such a breakthrough, no $O((n + \out) \cdot \polylog n)$ time algorithms can exist for the 3D 3-SJ problem.
@@
 W.l.o.g., let us assume that each vertex of $V$ is represented as a unique integer in $[|V|]$.
-Initialize 3 empty sets of rectangles: $R_1, R_2$, and $R_3$.
-For each edge $\set{u, v}\in E$ where $u < v$, we add rectangles
+Initialize $R_1, R_2$, and $R_3$ as 3 empty sets of rectangles.
+For each edge $\set{u, v}\in E$ where $u < v$, we add
 \begin{itemize}
-    \item $(-\infty, \infty) \times [u,u] \times [v,v]$ to $R_1$;
+    \item a rectangle $(-\infty, \infty) \times [u,u] \times [v,v]$ to $R_1$;
     \item $[v,v] \times(-\infty, \infty)\times [u,u]$ to $R_2$;
@@
 \myitems{
-    \item Suppose that $G$ with triangle with vertices $u$, $v$, and $w$ such that $u < v  < w$.
+    \item Suppose that $G$ has a triangle with vertices $u$, $v$, and $w$ such that $u < v  < w$.
 Then, by our construction, $r_1 = (-\infty, \infty) \times [u,u] \times [v,v] \in R_1$, $r_2 = [w, w] \times (-\infty, \infty) \times[v,v] \in R_2$, and $r_3 = [w,w] \times [u, u] \times (-\infty, \infty) \in r_3$.
@@
 Now, assume that there exists an algorithm $\A$ capable of solving the 3D 3-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
-For $\out = 0$, this algorithm must perform at most $c \cdot n \polylog n$ steps when executed on an input with input size $n$.
+For $\out = 0$, this algorithm must perform at most $c \cdot n \polylog n$ steps when the input size is $n$.
 We run $\A$ on the 3D 3-SJ instance $R_1, R_2, R_3$ constructed earlier in a {\em cost-monitoring manner}:
 \myitems{
-    \item If $\A$ terminates within $c \cdot (3m) \polylog (3m)$ steps, we check whether it finds any result tuple in $\J(R_1, R_2, R_3)$.
+    \item If $\A$ terminates within $c \cdot (3m) \polylog (3m)$ steps, we check whether it has output any result tuple in $\J(R_1, R_2, R_3)$.
 If so, a triangle has been found in $G$; otherwise, we declare that $G$ has no triangles.
 
-    \item If $\A$ has performed $1 + c \cdot (3m) \polylog (3m)$ steps, we manually terminate the algorithm and declare that $\J(R_1, R_2, R_3) \ne \empty$, meaning that $G$ must have at least one triangle.
+    \item If $\A$ has performed $1 + c \cdot (3m) \polylog (3m)$ steps, we manually terminate the algorithm and declare that $\J(R_1, R_2, R_3) \ne \emptyset$, meaning that $G$ must have at least one triangle.
 }

### 2024-05-28 msj.tex
yufei

@@
 \def\vgap{\vspace{0mm}}
-\def\extraspacing{\vspace{2mm} \noindent}
+\def\extraspacing{\vspace{1.5mm} \noindent}
 \def\figcapup{\vspace{-2mm}}
-\def\figcapdown{\vspace{-4mm}}
+\def\figcapdown{\vspace{-2mm}}
 
@@
 
-In the {\em triangle detection}, we are given an undirected graph $G = (V, E)$ and need to determine whether $G$ has a triangle (a.k.a.\, 3-clique).
+In the {\em triangle detection} problem, we are given an undirected graph $G = (V, E)$ and need to determine whether $G$ has a triangle (a.k.a.\, 3-clique).
 Set $m = |E|$.

### 2024-05-28 4-comp.tex
4-comp

We will first compute $\maxtop(r')$ for each rectangle $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$ by scanning the set $\J(R_1',...,R_{k-2}')$. Initially, we set $\maxtop(r') = -\infty$ for each $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$. For each tuple $\bm{t} = (r_1',...,r_{k-2}') \in \J(R_1',...,R_{k-2}')$, we update $\maxtop(\gbot(\bm{t}))$ to the maximum between its current value and $\ytop(B_{\bm{t}})$. Equipped with the values of $\maxtop(r')$ for each $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$, we can construct the set $R_i^*$ for each $i \in [k-2]$ by collecting the top-sliced rectangles for all rectangles in $R_i'$.

### 2024-05-28 4-comp.tex
4-comp

@@
 Initially, we set $\maxtop(r') = -\infty$ for each $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$.
-For each tuple $\bm{t} = (r_1',...,r_{k-2}') \in \J(R_1',...,R_{k-2}')$, we update $\maxtop(\gbot(\bm{t}))$ to the maximum between its current value and $\ytop(B_{\bm{t}})$.
-Equipped with the values of $\maxtop(r')$ for each $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$, we can construct the set $R_i^*$ for each $i \in [k-2]$ by collecting the top-sliced rectangles for all rectangles in $R_i'$.
+For each tuple $\bm{t} = (r_1',...,r_{k-2}') \in \J(R_1',...,R_{k-2}')$, we identify the rectangle $\gbot(\bm{t})$ in $O(k)$ time and
+ update $\maxtop(\gbot(\bm{t}))$ to the maximum between its current value and $\ytop(B_{\bm{t}})$. The time complexity so far is $O(n + k\cdot \out)$.
+Then, we can construct the set $R_i^*$ for every $i \in [k-2]$ by collecting the top-sliced rectangles of all rectangles $r_i' \in R_i'$ satisfying that $\maxtop(r_i')\neq -\infty$.
+This step takes $O(n)$ additional time.
+The overall time complexity is $O(n + k\cdot \out)$.

### 2024-05-28 4-comp.tex
4-comp

@@
 Initially, we set $\maxtop(r') = -\infty$ for each $r' \in \bigcup_{i = 1}^{k-2}R_{i}'$.
-For each tuple $\bm{t} = (r_1',...,r_{k-2}') \in \J(R_1',...,R_{k-2}')$, we identify the rectangle $\gbot(\bm{t})$ in $O(k)$ time and
+Then, for each tuple $\bm{t} = (r_1',...,r_{k-2}') \in \J(R_1',...,R_{k-2}')$, we identify the rectangle $\gbot(\bm{t})$ in $O(k)$ time and
  update $\maxtop(\gbot(\bm{t}))$ to the maximum between its current value and $\ytop(B_{\bm{t}})$. The time complexity so far is $O(n + k\cdot \out)$.
-Then, we can construct the set $R_i^*$ for every $i \in [k-2]$ by collecting the top-sliced rectangles of all rectangles $r_i' \in R_i'$ satisfying that $\maxtop(r_i')\neq -\infty$.
+Finally, we can construct the set $R_i^*$ for every $i \in [k-2]$ by collecting the top-sliced rectangles of all rectangles $r_i' \in R_i'$ satisfying that $\maxtop(r_i')\neq -\infty$.
 This step takes $O(n)$ additional time.

### 2024-05-29 4.3.tex
4-comp

First, we prove (1). For any $i \in [k-2]$, $r_i^*\in R_i^*$, and $h \in \cross_H(r_i^*)$, we will identify a unique tuple $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$ such that $\bm{t}[k-1] = h$, and $\bm{t}[i]$ is the full rectangle of $r_i'$, whose top-slice is $r_i^*$. Let $\maxtop(r_i') = \ytop(B_\bm{t'})$ for some tuple $\bm{t'} \in \J(R_1',...,R_{k-2}')$. Let $\bm{t'} = (r'_1,...,r'_{k-2})$, and let $r_i$ be the full rectangle of $r_i'$ for $i \in [k-2]$. Since $h$ crosses $r_i^*$, and $B_\bm{t'} \subseteq r_i^*$, $h$ crosses $B_{\bm{t'}}$. Let $v$ be the segment $v_\vdash$ of $\gleft(\bm{t'})$. The intersection of $h$ and $v$ is a point on the left boundary of $B_{\bm{t'}}$, which means that $h\cap v \in B_{\bm{t'}} \subseteq \cap_{i = 1}^{k-2}r_i$. Therefore, $\bm{t} = (r_1,...,r_{k-2},h,v)\in \J(R_1,...,R_{k-2},H,V)$ is the desired tuple. As a result, $\sum_{i \in [k-2]}\sum_{r_i^*\in R_i^*} |\cross_H(r_i^*)| \leq \out$.

### 2024-05-29 4.3.tex
4.3

@@
 The intersection of $h$ and $v$ is a point on the left boundary of $B_{\bm{t'}}$, which means that $h\cap v \in B_{\bm{t'}} \subseteq \cap_{i = 1}^{k-2}r_i$.
-Therefore, $\bm{t} = (r_1,...,r_{k-2},h,v)\in \J(R_1,...,R_{k-2},H,V)$ is the desired tuple.
-As a result, $\sum_{i \in [k-2]}\sum_{r_i^*\in R_i^*} |\cross_H(r_i^*)| \leq \out$.
+Therefore, $\bm{t} = (r_1,...,r_{k-2},h,v)\in \J(R_1,...,R_{k-2},H,V)$ is the desired tuple, and we have $\sum_{i \in [k-2]}\sum_{r_i^*\in R_i^*} |\cross_H(r_i^*)| \leq \out$.
+
+Next, we prove (2).
+Let $r' = [x_1,x_2]\times [y_1,y_2]$, and $r^* = [x_1,x_2] \times [y_1,\maxtop(r')]$.
+Since $\gbot(\bm{t}) = r'$, $\dcross_H(\bm{t})$ contains exactly every segment $h \in H$ that crosses the rectangle $[x_1,x_2]\times [y_1, \ytop(B_\bm{t})]$, where $\ytop(B_{\bm{t}})\leq \maxtop(r')$.
+Therefore, $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
+Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.

### 2024-05-29 4.3.tex

@@
 Let $r' = [x_1,x_2]\times [y_1,y_2]$, and $r^* = [x_1,x_2] \times [y_1,\maxtop(r')]$.
-Since $\gbot(\bm{t}) = r'$, $\dcross_H(\bm{t})$ contains exactly every segment $h \in H$ that crosses the rectangle $[x_1,x_2]\times [y_1, \ytop(B_\bm{t})]$, where $\ytop(B_{\bm{t}})\leq \maxtop(r')$.
+Since $\gbot(\bm{t}) = r'$, $\dcross_H(\bm{t})$ contains exactly the set of segments segment $h \in H$ that cross the rectangle $[x_1,x_2]\times [y_1, \ytop(B_\bm{t})]$, where $\ytop(B_{\bm{t}})\leq \maxtop(r')$.
 Therefore, $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.

### 2024-05-29 4.2.tex
4.2

@@
 First, we prove (1).
-By the definition of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$, $h$ and $v$ crosses $B_\bm{t}$.
+By the definition of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$, $h$ and $v$ cross $B_\bm{t}$.
 Thus, the intersection of $h$ and $v$ is a point in $B_\bm{t}$.
@@
 $v$) crosses the rectangles $r_i'$.
-Hence, the point $h \cap v$ is in $B_\bm{t}$, and $h \cap v$ crosses $B_{\bm{t}}$.
+Hence, the point $h \cap v$ is in $B_\bm{t}$, and $h, v$ cross $B_{\bm{t}}$.
 The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.

### 2024-05-29 4.2.tex
4.2

@@
 Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
-Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$. As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
+Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
+he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
+the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th component of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different.
+As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
 

### 2024-05-29 4.2.tex
4.2

--- 2024-05-29 4.2.tex
@@
 $v$) crosses the rectangles $r_i'$.
-Hence, the point $h \cap v$ is in $B_\bm{t}$, and $h, v$ cross $B_{\bm{t}}$.
+Hence, the point $h \cap v$ is in $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$, and $h, v$ cross $B_{\bm{t}}$.
 The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
@@
 Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
-Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$. As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
+Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
+he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
+the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th components of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different.
+As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
 

### 2024-05-29 msj.tex
yufei

@@
     \myenums{
-        \item $\sum_{i=1}^k \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
+        \item $\sum_{i=1}^{k-2} \sum_{r^* \in R^*_i} |\cross_H(r^*)| \le \out$.
 
@@
 
+\def\vgap{\vspace{2mm}}
+
 \section*{Appendix}
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}
-
-\input{4.1.tex}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} We will prove the lemma by providing a way to map each $(k-2)$-tuple $\J(R_1',...,R_{k-2}')$ to a unique $k$-tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+To explain the mapping, consider any tuple $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+Recall that $B_\bm{t}$ is the intersection of the $k-2$ rectangles $\bm{t}[1], ..., \bm{t}[k-2]$.
+Let $p$ be the bottom-left corner of $B_\bm{t}$.
+Define $i$ and $j$ to be values in $[k-2]$ such that $\bm{t}[i] = \gbot(\bm{t})$ and $\bm{t}[j] = \gleft(\bm{t})$; note that the values of $i$ and $j$ can be the same.
+Point $p$ is the intersection of the bottom edge of $\bm{t}[i]$ and the left edge of $\bm{t}[j]$.
+
+\vgap
+
+Recall that each $\bm{t}[z]$ (for $z \in [1, k-2]$) is a trimmed rectangle; let us denote by $r_z$ the full rectangle of $\bm{t}[z]$. Regarding the values $i, j$ decided earlier, let
+\myitems{
+    \item $h$ be the lowest segment in $H$ that crosses $r_i$;
+    \item $v$ be the leftmost segment in $V$ that crosses $r_j$.
+}
+By the way trimmed rectangles are computed, we know that $h$ must contain the bottom edge of rectangle $\bm{t}[i]$, and $v$ must contain the left edge of rectangle $\bm{t}[j]$.
+This means that point $p$ must be the intersection point of $h$ and $v$.
+
+\vgap
+
+We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ belongs to $\J(R_1,...,$ $R_{k-2},H,V)$.
+Clearly, $r_z$ covers $\bm{t}[z]$ for all $z \in [k-2]$.
+Thus, $\bigcap_{z=1}^{k-2} r_z$ covers $\bigcap_{z=1}^{k-2} \bm{t}[z]$, which is $B_\bm{t}$.
+Hence, point $p$ --- which is the intersection point of $h$ and $v$ and a corner of $B_\bm{t}$ --- falls in $\bigcap_{z=1}^{k-2} r_z$, indicating that $h \cap v \cap \bigcap_{z=1}^{k-2} r_z \ne \emptyset$.
+We now map $\bm{t}$ to $(r_1, ..., r_{k-2}, h, v)$.
+
+\vgap
+
+s
 
@@
 
-An axis-parallel rectangle in 3D has the form $r = [x_1, x_2] \times [y_1, y_2] \times [z_1, z_2]$.
+An axis-parallel rectangle in 3D space has the form $r = [x_1, x_2] \times [y_1, y_2] \times [z_1, z_2]$.
 We will refer to $[x_1, x_2]$ as the {\em x-projection} of $r$, and define its {\em y-} and {\em z-projections} analogously.

### 2024-05-29 msj.tex
yufei

@@
 Thus, $\bigcap_{z=1}^{k-2} r_z$ covers $\bigcap_{z=1}^{k-2} \bm{t}[z]$, which is $B_\bm{t}$.
-Hence, point $p$ --- which is the intersection point of $h$ and $v$ and a corner of $B_\bm{t}$ --- falls in $\bigcap_{z=1}^{k-2} r_z$, indicating that $h \cap v \cap \bigcap_{z=1}^{k-2} r_z \ne \emptyset$.
-We now map $\bm{t}$ to $(r_1, ..., r_{k-2}, h, v)$.
+Hence, point $p$ --- which is the intersection point of $h$ and $v$ and a corner of $B_\bm{t}$ --- falls in $\bigcap_{z=1}^{k-2} r_z$, indicating $h \cap v \cap \bigcap_{z=1}^{k-2} r_z \ne \emptyset$.
+We map $\bm{t}$ to $(r_1, ..., r_{k-2}, h, v)$.
 

### 2024-05-29 msj.tex
yufei

@@
 Thus, $\bigcap_{z=1}^{k-2} r_z$ covers $\bigcap_{z=1}^{k-2} \bm{t}[z]$, which is $B_\bm{t}$.
-Hence, point $p$ --- which is the intersection point of $h$ and $v$ and a corner of $B_\bm{t}$ --- falls in $\bigcap_{z=1}^{k-2} r_z$, indicating $h \cap v \cap \bigcap_{z=1}^{k-2} r_z \ne \emptyset$.
+Hence, point $p$ --- which is the intersection point of $h$ and $v$, and also a corner of $B_\bm{t}$ --- falls in $\bigcap_{z=1}^{k-2} r_z$, indicating $h \cap v \cap \bigcap_{z=1}^{k-2} r_z \ne \emptyset$.
 We map $\bm{t}$ to $(r_1, ..., r_{k-2}, h, v)$.
@@
 
-s
+It remains to prove that no two distinct tuples $\bm{t}_1, \bm{t}_2 \in \J(R_1',...,R_{k-2}')$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},$ $H, V)$.
+Assume that this is not true, namely, both $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
+However, in this case, $\bm{t}_1[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$, and $\bm{t}_2[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$.
+This indicates $\bm{t}_1 = \bm{t}_2$, giving a contradiction.
 

### 2024-05-29 msj.tex
yufei

@@
 
-It remains to prove that no two distinct tuples $\bm{t}_1, \bm{t}_2 \in \J(R_1',...,R_{k-2}')$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},$ $H, V)$.
+It remains to prove that no two distinct tuples $\bm{t}_1, \bm{t}_2 \in \J(R_1',...,$ $R_{k-2}')$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},$ $H, V)$.
 Assume that this is not true, namely, both $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
-However, in this case, $\bm{t}_1[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$, and $\bm{t}_2[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$.
-This indicates $\bm{t}_1 = \bm{t}_2$, giving a contradiction.
+However, in this case, $\bm{t}_1[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$ and, at the same time, $\bm{t}_2[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$.
+We thus have $\bm{t}_1 = \bm{t}_2$, giving a contradiction.
 

### 2024-05-29 4.2.tex
4.3

@@
 Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
-he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
+The result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
 the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th components of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different.

### 2024-05-29 4.2.tex
4.2

--- 2024-05-29 4.2.tex
@@
 Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
-Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
-he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
-the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th components of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different.
-As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
+Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$. As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
 

### 2024-05-29 msj.tex
yufei

@@
 
-Recall that each $\bm{t}[z]$ (for $z \in [1, k-2]$) is a trimmed rectangle; let us denote by $r_z$ the full rectangle of $\bm{t}[z]$. Regarding the values $i, j$ decided earlier, let
+Recall that each $\bm{t}[z]$ (for $z \in [k-2]$) is a trimmed rectangle; let us denote by $r_z$ the full rectangle of $\bm{t}[z]$. Regarding the values $i, j$ decided earlier, let
 \myitems{
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
-
-\input{4.2.tex}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} \\
+\underline{\em Proof of Statement (1).} By the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$, both segments $h$ and $v$ cross $B_\bm{t}$.
+Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$.
+As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know $\bigcap_{i=1}^{k-2} r_i$ covers $\bigcap_{i=1}^{k-2} \bm{t}[i]$, which is $B_\bm{t}$.
+This means that point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$ and thus $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
+
+Next, we prove (2). Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both $h$
+ and $v$ cross the rectangles $r_1,...,r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
+Additionally, for each $i\in[k-2]$, the x-range (resp.
+y-range) of $r_i'$ is a subset of the x-range (resp.
+y-range) of $r_i$, so $h$ (resp.
+$v$) crosses the rectangles $r_i'$.
+Hence, the point $h \cap v$ is in $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$, and $h, v$ cross $B_{\bm{t}}$.
+The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+
+Finally, we prove (3).
+For each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $h \in \dcross_H(\bm{t'})$, we will identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
+Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $v$ be the segment $v_\vdash$ of $\gleft(\bm{t'})$.
+Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
+Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
+Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
+he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
+the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th components of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different.
+As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
+
+Similarly, for each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $v \in \dcross_V(\bm{t'})$, we can identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
+Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $h$ be the segment $h_\bot$ of $\gbot(\bm{t'})$.
+Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
+Since $v$ crosses $B_{\bm{t'}}$, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
+As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq \out$.
 

### 2024-05-29 4.3.tex

@@
 First, we prove (1).
-For any $i \in [k-2]$, $r_i^*\in R_i^*$, and $h \in \cross_H(r_i^*)$, we will identify a unique tuple $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$ such that $\bm{t}[k-1] = h$, and $\bm{t}[i]$ is the full rectangle of $r_i'$, whose top-slice is $r_i^*$.
+For any $i \in [k-2]$, $r_i^*\in R_i^*$, and $h \in \cross_H(r_i^*)$, we will identify a unique tuple $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$ such that $\bm{t}[k-1] = h$, and $\bm{t}[i]$ is the full rectangle of $r_i'$, whose top-sliced rectangle is $r_i^*$.
 Let $\maxtop(r_i') = \ytop(B_\bm{t'})$ for some tuple $\bm{t'} \in \J(R_1',...,R_{k-2}')$.

### 2024-05-29 msj.tex

@@
 As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know $\bigcap_{i=1}^{k-2} r_i$ covers $\bigcap_{i=1}^{k-2} \bm{t}[i]$, which is $B_\bm{t}$.
-This means that point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$ and thus $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
-
-Next, we prove (2). Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both $h$
- and $v$ cross the rectangles $r_1,...,r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
+Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$.
+It follows that $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+\noindent \underline{\em Proof of Statement (2).} The fact $(r_1,...,r_{k-2},h,v)\in \J_1$ indicates that both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
 Additionally, for each $i\in[k-2]$, the x-range (resp.
@@
 
-Finally, we prove (3).
+\vgap
+
+\noindent \underline{\em Proof of Statement (3).}
 For each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $h \in \dcross_H(\bm{t'})$, we will identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
@@
 As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
+
+\vgap
 

### 2024-05-29 4.2.tex
yufei

--- 2024-05-29 4.2.tex
@@
-First, we prove (1).
-By the definition of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$, $h$ and $v$ cross $B_\bm{t}$.
-Thus, the intersection of $h$ and $v$ is a point in $B_\bm{t}$.
-Since $B_\bm{t}$ is the intersection of $\bm{t}[1],...,\bm{t}[k-2]$, the intersection of $h$ and $v$ lies in all the rectangles $r_1,...,r_{k-2}$.
-Therefore, $h\cap v\bigcap_{i = 1}^{k-2}r_i \neq \emptyset$, and  the tuple $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
-
-Next, we prove (2). Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both $h$
- and $v$ cross the rectangles $r_1,...,r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
+\noindent \underline{\em Proof of Statement (2).} The fact $(r_1,...,r_{k-2},h,v)\in \J_1$ indicates that both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
 Additionally, for each $i\in[k-2]$, the x-range (resp.
@@
 
-Finally, we prove (3). 
+\vgap
+
+\noindent \underline{\em Proof of Statement (3).}
 For each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $h \in \dcross_H(\bm{t'})$, we will identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
@@
 
+\vgap
+
 Similarly, for each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $v \in \dcross_V(\bm{t'})$, we can identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.

### 2024-05-29 msj.tex
yufei

--- 2024-05-29 msj.tex
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
-
-\input{4.2.tex}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} \\
+\underline{\em Proof of Statement (1).} By the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$, both segments $h$ and $v$ cross $B_\bm{t}$.
+Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$.
+As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $\bigcap_{i=1}^{k-2} \bm{t}[i]$, which is $B_\bm{t}$.
+Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$.
+It follows that $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
 

### 2024-05-29 msj.tex
yufei

@@
 
-\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}
-
-\input{4-comp.tex}
+\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} We consider, w.l.o.g., that each rectangle $r$ in the input $R_1 \cup ...
+\cup R_{k-2}$ is given a distinct integer ID in $[n]$.
+This allows us to create an array $A$ of size $n$ and store in $A[i]$ ($i \in [n]$) a constant amount of information for each $r$.
+
+\vgap
+
+To compute $R^*_1, R^*_2, ..., R^*_{k-2}$, we start by deriving $\maxtop(r')$ for each rectangle $r'$ in $R'_1 \cup  ...
+\cup R'_{k-2}$.
+For this purpose, first initialize $\maxtop(r') = -\infty$ for each such $r'$.
+Recall that $r'$ is the trimmed rectangle of some rectangle $r$ in $R_1 \cup ...
+\cup R_{k-2}$.
+We store $\maxtop(r')$ in the cell of array $A$ corresponding to $r$; the cell can be accessed in $O(1)$ time using the ID of $r$.
+
+\vgap
+
+Then, we scan $\J(R_1',...,R_{k-2}')$.
+For each tuple $\bm{t}$ therein, use $O(k)$ time to identify the rectangle $r' = \gbot(\bm{t})$, and then update $\maxtop(r')$ to the maximum between its current value and $\ytop(r')$.
+The scan requires $O(n + k \cdot \out)$ time.
+
+\vgap
+
+Finally, we can construct the set $R_i^*$ for every $i \in [k-2]$ by collecting the top-sliced rectangles of all rectangles $r_i' \in R_i'$ satisfying that $\maxtop(r_i')\neq -\infty$.
+This step takes $O(n)$ additional time.
+The overall time complexity is $O(n + k\cdot \out)$.
 

### 2024-05-29 msj.tex
yufei

@@
 \myeqn{
-    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
+    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \label{eqn:hv:type1:topsliced}
 }
@@
 
-Finally, we can construct the set $R_i^*$ for every $i \in [k-2]$ by collecting the top-sliced rectangles of all rectangles $r_i' \in R_i'$ satisfying that $\maxtop(r_i')\neq -\infty$.
-This step takes $O(n)$ additional time.
-The overall time complexity is $O(n + k\cdot \out)$.
+Finally, for each $i \in [k-2]$, we construct $R_i^*$ by collecting the top-sliced rectangle of every rectangle $r' \in R_i'$ (see definition in \eqref{eqn:hv:type1:topsliced}) with $\maxtop(r') \neq -\infty$.
+This step takes $O(|R_i'|)$ time for each $i \in [k-2]$, or $O(n)$ total time for all $i$.
 

### 2024-05-29 msj.tex
yufei

@@
 
-\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} We consider, w.l.o.g., that each rectangle $r$ in the input $R_1 \cup ...
+\input{4.2}
+
+\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} We consider, w.l.o.g., that each rectangle in the input $R_1 \cup ...
 \cup R_{k-2}$ is given a distinct integer ID in $[n]$.
-This allows us to create an array $A$ of size $n$ and store in $A[i]$ ($i \in [n]$) a constant amount of information for each $r$.
+This allows us to create an array $A$ of size $n$ and allocate a cell of $A$ to each $r \in R_1 \cup ...
+\cup R_{k-2}$.
+The cell can be accessed by the ID of $r$ in constant time.
 

### 2024-05-29 msj.tex
yufei

@@
 \cup R_{k-2}$.
-We store $\maxtop(r')$ in the cell of array $A$ corresponding to $r$; the cell can be accessed in $O(1)$ time using the ID of $r$.
+We store $\maxtop(r')$ in the cell of array $A$ allocated to $r$; the cell can be accessed in $O(1)$ time using the ID of $r$.
 
@@
 Then, we scan $\J(R_1',...,R_{k-2}')$.
-For each tuple $\bm{t}$ therein, use $O(k)$ time to identify the rectangle $r' = \gbot(\bm{t})$, and then update $\maxtop(r')$ to the maximum between its current value and $\ytop(r')$.
+For each tuple $\bm{t}$ therein, use $O(k)$ time to identify the rectangle $r' = \gbot(\bm{t})$, and then update in constant time $\maxtop(r')$ to the maximum between its current value and $\ytop(r')$.
 The scan requires $O(n + k \cdot \out)$ time.
@@
 
-Finally, for each $i \in [k-2]$, we construct $R_i^*$ by collecting the top-sliced rectangle of every rectangle $r' \in R_i'$ (see definition in \eqref{eqn:hv:type1:topsliced}) with $\maxtop(r') \neq -\infty$.
+Finally, for each $i \in [k-2]$, we construct $R_i^*$ by collecting the top-sliced rectangle (see definition in \eqref{eqn:hv:type1:topsliced}) of every rectangle $r' \in R_i'$ with $\maxtop(r') \neq -\infty$.
 This step takes $O(|R_i'|)$ time for each $i \in [k-2]$, or $O(n)$ total time for all $i$.

### 2024-05-29 4.2.tex
4.2

@@
-\noindent \underline{\em Proof of Statement (2).} The fact $(r_1,...,r_{k-2},h,v)\in \J_1$ indicates that both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
+\noindent \underline{\em Proof of Statement (2).} 
+First, we deviate a little to prove the following statement.
+For any rectangle $r\in \bigcup_{i = 1}^{k-2}$, if a segment $h\in H$ crosses $r$, then $h$ also crosses the trimmed rectangle of $r$, denoted as $r'$.
+Since the x-range of $r'$ is a subset of the x-range of $r$, it suffices to show that the y-coordinate of $h$ is in the range $[\ybot(r'), \ytop(r')]$.
+Since $\ybot(r')$ is the y-coordinate of the lowest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ybot(r') \leq \ybot(h)$.
+Moreover, since $\ytop(r')$ is the y-coordinate of the highest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ytop(h) \leq \ytop(r')$.
+Therefore, $h$ crosses $r'$.
+We can prove that if a segment $v\in V$ crosses $r$, then $v$ also crosses the trimmed rectangle of $r$ analogously.
+
+\vgap
+
+Now, we prove Statement (2).
+Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$.
+By the discussion above, $h$ and $v$ also cross the trimmed rectangles $r_1',..., r_{k-2}'$.
+Therefore, the point $h \cap v$ is in $r_i'$, for each $i \in [k-2]$.
+This implies that $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$, and $(r_1',..., r_{k-2}')\in \J(R_1',...,R_{k-2}')$.
+
+% The fact $(r_1,...,r_{k-2},h,v)\in \J_1$ indicates that both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
 Additionally, for each $i\in[k-2]$, the x-range (resp.

### 2024-05-29 msj.tex
yufei

@@
 
-\def\vgap{\vspace{2mm}}
+\def\vgap{\vspace{1mm}}
 
@@
 \cup R_{k-2}$ is given a distinct integer ID in $[n]$.
-This allows us to create an array $A$ of size $n$ and allocate a cell of $A$ to each $r \in R_1 \cup ...
-\cup R_{k-2}$.
-The cell can be accessed by the ID of $r$ in constant time.
+This allows us to create an array of size $n$ and allocate an array cell to each $r \in R_1 \cup ...
+\cup R_{k-2}$, with the property that the cell can be accessed by the ID of $r$ in constant time.
 
@@
 \cup R_{k-2}$.
-We store $\maxtop(r')$ in the cell of array $A$ allocated to $r$; the cell can be accessed in $O(1)$ time using the ID of $r$.
+We store $\maxtop(r')$ in the array cell allocated to $r$.
 

### 2024-05-29 4.2.tex
4.2

@@
 Since the x-range of $r'$ is a subset of the x-range of $r$, it suffices to show that the y-coordinate of $h$ is in the range $[\ybot(r'), \ytop(r')]$.
-Since $\ybot(r')$ is the y-coordinate of the lowest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ybot(r') \leq \ybot(h)$.
-Moreover, since $\ytop(r')$ is the y-coordinate of the highest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ytop(h) \leq \ytop(r')$.
-Therefore, $h$ crosses $r'$.
-We can prove that if a segment $v\in V$ crosses $r$, then $v$ also crosses the trimmed rectangle of $r$ analogously.
+Let $h = [x_1,x_2]\times y$.
+Since $\ybot(r')$ is the y-coordinate of the lowest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ybot(r') \leq y$.
+Moreover, since $\ytop(r')$ is the y-coordinate of the highest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $y \leq \ytop(r')$.
+Therefore, $y \in [\ybot(r'), \ytop(r')]$ and $h$ crosses $r'$.
+Analogously, we can prove that if a segment $v\in V$ crosses $r$, then $v$ also crosses the trimmed rectangle of $r$.
 

### 2024-05-29 4.2.tex

@@
 
-Now, we prove Statement (2).
+Now, we prove the first bullet of Statement (2).
 Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$.
 By the discussion above, $h$ and $v$ also cross the trimmed rectangles $r_1',..., r_{k-2}'$.
-Therefore, the point $h \cap v$ is in $r_i'$, for each $i \in [k-2]$.
-This implies that $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$, and $(r_1',..., r_{k-2}')\in \J(R_1',...,R_{k-2}')$.
-
-% The fact $(r_1,...,r_{k-2},h,v)\in \J_1$ indicates that both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
-Additionally, for each $i\in[k-2]$, the x-range (resp.
-y-range) of $r_i'$ is a subset of the x-range (resp.
-y-range) of $r_i$, so $h$ (resp.
-$v$) crosses the rectangles $r_i'$.
-Hence, the point $h \cap v$ is in $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$, and $h, v$ cross $B_{\bm{t}}$.
-The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+Let $h = [x_1,x_2]\times y$, $v = x \times [y_1,y_2]$, and $p = (x,y)$ be the intersection point of $h$ and $v$.
+Since $h$ crosses $r_i'$, we have $y \in [\ybot(r_i'), \ytop(r_i')]$, and since $v$ crosses $r_i'$, we have $x \in [\xleft(r_i'), \xright(r_i')]$.
+Hence, the point $p$ in $r_i'$, for each $i \in [k-2]$.
+This implies that $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$, and $\bm{t}\in \J(R_1',...,R_{k-2}')$.
 
@@
 
-\noindent \underline{\em Proof of Statement (3).}
-For each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $h \in \dcross_H(\bm{t'})$, we will identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
-Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $v$ be the segment $v_\vdash$ of $\gleft(\bm{t'})$.
-Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
-Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
-Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
-he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
-the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th components of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different.
-As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
+Finally, we prove the second bullet of Statement (2).
+As we proved that $h$ and $v$ cross $r_i'$ for all $i \in [k-2]$, they must cross $\gbot(\bm{t})$ and $\gleft(\bm{t})$ as well.
+Hence, it suffices to show that both $h$ and $v$ cross $B_\bm{t}$.
+As proved in the last paragraph, the point $p = h \cap v$ is in $B_\bm{t}$.
+So, $h\cap B_\bm{t}$ and $v \cap B_\bm{t}$ are non-empty.
+Since $h$ (resp.
+$v$) crosses $r_i'$ for some $i \in [k-2]$, $r_i'$ does not cover any of the endpoints of $h$ (resp.
+$v$).
+Additonally, $B_\bm{t}$ is a subset of $r_i'$, so $B_\bm{t}$ does not cover any of the endpoints of $h$ or $v$.
+Therefore, $h$ and $v$ cross $B_\bm{t}$.
 
@@
 
-Similarly, for each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $v \in \dcross_V(\bm{t'})$, we can identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
-Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $h$ be the segment $h_\bot$ of $\gbot(\bm{t'})$.
-Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
-Since $v$ crosses $B_{\bm{t'}}$, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
-As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq \out$.
+\noindent \underline{\em Proof of Statement (3).}  First, we prove that $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})|$ does not exceeds $\out$.
+Then, we can apply similar arguments to show that $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq |\out|$.
+
+\vgap
+
+We will provide a way to map each pair $(\bm{t},h)$ in $\J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$ to a unique tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+To explain the mapping, consider any pair $(\bm{t},h) \in \J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$.
+Recall that $B_\bm{t}$ is the intersection of the $k-2$ rectangles $\bm{t}[1], ..., \bm{t}[k-2]$.
+Since $h \in \dcross_H(\bm{t})$ (see definition~\eqref{eqn:dcross-H}), $h$ crosses $B_\bm{t}$.
+Let $p$ be the intersection point of $h$ and the left edge of $B_\bm{t}$.
+Define $i$ to be the value in $[k-2]$ such that $\bm{t}[i] = \gleft(\bm{t})$.
+
+\vgap
+
+Recall that each $\bm{t}[j]$ (for $j \in [k-2]$) is a trimmed rectangle; let us denote by $r_j$ the full rectangle of $\bm{t}[j]$.
+Regarding the value $i$ decided earlier, let $v$ be the leftmost segment in $V$ that crosses $r_i$.
+By the way trimmed rectangles are computed, we know that $v$ must contain the left edge of rectangle $\bm{t}[i]$.
+This means that point $p$ must be the intersection point of $h$ and $v$.
+
+\vgap
+
+We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)\in \J(R_1,..., R_{k-2},H,V)$.
+Clearly, $r_j$ covers $\bm{t}[j]$ for all $j \in [k-2]$.
+Thus, $\bigcap_{j=1}^{k-2} r_j$ covers $\bigcap_{j=1}^{k-2} \bm{t}[j]$, which is $B_\bm{t}$.
+Hence, point $p$ --- which is the intersection point of $h$ and the left edge of $B_\bm{t}$ --- falls in $\bigcap_{j=1}^{k-2} r_j$, indicating $h \cap v \cap \bigcap_{j=1}^{k-2} r_j \ne \emptyset$.
+We map $(\bm{t},h)$ to $(r_1, ..., r_{k-2}, h, v)$.
+
+\vgap
+
+It remains to prove that no two distinct pairs $(\bm{t}_1, h_1), (\bm{t}_2, h_2) \in \J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},H,V)$.
+Assume that this is not true, namely, both $(\bm{t}_1, h_1)$ and $(\bm{t}_2, h_2)$ are mapped to the tuple $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
+However, in this case, $\bm{t}_1[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-1] = h_1$.
+At the same time, $\bm{t}_2[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-1] = h_2$.
+We thus have $\bm{t}_1 = \bm{t}_2$ and $h_1 = h_2$, giving a contradiction.
+Therefore, we have $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq |\out|$.
+
+\vgap
+
+Analogously, we can map each pair $(\bm{t},v)$ in $\J(R_1',...,R_{k-2}')\times \dcross_V(\bm{t})$ to a unique tuple $(r_1,...,r_{k-2},h,v)$, where 
+\begin{itemize}
+    \item $r_j$ is the full rectangle of $\bm{t}[j]$ for all $j \in [k-2]$,
+    \item $h$ is the lowest segment in $H$ that crosses $r_i$, where $i$ is the value in $[k-2]$ such that $\bm{t}[i] = \gbot(\bm{t})$.
+\end{itemize}
+Let $p$ be the intersection of $v$ and the bottom edge of $B_\bm{t}$.
+We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)\in \J(R_1,..., R_{k-2},H,V)$.
+By the way trimmed rectangles are computed, we know that $h$ must contain the bottom edge of rectangle $\bm{t}[i]$, which is a superset of the bottom edge of $B_\bm{t}$.
+This means that point $p$ must be the intersection point of $h$ and $v$.
+As shown before, $r_j$ covers $B_\bm{t}$.
+Hence, point $p$ falls in $\bigcap_{j=1}^{k-2} r_j$, indicating $h \cap v \cap \bigcap_{j=1}^{k-2} r_j \ne \emptyset$.
+
+\vgap
+
+It suffices to show that no two distinct pairs $(\bm{t}_1, v_1), (\bm{t}_2, v_2) \in \J(R_1',...,R_{k-2}')\times \dcross_V(\bm{t})$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},H,V)$.
+Assume the contrary, namely, both $(\bm{t}_1, v_1)$ and $(\bm{t}_2, v_2)$ are mapped to the tuple $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
+However, in this case, $\bm{t}_1[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-2] = v_1$.
+At the same time, $\bm{t}_2[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-2] = v_2$.
+We thus have $\bm{t}_1 = \bm{t}_2$ and $v_1 = v_2$, giving a contradiction.
+Therefore, we have $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq |\out|$.

### 2024-05-29 msj.tex
yufei

@@
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} \\
-\underline{\em Proof of Statement (1).} By the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$, both segments $h$ and $v$ cross $B_\bm{t}$.
+\underline{\em Proof of Statement (1).} As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
 Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$.
-As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $\bigcap_{i=1}^{k-2} \bm{t}[i]$, which is $B_\bm{t}$.
+As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
 Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$.
@@
 
-\input{4.2}
+\noindent \underline{\em Proof of Statement (2).} Fix any rectangle $r \in R_1 \cup ...
+\cup R_{k-2}$ and any horizontal segment $h = [x_1, x_2] \times y$ from $H$ that crosses $r$.
+We will show that $h$ must also cross the trimmed rectangle $r'$ of $r$.
+From the fact of $h$ crossing $r$, it is clear that $[x_1, x_2]$ contains the x-range of $r$.
+As the x-range of $r$ contains that of $r'$, we can assert that $[x_1, x_2]$ contains the x-range of $r'$.
+To prove that $h$ crosses $r'$, we still need to show $y \in [\ybot(r'), \ytop(r')]$.
+Recall that $\ybot(r')$ is the y-coordinate of the lowest segment in $H$ crossing $r$.
+This implies $y \ge \ybot(r')$ because $h$ itself is a segment in $H$ crossing $r$.
+Analogously, it also holds that $y \le \ytop(r')$.
+We can now conclude that $h$ crosses $r'$.
+
+\vgap
+
+Fix any rectangle $r \in R_1 \cup ... \cup R_{k-2}$ and any vertical segment $v$ from $V$ that crosses $r$. An argument symmetric to the above shows that $v \in V$ must also cross $r'$.
+
+\vgap
+
+Next, we prove the first bullet of statement (2), namely, $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+Consider the $k$-tuple $(r_1,...,r_{k-2},h,v)\in \J_1$ given in statement (2).
+By definition of $\J_1$, segments $h$ and $v$ both cross each of the rectangles $r_1,..., r_{k-2}$.
+By the earlier discussion, $h$ and $v$ also both cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
+This means that the intersection of $h$ and $v$ must be a point in each of $r_1',..., r_{k-2}'$, suggesting $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$.
+It thus holds that $\bm{t} = (r_1', ..., r_{k-2}')$ is a result tuple in $\J(R_1',...,R_{k-2}')$.
+
+\vgap
+
+Finally, we prove the second bullet of Statement (2).
+As we proved that $h$ and $v$ cross $r_i'$ for all $i \in [k-2]$, they must cross $\gbot(\bm{t})$ and $\gleft(\bm{t})$ as well.
+Hence, it suffices to show that both $h$ and $v$ cross $B_\bm{t}$.
+As proved in the last paragraph, the point $p = h \cap v$ is in $B_\bm{t}$.
+So, $h\cap B_\bm{t}$ and $v \cap B_\bm{t}$ are non-empty.
+Since $h$ (resp.
+$v$) crosses $r_i'$ for some $i \in [k-2]$, $r_i'$ does not cover any of the endpoints of $h$ (resp.
+$v$).
+Additonally, $B_\bm{t}$ is a subset of $r_i'$, so $B_\bm{t}$ does not cover any of the endpoints of $h$ or $v$.
+Therefore, $h$ and $v$ cross $B_\bm{t}$.
 

### 2024-05-29 msj.tex
yufei

@@
 
-Next, we prove the first bullet of statement (2), namely, $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+We now prove the first bullet of statement (2), namely, $\bm{t}\in \J(R_1',...,R_{k-2}')$.
 Consider the $k$-tuple $(r_1,...,r_{k-2},h,v)\in \J_1$ given in statement (2).
 By definition of $\J_1$, segments $h$ and $v$ both cross each of the rectangles $r_1,..., r_{k-2}$.
-By the earlier discussion, $h$ and $v$ also both cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
-This means that the intersection of $h$ and $v$ must be a point in each of $r_1',..., r_{k-2}'$, suggesting $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$.
-It thus holds that $\bm{t} = (r_1', ..., r_{k-2}')$ is a result tuple in $\J(R_1',...,R_{k-2}')$.
-
-\vgap
-
-Finally, we prove the second bullet of Statement (2).
-As we proved that $h$ and $v$ cross $r_i'$ for all $i \in [k-2]$, they must cross $\gbot(\bm{t})$ and $\gleft(\bm{t})$ as well.
-Hence, it suffices to show that both $h$ and $v$ cross $B_\bm{t}$.
-As proved in the last paragraph, the point $p = h \cap v$ is in $B_\bm{t}$.
-So, $h\cap B_\bm{t}$ and $v \cap B_\bm{t}$ are non-empty.
-Since $h$ (resp.
-$v$) crosses $r_i'$ for some $i \in [k-2]$, $r_i'$ does not cover any of the endpoints of $h$ (resp.
-$v$).
-Additonally, $B_\bm{t}$ is a subset of $r_i'$, so $B_\bm{t}$ does not cover any of the endpoints of $h$ or $v$.
-Therefore, $h$ and $v$ cross $B_\bm{t}$.
+By the earlier discussion, $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
+Hence, the intersection of $h$ and $v$ must be a point covered by each of $r_1',..., r_{k-2}'$, suggesting the point is in $\bigcap_{i = 1}^{k-2}r_i'$, which is thus not empty.
+Therefore, $\bm{t} = (r_1', ..., r_{k-2}')$ is a result tuple in $\J(R_1',...,R_{k-2}')$.
+
+\vgap
+
+Next, we prove the second bullet, namely, $h \in \dcross_H(\bm{t})$ and $v \in \dcross_H(\bm{t})$.
+It suffices to show only the former due to symmetry.
+To prove $h \in \dcross_H(\bm{t})$, we need to argue that $h$ crosses $\gbot(\bm{t})$ and $B_\bm{t}$.
+The first part, $h$ crossing $\gbot(\bm{t})$, is done because, as already explained, $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is one of those $k-2$ rectangles.
+To prove that $h$ crosses $B_\bm{t}$, first note that $B_\bm{t}$ (which is the intersection of $r_1', ..., r_{k-2}'$) is covered by $\gbot(\bm{t})$.
+From the fact that $h$ crosses $\gbot(\bm{t})$, we can assert that the y-range of $h$ contains that of $B_\bm{t}$.
+Hence, either $h$ is disjoint with $B_\bm{t}$ or $h$ crosses $B_\bm{t}$.
+However, $h$ cannot be disjoint with $B_\bm{t}$ because, as shown earlier, the intersection of $h$ and $v$ falls in $B_\bm{t}$.
+It thus follows that $h$ crosses $B_\bm{t}$.
 

### 2024-05-29 msj.tex
yufei

@@
 Analogously, it also holds that $y \le \ytop(r')$.
-We can now conclude that $h$ crosses $r'$.
-
-\vgap
-
-Fix any rectangle $r \in R_1 \cup ... \cup R_{k-2}$ and any vertical segment $v$ from $V$ that crosses $r$. An argument symmetric to the above shows that $v \in V$ must also cross $r'$.
+We can thus conclude that $h$ crosses $r'$.
+
+\vgap
+
+Fix any rectangle $r \in R_1 \cup ... \cup R_{k-2}$ and any vertical segment $v$ from $V$ that crosses $r$. An argument symmetric to the above shows that $v$ must also cross $r'$.
 
@@
 By the earlier discussion, $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
-Hence, the intersection of $h$ and $v$ must be a point covered by each of $r_1',..., r_{k-2}'$, suggesting the point is in $\bigcap_{i = 1}^{k-2}r_i'$, which is thus not empty.
+Hence, the intersection of $h$ and $v$ must be a point covered by each of $r_1',..., r_{k-2}'$.
+This suggests that the point is in $\bigcap_{i = 1}^{k-2}r_i'$, which thus cannot be empty.
 Therefore, $\bm{t} = (r_1', ..., r_{k-2}')$ is a result tuple in $\J(R_1',...,R_{k-2}')$.
@@
 To prove that $h$ crosses $B_\bm{t}$, first note that $B_\bm{t}$ (which is the intersection of $r_1', ..., r_{k-2}'$) is covered by $\gbot(\bm{t})$.
-From the fact that $h$ crosses $\gbot(\bm{t})$, we can assert that the y-range of $h$ contains that of $B_\bm{t}$.
+The fact that $h$ crosses $\gbot(\bm{t})$ indicates that the y-range of $h$ contains that of $B_\bm{t}$.
 Hence, either $h$ is disjoint with $B_\bm{t}$ or $h$ crosses $B_\bm{t}$.

### 2024-05-29 msj.tex
yufei

@@
 
-Next, we prove the second bullet, namely, $h \in \dcross_H(\bm{t})$ and $v \in \dcross_H(\bm{t})$.
+Next, we prove the second bullet, namely, $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
 It suffices to show only the former due to symmetry.
@@
 It thus follows that $h$ crosses $B_\bm{t}$.
+
+\vgap
+
+\noindent \underline{\em Proof of Statement (3).} Our discussion will focus on explaining why $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})|$ $\le \out$ because a similar argument can then used to prove $\sum_{\bm{t} \in \J(R_1',...,R_{k-2}')}$ $|\dcross_V(\bm{t})| \le \out$.
+
+\vgap
+
+We will provide a way to map each pair $(\bm{t},h)$ in $\J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$ to a unique tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+To explain the mapping, consider any pair $(\bm{t},h) \in \J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$.
+Recall that $B_\bm{t}$ is the intersection of the $k-2$ rectangles $\bm{t}[1], ..., \bm{t}[k-2]$.
+Since $h \in \dcross_H(\bm{t})$ (see definition~\eqref{eqn:dcross-H}), $h$ crosses $B_\bm{t}$.
+Let $p$ be the intersection point of $h$ and the left edge of $B_\bm{t}$.
+Define $i$ to be the value in $[k-2]$ such that $\bm{t}[i] = \gleft(\bm{t})$.
+
+\vgap
+
+Recall that each $\bm{t}[j]$ (for $j \in [k-2]$) is a trimmed rectangle; let us denote by $r_j$ the full rectangle of $\bm{t}[j]$.
+Regarding the value $i$ decided earlier, let $v$ be the leftmost segment in $V$ that crosses $r_i$.
+By the way trimmed rectangles are computed, we know that $v$ must contain the left edge of rectangle $\bm{t}[i]$.
+This means that point $p$ must be the intersection point of $h$ and $v$.
+
+\vgap
+
+We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)\in \J(R_1,..., R_{k-2},H,V)$.
+Clearly, $r_j$ covers $\bm{t}[j]$ for all $j \in [k-2]$.
+Thus, $\bigcap_{j=1}^{k-2} r_j$ covers $\bigcap_{j=1}^{k-2} \bm{t}[j]$, which is $B_\bm{t}$.
+Hence, point $p$ --- which is the intersection point of $h$ and the left edge of $B_\bm{t}$ --- falls in $\bigcap_{j=1}^{k-2} r_j$, indicating $h \cap v \cap \bigcap_{j=1}^{k-2} r_j \ne \emptyset$.
+We map $(\bm{t},h)$ to $(r_1, ..., r_{k-2}, h, v)$.
+
+\vgap
+
+It remains to prove that no two distinct pairs $(\bm{t}_1, h_1), (\bm{t}_2, h_2) \in \J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},H,V)$.
+Assume that this is not true, namely, both $(\bm{t}_1, h_1)$ and $(\bm{t}_2, h_2)$ are mapped to the tuple $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
+However, in this case, $\bm{t}_1[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-1] = h_1$.
+At the same time, $\bm{t}_2[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-1] = h_2$.
+We thus have $\bm{t}_1 = \bm{t}_2$ and $h_1 = h_2$, giving a contradiction.
+Therefore, we have $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq |\out|$.
+
+\vgap
+
+Analogously, we can map each pair $(\bm{t},v)$ in $\J(R_1',...,R_{k-2}')\times \dcross_V(\bm{t})$ to a unique tuple $(r_1,...,r_{k-2},h,v)$, where 
+\begin{itemize}
+    \item $r_j$ is the full rectangle of $\bm{t}[j]$ for all $j \in [k-2]$,
+    \item $h$ is the lowest segment in $H$ that crosses $r_i$, where $i$ is the value in $[k-2]$ such that $\bm{t}[i] = \gbot(\bm{t})$.
+\end{itemize}
+Let $p$ be the intersection of $v$ and the bottom edge of $B_\bm{t}$.
+We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)\in \J(R_1,..., R_{k-2},H,V)$.
+By the way trimmed rectangles are computed, we know that $h$ must contain the bottom edge of rectangle $\bm{t}[i]$, which is a superset of the bottom edge of $B_\bm{t}$.
+This means that point $p$ must be the intersection point of $h$ and $v$.
+As shown before, $r_j$ covers $B_\bm{t}$.
+Hence, point $p$ falls in $\bigcap_{j=1}^{k-2} r_j$, indicating $h \cap v \cap \bigcap_{j=1}^{k-2} r_j \ne \emptyset$.
+
+\vgap
+
+It suffices to show that no two distinct pairs $(\bm{t}_1, v_1), (\bm{t}_2, v_2) \in \J(R_1',...,R_{k-2}')\times \dcross_V(\bm{t})$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},H,V)$.
+Assume the contrary, namely, both $(\bm{t}_1, v_1)$ and $(\bm{t}_2, v_2)$ are mapped to the tuple $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
+However, in this case, $\bm{t}_1[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-2] = v_1$.
+At the same time, $\bm{t}_2[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-2] = v_2$.
+We thus have $\bm{t}_1 = \bm{t}_2$ and $v_1 = v_2$, giving a contradiction.
+Therefore, we have $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq |\out|$.
 

### 2024-05-29 msj.tex
yufei

@@
 
-We will provide a way to map each pair $(\bm{t},h)$ in $\J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$ to a unique tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
-
-\vgap
-
-To explain the mapping, consider any pair $(\bm{t},h) \in \J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$.
+To argue for $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})|$ $\le \out$, we will provide a way to map each pair in $\J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$ to a unique tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+To explain the mapping, consider an arbitrary pair $(\bm{t},h) \in \J(R_1',...,R_{k-2}')\times \dcross_H(\bm{t})$.
 Recall that $B_\bm{t}$ is the intersection of the $k-2$ rectangles $\bm{t}[1], ..., \bm{t}[k-2]$.
-Since $h \in \dcross_H(\bm{t})$ (see definition~\eqref{eqn:dcross-H}), $h$ crosses $B_\bm{t}$.
-Let $p$ be the intersection point of $h$ and the left edge of $B_\bm{t}$.
+As $h \in \dcross_H(\bm{t})$, we know that $h$ crosses $B_\bm{t}$ and hence must intersect with the left edge of $B_\bm{t}$.
+Let $p$ denote that intersection point.
+
  Define $i$ to be the value in $[k-2]$ such that $\bm{t}[i] = \gleft(\bm{t})$.
@@
 Therefore, we have $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq |\out|$.
-
-\vgap
-
-Analogously, we can map each pair $(\bm{t},v)$ in $\J(R_1',...,R_{k-2}')\times \dcross_V(\bm{t})$ to a unique tuple $(r_1,...,r_{k-2},h,v)$, where 
-\begin{itemize}
-    \item $r_j$ is the full rectangle of $\bm{t}[j]$ for all $j \in [k-2]$,
-    \item $h$ is the lowest segment in $H$ that crosses $r_i$, where $i$ is the value in $[k-2]$ such that $\bm{t}[i] = \gbot(\bm{t})$.
-\end{itemize}
-Let $p$ be the intersection of $v$ and the bottom edge of $B_\bm{t}$.
-We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)\in \J(R_1,..., R_{k-2},H,V)$.
-By the way trimmed rectangles are computed, we know that $h$ must contain the bottom edge of rectangle $\bm{t}[i]$, which is a superset of the bottom edge of $B_\bm{t}$.
-This means that point $p$ must be the intersection point of $h$ and $v$.
-As shown before, $r_j$ covers $B_\bm{t}$.
-Hence, point $p$ falls in $\bigcap_{j=1}^{k-2} r_j$, indicating $h \cap v \cap \bigcap_{j=1}^{k-2} r_j \ne \emptyset$.
-
-\vgap
-
-It suffices to show that no two distinct pairs $(\bm{t}_1, v_1), (\bm{t}_2, v_2) \in \J(R_1',...,R_{k-2}')\times \dcross_V(\bm{t})$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},H,V)$.
-Assume the contrary, namely, both $(\bm{t}_1, v_1)$ and $(\bm{t}_2, v_2)$ are mapped to the tuple $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
-However, in this case, $\bm{t}_1[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-2] = v_1$.
-At the same time, $\bm{t}_2[j]$ is the trimmed rectangle of $r_j$, and $\bm{t}[k-2] = v_2$.
-We thus have $\bm{t}_1 = \bm{t}_2$ and $v_1 = v_2$, giving a contradiction.
-Therefore, we have $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq |\out|$.
 

### 2024-05-29 4.2.tex
yufei

@@
-\noindent \underline{\em Proof of Statement (2).} 
-First, we deviate a little to prove the following statement.
-For any rectangle $r\in \bigcup_{i = 1}^{k-2}$, if a segment $h\in H$ crosses $r$, then $h$ also crosses the trimmed rectangle of $r$, denoted as $r'$.
-Since the x-range of $r'$ is a subset of the x-range of $r$, it suffices to show that the y-coordinate of $h$ is in the range $[\ybot(r'), \ytop(r')]$.
-Let $h = [x_1,x_2]\times y$.
-Since $\ybot(r')$ is the y-coordinate of the lowest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ybot(r') \leq y$.
-Moreover, since $\ytop(r')$ is the y-coordinate of the highest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $y \leq \ytop(r')$.
-Therefore, $y \in [\ybot(r'), \ytop(r')]$ and $h$ crosses $r'$.
-Analogously, we can prove that if a segment $v\in V$ crosses $r$, then $v$ also crosses the trimmed rectangle of $r$.
-
-\vgap
-
-Now, we prove the first bullet of Statement (2).
-Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$.
-By the discussion above, $h$ and $v$ also cross the trimmed rectangles $r_1',..., r_{k-2}'$.
-Let $h = [x_1,x_2]\times y$, $v = x \times [y_1,y_2]$, and $p = (x,y)$ be the intersection point of $h$ and $v$.
-Since $h$ crosses $r_i'$, we have $y \in [\ybot(r_i'), \ytop(r_i')]$, and since $v$ crosses $r_i'$, we have $x \in [\xleft(r_i'), \xright(r_i')]$.
-Hence, the point $p$ in $r_i'$, for each $i \in [k-2]$.
-This implies that $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$, and $\bm{t}\in \J(R_1',...,R_{k-2}')$.
-
-\vgap
-
-Finally, we prove the second bullet of Statement (2).
-As we proved that $h$ and $v$ cross $r_i'$ for all $i \in [k-2]$, they must cross $\gbot(\bm{t})$ and $\gleft(\bm{t})$ as well.
-Hence, it suffices to show that both $h$ and $v$ cross $B_\bm{t}$.
-As proved in the last paragraph, the point $p = h \cap v$ is in $B_\bm{t}$.
-So, $h\cap B_\bm{t}$ and $v \cap B_\bm{t}$ are non-empty.
-Since $h$ (resp.
-$v$) crosses $r_i'$ for some $i \in [k-2]$, $r_i'$ does not cover any of the endpoints of $h$ (resp.
-$v$).
-Additonally, $B_\bm{t}$ is a subset of $r_i'$, so $B_\bm{t}$ does not cover any of the endpoints of $h$ or $v$.
-Therefore, $h$ and $v$ cross $B_\bm{t}$.
-
 \vgap

### 2024-05-29 msj.tex
yufei

--- 2024-05-29 msj.tex
@@
 
-\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} We consider, w.l.o.g., that each rectangle in the input $R_1 \cup ...
-\cup R_{k-2}$ is given a distinct integer ID in $[n]$.
-This allows us to create an array of size $n$ and allocate an array cell to each $r \in R_1 \cup ...
-\cup R_{k-2}$, with the property that the cell can be accessed by the ID of $r$ in constant time.
-
-\vgap
-
-To compute $R^*_1, R^*_2, ..., R^*_{k-2}$, we start by deriving $\maxtop(r')$ for each rectangle $r'$ in $R'_1 \cup  ...
-\cup R'_{k-2}$.
-For this purpose, first initialize $\maxtop(r') = -\infty$ for each such $r'$.
-Recall that $r'$ is the trimmed rectangle of some rectangle $r$ in $R_1 \cup ...
-\cup R_{k-2}$.
-We store $\maxtop(r')$ in the array cell allocated to $r$.
-
-\vgap
-
-Then, we scan $\J(R_1',...,R_{k-2}')$.
-For each tuple $\bm{t}$ therein, use $O(k)$ time to identify the rectangle $r' = \gbot(\bm{t})$, and then update in constant time $\maxtop(r')$ to the maximum between its current value and $\ytop(r')$.
-The scan requires $O(n + k \cdot \out)$ time.
-
-\vgap
-
-Finally, for each $i \in [k-2]$, we construct $R_i^*$ by collecting the top-sliced rectangle (see definition in \eqref{eqn:hv:type1:topsliced}) of every rectangle $r' \in R_i'$ with $\maxtop(r') \neq -\infty$.
-This step takes $O(|R_i'|)$ time for each $i \in [k-2]$, or $O(n)$ total time for all $i$.
+\vgap 
+\input{4.2.tex}
 

### 2024-05-29 4.2.tex

--- 2024-05-29 4.2.tex
@@
-\noindent \underline{\em Proof of Statement (2).} 
-First, we deviate a little to prove the following statement.
-For any rectangle $r\in \bigcup_{i = 1}^{k-2}$, if a segment $h\in H$ crosses $r$, then $h$ also crosses the trimmed rectangle of $r$, denoted as $r'$.
-Since the x-range of $r'$ is a subset of the x-range of $r$, it suffices to show that the y-coordinate of $h$ is in the range $[\ybot(r'), \ytop(r')]$.
-Let $h = [x_1,x_2]\times y$.
-Since $\ybot(r')$ is the y-coordinate of the lowest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $\ybot(r') \leq y$.
-Moreover, since $\ytop(r')$ is the y-coordinate of the highest segment in $H$ that crosses $r$, and $h$ crosses $r$, we have $y \leq \ytop(r')$.
-Therefore, $y \in [\ybot(r'), \ytop(r')]$ and $h$ crosses $r'$.
-Analogously, we can prove that if a segment $v\in V$ crosses $r$, then $v$ also crosses the trimmed rectangle of $r$.
+\vgap
+
+\noindent \underline{\em Proof of Statement (3).} Consider an arbitrary $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+First, we claim that $|\dcross_H(\bm{t})| \neq \emptyset$.
+Let $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.
+Let $h$ be the lowest segment that crosses $r$.
+We will prove the claim by showing that $h \in \dcross_H(\bm{t})$.
+More specifically, we will prove that $h$ crosses $B_{\bm{t}}$ and $r'$.
+From the first paragraph of the proof of Statement (2), we know that $h$ crosses $r'$ because $h$ crosses $r$.
+By how $r'$ is defined, the bottom edge of $r'$ is contained in $h$.
+Since $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$, and hence in $h$.
+Thus, $h \cap B_{\bm{t}}\neq \emptyset$.
+Since $h$ crosses $r'$, $r'$ does not contain any of the endpoints of $h$.
+As $r'$ covers $B_{\bm{t}}$, $B_{\bm{t}}$ does not contain any of the endpoints of $h$.
+Thus, $h$ crosses $B_{\bm{t}}$.
+This proves that $h \in \dcross_H(\bm{t})$, and hence $|\dcross_H(\bm{t})| \neq \emptyset$.
+By symmetry, $|\dcross_V(\bm{t})|$ is also nonempty.
+Equivalently, we have $|\dcross_H(\bm{t})| \geq 1$ and $|\dcross_V(\bm{t})| \geq 1$.
 
@@
 
-Now, we prove Statement (2).
-Since $(r_1,...,r_{k-2},h,v)\in \J_1$, both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$.
-By the discussion above, $h$ and $v$ also cross the trimmed rectangles $r_1',..., r_{k-2}'$.
-Therefore, the point $h \cap v$ is in $r_i'$, for each $i \in [k-2]$.
-This implies that $\bigcap_{i = 1}^{k-2}r_i' \neq \emptyset$, and $(r_1',..., r_{k-2}')\in \J(R_1',...,R_{k-2}')$.
-
-% The fact $(r_1,...,r_{k-2},h,v)\in \J_1$ indicates that both segments $h$ and $v$ cross all of the rectangles $r_1,..., r_{k-2}$ which means that for $i\in[k-2]$, the y-coorninate of $h$ is in the range $[\ybot(r_i'), \ytop(r_i')]$ and the x-coordinate of $v$ is in $[\xleft(r_i'), \xright(r_i')]$.
-Additionally, for each $i\in[k-2]$, the x-range (resp.
-y-range) of $r_i'$ is a subset of the x-range (resp.
-y-range) of $r_i$, so $h$ (resp.
-$v$) crosses the rectangles $r_i'$.
-Hence, the point $h \cap v$ is in $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$, and $h, v$ cross $B_{\bm{t}}$.
-The former implies that $\bm{t} \in \J(R_1',...,R_{k-2}')$, and the latter implies that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
-
-\vgap
-
-\noindent \underline{\em Proof of Statement (3).}
-For each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $h \in \dcross_H(\bm{t'})$, we will identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
-Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $v$ be the segment $v_\vdash$ of $\gleft(\bm{t'})$.
-Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
-Since $h$ crosses $B_{\bm{t'}}$ by the definition of $\dcross_H(\bm{t'})$, the $h\cap v$ is a point in $B_{\bm{t'}}$.
-Therefore, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
-he result tuple is unique because for any two distinct tuples $\bm{t'_1}$ and $\bm{t'_2}$ in $\J(R_1',...,R_{k-2}')$, there exists an integer $i \in [k-2]$ s.t.
-the full rectangles of $\bm{t'_1}[i]$ and $\bm{t'_2}[i]$ are different, which means the $i$-th components of their associating tuples $\bm{t_1}, \bm{t_2} \in \J(R_1,...,R_{k-2},H,V)$ are different.
-As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_H(\bm{t})| \leq \out$.
-
-\vgap
-
-Similarly, for each $\bm{t'}\in  \J(R_1',...,R_{k-2}')$ and each $v \in \dcross_V(\bm{t'})$, we can identify a unique tuple $\bm{t}$ in $\J(R_1,...,R_{k-2},H,V)$ defined as follows.
-Let $r_i$ be the full rectangle of $r_i'$ for $i \in [k]$, and $h$ be the segment $h_\bot$ of $\gbot(\bm{t'})$.
-Let $\bm{t} = (r_1,...,r_{k-2},h,v)$.
-Since $v$ crosses $B_{\bm{t'}}$, $h\cap v \in B_{\bm{t'}} \subseteq B_{\bm{t}}$, and $\bm{t}\in \J(R_1,...,R_{k-2},H,V)$.
-As a result, $\sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}  |\dcross_V(\bm{t})| \leq \out$.
+Therefore, to prove Statement (3),
+it suffices to show that \\
+\myeqn{
+    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}(|\dcross_H(\bm{t})|\cdot |\dcross_V(\bm{t})|) \leq \out. \label{eqn:lmm4.2.3}
+    } Recall that $\J^*$ is the set of tuples generated by the algorithm \textbf{generate-$\J^*$}(see below Example 4.3).
+By the discussion following the algorithm, we have $\J^*\subseteq \J(R_1,...,R_{k-2},H,V)$, which means $|\J^*|\leq \out$.
+Thus, we can prove Inequality~\eqref{eqn:lmm4.2.3} by showing that its left-hand side is equal to $|\J^*|$.
+ By the line 2-4 of \textbf{generate-$\J^*$}, the number of tuples generated in line 5 is exactly the left-hand side of Inequality~\eqref{eqn:lmm4.2.3}.
+Besides, it is obvious that every tuple $(r_1,...,r_{k-2},h,v)$ generated in line 5 of \textbf{generate-$\J^*$} is unique.
+Therefore, Inequality~\eqref{eqn:lmm4.2.3} holds.
+This completes the proof of Statement (3).

### 2024-05-29 msj.tex
yufei

@@
 
-Fix any rectangle $r \in R_1 \cup ... \cup R_{k-2}$ and any vertical segment $v$ from $V$ that crosses $r$. An argument symmetric to the above shows that $v$ must also cross $r'$.
+Fix any rectangle $r \in R_1 \cup ...
+\cup R_{k-2}$ and any vertical segment $v$ from $V$ that crosses $r$.
+An argument symmetric to the above shows that $v$ must also cross the trimmed rectangle $r'$.
 
@@
 \vgap
-\input{4.2.tex}
+
+\noindent \underline{\em Proof of Statement (3).} Consider an arbitrary $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+First, we claim that $|\dcross_H(\bm{t})| \neq \emptyset$.
+Let $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.
+Let $h$ be the lowest segment that crosses $r$.
+We will prove the claim by showing that $h \in \dcross_H(\bm{t})$.
+More specifically, we will prove that $h$ crosses $B_{\bm{t}}$ and $r'$.
+From the first paragraph of the proof of Statement (2), we know that $h$ crosses $r'$ because $h$ crosses $r$.
+By how $r'$ is defined, the bottom edge of $r'$ is contained in $h$.
+Since $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$, and hence in $h$.
+Thus, $h \cap B_{\bm{t}}\neq \emptyset$.
+Since $h$ crosses $r'$, $r'$ does not contain any of the endpoints of $h$.
+As $r'$ covers $B_{\bm{t}}$, $B_{\bm{t}}$ does not contain any of the endpoints of $h$.
+Thus, $h$ crosses $B_{\bm{t}}$.
+This proves that $h \in \dcross_H(\bm{t})$, and hence $|\dcross_H(\bm{t})| \neq \emptyset$.
+By symmetry, $|\dcross_V(\bm{t})|$ is also nonempty.
+Equivalently, we have $|\dcross_H(\bm{t})| \geq 1$ and $|\dcross_V(\bm{t})| \geq 1$.
+
+\vgap
+
+Therefore, to prove Statement (3),
+it suffices to show that \\
+\myeqn{
+    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}(|\dcross_H(\bm{t})|\cdot |\dcross_V(\bm{t})|) \leq \out. \label{eqn:lmm4.2.3}
+    } Recall that $\J^*$ is the set of tuples generated by the algorithm \textbf{generate-$\J^*$}(see below Example 4.3).
+By the discussion following the algorithm, we have $\J^*\subseteq \J(R_1,...,R_{k-2},H,V)$, which means $|\J^*|\leq \out$.
+Thus, we can prove Inequality~\eqref{eqn:lmm4.2.3} by showing that its left-hand side is equal to $|\J^*|$.
+ By the line 2-4 of \textbf{generate-$\J^*$}, the number of tuples generated in line 5 is exactly the left-hand side of Inequality~\eqref{eqn:lmm4.2.3}.
+Besides, it is obvious that every tuple $(r_1,...,r_{k-2},h,v)$ generated in line 5 of \textbf{generate-$\J^*$} is unique.
+Therefore, Inequality~\eqref{eqn:lmm4.2.3} holds.
+This completes the proof of Statement (3).
 

### 2024-05-29 msj.tex
yufei

--- 2024-05-29 msj.tex
@@
 Let $p$ be the bottom-left corner of $B_\bm{t}$.
-Define $i$ and $j$ to be values in $[k-2]$ such that $\bm{t}[i] = \gbot(\bm{t})$ and $\bm{t}[j] = \gleft(\bm{t})$; note that the values of $i$ and $j$ can be the same.
+Define $i$ and $j$ to be values in $[k-2]$ such that $\bm{t}[i] = \gbot(\bm{t})$ and $\bm{t}[j] = \gleft(\bm{t})$; note that $i$ and $j$ can be the same.
 Point $p$ is the intersection of the bottom edge of $\bm{t}[i]$ and the left edge of $\bm{t}[j]$.
@@
 \vgap
-\input{4.2.tex}
+
+\noindent \underline{\em Proof of Statement (3).} We first prove that $\dcross_H(\bm{t}) \neq \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+Set $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.
+Define $h$ as the lowest segment crossing $r$ (note that $h$ definitely exists because otherwise $r$ has no trimmed rectangle).
+We will show $h \in \dcross_H(\bm{t})$, and hence $\dcross_H(\bm{t})$ cannot be empty.
+For this purpose, we need to prove that $h$ crosses both $B_{\bm{t}}$ and $r'$.
+However, $h$ crossing $r$ implies $h$ crossing $r'$, as has already been argued in our proof of statement (2).
+It thus remains to show that $h$ crosses $B_\bm{t}$.
+
+\vgap
+
+By the way $r'$ is defined, the bottom edge of $r'$ must be contained in $h$.
+Because $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$ and hence also contained in $h$.
+This means that $h \cap B_{\bm{t}}\neq \emptyset$.
+On the other hand, the x-range of $h$ must contain that of $r'$ (because $h$ crosses $r'$), which in turn must contain that of $B_\bm{t}$ (because $r'$ covers $B_\bm{t}$).
+Thus, the x-range of $h$ covers that of $B_\bm{t}$.
+Combining this with $h \cap B_{\bm{t}}\neq \emptyset$ shows that $h$ must cross $B_\bm{t}$.
+
+\vgap
+
+A symmetric argument shows that $\dcross_V(\bm{t}) \ne \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+
+%In other words, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ must have size at least 1.
+
+\vgap
+
+Next, we will prove 
+\myeqn{
+    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})| \cdot |\dcross_V(\bm{t})| \leq \out. \label{eqn:hv:type1:properties:proof-help1}
+} 
+which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$.
+Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}).
+This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}.
+By statement (2), the $\J^*$ computed must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$.
+This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).
 

### 2024-05-30 4.3.tex
4.3(1)

--- 2024-05-29 4.3.tex
@@
-First, we prove (1).
-For any $i \in [k-2]$, $r_i^*\in R_i^*$, and $h \in \cross_H(r_i^*)$, we will identify a unique tuple $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$ such that $\bm{t}[k-1] = h$, and $\bm{t}[i]$ is the full rectangle of $r_i'$, whose top-slice is $r_i^*$.
+% First, we prove (1).
+For any $i \in [k-2]$, $r_i^*\in R_i^*$, and $h \in \cross_H(r_i^*)$, we will identify a unique tuple $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$ such that $\bm{t}[k-1] = h$, and $\bm{t}[i]$ is the full rectangle of $r_i'$, whose top-sliced rectangle is $r_i^*$.
 Let $\maxtop(r_i') = \ytop(B_\bm{t'})$ for some tuple $\bm{t'} \in \J(R_1',...,R_{k-2}')$.
@@
 
-Next, we prove (2).
+% Next, we prove (2).
 Let $r' = [x_1,x_2]\times [y_1,y_2]$, and $r^* = [x_1,x_2] \times [y_1,\maxtop(r')]$.
-Since $\gbot(\bm{t}) = r'$, $\dcross_H(\bm{t})$ contains exactly every segment $h \in H$ that crosses the rectangle $[x_1,x_2]\times [y_1, \ytop(B_\bm{t})]$, where $\ytop(B_{\bm{t}})\leq \maxtop(r')$.
+Since $\gbot(\bm{t}) = r'$, $\dcross_H(\bm{t})$ contains exactly the set of segments segment $h \in H$ that cross the rectangle $[x_1,x_2]\times [y_1, \ytop(B_\bm{t})]$, where $\ytop(B_{\bm{t}})\leq \maxtop(r')$.
 Therefore, $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
 Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+
+\noindent \underline{\em Proof of Statement (1).}
+We will map each $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ such that $\dcross_H(\bm{t}) = \cross_H(r^*)$.
+Then, by Lemma~\ref{lmm:hv:type1:properties} (3), we have $\sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out$.
+
+\vgap
+
+Now we explain the mapping. Let $r^*$ be the top-sliced rectangle of $r'$. By how $\maxtop(r')$ is defined, we can identify a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ such that 
+\begin{itemize}
+    \item $\gbot(\bm{t}) = r'$,
+    \item $\maxtop(r') = \ytop(B_{\bm{t}})$.
+\end{itemize}
+Let $r' = [x_1,x_2]\times [y_1,y_2]$, and $r^* = [x_1,x_2]\times [y_1,\maxtop(r')]$.
+Since $\gbot(\bm{t}) = r'$, we can represent $B_{\bm{t}}$ as $[\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1, \maxtop(r')]$.
+We will prove that a segment $h\in \cross_H(r^*)$ if and only if $h\in \dcross_H(\bm{t})$.
+
+\vgap
+
+We show the ``if'' direction as follows.
+Let $h\in \cross_H(r^*)$.
+First, we prove that $h$ crosses $r'$.
+Since $h$ crosses $r^*$, $h \cap r^* \neq \emptyset$.
+As $r^* \subseteq r'$, $r'\cap h \neq \emptyset$.
+It suffices to show that $r'$ does not containe any endpoints of $h$.
+Suppose to the contrary that $r'$ contains at least an endpoint of $h$. W.l.o.g, assume that the left endpoint of $h$ is in $r'$.
+Then, $\xleft(h)\in[x_1,x_2]$.
+Since $h\cap r^*\neq \emptyset$, the y-coordinate of $h$ is in $[y_1,\maxtop(r')]$.
+Since $r^*$ does not contain the any endpoint of $h$, we have $\xleft(h)<x_1<x_2<\xright(h)$,
+which is a contradiction. Therefore, $h$ crosses $r'$.
+
+\vgap
+
+Then, we prove that $h$ crosses $B_{\bm{t}}$.
+Since $h$ crosses $r'$ and $B_{\bm{t}}\subseteq r'$, $B_{\bm{t}}$ does not contain any endpoints of $h$.
+Let the y-coordinate of $h$ be $y$.
+Let $p = (\xleft(B_{\bm{t}}), y)$.
+We will prove that $p \in h$ and $p \in B_{\bm{t}}$, which means $h\cap B_{\bm{t}} \neq \emptyset$ and $h$ crosses $B_{\bm{t}}$.
+Since $\gbot(\bm{t}) = r'$, the bottom segment of $B_{\bm{t}}$ is contained in $r'$.
+Therefore,
+$\xleft(B_{\bm{t}})\in [x_1,x_2]$.
+We proved in the last paragraph that $\xleft(h)<x_1<x_2<\xright(h)$, so  $\xleft(B_{\bm{t}}) \in (\xleft(h),\xright(h))$, and $p\in h$.
+To prove that $p\in B_{\bm{t}}$, it suffices to show that $y$ is in the y-range of $B_{\bm{t}}$, i.e.
+$y \in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$.
+ Since $h$ intersects $r^*$, $y \in [y_1 , \maxtop(r')]$, where $\maxtop(r') = \ytop(B_{\bm{t}})$.
+Since $\gbot(\bm{t}) = r'$, $y_1 = \ybot(B_{\bm{t}})$.
+Hence, $y\in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$, and $p\in B_{\bm{t}}$.
+
+\vgap
+
+Conversely, let $h\in \dcross_H(\bm{t})$, which means that $h$ crosses $r'$ and $B_{\bm{t}}$.
+Thus, $r'$ does not contain any endpoints of $h$.
+Since $r^*\subseteq r'$, $r^*$ does not contain any endpoints of $h$ neither.
+Next, we prove that $B_{\bm{t}} \subseteq r^*$.
+Since $h \cap B_{\bm{t}}\neq \emptyset$, this statement leads to $h \cap r^* \neq \emptyset$ and $h$ crosses $r^*$.
+Recall that $r^* = [x_1,x_2], [y_1, \maxtop(r')]$, and $B_{\bm{t}} = [\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1,\maxtop(r')]$.
+Since $B_{\bm{t}}$ is a subset of $r'$, we have $\xleft(B_{\bm{t}}) \subseteq [x_1,x_2]$.
+Therefore, $B_{\bm{t}} \subseteq r^*$, and $h\in \cross_H(r^*)$.
+
+\vgap
+
+It remains to show that we will not map two distinct rectangle $r_1^*, r_2^*\in \bigcup_{i = 1}^{k-2}R_i^*$ to the same tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+Suppose to the contrary that $r_1^*, r_2^*$ are mapped to the same tuple $\bm{t}$.
+Let $r_1^*, r_2^*$ be the top-sliced rectangles of $r_1'$ and $r_2'$ respectively.
+Since $r_1^*$ and $r_2^*$ are disinct, $r_1'$ is not the same as $r_2'$.
+By our mapping, $\gbot(B_{\bm{t}})$ is $r_1'$ and $r_2'$.
+Since $\gbot(B_{\bm{t}})$ is unique for each tuple $\bm{t}$, we have a contradiction.
+Therefore, each rectangle $r^*\in \bigcup_{i = 1}^{k-2}R_i^*$ is mapped to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+
+\vgap
+
+\noindent \underline{\em Proof of Statement (2).}

### 2024-05-30 4.3.tex

@@
 
-It remains to show that we will not map two distinct rectangle $r_1^*, r_2^*\in \bigcup_{i = 1}^{k-2}R_i^*$ to the same tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
-Suppose to the contrary that $r_1^*, r_2^*$ are mapped to the same tuple $\bm{t}$.
-Let $r_1^*, r_2^*$ be the top-sliced rectangles of $r_1'$ and $r_2'$ respectively.
-Since $r_1^*$ and $r_2^*$ are disinct, $r_1'$ is not the same as $r_2'$.
-By our mapping, $\gbot(B_{\bm{t}})$ is $r_1'$ and $r_2'$.
-Since $\gbot(B_{\bm{t}})$ is unique for each tuple $\bm{t}$, we have a contradiction.
-Therefore, each rectangle $r^*\in \bigcup_{i = 1}^{k-2}R_i^*$ is mapped to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
-
-\vgap
-
 \noindent \underline{\em Proof of Statement (2).}

### 2024-05-30 msj.tex

@@
 Define $h$ as the lowest segment crossing $r$ (note that $h$ definitely exists because otherwise $r$ has no trimmed rectangle).
-We will show $h \in \dcross_H(\bm{t})$, and hence $\dcross_H(\bm{t})$ cannot be empty.
-For this purpose, we need to prove that $h$ crosses both $B_{\bm{t}}$ and $r'$.
+We will show that $h \in \dcross_H(\bm{t})$, and hence $\dcross_H(\bm{t})$ cannot be empty.
+For this purpose, we should explain why $h$ crosses both $B_{\bm{t}}$ and $r'$.
 However, $h$ crossing $r$ implies $h$ crossing $r'$, as has already been argued in our proof of statement (2).
-It thus remains to show that $h$ crosses $B_\bm{t}$.
+It remains to show that $h$ crosses $B_\bm{t}$.
 
@@
 Thus, the x-range of $h$ covers that of $B_\bm{t}$.
-Combining this with $h \cap B_{\bm{t}}\neq \emptyset$ shows that $h$ must cross $B_\bm{t}$.
-
-\vgap
-
-A symmetric argument shows that $\dcross_V(\bm{t}) \ne \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+Combining this with $h \cap B_{\bm{t}}\neq \emptyset$ yields that $h$ must cross $B_\bm{t}$.
+
+\vgap
+
+A symmetric argument assures us that $\dcross_V(\bm{t}) \ne \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
 

### 2024-05-30 4.3.tex
4.3
yufei

--- 2024-05-29 4.3.tex
@@
-First, we prove (1).
+% First, we prove (1).
 For any $i \in [k-2]$, $r_i^*\in R_i^*$, and $h \in \cross_H(r_i^*)$, we will identify a unique tuple $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$ such that $\bm{t}[k-1] = h$, and $\bm{t}[i]$ is the full rectangle of $r_i'$, whose top-sliced rectangle is $r_i^*$.
@@
 
-Next, we prove (2).
+% Next, we prove (2).
 Let $r' = [x_1,x_2]\times [y_1,y_2]$, and $r^* = [x_1,x_2] \times [y_1,\maxtop(r')]$.
@@
 Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+
+\noindent \underline{\em Proof of Statement (2).}

### 2024-05-30 msj.tex
yufei

@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.} \\
+\underline{\em Proof of Statement (1).}
+We will map each $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t}) = \cross_H(r^*)$.
+After this done, we can derive $\sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out$, where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
+
+\vgap
+
+Now, we explain the mapping.
+Let $r^*$ be the top-sliced rectangle of $r'$.
+By how $\maxtop(r')$ is defined, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_{\bm{t}})$.
+More specifically, if $r' = [x_1,x_2]\times [y_1,y_2]$, then $r^* = [x_1,x_2]\times [y_1,\maxtop(r')]$.
+We will prove that a segment $h\in \cross_H(r^*)$ if and only if $h\in \dcross_H(\bm{t})$.
+
+\vgap
+
+As $\gbot(\bm{t}) = r'$, the rectangle $B_{\bm{t}}$ can be represented as $[\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1, \maxtop(r')]$.
+
+\vgap
+
+We show the ``if'' direction as follows.
+Let $h\in \cross_H(r^*)$.
+First, we prove that $h$ crosses $r'$.
+Since $h$ crosses $r^*$, $h \cap r^* \neq \emptyset$.
+As $r^* \subseteq r'$, $r'\cap h \neq \emptyset$.
+It suffices to show that $r'$ does not containe any endpoints of $h$.
+Suppose to the contrary that $r'$ contains at least an endpoint of $h$. W.l.o.g, assume that the left endpoint of $h$ is in $r'$.
+Then, $\xleft(h)\in[x_1,x_2]$.
+Since $h\cap r^*\neq \emptyset$, the y-coordinate of $h$ is in $[y_1,\maxtop(r')]$.
+Since $r^*$ does not contain the any endpoint of $h$, we have $\xleft(h)<x_1<x_2<\xright(h)$,
+which is a contradiction. Therefore, $h$ crosses $r'$.
+
+\vgap
+
+Then, we prove that $h$ crosses $B_{\bm{t}}$.
+Since $h$ crosses $r'$ and $B_{\bm{t}}\subseteq r'$, $B_{\bm{t}}$ does not contain any endpoints of $h$.
+Let the y-coordinate of $h$ be $y$.
+Let $p = (\xleft(B_{\bm{t}}), y)$.
+We will prove that $p \in h$ and $p \in B_{\bm{t}}$, which means $h\cap B_{\bm{t}} \neq \emptyset$ and $h$ crosses $B_{\bm{t}}$.
+Since $\gbot(\bm{t}) = r'$, the bottom segment of $B_{\bm{t}}$ is contained in $r'$.
+Therefore,
+$\xleft(B_{\bm{t}})\in [x_1,x_2]$.
+We proved in the last paragraph that $\xleft(h)<x_1<x_2<\xright(h)$, so  $\xleft(B_{\bm{t}}) \in (\xleft(h),\xright(h))$, and $p\in h$.
+To prove that $p\in B_{\bm{t}}$, it suffices to show that $y$ is in the y-range of $B_{\bm{t}}$, i.e.
+$y \in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$.
+ Since $h$ intersects $r^*$, $y \in [y_1 , \maxtop(r')]$, where $\maxtop(r') = \ytop(B_{\bm{t}})$.
+Since $\gbot(\bm{t}) = r'$, $y_1 = \ybot(B_{\bm{t}})$.
+Hence, $y\in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$, and $p\in B_{\bm{t}}$.
+
+\vgap
+
+Conversely, let $h\in \dcross_H(\bm{t})$, which means that $h$ crosses $r'$ and $B_{\bm{t}}$.
+Thus, $r'$ does not contain any endpoints of $h$.
+Since $r^*\subseteq r'$, $r^*$ does not contain any endpoints of $h$ neither.
+Next, we prove that $B_{\bm{t}} \subseteq r^*$.
+Since $h \cap B_{\bm{t}}\neq \emptyset$, this statement leads to $h \cap r^* \neq \emptyset$ and $h$ crosses $r^*$.
+Recall that $r^* = [x_1,x_2], [y_1, \maxtop(r')]$, and $B_{\bm{t}} = [\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1,\maxtop(r')]$.
+Since $B_{\bm{t}}$ is a subset of $r'$, we have $\xleft(B_{\bm{t}}) \subseteq [x_1,x_2]$.
+Therefore, $B_{\bm{t}} \subseteq r^*$, and $h\in \cross_H(r^*)$.
+
+\vgap
 

### 2024-05-30 msj.tex

@@
 
+It remains to show that we will not map two distinct rectangle $r_1^*, r_2^*\in \bigcup_{i = 1}^{k-2}R_i^*$ to the same tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+Suppose to the contrary that $r_1^*, r_2^*$ are mapped to the same tuple $\bm{t}$.
+Let $r_1^*, r_2^*$ be the top-sliced rectangles of $r_1'$ and $r_2'$ respectively.
+Since $r_1^*$ and $r_2^*$ are disinct, $r_1'$ is not the same as $r_2'$.
+By our mapping, $\gbot(B_{\bm{t}})$ is $r_1'$ and $r_2'$.
+Since $\gbot(B_{\bm{t}})$ is unique for each tuple $\bm{t}$, we have a contradiction.
+Therefore, each rectangle $r^*\in \bigcup_{i = 1}^{k-2}R_i^*$ is mapped to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+
 \input{4.3.tex}

### 2024-05-30 msj.tex
yufei

--- 2024-05-30 msj.tex
@@
 }
-See Figure~\ref{fig:guard} for an illustration.
+See Figure~\ref{fig:guard} for an illustration. It is worth mentioning that since a horizontal segment $h$ is a degenerated rectangle, notations such as $\xleft(h)$ and $\xright(h)$ are well-defined.
 
@@
     \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
+    \label{eqn:hv:type1:maxtop}
 }
@@
 This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}.
-By statement (2), the $\J^*$ computed must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$.
+By statement (2), the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$.
 This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.} \\
+\underline{\em Proof of Statement (1).} We will map each $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t}) = \cross_H(r^*)$.
+After this, we can then derive $\sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out$, where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
+
+\vgap
+
+The mapping is designed as follows.
+Consider an arbitrary $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$.
+Recall that $r^*$ is the top-sliced of some rectangle $r'$.
+Specifically, if $r' = [x_1,x_2]\times [y_1,y_2]$, then $r^* = [x_1,x_2]\times [y_1,\maxtop(r')]$.
+By the definition of $\maxtop(r')$ in \eqref{eqn:hv:type1:topsliced}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_{\bm{t}})$.
+We map $r^*$ to $\bm{t}$.
+
+\vgap
+
+%As $\gbot(\bm{t}) = r'$, the rectangle $B_{\bm{t}}$ can be represented as $[\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1, \maxtop(r')]$.
+
+Next, we will prove that a segment $h \in \cross_H(r^*)$ if and only if $h \in \dcross_H(\bm{t})$.
+Let us start with the ``only-if direction'': assuming $h \in \cross_H(r^*)$, we aim to show $h \in \dcross_H(\bm{t})$.
+For this purpose, we need to show that $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
+\myitems {
+    \item First, we explain why $h$ crosses $r'$.
+Because $r'$ covers $r^*$, by the fact $h$ crossing $r^*$, we know that $h$ must intersect with $r'$.
+On the other hand, as $h$ crosses $r^*$, it must hold that $\xleft(h) < x_1 < x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$).
+As $[x_1, x_2]$ is also the x-range of $r'$, using the fact of $h \cap r' \ne \emptyset$, we can assert that $h$ crosses $r'$.
+
+    \vgap
+
+    \item Then, we explain why $h$ crosses $B_{\bm{t}}$.
+Note that because $h$ crosses $r'$ (which we just proved) and $B_{\bm{t}}\subseteq r'$, we know that $B_{\bm{t}}$ cannot contain any endpoint of $h$.
+Let $y_h$ be the y-coordinate of $h$, and take the point $p = (\xleft(B_{\bm{t}}), y_h)$.
+We will prove $p \in h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$ and, hence, $h$ must cross $B_{\bm{t}}$ (because no endpoint of $h$ falls in $B_\bm{t}$, as mentioned).
+
+    \vgap
+
+    Since $\gbot(\bm{t}) = r'$, the bottom segment of $B_{\bm{t}}$ must be contained in $r'$.
+Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$.
+As proved earlier, $\xleft(h) < x_1 < x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting $p \in h$.
+To prove $p \in B_{\bm{t}}$, it remains to show $y_h \in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$.
+Since $h$ intersects $r^*$, $y \in [y_1 , \maxtop(r')]$ --- recall that $[y_1, \maxtop(r')]$ is the y-range of $r^*$ --- where $\maxtop(r') = \ytop(B_{\bm{t}})$.
+Since $\gbot(\bm{t}) = r'$, $y_1 = \ybot(B_{\bm{t}})$.
+Hence, $y\in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$, and $p\in B_{\bm{t}}$.
+
+}
+
+\vgap
+
+Conversely, let $h\in \dcross_H(\bm{t})$, which means that $h$ crosses $r'$ and $B_{\bm{t}}$.
+Thus, $r'$ does not contain any endpoints of $h$.
+Since $r^*\subseteq r'$, $r^*$ does not contain any endpoints of $h$ neither.
+Next, we prove that $B_{\bm{t}} \subseteq r^*$.
+Since $h \cap B_{\bm{t}}\neq \emptyset$, this statement leads to $h \cap r^* \neq \emptyset$ and $h$ crosses $r^*$.
+Recall that $r^* = [x_1,x_2], [y_1, \maxtop(r')]$, and $B_{\bm{t}} = [\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1,\maxtop(r')]$.
+Since $B_{\bm{t}}$ is a subset of $r'$, we have $\xleft(B_{\bm{t}}) \subseteq [x_1,x_2]$.
+Therefore, $B_{\bm{t}} \subseteq r^*$, and $h\in \cross_H(r^*)$.
+
+\vgap
+
+It remains to show that we will not map two distinct rectangle $r_1^*, r_2^*\in \bigcup_{i = 1}^{k-2}R_i^*$ to the same tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+Suppose to the contrary that $r_1^*, r_2^*$ are mapped to the same tuple $\bm{t}$.
+Let $r_1^*, r_2^*$ be the top-sliced rectangles of $r_1'$ and $r_2'$ respectively.
+Since $r_1^*$ and $r_2^*$ are disinct, $r_1'$ is not the same as $r_2'$.
+By our mapping, $\gbot(B_{\bm{t}})$ is $r_1'$ and $r_2'$.
+Since $\gbot(B_{\bm{t}})$ is unique for each tuple $\bm{t}$, we have a contradiction.
+Therefore, each rectangle $r^*\in \bigcup_{i = 1}^{k-2}R_i^*$ is mapped to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
 

### 2024-05-30 msj.tex
yufei

@@
 
-%As $\gbot(\bm{t}) = r'$, the rectangle $B_{\bm{t}}$ can be represented as $[\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1, \maxtop(r')]$.
-
 Next, we will prove that a segment $h \in \cross_H(r^*)$ if and only if $h \in \dcross_H(\bm{t})$.
@@
 To prove $p \in B_{\bm{t}}$, it remains to show $y_h \in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$.
-Since $h$ intersects $r^*$, $y \in [y_1 , \maxtop(r')]$ --- recall that $[y_1, \maxtop(r')]$ is the y-range of $r^*$ --- where $\maxtop(r') = \ytop(B_{\bm{t}})$.
-Since $\gbot(\bm{t}) = r'$, $y_1 = \ybot(B_{\bm{t}})$.
-Hence, $y\in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$, and $p\in B_{\bm{t}}$.
-
-}
-
-\vgap
+Since $h$ intersects $r^*$ and $[y_1, \maxtop(r')]$ is the y-range of $r^*$, we know $y_h \in [y_1 , \maxtop(r')]$.
+Because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$.
+Therefore, $y_h \in [y_1 , \maxtop(r')]$ tells us that the point $p$ --- which is $(\xleft(B_{\bm{t}}), y_h)$ --- must fall in $B_{\bm{t}}$.
+}
 

### 2024-05-30 msj.tex
yufei

@@
 
-    Since $\gbot(\bm{t}) = r'$, the bottom segment of $B_{\bm{t}}$ must be contained in $r'$.
+    Since $\gbot(\bm{t}) = r'$, the bottom edge of $B_{\bm{t}}$ must be contained in $r'$.
 Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$.
 As proved earlier, $\xleft(h) < x_1 < x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting $p \in h$.
-To prove $p \in B_{\bm{t}}$, it remains to show $y_h \in [\ybot(B_{\bm{t}}), \ytop(B_{\bm{t}})]$.
-Since $h$ intersects $r^*$ and $[y_1, \maxtop(r')]$ is the y-range of $r^*$, we know $y_h \in [y_1 , \maxtop(r')]$.
+
+    \vgap
+
     Because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$.
-Therefore, $y_h \in [y_1 , \maxtop(r')]$ tells us that the point $p$ --- which is $(\xleft(B_{\bm{t}}), y_h)$ --- must fall in $B_{\bm{t}}$.
+To prove $p \in B_{\bm{t}}$, it remains to show $y_h \in [y_1, \maxtop(r')]$.
+This is true because $h$ intersects $r^*$ and $[y_1, \maxtop(r')]$ is the y-range of $r^*$.
 }

### 2024-05-30 msj.tex
yufei

@@
 
-Conversely, let $h\in \dcross_H(\bm{t})$, which means that $h$ crosses $r'$ and $B_{\bm{t}}$.
-Thus, $r'$ does not contain any endpoints of $h$.
-Since $r^*\subseteq r'$, $r^*$ does not contain any endpoints of $h$ neither.
-Next, we prove that $B_{\bm{t}} \subseteq r^*$.
-Since $h \cap B_{\bm{t}}\neq \emptyset$, this statement leads to $h \cap r^* \neq \emptyset$ and $h$ crosses $r^*$.
-Recall that $r^* = [x_1,x_2], [y_1, \maxtop(r')]$, and $B_{\bm{t}} = [\xleft(B_{\bm{t}}), \xright(B_{\bm{t}})]\times [y_1,\maxtop(r')]$.
-Since $B_{\bm{t}}$ is a subset of $r'$, we have $\xleft(B_{\bm{t}}) \subseteq [x_1,x_2]$.
-Therefore, $B_{\bm{t}} \subseteq r^*$, and $h\in \cross_H(r^*)$.
-
-\vgap
+Our argument now proceeds to prove the ``if direction'': assuming $h \in \dcross_H(\bm{t})$, we aim to show $h \in \cross_H(r^*)$.
+Recall that $r^* = [x_1,x_2] \times [y_1, \maxtop(r')]$.
+To prove that $h$ crosses $r^*$, we will show $\xleft(h) \le x_1 \le x_2 \le \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
+\myitems{
+    \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $r'$. As $[x_1, x_2]$ is the x-range of $r'$, we know $\xleft(h) \le x_1 \le x_2 \le \xright(h)$.
+
+    \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$.
+As $[y_1, \maxtop(r')]$ is the y-range of $B_\bm{t}$ (because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$), we know $y_h \in [y_1, \maxtop(r')]$.
+}
 

### 2024-05-30 msj.tex
yufei

@@
 
+    \vgap
+
     \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$.
@@
 
-It remains to show that we will not map two distinct rectangle $r_1^*, r_2^*\in \bigcup_{i = 1}^{k-2}R_i^*$ to the same tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
-Suppose to the contrary that $r_1^*, r_2^*$ are mapped to the same tuple $\bm{t}$.
-Let $r_1^*, r_2^*$ be the top-sliced rectangles of $r_1'$ and $r_2'$ respectively.
-Since $r_1^*$ and $r_2^*$ are disinct, $r_1'$ is not the same as $r_2'$.
-By our mapping, $\gbot(B_{\bm{t}})$ is $r_1'$ and $r_2'$.
-Since $\gbot(B_{\bm{t}})$ is unique for each tuple $\bm{t}$, we have a contradiction.
-Therefore, each rectangle $r^*\in \bigcup_{i = 1}^{k-2}R_i^*$ is mapped to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
-
-\input{4.3.tex}
+It remains to show that no two distinct rectangles $r_1^*, r_2^* \in \bigcup_{i = 1}^{k-2}R_i^*$ can be mapped to the same tuple in $\J(R_1',...,R_{k-2}')$.
+Assume, on the contrary, that $r_1^*$ and $r_2^*$ are mapped to the same tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+Suppose that $r_1^*$ (resp., $r_2^*$) is the top-sliced rectangle of $r_1'$ (resp., $r_2'$).
+Under our mapping, it must be true that  $r_1' = r_2' = \gbot(B_{\bm{t}})$.
+However, the distinctness of $r_1^*$ and $r_2^*$ requires $r_1' \ne r_2'$, thus giving a contradiction.
+
+%\input{4.3.tex}
 

### 2024-05-30 msj.tex
yufei

@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of $N$ intervals in $O(N)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log N + K)$ time, where $K$ is the number of intervals reported.
+It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log N)$ time.
+Each interval insertion or deletion on $S$ can be supported in $O(\log N)$ time.
+
+\vgap
+
+Let us first consider the detection version of Problem $\mathscr{A}$.
+Each point $p = (x, y)$ of $P$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+We start by sorting the set of y-coordinates defined by all the points and rectangles.
+
+\vgap
+
+Next, scan (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
+Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in an interval tree $\T$.
+Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
+When $\ell$ hits a point $p = (x, y)$ of $P$, search $\T$ to determine if any interval in $S_\ell$ contains the value $x$.
+If so, point $p$ is covered by at least one rectangle in $R$; otherwise, it is not.
+The overall running time is $O(n \log n)$.
+
+\vgap
+
+The algorithm for the reporting version of Problem $\mathscr{A}$ is similar.
+The only difference is that, when $\ell$ hits a point $p = (x, y)$ of $P$, we use $\T$ to report all the intervals in $S_\ell$ that contain $x$; the cost is $O(n \log n + K_p)$, where $K_p$ is the number of such intervals.
+Every interval corresponds to a rectangle in $R$ that contains $p$.
+The total running time is $O(n \log n + \sum_p K_p) = O(n \log n + \out)$.
 

### 2024-05-30 msj.tex
yufei

@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of $N$ intervals in $O(N)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log N + K)$ time, where $K$ is the number of intervals reported.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of $N$ intervals in $\real$ using $O(N)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log N + K)$ time, where $K$ is the number of intervals reported.
 It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log N)$ time.
@@
 
-Next, scan (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
 At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.}
-
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each vertical segment $v = x \times [y_1, y_2]$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+We start by sorting the set of y-coordinates defined by all the horizontal and vertical segments.
+
+\vgap
+
+Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+At all times, maintain the set $V_\ell$ of rectangles in $R$ that intersect with $\ell$.
+Let $S_\ell$ be the set of x-coordinates of the segments in $V_\ell$; we store $S_\ell$ in a binary search tree (BST) $\T$.
+Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
+When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$.
+If $x' \le x_2$, then then we output a pair $(h, p)$, where $p$ is the point $(x', y)$.
+The overall running time is $O(n \log n)$.
+
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Let $S$ be a set of $N$ intervals in $\real$.
 

### 2024-05-30 4.3.tex
4.3
yufei

@@
-% First, we prove (1).
-For any $i \in [k-2]$, $r_i^*\in R_i^*$, and $h \in \cross_H(r_i^*)$, we will identify a unique tuple $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$ such that $\bm{t}[k-1] = h$, and $\bm{t}[i]$ is the full rectangle of $r_i'$, whose top-sliced rectangle is $r_i^*$.
-Let $\maxtop(r_i') = \ytop(B_\bm{t'})$ for some tuple $\bm{t'} \in \J(R_1',...,R_{k-2}')$.
-Let $\bm{t'} = (r'_1,...,r'_{k-2})$, and let $r_i$ be the full rectangle of $r_i'$ for $i \in [k-2]$.
-Since $h$ crosses $r_i^*$, and $B_\bm{t'} \subseteq r_i^*$, $h$ crosses $B_{\bm{t'}}$.
-Let $v$ be the segment $v_\vdash$ of $\gleft(\bm{t'})$.
-The intersection of $h$ and $v$ is a point on the left boundary of $B_{\bm{t'}}$, which means that $h\cap v \in B_{\bm{t'}} \subseteq \cap_{i = 1}^{k-2}r_i$.
-Therefore, $\bm{t} = (r_1,...,r_{k-2},h,v)\in \J(R_1,...,R_{k-2},H,V)$ is the desired tuple, and we have $\sum_{i \in [k-2]}\sum_{r_i^*\in R_i^*} |\cross_H(r_i^*)| \leq \out$.
+\noindent \underline{\em Proof of Statement (2).} 
+Let $r' = [x_1,x_2] \times [y_1,y_2]$.
+Then, $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$.
+Let $h$ be an arbitrary segment in $\dcross_H(\bm{t})$, whose y-coordinate is $y_h$.
+We will prove that $h$ crosses $r^*$, which leads to $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
+It suffies to show that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$.
+\begin{itemize}
+    \item Since $h$ crosses $r' = [x_1,x_2]\times [y_1,y_2]$, we have $\xleft(h) < x_1 \le x_2 < \xright(h)$.
+    \item Since $h$ crosses $B_{\bm{t}}$, we have $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+By the definition of $\maxtop(r')$ (see Definition \eqref{eqn:hv:type1:maxtop}), and the fact that $r' = \gbot(B_{\bm{t}})$, we have $\ytop(B_{\bm{t}}) \leq \maxtop(r')$.
+Therefore, $y_h \in [y_1, \maxtop(r')]$.
+\end{itemize}
 
-% Next, we prove (2).
-Let $r' = [x_1,x_2]\times [y_1,y_2]$, and $r^* = [x_1,x_2] \times [y_1,\maxtop(r')]$.
-Since $\gbot(\bm{t}) = r'$, $\dcross_H(\bm{t})$ contains exactly the set of segments segment $h \in H$ that cross the rectangle $[x_1,x_2]\times [y_1, \ytop(B_\bm{t})]$, where $\ytop(B_{\bm{t}})\leq \maxtop(r')$.
-Therefore, $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
-Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+\vgap
 
-\noindent \underline{\em Proof of Statement (2).}
+Then, we will prove that each segment $h \in \cross_H(r^*)$ satisfying that the y-coordinate $h$ is no more than $\ytop(B_{\bm{t}})$ is in $\dcross_H(\bm{t})$.
+This claim leads to the last sentence of Statement (3).
+Consider an arbitrary $h\in \cross_H(r^*)$, whose y-coordinate $y_h$ satisfies that $y_h \le \ytop(B_{\bm{t}})$.
+We will prove that $h$ crosses $r'$ and $B_{\bm{t}}$, and therefore $h \in \dcross_H(\bm{t})$.
+
+\vgap
+
+First, we prove that $h$ crosses $r'$.
+It suffices to show that $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$.
+Both conditions are satisfied since $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$.
+Then, we prove that $h$ crosses $B_{\bm{t}}$.
+It suffices to show that $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+\begin{itemize}
+    \item Since $h$ crosses $r'$, we have $\xleft(h) < x_1 \leq x_2 < \xright(h)$.
+Since $r' = \gbot(B_{\bm{t}})$, we have the x-range of $B_{\bm{t}}$ is contained in $[x_1,x_2]$.
+Therefore, $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$.
+    \item Since $h$ crosses $r^*$, we have $y_h \in [y_1, \maxtop(r')]$. By the definition of $h$, we have $y_h \leq \ytop(B_{\bm{t}})$. Therefore, $y_h \in [y_1, \ytop(B_{\bm{t}})]$. 
+\end{itemize}

### 2024-05-30 msj.tex
yufei

@@
 Let us first consider the detection version of Problem $\mathscr{A}$.
-Each point $p = (x, y)$ of $P$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+Each point $p = (x, y)$ of $P$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
 We start by sorting the set of y-coordinates defined by all the points and rectangles.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each vertical segment $v = x \times [y_1, y_2]$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each vertical segment $v = x \times [y_1, y_2]$ of $V$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
 We start by sorting the set of y-coordinates defined by all the horizontal and vertical segments.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Let $S$ be a set of $N$ intervals in $\real$.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Let $S$ be a set of $N$ intervals in $\real$, each of which is associated with a real-valued {\em weight}.
+Given a search value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$).
+We can store $S$ in a structure of \cite{aak+12} using $O(N)$ space that can answer every query in $O(\log N)$ time.
+The structure also supports an insertion or deletion on $S$ in $O(\log N)$ amortized time.
+
+\vgap
+
+For Problem $\mathscr{C}$, each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+We start by sorting the set of y-coordinates defined by all the horizontal segments and rectangles.
+
+\vgap
+
+Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
+Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a structure $\T$ of \cite{aak+12}.
+Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ with {\em weight} $x_2$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
+When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $P$, search $\T$ to determine if any interval in $S_\ell$ contains the value $x$.
+If so, point $p$ is covered by at least one rectangle in $R$; otherwise, it is not.
+The overall running time is $O(n \log n)$.
 

### 2024-05-30 msj.tex
yufei

@@
 Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ with {\em weight} $x_2$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
-When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $P$, search $\T$ to determine if any interval in $S_\ell$ contains the value $x$.
-If so, point $p$ is covered by at least one rectangle in $R$; otherwise, it is not.
+When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $P$, search $\T$ to determine the maximum weight $w$ of all the intervals in $S_\ell$ containing the value $x_1$.
+If $w \ne -\infty$, we output $(h, p)$ where the point $p$ is defined in a way depending on $w$: if $w \le x_2$, then $p = (w, y)$; otherwise $p = (x_2, y)$.
 The overall running time is $O(n \log n)$.

### 2024-05-30 msj.tex
yufei

@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of $N$ intervals in $\real$ using $O(N)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log N + K)$ time, where $K$ is the number of intervals reported.
-It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log N)$ time.
-Each interval insertion or deletion on $S$ can be supported in $O(\log N)$ time.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
+It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log |S|)$ time.
+Each interval insertion or deletion on $S$ can be supported in $O(\log |S|)$ time.
 
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Let $S$ be a set of $N$ intervals in $\real$, each of which is associated with a real-valued {\em weight}.
-Given a search value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$).
-We can store $S$ in a structure of \cite{aak+12} using $O(N)$ space that can answer every query in $O(\log N)$ time.
-The structure also supports an insertion or deletion on $S$ in $O(\log N)$ amortized time.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Let $S$ be a set of intervals in $\real$, each of which is associated with a real-valued {\em weight}.
+Given a real value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$).
+We can store $S$ in a structure of \cite{aak+12} using $O(|S|)$ space that can answer every query in $O(\log |S|)$ time.
+The structure also supports an insertion or deletion on $S$ in $O(\log |S|)$ amortized time.
 
@@
 
-%\input{4.3.tex}
+\input{4.3.tex}
 

### 2024-05-30 msj.tex
yufei

@@
 Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
-At all times, maintain the set $V_\ell$ of rectangles in $R$ that intersect with $\ell$.
+At all times, maintain the set $V_\ell$ of segments in $V$ that intersect with $\ell$.
 Let $S_\ell$ be the set of x-coordinates of the segments in $V_\ell$; we store $S_\ell$ in a binary search tree (BST) $\T$.
-Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
+Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ amortized time.
 When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$.
-If $x' \le x_2$, then then we output a pair $(h, p)$, where $p$ is the point $(x', y)$.
+If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$.
 The overall running time is $O(n \log n)$.

### 2024-05-30 msj.tex
yufei

@@
 
-\input{4.3.tex}
+\extraspacing \underline{\em Proof of Statement (2).}
+Let $r' = [x_1,x_2] \times [y_1,y_2]$.
+Then, $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$.
+Let $h$ be an arbitrary segment in $\dcross_H(\bm{t})$, whose y-coordinate is $y_h$.
+We will prove that $h$ crosses $r^*$, which leads to $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
+It suffies to show that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$.
+\begin{itemize}
+    \item Since $h$ crosses $r' = [x_1,x_2]\times [y_1,y_2]$, we have $\xleft(h) < x_1 \le x_2 < \xright(h)$.
+    \item Since $h$ crosses $B_{\bm{t}}$, we have $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+By the definition of $\maxtop(r')$ (see Definition \eqref{eqn:hv:type1:maxtop}), and the fact that $r' = \gbot(B_{\bm{t}})$, we have $\ytop(B_{\bm{t}}) \leq \maxtop(r')$.
+Therefore, $y_h \in [y_1, \maxtop(r')]$.
+\end{itemize}
+
+Then, we will prove that each segment $h \in \cross_H(r^*)$ satisfying that the y-coordinate $h$ is no more than $\ytop(B_{\bm{t}})$ is in $\dcross_H(\bm{t})$.
+This claim leads to the last sentence of Statement (3).
+Consider an arbitrary $h\in \cross_H(r^*)$, whose y-coordinate $y_h$ satisfies that $y_h \le \ytop(B_{\bm{t}})$.
+We will prove that $h$ crosses $r'$ and $B_{\bm{t}}$, and therefore $h \in \dcross_H(\bm{t})$.
+
+\vgap
+
+First, we prove that $h$ crosses $r'$.
+It suffices to show that $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$.
+Both conditions are satisfied since $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$.
+Then, we prove that $h$ crosses $B_{\bm{t}}$.
+It suffices to show that $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+\begin{itemize}
+    \item Since $h$ crosses $r'$, we have $\xleft(h) < x_1 \leq x_2 < \xright(h)$.
+Since $r' = \gbot(B_{\bm{t}})$, we have the x-range of $B_{\bm{t}}$ is contained in $[x_1,x_2]$.
+Therefore, $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$.
+    \item Since $h$ crosses $r^*$, we have $y_h \in [y_1, \maxtop(r')]$. By the definition of $h$, we have $y_h \leq \ytop(B_{\bm{t}})$. Therefore, $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+\end{itemize}
 

### 2024-05-30 msj.tex
yufei

@@
 Recall that $r^* = [x_1,x_2] \times [y_1, \maxtop(r')]$.
-To prove that $h$ crosses $r^*$, we will show $\xleft(h) \le x_1 \le x_2 \le \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
+To prove that $h$ crosses $r^*$, we will show $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
 \myitems{
-    \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $r'$. As $[x_1, x_2]$ is the x-range of $r'$, we know $\xleft(h) \le x_1 \le x_2 \le \xright(h)$.
+    \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $r'$. As $[x_1, x_2]$ is the x-range of $r'$, we know $\xleft(h) < x_1 \le x_2 < \xright(h)$.
 
@@
 \extraspacing \underline{\em Proof of Statement (2).}
-Let $r' = [x_1,x_2] \times [y_1,y_2]$.
-Then, $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$.
-Let $h$ be an arbitrary segment in $\dcross_H(\bm{t})$, whose y-coordinate is $y_h$.
-We will prove that $h$ crosses $r^*$, which leads to $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
-It suffies to show that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$.
+Suppose that $r' = [x_1,x_2] \times [y_1, y_2]$.
+Accordingly, $r^*$ can be written as $[x_1, x_2] \times [y_1, \maxtop(r')]$.
+We will first prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
+For this purpose, given an arbitrary segment $h \in \dcross_H(\bm{t})$, we will show that $h$ crosses $r^*$.
+For this purpose, we need to explain why $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$.
 \begin{itemize}
-    \item Since $h$ crosses $r' = [x_1,x_2]\times [y_1,y_2]$, we have $\xleft(h) < x_1 \le x_2 < \xright(h)$.
-    \item Since $h$ crosses $B_{\bm{t}}$, we have $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
-By the definition of $\maxtop(r')$ (see Definition \eqref{eqn:hv:type1:maxtop}), and the fact that $r' = \gbot(B_{\bm{t}})$, we have $\ytop(B_{\bm{t}}) \leq \maxtop(r')$.
-Therefore, $y_h \in [y_1, \maxtop(r')]$.
+    \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $\gbot(\bm{t})$, which is $r'$.
+As the x-range of $r'$ is $[x_1,x_2]$, it must hold true that $\xleft(h) < x_1 \le x_2 < \xright(h)$.
+
+    \vgap
+
+    \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$.
+Hence, $y_h \in [\ybot(B_\bm{t}), \ytop(B_{\bm{t}})]$.
+As $r' = \gbot(\bm{t})$, we know $\ybot(B_\bm{t}) = \ybot(r') = y_1$.
+By the definition of $\maxtop(r')$ in \eqref{eqn:hv:type1:maxtop}, we know $\ytop(B_{\bm{t}}) \leq \maxtop(r')$.
+It thus follows that $y_h \in [y_1, \maxtop(r')]$.
 \end{itemize}
+We can now conclude that $h$ crosses $r^*$.
+
+\vgap
 

### 2024-05-30 msj.tex
yufei

@@
 Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$.
-Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
+Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
     }
@@
 
-Then, we will prove that each segment $h \in \cross_H(r^*)$ satisfying that the y-coordinate $h$ is no more than $\ytop(B_{\bm{t}})$ is in $\dcross_H(\bm{t})$.
-This claim leads to the last sentence of Statement (3).
-Consider an arbitrary $h\in \cross_H(r^*)$, whose y-coordinate $y_h$ satisfies that $y_h \le \ytop(B_{\bm{t}})$.
-We will prove that $h$ crosses $r'$ and $B_{\bm{t}}$, and therefore $h \in \dcross_H(\bm{t})$.
+Next, assuming that the segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates,
+we will prove that $\dcross_H(\bm{t})$ includes a prefix of the sorted order. It suffices to establish the following equivalent statement:
+
+\vgap
+
+\minipg{0.8\linewidth} {
+    If a segment $h \in \cross_H(r^*)$ has y-coordinate $y_h \le \ytop(B_{\bm{t}})$, then $h$ must be in $\dcross_H(\bm{t})$.
+}
+
+\vgap
+
+\noindent To prove the above, we must explain why $h$ crosses both $r'$ and $B_{\bm{t}}$.
 
@@
 Both conditions are satisfied since $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$.
-Then, we prove that $h$ crosses $B_{\bm{t}}$.
-It suffices to show that $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+
+\vgap
+
+Then, we prove that $h$ crosses $B_{\bm{t}}$. It suffices to show that $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
 \begin{itemize}

### 2024-05-30 msj.tex
yufei

@@
 
-First, we prove that $h$ crosses $r'$.
-It suffices to show that $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$.
-Both conditions are satisfied since $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$.
-
-\vgap
-
-Then, we prove that $h$ crosses $B_{\bm{t}}$. It suffices to show that $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+We first show that $h$ crosses $r'$, or equivalently: $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$.
+These conditions hold true because $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$ (due to the definition of top-sliced rectangle in \eqref{eqn:hv:type1:topsliced}).
+
+\vgap
+
+Finally, we show that $h$ crosses $B_{\bm{t}}$, or equivalently: $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [\ybot(B_\bm{t}), \ytop(B_{\bm{t}})] = [y_1, \ytop(B_{\bm{t}})]$.
 \begin{itemize}
-    \item Since $h$ crosses $r'$, we have $\xleft(h) < x_1 \leq x_2 < \xright(h)$.
+    \item The fact of $h$ crossing $r'$ tells us $\xleft(h) < x_1 \leq x_2 < \xright(h)$.
 Since $r' = \gbot(B_{\bm{t}})$, we have the x-range of $B_{\bm{t}}$ is contained in $[x_1,x_2]$.

### 2024-05-30 msj.tex
yufei

@@
     \item The fact of $h$ crossing $r'$ tells us $\xleft(h) < x_1 \leq x_2 < \xright(h)$.
-Since $r' = \gbot(B_{\bm{t}})$, we have the x-range of $B_{\bm{t}}$ is contained in $[x_1,x_2]$.
+Since $r' = \gbot(B_{\bm{t}})$, the x-range of $B_{\bm{t}}$ must be contained in $[x_1,x_2]$.
 Therefore, $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$.
-    \item Since $h$ crosses $r^*$, we have $y_h \in [y_1, \maxtop(r')]$. By the definition of $h$, we have $y_h \leq \ytop(B_{\bm{t}})$. Therefore, $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
+
+    \vgap
+
+    \item The fact of $h$ crossing $r^*$ also tells us that $y_h \in [y_1, \maxtop(r')]$.
+Moreover, by the definition of $h$, we have $y_h \leq \ytop(B_{\bm{t}})$.
+It thus follows that $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
 \end{itemize}

### 2024-05-30 msj.tex
yufei

@@
 \extraspacing \underline{\em Proof of Statement (2).}
-Suppose that $r' = [x_1,x_2] \times [y_1, y_2]$.
-Accordingly, $r^*$ can be written as $[x_1, x_2] \times [y_1, \maxtop(r')]$.
+Let $r', r^*$, and $\bm{t}$ be as defined in statement (2). Let us represent $r'$ as $[x_1,x_2] \times [y_1, y_2]$. Accordingly, $r^*$ can be written as $[x_1, x_2] \times [y_1, \maxtop(r')]$.
+
+\vgap
+
 We will first prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
-For this purpose, given an arbitrary segment $h \in \dcross_H(\bm{t})$, we will show that $h$ crosses $r^*$.
-For this purpose, we need to explain why $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$.
+For this purpose, given any segment $h \in \dcross_H(\bm{t})$, we will show that $h$ crosses $r^*$.
+To do so, we need to explain why $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$.
 \begin{itemize}
@@
 Hence, $y_h \in [\ybot(B_\bm{t}), \ytop(B_{\bm{t}})]$.
-As $r' = \gbot(\bm{t})$, we know $\ybot(B_\bm{t}) = \ybot(r') = y_1$.
+From $r' = \gbot(\bm{t})$, we get $\ybot(B_\bm{t}) = \ybot(r') = y_1$.
 By the definition of $\maxtop(r')$ in \eqref{eqn:hv:type1:maxtop}, we know $\ytop(B_{\bm{t}}) \leq \maxtop(r')$.

### 2024-05-30 msj.tex
yufei

@@
 \myeqn{
-    \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers} \nn \\[-1mm]
+    \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers} \nn \\[-0.5mm]
     && \text{the left endpoint of $h$}\}. \label{eqn:hv:type2:J2}
 }
-Other sub-types can be found analogously.
+The other sub-types can be found analogously.
 
@@
     \item The fact of $h$ crossing $r'$ tells us $\xleft(h) < x_1 \leq x_2 < \xright(h)$.
-Since $r' = \gbot(B_{\bm{t}})$, the x-range of $B_{\bm{t}}$ must be contained in $[x_1,x_2]$.
+As $r' = \gbot(B_{\bm{t}})$, the x-range of $B_{\bm{t}}$ must be contained in $[x_1,x_2]$.
 Therefore, $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$.

### 2024-05-30 msj.tex
yufei

@@
 Because $r'$ covers $r^*$, by the fact $h$ crossing $r^*$, we know that $h$ must intersect with $r'$.
-On the other hand, as $h$ crosses $r^*$, it must hold that $\xleft(h) < x_1 < x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$).
+On the other hand, as $h$ crosses $r^*$, it must hold that $\xleft(h) < x_1 \le x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$).
 As $[x_1, x_2]$ is also the x-range of $r'$, using the fact of $h \cap r' \ne \emptyset$, we can assert that $h$ crosses $r'$.
@@
 Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$.
-As proved earlier, $\xleft(h) < x_1 < x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting $p \in h$.
+As proved earlier, $\xleft(h) < x_1 \le x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting $p \in h$.
 

### 2024-05-30 5.1.tex
5.1

We prove by mapping each tuple $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$. 
To define the mapping, consider an arbitrary tuple $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$. Let $h$ be the full segment of $h'$. By the definition of $h'$, there exists a rectangle $r_{k-2}\in R_{k-2}$ that covers the left endpoint of $h$ and the right endpoint of $h'$. Let $\bm{t} = (r_1,...,r_{k-3}, r_{k-2},h,v)$. We map the tuple $(r_1,...,r_{k-3}, h', v)$ to $\bm{t}$.

\vgap

We argue that $\bm{t}\in \J_2$. Let $p$ be the intersection of $h'$ and $v$. Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, $p \in r_i$ for $i \in [k-3]$.

### 2024-05-30 5.1.tex
5.1

@@
 By the definition of $h'$, there exists a rectangle $r_{k-2}\in R_{k-2}$ that covers the left endpoint of $h$ and the right endpoint of $h'$.
+In another word, $r_{k-2}$ covers $h'$.
 Let $\bm{t} = (r_1,...,r_{k-3}, r_{k-2},h,v)$.
@@
 
-We argue that $\bm{t}\in \J_2$. Let $p$ be the intersection of $h'$ and $v$. Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, $p \in r_i$ for $i \in [k-3]$.
+We argue that $\bm{t}\in \J_2$.
+Let $p$ be the intersection of $h'$ and $v$.
+Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, $p \in r_i$ for $i \in [k-3]$.
+As we proved in the last paragraph, $r_{k-2}$ covers $h'$, so $p \in r_{k-2}$.
+Additionally, $r_{k-2}$ covers the left endpoint of $h$.
+Therefore, $\bm{t} \in \J_2$.
+
+\vgap
+
+It remains to prove that no two distinct tuple $\bm{t}_1, \bm{t}_2$ will be mapped to the same tuple in $\J_2$.
+Assume to the contrary, $\bm{t}_1, \bm{t}_2$ are mapped to the same tuple $\bm{t} \in \J_2$.
+Then, $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i\in [k-3]$ and $i = k-1$.
+Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both trimmed segment of $bm{t}[k-2]$.
+Since the trimmed segment is unique, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$.
+Hence, $\bm{t}_1 = \bm{t}_2$, which contradicts the assumption.
+
+\vgap
+
+Therefore, we have $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality is due to Definition \eqref{eqn:hv:type2:J2}.

### 2024-05-31 5.2.tex
5.2

\noindent \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$. Let $h$ be the full segment of $h'$. Since $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, $B_\bm{t}$, which is the intersection point of $h'$ and $v$, is in $r_i$ for $i \in [k-3]$. As $h$ is the full segment of $h'$, $B_\bm{t} \in h$. Since $v$ also contains $B_\bm{t}$, $B_\bm{t}$ is also the intersection of $h$ and $v$. Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$. By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers effective segment of $h'$, whose left enpoint is $B_{\bm{t}}$. Therefore, $B_{\bm{t}}$ is in $r_{i}$ for $i \in [k-2]$, and $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$. Moreover, since the left endpoint of the effective segment of $h'$ is the same as the left endpoint of $h$, $r_{k-2}$ contains the left endpoint of $h$. Therefore, $\bm{t} \in \J_2$.

\vgap

\noindent \underline{\em Proof of Statement (2).}

\vgap

\noindent \underline{\em Proof of Statement (3).}

\vgap

### 2024-05-31 msj.tex
yufei

@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.} The {\em priority search tree} (PST) \cite{m85} stores a set $P$ of points using $O(|P|)$ space such that, given any 3-sided rectangle $q = [x_1, x_2] \times [y, \infty)$, the points of $S$ covered by $q$ can be found in $O(\log |P| + K)$ time, where $K$ is the number of points reported.
+This structure can be used to answer the following {\em containment query} on intervals: store a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any interval $q' = [z_1, z_2]$, the intervals of $S$ that are contained in $q'$ can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
+To see why, observe that an interval $[z_1, z_2]$ contains another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $[z_1, \infty) \times (-\infty, z_2]$.
+Thus, we can create from $S$ a set of points $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
+Given an interval $q' = [z_1, z_2]$, we use the PST to find all the points in $P$ covered by $q = [z_1, \infty) \times (-\infty, z_2]$ and, for each such point $(x, y)$, report $[x, y]$.
+
+%\vgap
+
+%A PST built on a set $P$ of points can also answer the following detection query in $O(\log |P|)$ time: given any 3-sided rectangle $q = [x_1, x_2] \times [y, \infty)$, decide whether $P$ contains at least one point covered by $q$.
+Accordingly, a PST built on a set $S$ of intervals can also answer the detection version of the containment query in $O(\log |S|)$ time: specifically, given any interval $q' = [z_1, z_2]$, decide whether $S$ has at least one interval that is contained in $q'$.
+
+\vgap
+
+Let us now return to the context of $\mathscr{D}$.
+We assume, w.l.o.g., that each rectangle of $R$ is given a distinct ID in $[n]$.
+Let us first consider the detection version.
+Initialize an array $A$ of size $|R|$, where cell $A[i]$, $i \in [|R|]$, is allocated to the rectangle $r$ with ID $i$ and will be used to store a constant amount of information for $r$.
+In the outset, all rectangles of $R$ are marked as inactive (in $A$).
+During our algorithm, the status of each rectangle will turn from inactive to active at some point, turn from active back to inactive at a later point, and then stay that way forever.
+
+\vgap
+
+Sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+At all times, we maintain the set $R_\ell$ of {\em active} rectangles in $R$ that intersect with $\ell$.
+Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$.
+Specifically, when $\ell$ hits the bottom edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert $[x_1, x_2]$ into $\T$ and mark $r$ as active, which can be done in $O(\log n)$ time.
+The rectangle $r$ will be referred to as the {\em host} of $[x_1, x_2]$ and is stored together with $[x_1, x_2]$ in $\T$.
+
+\vgap
+
+When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a containment query on $\T$ to find all the intervals in $S_\ell$ that are contained by $[z_1, z_2]$; if $K_h$ is the number of such intervals, this retrieval takes $O(\log n + K_h)$ time.
+For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$ (stored along with $[x_1, x_2]$ in $\T$).
+Observe that $h$ is the lowest segment in $H$ that crosses $r$.
+We store $h$ in the cell of array $A$ allocated to $r$ and then mark $r$ as inactive; accordingly, we remove $[x_1, x_2]$ from $T$ in $O(\log n)$ time.
+As $r$ will remain inactive in the rest of the execution, its x-range $[x_1, x_2]$ will not be retrieved by another containment query in the future.
+
+\vgap
+
+When $\ell$ hits the bottom edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we check whether $r$ is active.
+If so, delete $[x_1, x_2]$ from $\T$ and mark $r$ as inactive; otherwise, do nothing.
+
+\vgap
+
+Overall, each rectangle of $R$ necessitates one insertion and one deletion in $\T$.
+All these insertions and deletions take $O(n \log n)$ time in total.
+Each segment $h$ of $H$ performs a query on $\T$, which has a cost of $O(\log n + K_h)$.
+All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.
+Recall that the x-range of a rectangle in $R$ can be retrieved by at most one query.
+Hence, $\sum_h K_h \le n$ and the runtime of our algorithm is $O(n \log n)$.
 

### 2024-05-31 msj.tex
yufei

@@
 Observe that $h$ is the lowest segment in $H$ that crosses $r$.
-We store $h$ in the cell of array $A$ allocated to $r$ and then mark $r$ as inactive; accordingly, we remove $[x_1, x_2]$ from $T$ in $O(\log n)$ time.
-As $r$ will remain inactive in the rest of the execution, its x-range $[x_1, x_2]$ will not be retrieved by another containment query in the future.
+We store $h$ in the cell of array $A$ allocated to $r$, after which $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time.
+As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future.
 
@@
 All these insertions and deletions take $O(n \log n)$ time in total.
-Each segment $h$ of $H$ performs a query on $\T$, which has a cost of $O(\log n + K_h)$.
+Each segment $h$ of $H$ performs a containment query on $\T$, which has a cost of $O(\log n + K_h)$.
 All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.
-Recall that the x-range of a rectangle in $R$ can be retrieved by at most one query.
+Recall that the x-range of a rectangle in $R$ can be retrieved by at most one containment query.
 Hence, $\sum_h K_h \le n$ and the runtime of our algorithm is $O(n \log n)$.

### 2024-05-31 msj.tex
yufei

@@
 
+\vgap
+
+Next, we consider the find-all-sorted version of Problem $\mathscr{D}$.
+Again, initialize an array $A$ of size $|R|$, where each rectangle of $R$ is allocated a cell.
+For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates.
+A pointer referencing the linked list is stored in the array cell of $r$.
+In the outset, all the linked lists are empty.
+Unlike the detection version, we do not need to record whether a rectangle is active or not.
+
+\vgap
+
+Again, sweep a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+At all times, we maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
+Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$.
+Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
+The rectangle $r$ will be referred to as the {\em host} of $[x_1, x_2]$ and is stored together with $[x_1, x_2]$ in $\T$.
+
+\vgap
+
+When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a containment query on $\T$ to find all the intervals in $S_\ell$ that are contained by $[z_1, z_2]$; the query cost is $O(\log n + K_h)$, where $K_h$ is the number of intervals reported.
+For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$.
+It is clear that $h$ is a segment crossing $r$ and is thus appended to the linked list of $r$.
+Note that $h$ is higher than all the segments already existing in that linked list.
+
+\vgap
+
+Overall, each rectangle of $R$ necessitates one insertion and one deletion in $\T$.
+All these insertions and deletions take $O(n \log n)$ time in total.
+Each segment $h$ of $H$ performs a containment query on $\T$, which has a cost of $O(\log n + K_h)$.
+All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.
+Recall that the x-range of a rectangle in $R$ can be retrieved by at most one containment query.
+Hence, $\sum_h K_h \le n$ and the runtime of our algorithm is $O(n \log n)$.
+
 \extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.}

### 2024-05-31 msj.tex
yufei

@@
 For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates.
-A pointer referencing the linked list is stored in the array cell of $r$.
-In the outset, all the linked lists are empty.
-Unlike the detection version, we do not need to record whether a rectangle is active or not.
+A pointer referencing the linked list is stored in the cell of $A$ allocated to $r$.
+In the outset, all linked lists are empty.
+Unlike the detection version, we will not need to keep the active status for the rectangles.
 
@@
 It is clear that $h$ is a segment crossing $r$ and is thus appended to the linked list of $r$.
-Note that $h$ is higher than all the segments already existing in that linked list.
+Note that $h$ is higher than all the segments already in that linked list.
 
@@
 All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.
-Recall that the x-range of a rectangle in $R$ can be retrieved by at most one containment query.
-Hence, $\sum_h K_h \le n$ and the runtime of our algorithm is $O(n \log n)$.
+However, unlike the detection version, the sum $\sum_h K_h$ is equal to the total size of $\cross_H(r)$ for all the $r \in R$.
+The total size is equivalent to $\out$.
 

### 2024-05-31 msj.tex

@@
 Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
-The rectangle $r$ will be referred to as the {\em host} of $[x_1, x_2]$ and is stored together with $[x_1, x_2]$ in $\T$.
+Again, the rectangle $r$ --- the host of $[x_1, x_2]$ --- is stored together with $[x_1, x_2]$ in $\T$.
 
@@
 All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.
-However, unlike the detection version, the sum $\sum_h K_h$ is equal to the total size of $\cross_H(r)$ for all the $r \in R$.
+However, unlike the detection version, the sum $\sum_h K_h$ here is equal to the total size of $\cross_H(r)$ for all the $r \in R$.
 The total size is equivalent to $\out$.

### 2024-05-31 5-comp.tex
5-comp

\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} We consider, w.l.o.g., that each segment in the input $H$ is given a distinct integer ID in $[|H|]$. This allows us to create an array of size $|H|$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.

\vgap

To compute $H*$, we start by deriving $\minleft(h')$ for each segment $h'$ in $H'$. For this purpose, first initialize $\minleft(h') = \infty$ for each such $h'$. Recall that $h'$ is the trimmed segment of some segment $h$ in $H$. We store $\minleft(h')$ in the array cell allocated to $h$.

\vgap

Then, we scan $\J(R_1,...,R_{k-3},H',V)$. For each tuple $\bm{t}$ therein, update in constant time $\minleft(h')$ to the minimum between its current value and the x-coordinate of $\bm{t}[k-1]$. The scan requires $O(n + \out)$ time.

\vgap

Finally, we construct $H^*$ by collecting the minimal segment (see definition in \eqref{eqn:hv:type2:H*}) of every segment $h' \in H'$ with $\minleft(h') \neq \infty$. This step takes $O(|H'|) = O(n)$ time.

### 2024-05-31 5.3.tex
5.3

\noindent \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple in $\J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$. We can then derive that 
\myeqn{
    \sum_{h^* \in H^*} |\cross_{R_{k-2}}(h^*)| \le \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\cross_{R_{k-2}}(\bm{t})| \leq \out, \nn
} where the last inequality is due to Statement (3) of Lemma~\ref{lmm:hv:type2:properties}.

\vgap

Now, we explain the mapping. Consider an arbitrary $h^* \in H^*$. Recall that $h^*$ is the minimal segment of some segment $h' \in H'$. Specifically, if $h' = [x_1,x_2]\times y$, then $h^* = [x_1,\minleft(h')]\times y$. By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$. We map $h^*$ to $\bm{t}$.

\vgap

Next, we will prove that a rectangle $r_{k-2} \in \cross_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \cross_{R_{k-2}}(\bm{t})$. By the definition of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ (see Definition~\eqref{eqn:contained} and \eqref{eqn:type2:contained-t}), it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$. Let $x$ be the $x$-coordinate of the vertical segment $\bm{t}[k-1]$. Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$, where $y$ is the $y$-coordinate of $h'$. The effective horizontal segment of $\bm{t}$ is therefore $[x_1,x]\times y$. Recall that $h^* = [x_1,\minleft(h')]\times y$, where $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$} = x$. Therefore, $h^*$ is the same as the effective horizontal segment of $\bm{t}$.

It remains to show that no two distinct rectangles $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$. Assume, on the contrary, that $h_1^*$ and $h_2^*$ are mapped to the same tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H',V)$. Suppose that $h_1^*$ (resp., $h_2^*$) is the minimal segment of $r_1'$ (resp., $r_2'$). Under our mapping, it must be true that  $h_1' = h_2' = \bm{t}[k-2]$. However, the distinctness of $h_1^*$ and $h_2^*$ requires $h_1' \ne h_2'$, which yields a contradiction.

\extraspacing \underline{\em Proof of Statement (2).}
Let $h', h^*$, and $\bm{t}$ be as defined in statement (2). Let us represent $h'$ as $[x_1,x_2] \times y$. Accordingly, $h^*$ can be written as $[x_1, \minleft(h')] \times y$. Additionally, let $x$ be the x-coordinate of the vertical segment $\bm{t}[k-1]$. Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$. The effective horizontal segment of $\bm{t}$ can be denoted as $[x_1,x] \times y$.

\vgap

We will first prove $\contained_{R_{k-2}}(\bm{t})\subseteq \cross_{R_{k-2}}(r^*)$. For this purpose, we will show that any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ also covers $h^*$. It suffices to show that $h^*$ is contained in the effective horizontal segment of $\bm{t}$. Combining the definition of $\minleft(h')$ (see \eqref{eqn:minleft}), and the fact that $x$ is the x-coordinate of $\bm{t}[k-1]$, where $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, we have $\minleft(h') \le x$. This implies that $h^* = [x_1, \minleft(h')] \times y$ is contained in $[x_1, x] \times y$, i.e. the effective horizontal segment of $\bm{t}$.

\vgap

Next, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order. It suffices to establish the following equivalent statement:

\vgap

\minipg{0.8\linewidth} {
    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies that $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
}

\vgap

\noindent To prove the above, we must explain why $r$ contains the effective horizontal segment of $\bm{t}$. Since $r \in \contained_{R_{k-2}}(h^*)$, $r$ covers the segment $[x_1, \minleft(h')] \times y$. Therefore, $[x_1, \minleft(h')] \subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
As $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, which we denote as $x$, $[x_1, x] \subseteq [\xleft(r), \xright(r)]$. This implies that $[x_1, x] \times y$, i.e. the effective horizontal segment of $\bm{t}$, is contained in $r$.

### 2024-05-31 msj.tex
yufei
Computing R^*

@@
 
+\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} We consider, w.l.o.g., that each rectangle in the input $R_1 \cup ...
+\cup R_{k-2}$ is given a distinct integer ID in $[n]$.
+This allows us to create an array of size $n$ and allocate an array cell to each $r \in R_1 \cup ...
+\cup R_{k-2}$, with the property that the cell can be accessed by the ID of $r$ in constant time.
+
+\vgap
+
+To compute $R^*_1, R^*_2, ..., R^*_{k-2}$, we start by deriving $\maxtop(r')$ for each rectangle $r'$ in $R'_1 \cup  ...
+\cup R'_{k-2}$.
+For this purpose, first initialize $\maxtop(r') = -\infty$ for each such $r'$.
+Recall that $r'$ is the trimmed rectangle of some rectangle $r$ in $R_1 \cup ...
+\cup R_{k-2}$.
+We store $\maxtop(r')$ in the array cell allocated to $r$.
+
+\vgap
+
+Then, we scan $\J(R_1',...,R_{k-2}')$.
+For each tuple $\bm{t}$ therein, use $O(k)$ time to identify the rectangle $r' = \gbot(\bm{t})$, and then update in constant time $\maxtop(r')$ to the maximum between its current value and $\ytop(r')$.
+The scan requires $O(n + k \cdot \out)$ time.
+
+\vgap
+
+Finally, for each $i \in [k-2]$, we construct $R_i^*$ by collecting the top-sliced rectangle (see definition in \eqref{eqn:hv:type1:topsliced}) of every rectangle $r' \in R_i'$ with $\maxtop(r') \neq -\infty$.
+This step takes $O(|R_i'|)$ time for each $i \in [k-2]$, or $O(n)$ total time for all $i$.
+
 \extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.} \\

### 2024-06-01 5-comp.tex
5-comp

@@
-\extraspacing {\bf Computing $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} We consider, w.l.o.g., that each segment in the input $H$ is given a distinct integer ID in $[|H|]$.
+W.l.o.g., we assume that each segment in the input $H$ is given a distinct integer ID in $[|H|]$.
 This allows us to create an array of size $|H|$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.
@@
 
-To compute $H*$, we start by deriving $\minleft(h')$ for each segment $h'$ in $H'$.
+To compute $H^*$, we start by deriving $\minleft(h')$ for each segment $h'$ in $H'$.
 For this purpose, first initialize $\minleft(h') = \infty$ for each such $h'$.

### 2024-06-01 5.3.tex
5.3

@@
-\noindent \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple in $\J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$.
+\noindent \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$.
 We can then derive that
@@
 Therefore, $[x_1, \minleft(h')] \subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
-As $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, which we denote as $x$, $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
+As $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, which is denoted as $x$, we have $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
 This implies that $[x_1, x] \times y$, i.e.

### 2024-06-01 5.2.tex

@@
 Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$.
-By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers effective segment of $h'$, whose left enpoint is $B_{\bm{t}}$.
+By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers effective segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
 Therefore, $B_{\bm{t}}$ is in $r_{i}$ for $i \in [k-2]$, and $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
@@
 
-\noindent \underline{\em Proof of Statement (2).}
+\noindent \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement. First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$.
+Let $p$ be the intersection of $h$ and $v$. Since $(r_1,...,r_{k-2},h,v)\in \J_2$, $p$ is in $r_i$ for $i \in [k-2]$.
+We argue that $p$ is also the intersection of $h'$ and $v$, which implies that $h'\cap v$ is covered by $r_i$, $i \in [k-3]$ and $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
+Since $p \in v$, it suffices to show that $p \in h'$.
+Let $h = [x_1,x_2]\times y$ and $p = (x,y)$.
+Since $h' \subseteq h$, we can denote $h'$ as $[x_1, x_2'] \times y$, where $x_1 \le x_2' \le x_2$.
+Then, we only need to show that $x \in [x_1, x_2']$.
+Since $p \in h$, $x \ge x_1$.
+Since $(r_1,...,r_{k-2},h,v) \in \J_2$, $r_{k-2}$ is a left-end covering rectangle of $h$ in $R_{k-2}$, that also covers the point $p \in h$.
+By the definition of trimmed segment (see Secton~\ref{sec:hv:type2}), the x-coordinate of $p$ should not exceed $x_2'$.
+Hence, $x \in [x_1, x_2']$, and $p \in h'$.
+
+\vgap
+
+Next, we will show that $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
+As proven above, $B_\bm{t} = h'\cap v = p$.
+Recall that $p = (x,y)$.
+We can denote the effective horizontal segment of $\bm{t}$ as $[x_1, x] \times y$.
+Then, we only need to prove that $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
+Since $r_{k-2}$ is a left-end covering rectangle of $h$, we have  $x_1 \in [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
+As we show in the last paragrah, $r_{k-2}$ also covers $p$.
+Therefore, $x \in [\xleft(r_2), \xright(r_2)]$, and $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$.
+Hence, $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
 
@@
 \noindent \underline{\em Proof of Statement (3).}
-
-\vgap
+ This statement follows from Statement (1).
+More specifically, we can map each pair $(\bm{t}, r)\in \J(R_1,...,R_{k-3},H',V) \times \contained_{R_{k-2}}(\bm{t})$ to the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ defined in Statement (1).
+By Statement (1), this $k$-tuple belongs to $\J_2$.
+Then, it remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$.
+Assume to the contrary, $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ are mapped to the same tuple $\bm{t} \in \J_2$.
+Then, $r_1 = \bm{t}[k-2] = r_2$, and $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i \in [k-3]$.
+Moreover, both $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are the trimmed segment of $\bm{t}[k-1]$.
+Since the trimmed segment is unique, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$.
+Hence, $\bm{t}_1 = \bm{t}_2$, which contradicts the assumption.
+Therefore, we have $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality is due to Definition \eqref{eqn:hv:type2:J2}.

### 2024-06-01 5-comp.tex
5-comp

@@
 W.l.o.g., we assume that each segment in the input $H$ is given a distinct integer ID in $[|H|]$.
-This allows us to create an array of size $|H|$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.
+This allows us to create an array of size $|H| < n$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.
 

### 2024-06-01 msj.tex
yufei

@@
 
-It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm solving the $k$-SJ problem, even for $k = 2$.
+It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm  solving the $k$-SJ problem (see Appendix~\ref{app:comp-class} for a formal definition of such algorithms), even for $k = 2$.
 This can be established via a reduction from the {\em element distinctness} (ED) problem; see \cite{dl79}.
@@
 
-In this paper, we solve the $k$-SJ problem with a comparison-based algorithm that runs in $O(n \log n + \out)$ time regardless of the constant $k$. The time complexity is asymptotically optimal.
+In this paper, we solve the $k$-SJ problem with a comparison-based algorithm that runs in $O(n \log n + \out)$ time regardless of the constant $k$.
+The time complexity is asymptotically optimal in the class of comparison-based algorithms.
 
@@
 Let $h'$ be the trimmed segment of $h$ and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
-Then, $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ and $h' \in \contained_{R_{k-2}}(\bm{t})$.
+Then, $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ and $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
 
@@
 
+\section{The Comparison-Based Class for $k$-SJ} \label{app:comp-class}
+
 \section{Building Brick Algorithms} \label{app:bricks}
@@
 
-%\vgap
-
-%A PST built on a set $P$ of points can also answer the following detection query in $O(\log |P|)$ time: given any 3-sided rectangle $q = [x_1, x_2] \times [y, \infty)$, decide whether $P$ contains at least one point covered by $q$.
-Accordingly, a PST built on a set $S$ of intervals can also answer the detection version of the containment query in $O(\log |S|)$ time: specifically, given any interval $q' = [z_1, z_2]$, decide whether $S$ has at least one interval that is contained in $q'$.
-
-\vgap
-
-Let us now return to the context of $\mathscr{D}$.
+\vgap
+
+Let us now return to the context of Problem $\mathscr{D}$.
+Let us first consider the detection version.
+Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+We start by sorting the set of y-coordinates defined by all the horizontal segments and rectangles.
+
+\vgap
+
 We assume, w.l.o.g., that each rectangle of $R$ is given a distinct ID in $[n]$.
-Let us first consider the detection version.
 Initialize an array $A$ of size $|R|$, where cell $A[i]$, $i \in [|R|]$, is allocated to the rectangle $r$ with ID $i$ and will be used to store a constant amount of information for $r$.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.}
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} The PST introduced earlier in Problem $\mathscr{D}$ can be used to answer the following {\em reverse-containment query} on intervals: store a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any interval $q' = [z_1, z_2]$, the intervals of $S$ that contain  $q'$ (rather than ``being contained by $q'$'') can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
+To see why, observe that an interval $[z_1, z_2]$ contains another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.
+Thus, we create from $S$ a set of points $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
+Given an interval $q' = [z_1, z_2]$, we use the PST to find all the points in $P$ covered by $(-\infty, z_1] \times [z_2, \infty)$ and, for each such point $(x, y)$, report $[x, y]$.
+
+\vgap
+
+Let us now return to the context of Problem $\mathscr{E}$.
+We sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$).
+Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position in the sorted list.
+We create an array $A$ and assign the $i$-th ($i \in [R]$) cell of $A$ to the rectangle $r \in R$ with ID $i$.
+
+\vgap
+
+Next, we aim to compute, for each pair $(h, r) \in H \times r$, we output a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$. These pairs may be output in an arbitrary order.
+
+\vgap
+
+Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+We sort the set of y-coordinates defined by all the horizontal segments and rectangles.
+Then, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+At all times, we maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
+Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$.
+Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
+We call the rectangle $r$ the {\em host} of $[x_1, x_2]$ and store its ID together with $[x_1, x_2]$ in $\T$.
+When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a reverse-containment query on $\T$ to find all the intervals in $S_\ell$ that contain $[z_1, z_2]$; the query cost is $O(\log n + K_h)$, where $K_h$ is the number of intervals reported.
+For each retrieved interval $[x_1, x_2]$, we also obtain the ID $\lambda$ of its host rectangle $r$, and output the pair $(h, \lambda)$.
+
+\vgap
+
+Each rectangle of $R$ necessitates one insertion and one deletion in $\T$.
+All these insertions and deletions take $O(n \log n)$ time in total.
+Each segment $h$ of $H$ performs a containment query on $\T$, which has a cost of $O(\log n + K_h)$.
+All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.
+The sum $\sum_h K_h$ here is equal to the total size of $\contained_R(h)$ for all the $h \in H$.
+The total size is equivalent to $\out$.
+The cost so far is therefore $O(n \log n + \out)$.
 

### 2024-06-01 5.1.tex
5.1

@@
 Let $h$ be the full segment of $h'$.
-By the definition of $h'$, there exists a rectangle $r_{k-2}\in R_{k-2}$ that covers the left endpoint of $h$ and the right endpoint of $h'$.
+By the definition of the trimmed segment (see Section~\ref{sec:hv:type2}), there exists a rectangle $r_{k-2}\in R_{k-2}$ that covers the left endpoint of $h$ and the right endpoint of $h'$.
 In another word, $r_{k-2}$ covers $h'$.
@@
 Then, $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i\in [k-3]$ and $i = k-1$.
-Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both trimmed segment of $bm{t}[k-2]$.
+Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both trimmed segment of $\bm{t}[k-2]$.
 Since the trimmed segment is unique, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$.

### 2024-06-01 msj.tex
yufei

@@
 
+\vgap
+
+Let $L$ be the list of $(h, \lambda)$ pairs produced (the size of $L$ is $\out$).
+We now proceed to sort $L$ in ascending order of the $\lambda$-field (which is a rectangle ID).
+Because the IDs are integers created by the algorithm, we are permitted to sort $L$ using {\em counting sort} without violating the requirements of the comparison-based class (see Appendix~\ref{app:comp-class}).
+The counting sort finishes in $O(|R| + |L|) = O(n + \out)$ time, recalling that all the IDs are in $[|R|]$.
+
 \section{Completing the Proof of Lemma~\ref{lmm:hv}} \label{app:hv}

### 2024-06-01 5.1.tex
5.1

@@
 Let $h$ be the full segment of $h'$.
-By the definition of the trimmed segment (see Section~\ref{sec:hv:type2}), there exists a rectangle $r_{k-2}\in R_{k-2}$ that covers the left endpoint of $h$ and the right endpoint of $h'$.
+By the definition of the trimmed segment (see Section~\ref{sec:hv:type2}), there exists a rectangle $r_{k-2}\in R_{k-2}$ that covers the right endpoint of $h'$, and the left endpoint of $h$, which is the same as the left endpoint of $h'$.
 In another word, $r_{k-2}$ covers $h'$.

### 2024-06-01 msj.tex
yufei

@@
 
+\vgap
+
+Finally, we generate, for each $h$, its set $\contained_R(h)$ (i.e., the set of rectangles covering $h$), where the rectangles $r$ are sorted by $\xright(r)$.
+To start with, initialize an empty linked list for every $h \in H$.
+Inspect the pairs $(h, \lambda) \in L$ in ascending order of the ID-field $\lambda$.
+For each pair $(h, \lambda)$ examined, identify the rectangle $r$ whose ID is $\lambda$ and add $r$ to the linked list of $h$.
+By the way the rectangle IDs were generated, it is clear that $\xright(r)$ is larger than or equal to the x-coordinates of the right edges of the rectangles already in the linked list.
+The whole scan over $L$ finishes in $O(1 + \out)$ time, and produces the correct output for Problem $\mathscr{E}$.
+
 \section{Completing the Proof of Lemma~\ref{lmm:hv}} \label{app:hv}

### 2024-06-01 5.1.tex
5.1

@@
 
-We argue that $\bm{t}\in \J_2$.
-Let $p$ be the intersection of $h'$ and $v$.
+We argue that $\bm{t}\in \J_2$. Let $p$ be the intersection point of $h'$ and $v$. As $h' \subseteq h$, the intersection point of $h$ and $v$ is also $p$.
 Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, $p \in r_i$ for $i \in [k-3]$.
-As we proved in the last paragraph, $r_{k-2}$ covers $h'$, so $p \in r_{k-2}$.
-Additionally, $r_{k-2}$ covers the left endpoint of $h$.
-Therefore, $\bm{t} \in \J_2$.
+As shown in the last paragraph, $r_{k-2}$ covers $h'$, so $p \in r_{k-2}$.
+Thus, $h \cap v \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$, and $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
+ Additionally, $r_{k-2}$ covers the left endpoint of $h$. Therefore, $\bm{t} \in \J_2$.
 

### 2024-06-01 5.1.tex
5.1

@@
 
-We argue that $\bm{t}\in \J_2$. Let $p$ be the intersection point of $h'$ and $v$. As $h' \subseteq h$, the intersection point of $h$ and $v$ is also $p$.
+We argue that $\bm{t}\in \J_2$. Let $p$ be the intersection point of $h'$ and $v$. As $h' \subseteq h$, the point $h\cap v$ is the same as $p$.
 Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, $p \in r_i$ for $i \in [k-3]$.
@@
 Thus, $h \cap v \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$, and $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
- Additionally, $r_{k-2}$ covers the left endpoint of $h$. Therefore, $\bm{t} \in \J_2$.
+ Additionally, by the definiton of $r_{k-2}$ in the last paragraph, $r_{k-2}$ covers the left endpoint of $h$. Therefore, $\bm{t} \in \J_2$.
 
@@
 It remains to prove that no two distinct tuple $\bm{t}_1, \bm{t}_2$ will be mapped to the same tuple in $\J_2$.
-Assume to the contrary, $\bm{t}_1, \bm{t}_2$ are mapped to the same tuple $\bm{t} \in \J_2$.
+Assume,on the contrary, $\bm{t}_1, \bm{t}_2$ are mapped to the same tuple $\bm{t} \in \J_2$.
 Then, $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i\in [k-3]$ and $i = k-1$.

### 2024-06-01 msj.tex
yufei

@@
 
+\extraspacing {\bf Terminology.} Each point $(x, y)$ is said to {\em define} a y-coordinate $y$, a horizontal segment $[x_1, x_2] \times y$ is said to {\em define} a y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
+These definitions permit us to convenient specify a set of y-coordinates using expressions like ``the set of four y-coordinates defined by point $p$, horizontal segment $h$, and rectangle $r$''.
+
+\extraspacing {\bf Fundamental Data Structures.} The {\em interval tree} \cite{bcko08} stores a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
+It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log |S|)$ time.
+The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ time.
+
+\vgap
+
+Now, suppose that each interval of $S$ is associated with a real-valued {\em weight}.
+Given a real value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$).
+We can store $S$ in a structure of \cite{aak+12} using $O(|S|)$ space that can answer such a query in $O(\log |S|)$ time.
+The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ amortized time.
+
+\vgap
+
+The {\em priority search tree} (PST) \cite{m85} stores a set $P$ of points using $O(|P|)$ space such that, given a 3-sided rectangle $q = [x_1, x_2] \times [y, \infty)$, the points of $S$ covered by $q$ can be found in $O(\log |P| + K)$ time, where $K$ is the number of points reported.
+The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ time.
+
+\vgap
+
+The PST can be deployed to answer queries on intervals.
+Let $S$ be a set of intervals in $\real$.
+Given an interval $q = [z_1, z_2]$, a {\em containment query} reports all the intervals of $S$ that are contained in $q$.
+We can store $S$ in a PST of $O(|S|)$ space that solves such a query in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
+To see why, observe that an interval $[z_1, z_2]$ is contained by another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $[z_1, \infty) \times (-\infty, z_2]$.
+Thus, we create from $S$ a point set $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
+Given an interval $q = [z_1, z_2]$, we can answer the containment query by using the PST to find all the points in $P$ covered by $[z_1, \infty) \times (-\infty, z_2]$ and, for each such point $(x, y)$, report $[x, y]$.
+
+\vgap
+
+Another closely related query is the {\em reverse-containment query}, which, given an interval $q = [z_1, z_2]$, finds all the intervals of $S$ that contain $q$ (rather than ``being contained by $q$'').
+Again, we can store $S$ in a PST using $O(|S|)$ space such that any such query can be answered in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
+In general, an interval $[z_1, z_2]$ contains another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.
+Thus, we create from $S$ a point set $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
+Given an interval $q = [z_1, z_2]$, we use the PST to find all the points in $P$ covered by $(-\infty, z_1] \times [z_2, \infty)$ and, for each such point $(x, y)$, report $[x, y]$.
+
 \extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Let $S$ be a set of intervals in $\real$, each of which is associated with a real-valued {\em weight}.
-Given a real value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$).
-We can store $S$ in a structure of \cite{aak+12} using $O(|S|)$ space that can answer every query in $O(\log |S|)$ time.
-The structure also supports an insertion or deletion on $S$ in $O(\log |S|)$ amortized time.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.}
 
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.} The {\em priority search tree} (PST) \cite{m85} stores a set $P$ of points using $O(|P|)$ space such that, given any 3-sided rectangle $q = [x_1, x_2] \times [y, \infty)$, the points of $S$ covered by $q$ can be found in $O(\log |P| + K)$ time, where $K$ is the number of points reported.
-This structure can be used to answer the following {\em containment query} on intervals: store a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any interval $q' = [z_1, z_2]$, the intervals of $S$ that are contained in $q'$ can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
-To see why, observe that an interval $[z_1, z_2]$ contains another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $[z_1, \infty) \times (-\infty, z_2]$.
-Thus, we can create from $S$ a set of points $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
-Given an interval $q' = [z_1, z_2]$, we use the PST to find all the points in $P$ covered by $q = [z_1, \infty) \times (-\infty, z_2]$ and, for each such point $(x, y)$, report $[x, y]$.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.}
 
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} The PST introduced earlier in Problem $\mathscr{D}$ can be used to answer the following {\em reverse-containment query} on intervals: store a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any interval $q' = [z_1, z_2]$, the intervals of $S$ that contain  $q'$ (rather than ``being contained by $q'$'') can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
-To see why, observe that an interval $[z_1, z_2]$ contains another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.
-Thus, we create from $S$ a set of points $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
-Given an interval $q' = [z_1, z_2]$, we use the PST to find all the points in $P$ covered by $(-\infty, z_1] \times [z_2, \infty)$ and, for each such point $(x, y)$, report $[x, y]$.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.}
 

### 2024-06-01 5.1.tex
5.1

@@
 It remains to prove that no two distinct tuple $\bm{t}_1, \bm{t}_2$ will be mapped to the same tuple in $\J_2$.
-Assume,on the contrary, $\bm{t}_1, \bm{t}_2$ are mapped to the same tuple $\bm{t} \in \J_2$.
-Then, $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i\in [k-3]$ and $i = k-1$.
-Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both trimmed segment of $\bm{t}[k-2]$.
-Since the trimmed segment is unique, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$.
+Assume, on the contrary, $\bm{t}_1, \bm{t}_2$ are mapped to the same tuple $\bm{t} \in \J_2$.
+Then, $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for each $i\in [k-3]$ and $i = k-1$.
+Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both the trimmed segment of $\bm{t}[k-2]$.
+By the uniqueness of the trimmed segment, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$.
 Hence, $\bm{t}_1 = \bm{t}_2$, which contradicts the assumption.

### 2024-06-01 5.2.tex

@@
 Let $h$ be the full segment of $h'$.
-Since $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, $B_\bm{t}$, which is the intersection point of $h'$ and $v$, is in $r_i$ for $i \in [k-3]$.
-As $h$ is the full segment of $h'$, $B_\bm{t} \in h$.
-Since $v$ also contains $B_\bm{t}$, $B_\bm{t}$ is also the intersection of $h$ and $v$.
+Folloing our conventions, $B_\bm{t}$ denotes the intersection point of $h'$ and $v$, and it is covered by the rectangles $r_1,...,r_{k-3}$.
+As $h$ is the full segment of $h'$, $B_\bm{t} \in h$, and $B_\bm{t}$ is also the intersection of $h$ and $v$.
 Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$.
-By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers effective segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
-Therefore, $B_{\bm{t}}$ is in $r_{i}$ for $i \in [k-2]$, and $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
-Moreover, since the left endpoint of the effective segment of $h'$ is the same as the left endpoint of $h$, $r_{k-2}$ contains the left endpoint of $h$.
+By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers the effective segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
+Therefore, $B_{\bm{t}}$ is in $r_{i}$ for $i \in [k-2]$.
+Equivalently, $h \cap v \bigcap_{i = 1}^{k-2} r_i = B_\bm{t} \neq \emptyset$.
+Hence,
+$\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
+Moreover, since $r_{k-2}$ covers the effective segment of $h'$, whose left endpoint is the same as the left endpoint of $h'$ and $h$, $r_{k-2}$ also covers the left endpoint of $h$.
 Therefore, $\bm{t} \in \J_2$.

### 2024-06-01 msj.tex
yufei

@@
 
-\def\vgap{\vspace{1mm}}
+\def\vgap{\vspace{2mm}}
+\def\extraspacing{\vspace{2mm}}
 
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
-It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log |S|)$ time.
-Each interval insertion or deletion on $S$ can be supported in $O(\log |S|)$ time.
-
-\vgap
-
-Let us first consider the detection version of Problem $\mathscr{A}$.
-Each point $p = (x, y)$ of $P$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
-We start by sorting the set of y-coordinates defined by all the points and rectangles.
-
-\vgap
-
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} Consider first the detection version.
+Sort the set of y-coordinates defined by the points of $P$ and the rectangles of $R$.
 Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each vertical segment $v = x \times [y_1, y_2]$ of $V$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
-We start by sorting the set of y-coordinates defined by all the horizontal and vertical segments.
-
-\vgap
-
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Sort the set of y-coordinates defined by all the segments of $H$ and $V$.
 Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.}
-
-\vgap
-
-For Problem $\mathscr{C}$, each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
-We start by sorting the set of y-coordinates defined by all the horizontal segments and rectangles.
-
-\vgap
-
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Sort the set of y-coordinates defined by all the segments in $H$ and all the rectangles in $R$.
 Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.}
-
-\vgap
-
-Let us now return to the context of Problem $\mathscr{D}$.
-Let us first consider the detection version.
-Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
-We start by sorting the set of y-coordinates defined by all the horizontal segments and rectangles.
-
-\vgap
-
-We assume, w.l.o.g., that each rectangle of $R$ is given a distinct ID in $[n]$.
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.} Consider first the find-lowest version.
+Sort the set of y-coordinates defined by all the horizontal segments and rectangles.
+In the outset, all rectangles of $R$ are marked as inactive.
+During our algorithm, the status of each rectangle will turn from inactive to active at some point, turn from active back to inactive at a later point, and then stay that way forever.
+
+%We assume, w.l.o.g., that each rectangle of $R$ is given a distinct ID in $[n]$.
 Initialize an array $A$ of size $|R|$, where cell $A[i]$, $i \in [|R|]$, is allocated to the rectangle $r$ with ID $i$ and will be used to store a constant amount of information for $r$.
-In the outset, all rectangles of $R$ are marked as inactive (in $A$).
-During our algorithm, the status of each rectangle will turn from inactive to active at some point, turn from active back to inactive at a later point, and then stay that way forever.
 
@@
 For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$ (stored along with $[x_1, x_2]$ in $\T$).
-Observe that $h$ is the lowest segment in $H$ that crosses $r$.
-We store $h$ in the cell of array $A$ allocated to $r$, after which $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time.
+Observe that $h$ is the lowest segment in $H$ that crosses $r$; we therefore output $(r, h)$.
+After that, $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time.
 As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future.
@@
 For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates.
-A pointer referencing the linked list is stored in the cell of $A$ allocated to $r$.
 In the outset, all linked lists are empty.
@@
 
-\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.}
-
-\vgap
-
-Let us now return to the context of Problem $\mathscr{E}$.
-We sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$).
+\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} Sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$).
 Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position in the sorted list.
-We create an array $A$ and assign the $i$-th ($i \in [R]$) cell of $A$ to the rectangle $r \in R$ with ID $i$.
-
-\vgap
-
-Next, we aim to compute, for each pair $(h, r) \in H \times r$, we output a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$. These pairs may be output in an arbitrary order.
-
-\vgap
-
-Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
-We sort the set of y-coordinates defined by all the horizontal segments and rectangles.
-Then, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+
+%We create an array $A$ and assign the $i$-th ($i \in [R]$) cell of $A$ to the rectangle $r \in R$ with ID $i$.
+
+\vgap
+
+Next, we aim to produce, for each pair $(h, r) \in H \times r$, a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$.
+These pairs may be output in an arbitrary order.
+Sort the set of y-coordinates defined by all the horizontal segments and rectangles.
+Sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
 At all times, we maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.

### 2024-06-01 msj.tex
yufei

@@
 
-Now, suppose that each interval of $S$ is associated with a real-valued {\em weight}.
+Now, let us assume that each interval of $S$ is associated with a real-valued {\em weight}.
 Given a real value $q$, a {\em stabbing max} query returns the maximum weight of all the intervals in $S$ covering $q$ (if no such intervals exist, the query returns $-\infty$).
@@
 When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$.
-If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$.
+If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$; otherwise, output nothing for $h$.
 The overall running time is $O(n \log n)$.
@@
 For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$ (stored along with $[x_1, x_2]$ in $\T$).
-Observe that $h$ is the lowest segment in $H$ that crosses $r$; we therefore output $(r, h)$.
+As can be verified shortly, $h$ is the lowest segment in $H$ that crosses $r$; we therefore output $(r, h)$.
 After that, $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time.
 As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future.
+This implies that $h$ is indeed the lowest segment in $H$ crossing $r$.
 

### 2024-06-01 msj.tex
yufei

@@
 \def\A{\mathcal{A}}
-\def\B{\mathcal{B}}
-\def\C{\mathcal{C}}
-\def\E{\mathcal{E}}
-\def\G{\mathcal{G}}
+%\def\B{\mathcal{B}}
+%\def\C{\mathcal{C}}
+%\def\E{\mathcal{E}}
+%\def\G{\mathcal{G}}
 \def\I{\mathcal{I}}
 \def\J{\mathcal{J}}
-\def\II{\mathscr{I}}
-\def\L{\mathcal{L}}
-\def\P{\mathcal{P}}
-\def\Q{\mathcal{Q}}
-\def\R{\mathcal{R}}
+%\def\II{\mathscr{I}}
+%\def\L{\mathcal{L}}
+%\def\P{\mathcal{P}}
+%\def\Q{\mathcal{Q}}
+%\def\R{\mathcal{R}}
 \def\T{\mathcal{T}}
-\def\U{\mathcal{U}}
-\def\V{\mathcal{V}}
-\def\X{\mathcal{X}}
-\def\XX{\mathscr{X}}
-\def\Y{\mathcal{Y}}
-\def\YY{\mathscr{Y}}
-\def\Z{\mathcal{Z}}
+%\def\U{\mathcal{U}}
+%\def\V{\mathcal{V}}
+%\def\X{\mathcal{X}}
+%\def\XX{\mathscr{X}}
+%\def\Y{\mathcal{Y}}
+%\def\YY{\mathscr{Y}}
+%\def\Z{\mathcal{Z}}
 
@@
 \def\out{\mathrm{OUT}}
+\def\dim{\mathrm{dim}}
 
@@
 \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$.
-Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$.
+Unless otherwise specified, every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$.
 All logarithms have base 2 by default.
@@
 \def\vgap{\vspace{2mm}}
-\def\extraspacing{\vspace{2mm}}
+\def\extraspacing{\vspace{2mm} \noindent}
 
@@
 \section{The Comparison-Based Class for $k$-SJ} \label{app:comp-class}
+
+Fix an integer $k \ge 3$, as well as $k$ integers $n_1, n_2, ..., n_k$ at least 1.
+Define $n = \sum_{i=1}^k n_i$.
+An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$.
+
+\vgap
+
+A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a decision tree $\T$ defined as follows.
+Each internal node $u$ of $\T$ has three child nodes, referred as the {\em left}, {\em middle}, and {\em right} child, respectively.
+Each node is associated with an {\em instruction} of the form $(i_1, \dim_1, \lambda_1; i_2, \dim_2, \lambda_2)$.
 
@@
 Next, we consider the find-all-sorted version of Problem $\mathscr{D}$.
-Again, initialize an array $A$ of size $|R|$, where each rectangle of $R$ is allocated a cell.
 For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates.
@@
 \extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} Sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$).
-Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position in the sorted list.
+Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position of the sorted list.
 

### 2024-06-01 5.2.tex
5.2

@@
 By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers the effective segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
-Therefore, $B_{\bm{t}}$ is in $r_{i}$ for $i \in [k-2]$.
+Therefore, $B_{\bm{t}}$ is covered by each of the rectangles $r_1,...,r_{k-2}$.
 Equivalently, $h \cap v \bigcap_{i = 1}^{k-2} r_i = B_\bm{t} \neq \emptyset$.
@@
 $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
-Moreover, since $r_{k-2}$ covers the effective segment of $h'$, whose left endpoint is the same as the left endpoint of $h'$ and $h$, $r_{k-2}$ also covers the left endpoint of $h$.
-Therefore, $\bm{t} \in \J_2$.
+To show that $\bm{t} \in \J_2$, we need to prove that $r_{k-2}$ covers the left endpoint of $h$.
+It is true because $r_{k-2}$ covers the effective segment of $h'$, whose left endpoint is the same as the left endpoint of $h$.
 

### 2024-06-01 5.2.tex
5.2

@@
 Let $h$ be the full segment of $h'$.
-Folloing our conventions, $B_\bm{t}$ denotes the intersection point of $h'$ and $v$, and it is covered by the rectangles $r_1,...,r_{k-3}$.
+Following our conventions, $B_\bm{t}$ denotes the intersection point of $h'$ and $v$, and it is covered by the rectangles $r_1,...,r_{k-3}$.
 As $h$ is the full segment of $h'$, $B_\bm{t} \in h$, and $B_\bm{t}$ is also the intersection of $h$ and $v$.
@@
 $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
-To show that $\bm{t} \in \J_2$, we need to prove that $r_{k-2}$ covers the left endpoint of $h$.
+To show that $\bm{t} \in \J_2$, we only need to prove that $r_{k-2}$ covers the left endpoint of $h$.
 It is true because $r_{k-2}$ covers the effective segment of $h'$, whose left endpoint is the same as the left endpoint of $h$.

### 2024-06-01 5.2.tex
5.2

@@
 Following our conventions, $B_\bm{t}$ denotes the intersection point of $h'$ and $v$, and it is covered by the rectangles $r_1,...,r_{k-3}$.
-As $h$ is the full segment of $h'$, $B_\bm{t} \in h$, and $B_\bm{t}$ is also the intersection of $h$ and $v$.
+As $h$ is the full segment of $h'$, we have $B_\bm{t} \in h$, and $B_\bm{t}$ is also the intersection of $h$ and $v$.
 Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$.
@@
 
-\noindent \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement. First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$.
-Let $p$ be the intersection of $h$ and $v$. Since $(r_1,...,r_{k-2},h,v)\in \J_2$, $p$ is in $r_i$ for $i \in [k-2]$.
-We argue that $p$ is also the intersection of $h'$ and $v$, which implies that $h'\cap v$ is covered by $r_i$, $i \in [k-3]$ and $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
-Since $p \in v$, it suffices to show that $p \in h'$.
+\noindent \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement.
+First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$.
+It suffices to show that the intersection of $h'$ and $v$ is contained in $r_1,...,r_{k-3}$.
+We will prove this by showing that $h'\cap v$ is the same as $h\cap v$, which is covered by $r_1,...,r_{k-2}$, because $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+Let $p$ be the intersection of $h$ and $v$. Since $h' \subseteq h$, the intersection of $h'$ and $v$ is either empty or the same as $p$.
+Therefore, to prove that $h' \cap v = p$, we only need to show that $p \in h'$.
 Let $h = [x_1,x_2]\times y$ and $p = (x,y)$.
 Since $h' \subseteq h$, we can denote $h'$ as $[x_1, x_2'] \times y$, where $x_1 \le x_2' \le x_2$.
-Then, we only need to show that $x \in [x_1, x_2']$.
-Since $p \in h$, $x \ge x_1$.
-Since $(r_1,...,r_{k-2},h,v) \in \J_2$, $r_{k-2}$ is a left-end covering rectangle of $h$ in $R_{k-2}$, that also covers the point $p \in h$.
-By the definition of trimmed segment (see Secton~\ref{sec:hv:type2}), the x-coordinate of $p$ should not exceed $x_2'$.
-Hence, $x \in [x_1, x_2']$, and $p \in h'$.
+Then, $p \in h'$ is equivalent to $x \in [x_1, x_2']$.
+Since $p \in h$, we have $x \ge x_1$.
+By the definition of trimmed segmemt, $(x_2',y)$ is the rightmost point on $h$ satisfying that a left-end covering rectangle of $h$ in $R_{k-2}$ covers it.
+As $(r_1,...,r_{k-2},h,v) \in \J_2$, the point $p \in h$ is covered by the rectangle $r_{k-2}$, which is a left-end covering rectangle of $h$.
+Therefore, the x-coordinate of $p$ should not exceed $x_2'$.
+As a result, we have $x \in [x_1, x_2']$, and $p \in h'$.
 

### 2024-06-01 5.2.tex
5.2

@@
 
-Next, we will show that $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
-As proven above, $B_\bm{t} = h'\cap v = p$.
-Recall that $p = (x,y)$.
-We can denote the effective horizontal segment of $\bm{t}$ as $[x_1, x] \times y$.
-Then, we only need to prove that $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
-Since $r_{k-2}$ is a left-end covering rectangle of $h$, we have  $x_1 \in [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
-As we show in the last paragrah, $r_{k-2}$ also covers $p$.
+Next, we aim at proving $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
+As proven above, $h'\cap v= p$.
+In another word, $B_\bm{t} = p$.
+Recall that we defind $p = (x,y)$, and $h' = [x_1,x_2']\times y$.
+Thus, we can denote the effective horizontal segment of $\bm{t}$ as $[x_1, x] \times y$.
+For our goal, we only need to show that $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
+\begin{itemize}
+    \item Since $r_{k-2}$ is a left-end covering rectangle of $h$, we have  $x_1 \in [\xleft(r_2), \xright(r_2)]$.
+As we show in the last paragrah, $r_{k-2}$ also covers $p$, whose x-coordinate, denoted as $x$, satisfies $x_1 \le x$.
 Therefore, $x \in [\xleft(r_2), \xright(r_2)]$, and $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$.
-Hence, $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
+    \item The y-coordinate of $h$ is $y$. Since $r_{k-2}$ intersects $h$, $y \in [\ybot(r_2), \ytop(r_2)]$.
+\end{itemize}
 

### 2024-06-01 5.2.tex
5.2

@@
 \noindent \underline{\em Proof of Statement (3).}
- This statement follows from Statement (1).
-More specifically, we can map each pair $(\bm{t}, r)\in \J(R_1,...,R_{k-3},H',V) \times \contained_{R_{k-2}}(\bm{t})$ to the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ defined in Statement (1).
-By Statement (1), this $k$-tuple belongs to $\J_2$.
-Then, it remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$.
-Assume to the contrary, $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ are mapped to the same tuple $\bm{t} \in \J_2$.
+We will provide a way to map each pair $(\bm{t'}, r)\in \J(R_1,...,R_{k-3},H',V) \times \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.
+Consider an arbitrary pair $(\bm{t'}, r)$.
+Let $h$ be the full segment of $\bm{t'}[k-2]$.
+We map $(\bm{t'}, r)$ to the tuple $\bm{t} = (\bm{t}'[1],...,\bm{t}'[k-3], r, h, \bm{t}'[k-1])$.
+By Statement (1) and the fact that $r\in \contained_{R_{k-2}}(\bm{t'})$, the $k$-tuple $\bm{t}$ belongs to $\J_2$.
+
+\vgap
+
+It remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$.
+Assume, on the contrary, $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ are mapped to the same tuple $\bm{t} \in \J_2$.
 Then, $r_1 = \bm{t}[k-2] = r_2$, and $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i \in [k-3]$.

### 2024-06-01 msj.tex
yufei

@@
 Each node is associated with an {\em instruction} of the form $(i_1, \dim_1, \lambda_1; i_2, \dim_2, \lambda_2)$.
+Each leaf is associated with an {\em output}, which is defined to be a subset of $[n_1] \times [n_2] \times ...
+\times [n_k]$.
+
+\vgap
+
+Given an $(n_1, n_2, ..., n_k)$-input, we execute the algorithm from the root of $\T$.
 
@@
 \extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} Sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$).
-Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position of the sorted list.
+To each $r \in R$, we assign an ID $i \in [R]$ if $r$ is at the $i$-th position of the sorted list.
 

### 2024-06-01 msj.tex
yufei

@@
 \def\out{\mathrm{OUT}}
-\def\dim{\mathrm{dim}}
+\def\side{\mathrm{side}}
 
@@
 
-\section{The Comparison-Based Class for $k$-SJ} \label{app:comp-class}
+\section{Comparison-Based $k$-SJ Algorithms} \label{app:comp-class}
 
@@
 Define $n = \sum_{i=1}^k n_i$.
-An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$.
+An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$ such that $|R_i| = n_i$ for each $i \in [k]$.
+We assume that each rectangle in $R_i$ ($i \in [k]$) is given a distinct ID from $[n_i]$.
 
@@
 Each internal node $u$ of $\T$ has three child nodes, referred as the {\em left}, {\em middle}, and {\em right} child, respectively.
-Each node is associated with an {\em instruction} of the form $(i_1, \dim_1, \lambda_1; i_2, \dim_2, \lambda_2)$.
-Each leaf is associated with an {\em output}, which is defined to be a subset of $[n_1] \times [n_2] \times ...
-\times [n_k]$.
-
-\vgap
-
-Given an $(n_1, n_2, ..., n_k)$-input, we execute the algorithm from the root of $\T$.
+Each node is associated with an {\em instruction} of the form $(i, \lambda, \side; i', \lambda', \side')$ where
+\myitems{
+    \item $i, i' \in [1, k]$;
+    \item $\lambda \in [n_i], \lambda' \in [n_{i'}]$;
+    \item $\side$ and $\side'$ are functions chosen from the set $\{\xleft, \xright, \ybot$, $\ytop\}$.
+}
+Each leaf is associated with an {\em result}, which is defined to be a subset of $[n_1] \times [n_2] \times ... \times [n_k]$.
+
+\vgap
+
+Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$.
+In general, at an internal node with the instruction $(i, \lambda, \side; i', \lambda', \side')$, we first identify
+\myitems{
+    \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
+    \item $r'$ as the rectangle with ID $\lambda'$ in $R_i'$.
+}
+Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
+\myitems{
+    \item the left child if $\side(r) < \side'(r')$;
+    \item the middle child if $\side(r) = \side'(r')$;
+    \item the right child if $\side(r) > \side'(r')$.
+}
+At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in the result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.
+
+\vgap
+
+An algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.
 

### 2024-06-01 5.2.tex

@@
 Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$.
-By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers the effective segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
+By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers the effective horizontal segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
 Therefore, $B_{\bm{t}}$ is covered by each of the rectangles $r_1,...,r_{k-2}$.
@@
 To show that $\bm{t} \in \J_2$, we only need to prove that $r_{k-2}$ covers the left endpoint of $h$.
-It is true because $r_{k-2}$ covers the effective segment of $h'$, whose left endpoint is the same as the left endpoint of $h$.
+It is true because $r_{k-2}$ covers the effective horizontal segment of $h'$, whose left endpoint is the same as the left endpoint of $h$.
 

### 2024-06-01 msj.tex
yufei

@@
 It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm  solving the $k$-SJ problem (see Appendix~\ref{app:comp-class} for a formal definition of such algorithms), even for $k = 2$.
-This can be established via a reduction from the {\em element distinctness} (ED) problem; see \cite{dl79}.
+This can be established via a reduction from the {\em element distinctness} problem; see \cite{dl79}.
 
@@
 
-An algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.
+The algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.
+
+\vgap
+
+As mentioned in Section~\ref{sec:intro:prev}, through a reduction from the element distinctness problem \cite{dl79}, one can show that when $k = 2$, any comparison-based algorithm capable of solving all $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.
 

### 2024-06-01 msj.tex

@@
     \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
-    \item $r'$ as the rectangle with ID $\lambda'$ in $R_i'$.
+    \item $r'$ as the rectangle with ID $\lambda'$ in $R_{i'}$.
 }
@@
 }
-At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in the result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.
-
-\vgap
-
-The algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.
-
-\vgap
-
-As mentioned in Section~\ref{sec:intro:prev}, through a reduction from the element distinctness problem \cite{dl79}, one can show that when $k = 2$, any comparison-based algorithm capable of solving all $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.
+At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.
+The algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs.
+Its {\em worst-case cost} is the height of $\T$.
+
+\vgap
+
+Through a reduction from the element distinctness problem \cite{dl79}, one can show that, for $k = 2$, any comparison-based algorithm capable of solving $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.
 

### 2024-06-01 msj.tex
yufei

@@
 Again, we can store $S$ in a PST using $O(|S|)$ space such that any such query can be answered in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
-In general, an interval $[z_1, z_2]$ contains another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.
+In general, an interval $[x, y]$ contains another $[z_1, z_2]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.
 Thus, we create from $S$ a point set $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
@@
 \extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} Sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$).
-To each $r \in R$, we assign an ID $i \in [R]$ if $r$ is at the $i$-th position of the sorted list.
+To each $r \in R$, we assign an ID $i \in [|R|]$ if $r$ is at the $i$-th position of the sorted list.
 
@@
 
-Next, we aim to produce, for each pair $(h, r) \in H \times r$, a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$.
+Next, we aim to produce, for each pair $(h, r) \in H \times r$ such that $r$ crosses $h$, a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$.
 These pairs may be output in an arbitrary order.
@@
 
-\section{Completing the Proof of Lemma~\ref{lmm:hv}} \label{app:hv}
+\section{Supplementary Proofs for Section~\ref{sec:hv:type1}} \label{app:hv:type1}
 
@@
 \end{itemize}
+
+\section{Supplementary Proofs for Section~\ref{sec:hv:type2}} \label{app:hv:type2}
 

### 2024-06-01 5.2.tex

@@
 \noindent \underline{\em Proof of Statement (3).}
-We will provide a way to map each pair $(\bm{t'}, r)\in \J(R_1,...,R_{k-3},H',V) \times \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.
-Consider an arbitrary pair $(\bm{t'}, r)$.
+We will provide a way to map each pair $(\bm{t'}, r)$ satisfyint that $\bm{t'}\in \J(R_1,...,R_{k-3},H',V)$ and $r \in \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.
+Consider an arbitrary such pair $(\bm{t'}, r)$.
 Let $h$ be the full segment of $\bm{t'}[k-2]$.

### 2024-06-01 msj.tex
yufei

@@
 This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$.
-We prove in Appendix~\ref{app:hv}:
+We prove in Appendix~\ref{app:hv:type1}:
 
@@
 }
-The prefix ``d-'' stands for ``double''. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:
+The prefix ``d-'' stands for ``double''. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv:type1}:
 
@@
 }
-In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
+In Appendix~\ref{app:hv:type1}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
 
@@
 Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
-The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties of these sets.
+The following lemma, proven in Appendix~\ref{app:hv:type1}, presents some useful properties of these sets.
 
@@
 Let $\J(R_1, ...,$ $R_{k-3}, H', V)$ represent the result of this $(k-1)$-SJ, whose input size is at most $n$.
-Given the lemma below (which is proved in Appendix~\ref{app:hv}), we assert that $\J(R_1, ..., R_{k-3}, H', V)$ can be computed in $F_{k-1}(n, \out)$ time.
+Given the lemma below (which is proved in Appendix~\ref{app:hv:type2}), we assert that $\J(R_1, ..., R_{k-3}, H', V)$ can be computed in $F_{k-1}(n, \out)$ time.
 
@@
 
-We prove the next lemma in Appendix~\ref{app:hv} (to avoid confusion, the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).
+We prove the next lemma in Appendix~\ref{app:hv:type2} (to avoid confusion, the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).
 
@@
 }
-This can be done in $O(n + k \cdot \out)$ time, as shown in Appendix~\ref{app:hv}.
+This can be done in $O(n + k \cdot \out)$ time, as shown in Appendix~\ref{app:hv:type2}.
 
@@
 
-The proof can be found in Appendix~\ref{app:hv}.
+The proof can be found in Appendix~\ref{app:hv:type2}.
 
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} \\
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
 \underline{\em Proof of Statement (1).} As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.} \\
-\underline{\em Proof of Statement (1).} We will map each $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t}) = \cross_H(r^*)$.
-After this, we can then derive $\sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out$, where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
+\underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$.
+After this, we can then derive
+\myeqn{
+    \sum_{i\in[k-2]}\sum_{r^*\in R_i^*} |\cross_H(r^*)| \leq \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')}|\dcross_H(\bm{t})|\leq \out \nn
+}
+where the last step used  statement (3) of Lemma~\ref{lmm:hv:type1:properties}.
 

### 2024-06-01 msj.tex
yufei

@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.}
-\underline{\em Proof of Statement (1).} As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} We will prove each statement in turn.
+
+\extraspacing \underline{\em Proof of Statement (1).} As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
 Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$.
@@
 
-\vgap
-
-\noindent \underline{\em Proof of Statement (2).} Fix any rectangle $r \in R_1 \cup ...
+\extraspacing \underline{\em Proof of Statement (2).} Fix any rectangle $r \in R_1 \cup ...
 \cup R_{k-2}$ and any horizontal segment $h = [x_1, x_2] \times y$ from $H$ that crosses $r$.
@@
 
-\vgap
-
-\noindent \underline{\em Proof of Statement (3).} We first prove that $\dcross_H(\bm{t}) \neq \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+\extraspacing \underline{\em Proof of Statement (3).} We first prove that $\dcross_H(\bm{t}) \neq \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
 Set $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
-\underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$.
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.} We will prove each statement in turn.
+
+\extraspacing{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$.
 After this, we can then derive

### 2024-06-01 msj.tex
yufei

@@
 
-A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a decision tree $\T$ defined as follows.
+A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a tree $\T$ defined as follows.
 Each internal node $u$ of $\T$ has three child nodes, referred as the {\em left}, {\em middle}, and {\em right} child, respectively.
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
-
-\input{5.1.tex}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.
+
+\vgap
+
+To describe the mapping, consider an arbitrary $(k-1)$-tuple $\bm{t} = (r_1,...,$ $r_{k-3},$ $h', v) \in \J(R_1,...,$ $R_{k-3},H',V)$.
+Let $h$ be the full segment of $h'$.
+By the definition of trimmed segments (see Section~\ref{sec:hv:type2}), at least one rectangle $r_{k-2} \in R_{k-2}$ covers both the left endpoint of $h$ and the right endpoint of $h'$.
+We map $\bm{t}$ to the $k$-tuple $(r_1,...,r_{k-3}, r_{k-2},h,v)$.
+It is worth noting that $r_{k-2}$ definitely contains the entire $h'$ (which has the same left point as $h$, implying that $r_{k-2}$ covers both endpoints of $h'$).
+
+\vgap
+
+We argue that $(r_1,...,r_{k-3}, r_{k-2},h,v)$ belongs to $\J_2$.
+Let $p$ be the intersection point of $h'$ and $v$.
+As $h'$ is contained in $h$, the intersection point between $h$ and $v$ is also $p$.
+Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, it must hold that $p$ falls in $r_i$ for $i \in [k-3]$.
+From our earlier discussion, we already know that $r_{k-2}$ contains $h'$; hence, $p$ falls in $r_{k-2}$ as well.
+Thus, $h \cap v \cap \bigcap_{i = 1}^{k-2}r_i$ is non-empty, indicating that $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$.
+This, together with the fact that $r_{k-2}$ covers the left endpoint of $h$, tells us $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.
+
+\vgap
+
+It remains to show that no two distinct $(k-1)$-tuples $\bm{t}_1$ and $\bm{t}_2$ can be mapped to the same tuple in $\J_2$.
+Assume, on the contrary, that $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.
+Then, $\bm{t}_1[i] = \bm{t}_2[i] = r_i$ for each $i \in [k-3]$, and $\bm{t}_1[k-1] = \bm{t}_2[k-1] = v$.
+Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both the trimmed segment of $h$.
+But then this means $\bm{t}_1 = \bm{t}_2$, giving a contradiction.
+
+\vgap
+
+We can now conclude that $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality used the definition in \eqref{eqn:hv:type2:J2}.
 

### 2024-06-01 msj.tex
yufei

@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.}
-
-\input{5.2.tex}
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.} We will prove each statement in turn.
+
+\noindent \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$.
+Let $h$ be the full segment of $h'$.
+Following our conventions, $B_\bm{t}$ denotes the intersection point of $h'$ and $v$, and it is covered by the rectangles $r_1,...,r_{k-3}$.
+As $h$ is the full segment of $h'$, we have $B_\bm{t} \in h$, and $B_\bm{t}$ is also the intersection of $h$ and $v$.
+Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$.
+By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers the effective horizontal segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
+Therefore, $B_{\bm{t}}$ is covered by each of the rectangles $r_1,...,r_{k-2}$.
+Equivalently, $h \cap v \bigcap_{i = 1}^{k-2} r_i = B_\bm{t} \neq \emptyset$.
+Hence,
+$\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
+To show that $\bm{t} \in \J_2$, we only need to prove that $r_{k-2}$ covers the left endpoint of $h$.
+It is true because $r_{k-2}$ covers the effective horizontal segment of $h'$, whose left endpoint is the same as the left endpoint of $h$.
+
+\vgap
+
+\noindent \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement.
+First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$.
+It suffices to show that the intersection of $h'$ and $v$ is contained in $r_1,...,r_{k-3}$.
+We will prove this by showing that $h'\cap v$ is the same as $h\cap v$, which is covered by $r_1,...,r_{k-2}$, because $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+Let $p$ be the intersection of $h$ and $v$. Since $h' \subseteq h$, the intersection of $h'$ and $v$ is either empty or the same as $p$.
+Therefore, to prove that $h' \cap v = p$, we only need to show that $p \in h'$.
+Let $h = [x_1,x_2]\times y$ and $p = (x,y)$.
+Since $h' \subseteq h$, we can denote $h'$ as $[x_1, x_2'] \times y$, where $x_1 \le x_2' \le x_2$.
+Then, $p \in h'$ is equivalent to $x \in [x_1, x_2']$.
+Since $p \in h$, we have $x \ge x_1$.
+By the definition of trimmed segmemt, $(x_2',y)$ is the rightmost point on $h$ satisfying that a left-end covering rectangle of $h$ in $R_{k-2}$ covers it.
+As $(r_1,...,r_{k-2},h,v) \in \J_2$, the point $p \in h$ is covered by the rectangle $r_{k-2}$, which is a left-end covering rectangle of $h$.
+Therefore, the x-coordinate of $p$ should not exceed $x_2'$.
+As a result, we have $x \in [x_1, x_2']$, and $p \in h'$.
+
+\vgap
+
+Next, we aim at proving $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
+As proven above, $h'\cap v= p$.
+In another word, $B_\bm{t} = p$.
+Recall that we defind $p = (x,y)$, and $h' = [x_1,x_2']\times y$.
+Thus, we can denote the effective horizontal segment of $\bm{t}$ as $[x_1, x] \times y$.
+For our goal, we only need to show that $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
+\begin{itemize}
+    \item Since $r_{k-2}$ is a left-end covering rectangle of $h$, we have  $x_1 \in [\xleft(r_2), \xright(r_2)]$.
+As we show in the last paragrah, $r_{k-2}$ also covers $p$, whose x-coordinate, denoted as $x$, satisfies $x_1 \le x$.
+Therefore, $x \in [\xleft(r_2), \xright(r_2)]$, and $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$.
+    \item The y-coordinate of $h$ is $y$. Since $r_{k-2}$ intersects $h$, $y \in [\ybot(r_2), \ytop(r_2)]$.
+\end{itemize}
+
+\vgap
+
+\noindent \underline{\em Proof of Statement (3).}
+We will provide a way to map each pair $(\bm{t'}, r)$ satisfyint that $\bm{t'}\in \J(R_1,...,R_{k-3},H',V)$ and $r \in \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.
+Consider an arbitrary such pair $(\bm{t'}, r)$.
+Let $h$ be the full segment of $\bm{t'}[k-2]$.
+We map $(\bm{t'}, r)$ to the tuple $\bm{t} = (\bm{t}'[1],...,\bm{t}'[k-3], r, h, \bm{t}'[k-1])$.
+By Statement (1) and the fact that $r\in \contained_{R_{k-2}}(\bm{t'})$, the $k$-tuple $\bm{t}$ belongs to $\J_2$.
+
+\vgap
+
+It remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$.
+Assume, on the contrary, $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ are mapped to the same tuple $\bm{t} \in \J_2$.
+Then, $r_1 = \bm{t}[k-2] = r_2$, and $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i \in [k-3]$.
+Moreover, both $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are the trimmed segment of $\bm{t}[k-1]$.
+Since the trimmed segment is unique, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$.
+Hence, $\bm{t}_1 = \bm{t}_2$, which contradicts the assumption.
+Therefore, we have $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality is due to Definition \eqref{eqn:hv:type2:J2}.
 

### 2024-06-01 msj.tex
yufei

@@
 At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.
-The algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs.
+The algorithm must produce the correct output on all $(n_1, n_2, ..., n_k)$-inputs.
 Its {\em worst-case cost} is the height of $\T$.
@@
 
-Through a reduction from the element distinctness problem \cite{dl79}, one can show that, for $k = 2$, any comparison-based algorithm capable of solving $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.
+Through a reduction from element distinctness \cite{dl79}, one can show that, for $k = 2$, any comparison-based algorithm correctly solving all $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.
 
@@
 
-\noindent \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$.
+\extraspacing \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$.
 Let $h$ be the full segment of $h'$.
-Following our conventions, $B_\bm{t}$ denotes the intersection point of $h'$ and $v$, and it is covered by the rectangles $r_1,...,r_{k-3}$.
-As $h$ is the full segment of $h'$, we have $B_\bm{t} \in h$, and $B_\bm{t}$ is also the intersection of $h$ and $v$.
-Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$.
-By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers the effective horizontal segment of $h'$, whose right enpoint is $B_{\bm{t}}$.
-Therefore, $B_{\bm{t}}$ is covered by each of the rectangles $r_1,...,r_{k-2}$.
-Equivalently, $h \cap v \bigcap_{i = 1}^{k-2} r_i = B_\bm{t} \neq \emptyset$.
-Hence,
-$\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
-To show that $\bm{t} \in \J_2$, we only need to prove that $r_{k-2}$ covers the left endpoint of $h$.
-It is true because $r_{k-2}$ covers the effective horizontal segment of $h'$, whose left endpoint is the same as the left endpoint of $h$.
-
-\vgap
-
-\noindent \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement.
+The rectangle $B_\bm{t}$ defined in \eqref{eqn:B_t} is the intersection point of $h'$ and $v$ (this point is covered by all of $r_1,...,r_{k-3}$).
+As $h$ is the full segment of $h'$, the point $B_\bm{t}$ must also be the intersection point of $h$ and $v$.
+
+\vgap
+
+Let $r_{k-2}$ be an arbitrary rectangle in the set $\contained_{R_{k-2}}(\bm{t})$.
+The definition of $\contained_{R_{k-2}}(\bm{t})$ tells us that $r_{k-2}$ covers the effective horizontal segment of $h'$ (see Section~\ref{sec:hv:type2} for the definition of ``effective horizontal segment''), whose right endpoint is $B_{\bm{t}}$.
+This means that $B_\bm{t}$ falls in $r_{k-2}$ as well.
+
+\vgap
+
+Now we know that $h \cap v \cap \bigcap_{i = 1}^{k-2} r_i$ contains $B_\bm{t}$ and is thus non-empty. Hence,
+$(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
+To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to prove that $r_{k-2}$ covers the left endpoint of $h$.
+This follows immediately from the fact that $r_{k-2}$ covers the effective horizontal segment of $h'$ (which shares the same left endpoint as $h$).
+
+\extraspacing \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement.
 First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$.
@@
 
-\vgap
-
 \noindent \underline{\em Proof of Statement (3).}

### 2024-06-01 msj.tex
yufei

@@
 Let $h$ be the full segment of $h'$.
-By the definition of trimmed segments (see Section~\ref{sec:hv:type2}), at least one rectangle $r_{k-2} \in R_{k-2}$ covers both the left endpoint of $h$ and the right endpoint of $h'$.
+By the definition of a trimmed segment (see Section~\ref{sec:hv:type2}), at least one rectangle $r_{k-2} \in R_{k-2}$ covers both the left endpoint of $h$ and the right endpoint of $h'$.
 We map $\bm{t}$ to the $k$-tuple $(r_1,...,r_{k-3}, r_{k-2},h,v)$.
@@
 Let $r_{k-2}$ be an arbitrary rectangle in the set $\contained_{R_{k-2}}(\bm{t})$.
-The definition of $\contained_{R_{k-2}}(\bm{t})$ tells us that $r_{k-2}$ covers the effective horizontal segment of $h'$ (see Section~\ref{sec:hv:type2} for the definition of ``effective horizontal segment''), whose right endpoint is $B_{\bm{t}}$.
+The definition of $\contained_{R_{k-2}}(\bm{t})$ tells us that $r_{k-2}$ covers the effective horizontal segment of $h'$ (see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment), whose right endpoint is $B_{\bm{t}}$.
 This means that $B_\bm{t}$ falls in $r_{k-2}$ as well.
@@
 
-\extraspacing \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement.
-First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$.
-It suffices to show that the intersection of $h'$ and $v$ is contained in $r_1,...,r_{k-3}$.
-We will prove this by showing that $h'\cap v$ is the same as $h\cap v$, which is covered by $r_1,...,r_{k-2}$, because $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
-
-\vgap
-
-Let $p$ be the intersection of $h$ and $v$. Since $h' \subseteq h$, the intersection of $h'$ and $v$ is either empty or the same as $p$.
-Therefore, to prove that $h' \cap v = p$, we only need to show that $p \in h'$.
-Let $h = [x_1,x_2]\times y$ and $p = (x,y)$.
-Since $h' \subseteq h$, we can denote $h'$ as $[x_1, x_2'] \times y$, where $x_1 \le x_2' \le x_2$.
-Then, $p \in h'$ is equivalent to $x \in [x_1, x_2']$.
-Since $p \in h$, we have $x \ge x_1$.
-By the definition of trimmed segmemt, $(x_2',y)$ is the rightmost point on $h$ satisfying that a left-end covering rectangle of $h$ in $R_{k-2}$ covers it.
-As $(r_1,...,r_{k-2},h,v) \in \J_2$, the point $p \in h$ is covered by the rectangle $r_{k-2}$, which is a left-end covering rectangle of $h$.
-Therefore, the x-coordinate of $p$ should not exceed $x_2'$.
-As a result, we have $x \in [x_1, x_2']$, and $p \in h'$.
-
-\vgap
-
-Next, we aim at proving $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
-As proven above, $h'\cap v= p$.
-In another word, $B_\bm{t} = p$.
-Recall that we defind $p = (x,y)$, and $h' = [x_1,x_2']\times y$.
-Thus, we can denote the effective horizontal segment of $\bm{t}$ as $[x_1, x] \times y$.
-For our goal, we only need to show that $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
+\extraspacing \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, and $\bm{t}$ be as defined in statement (2).
+Firstly, we will argue that the $(k-1)$-tuple $\bm{t}$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, which is equivalent to saying that that the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
+The subsequent discussion will prove that $h' \cap v$ is in fact the same point as $h \cap v$.
+From $(r_1,...,r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$, we already know that $h \cap v$ falls in all of $r_1,...,r_{k-2}$.
+Therefore, obtaining $h' \cap v = h \cap v$ will allow us to assert that $r_1,...,r_{k-3}$ all cover $h' \cap v$.
+
+\vgap
+
+Denote by $p$ the intersection point of $h$ and $v$.
+As $h'$ is contained in $h$, to prove $h' \cap v = p$, it suffices to explain why $p$ is on $h'$.
+Let us represent $h$ as $[x_1,x_2] \times y$ and $p$ as $(x, y)$ for some $x \in [x_1, x_2]$.
+Accordingly, we can represent $h'$ as $[x_1, x_2'] \times y$ for some $x_2' \in [x_1, x_2]$.
+By the definition of a trimmed segment, we know that $(x_2',y)$ is the rightmost point on $h$ that falls in a left-end covering rectangle of $h$ in $R_{k-2}$.
+From statement (2), we have $(r_1,...,r_{k-2},h,v) \in \J_2$, meaning that $r_{k-2}$ (i) is a left-end covering rectangle of $h$, and (ii) covers $p$ ($= h \cap v$).
+Therefore, the x-coordinate of $p$ cannot exceed $x_2'$, allowing us to assert that $p$ is on $h'$.
+
+\vgap
+
+%This tells us $B_\bm{t} = p$.
+
+Secondly, we will argue that $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
+The above discussion has shown that $h \cap v = h' \cap v = p$.
+If (as before) we write $h = [x_1,x_2] \times y$, $h' = [x_1,x_2'] \times y$, and $p = (x,y)$, then the effective horizontal segment of $\bm{t}$ can be written as $[x_1, x] \times y$ (due to the definition of an effective horizontal segment).
+To prove $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers the entire segment $[x_1, x] \times y$.
+We do so in two steps:
+
+% $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
+
 \begin{itemize}
-    \item Since $r_{k-2}$ is a left-end covering rectangle of $h$, we have  $x_1 \in [\xleft(r_2), \xright(r_2)]$.
-As we show in the last paragrah, $r_{k-2}$ also covers $p$, whose x-coordinate, denoted as $x$, satisfies $x_1 \le x$.
-Therefore, $x \in [\xleft(r_2), \xright(r_2)]$, and $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$.
-    \item The y-coordinate of $h$ is $y$. Since $r_{k-2}$ intersects $h$, $y \in [\ybot(r_2), \ytop(r_2)]$.
+    \item As $r_{k-2}$ is a left-end covering rectangle of $h$, we must have $x_1 \in [\xleft(r_2), \xright(r_2)]$.
+On the other hand, the fact $r_{k-2}$ covering $p$ gives $x \in [\xleft(r_2), \xright(r_2)]$.
+It thus follows that $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$.
+
+    \vgap
+
+    \item The fact $h$ intersecting with $r_{k-2}$ gives $y \in [\ybot(r_2), \ytop(r_2)]$.
 \end{itemize}
 
-\noindent \underline{\em Proof of Statement (3).}
+\noindent Combining the above tells us that  $r_{k-2}$ covers $[x_1, x] \times y$.
+
+\extraspacing \underline{\em Proof of Statement (3).}
 We will provide a way to map each pair $(\bm{t'}, r)$ satisfyint that $\bm{t'}\in \J(R_1,...,R_{k-3},H',V)$ and $r \in \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.

### 2024-06-01 msj.tex
yufei

@@
 \extraspacing \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$.
+The point $h' \cap v$ (intersection of $h'$ and $v$) is covered by all of $r_1,...,r_{k-3}$.
 Let $h$ be the full segment of $h'$.
-The rectangle $B_\bm{t}$ defined in \eqref{eqn:B_t} is the intersection point of $h'$ and $v$ (this point is covered by all of $r_1,...,r_{k-3}$).
-As $h$ is the full segment of $h'$, the point $B_\bm{t}$ must also be the intersection point of $h$ and $v$.
+As $h$ contains $h'$, we must have $h \cap v = h' \cap v$.
 
@@
 Let $r_{k-2}$ be an arbitrary rectangle in the set $\contained_{R_{k-2}}(\bm{t})$.
-The definition of $\contained_{R_{k-2}}(\bm{t})$ tells us that $r_{k-2}$ covers the effective horizontal segment of $h'$ (see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment), whose right endpoint is $B_{\bm{t}}$.
-This means that $B_\bm{t}$ falls in $r_{k-2}$ as well.
-
-\vgap
-
-Now we know that $h \cap v \cap \bigcap_{i = 1}^{k-2} r_i$ contains $B_\bm{t}$ and is thus non-empty. Hence,
+The definition of $\contained_{R_{k-2}}(\bm{t})$ tells us that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ (see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment), whose right endpoint is $h' \cap v$.
+This means that $h \cap v$ falls in $r_{k-2}$ as well.
+
+\vgap
+
+Now we know that $h \cap v \cap \bigcap_{i = 1}^{k-2} r_i$ is not empty because it equals $h' \cap v$. Hence,
 $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
-To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to prove that $r_{k-2}$ covers the left endpoint of $h$.
+To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to explain why $r_{k-2}$ covers the left endpoint of $h$.
 This follows immediately from the fact that $r_{k-2}$ covers the effective horizontal segment of $h'$ (which shares the same left endpoint as $h$).
@@
 From statement (2), we have $(r_1,...,r_{k-2},h,v) \in \J_2$, meaning that $r_{k-2}$ (i) is a left-end covering rectangle of $h$, and (ii) covers $p$ ($= h \cap v$).
-Therefore, the x-coordinate of $p$ cannot exceed $x_2'$, allowing us to assert that $p$ is on $h'$.
+Therefore, the x-coordinate of $p$ cannot exceed $x_2'$ (i.e., $x \le x_2'$), allowing us to assert that $p$ is on $h'$.
 

### 2024-06-01 msj.tex
yufei

@@
 
-\extraspacing \underline{\em Proof of Statement (3).}
-We will provide a way to map each pair $(\bm{t'}, r)$ satisfyint that $\bm{t'}\in \J(R_1,...,R_{k-3},H',V)$ and $r \in \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.
+\extraspacing \underline{\em Proof of Statement (3).} Our proof resorts to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}). This algorithm adds to $\J_2$ exactly
+\myeqn{
+    \sum_{\bm{t} \in \J(R_1, ..., R_3, H', V)} |\contained_{R_{k-2}}(\bm{t})| \nn
+}
+tuples. By statement (2), the $\J_2$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$.
+Therefore, the above expression is at most $\out$, as claimed in statement (3).
+
+% We will provide a way to map each pair $(\bm{t'}, r)$ satisfyint that $\bm{t'}\in \J(R_1,...,R_{k-3},H',V)$ and $r \in \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.
 Consider an arbitrary such pair $(\bm{t'}, r)$.
@@
 By Statement (1) and the fact that $r\in \contained_{R_{k-2}}(\bm{t'})$, the $k$-tuple $\bm{t}$ belongs to $\J_2$.
-
-\vgap
-
-It remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$.
+%
+% \vgap
+%
+% It remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$.
 Assume, on the contrary, $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ are mapped to the same tuple $\bm{t} \in \J_2$.

### 2024-06-01 msj.tex
yufei

@@
 We argue that $(r_1,...,r_{k-3}, r_{k-2},h,v)$ belongs to $\J_2$.
-Let $p$ be the intersection point of $h'$ and $v$.
-As $h'$ is contained in $h$, the intersection point between $h$ and $v$ is also $p$.
-Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, it must hold that $p$ falls in $r_i$ for $i \in [k-3]$.
-From our earlier discussion, we already know that $r_{k-2}$ contains $h'$; hence, $p$ falls in $r_{k-2}$ as well.
-Thus, $h \cap v \cap \bigcap_{i = 1}^{k-2}r_i$ is non-empty, indicating that $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$.
+Let $p$ be $h' \cap v$ (the intersection point of $h'$ and $v$).
+As $h'$ is contained in $h$, it must hold that $h \cap v = p$.
+Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, we know that $p$ falls in all of $r_1, ..., r_{k-3}$.
+Our earlier discussion already shows that $r_{k-2}$ contains $h'$; this means that $p$ falls in $r_{k-2}$.
+Thus, $h \cap v \cap \bigcap_{i = 1}^{k-2}r_i$ cannot be empty as it equals $p$.
+We can now conclude that $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$.
 This, together with the fact that $r_{k-2}$ covers the left endpoint of $h$, tells us $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.

### 2024-06-01 msj.tex
yufei

@@
 
-Equipped with Lemma~\ref{lmm:hv:type1:properties}, we generate our target $\J_2$ as follows:
+Equipped with Lemma~\ref{lmm:hv:type2:properties}, we generate our target $\J_2$ as follows:
 
@@
 \extraspacing \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$.
-The point $h' \cap v$ (intersection of $h'$ and $v$) is covered by all of $r_1,...,r_{k-3}$.
+The point $h' \cap v$ is covered by all of $r_1,...,r_{k-3}$.
 Let $h$ be the full segment of $h'$.
@@
 
-\extraspacing {\bf Computing $\bm{H^*}$.}
-
-\input{5-comp.tex}
+\extraspacing {\bf Computing $\bm{H^*}$.} W.l.o.g., we assume that each segment in the input $H$ is given a distinct integer ID in $[|H|]$.
+This allows us to create an array of size $|H| < n$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.
+
+\vgap
+
+To compute $H^*$, we start by deriving $\minleft(h')$ for each segment $h'$ in $H'$.
+For this purpose, first initialize $\minleft(h') = \infty$ for each such $h'$.
+Recall that $h'$ is the trimmed segment of some segment $h$ in $H$.
+We store $\minleft(h')$ in the array cell allocated to $h$.
+
+\vgap
+
+Then, we scan $\J(R_1,...,R_{k-3},H',V)$.
+For each tuple $\bm{t}$ therein, update in constant time $\minleft(h')$ to the minimum between its current value and the x-coordinate of $\bm{t}[k-1]$.
+The scan requires $O(n + k \cdot \out)$ time.
+
+\vgap
+
+Finally, we construct $H^*$ by collecting the minimal segment (see definition in \eqref{eqn:hv:type2:H*}) of every segment $h' \in H'$ with $\minleft(h') \neq \infty$.
+This step takes $O(|H'|) = O(n)$ time.
 

### 2024-06-02 msj.tex
yufei

@@
 
-%In other words, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ must have size at least 1.
-
 \vgap
@@
 We store $\maxtop(r')$ in the array cell allocated to $r$.
-
-\vgap
-
 Then, we scan $\J(R_1',...,R_{k-2}')$.
@@
 
-% We will provide a way to map each pair $(\bm{t'}, r)$ satisfyint that $\bm{t'}\in \J(R_1,...,R_{k-3},H',V)$ and $r \in \contained_{R_{k-2}}(\bm{t})$ to a unique tuple in $\J_2$.
-Consider an arbitrary such pair $(\bm{t'}, r)$.
-Let $h$ be the full segment of $\bm{t'}[k-2]$.
-We map $(\bm{t'}, r)$ to the tuple $\bm{t} = (\bm{t}'[1],...,\bm{t}'[k-3], r, h, \bm{t}'[k-1])$.
-By Statement (1) and the fact that $r\in \contained_{R_{k-2}}(\bm{t'})$, the $k$-tuple $\bm{t}$ belongs to $\J_2$.
-%
-% \vgap
-%
-% It remains to prove that no two distinct pair $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ will be mapped to the same tuple in $\J_2$.
-Assume, on the contrary, $(\bm{t}_1, r_1)$ and $(\bm{t}_2, r_2)$ are mapped to the same tuple $\bm{t} \in \J_2$.
-Then, $r_1 = \bm{t}[k-2] = r_2$, and $\bm{t}_1[i] = \bm{t}[i] = \bm{t}_2[i]$ for $i \in [k-3]$.
-Moreover, both $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are the trimmed segment of $\bm{t}[k-1]$.
-Since the trimmed segment is unique, $\bm{t}_1[k-2] = \bm{t}_2[k-2]$.
-Hence, $\bm{t}_1 = \bm{t}_2$, which contradicts the assumption.
-Therefore, we have $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality is due to Definition \eqref{eqn:hv:type2:J2}.
-
-\extraspacing {\bf Computing $\bm{H^*}$.} W.l.o.g., we assume that each segment in the input $H$ is given a distinct integer ID in $[|H|]$.
+\extraspacing {\bf Computing $\bm{H^*}$.} We assume, w.l.o.g., that each segment in the input $H$ is given a distinct integer ID in $[|H|]$.
 This allows us to create an array of size $|H| < n$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.
@@
 We store $\minleft(h')$ in the array cell allocated to $h$.
-
-\vgap
-
 Then, we scan $\J(R_1,...,R_{k-3},H',V)$.
@@
 
-\input{5.3.tex}
+\noindent \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$.
+We can then derive that
+\myeqn{
+    \sum_{h^* \in H^*} |\cross_{R_{k-2}}(h^*)| \le \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\cross_{R_{k-2}}(\bm{t})| \leq \out, \nn
+} where the last inequality is due to Statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+
+\vgap
+
+Now, we explain the mapping.
+Consider an arbitrary $h^* \in H^*$.
+Recall that $h^*$ is the minimal segment of some segment $h' \in H'$.
+Specifically, if $h' = [x_1,x_2]\times y$, then $h^* = [x_1,\minleft(h')]\times y$.
+By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$.
+We map $h^*$ to $\bm{t}$.
+
+\vgap
+
+Next, we will prove that a rectangle $r_{k-2} \in \cross_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \cross_{R_{k-2}}(\bm{t})$.
+By the definition of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ (see Definition~\eqref{eqn:contained} and \eqref{eqn:type2:contained-t}), it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
+Let $x$ be the $x$-coordinate of the vertical segment $\bm{t}[k-1]$.
+Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$, where $y$ is the $y$-coordinate of $h'$.
+The effective horizontal segment of $\bm{t}$ is therefore $[x_1,x]\times y$.
+Recall that $h^* = [x_1,\minleft(h')]\times y$, where $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$} = x$.
+Therefore, $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
+
+It remains to show that no two distinct rectangles $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$.
+Assume, on the contrary, that $h_1^*$ and $h_2^*$ are mapped to the same tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H',V)$.
+Suppose that $h_1^*$ (resp., $h_2^*$) is the minimal segment of $r_1'$ (resp., $r_2'$).
+Under our mapping, it must be true that  $h_1' = h_2' = \bm{t}[k-2]$.
+However, the distinctness of $h_1^*$ and $h_2^*$ requires $h_1' \ne h_2'$, which yields a contradiction.
+
+\extraspacing \underline{\em Proof of Statement (2).}
+Let $h', h^*$, and $\bm{t}$ be as defined in statement (2).
+Let us represent $h'$ as $[x_1,x_2] \times y$.
+Accordingly, $h^*$ can be written as $[x_1, \minleft(h')] \times y$.
+Additionally, let $x$ be the x-coordinate of the vertical segment $\bm{t}[k-1]$.
+Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$.
+The effective horizontal segment of $\bm{t}$ can be denoted as $[x_1,x] \times y$.
+
+\vgap
+
+We will first prove $\contained_{R_{k-2}}(\bm{t})\subseteq \cross_{R_{k-2}}(h^*)$.
+For this purpose, we will show that any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ also covers $h^*$.
+It suffices to show that $h^*$ is contained in the effective horizontal segment of $\bm{t}$.
+Combining the definition of $\minleft(h')$ (see \eqref{eqn:minleft}), and the fact that $x$ is the x-coordinate of $\bm{t}[k-1]$, where $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, we have $\minleft(h') \le x$.
+This implies that $h^* = [x_1, \minleft(h')] \times y$ is contained in $[x_1, x] \times y$, i.e.
+the effective horizontal segment of $\bm{t}$.
+
+\vgap
+
+Next, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
+It suffices to establish the following equivalent statement:
+
+\vgap
+
+\minipg{0.8\linewidth} {
+    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies that $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
+}
+
+\vgap
+
+\noindent To prove the above, we must explain why $r$ contains the effective horizontal segment of $\bm{t}$.
+Since $r \in \contained_{R_{k-2}}(h^*)$, $r$ covers the segment $[x_1, \minleft(h')] \times y$.
+Therefore, $[x_1, \minleft(h')] \subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
+As $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, which is denoted as $x$, we have $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
+This implies that $[x_1, x] \times y$, i.e.
+the effective horizontal segment of $\bm{t}$, is contained in $r$.
 

### 2024-06-02 msj.tex
yufei

@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.}
-
-\noindent \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$.
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:contained-h*}.} We will prove each statement in turn.
+
+\extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$.
 We can then derive that
@@
 Specifically, if $h' = [x_1,x_2]\times y$, then $h^* = [x_1,\minleft(h')]\times y$.
-By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$.
+By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$ (recall that $\bm{t}[k-1]$ is a vertical segment).
 We map $h^*$ to $\bm{t}$.
@@
 Next, we will prove that a rectangle $r_{k-2} \in \cross_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \cross_{R_{k-2}}(\bm{t})$.
-By the definition of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ (see Definition~\eqref{eqn:contained} and \eqref{eqn:type2:contained-t}), it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
+By the definition of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ --- defined in  \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
 Let $x$ be the $x$-coordinate of the vertical segment $\bm{t}[k-1]$.

### 2024-06-02 msj.tex
yufei

@@
 Specifically, if $h' = [x_1,x_2]\times y$, then $h^* = [x_1,\minleft(h')]\times y$.
-By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1',...,R_{k-2}')$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$ (recall that $\bm{t}[k-1]$ is a vertical segment).
+By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$ (recall that $\bm{t}[k-1]$ is a vertical segment).
 We map $h^*$ to $\bm{t}$.
@@
 Next, we will prove that a rectangle $r_{k-2} \in \cross_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \cross_{R_{k-2}}(\bm{t})$.
-By the definition of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ --- defined in  \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
-Let $x$ be the $x$-coordinate of the vertical segment $\bm{t}[k-1]$.
-Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$, where $y$ is the $y$-coordinate of $h'$.
+According to the definitions of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ --- see \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
+Denote by $x$ the x-coordinate of the vertical segment $\bm{t}[k-1]$.
+Then, the point $h' \cap v$ (intersection between $h'$ and $v$) is $(x,y)$, where as mentioned $y$ is the y-coordinate of $h'$.
 The effective horizontal segment of $\bm{t}$ is therefore $[x_1,x]\times y$.
@@
 
-It remains to show that no two distinct rectangles $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$.
+\vgap
+
+It remains to show that no two distinct segments $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$.
 Assume, on the contrary, that $h_1^*$ and $h_2^*$ are mapped to the same tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H',V)$.
-Suppose that $h_1^*$ (resp., $h_2^*$) is the minimal segment of $r_1'$ (resp., $r_2'$).
+Suppose that $h_1^*$ (resp., $h_2^*$) is the minimal segment of $h_1'$ (resp., $h_2'$).
 Under our mapping, it must be true that  $h_1' = h_2' = \bm{t}[k-2]$.

### 2024-06-02 msj.tex
yufei

@@
 
-\extraspacing \underline{\em Proof of Statement (2).}
-Let $h', h^*$, and $\bm{t}$ be as defined in statement (2).
+\extraspacing \underline{\em Proof of Statement (2).} Let $h', h^*$, and $\bm{t}$ be as defined in statement (2).
 Let us represent $h'$ as $[x_1,x_2] \times y$.
@@
 Additionally, let $x$ be the x-coordinate of the vertical segment $\bm{t}[k-1]$.
-Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$.
-The effective horizontal segment of $\bm{t}$ can be denoted as $[x_1,x] \times y$.
+The effective horizontal segment of $\bm{t}$ can be represented as $[x_1,x] \times y$.
+
+%Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$.
 
@@
 It suffices to show that $h^*$ is contained in the effective horizontal segment of $\bm{t}$.
-Combining the definition of $\minleft(h')$ (see \eqref{eqn:minleft}), and the fact that $x$ is the x-coordinate of $\bm{t}[k-1]$, where $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, we have $\minleft(h') \le x$.
+Combining the definition of $\minleft(h')$ (see \eqref{eqn:minleft}), and the fact that $x$ is the x-coordinate of $\bm{t}[k-1]$, we have $\minleft(h') \le x$.
 This implies that $h^* = [x_1, \minleft(h')] \times y$ is contained in $[x_1, x] \times y$, i.e.
@@
 
-Next, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
+Next, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
 It suffices to establish the following equivalent statement:
@@
 \minipg{0.8\linewidth} {
-    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies that $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
+    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies the condition that $\xright(r) \ge$ the x-coordinate of $\bm{t}[k-1]$, then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
 }
@@
 \noindent To prove the above, we must explain why $r$ contains the effective horizontal segment of $\bm{t}$.
-Since $r \in \contained_{R_{k-2}}(h^*)$, $r$ covers the segment $[x_1, \minleft(h')] \times y$.
+Since $r \in \contained_{R_{k-2}}(h^*)$, we know that $r$ covers the segment $[x_1, \minleft(h')] \times y$.
 Therefore, $[x_1, \minleft(h')] \subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
-As $\xright(r)$ is no less than the x-coordinate of $\bm{t}[k-1]$, which is denoted as $x$, we have $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
-This implies that $[x_1, x] \times y$, i.e.
+As $\xright(r) \ge x$ (recall that $x$ is the x-coordinate of $\bm{t}[k-1]$), we have $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
+This means that $[x_1, x] \times y$, i.e.
 the effective horizontal segment of $\bm{t}$, is contained in $r$.

### 2024-06-02 msj.tex
yufei

@@
 \extraspacing \underline{\em Proof of Statement (2).} Let $h', h^*$, and $\bm{t}$ be as defined in statement (2).
-Let us represent $h'$ as $[x_1,x_2] \times y$.
-Accordingly, $h^*$ can be written as $[x_1, \minleft(h')] \times y$.
+Let us represent $h'$ as $[x_1,x_2] \times y$; accordingly, $h^*$ can be written as $[x_1, \minleft(h')] \times y$.
 Additionally, let $x$ be the x-coordinate of the vertical segment $\bm{t}[k-1]$.
-The effective horizontal segment of $\bm{t}$ can be represented as $[x_1,x] \times y$.
+The effective horizontal segment of $\bm{t}$ can then be represented as $[x_1,x] \times y$.
 
@@
 
-We will first prove $\contained_{R_{k-2}}(\bm{t})\subseteq \cross_{R_{k-2}}(h^*)$.
+Firstly, we will prove $\contained_{R_{k-2}}(\bm{t})\subseteq \cross_{R_{k-2}}(h^*)$.
 For this purpose, we will show that any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ also covers $h^*$.
@@
 
-Next, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
+Secondly, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ have been sorted in descending order of $\xright(r)$, we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
 It suffices to establish the following equivalent statement:
@@
 \minipg{0.8\linewidth} {
-    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies the condition that $\xright(r) \ge$ the x-coordinate of $\bm{t}[k-1]$, then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
+    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies the condition that $\xright(r) \ge x$ (where as defined before $x$ is the x-coordinate of $\bm{t}[k-1]$), then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
 }
@@
 Therefore, $[x_1, \minleft(h')] \subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
-As $\xright(r) \ge x$ (recall that $x$ is the x-coordinate of $\bm{t}[k-1]$), we have $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
+Using also the given condition $\xright(r) \ge x$, we can derive $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
 This means that $[x_1, x] \times y$, i.e.

### 2024-06-02 msj.tex
yufei

@@
 
-\extraspacing \underline{\em Proof of Statement (2).}
-Let $r', r^*$, and $\bm{t}$ be as defined in statement (2). Let us represent $r'$ as $[x_1,x_2] \times [y_1, y_2]$. Accordingly, $r^*$ can be written as $[x_1, x_2] \times [y_1, \maxtop(r')]$.
-
-\vgap
-
-We will first prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
+\extraspacing \underline{\em Proof of Statement (2).} Let $r', r^*$, and $\bm{t}$ be as defined in statement (2).
+Let us represent $r'$ as $[x_1,x_2] \times [y_1, y_2]$.
+Accordingly, $r^*$ can be written as $[x_1, x_2] \times [y_1, \maxtop(r')]$.
+
+\vgap
+
+Firstly, we will prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
 For this purpose, given any segment $h \in \dcross_H(\bm{t})$, we will show that $h$ crosses $r^*$.
@@
 
-Next, assuming that the segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates,
+Secondly, assuming that the segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates,
 we will prove that $\dcross_H(\bm{t})$ includes a prefix of the sorted order. It suffices to establish the following equivalent statement:

### 2024-06-02 msj.tex
yufei

@@
 \def\vgap{\vspace{0mm}}
-\def\extraspacing{\vspace{1.5mm} \noindent}
+\def\extraspacing{\vspace{2mm} \noindent}
 \def\figcapup{\vspace{-2mm}}
-\def\figcapdown{\vspace{-2mm}}
+\def\figcapdown{\vspace{-3mm}}
 
@@
 \begin{figure}
-    \includegraphics[height=27mm]{./artwork/guard}
+    \includegraphics[height=25mm]{./artwork/guard}
 
@@
 
-\def\vgap{\vspace{2mm}}
+\def\vgap{\vspace{1mm}}
 \def\extraspacing{\vspace{2mm} \noindent}
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} We will prove the lemma by providing a way to map each $(k-2)$-tuple $\J(R_1',...,R_{k-2}')$ to a unique $k$-tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
+\noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} We will prove the lemma by providing a way to map each $(k-2)$-tuple $\J(R_1',...,R_{k-2}')$ to a unique $k$-tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
 
@@
 
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.
+\noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.
 

### 2024-06-02 msj.tex
yufei

@@
 \def\vgap{\vspace{0mm}}
-\def\extraspacing{\vspace{2mm} \noindent}
+\def\extraspacing{\vspace{1.5mm} \noindent}
 \def\figcapup{\vspace{-2mm}}
@@
 
-    \vspace{3mm}
+    \vspace{2mm}
     \figcapup 
@@
     where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the $k$-SJ algorithm.
-
-%     time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
 \end{theorem}
@@
 Let $r'_i$ ($i \in [k-2])$ be the trimmed rectangle of $r_i$, and set $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2})$.
-Then, we must have
-        \myitems{
-            \item $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$;
-            \item $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$. 
-        }
-
-        \vspace{2mm}
+Then, we must have (i) $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, and (ii) $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
 
@@
 \def\vgap{\vspace{1mm}}
-\def\extraspacing{\vspace{2mm} \noindent}
+\def\extraspacing{\vspace{1.5mm} \noindent}
 
@@
 
-\extraspacing{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$.
+\extraspacing \underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$.
 After this, we can then derive

### 2024-06-02 msj.tex
yufei

@@
 
-We now prove the first bullet of statement (2), namely, $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+We now prove the first claim of statement (2), namely, $\bm{t}\in \J(R_1',...,R_{k-2}')$.
 Consider the $k$-tuple $(r_1,...,r_{k-2},h,v)\in \J_1$ given in statement (2).
@@
 
-Next, we prove the second bullet, namely, $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+Next, we prove the second claim, namely, $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
 It suffices to show only the former due to symmetry.

### 2024-06-02 msj.tex
lmm 4.1

@@
 \section{Supplementary Proofs for Section~\ref{sec:hv:type1}} \label{app:hv:type1}
-
-\noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} We will prove the lemma by providing a way to map each $(k-2)$-tuple $\J(R_1',...,R_{k-2}')$ to a unique $k$-tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
-
-\vgap
-
-To explain the mapping, consider any tuple $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+We defer the proof of Lemma~\ref{lmm:hv:type1:recur-output} to the paragraph following the proof of Lemma~\ref{lmm:hv:type1:properties}, as the latter proof requires results in the former.
+
+% \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} We will prove the lemma by providing a way to map each $(k-2)$-tuple $\J(R_1',...,R_{k-2}')$ to a unique $k$-tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
+
+% \vgap
+
+% To explain the mapping, consider any tuple $\bm{t} \in \J(R_1',...,R_{k-2}')$.
 Recall that $B_\bm{t}$ is the intersection of the $k-2$ rectangles $\bm{t}[1], ..., \bm{t}[k-2]$.
@@
 
-\vgap
-
-Recall that each $\bm{t}[z]$ (for $z \in [k-2]$) is a trimmed rectangle; let us denote by $r_z$ the full rectangle of $\bm{t}[z]$. Regarding the values $i, j$ decided earlier, let
-\myitems{
-    \item $h$ be the lowest segment in $H$ that crosses $r_i$;
-    \item $v$ be the leftmost segment in $V$ that crosses $r_j$.
-}
-By the way trimmed rectangles are computed, we know that $h$ must contain the bottom edge of rectangle $\bm{t}[i]$, and $v$ must contain the left edge of rectangle $\bm{t}[j]$.
+% \vgap
+
+% Recall that each $\bm{t}[z]$ (for $z \in [k-2]$) is a trimmed rectangle; let us denote by $r_z$ the full rectangle of $\bm{t}[z]$. Regarding the values $i, j$ decided earlier, let
+% \myitems{
+%     \item $h$ be the lowest segment in $H$ that crosses $r_i$;
+%     \item $v$ be the leftmost segment in $V$ that crosses $r_j$.
+% }
+% By the way trimmed rectangles are computed, we know that $h$ must contain the bottom edge of rectangle $\bm{t}[i]$, and $v$ must contain the left edge of rectangle $\bm{t}[j]$.
 This means that point $p$ must be the intersection point of $h$ and $v$.
 
-\vgap
-
-We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ belongs to $\J(R_1,...,$ $R_{k-2},H,V)$.
+% \vgap
+
+% We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ belongs to $\J(R_1,...,$ $R_{k-2},H,V)$.
 Clearly, $r_z$ covers $\bm{t}[z]$ for all $z \in [k-2]$.
@@
 
-\vgap
-
-It remains to prove that no two distinct tuples $\bm{t}_1, \bm{t}_2 \in \J(R_1',...,$ $R_{k-2}')$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},$ $H, V)$.
+% \vgap
+
+% It remains to prove that no two distinct tuples $\bm{t}_1, \bm{t}_2 \in \J(R_1',...,$ $R_{k-2}')$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},$ $H, V)$.
 Assume that this is not true, namely, both $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
@@
 This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).
+
+\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}).
+As we proved in Statement (3) of Lemma~\ref{lmm:hv:type1:properties}, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ are non-empty.
+Therefore, for each $\bm{t} \in \J(R_1',...,R_{k-2}')$, the algorithm generate-$\J^*$ will add to $\J^*$ at least one distinct tuple.
+This implies that $|\J(R_1',...,R_{k-2}')| \le |\J^*|\le \out$, where the second inequality follows from Statement (1) of Lemma~\ref{lmm:hv:type1:properties}.
 

### 2024-06-02 msj.tex
lmm 5.1

@@
 
-\noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.
-
-\vgap
-
-To describe the mapping, consider an arbitrary $(k-1)$-tuple $\bm{t} = (r_1,...,$ $r_{k-3},$ $h', v) \in \J(R_1,...,$ $R_{k-3},H',V)$.
+We first prove Lemma~\ref{lmm:hv:type2:properties} and then proceed to prove Lemma~\ref{lmm:hv:type2:recur-output}, because the latter proof requires results in the former.
+
+% \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.
+
+% \vgap
+
+% To describe the mapping, consider an arbitrary $(k-1)$-tuple $\bm{t} = (r_1,...,$ $r_{k-3},$ $h', v) \in \J(R_1,...,$ $R_{k-3},H',V)$.
 Let $h$ be the full segment of $h'$.
@@
 
-\vgap
-
-We argue that $(r_1,...,r_{k-3}, r_{k-2},h,v)$ belongs to $\J_2$.
+% \vgap
+
+% We argue that $(r_1,...,r_{k-3}, r_{k-2},h,v)$ belongs to $\J_2$.
 Let $p$ be $h' \cap v$ (the intersection point of $h'$ and $v$).
@@
 
-\vgap
-
-It remains to show that no two distinct $(k-1)$-tuples $\bm{t}_1$ and $\bm{t}_2$ can be mapped to the same tuple in $\J_2$.
+% \vgap
+
+% It remains to show that no two distinct $(k-1)$-tuples $\bm{t}_1$ and $\bm{t}_2$ can be mapped to the same tuple in $\J_2$.
 Assume, on the contrary, that $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.
@@
 
-\vgap
-
-We can now conclude that $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality used the definition in \eqref{eqn:hv:type2:J2}.
+% \vgap
+
+% We can now conclude that $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality used the definition in \eqref{eqn:hv:type2:J2}.
 
@@
 Therefore, the above expression is at most $\out$, as claimed in statement (3).
+
+\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
+First, we prove that $\cross_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
+Let $\bm{t} = (r_1,...,r_{k-3},h',v)$.
+By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ that covers both the left endpoint of $h$ and the right endpoint of $h'$.
+Thus, $r_{k-2}$ contains $h'$, because $h'$ shares the same left endpoint as $h$.
+Since the effective horizontal segment of $\bm{t}$ is contained in $h'$, it is also covered by $r_{k-2}$.
+This implies that $\cross_{R_{k-2}}(\bm{t})$ includes $r_{k-2}$, and is therefore non-empty.
+
+\vgap
+
+Then, we resorts to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}).
+For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, this algorithm adds to $\J_2$ exactly $|\cross_{R_{k-2}}(\bm{t})|\geq 1$ tuples.
+Therefore, $|\J(R_1, ..., R_{k-3}, H', V)| \le |\J_2| \le \out$, where the last inequality follows from the definition of $\J_2$ (see \eqref{eqn:hv:type2:J2}).
 

### 2024-06-02 msj.tex
lmm 5.1
lmm4.1

@@
 First, we prove that $\cross_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
-Let $\bm{t} = (r_1,...,r_{k-3},h',v)$.
+Let $h$ be the full segment of $\bm{t}[k-2]$.
 By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ that covers both the left endpoint of $h$ and the right endpoint of $h'$.
@@
 For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, this algorithm adds to $\J_2$ exactly $|\cross_{R_{k-2}}(\bm{t})|\geq 1$ tuples.
-Therefore, $|\J(R_1, ..., R_{k-3}, H', V)| \le |\J_2| \le \out$, where the last inequality follows from the definition of $\J_2$ (see \eqref{eqn:hv:type2:J2}).
+By the correctness of the algorithm, we have $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the last inequality follows from the definition of $\J_2$ (see \eqref{eqn:hv:type2:J2}).
 

### 2024-06-02 msj.tex
yufei

@@
 \section{Supplementary Proofs for Section~\ref{sec:hv:type1}} \label{app:hv:type1}
-We defer the proof of Lemma~\ref{lmm:hv:type1:recur-output} to the paragraph following the proof of Lemma~\ref{lmm:hv:type1:properties}, as the latter proof requires results in the former.
-
-% \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} We will prove the lemma by providing a way to map each $(k-2)$-tuple $\J(R_1',...,R_{k-2}')$ to a unique $k$-tuple in the H-V $k$-SJ result $\J(R_1,...,R_{k-2},H,V)$.
-
-% \vgap
-
-% To explain the mapping, consider any tuple $\bm{t} \in \J(R_1',...,R_{k-2}')$.
-Recall that $B_\bm{t}$ is the intersection of the $k-2$ rectangles $\bm{t}[1], ..., \bm{t}[k-2]$.
-Let $p$ be the bottom-left corner of $B_\bm{t}$.
-Define $i$ and $j$ to be values in $[k-2]$ such that $\bm{t}[i] = \gbot(\bm{t})$ and $\bm{t}[j] = \gleft(\bm{t})$; note that $i$ and $j$ can be the same.
-Point $p$ is the intersection of the bottom edge of $\bm{t}[i]$ and the left edge of $\bm{t}[j]$.
-
-% \vgap
-
-% Recall that each $\bm{t}[z]$ (for $z \in [k-2]$) is a trimmed rectangle; let us denote by $r_z$ the full rectangle of $\bm{t}[z]$. Regarding the values $i, j$ decided earlier, let
-% \myitems{
-%     \item $h$ be the lowest segment in $H$ that crosses $r_i$;
-%     \item $v$ be the leftmost segment in $V$ that crosses $r_j$.
-% }
-% By the way trimmed rectangles are computed, we know that $h$ must contain the bottom edge of rectangle $\bm{t}[i]$, and $v$ must contain the left edge of rectangle $\bm{t}[j]$.
-This means that point $p$ must be the intersection point of $h$ and $v$.
-
-% \vgap
-
-% We argue that the $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ belongs to $\J(R_1,...,$ $R_{k-2},H,V)$.
-Clearly, $r_z$ covers $\bm{t}[z]$ for all $z \in [k-2]$.
-Thus, $\bigcap_{z=1}^{k-2} r_z$ covers $\bigcap_{z=1}^{k-2} \bm{t}[z]$, which is $B_\bm{t}$.
-Hence, point $p$ --- which is the intersection point of $h$ and $v$, and also a corner of $B_\bm{t}$ --- falls in $\bigcap_{z=1}^{k-2} r_z$, indicating $h \cap v \cap \bigcap_{z=1}^{k-2} r_z \ne \emptyset$.
-We map $\bm{t}$ to $(r_1, ..., r_{k-2}, h, v)$.
-
-% \vgap
-
-% It remains to prove that no two distinct tuples $\bm{t}_1, \bm{t}_2 \in \J(R_1',...,$ $R_{k-2}')$ can be mapped to the same tuple in $\J(R_1,..., R_{k-2},$ $H, V)$.
-Assume that this is not true, namely, both $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1, ..., r_{k-2}, h, v) \in \J(R_1,..., R_{k-2}, H, V)$.
-However, in this case, $\bm{t}_1[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$ and, at the same time, $\bm{t}_2[z]$ is the trimmed rectangle of $r_z$ for all $z \in [k-2]$.
-We thus have $\bm{t}_1 = \bm{t}_2$, giving a contradiction.
+
+This section supplies the proofs postponed from Section~\ref{sec:hv:type1}.
+The order of the proofs here will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type1}.
+In particular, we will prove Lemma~\ref{lmm:hv:type1:recur-output} before Lemma~\ref{lmm:hv:type1:properties}, because the claims of the former lemma can be used to construct a simple argument for the latter.
 
@@
 This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}.
-By statement (2), the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$.
+By statement (1) of Lemma~\ref{lmm:hv:type2:properties}, the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$.
 This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).
 
-\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}).
-As we proved in Statement (3) of Lemma~\ref{lmm:hv:type1:properties}, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ are non-empty.
-Therefore, for each $\bm{t} \in \J(R_1',...,R_{k-2}')$, the algorithm generate-$\J^*$ will add to $\J^*$ at least one distinct tuple.
-This implies that $|\J(R_1',...,R_{k-2}')| \le |\J^*|\le \out$, where the second inequality follows from Statement (1) of Lemma~\ref{lmm:hv:type1:properties}.
+\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} Our proof again resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}).
+As argued in Statement (3) of Lemma~\ref{lmm:hv:type1:properties}, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ are non-empty for any $\bm{t} \in \J(R_1',$ $...,R_{k-2}')$.
+Therefore, in processing this $\bm{t}$, the algorithm adds at least a new tuple to $\J^*$.
+Hence, $|\J(R_1',...,R_{k-2}')| \le |\J^*|\le \out$, where the second inequality follows from Statement (1) of Lemma~\ref{lmm:hv:type1:properties}.
 
@@
 
-We first prove Lemma~\ref{lmm:hv:type2:properties} and then proceed to prove Lemma~\ref{lmm:hv:type2:recur-output}, because the latter proof requires results in the former.
-
-% \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} We will prove the lemma by mapping each tuple of $\J(R_1,...,R_{k-3},H',V)$ to a unique tuple in $\J_2$.
-
-% \vgap
-
-% To describe the mapping, consider an arbitrary $(k-1)$-tuple $\bm{t} = (r_1,...,$ $r_{k-3},$ $h', v) \in \J(R_1,...,$ $R_{k-3},H',V)$.
-Let $h$ be the full segment of $h'$.
-By the definition of a trimmed segment (see Section~\ref{sec:hv:type2}), at least one rectangle $r_{k-2} \in R_{k-2}$ covers both the left endpoint of $h$ and the right endpoint of $h'$.
-We map $\bm{t}$ to the $k$-tuple $(r_1,...,r_{k-3}, r_{k-2},h,v)$.
-It is worth noting that $r_{k-2}$ definitely contains the entire $h'$ (which has the same left point as $h$, implying that $r_{k-2}$ covers both endpoints of $h'$).
-
-% \vgap
-
-% We argue that $(r_1,...,r_{k-3}, r_{k-2},h,v)$ belongs to $\J_2$.
-Let $p$ be $h' \cap v$ (the intersection point of $h'$ and $v$).
-As $h'$ is contained in $h$, it must hold that $h \cap v = p$.
-Since $(r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$, we know that $p$ falls in all of $r_1, ..., r_{k-3}$.
-Our earlier discussion already shows that $r_{k-2}$ contains $h'$; this means that $p$ falls in $r_{k-2}$.
-Thus, $h \cap v \cap \bigcap_{i = 1}^{k-2}r_i$ cannot be empty as it equals $p$.
-We can now conclude that $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$.
-This, together with the fact that $r_{k-2}$ covers the left endpoint of $h$, tells us $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.
-
-% \vgap
-
-% It remains to show that no two distinct $(k-1)$-tuples $\bm{t}_1$ and $\bm{t}_2$ can be mapped to the same tuple in $\J_2$.
-Assume, on the contrary, that $\bm{t}_1$ and $\bm{t}_2$ are mapped to $(r_1,...,r_{k-3}, r_{k-2},h,v) \in \J_2$.
-Then, $\bm{t}_1[i] = \bm{t}_2[i] = r_i$ for each $i \in [k-3]$, and $\bm{t}_1[k-1] = \bm{t}_2[k-1] = v$.
-Moreover, $\bm{t}_1[k-2]$ and $\bm{t}_2[k-2]$ are both the trimmed segment of $h$.
-But then this means $\bm{t}_1 = \bm{t}_2$, giving a contradiction.
-
-% \vgap
-
-% We can now conclude that $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the second inequality used the definition in \eqref{eqn:hv:type2:J2}.
+This section supplies the proofs postponed from Section~\ref{sec:hv:type2}.
+The order of the proofs here will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type2}.
+In particular, we will prove Lemma~\ref{lmm:hv:type2:recur-output} before Lemma~\ref{lmm:hv:type2:properties}, because the claims of the former lemma can be used to construct a simple argument for the latter.
 
@@
 
-\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.}
-First, we prove that $\cross_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
+\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} First, we prove that $\cross_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
 Let $h$ be the full segment of $\bm{t}[k-2]$.

### 2024-06-02 msj.tex
yufei

@@
 
-\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} First, we prove that $\cross_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
+\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} Before proving the lemma, let us prove a fact that will be useful later: the set $\contained_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
 Let $h$ be the full segment of $\bm{t}[k-2]$.
-By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ that covers both the left endpoint of $h$ and the right endpoint of $h'$.
-Thus, $r_{k-2}$ contains $h'$, because $h'$ shares the same left endpoint as $h$.
-Since the effective horizontal segment of $\bm{t}$ is contained in $h'$, it is also covered by $r_{k-2}$.
-This implies that $\cross_{R_{k-2}}(\bm{t})$ includes $r_{k-2}$, and is therefore non-empty.
-
-\vgap
-
-Then, we resorts to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}).
-For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, this algorithm adds to $\J_2$ exactly $|\cross_{R_{k-2}}(\bm{t})|\geq 1$ tuples.
-By the correctness of the algorithm, we have $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the last inequality follows from the definition of $\J_2$ (see \eqref{eqn:hv:type2:J2}).
+By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ that covers both endpoints of $h$.
+Thus, the effective horizontal segment of $\bm{t}$, which is contained in $h'$, must also covered by $r_{k-2}$.
+This implies that $\cross_{R_{k-2}}(\bm{t})$ cannot be empty (as it includes at least $r_{k-2}$).
+
+\vgap
+
+With the above fact, next we prove Lemma~\ref{lmm:hv:type1:recur-output} by resorting to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}).
+For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, this algorithm adds $|\contained_{R_{k-2}}(\bm{t})|\geq 1$ tuples to $\J_2$.
+Therefore, $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the last inequality used the definition of $\J_2$; see \eqref{eqn:hv:type2:J2}.
 

### 2024-06-02 msj.tex
yufei

@@
 
-With the above fact, next we prove Lemma~\ref{lmm:hv:type1:recur-output} by resorting to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}).
+With the above fact, next we prove Lemma~\ref{lmm:hv:type2:recur-output} by resorting to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}).
 For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, this algorithm adds $|\contained_{R_{k-2}}(\bm{t})|\geq 1$ tuples to $\J_2$.

### 2024-06-02 msj.tex
yufei

@@
 The order of the proofs here will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type1}.
-In particular, we will prove Lemma~\ref{lmm:hv:type1:recur-output} before Lemma~\ref{lmm:hv:type1:properties}, because the claims of the former lemma can be used to construct a simple argument for the latter.
+In particular, we will prove Lemma~\ref{lmm:hv:type1:properties} before Lemma~\ref{lmm:hv:type1:recur-output}, because the claims of the former lemma can be used to construct a simple argument for the latter.
 

### 2024-06-02 msj.tex
yufei

@@
 The order of the proofs here will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type2}.
-In particular, we will prove Lemma~\ref{lmm:hv:type2:recur-output} before Lemma~\ref{lmm:hv:type2:properties}, because the claims of the former lemma can be used to construct a simple argument for the latter.
+In particular, we will prove Lemma~\ref{lmm:hv:type2:properties} before Lemma~\ref{lmm:hv:type2:recur-output}, because the claims of the former lemma can be used to construct a simple argument for the latter.
 

### 2024-06-02 msj.tex

@@
 Thus, the effective horizontal segment of $\bm{t}$, which is contained in $h'$, must also covered by $r_{k-2}$.
-This implies that $\cross_{R_{k-2}}(\bm{t})$ cannot be empty (as it includes at least $r_{k-2}$).
+This implies that $\contained_{R_{k-2}}(\bm{t})$ cannot be empty (as it includes at least $r_{k-2}$).
 
@@
 
-\extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\cross_{R_{k-2}}(\bm{t}) = \cross_{R_{k-2}}(h^*)$.
+\extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\contained_{R_{k-2}}(\bm{t}) = \contained_{R_{k-2}}(h^*)$.
 We can then derive that
 \myeqn{
-    \sum_{h^* \in H^*} |\cross_{R_{k-2}}(h^*)| \le \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\cross_{R_{k-2}}(\bm{t})| \leq \out, \nn
+    \sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)| 
+    &\le& \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\contained_{R_{k-2}}(\bm{t})| \nn \\
+    &\leq& \out, \nn
 } where the last inequality is due to Statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
@@
 
-Next, we will prove that a rectangle $r_{k-2} \in \cross_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \cross_{R_{k-2}}(\bm{t})$.
-According to the definitions of $\cross_{R_{k-2}}(h^*)$ and $\cross_{R_{k-2}}(\bm{t})$ --- see \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
+Next, we will prove that a rectangle $r_{k-2} \in \contained_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
+According to the definitions of $\contained_{R_{k-2}}(h^*)$ and $\contained_{R_{k-2}}(\bm{t})$ --- see \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
 Denote by $x$ the x-coordinate of the vertical segment $\bm{t}[k-1]$.
@@
 
-Firstly, we will prove $\contained_{R_{k-2}}(\bm{t})\subseteq \cross_{R_{k-2}}(h^*)$.
+Firstly, we will prove $\contained_{R_{k-2}}(\bm{t})\subseteq \contained_{R_{k-2}}(h^*)$.
 For this purpose, we will show that any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ also covers $h^*$.

### 2024-06-02 msj.tex
yufei

@@
 \extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\contained_{R_{k-2}}(\bm{t}) = \contained_{R_{k-2}}(h^*)$.
-We can then derive that
-\myeqn{
-    \sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)| 
-    &\le& \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\contained_{R_{k-2}}(\bm{t})| \nn \\
-    &\leq& \out, \nn
+We can then derive
+\myeqn{
+    && \sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)|  \nn \\ 
+    &\le& \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\contained_{R_{k-2}}(\bm{t})| 
+    \leq \out \nn
 } where the last inequality is due to Statement (3) of Lemma~\ref{lmm:hv:type2:properties}.

### 2024-06-03 msj.tex
yufei

@@
 \begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^{k-1} (k!)^3 \cdot (n \log n + k \cdot \out))$ time.
-    %and $O(k \cdot n + \out)$ space, where $c > 1$ is a positive constant.
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k \cdot (k!)^3 \cdot (n \log n + k \cdot \out))$ time, where $c > 1$ is a positive constant.
+    %and $O(k \cdot n + \out)$ space,
 \end{theorem}
@@
 
-\def\vgap{\vspace{1mm}}
+\def\vgap{\vspace{0.5mm}}
 \def\extraspacing{\vspace{1.5mm} \noindent}

### 2024-06-03 msj.tex
yufei

@@
-\documentclass[sigconf]{acmart}
+%\documentclass[sigconf]{acmart}
+\documentclass[acmsmall,review,anonymous]{acmart}
 % \documentclass{...}
@@
 
-\def\vgap{\vspace{0mm}}
-\def\extraspacing{\vspace{1.5mm} \noindent}
+\def\vgap{\vspace{1mm}}
+\def\extraspacing{\vspace{3mm} \noindent}
 \def\figcapup{\vspace{-2mm}}
@@
 
-    \vspace{2mm}
+    \vspace{4mm}
     \figcapup 
@@
 We simplify the argument for 3D 3-SJ and present the full reduction in Appendix~\ref{app:lb-cond}.
+\section{Preliminaries in Geometry} \label{sec:bricks}
+
+This section will first define some notions frequently used in our presentation and then introduce several  geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
+
+\extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$.
+We say that a horizontal segment $h_1$ is {\em lower} (resp., {\em higher}) than another horizontal segment $h_2$ if the y-coordinate of $h_1$ is smaller (resp., larger) than that of $h_2$.
+Similarly, a vertical segment $v_1$ is {\em to the left} (resp., {\em right}) {\em of} another vertical segment $v_2$ if the x-coordinate of $v_1$ is smaller (resp., larger) than that of $v_2$.
+
+\vgap
+
+%Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. \yf{need this?}
+
+Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
+A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s \cap r \ne \emptyset$ but $r$ covers neither of the two endpoints of $s$.
+A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
+
+\vgap
+
+Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
+\myeqn{
+    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}
+    \label{eqn:cross}
+}
+namely, the set of segments in $S$ crossing $r$.
+Let $R$ be a set of rectangles. Given a horizontal segment, we define 
+\myeqn{
+    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}
+    \label{eqn:contained}
+}
+namely, the set of rectangles in $R$ containing $h$.
 
@@
 
+\vgap
+
+Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$.
+Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
+We define
+\myeqn{
+    B_\bm{t} &=& \cap_{i=1}^t r_i \label{eqn:B_t}
+}
+namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, define:
+\myitems{
+    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$.
+In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
+
+    \item $\gbot(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$.
+In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
+}
+See Figure~\ref{fig:guard} for an illustration. It is worth mentioning that since a horizontal segment $h$ is a degenerated rectangle, notations such as $\xleft(h)$ and $\xright(h)$ are well-defined.
+
+\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of 2D points and set $R$ of rectangles.
+In the {\em detection version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$.
+Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$.
+This can be achieved in $O(n \log n)$ time where $n = |P| + |R|$ as shown
+in Appendix~\ref{app:bricks}.
+
+\vgap
+
+In the {\em reporting version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$.
+In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_1, \emptyset), (p_2: r_1, r_2), (p_3: r_2)}$.
+As shown in Appendix~\ref{app:bricks}, this can be achieved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
+
+\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments.
+The goal is to report, for each segment $h \in H$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$.
+If $h$ does not intersect with any segment in $V$, report nothing for $h$.
+Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+This problem can be solved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.
+
+\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
+The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ such that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that at least one rectangle $r \in R$ covers both the point $(x_1, y)$ and the point $(x, y)$.
+If the point $p$ exists (i.e., $h$ has at least one left-end covering rectangle in $R$), we should output a tuple $(h, p)$; otherwise, output nothing for $h$.
+Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+This problem can be solved in in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
+
+\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
+In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see its definition in \eqref{eqn:cross}).
+If no segment in $H$ crosses $r$, output nothing for $r$.
+Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
+This can be achieved in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
+
 \begin{figure*}
-    \begin{tabular}{ccccc}
+    \begin{tabular}{cc}
         \includegraphics[height=30mm]{./artwork/prob-a} &
-        \hspace{3mm}
-        \includegraphics[height=27mm]{./artwork/prob-b} &
-        \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-c} &
-        \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-d} &
-        \hspace{3mm}
-        \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
+        \includegraphics[height=27mm]{./artwork/prob-b}  \\[2mm]
         (a) Problems $\mathscr{A}$ &
         \hspace{3mm}
-        (b) Problem $\mathscr{B}$ &
-        \hspace{3mm}
+        (b) Problem $\mathscr{B}$ \\[3mm]
+    \end{tabular}
+    %\vspace{3mm}
+    \begin{tabular}{ccc}
+        \includegraphics[height=30mm]{./artwork/prob-c} &
+        \includegraphics[height=30mm]{./artwork/prob-d} &
+        \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
         (c) Problem $\mathscr{C}$ &
-        \hspace{3mm}
         (d) Problem $\mathscr{D}$ &
-        \hspace{3mm}
         (e) Problem $\mathscr{E}$
@@
 
-\section{Preliminaries in Geometry} \label{sec:bricks}
-
-This section will first define some notions frequently used in our presentation and then introduce several  geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
-
-\extraspacing {\bf Terminology.} A {\em horizontal} segment is a segment of the form $[x_1, x_2] \times y$, and a {\em vertical} segment is a segment of the form $x \times [y_1, y_2]$.
-We say that a horizontal segment $h_1$ is {\em lower} (resp., {\em higher}) than another horizontal segment $h_2$ if the y-coordinate of $h_1$ is smaller (resp., larger) than that of $h_2$.
-Similarly, a vertical segment $v_1$ is {\em to the left} (resp., {\em right}) {\em of} another vertical segment $v_2$ if the x-coordinate of $v_1$ is smaller (resp., larger) than that of $v_2$.
-
-\vgap
-
-%Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. \yf{need this?}
-
-Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
-A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s \cap r \ne \emptyset$ but $r$ covers neither of the two endpoints of $s$.
-A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
-
-\vgap
-
-Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
-\myeqn{
-    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}
-    \label{eqn:cross}
-}
-namely, the set of segments in $S$ crossing $r$.
-Let $R$ be a set of rectangles. Given a horizontal segment, we define 
-\myeqn{
-    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}
-    \label{eqn:contained}
-}
-namely, the set of rectangles in $R$ containing $h$.
-
-\vgap
-
-Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$.
-Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
-We define
-\myeqn{
-    B_\bm{t} &=& \cap_{i=1}^t r_i \label{eqn:B_t}
-}
-namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, define:
-\myitems{
-    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$.
-In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
-
-    \item $\gbot(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\ybot(r_i)$ $= \ybot(B_\bm{t})$.
-In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.
-}
-See Figure~\ref{fig:guard} for an illustration. It is worth mentioning that since a horizontal segment $h$ is a degenerated rectangle, notations such as $\xleft(h)$ and $\xright(h)$ are well-defined.
-
-\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of 2D points and set $R$ of rectangles.
-In the {\em detection version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$.
-Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$.
-This can be achieved in $O(n \log n)$ time where $n = |P| + |R|$ as shown
-in Appendix~\ref{app:bricks}.
-
-\vgap
-
-In the {\em reporting version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$.
-In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_1, \emptyset), (p_2: r_1, r_2), (p_3: r_2)}$.
-As shown in Appendix~\ref{app:bricks}, this can be achieved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
-
-\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments.
-The goal is to report, for each segment $h \in H$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$.
-If $h$ does not intersect with any segment in $V$, report nothing for $h$.
-Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-This problem can be solved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.
-
-\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
-The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ such that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that at least one rectangle $r \in R$ covers both the point $(x_1, y)$ and the point $(x, y)$.
-If the point $p$ exists (i.e., $h$ has at least one left-end covering rectangle in $R$), we should output a tuple $(h, p)$; otherwise, output nothing for $h$.
-Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-This problem can be solved in in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
-
-\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
-In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see its definition in \eqref{eqn:cross}).
-If no segment in $H$ crosses $r$, output nothing for $r$.
-Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
-This can be achieved in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
-
 \vgap
@@
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
-
-\section{The Core: H-V Multiway Spatial Joins} \label{sec:hv}
-
-Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ...,$ $R_k$.
-We now formulate a special version of $k$-SJ, named the {\em H-V $k$-SJ problem}.
-The special nature is reflected in the introduction of three constraints: (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
-For better clarity, we will represent the input sets as $R_1, R_2, ..., R_{k-2}, H$ ($=R_{k-1}$), and $V$ ($=R_k$).
-The goal is to output the join result $\J(R_1, ..., R_{k-2},$ $H, V)$, including every $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
-\times R_{k-2} \times H \times V$ such that $h \cap v \cap \bigcap_{i=1}^{k-2} r_i$ is not empty.
-
-\vgap
-
-Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
-To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time.
-We assume the existence of an algorithm $\A$ that can settle $\kappa$-SJ for all $\kappa \in [1, k-1]$.
-Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
-We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \ge 1$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
-
-\vgap
-
-We will establish:
-
-\begin{lemma} \label{lmm:hv}
-    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
-    \myeqn{
-        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
-    }
-    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the H-V $k$-SJ algorithm.
-\end{lemma}
-
-The part of the paper from this point till the end of Section~\ref{sec:hv:type2} will be devoted to proving the above lemma.
-This is the most challenging step in solving the general $k$-SJ problem optimally, as will be discussed in Section~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
-
-\vgap
-
-%The condition is equivalent to saying that the intersection point of $h$ and $v$ falls in every $r_i$ of $i \in [k-2]$.
-
-Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ in the join result $\J(R_1, ...,$ $R_{k-2}, H, V)$. We classify the tuple into one of the two types below:
-
-\myitems{
-    \item {{\bf Type 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$ and, at the same time, $v$ crosses all of $r_1, ..., r_{k-2}$;
-
-    \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
-    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
-}
-Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.
-In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.
-
-\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently.
-They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
-Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
-We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.
-Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.
 
@@
 
+\section{The Core: H-V Multiway Spatial Joins} \label{sec:hv}
+
+Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ...,$ $R_k$.
+We now formulate a special version of $k$-SJ, named the {\em H-V $k$-SJ problem}.
+The special nature is reflected in the introduction of three constraints: (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
+For better clarity, we will represent the input sets as $R_1, R_2, ..., R_{k-2}, H$ ($=R_{k-1}$), and $V$ ($=R_k$).
+The goal is to output the join result $\J(R_1, ..., R_{k-2},$ $H, V)$, including every $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
+\times R_{k-2} \times H \times V$ such that $h \cap v \cap \bigcap_{i=1}^{k-2} r_i$ is not empty.
+
+\vgap
+
+Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
+To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time.
+We assume the existence of an algorithm $\A$ that can settle $\kappa$-SJ for all $\kappa \in [1, k-1]$.
+Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
+We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \ge 1$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
+
+\vgap
+
+We will establish:
+
+\begin{lemma} \label{lmm:hv}
+    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
+    \myeqn{
+        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
+    }
+    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the H-V $k$-SJ algorithm.
+\end{lemma}
+
+The part of the paper from this point till the end of Section~\ref{sec:hv:type2} will be devoted to proving the above lemma.
+This is the most challenging step in solving the general $k$-SJ problem optimally, as will be discussed in Section~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
+
+\vgap
+
+%The condition is equivalent to saying that the intersection point of $h$ and $v$ falls in every $r_i$ of $i \in [k-2]$.
+
+Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ in the join result $\J(R_1, ...,$ $R_{k-2}, H, V)$. We classify the tuple into one of the two types below:
+
+\myitems{
+    \item {{\bf Type 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$ and, at the same time, $v$ crosses all of $r_1, ..., r_{k-2}$;
+
+    \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
+    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
+}
+Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.
+In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.
+
+\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently.
+They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
+Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
+Another structure with the same guarantees can be found in \cite{oa19}.
+We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.
+Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.
+
+\section{H-V $k$-SJ: Result Tuples of Type 1} \label{sec:hv:type1}
+
+As before, let $R_1, ..., R_{k-2}, H$, and $V$ be the input sets of the H-V $k$-SJ problem.
+Denote by $\J_1$ the set of type-1 result tuples defined in Section~\ref{sec:hv}.
+In this section, we aim to compute a set $\J^*$ satisfying
+\myeqn{
+    \J_1
+    \subseteq
+    \J^*
+    \subseteq
+    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:J*}
+}
+where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ.
+Remember that the output size $\out$ is defined as $|\J(R_1, ..., R_{k-2}, H, V)|$.
+From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.
+
+\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
+The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
+\qed
+
+\extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r \in R_i$, we compute four segments:
+\myitems{
+    \item $h_\bot$ (resp., $h_\top$): the lowest (resp., highest) segment in $H$ that crosses $r$;
+
+    \item $v_\vdash$ (resp., $v_\dashv$): the leftmost (resp., rightmost) segment in $V$ that crosses $r$.
+}
+Define $r' = [x_\vdash, x_\dashv] \times [y_\bot, y_\top]$, where $x_\vdash$ (resp., $x_\dashv$) is the x-coordinate of $v_\vdash$ (resp., $v_\dashv$), and $y_\bot$ (resp., $y_\top$) is the y-coordinate of $h_\bot$ (resp., $h_\top$).
+We say that $r'$ is the {\em trimmed rectangle} of $r$, and conversely, $r$ is the {\em full rectangle} of $r'$.
+Note that $r'$ exists if and only if $r$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
+
+%Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$.
+
+\vgap
+
+Construct
+\myeqn{
+    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \nn
+}
+Computing the ``segment $h_\bot$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input).
+By symmetry, so is the computation of $h_\top, v_\vdash$, and $v_\dashv$ segments for each $r \in R_i$.
+It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(kn \log n)$ total time.
+
+\vgap
+
+We now solve a $(k-2)$-SJ problem on the input $\set{R_1', R_2', ..., R_{k-2}'}$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
+This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$.
+We prove in Appendix~\ref{app:hv:type1}:
+
+\begin{lemma} \label{lmm:hv:type1:recur-output}
+    $|\J(R_1', R'_2, ..., R'_{k-2})| \le \out$.
+\end{lemma}
+
+As a corollary of Lemma~\ref{lmm:hv:type1:recur-output}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
+
+\extraspacing {\underline{\em Example 4.2.}} Figure~\ref{fig:hv:type1:ex}b shows the rectangles in $R'_1 = \set{\alpha'}$ and $R'_2$ $=$ $\{\beta'_1,$ $\beta'_2\}$.
+For instance, $\alpha'$, which is trimmed from rectangle $\alpha$, is decided by $h_\bot$ $=$ $h_2$, $h_\top$ $=$ $h_6$, $v_\vdash = v_3$, and $v_\dashv = v_5$.
+The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2, ..., R'_{k-2}) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
+\qed
+
 \begin{figure*}
-    \begin{tabular}{cccc}
-        \hspace{-5mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex1-a} &
-        \hspace{-4mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex1-b} & 
-        \hspace{-4mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex1-c} &
-        \hspace{-4mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex1-d} \\ 
-        \hspace{-5mm} (a) $R_1$, $R_2$, $H$, and $V$ &
-        \hspace{-4mm} (b) $R_1'$ and $R_2'$ &
-        \hspace{-4mm} (c) $R_1'$, $R_2'$, $H$, and $V$ &
-        \hspace{-4mm} (d) $R_1^*$, $R_2^*$ ($= \emptyset$), and $H$
+    \begin{tabular}{cc}
+        %\hspace{-5mm}
+        \includegraphics[height=48mm]{./artwork/alg-ex1-a} &
+        %\hspace{-4mm}
+        \includegraphics[height=48mm]{./artwork/alg-ex1-b} \\
+        %\hspace{-5mm}
+        (a) $R_1$, $R_2$, $H$, and $V$ &
+        %\hspace{-4mm}
+        (b) $R_1'$ and $R_2'$ \\[2mm]
+        %\hspace{-4mm}
+        \includegraphics[height=48mm]{./artwork/alg-ex1-c} &
+        %\hspace{-4mm}
+        \includegraphics[height=48mm]{./artwork/alg-ex1-d} \\
+        (c) $R_1'$, $R_2'$, $H$, and $V$ &
+        (d) $R_1^*$, $R_2^*$ ($= \emptyset$), and $H$
     \end{tabular}
@@
 \end{figure*}
-
-\section{H-V $k$-SJ: Result Tuples of Type 1} \label{sec:hv:type1}
-
-As before, let $R_1, ..., R_{k-2}, H$, and $V$ be the input sets of the H-V $k$-SJ problem.
-Denote by $\J_1$ the set of type-1 result tuples defined in Section~\ref{sec:hv}.
-In this section, we aim to compute a set $\J^*$ satisfying
-\myeqn{
-    \J_1
-    \subseteq
-    \J^*
-    \subseteq
-    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:J*}
-}
-where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ.
-Remember that the output size $\out$ is defined as $|\J(R_1, ..., R_{k-2}, H, V)|$.
-From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.
-
-\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
-The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
-\qed
-
-\extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r \in R_i$, we compute four segments:
-\myitems{
-    \item $h_\bot$ (resp., $h_\top$): the lowest (resp., highest) segment in $H$ that crosses $r$;
-
-    \item $v_\vdash$ (resp., $v_\dashv$): the leftmost (resp., rightmost) segment in $V$ that crosses $r$.
-}
-Define $r' = [x_\vdash, x_\dashv] \times [y_\bot, y_\top]$, where $x_\vdash$ (resp., $x_\dashv$) is the x-coordinate of $v_\vdash$ (resp., $v_\dashv$), and $y_\bot$ (resp., $y_\top$) is the y-coordinate of $h_\bot$ (resp., $h_\top$).
-We say that $r'$ is the {\em trimmed rectangle} of $r$, and conversely, $r$ is the {\em full rectangle} of $r'$.
-Note that $r'$ exists if and only if $r$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$.
-
-%Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$.
-
-\vgap
-
-Construct
-\myeqn{
-    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \nn
-}
-Computing the ``segment $h_\bot$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input).
-By symmetry, so is the computation of $h_\top, v_\vdash$, and $v_\dashv$ segments for each $r \in R_i$.
-It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(kn \log n)$ total time.
-
-\vgap
-
-We now solve a $(k-2)$-SJ problem on the input $\set{R_1', R_2', ..., R_{k-2}'}$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}).
-This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$.
-We prove in Appendix~\ref{app:hv:type1}:
-
-\begin{lemma} \label{lmm:hv:type1:recur-output}
-    $|\J(R_1', R'_2, ..., R'_{k-2})| \le \out$.
-\end{lemma}
-
-As a corollary of Lemma~\ref{lmm:hv:type1:recur-output}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.
-
-\extraspacing {\underline{\em Example 4.2.}} Figure~\ref{fig:hv:type1:ex}b shows the rectangles in $R'_1 = \set{\alpha'}$ and $R'_2$ $=$ $\{\beta'_1,$ $\beta'_2\}$.
-For instance, $\alpha'$, which is trimmed from rectangle $\alpha$, is decided by $h_\bot$ $=$ $h_2$, $h_\top$ $=$ $h_6$, $v_\vdash = v_3$, and $v_\dashv = v_5$.
-The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2, ..., R'_{k-2}) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
-\qed
-
-%\yf{example}
 
@@
 \begin{figure*}
-    \begin{tabular}{cccc}
-        \hspace{-5mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex2-a} &
-        \hspace{-4mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex2-b} &
-        \hspace{-4mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex2-c} &
-        \hspace{-4mm}
-        \includegraphics[height=45mm]{./artwork/alg-ex2-d} \\
-        \hspace{-5mm} (a) $R_1$, $R_2$, $H$, and $V$ &
-        \hspace{-4mm} (b) $R_1$, $H'$, and $V$ &
-        \hspace{-4mm} (c) $R_2$, $H'$, and $V$ &
-        \hspace{-4mm} (d) $R_2$ and $H^*$
+    \begin{tabular}{cc}
+        %\hspace{-5mm}
+        \includegraphics[height=48mm]{./artwork/alg-ex2-a} &
+        %\hspace{-4mm}
+        \includegraphics[height=48mm]{./artwork/alg-ex2-b} \\
+        %\hspace{-5mm}
+        (a) $R_1$, $R_2$, $H$, and $V$ &
+        %\hspace{-4mm}
+        (b) $R_1$, $H'$, and $V$ \\ [2mm]
+        \includegraphics[height=48mm]{./artwork/alg-ex2-c} &
+        \includegraphics[height=48mm]{./artwork/alg-ex2-d} \\
+        %\hspace{-4mm}
+        (c) $R_2$, $H'$, and $V$ &
+        %\hspace{-4mm}
+        (d) $R_2$ and $H^*$
     \end{tabular}

### 2024-06-03 msj.tex
yufei

@@
 \begin{abstract}
-    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets containing axis-parallel rectangles in a 2D space, denoted as $R_1, R_2, ..., R_k$.
+    In a {\em spatial join}, we are given a constant number $k \geq 2$ of sets --- denoted as $R_1, R_2, ..., R_k$ --- containing axis-parallel rectangles in a 2D space.
 The objective is to report all $k$-tuples $(r_1, r_2, ..., r_k) \in R_1 \times R_2 \times ...
@@
 The runtime  is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
-Our result significantly improves the state of the art, which is an algorithm with running time $O(n \log^{2k} n + \out)$.
+Previously, the state of the art was an algorithm with running time $O(n \log^{2k} n + \out)$.
 \end{abstract}
@@
 \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$.
-Unless otherwise specified, every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$.
+Unless otherwise stated, every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$.
 All logarithms have base 2 by default.
@@
 First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with a ``non-empty intersection'' predicate on the attributes involved.
-The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}.
+The $k$-SJ problem can be converted to a join under the framework of \cite{kcko22} as defined next.
 For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$.
@@
 
-It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based algorithm  solving the $k$-SJ problem (see Appendix~\ref{app:comp-class} for a formal definition of such algorithms), even for $k = 2$.
+It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based $k$-SJ algorithms (see Appendix~\ref{app:comp-class} for a formal definition of such algorithms), even for $k = 2$.
 This can be established via a reduction from the {\em element distinctness} problem; see \cite{dl79}.
@@
 The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
-\qed
+\done
 
@@
 The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2, ..., R'_{k-2}) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
-\qed
+\done
 
@@
 Accordingly, $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ and $\dcross_V(\bm{t}_2) = \set{v_3, v_5}$.
-\qed
+\done
 
@@
 All these 7 tuples belong to $\J(R_1, R_2, H, V)$ and include the 4 tuples in $\J_1$ (see Example 4.1).
-\qed
+\done
 
@@
 Rectangles $\beta'_1$ and $\beta'_2$ do not have top-sliced rectangles.
-\qed
+\done
 
@@
 Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list, as is consistent with Lemma~\ref{lmm:hv:type1:cross-r*}.
-\qed
+\done
 
@@
 The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$.
-\qed
+\done
 
@@
 It is clear from Figure~\ref{fig:hv:type2:ex}b that $\J(R_1, H', V)$ has 3 tuples: $\bm{t}_1 = (\alpha, h_2', v_1), \bm{t}_2 = (\alpha, h_3', v_1)$, and $\bm{t}_3 = (\alpha, h_3', v_2)$.
-\qed
+\done
 
@@
 $\contained_{R_{k-2}}(\bm{t}_1) = \contained_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contained_{R_2}(\bm{t}_3) = \set{\beta_3}$.
-\qed
+\done
 
@@
 Finally, the processing of $\bm{t}_3 = (\alpha, h_3', v_2)$ adds $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
-\qed
+\done
 
@@
 The minimal segment $h_2^*$ of $h_2'$ is also shown in Figure~\ref{fig:hv:type2:ex}(c).
-\qed
+\done
 
@@
 For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contained_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contained_{R_2}(h_2^*)$, as is consistent with the lemma.
-\qed
+\done
 
@@
 
-\def\vgap{\vspace{0.5mm}}
-\def\extraspacing{\vspace{1.5mm} \noindent}
+% \def\vgap{\vspace{0.5mm}}
+% \def\extraspacing{\vspace{1.5mm} \noindent}
 

### 2024-06-03 msj.tex
yufei

@@
 
+\subsection{Our Results} \label{sec:intro:ours}
+
+In this paper, we solve the $k$-SJ problem with a comparison-based algorithm that runs in $O(n \log n + \out)$ time regardless of the constant $k$.
+The time complexity is asymptotically optimal in the class of comparison-based algorithms.
+
+\vgap
+
+Our primary technical contribution is the revelation of a new property on the problem's mathematical structure.
+Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem.
+Define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$.
+We will establish:
+
+\begin{theorem} \label{thm:main-recur}
+    Equipped with the algorithm $\A$ as described above, the $k$-SJ problem with $k \ge 3$ can be solved in time
+    \myeqn{
+        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big)
+        %\nn
+        \label{eqn:main:reccurrence}
+    }
+    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, the obtained $k$-SJ algorithm is also comparison-based.
+\end{theorem}
+
+The theorem implies a recursive nature of $k$-SJ.
+Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + k \cdot \out)$.
+For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.
+Combining this with \eqref{eqn:main:reccurrence} gives a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ.
+Solving the recurrence yields:
+
+\begin{theorem} \label{thm:main-alg}
+    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in
+    \myeqn{
+        O( c^k \cdot (k!)^3 \cdot (n \log n + k \cdot \out))
+        \nn
+    }
+    time, where $c > 1$ is a positive constant.
+    %and $O(k \cdot n + \out)$ space,
+\end{theorem}
+
 \begin{table}
@@
 
-\subsection{Our Results} \label{sec:intro:ours}
-
-In this paper, we solve the $k$-SJ problem with a comparison-based algorithm that runs in $O(n \log n + \out)$ time regardless of the constant $k$.
-The time complexity is asymptotically optimal in the class of comparison-based algorithms.
-
-\vgap
-
-In terms of techniques, our primary contribution is the revelation of a new property on the problem's mathematical structure.
-Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem.
-Define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$.
-Our core technical contribution is to establish:
-
-\begin{theorem} \label{thm:main-recur}
-    Equipped with the algorithm $\A$ as described above, the $k$-SJ problem with $k \ge 3$ can be solved in time
-    \myeqn{
-        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big)
-        %\nn
-        \label{eqn:main:reccurrence}
-    }
-    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the $k$-SJ algorithm.
-\end{theorem}
-
-The theorem implies a recursive nature of $k$-SJ.
-Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$.
-For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.
-Combining this with \eqref{eqn:main:reccurrence} gives a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ.
-Solving the recurrence yields:
-
-\begin{theorem} \label{thm:main-alg}
-    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k \cdot (k!)^3 \cdot (n \log n + k \cdot \out))$ time, where $c > 1$ is a positive constant.
-    %and $O(k \cdot n + \out)$ space,
-\end{theorem}
-
 When $k = O(1)$, the time complexity becomes $O(n \log n$ $+ \out)$, as promised; the space consumption of our algorithm is $O(n + \out)$.
@@
 Interestingly, the answer is ``yes'' as far as $k \ge 3$ is concerned, subject to the absence of breakthroughs on a classical problem in graph theory.
-Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O((n + \out) \cdot \polylog n)$ time, we would obtain an algorithm that detects the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a truly remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
+Specifically, if the 3D version of the 3-SJ problem (which we will formally define in Appendix~\ref{app:lb-cond}) could be solved in $O((n + \out) \cdot \polylog n)$ time, we would be able to detect the presence of a triangle (i.e., 3-clique) in a graph of $m$ edges in $O(m \polylog m)$ time, which would make a remarkable breakthrough because the state of the art needs $O(m^{1.41})$ time \cite{ayz97}.
 This reduction can be inferred from an argument in \cite{kcko22} used to prove a more generic result.
 We simplify the argument for 3D 3-SJ and present the full reduction in Appendix~\ref{app:lb-cond}.
+
 \section{Preliminaries in Geometry} \label{sec:bricks}

### 2024-06-03 msj.tex
yufei

@@
     \label{tab:results-com}
-    \figcapdown \vspace{-5mm}
+    \figcapdown \vspace{-3mm}
 \end{table}
@@
 
-This section will first define some notions frequently used in our presentation and then introduce several  geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
+This section will first introduce some definitions and notations to be frequently used in our presentation and then formulate several computational geometry problems, whose solutions will serve as building bricks for our $k$-SJ algorithm.
 

### 2024-06-03 msj.tex
yufei

@@
 
-%Similarly, given a vertical segment $v$, we call a rectangle $r$ a {\em bottom-end covering rectangle} of $v$, if $r$ contains the bottom endpoint of $v$. \yf{need this?}
-
-Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
+Given a horizontal segment $h = [x_1, x_2] \times y$, we say that a rectangle $r$ is a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
 A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s \cap r \ne \emptyset$ but $r$ covers neither of the two endpoints of $s$.
@@
 \myeqn{
-    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}
+    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \};
     \label{eqn:cross}
 }
-namely, the set of segments in $S$ crossing $r$.
+namely, $\cross_S(r)$ is the set of segments in $S$ crossing $r$.
 Let $R$ be a set of rectangles. Given a horizontal segment, we define 
 \myeqn{
-    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}
+    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \};
     \label{eqn:contained}
 }
-namely, the set of rectangles in $R$ containing $h$.
+namely, $\contained_R(h)$ is the set of rectangles in $R$ containing $h$.
 
@@
     \figcapup
-    \caption{Consider a 4-tuple $\bm{t} = \set{r_1, r_2, r_3, r_4}$. $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
+    \caption{For 4-tuple $\bm{t} = \set{r_1, r_2, r_3, r_4}$, $B_\bm{t}$ is the rectangle in gray, $\gleft(\bm{t}) = r_3$ and $\gbot(\bm{t}) = r_2$.}
     \label{fig:guard}
@@
 Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$.
-Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
+Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$ and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle.
 We define
 \myeqn{
-    B_\bm{t} &=& \cap_{i=1}^t r_i \label{eqn:B_t}
+    B_\bm{t} &=& \bigcap_{i=1}^t r_i; \label{eqn:B_t}
 }

### 2024-06-03 msj.tex
yufei

@@
 Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$.
-This can be achieved in $O(n \log n)$ time where $n = |P| + |R|$ as shown
+The problem can be solved in $O(n \log n)$ time where $n = |P| + |R|$ as shown
 in Appendix~\ref{app:bricks}.
@@
 In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_1, \emptyset), (p_2: r_1, r_2), (p_3: r_2)}$.
-As shown in Appendix~\ref{app:bricks}, this can be achieved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
+As shown in Appendix~\ref{app:bricks}, The problem can be solved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
 
@@
 Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-This problem can be solved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.
+The problem can be solved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.
 
@@
 Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
-This problem can be solved in in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
+The problem can be solved in in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
 
@@
 Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.
-This can be achieved in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
+The problem can be solved in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.
 

### 2024-06-03 msj.tex
yufei

@@
 If the point $p$ exists (i.e., $h$ has at least one left-end covering rectangle in $R$), we should output a tuple $(h, p)$; otherwise, output nothing for $h$.
-Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
+Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2, h_3}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.
 The problem can be solved in in $O(n \log n)$ time where $n = |H| + |R|$, as shown in Appendix~\ref{app:bricks}.

### 2024-06-03 msj.tex
yufei

@@
 \extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
-In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see its definition in \eqref{eqn:cross}).
+In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$; see \eqref{eqn:cross} for the definition of $\cross_H(r)$.
 If no segment in $H$ crosses $r$, output nothing for $r$.
@@
 
+\section{The Core: H-V Multiway Spatial Joins} \label{sec:hv}
+
+Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ...,$ $R_k$.
+We now formulate a special version of $k$-SJ, named the {\em H-V $k$-SJ problem}.
+The special nature is reflected in the introduction of three constraints: (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
+For better clarity, we will represent the input sets as $R_1, R_2, ..., R_{k-2}, H$ ($=R_{k-1}$), and $V$ ($=R_k$).
+The goal is to output the join result $\J(R_1, ..., R_{k-2},$ $H, V)$, including every $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
+\times R_{k-2} \times H \times V$ such that $h \cap v \cap \bigcap_{i=1}^{k-2} r_i$ is not empty.
+
+\vgap
+
+Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
+To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time.
+We assume the existence of an algorithm $\A$ that can settle $\kappa$-SJ for all $\kappa \in [1, k-1]$.
+Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
+We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \ge 1$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
+
+\vgap
+
+We will establish:
+
+\begin{lemma} \label{lmm:hv}
+    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
+    \myeqn{
+        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
+    }
+    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the H-V $k$-SJ algorithm.
+\end{lemma}
+
+The part of the paper from this point till the end of Section~\ref{sec:hv:type2} will be devoted to proving the above lemma.
+This is the most challenging step in solving the general $k$-SJ problem optimally, as will be discussed in Section~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
+
+\vgap
+
+%The condition is equivalent to saying that the intersection point of $h$ and $v$ falls in every $r_i$ of $i \in [k-2]$.
+
+Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ in the join result $\J(R_1, ...,$ $R_{k-2}, H, V)$. We classify the tuple into one of the two types below:
+
+\myitems{
+    \item {{\bf Type 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$ and, at the same time, $v$ crosses all of $r_1, ..., r_{k-2}$;
+
+    \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
+    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
+}
+Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.
+In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.
+
+\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently.
+They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
+Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
+Another structure with the same guarantees can be found in \cite{oa19}.
+We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.
+Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.
+
 \begin{figure}
@@
 \end{figure}
-
-\section{The Core: H-V Multiway Spatial Joins} \label{sec:hv}
-
-Recall that the input of $k$-SJ comprises $k$ sets of rectangles: $R_1, R_2, ...,$ $R_k$.
-We now formulate a special version of $k$-SJ, named the {\em H-V $k$-SJ problem}.
-The special nature is reflected in the introduction of three constraints: (i) $k \ge 3$, (ii) $R_{k-1}$ should be a set of horizontal segments, and (iii) $R_k$ should be a set of vertical segments.
-For better clarity, we will represent the input sets as $R_1, R_2, ..., R_{k-2}, H$ ($=R_{k-1}$), and $V$ ($=R_k$).
-The goal is to output the join result $\J(R_1, ..., R_{k-2},$ $H, V)$, including every $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in R_1 \times ...
-\times R_{k-2} \times H \times V$ such that $h \cap v \cap \bigcap_{i=1}^{k-2} r_i$ is not empty.
-
-\vgap
-
-Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
-To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time.
-We assume the existence of an algorithm $\A$ that can settle $\kappa$-SJ for all $\kappa \in [1, k-1]$.
-Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$.
-We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \ge 1$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
-
-\vgap
-
-We will establish:
-
-\begin{lemma} \label{lmm:hv}
-    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
-    \myeqn{
-        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
-    }
-    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the H-V $k$-SJ algorithm.
-\end{lemma}
-
-The part of the paper from this point till the end of Section~\ref{sec:hv:type2} will be devoted to proving the above lemma.
-This is the most challenging step in solving the general $k$-SJ problem optimally, as will be discussed in Section~\ref{sec:ksj}, where we will prove Theorems~\ref{thm:main-recur} and \ref{thm:main-alg} based on Lemma~\ref{lmm:hv}.
-
-\vgap
-
-%The condition is equivalent to saying that the intersection point of $h$ and $v$ falls in every $r_i$ of $i \in [k-2]$.
-
-Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v)$ in the join result $\J(R_1, ...,$ $R_{k-2}, H, V)$. We classify the tuple into one of the two types below:
-
-\myitems{
-    \item {{\bf Type 1:}} $h$ crosses all of $r_1, ..., r_{k-2}$ and, at the same time, $v$ crosses all of $r_1, ..., r_{k-2}$;
-
-    \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
-    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
-}
-Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.
-In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.
-
-\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently.
-They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
-Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
-Another structure with the same guarantees can be found in \cite{oa19}.
-We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.
-Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.
 

### 2024-06-03 msj.tex
yufei

@@
 \begin{lemma} \label{lmm:hv}
-    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
+    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in
     \myeqn{
@@
     }
-    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the H-V $k$-SJ algorithm.
+    time where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, the H-V $k$-SJ algorithm obtained is also comparison-based.
 \end{lemma}
@@
     \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
-    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
+    at least a rectangle $r_i$ (for some $i \in [k-2]$) covers an endpoint of either $h$ or $v$ or both.
 }
@@
 Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
-Another structure with the same guarantees can be found in \cite{oa19}.
+Oh and Ahn \cite{oa19} developed a structure for a more general version of the problem; in the specific scenario of \cite{rdr+11}, the structure of \cite{oa19} matches the performance guarantees of \cite{rdr+11}.
 We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.

### 2024-06-03 msj.tex
yufei

@@
 \extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently.
-They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
+They gave a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported.
 Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
 Oh and Ahn \cite{oa19} developed a structure for a more general version of the problem; in the specific scenario of \cite{rdr+11}, the structure of \cite{oa19} matches the performance guarantees of \cite{rdr+11}.
-We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$.
-Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.
+We are unaware of a way to extend these solutions to handle H-V $k$-SJ of $k > 3$.
+Our method for proving Lemma~\ref{lmm:hv} is based on drastically different ideas even for $k = 3$.
 

### 2024-06-03 msj.tex
yufei

@@
 Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time.
-Oh and Ahn \cite{oa19} developed a structure for a more general version of the problem; in the specific scenario of \cite{rdr+11}, the structure of \cite{oa19} matches the performance guarantees of \cite{rdr+11}.
+Oh and Ahn \cite{oa19} developed a structure for solving a problem more general than that of \cite{rdr+11}; however, in the specific scenario of \cite{rdr+11}, the structure of \cite{oa19} offers the same guarantees as \cite{rdr+11}.
 We are unaware of a way to extend these solutions to handle H-V $k$-SJ of $k > 3$.

### 2024-06-03 msj.tex
yufei

@@
 
-\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2}$, $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
+\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$ (the solid rectangle), $R_2 = \set{\beta_2, \beta_2}$ (the dashed rectangles), $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
 The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
@@
 \myeqn{
-    \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers} \nn \\[-0.5mm]
-    && \text{the left endpoint of $h$}\}. \label{eqn:hv:type2:J2}
+    \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers the left endpoint of $h$}\}. \label{eqn:hv:type2:J2}
 }

### 2024-06-03 msj.tex
yufei

@@
 \myeqn{
-    \contained_{R_{k-2}}(\bm{t}) &=& \{r \in R_{k-2} \mid \text{ the effective horizontal } \nn \\[-1mm]
-    && \text{segment of $\bm{t}$ is contained in $r$}   \}
+    \contained_{R_{k-2}}(\bm{t}) = \{r \in R_{k-2} \mid \text{ the effective horizontal segment of $\bm{t}$ is contained in $r$}   \}
     \label{eqn:type2:contained-t}
@@
 
-\extraspacing {\underline{\em Example 5.4.}} For $\bm{t}_1 = (\alpha, h_2', v_1)$, first find the full segment of $h_2'$, which is $h_2$.
+\extraspacing {\underline{\em Example 5.4.}} For $\bm{t}_1 = (\alpha, h_2', v_1)$, the full segment of $h_2'$ is $h_2$.
 As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 of the algorithm adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$.

### 2024-06-03 msj.tex
yufei

@@
 \def\vgap{\vspace{1mm}}
-\def\extraspacing{\vspace{3mm} \noindent}
+\def\extraspacing{\vspace{2mm} \noindent}
 \def\figcapup{\vspace{-2mm}}
@@
 \myeqn{
-    && \sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)|  \nn \\ 
-    &\le& \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\contained_{R_{k-2}}(\bm{t})| 
+    \sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)|
+    \le
+    \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\contained_{R_{k-2}}(\bm{t})|
     \leq \out \nn
@@
     &\le& c \cdot k^3 \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big) \nn \\
-    &=& c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out) + \nn \\
-    && c \cdot k^3 \cdot (n\log n + k\cdot \out ) \nn\\
-    &<& c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) +  \nn\\
-    && (c+c^2)\cdot k^3(k-1)^3 \cdot (n\log n + k \cdot\out) \nn \\
-    &\leq& c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out) + \nn\\
-    && (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3 \cdot (n\log n + k \cdot\out) \nn\\
+    &=& c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out) + c \cdot k^3 \cdot (n\log n + k\cdot \out ) \nn\\
+    &<& c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) +  (c+c^2)\cdot k^3(k-1)^3 \cdot (n\log n + k \cdot\out) \nn \\
+    &\leq& c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out) +
+    (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3 \cdot (n\log n + k \cdot\out) \nn\\
     &\leq& ... \nn \\
-    &\leq& c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \nn \\
-    && \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3 \cdot (n\log n + k \cdot\out) \nn \\
+    &\leq& c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3 \cdot (n\log n + k \cdot\out) \nn \\
     &\leq & 2c^{k-1} \cdot (k!)^3 \cdot O(n\log n + k \cdot\out) \nn

### 2024-06-03 msj.tex
yufei

@@
 \def\figcapup{\vspace{-2mm}}
-\def\figcapdown{\vspace{-3mm}}
+\def\figcapdown{\vspace{-4mm}}
 
@@
 
-        \item Consider any a $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v) \in \J_1$.
+        \item Consider any $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v) \in \J_1$.
 Let $r'_i$ ($i \in [k-2])$ be the trimmed rectangle of $r_i$, and set $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2})$.
-Then, we must have (i) $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$, and (ii) $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+Then, we must have
+        \myitems{
+            \item $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$;
+            \item $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+        }
 
@@
 Equipped with Lemma~\ref{lmm:hv:type1:properties}, we generate our target $\J^*$ as follows:
+
+\vgap
 
@@
 
+\vgap
+
 By statements (1) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
@@
 }
-We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
+We set $\maxtop(r')$ to $-\infty$ if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$.
 When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
@@
 }
-In Appendix~\ref{app:hv:type1}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.
+In Appendix~\ref{app:hv:type1}, we show how to compute $R^*_1, ..., R^*_{k-2}$ altogether in $O(n + k \cdot \out)$ total time.
 
@@
 
+\vgap
+
 \mytab{
@@
 }
+
+\vgap
 

### 2024-06-03 msj.tex
yufei

@@
 
-Our focus lies specifically on the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
+Our interest lies specifically in the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$.
 The following lemma, proven in Appendix~\ref{app:hv:type1}, presents some useful properties of these sets.

### 2024-06-03 msj.tex

@@
 
-\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of y-coordinate.
-Recall that
-$\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$.
-Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are prefixes of the sorted list, as is consistent with Lemma~\ref{lmm:hv:type1:cross-r*}.
+\extraspacing {\underline{\em Example 4.6.}} It is clear from Figure~\ref{fig:hv:type1:ex}d that $\cross_H(\alpha^*)$ contains $h_2, h_4$, and $h_5$, sorted in ascending order of their y-coordinates.
+Recall that $\gbot(\bm{t}_1) = \gbot(\bm{t}_2) = \alpha'$.
+Both $\dcross_H(\bm{t}_1) = \set{h_2}$ and $\dcross_H(\bm{t}_2) = \set{h_2, h_4, h_5}$ are indeed prefixes of the sorted $\cross_H(\alpha^*)$, as stated in Lemma~\ref{lmm:hv:type1:cross-r*}.
 \done

### 2024-06-03 msj.tex
yufei

@@
 
-\extraspacing {\underline{\em Example 5.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_2, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$.
+\extraspacing {\underline{\em Example 5.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type2:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$, $R_2 = \set{\beta_1, \beta_2, \beta_3}$, $H$ $=$ $\set{h_1,$ $h_2,$ $h_3}$, and $V = \set{v_1, v_2}$.
 The set $\J_2$ contains the following tuples: $(\alpha, \beta_1, h_2, v_1)$, $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, $(\alpha, \beta_3, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_2)$.

### 2024-06-03 msj.tex
yufei

@@
 \myeqn{
-    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \};
+    \contained_R(h) = \{r \in R \mid \text{$r$ contains $h$} \};
     \label{eqn:contained}
@@
 \extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$.
-Note that $B_\bm{t}$ --- defined in \eqref{eqn:B_t} --- is a point (due to $h'$ and $v$).
+Note that $B_\bm{t}$ --- defined in \eqref{eqn:B_t} --- is the point $h' \cap v$ (the intersection of $h'$ and $v$).
 Suppose $h' = [x_1, x_2] \times y$ and $B_\bm{t} = (x, y)$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $[x_1, x] \times y$.
@@
 \myeqn{
-    \contained_{R_{k-2}}(\bm{t}) = \{r \in R_{k-2} \mid \text{ the effective horizontal segment of $\bm{t}$ is contained in $r$}   \}
+    \contained_{R_{k-2}}(\bm{t}) = \{r \in R_{k-2} \mid \text{$r$ contains the effective horizontal segment of $\bm{t}$}   \}
     \label{eqn:type2:contained-t}
@@
 
-We prove the next lemma in Appendix~\ref{app:hv:type2} (to avoid confusion, the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).
+We prove the next lemma in Appendix~\ref{app:hv:type2} (the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).
 
@@
 }
-We set $\minleft(h')$ to $\infty$, if no $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ has $h'$ in its field $\bm{t}[k-2]$.
+We set $\minleft(h')$ to $\infty$ if no $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ has $h'$ in its field $\bm{t}[k-2]$.
 When $\minleft(h') \ne \infty$, assuming $h' = [x_1, x_2] \times y$, we introduce a horizontal segment
@@
 Both $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$ have $h_3'$ as the horizontal segment.
-Therefore, $\minleft(h_3')$ equals the smaller between the x-coordinate of $v_1$ and that of $v_2$.
+Therefore, $\minleft(h_3')$ equals 5, which is the smaller between the x-coordinate of $v_1$ and that of $v_2$.
 The minimal segment $h_3^*$ of $h_3'$ is shown in Figure~\ref{fig:hv:type2:ex}(c).
@@
 
-%Computing $H^*$ from $H'$ and $V$ is an instance of Problem $\mathscr{B}$, and can be achieved in $O(n \log n)$ as per the discussion from Section~\ref{sec:bricks}.
-
-\vgap
-
-We will utilize the sets $\contained_{R_{k-2}}(h^*)$ of the segments $h^*$ in $H^*$, where $\contained_{R_{k-2}}(h^*)$ --- defined in \eqref{eqn:contained} --- is the set of rectangles in $R_{k-2}$ containing $h^*$.
+\vgap
+
+We are interested in the sets $\contained_{R_{k-2}}(h^*)$ of the segments $h^*$ in $H^*$, where $\contained_{R_{k-2}}(h^*)$ --- defined in \eqref{eqn:contained} --- is the set of rectangles in $R_{k-2}$ containing $h^*$.
 These sets have some useful properties:

### 2024-06-03 msj.tex
yufei

@@
 Regarding $h_2^*$, it is the minimal segment of $h_2'$, and $\contained_{R_2}(h_2^*)$ contains only  $\beta_1$.
-For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contained_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contained_{R_2}(h_2^*)$, as is consistent with the lemma.
+For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contained_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contained_{R_2}(h_2^*)$, as is also consistent with the lemma.
 \done

### 2024-06-03 msj.tex
yufei

@@
 For every $\bm{t} = (r_1, r_2, ..., r_k) \in \J^\cat_k$, the top-left corner of $r_k$ must be covered by all of $r_1, ..., r_{k-1}$.
-This observation motivates us to find $\J^\cat_i$ as follows.
+This observation motivates us to find $\J^\cat_k$ as follows.
 First, collect the set $P$ of top-left corners of all the rectangles in $R_k$.
@@
 we add $(r_1, ..., r_{k-1}, r)$ to $\J^\cat_k$.
-Performing the above for all $p \in P'$ generates the whole $\J^\cat_k$ in $O(1 + |\J^\cat_k|)$ time.
+Performing the above for all $p \in P'$ generates the whole $\J^\cat_k$ in $O(1 + k \cdot|\J^\cat_k|)$ time.
 The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for each $j \in [k-1]$.

### 2024-06-03 msj.tex
yufei

@@
 namely, $\cross_S(r)$ is the set of segments in $S$ crossing $r$.
-Let $R$ be a set of rectangles. Given a horizontal segment, we define 
+Let $R$ be a set of rectangles. Given a horizontal segment $h$, we define 
 \myeqn{
@@
 Thus, by Lemma~\ref{lmm:hv}, the H-V $k$-SJ can be solved in $O(k) \cdot (F_{k-1}(n, |\J^\catt_{i,j}|) + n \log n + k \cdot |\J^\catt_{i,j}|)$ time.
-Converting the output into $\J^\catt_{i,j}$ takes another $O(|\J^\catt_{i,j}|)$ time.
+Converting the output into $\J^\catt_{i,j}$ takes another $O(k \cdot |\J^\catt_{i,j}|)$ time.
 Applying $|\J^\catt_{i,j}| \le \out$, we know that $\J^\catt_{i,j}$ can be produced in $O(k) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$ time
@@
 
-Fix an integer $k \ge 3$, as well as $k$ integers $n_1, n_2, ..., n_k$ at least 1.
+Fix an integer $k \ge 3$ and $k$ integers $n_1, n_2, ..., n_k$ at least 1.
 Define $n = \sum_{i=1}^k n_i$.
@@
 }
-Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
+Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side' = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
 \myitems{

### 2024-06-03 msj.tex
yufei

@@
 \extraspacing {\bf Terminology.} Each point $(x, y)$ is said to {\em define} a y-coordinate $y$, a horizontal segment $[x_1, x_2] \times y$ is said to {\em define} a y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ is said to {\em define} two y-coordinates $y_1$ and $y_2$.
-These definitions permit us to convenient specify a set of y-coordinates using expressions like ``the set of four y-coordinates defined by point $p$, horizontal segment $h$, and rectangle $r$''.
+These definitions permit us to conveniently specify a set of y-coordinates using expressions like ``the set of four y-coordinates defined by point $p$, horizontal segment $h$, and rectangle $r$''.
 

### 2024-06-03 msj.tex
yufei

@@
 It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log |S|)$ time.
-The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ time.
+The structure supports insertions and deletions on $S$ in $O(\log |S|)$ time.
 
@@
 We can store $S$ in a structure of \cite{aak+12} using $O(|S|)$ space that can answer such a query in $O(\log |S|)$ time.
-The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ amortized time.
+The structure supports insertions and deletions on $S$ in $O(\log |S|)$ amortized time.
 
@@
 The {\em priority search tree} (PST) \cite{m85} stores a set $P$ of points using $O(|P|)$ space such that, given a 3-sided rectangle $q = [x_1, x_2] \times [y, \infty)$, the points of $S$ covered by $q$ can be found in $O(\log |P| + K)$ time, where $K$ is the number of points reported.
-The structure supports an insertion or deletion on $S$ in $O(\log |S|)$ time.
+The structure supports insertions and deletions on $S$ in $O(\log |S|)$ time.
 

### 2024-06-03 msj.tex
yufei

@@
 Let $S$ be a set of intervals in $\real$.
-Given an interval $q = [z_1, z_2]$, a {\em containment query} reports all the intervals of $S$ that are contained in $q$.
+Given an interval $q = [z_1, z_2]$, a {\em containment query} reports all the intervals of $S$ that are contained by $q$.
 We can store $S$ in a PST of $O(|S|)$ space that solves such a query in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
-To see why, observe that an interval $[z_1, z_2]$ is contained by another $[x, y]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $[z_1, \infty) \times (-\infty, z_2]$.
+To see why, observe that an interval $[z_1, z_2]$ is contained by another $[x, y]$ if and only if the point $(x, y)$ falls in the 3-sided rectangle $[z_1, \infty) \times (-\infty, z_2]$.
 Thus, we create from $S$ a point set $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.
@@
 Again, we can store $S$ in a PST using $O(|S|)$ space such that any such query can be answered in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
-In general, an interval $[x, y]$ contains another $[z_1, z_2]$ if and only if the point $(x, y)$ is covered by the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.
+In general, an interval $[x, y]$ contains another $[z_1, z_2]$ if and only if the point $(x, y)$ falls in the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.
 Thus, we create from $S$ a point set $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.

### 2024-06-03 msj.tex
yufei

@@
 We can store $S$ in a PST of $O(|S|)$ space that solves such a query in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
-To see why, observe that an interval $[z_1, z_2]$ is contained by another $[x, y]$ if and only if the point $(x, y)$ falls in the 3-sided rectangle $[z_1, \infty) \times (-\infty, z_2]$.
+To see why, observe that an interval $[x, y]$ is contained by another $[z_1, z_2]$ if and only if the point $(x, y)$ falls in the 3-sided rectangle $[z_1, \infty) \times (-\infty, z_2]$.
 Thus, we create from $S$ a point set $P = \set{(x, y) \mid [x, y] \in S}$ and store $P$ in a PST.

### 2024-06-03 msj.tex
yufei

@@
 Another closely related query is the {\em reverse-containment query}, which, given an interval $q = [z_1, z_2]$, finds all the intervals of $S$ that contain $q$ (rather than ``being contained by $q$'').
-Again, we can store $S$ in a PST using $O(|S|)$ space such that any such query can be answered in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
+Again, we can store $S$ in a PST of $O(|S|)$ space that answers such a query in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported.
 In general, an interval $[x, y]$ contains another $[z_1, z_2]$ if and only if the point $(x, y)$ falls in the 3-sided rectangle $(-\infty, z_1] \times [z_2, \infty)$.

### 2024-06-03 msj.tex
yufei

@@
 
-Next, we aim to produce, for each pair $(h, r) \in H \times r$ such that $r$ crosses $h$, a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$.
+Next, we aim to produce, for each pair $(h, r) \in H \times r$ such that $h$ crosses $r$, a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$.
 These pairs may be output in an arbitrary order.
@@
 All these insertions and deletions take $O(n \log n)$ time in total.
-Each segment $h$ of $H$ performs a containment query on $\T$, which has a cost of $O(\log n + K_h)$.
+Each segment $h$ of $H$ performs a reverse-containment query on $\T$, which has a cost of $O(\log n + K_h)$.
 All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.

### 2024-06-03 msj.tex
yufei

@@
 Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ amortized time.
-When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$.
+When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine the successor $x'$ of $x_1$ in $S_\ell$.
 If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$; otherwise, output nothing for $h$.
@@
 At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
-Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a structure $\T$ of \cite{aak+12}.
+Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a stabbing-max structure $\T$ of \cite{aak+12}.
 Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ with {\em weight} $x_2$ into (resp., from) $\T$, which can be done in $O(\log n)$ time.
@@
 We call the rectangle $r$ the {\em host} of $[x_1, x_2]$ and store its ID together with $[x_1, x_2]$ in $\T$.
-When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a reverse-containment query on $\T$ to find all the intervals in $S_\ell$ that contain $[z_1, z_2]$; the query cost is $O(\log n + K_h)$, where $K_h$ is the number of intervals reported.
+When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $H$, perform a reverse-containment query on $\T$ to find all the intervals in $S_\ell$ that contain $[z_1, z_2]$; the query cost is $O(\log n + K_h)$, where $K_h$ is the number of intervals reported.
 For each retrieved interval $[x_1, x_2]$, we also obtain the ID $\lambda$ of its host rectangle $r$, and output the pair $(h, \lambda)$.

### 2024-06-03 msj.tex
yufei

@@
 Recall that the x-range of a rectangle in $R$ can be retrieved by at most one containment query.
-Hence, $\sum_h K_h \le n$ and the runtime of our algorithm is $O(n \log n)$.
+Hence, $\sum_h K_h \le |R| \le n$ and the runtime of our algorithm is $O(n \log n)$.
 
@@
 Let $L$ be the list of $(h, \lambda)$ pairs produced (the size of $L$ is $\out$).
-We now proceed to sort $L$ in ascending order of the $\lambda$-field (which is a rectangle ID).
-Because the IDs are integers created by the algorithm, we are permitted to sort $L$ using {\em counting sort} without violating the requirements of the comparison-based class (see Appendix~\ref{app:comp-class}).
+We now proceed to sort $L$ in ascending order of the $\lambda$-field (which is a rectangle ID), breaking ties arbitrarily.
+Because the IDs are integers created by the algorithm, we are permitted to sort $L$ using {\em counting sort} without violating the comparison-based requirements (see Appendix~\ref{app:comp-class}).
 The counting sort finishes in $O(|R| + |L|) = O(n + \out)$ time, recalling that all the IDs are in $[|R|]$.

### 2024-06-03 msj.tex
yufei

@@
     \item a rectangle $(-\infty, \infty) \times [u,u] \times [v,v]$ to $R_1$;
-    \item $[v,v] \times(-\infty, \infty)\times [u,u]$ to $R_2$;
-    \item $[v,v] \times [u,u] \times(-\infty, \infty) $ to $R_3$.
+    \item a rectangle $[v,v] \times(-\infty, \infty)\times [u,u]$ to $R_2$;
+    \item a rectangle $[v,v] \times [u,u] \times(-\infty, \infty) $ to $R_3$.
 \end{itemize}
@@
 Assume, w.l.o.g., that $r_1 = (-\infty, \infty) \times [u, u] \times [v, v]$.
-    Because the z-projection of $r_2$ must match that of $r_1$, we assert that $r_2$ must have the form $[w, w] \times (-\infty, \infty) \times [v, v]$ where $w > u$.
+    Because the z-projection of $r_2$ must match that of $r_1$, we assert that $r_2$ must have the form $[w, w] \times (-\infty, \infty) \times [v, v]$ for some $w > u$.
 Because the x-projection of $r_3$ must match that of $r_2$ and the y-projection of $r_3$ must match that of $r_1$, it follows that $r_3$ must have the form $[w, w] \times [u, u] \times (-\infty, \infty)$.

### 2024-06-03 msj.tex
yufei

@@
 Consider an arbitrary result tuple $(r_1, r_2, r_3) \in \J(R_1, R_2, R_3)$.
-Assume, w.l.o.g., that $r_1 = (-\infty, \infty) \times [u, u] \times [v, v]$.
-    Because the z-projection of $r_2$ must match that of $r_1$, we assert that $r_2$ must have the form $[w, w] \times (-\infty, \infty) \times [v, v]$ for some $w > u$.
+Assume, w.l.o.g., that $r_1 = (-\infty, \infty) \times [u, u] \times [v, v]$, where $u < v$.
+    Because the z-projection of $r_2$ must match that of $r_1$, we assert that $r_2$ must have the form $[w, w] \times (-\infty, \infty) \times [v, v]$ for some $w > v$.
 Because the x-projection of $r_3$ must match that of $r_2$ and the y-projection of $r_3$ must match that of $r_1$, it follows that $r_3$ must have the form $[w, w] \times [u, u] \times (-\infty, \infty)$.

### 2024-06-03 msj.tex
yufei

@@
 By definition of $\J_1$, segments $h$ and $v$ both cross each of the rectangles $r_1,..., r_{k-2}$.
-By the earlier discussion, $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
-Hence, the intersection of $h$ and $v$ must be a point covered by each of $r_1',..., r_{k-2}'$.
-This suggests that the point is in $\bigcap_{i = 1}^{k-2}r_i'$, which thus cannot be empty.
+According to our earlier discussion, $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
+Hence, the intersection of $h$ and $v$ is a point covered by each of $r_1',..., r_{k-2}'$.
+Hence, $\bigcap_{i = 1}^{k-2}r_i'$ cannot be empty.
 Therefore, $\bm{t} = (r_1', ..., r_{k-2}')$ is a result tuple in $\J(R_1',...,R_{k-2}')$.

### 2024-06-03 msj.tex
yufei

@@
 Hence, the intersection of $h$ and $v$ is a point covered by each of $r_1',..., r_{k-2}'$.
-Hence, $\bigcap_{i = 1}^{k-2}r_i'$ cannot be empty.
-Therefore, $\bm{t} = (r_1', ..., r_{k-2}')$ is a result tuple in $\J(R_1',...,R_{k-2}')$.
+Hence, $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$ cannot be empty, meaning that $\bm{t} = (r_1', ..., r_{k-2}') \in \J(R_1',...,R_{k-2}')$.
 
@@
 The first part, $h$ crossing $\gbot(\bm{t})$, is done because, as already explained, $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is one of those $k-2$ rectangles.
-To prove that $h$ crosses $B_\bm{t}$, first note that $B_\bm{t}$ (which is the intersection of $r_1', ..., r_{k-2}'$) is covered by $\gbot(\bm{t})$.
-The fact that $h$ crosses $\gbot(\bm{t})$ indicates that the y-range of $h$ contains that of $B_\bm{t}$.
+To prove that $h$ crosses $B_\bm{t}$, first note that $B_\bm{t}$ is covered by $\gbot(\bm{t})$.
+The fact of $h$ crossing $\gbot(\bm{t})$ indicates that the x-range of $h$ contains that of $B_\bm{t}$.
 Hence, either $h$ is disjoint with $B_\bm{t}$ or $h$ crosses $B_\bm{t}$.
-However, $h$ cannot be disjoint with $B_\bm{t}$ because, as shown earlier, the intersection of $h$ and $v$ falls in $B_\bm{t}$.
+However, $h$ cannot be disjoint with $B_\bm{t}$ because as explained earlier $h \cap v$ falls in $B_\bm{t}$.
 It thus follows that $h$ crosses $B_\bm{t}$.

### 2024-06-03 msj.tex
yufei

@@
 
-\extraspacing \underline{\em Proof of Statement (1).} As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
+\extraspacing \underline{\em Proof of Statement (1).} Let $\bm{t}$, $h$, and $v$ be as defined in the statement. 
+As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
 Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$.

### 2024-06-03 msj.tex
yufei

@@
 
-We now prove the first claim of statement (2), namely, $\bm{t}\in \J(R_1',...,R_{k-2}')$.
+We now prove the first claim of statement (2), i.e., $\bm{t}\in \J(R_1',...,R_{k-2}')$.
 Consider the $k$-tuple $(r_1,...,r_{k-2},h,v)\in \J_1$ given in statement (2).
 By definition of $\J_1$, segments $h$ and $v$ both cross each of the rectangles $r_1,..., r_{k-2}$.
-According to our earlier discussion, $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
-Hence, the intersection of $h$ and $v$ is a point covered by each of $r_1',..., r_{k-2}'$.
-Hence, $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$ cannot be empty, meaning that $\bm{t} = (r_1', ..., r_{k-2}') \in \J(R_1',...,R_{k-2}')$.
-
-\vgap
-
-Next, we prove the second claim, namely, $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+As explained above, $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
+Hence, $h \cap v$ is a point covered by each of $r_1',..., r_{k-2}'$.
+Therefore, $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$ cannot be empty, meaning that $\bm{t} = (r_1', ..., r_{k-2}') \in \J(R_1',...,R_{k-2}')$.
+
+\vgap
+
+Next, we prove the second claim, i.e., $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
 It suffices to show only the former due to symmetry.
 To prove $h \in \dcross_H(\bm{t})$, we need to argue that $h$ crosses $\gbot(\bm{t})$ and $B_\bm{t}$.
-The first part, $h$ crossing $\gbot(\bm{t})$, is done because, as already explained, $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is one of those $k-2$ rectangles.
+The first part, $h$ crossing $\gbot(\bm{t})$, is done because as mentioned $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is one of those rectangles.
 To prove that $h$ crosses $B_\bm{t}$, first note that $B_\bm{t}$ is covered by $\gbot(\bm{t})$.
@@
 Hence, either $h$ is disjoint with $B_\bm{t}$ or $h$ crosses $B_\bm{t}$.
-However, $h$ cannot be disjoint with $B_\bm{t}$ because as explained earlier $h \cap v$ falls in $B_\bm{t}$.
+However, $h$ cannot be disjoint with $B_\bm{t}$ because as explained before $B_\bm{t}$ contains the point $h \cap v$.
 It thus follows that $h$ crosses $B_\bm{t}$.

### 2024-06-03 msj.tex
yufei

@@
 The first part, $h$ crossing $\gbot(\bm{t})$, is done because as mentioned $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is one of those rectangles.
-To prove that $h$ crosses $B_\bm{t}$, first note that $B_\bm{t}$ is covered by $\gbot(\bm{t})$.
-The fact of $h$ crossing $\gbot(\bm{t})$ indicates that the x-range of $h$ contains that of $B_\bm{t}$.
-Hence, either $h$ is disjoint with $B_\bm{t}$ or $h$ crosses $B_\bm{t}$.
-However, $h$ cannot be disjoint with $B_\bm{t}$ because as explained before $B_\bm{t}$ contains the point $h \cap v$.
-It thus follows that $h$ crosses $B_\bm{t}$.
+To prove that $h$ crosses $B_\bm{t}$, first note that $h \cap B_\bm{t} \ne \emptyset$ because as explained before $B_\bm{t}$ contains the point $h \cap v$.
+On the other hand, as $\gbot(\bm{t})$ covers $B_\bm{t}$, the fact of $h$ crossing $\gbot(\bm{t})$ indicates that the x-range of $h$ contains that of $B_\bm{t}$.
+Combining this with $h \cap B_\bm{t} \ne \emptyset$ shows that $h$ crosses $B_\bm{t}$.
 
@@
 Set $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.
-Define $h$ as the lowest segment crossing $r$ (note that $h$ definitely exists because otherwise $r$ has no trimmed rectangle).
-We will show that $h \in \dcross_H(\bm{t})$, and hence $\dcross_H(\bm{t})$ cannot be empty.
+Define $h$ as the lowest segment crossing $r$ (note that $h$ definitely exists because otherwise $r$ has no trimmed rectangle, contradicting the definition of $r'$).
+We will show that $h \in \dcross_H(\bm{t})$ and hence $\dcross_H(\bm{t}) \ne \emptyset$.
 For this purpose, we should explain why $h$ crosses both $B_{\bm{t}}$ and $r'$.
-However, $h$ crossing $r$ implies $h$ crossing $r'$, as has already been argued in our proof of statement (2).
+However, $h$ crossing $r$ implies $h$ crossing $r'$, as we already know from the above proof of statement (2).
 It remains to show that $h$ crosses $B_\bm{t}$.
@@
 
-By the way $r'$ is defined, the bottom edge of $r'$ must be contained in $h$.
+By the definitions of $r'$ and $h$, the bottom edge of $r'$ must be contained in $h$.
 Because $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$ and hence also contained in $h$.

### 2024-06-03 msj.tex
yufei

@@
 
-\extraspacing \underline{\em Proof of Statement (3).} We first prove that $\dcross_H(\bm{t}) \neq \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
+\extraspacing \underline{\em Proof of Statement (3).} We start by proving that $\dcross_H(\bm{t}) \neq \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
 Set $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.

### 2024-06-03 msj.tex
yufei

@@
 This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}.
-By statement (1) of Lemma~\ref{lmm:hv:type2:properties}, the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whose size is $\out$.
+By statement (1) of Lemma~\ref{lmm:hv:type2:properties}, the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whereas $\out = |\J(R_1,...,R_{k-2},H,V)|$.
 This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).

### 2024-06-03 msj.tex
yufei

@@
 This section supplies the proofs postponed from Section~\ref{sec:hv:type1}.
-The order of the proofs here will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type1}.
-In particular, we will prove Lemma~\ref{lmm:hv:type1:properties} before Lemma~\ref{lmm:hv:type1:recur-output}, because the claims of the former lemma can be used to construct a simple argument for the latter.
-
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} We will prove each statement in turn.
-
-\extraspacing \underline{\em Proof of Statement (1).} Let $\bm{t}$, $h$, and $v$ be as defined in the statement. 
-As $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$, the two segments both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
-Thus, the intersection of $h$ and $v$ is a point $p$ in $B_\bm{t}$.
-As $r_i$ (for each $i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
-Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$.
-It follows that $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
-
-\extraspacing \underline{\em Proof of Statement (2).} Fix any rectangle $r \in R_1 \cup ...
-\cup R_{k-2}$ and any horizontal segment $h = [x_1, x_2] \times y$ from $H$ that crosses $r$.
-We will show that $h$ must also cross the trimmed rectangle $r'$ of $r$.
-From the fact of $h$ crossing $r$, it is clear that $[x_1, x_2]$ contains the x-range of $r$.
-As the x-range of $r$ contains that of $r'$, we can assert that $[x_1, x_2]$ contains the x-range of $r'$.
+We will start by presenting a set of geometric properties underneath the procedures designed in Section~\ref{sec:hv:type1}.
+These properties are interesting on their own and will enable us to construct simpler proofs later.
+
+\begin{proposition} \label{prop:app:hv:type1:help1}
+    Consider any rectangle $r$ taken from $R_1, R_2, ...,$ or $R_{k-2}$. Let $r'$ be the trimmed rectangle of $r$ defined in Section~\ref{sec:hv:type1}. 
+    \myitems{
+        \item If a horizontal segment $h \in H$ crosses $r$, then $h$ must also cross $r'$.
+        \item If a vertical segment $v \in V$ crosses $r$, then $v$ must also cross $r'$.
+    }
+\end{proposition}
+
+\begin{proof}
+    We will prove only the first bullet due to symmetry.
+Let us represent $h$ as $[x_1, x_2] \times y$.
+The fact of $h$ crossing $r$ indicates that $[x_1, x_2]$ contains the x-range of $r$.
+Since the x-range of $r$ contains that of $r'$, we know that $[x_1, x_2]$ must also contain the x-range of $r'$.
 To prove that $h$ crosses $r'$, we still need to show $y \in [\ybot(r'), \ytop(r')]$.
@@
 Analogously, it also holds that $y \le \ytop(r')$.
-We can thus conclude that $h$ crosses $r'$.
-
-\vgap
-
-Fix any rectangle $r \in R_1 \cup ...
-\cup R_{k-2}$ and any vertical segment $v$ from $V$ that crosses $r$.
-An argument symmetric to the above shows that $v$ must also cross the trimmed rectangle $r'$.
-
-\vgap
-
-We now prove the first claim of statement (2), i.e., $\bm{t}\in \J(R_1',...,R_{k-2}')$.
-Consider the $k$-tuple $(r_1,...,r_{k-2},h,v)\in \J_1$ given in statement (2).
-By definition of $\J_1$, segments $h$ and $v$ both cross each of the rectangles $r_1,..., r_{k-2}$.
-As explained above, $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
-Hence, $h \cap v$ is a point covered by each of $r_1',..., r_{k-2}'$.
-Therefore, $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$ cannot be empty, meaning that $\bm{t} = (r_1', ..., r_{k-2}') \in \J(R_1',...,R_{k-2}')$.
+We can now conclude that $h$ crosses $r'$.
+\end{proof}
+
+We now proceed to elaborate the proofs deferred from Section~\ref{sec:hv:type1}.
+The order of the subsequent proofs will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type1}.
+In particular, we will prove Lemma~\ref{lmm:hv:type1:properties} before Lemma~\ref{lmm:hv:type1:recur-output}, because the claims of the former lemma can be used to produce a succinct argument for the latter.
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:properties}.} We will prove each statement in turn.
+
+\extraspacing \underline{\em Proof of Statement (1).} Let $\bm{t}$, $h$, and $v$ be as defined in the statement. 
+Recall that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+Hence, the segments $h$ and $v$ both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
+Thus, $h \cap v$ is a point $p$ in $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
+As $r_i$ ($i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
+Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$.
+It follows that $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.
+
+\extraspacing \underline{\em Proof of Statement (2).} Consider the $k$-tuple $(r_1,...,r_{k-2},h,v) \in \J_1$ in statement (2).
+By definition of $\J_1$, segments $h$ and $v$ cross each of the rectangles $r_1,..., r_{k-2}$.
+By Proposition~\ref{prop:app:hv:type1:help1}, segments $h$ and $v$ must also cross each of the trimmed rectangles $r_1',..., r_{k-2}'$.
+Hence, $B_\bm{t} = \bigcap_{i = 1}^{k-2}r_i'$ is non-empty as it contains the point $h \cap v$, which proves the first claim $\bm{t} = (r_1', ..., r_{k-2}') \in \J(R_1',...,R_{k-2}')$.
 
@@
 To prove $h \in \dcross_H(\bm{t})$, we need to argue that $h$ crosses $\gbot(\bm{t})$ and $B_\bm{t}$.
-The first part, $h$ crossing $\gbot(\bm{t})$, is done because as mentioned $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is one of those rectangles.
-To prove that $h$ crosses $B_\bm{t}$, first note that $h \cap B_\bm{t} \ne \emptyset$ because as explained before $B_\bm{t}$ contains the point $h \cap v$.
+The first part, $h$ crossing $\gbot(\bm{t})$, is done because as mentioned $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is merely one of those rectangles.
+To prove that $h$ crosses $B_\bm{t}$, first note that $h \cap B_\bm{t} \ne \emptyset$ because as explained before $B_\bm{t}$ contains $h \cap v$.
 On the other hand, as $\gbot(\bm{t})$ covers $B_\bm{t}$, the fact of $h$ crossing $\gbot(\bm{t})$ indicates that the x-range of $h$ contains that of $B_\bm{t}$.

### 2024-06-03 msj.tex
yufei

@@
 \myeqn{
-    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \nn
+    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \label{eqn:hv:type1:R_i'}
 }
@@
 This section supplies the proofs postponed from Section~\ref{sec:hv:type1}.
-We will start by presenting a set of geometric properties underneath the procedures designed in Section~\ref{sec:hv:type1}.
-These properties are interesting on their own and will enable us to construct simpler proofs later.
+We will start by presenting two properties underneath the procedures designed in Section~\ref{sec:hv:type1}.
+These properties will enable us to construct simpler proofs later.
 
@@
 
+\begin{proposition} \label{prop:app:hv:type1:help2}
+    Consider the sets $R_1', R_2', ..., R'_{k-2}$ defined in \eqref{eqn:hv:type1:R_i'}.
+Let $\bm{t}$ be any $(k-2)$-tuple in $\J(R_1', R_2', ..., R'_{k-2})$.
+Then, neither $\dcross_H(\bm{t})$ nor $\dcross_V(\bm{t})$ can be empty.
+\end{proposition}
+
+\begin{proof}
+    Due to symmetry, we will give the proof only for $\dcross_H(\bm{t}) \ne \emptyset$.
+Set $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.
+Define $h$ as the lowest segment crossing $r$ (note that $h$ definitely exists because otherwise $r$ has no trimmed rectangle, contradicting the definition of $r'$).
+We will show that $h \in \dcross_H(\bm{t})$, which indicates $\dcross_H(\bm{t}) \ne \emptyset$.
+For this purpose, we should explain why $h$ crosses both $B_{\bm{t}}$ and $r'$.
+However, by Proposition~\ref{prop:app:hv:type1:help1}, $h$ crossing $r$ directly implies $h$ crossing $r'$.
+It remains to prove that $h$ crosses $B_\bm{t}$.
+
+    \vgap
+
+    By the definitions of $r'$ and $h$, the bottom edge of $r'$ must be contained in $h$.
+Because $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$ and thus also contained in $h$.
+This means that $h \cap B_{\bm{t}}\neq \emptyset$.
+On the other hand, the x-range of $h$ must cover that of $r'$ (because $h$ crosses $r'$), which in turn must covers that of $B_\bm{t}$ (because $r'$ covers $B_\bm{t}$).
+Thus, the x-range of $h$ covers that of $B_\bm{t}$.
+This together with $h \cap B_{\bm{t}}\neq \emptyset$ tells us that $h$ must cross $B_\bm{t}$.
+\end{proof}
+
 We now proceed to elaborate the proofs deferred from Section~\ref{sec:hv:type1}.
@@
 
-\extraspacing \underline{\em Proof of Statement (3).} We start by proving that $\dcross_H(\bm{t}) \neq \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
-Set $r' = \gbot(\bm{t})$, and let $r$ be the full rectangle of $r'$.
-Define $h$ as the lowest segment crossing $r$ (note that $h$ definitely exists because otherwise $r$ has no trimmed rectangle, contradicting the definition of $r'$).
-We will show that $h \in \dcross_H(\bm{t})$ and hence $\dcross_H(\bm{t}) \ne \emptyset$.
-For this purpose, we should explain why $h$ crosses both $B_{\bm{t}}$ and $r'$.
-However, $h$ crossing $r$ implies $h$ crossing $r'$, as we already know from the above proof of statement (2).
-It remains to show that $h$ crosses $B_\bm{t}$.
-
-\vgap
-
-By the definitions of $r'$ and $h$, the bottom edge of $r'$ must be contained in $h$.
-Because $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$ and hence also contained in $h$.
-This means that $h \cap B_{\bm{t}}\neq \emptyset$.
-On the other hand, the x-range of $h$ must contain that of $r'$ (because $h$ crosses $r'$), which in turn must contain that of $B_\bm{t}$ (because $r'$ covers $B_\bm{t}$).
-Thus, the x-range of $h$ covers that of $B_\bm{t}$.
-Combining this with $h \cap B_{\bm{t}}\neq \emptyset$ yields that $h$ must cross $B_\bm{t}$.
-
-\vgap
-
-A symmetric argument assures us that $\dcross_V(\bm{t}) \ne \emptyset$ for any $\bm{t} \in \J(R_1',...,R_{k-2}')$.
-
-\vgap
-
-Next, we will prove 
+\extraspacing \underline{\em Proof of Statement (3).} We will prove 
 \myeqn{
@@
 } 
-which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$.
+which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$ by Proposition~\ref{prop:app:hv:type1:help2}.
 Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}).
@@
 This allows us to create an array of size $n$ and allocate an array cell to each $r \in R_1 \cup ...
-\cup R_{k-2}$, with the property that the cell can be accessed by the ID of $r$ in constant time.
+\cup R_{k-2}$.
+The cell can be accessed by the ID of $r$ in constant time.
 

### 2024-06-03 msj.tex
yufei

@@
 Next, we will prove that a segment $h \in \cross_H(r^*)$ if and only if $h \in \dcross_H(\bm{t})$.
-Let us start with the ``only-if direction'': assuming $h \in \cross_H(r^*)$, we aim to show $h \in \dcross_H(\bm{t})$.
-For this purpose, we need to show that $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
+Let us start with the ``only-if direction'': assuming $h \in \cross_H(r^*)$, we aim to show $h \in \dcross_H(\bm{t})$, or in other words, $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
 \myitems {
     \item First, we explain why $h$ crosses $r'$.
-Because $r'$ covers $r^*$, by the fact $h$ crossing $r^*$, we know that $h$ must intersect with $r'$.
+Because $r'$ covers $r^*$, by the fact of $h$ crossing $r^*$, we know that $h \cap r' \ne \emptyset$.
 On the other hand, as $h$ crosses $r^*$, it must hold that $\xleft(h) < x_1 \le x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$).
-As $[x_1, x_2]$ is also the x-range of $r'$, using the fact of $h \cap r' \ne \emptyset$, we can assert that $h$ crosses $r'$.
+As $[x_1, x_2]$ is also the x-range of $r'$, the fact of $h \cap r' \ne \emptyset$ indicates that $h$ must cross $r'$.
 

### 2024-06-03 msj.tex
yufei

@@
     \item First, we explain why $h$ crosses $r'$.
-Because $r'$ covers $r^*$, by the fact of $h$ crossing $r^*$, we know that $h \cap r' \ne \emptyset$.
-On the other hand, as $h$ crosses $r^*$, it must hold that $\xleft(h) < x_1 \le x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$).
+Because $r^* \subseteq r'$, the fact of $h$ crossing $r^*$ tells us that $h \cap r' \ne \emptyset$.
+On the other hand, the fact of $h$ crossing $r^*$ also tells us that $\xleft(h) < x_1 \le x_2 < \xright(h)$ (recall that $[x_1, x_2]$ is the x-range of $r^*$).
 As $[x_1, x_2]$ is also the x-range of $r'$, the fact of $h \cap r' \ne \emptyset$ indicates that $h$ must cross $r'$.
@@
     \item Then, we explain why $h$ crosses $B_{\bm{t}}$.
-Note that because $h$ crosses $r'$ (which we just proved) and $B_{\bm{t}}\subseteq r'$, we know that $B_{\bm{t}}$ cannot contain any endpoint of $h$.
-Let $y_h$ be the y-coordinate of $h$, and take the point $p = (\xleft(B_{\bm{t}}), y_h)$.
-We will prove $p \in h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$ and, hence, $h$ must cross $B_{\bm{t}}$ (because no endpoint of $h$ falls in $B_\bm{t}$, as mentioned).
-
-    \vgap
-
-    Since $\gbot(\bm{t}) = r'$, the bottom edge of $B_{\bm{t}}$ must be contained in $r'$.
+Because $B_{\bm{t}}\subseteq r'$ and $h$ crosses $r'$ (as was just proved), the rectangle $B_{\bm{t}}$ cannot contain any endpoint of $h$.
+Let us take the y-coordinate $y_h$ of $h$.
+We will prove that the point $p = (\xleft(B_{\bm{t}}), y_h)$ is on the segment $h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$.
+Once done, this will indicate that $h$ crosses $B_{\bm{t}}$ (because no endpoint of $h$ falls in $B_\bm{t}$).
+
+    \vgap
+
+    As $\gbot(\bm{t}) = r'$, the bottom edge of $B_{\bm{t}}$ is contained in $r'$.
 Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$.
-As proved earlier, $\xleft(h) < x_1 \le x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting $p \in h$.
-
-    \vgap
-
-    Because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$.
-To prove $p \in B_{\bm{t}}$, it remains to show $y_h \in [y_1, \maxtop(r')]$.
-This is true because $h$ intersects $r^*$ and $[y_1, \maxtop(r')]$ is the y-range of $r^*$.
+As proved earlier, $\xleft(h) < x_1 \le x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting that $p \in h$.
+
+    \vgap
+
+    As $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$.
+Because $h \cap r^* \ne \emptyset$ (as $h$ crosses $r^*$) and $[y_1, \maxtop(r')]$ is the y-range of $r^*$, we know that $y_h \in [y_1, \maxtop(r')]$, suggesting that $p \in B_{\bm{t}}$.
 }

### 2024-06-03 msj.tex
yufei

@@
 
-Our argument now proceeds to prove the ``if direction'': assuming $h \in \dcross_H(\bm{t})$, we aim to show $h \in \cross_H(r^*)$.
+We now proceed to establish the ``if direction'': assuming $h \in \dcross_H(\bm{t})$, we aim to show $h \in \cross_H(r^*)$.
 Recall that $r^* = [x_1,x_2] \times [y_1, \maxtop(r')]$.
-To prove that $h$ crosses $r^*$, we will show $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
+To prove that $h$ crosses $r^*$, we will argue that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
 \myitems{

### 2024-06-03 msj.tex
yufei

@@
 \myeqn{
-    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})| \cdot |\dcross_V(\bm{t})| \leq \out. \label{eqn:hv:type1:properties:proof-help1}
+    \sum_{\bm{t}\in \J(R_1',...,R_{k-2}')} |\dcross_H(\bm{t})| \cdot |\dcross_V(\bm{t})| \leq \out \label{eqn:hv:type1:properties:proof-help1}
 } 
@@
 
-\extraspacing \underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\dcross_H(\bm{t})$ $= \cross_H(r^*)$.
-After this, we can then derive
+\extraspacing \underline{\em Proof of Statement (1).} We will map each $r^*$ of $\bigcup_{i = 1}^{k-2} R_i^*$ to a unique tuple $\bm{t}$ in $\J(R_1',...,R_{k-2}')$ satisfying $\cross_H(r^*) \subseteq \dcross_H(\bm{t})$.
+The mapping allows us to derive
 \myeqn{
@@
 
-The mapping is designed as follows.
+The mapping is as follows.
 Consider an arbitrary $r^*\in \bigcup_{i = 1}^{k-2} R_i^*$.
@@
 
-Next, we will prove that a segment $h \in \cross_H(r^*)$ if and only if $h \in \dcross_H(\bm{t})$.
-Let us start with the ``only-if direction'': assuming $h \in \cross_H(r^*)$, we aim to show $h \in \dcross_H(\bm{t})$, or in other words, $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
+Next, we will prove 
+\begin{center}
+    {\bf Claim 1:} If a segment $h \in \cross_H(r^*)$, then $h \in \dcross_H(\bm{t})$.
+\end{center}
+To prove the claim, we must show $h \in \dcross_H(\bm{t})$, or in other words, $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
 \myitems {
@@
     \item Then, we explain why $h$ crosses $B_{\bm{t}}$.
-Because $B_{\bm{t}}\subseteq r'$ and $h$ crosses $r'$ (as was just proved), the rectangle $B_{\bm{t}}$ cannot contain any endpoint of $h$.
+Because $B_{\bm{t}}\subseteq r'$ and $h$ crosses $r'$ (just proved), the rectangle $B_{\bm{t}}$ cannot contain any endpoint of $h$.
 Let us take the y-coordinate $y_h$ of $h$.
@@
 
-    As $\gbot(\bm{t}) = r'$, the bottom edge of $B_{\bm{t}}$ is contained in $r'$.
+    \myitems{
+        \item As $\gbot(\bm{t}) = r'$, the bottom edge of $B_{\bm{t}}$ is contained in $r'$.
 Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$.
@@
 
-    As $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$.
+        \item As $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$.
 Because $h \cap r^* \ne \emptyset$ (as $h$ crosses $r^*$) and $[y_1, \maxtop(r')]$ is the y-range of $r^*$, we know that $y_h \in [y_1, \maxtop(r')]$, suggesting that $p \in B_{\bm{t}}$.
     }
-
-We now proceed to establish the ``if direction'': assuming $h \in \dcross_H(\bm{t})$, we aim to show $h \in \cross_H(r^*)$.
+}
+
+% We now turn to the ``if direction'': assuming $h \in \dcross_H(\bm{t})$, we aim to show $h \in \cross_H(r^*)$.
 Recall that $r^* = [x_1,x_2] \times [y_1, \maxtop(r')]$.
 To prove that $h$ crosses $r^*$, we will argue that $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$, where $y_h$ is the y-coordinate of $h$.
-\myitems{
-    \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $r'$. As $[x_1, x_2]$ is the x-range of $r'$, we know $\xleft(h) < x_1 \le x_2 < \xright(h)$.
-
-    \vgap
-
-    \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$.
-As $[y_1, \maxtop(r')]$ is the y-range of $B_\bm{t}$ (because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$), we know $y_h \in [y_1, \maxtop(r')]$.
-}
+% \myitems{
+%     \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $r'$. As $[x_1, x_2]$ is the x-range of $r'$, we have $\xleft(h) < x_1 \le x_2 < \xright(h)$.
+% 
+%     \vgap
+% 
+%     \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$.
+As $[y_1, \maxtop(r')]$ is the y-range of $B_\bm{t}$ (because $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$), we have $y_h \in [y_1, \maxtop(r')]$.
+% }
 

### 2024-06-03 msj.tex
yufei

@@
 \end{center}
-To prove the claim, we must show $h \in \dcross_H(\bm{t})$, or in other words, $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
+In other words, we must show that $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
 \myitems {
@@
 
-Firstly, we will prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$.
-For this purpose, given any segment $h \in \dcross_H(\bm{t})$, we will show that $h$ crosses $r^*$.
-To do so, we need to explain why $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1, \maxtop(r')]$.
+First, we will prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$ or equivalently: if a segment $h \in \dcross_H(\bm{t})$, then $h$ crosses $r^*$.
+To do so, we need to explain why $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1,$ $\maxtop(r')]$.
 \begin{itemize}
-    \item The fact $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $\gbot(\bm{t})$, which is $r'$.
+    \item The fact of $h \in \dcross_H(\bm{t})$ tells us that $h$ crosses $\gbot(\bm{t})$, which is $r'$.
 As the x-range of $r'$ is $[x_1,x_2]$, it must hold true that $\xleft(h) < x_1 \le x_2 < \xright(h)$.
@@
 
-    \item The fact $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$.
+    \item The fact of $h \in \dcross_H(\bm{t})$ also tells us that $h$ crosses $B_{\bm{t}}$.
 Hence, $y_h \in [\ybot(B_\bm{t}), \ytop(B_{\bm{t}})]$.
@@
 \end{itemize}
-We can now conclude that $h$ crosses $r^*$.
-
-\vgap
-
-Secondly, assuming that the segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates,
-we will prove that $\dcross_H(\bm{t})$ includes a prefix of the sorted order. It suffices to establish the following equivalent statement:
+
+Second, assuming that the segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates,
+we will prove that $\dcross_H(\bm{t})$ includes a prefix of the sorted order. It suffices to establish the following equivalent claim:
 
@@
 \minipg{0.8\linewidth} {
-    If a segment $h \in \cross_H(r^*)$ has y-coordinate $y_h \le \ytop(B_{\bm{t}})$, then $h$ must be in $\dcross_H(\bm{t})$.
+    {\bf Claim 2:} If a segment $h \in \cross_H(r^*)$ has y-coordinate $y_h \le \ytop(B_{\bm{t}})$, then $h$ must be in $\dcross_H(\bm{t})$.
 }
@@
 
-\vgap
-
-We first show that $h$ crosses $r'$, or equivalently: $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$.
-These conditions hold true because $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$ (due to the definition of top-sliced rectangle in \eqref{eqn:hv:type1:topsliced}).
-
-\vgap
-
-Finally, we show that $h$ crosses $B_{\bm{t}}$, or equivalently: $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [\ybot(B_\bm{t}), \ytop(B_{\bm{t}})] = [y_1, \ytop(B_{\bm{t}})]$.
+\myitems{
+
+\item We first show that $h$ crosses $r'$, or equivalently: $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$.
+These conditions hold true because $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$ (due to the definition of a top-sliced rectangle in \eqref{eqn:hv:type1:topsliced}).
+
+\vgap
+
+\item Next we show that $h$ crosses $B_{\bm{t}}$, or equivalently: $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$ and $y_h \in [\ybot(B_\bm{t}), \ytop(B_{\bm{t}})] = [y_1, \ytop(B_{\bm{t}})]$.
+
+    \vgap
+
 \begin{itemize}
@@
 As $r' = \gbot(B_{\bm{t}})$, the x-range of $B_{\bm{t}}$ must be contained in $[x_1,x_2]$.
-Therefore, $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$.
+Thus, $\xleft(h) < \xleft(B_{\bm{t}}) \le \xright(B_{\bm{t}}) < \xright(h)$.
 
@@
     \item The fact of $h$ crossing $r^*$ also tells us that $y_h \in [y_1, \maxtop(r')]$.
-Moreover, by the definition of $h$, we have $y_h \leq \ytop(B_{\bm{t}})$.
+Moreover, Claim 2 explicitly gives us the condition $y_h \leq \ytop(B_{\bm{t}})$.
 It thus follows that $y_h \in [y_1, \ytop(B_{\bm{t}})]$.
 \end{itemize}
+}
 

### 2024-06-03 msj.tex
yufei

@@
 
-    By the definitions of $r'$ and $h$, the bottom edge of $r'$ must be contained in $h$.
+    By the definitions of $r'$ and $h$, the bottom edge of $r'$ must be contained in $h$ (note that $h$ is one of the segments used to trim $r$ into $r'$).
 Because $r' = \gbot(\bm{t})$, the bottom edge of $B_\bm{t}$ is contained in the bottom edge of $r'$ and thus also contained in $h$.
 This means that $h \cap B_{\bm{t}}\neq \emptyset$.
-On the other hand, the x-range of $h$ must cover that of $r'$ (because $h$ crosses $r'$), which in turn must covers that of $B_\bm{t}$ (because $r'$ covers $B_\bm{t}$).
+On the other hand, the x-range of $h$ must cover that of $r'$ (because $h$ crosses $r'$), which in turn must cover that of $B_\bm{t}$ (because $r'$ covers $B_\bm{t}$).
 Thus, the x-range of $h$ covers that of $B_\bm{t}$.

### 2024-06-03 msj.tex
yufei

@@
 
-This section supplies the proofs postponed from Section~\ref{sec:hv:type1}.
-We will start by presenting two properties underneath the procedures designed in Section~\ref{sec:hv:type1}.
-These properties will enable us to construct simpler proofs later.
+%This section supplies the proofs postponed from Section~\ref{sec:hv:type1}.
+
+We start by presenting two properties underneath the procedures designed in Section~\ref{sec:hv:type1}. These properties will enable us to construct simpler proofs later.
 
@@
 
-We now proceed to elaborate the proofs deferred from Section~\ref{sec:hv:type1}.
+We now proceed to elaborate the proofs postponed from Section~\ref{sec:hv:type1}.
 The order of the subsequent proofs will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type1}.
@@
 Recall that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
-Hence, the segments $h$ and $v$ both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
-Thus, $h \cap v$ is a point $p$ in $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
+The segments $h$ and $v$ both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
+Thus, $h \cap v$ is a point in $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
 As $r_i$ ($i \in [k-2]$) is the full rectangle of $\bm{t}[i]$, we know that $\bigcap_{i=1}^{k-2} r_i$ covers $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
-Hence, point $p$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$.
+Hence, point $h \cap v$ falls in $\bigcap_{i=1}^{k-2} r_i$, indicating that $h\cap v \cap \bigcap_{i = 1}^{k-2}r_i \neq \emptyset$.
 It follows that $(r_1,...,r_{k-2},h,v)$ is a result tuple in $\J(R_1,...,R_{k-2},H,V)$.

### 2024-06-03 msj.tex
yufei

@@
 It suffices to show only the former due to symmetry.
-To prove $h \in \dcross_H(\bm{t})$, we need to argue that $h$ crosses $\gbot(\bm{t})$ and $B_\bm{t}$.
+For that purpose, we need to argue that $h$ crosses $\gbot(\bm{t})$ and $B_\bm{t}$.
 The first part, $h$ crossing $\gbot(\bm{t})$, is done because as mentioned $h$ crosses each of $r_1', ..., r_{k-2}'$, and $\gbot(\bm{t})$ is merely one of those rectangles.
@@
 } 
-which implies statement (3) because, for any $\bm{t}$ in the summation, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$ by Proposition~\ref{prop:app:hv:type1:help2}.
-Our proof resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}).
-This algorithm adds to $\J^*$ exactly as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}.
-By statement (1) of Lemma~\ref{lmm:hv:type2:properties}, the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whereas $\out = |\J(R_1,...,R_{k-2},H,V)|$.
-This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}, which completes the proof of statement (3).
-
-\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} Our proof again resorts to the algorithm generate-$\J^*$ (see Section~\ref{sec:hv:type1}).
+which implies statement (3) because, by Proposition~\ref{prop:app:hv:type1:help2}, $|\dcross_H(\bm{t})| \ge 1$ and $|\dcross_V(\bm{t})| \ge 1$  for any $\bm{t}$ in the summation.
+Our proof resorts to the algorithm generate-$\J^*$ given in Section~\ref{sec:hv:type1}.
+This algorithm adds to $\J^*$ as many tuples as calculated by the left hand side of \eqref{eqn:hv:type1:properties:proof-help1}.
+By statement (1) of Lemma~\ref{lmm:hv:type1:properties}, the $\J^*$ produced must be a subset of $\J(R_1,...,R_{k-2},H,V)$, whereas $\out = |\J(R_1,...,R_{k-2},H,V)|$.
+This establishes the inequality in \eqref{eqn:hv:type1:properties:proof-help1}.
+
+\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} Our proof again resorts to the algorithm generate-$\J^*$ in Section~\ref{sec:hv:type1}.
 As argued in Statement (3) of Lemma~\ref{lmm:hv:type1:properties}, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ are non-empty for any $\bm{t} \in \J(R_1',$ $...,R_{k-2}')$.
-Therefore, in processing this $\bm{t}$, the algorithm adds at least a new tuple to $\J^*$.
-Hence, $|\J(R_1',...,R_{k-2}')| \le |\J^*|\le \out$, where the second inequality follows from Statement (1) of Lemma~\ref{lmm:hv:type1:properties}.
+Therefore, in processing this $\bm{t}$, the algorithm adds at least one new tuple to $\J^*$.
+Hence, $|\J(R_1',...,R_{k-2}')| \le |\J^*|\le |\J(R_1,...,R_{k-2},H,V)| = \out$, where the second inequality used Statement (1) of Lemma~\ref{lmm:hv:type1:properties}.
 
@@
 \end{center}
-In other words, we must show that $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
+For this purpose, we need to show that $h$ crosses both $\gbot(\bm{t}) = r'$ and $B_\bm{t}$.
 \myitems {
@@
 We will prove that the point $p = (\xleft(B_{\bm{t}}), y_h)$ is on the segment $h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$.
-Once done, this will indicate that $h$ crosses $B_{\bm{t}}$ (because no endpoint of $h$ falls in $B_\bm{t}$).
+Once done, we can then assert that $h$ crosses $B_{\bm{t}}$ (as no endpoint of $h$ falls in $B_\bm{t}$).
 
@@
         \item As $\gbot(\bm{t}) = r'$ and $\maxtop(r') = \ytop(B_\bm{t})$, the y-range of $B_\bm{t}$ is $[y_1, \maxtop(r')]$.
-Because $h \cap r^* \ne \emptyset$ (as $h$ crosses $r^*$) and $[y_1, \maxtop(r')]$ is the y-range of $r^*$, we know that $y_h \in [y_1, \maxtop(r')]$, suggesting that $p \in B_{\bm{t}}$.
+Because $h \cap r^* \ne \emptyset$ (as $h$ crosses $r^*$) and $[y_1, \maxtop(r')]$ is also the y-range of $r^*$, we know that $y_h \in [y_1, \maxtop(r')]$, suggesting that $p \in B_{\bm{t}}$.
     }

### 2024-06-03 msj.tex
yufei

@@
 \myitems {
-    \item First, we explain why $h$ crosses $r'$.
+    \item First, $h$ crosses $r'$ follows from the facts that (i) $h$ crosses $r^*$, (ii) $r^* \subseteq r'$, and (iii) $r^*$ and $r'$ share the same x-range.
+
+    %.
 Because $r^* \subseteq r'$, the fact of $h$ crossing $r^*$ tells us that $h \cap r' \ne \emptyset$.
@@
     \item Then, we explain why $h$ crosses $B_{\bm{t}}$.
-Because $B_{\bm{t}}\subseteq r'$ and $h$ crosses $r'$ (just proved), the rectangle $B_{\bm{t}}$ cannot contain any endpoint of $h$.
+Because $B_{\bm{t}}\subseteq r'$ and $h$ crosses $r'$ (just proved), no endpoint of $h$ can fall in $B_\bm{t}$.
 Let us take the y-coordinate $y_h$ of $h$.
 We will prove that the point $p = (\xleft(B_{\bm{t}}), y_h)$ is on the segment $h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$.
-Once done, we can then assert that $h$ crosses $B_{\bm{t}}$ (as no endpoint of $h$ falls in $B_\bm{t}$).
+Once this is done, we can assert that $h$ crosses $B_{\bm{t}}$ (as no endpoint of $h$ falls in $B_\bm{t}$).
 
@@
 Thus, $\xleft(B_{\bm{t}})\in [x_1,x_2]$.
-As proved earlier, $\xleft(h) < x_1 \le x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting that $p \in h$.
+The fact of $h$ crossing $r^*$ tells us that $\xleft(h) < x_1 \le x_2 < \xright(h)$, which leads to $\xleft(B_{\bm{t}}) \in [\xleft(h), \xright(h)]$, suggesting that $p \in h$.
 

### 2024-06-04 msj.tex

@@
 
-First, we will prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$ or equivalently: if a segment $h \in \dcross_H(\bm{t})$, then $h$ crosses $r^*$.
+We will first prove $\dcross_H(\bm{t})\subseteq \cross_H(r^*)$ or equivalently: if a segment $h \in \dcross_H(\bm{t})$, then $h$ crosses $r^*$.
 To do so, we need to explain why $\xleft(h) < x_1 \le x_2 < \xright(h)$ and $y_h \in [y_1,$ $\maxtop(r')]$.
@@
 
-Second, assuming that the segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates,
+Next, assuming that the segments of $\cross_H(r^*)$ are sorted in ascending order of their y-coordinates,
 we will prove that $\dcross_H(\bm{t})$ includes a prefix of the sorted order. It suffices to establish the following equivalent claim:
@@
 \item We first show that $h$ crosses $r'$, or equivalently: $\xleft(h)  < x_1 \leq x_2 < \xright(h)$ and $y_h \in [y_1, y_2]$.
-These conditions hold true because $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and $\maxtop(r')\leq y_2$ (due to the definition of a top-sliced rectangle in \eqref{eqn:hv:type1:topsliced}).
+These conditions hold true because (i) $h$ crosses $r^* = [x_1,x_2]\times [y_1, \maxtop(r')]$, and (ii) the definition of a top-sliced rectangle in \eqref{eqn:hv:type1:topsliced} tells us $\maxtop(r')\leq y_2$.
 

### 2024-06-04 msj.tex
yufei

@@
 Let $h$ be the full segment of $h'$.
-As $h$ contains $h'$, we must have $h \cap v = h' \cap v$.
-
-\vgap
-
-Let $r_{k-2}$ be an arbitrary rectangle in the set $\contained_{R_{k-2}}(\bm{t})$.
-The definition of $\contained_{R_{k-2}}(\bm{t})$ tells us that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ (see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment), whose right endpoint is $h' \cap v$.
-This means that $h \cap v$ falls in $r_{k-2}$ as well.
-
-\vgap
-
-Now we know that $h \cap v \cap \bigcap_{i = 1}^{k-2} r_i$ is not empty because it equals $h' \cap v$. Hence,
-$(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
+As $h$ contains $h'$, we have $h \cap v = h' \cap v$.
+
+\vgap
+
+Consider any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
+From the definition of $\contained_{R_{k-2}}(\bm{t})$, we know that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ --- see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment --- whose right endpoint is $h' \cap v$.
+This means that $h \cap v$ falls in $r_{k-2}$.
+Therefore, $h \, \cap \, v \, \cap \, \bigcap_{i = 1}^{k-2} r_i$ equals $h' \cap v$ and hence is non-empty, meaning that $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
 To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to explain why $r_{k-2}$ covers the left endpoint of $h$.

### 2024-06-04 msj.tex
yufei

@@
 
-\extraspacing \underline{\em Proof of Statement (1).} Let $\bm{t}$, $h$, and $v$ be as defined in the statement. 
-Recall that $h \in \dcross_H(\bm{t})$ and $v \in \dcross_V(\bm{t})$.
+\extraspacing \underline{\em Proof of Statement (1).} Take any $(k-2)$-tuple $\bm{t} \in \J(R_1', R'_2, ..., R'_{k-2})$.
+Let $r_i$ ($i$ $\in$  $[k-2]$) be the full rectangle of $\bm{t}[i]$.
+Fix any $h \in \dcross_H(\bm{t})$ and any $v \in \dcross_V(\bm{t})$.
 The segments $h$ and $v$ both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
@@
 
-\extraspacing \underline{\em Proof of Statement (2).} Consider the $k$-tuple $(r_1,...,r_{k-2},h,v) \in \J_1$ in statement (2).
+\extraspacing \underline{\em Proof of Statement (2).} Take any $k$-tuple $(r_1,...,r_{k-2},h,v) \in \J_1$.
 By definition of $\J_1$, segments $h$ and $v$ cross each of the rectangles $r_1,..., r_{k-2}$.
@@
 
-\extraspacing \underline{\em Proof of Statement (2).} Let $r', r^*$, and $\bm{t}$ be as defined in statement (2).
-Let us represent $r'$ as $[x_1,x_2] \times [y_1, y_2]$.
-Accordingly, $r^*$ can be written as $[x_1, x_2] \times [y_1, \maxtop(r')]$.
+\extraspacing \underline{\em Proof of Statement (2).} Take any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$.
+Set $r' = \gbot(\bm{t})$, and let $r^*$ be the top-sliced rectangle of $r'$.
+If we represent $r'$ as $[x_1,x_2] \times [y_1, y_2]$, then $r^*$ can be written as $[x_1, x_2] \times [y_1, \maxtop(r')]$.
 
@@
 
-\extraspacing \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, and $\bm{t}$ be as defined in statement (2).
+\extraspacing \underline{\em Proof of Statement (2).} Take any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
+Let $h'$ be the trimmed segment of $h$ and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
 Firstly, we will argue that the $(k-1)$-tuple $\bm{t}$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, which is equivalent to saying that that the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.

### 2024-06-04 msj.tex
yufei

@@
 \extraspacing \underline{\em Proof of Statement (2).} Take any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
-Let $h'$ be the trimmed segment of $h$ and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
-Firstly, we will argue that the $(k-1)$-tuple $\bm{t}$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, which is equivalent to saying that that the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
-The subsequent discussion will prove that $h' \cap v$ is in fact the same point as $h \cap v$.
+Define $h'$ as the trimmed segment of $h$.
+We will first argue that the $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, or equivalently, the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
+The subsequent discussion will show that $h' \cap v$ is in fact the same point as $h \cap v$.
 From $(r_1,...,r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$, we already know that $h \cap v$ falls in all of $r_1,...,r_{k-2}$.
-Therefore, obtaining $h' \cap v = h \cap v$ will allow us to assert that $r_1,...,r_{k-3}$ all cover $h' \cap v$.
-
-\vgap
-
-Denote by $p$ the intersection point of $h$ and $v$.
+It will then follow that $r_1,...,r_{k-3}$ all cover $h' \cap v$.
+
+\vgap
+
 As $h'$ is contained in $h$, to prove $h' \cap v = p$, it suffices to explain why $p$ is on $h'$.
-Let us represent $h$ as $[x_1,x_2] \times y$ and $p$ as $(x, y)$ for some $x \in [x_1, x_2]$.
-Accordingly, we can represent $h'$ as $[x_1, x_2'] \times y$ for some $x_2' \in [x_1, x_2]$.
-By the definition of a trimmed segment, we know that $(x_2',y)$ is the rightmost point on $h$ that falls in a left-end covering rectangle of $h$ in $R_{k-2}$.
-From statement (2), we have $(r_1,...,r_{k-2},h,v) \in \J_2$, meaning that $r_{k-2}$ (i) is a left-end covering rectangle of $h$, and (ii) covers $p$ ($= h \cap v$).
-Therefore, the x-coordinate of $p$ cannot exceed $x_2'$ (i.e., $x \le x_2'$), allowing us to assert that $p$ is on $h'$.
+Let us represent $h$ as $[x_1,x_2] \times y$.
+Accordingly, the segment $h'$ can be written as $[x_1, x_2'] \times y$ for some $x_2' \in [x_1, x_2]$.
+By the definition of a trimmed segment, $(x_2',y)$ is the rightmost point on $h$ that falls in a left-end covering rectangle of $h$ in $R_{k-2}$.
+As $(r_1,...,r_{k-2},h,v) \in \J_2$, we know that $r_{k-2}$ is a left-end covering rectangle of $h$, and $r_{k-2}$ covers $p$.
+Therefore, the x-coordinate of $p$ cannot exceed $x_2'$, allowing us to assert that $p$ is on $h'$.
 

### 2024-06-04 msj.tex
yufei

@@
 \extraspacing \underline{\em Proof of Statement (2).} Take any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
-Define $h'$ as the trimmed segment of $h$.
-We will first argue that the $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, or equivalently, the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
-The subsequent discussion will show that $h' \cap v$ is in fact the same point as $h \cap v$.
-From $(r_1,...,r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$, we already know that $h \cap v$ falls in all of $r_1,...,r_{k-2}$.
+Define $h'$ as the trimmed segment of $h$, and point $p = h \cap v$.
+We first argue that the $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, or equivalently, the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
+The subsequent discussion will show that $h' \cap v$ is in fact the same point as $p = h \cap v$.
+From $(r_1,...,r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$, we already know that $p$ falls in all of $r_1,...,r_{k-2}$.
 It will then follow that $r_1,...,r_{k-3}$ all cover $h' \cap v$.
@@
 
-As $h'$ is contained in $h$, to prove $h' \cap v = p$, it suffices to explain why $p$ is on $h'$.
+As $h'$ is contained in $h$, for proving $h' \cap v = p$, it suffices to explain why $p$ is on $h'$.
 Let us represent $h$ as $[x_1,x_2] \times y$.
@@
 
-%This tells us $B_\bm{t} = p$.
-
-Secondly, we will argue that $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
-The above discussion has shown that $h \cap v = h' \cap v = p$.
-If (as before) we write $h = [x_1,x_2] \times y$, $h' = [x_1,x_2'] \times y$, and $p = (x,y)$, then the effective horizontal segment of $\bm{t}$ can be written as $[x_1, x] \times y$ (due to the definition of an effective horizontal segment).
-To prove $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers the entire segment $[x_1, x] \times y$.
-We do so in two steps:
-
-% $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$ and $y \in [\ybot(r_2), \ytop(r_2)]$.
+Next, we argue that $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
+Let $p'$ be the left endpoint of $h$.
+Because we already know $h' \cap v = h \cap v = p$, the effective horizontal segment of $\bm{t}$ is the segment $p'p$.
+To prove $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers this segment.
+This is true because:
 
 \begin{itemize}
-    \item As $r_{k-2}$ is a left-end covering rectangle of $h$, we must have $x_1 \in [\xleft(r_2), \xright(r_2)]$.
-On the other hand, the fact $r_{k-2}$ covering $p$ gives $x \in [\xleft(r_2), \xright(r_2)]$.
-It thus follows that $[x_1,x] \subseteq [\xleft(r_2), \xright(r_2)]$.
-
-    \vgap
-
-    \item The fact $h$ intersecting with $r_{k-2}$ gives $y \in [\ybot(r_2), \ytop(r_2)]$.
+    \item $r_{k-2}$ covers $p$, as we already know;
+    \item $r_{k-2}$ covers $p'$, as it is a left-end covering rectangle of $h$.
 \end{itemize}
-
-\noindent Combining the above tells us that  $r_{k-2}$ covers $[x_1, x] \times y$.
 

### 2024-06-04 msj.tex
yufei

@@
 This means that $h \cap v$ falls in $r_{k-2}$.
-Therefore, $h \, \cap \, v \, \cap \, \bigcap_{i = 1}^{k-2} r_i$ equals $h' \cap v$ and hence is non-empty, meaning that $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
+We now have $h \, \cap \, v \, \cap \, \bigcap_{i = 1}^{k-2} r_i$ equals $h' \cap v$ and hence is non-empty, meaning that $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
 To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to explain why $r_{k-2}$ covers the left endpoint of $h$.

### 2024-06-04 msj.tex
yufei

@@
 \extraspacing \underline{\em Proof of Statement (2).} Take any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
-Define $h'$ as the trimmed segment of $h$, and point $p = h \cap v$.
+Set $p = h \cap v$, and define $h'$ as the trimmed segment of $h$.
 We first argue that the $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, or equivalently, the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
@@
 By the definition of a trimmed segment, $(x_2',y)$ is the rightmost point on $h$ that falls in a left-end covering rectangle of $h$ in $R_{k-2}$.
-As $(r_1,...,r_{k-2},h,v) \in \J_2$, we know that $r_{k-2}$ is a left-end covering rectangle of $h$, and $r_{k-2}$ covers $p$.
+As $(r_1,...,r_{k-2},h,v) \in \J_2$, we know that $r_{k-2}$ is a left-end covering rectangle of $h$, and $r_{k-2}$ covers $p$ (which is a point on $h$).
 Therefore, the x-coordinate of $p$ cannot exceed $x_2'$, allowing us to assert that $p$ is on $h'$.

### 2024-06-04 msj.tex
yufei

@@
 
-This section supplies the proofs postponed from Section~\ref{sec:hv:type2}.
-The order of the proofs here will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type2}.
-In particular, we will prove Lemma~\ref{lmm:hv:type2:properties} before Lemma~\ref{lmm:hv:type2:recur-output}, because the claims of the former lemma can be used to construct a simple argument for the latter.
-
-\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.} We will prove each statement in turn.
-
-\extraspacing \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$.
-The point $h' \cap v$ is covered by all of $r_1,...,r_{k-3}$.
-Let $h$ be the full segment of $h'$.
-As $h$ contains $h'$, we have $h \cap v = h' \cap v$.
-
-\vgap
-
-Consider any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
-From the definition of $\contained_{R_{k-2}}(\bm{t})$, we know that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ --- see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment --- whose right endpoint is $h' \cap v$.
-This means that $h \cap v$ falls in $r_{k-2}$.
-We now have $h \, \cap \, v \, \cap \, \bigcap_{i = 1}^{k-2} r_i$ equals $h' \cap v$ and hence is non-empty, meaning that $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
-
-\vgap
-
-To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to explain why $r_{k-2}$ covers the left endpoint of $h$.
-This follows immediately from the fact that $r_{k-2}$ covers the effective horizontal segment of $h'$ (which shares the same left endpoint as $h$).
-
-\extraspacing \underline{\em Proof of Statement (2).} Take any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
-Set $p = h \cap v$, and define $h'$ as the trimmed segment of $h$.
-We first argue that the $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, or equivalently, the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
-The subsequent discussion will show that $h' \cap v$ is in fact the same point as $p = h \cap v$.
-From $(r_1,...,r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$, we already know that $p$ falls in all of $r_1,...,r_{k-2}$.
-It will then follow that $r_1,...,r_{k-3}$ all cover $h' \cap v$.
-
-\vgap
-
-As $h'$ is contained in $h$, for proving $h' \cap v = p$, it suffices to explain why $p$ is on $h'$.
+We start by presenting two properties underneath the procedures designed in Section~\ref{sec:hv:type2}. These properties will enable us to construct simpler proofs later.
+
+\begin{proposition} \label{prop:app:hv:type2:help1}
+    Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J2$. Define $h'$ as the trimmed segment of $h$. It holds that $h \cap v = h' \cap v$.
+\end{proposition}
+
+\begin{proof}
+    Set $p = h \cap v$.
+As $h'$ is contained in $h$, for proving $p = h' \cap v$, it suffices to explain why $p$ is on $h'$.
 Let us represent $h$ as $[x_1,x_2] \times y$.
@@
 Therefore, the x-coordinate of $p$ cannot exceed $x_2'$, allowing us to assert that $p$ is on $h'$.
+\end{proof}
+
+We now proceed to elaborate the proofs postponed from Section~\ref{sec:hv:type2}.
+The order of the proofs here will not strictly follow the sequence in which they are referenced in Section~\ref{sec:hv:type2}.
+In particular, we will prove Lemma~\ref{lmm:hv:type2:properties} before Lemma~\ref{lmm:hv:type2:recur-output}, because the claims of the former lemma can be used to construct a simple argument for the latter.
+
+\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type2:properties}.} We will prove each statement in turn.
+
+\extraspacing \underline{\em Proof of Statement (1).} Take any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$.
+The point $h' \cap v$ is covered by all of $r_1,...,r_{k-3}$.
+Let $h$ be the full segment of $h'$.
+As $h$ contains $h'$, we have $h \cap v = h' \cap v$.
+
+\vgap
+
+Consider any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
+From the definition of $\contained_{R_{k-2}}(\bm{t})$, we know that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ --- see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment --- whose right endpoint is $h' \cap v$.
+This means that $h \cap v$ falls in $r_{k-2}$.
+We now have $h \, \cap \, v \, \cap \, \bigcap_{i = 1}^{k-2} r_i$ equals $h' \cap v$ and hence is non-empty, meaning that $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$.
+
+\vgap
+
+To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to explain why $r_{k-2}$ covers the left endpoint of $h$.
+This follows immediately from the fact that $r_{k-2}$ covers the effective horizontal segment of $h'$ (which shares the same left endpoint as $h$).
+
+\extraspacing \underline{\em Proof of Statement (2).} Take any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$.
+Set $p = h \cap v$, and define $h'$ as the trimmed segment of $h$.
+We first argue that the $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v)$ belongs to $\J(R_1,...,R_{k-3}, H', V)$, or equivalently, the point $h' \cap v$ falls in all of $r_1,...,r_{k-3}$.
+By Proposition~\ref{prop:app:hv:type2:help1}, $h' \cap v$ is the same point as $p$.
+From $(r_1,...,r_{k-2},h,v) \in \J(R_1,...,R_{k-2},H,V)$, we know that $p = h \cap v$ falls in all of $r_1,...,r_{k-2}$.
+It thus follows that $r_1,...,r_{k-3}$ all cover $h' \cap v$.
 
@@
 Let $p'$ be the left endpoint of $h$.
-Because we already know $h' \cap v = h \cap v = p$, the effective horizontal segment of $\bm{t}$ is the segment $p'p$.
+Because$h' \cap v = h \cap v = p$, the effective horizontal segment of $\bm{t}$ is the segment $p'p$.
 To prove $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers this segment.

### 2024-06-04 msj.tex
yufei

@@
 \myeqn{
-    H' &=& \{h' \mid h \in H \text{ and its trimmed segment $h'$ exists} \}. \nn
+    H' &=& \{h' \mid h \in H \text{ and its trimmed segment $h'$ exists} \}. \label{eqn:hv:type2:H'}
 }
@@
 \begin{proposition} \label{prop:app:hv:type2:help1}
-    Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J2$. Define $h'$ as the trimmed segment of $h$. It holds that $h \cap v = h' \cap v$.
+    Consider any $k$-tuple $(r_1, ..., r_{k-2}, h, v) \in \J_2$. Define $h'$ as the trimmed segment of $h$. It holds that $h \cap v = h' \cap v$.
 \end{proposition}
@@
 
+\begin{proposition} \label{prop:app:hv:type2:help2}
+    For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$ (where $H'$ is defined in \eqref{eqn:hv:type2:H'}), the set $\contained_{R_{k-2}}(\bm{t})$ is non-empty.
+\end{proposition}
+
+\begin{proof}
+    Let $h$ be the full segment of $\bm{t}[k-2]$.
+By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ covering both endpoints of $h$.
+Thus, the effective horizontal segment of $\bm{t}$ (defined in Section~\ref{sec:hv:type2}), which is contained in $h'$, must also be covered by $r_{k-2}$.
+Hence, $\contained_{R_{k-2}}(\bm{t})$ contains $r_{k-2}$ and thus cannot be empty.
+\end{proof}
+
 We now proceed to elaborate the proofs postponed from Section~\ref{sec:hv:type2}.
@@
 Consider any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
-From the definition of $\contained_{R_{k-2}}(\bm{t})$, we know that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ --- see Section~\ref{sec:hv:type2} for the definition of an effective horizontal segment --- whose right endpoint is $h' \cap v$.
+From the definition of $\contained_{R_{k-2}}(\bm{t})$, we know that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ whose right endpoint is $h' \cap v$.
 This means that $h \cap v$ falls in $r_{k-2}$.
@@
 
-\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} Before proving the lemma, let us prove a fact that will be useful later: the set $\contained_{R_{k-2}}(\bm{t})$ is non-empty for each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$.
-Let $h$ be the full segment of $\bm{t}[k-2]$.
-By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ that covers both endpoints of $h$.
-Thus, the effective horizontal segment of $\bm{t}$, which is contained in $h'$, must also covered by $r_{k-2}$.
-This implies that $\contained_{R_{k-2}}(\bm{t})$ cannot be empty (as it includes at least $r_{k-2}$).
-
-\vgap
-
-With the above fact, next we prove Lemma~\ref{lmm:hv:type2:recur-output} by resorting to the algorithm generate-$\J_2$ (see Section~\ref{sec:hv:type2}).
-For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$, this algorithm adds $|\contained_{R_{k-2}}(\bm{t})|\geq 1$ tuples to $\J_2$.
-Therefore, $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the last inequality used the definition of $\J_2$; see \eqref{eqn:hv:type2:J2}.
+\extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} Again, we resort to the algorithm generate-$\J_2$ .
+For each tuple $\bm{t} \in \J(R_1,...,$ $R_{k-3},$ $H',V)$, we know by Proposition~\ref{prop:app:hv:type2:help2} that $\contained_{R_{k-2}}(\bm{t}) \ne \emptyset$; thus, the algorithm adds at least one tuple to $\J_2$.
+Therefore, $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the derivation used the definition of $\J_2$; see \eqref{eqn:hv:type2:J2}.
 
 \extraspacing {\bf Computing $\bm{H^*}$.} We assume, w.l.o.g., that each segment in the input $H$ is given a distinct integer ID in $[|H|]$.
-This allows us to create an array of size $|H| < n$ and allocate an array cell to each $h \in H$, with the property that the cell can be accessed by the ID of $h$ in constant time.
+This allows us to create an array of size $|H| < n$ and allocate an array cell to each $h \in H$.
+The cell can be accessed by the ID of $h$ in constant time.
 

### 2024-06-04 msj.tex
yufei

@@
 
-Finally, we construct $H^*$ by collecting the minimal segment (see definition in \eqref{eqn:hv:type2:H*}) of every segment $h' \in H'$ with $\minleft(h') \neq \infty$.
-This step takes $O(|H'|) = O(n)$ time.
+Finally, we construct $H^*$ by collecting the minimal segment (defined in \eqref{eqn:hv:type2:H*}) of every segment $h' \in H'$ with $\minleft(h') \neq \infty$. This step takes $O(|H'|) = O(n)$ time.
 

### 2024-06-04 msj.tex
yufei

@@
 
-\extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\contained_{R_{k-2}}(\bm{t}) = \contained_{R_{k-2}}(h^*)$.
-We can then derive
+\extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\contained_{R_{k-2}}(h^*) = \contained_{R_{k-2}}(\bm{t})$.
+The mapping allows us to derive
 \myeqn{
@@
     \leq \out \nn
-} where the last inequality is due to Statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
-
-\vgap
-
-Now, we explain the mapping.
-Consider an arbitrary $h^* \in H^*$.
+}
+where the last inequality used Statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+
+\vgap
+
+The mapping is as follows.
+Consider any $h^* \in H^*$.
 Recall that $h^*$ is the minimal segment of some segment $h' \in H'$.
 Specifically, if $h' = [x_1,x_2]\times y$, then $h^* = [x_1,\minleft(h')]\times y$.
-By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$}$ (recall that $\bm{t}[k-1]$ is a vertical segment).
+By the definition of $\minleft(h')$ in \eqref{eqn:hv:type2:H*}, there exists a tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$ satisfying $\bm{t}[k-2] = h'$ and $\minleft(h') = \text{$x$-coordinate of the vertical segment $\bm{t}[k-1]$}$.
 We map $h^*$ to $\bm{t}$.
@@
 
-Next, we will prove that a rectangle $r_{k-2} \in \contained_{R_{k-2}}(h^*)$ if and only if $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
+Next, we will prove
+\begin{center}
+    {\bf Claim 1:} a rectangle $r \in \contained_{R_{k-2}}(h^*)$ if and only if $r \in \contained_{R_{k-2}}(\bm{t})$.
+\end{center}
 According to the definitions of $\contained_{R_{k-2}}(h^*)$ and $\contained_{R_{k-2}}(\bm{t})$ --- see \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
-Denote by $x$ the x-coordinate of the vertical segment $\bm{t}[k-1]$.
-Then, the point $h' \cap v$ (intersection between $h'$ and $v$) is $(x,y)$, where as mentioned $y$ is the y-coordinate of $h'$.
+
+\vgap
+
+Denote by $x$ the x-coordinate of $\bm{t}[k-1]$.
+Then, the point $h' \cap v$ can be written as $(x,y)$, where as mentioned $y$ is the y-coordinate of $h'$.
 The effective horizontal segment of $\bm{t}$ is therefore $[x_1,x]\times y$.
-Recall that $h^* = [x_1,\minleft(h')]\times y$, where $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$} = x$.
-Therefore, $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
+Recall that $h^* = [x_1,\minleft(h')]\times y$, where (by definition of $\bm{t}$) $\minleft(h') = \text{$x$-coordinate of $\bm{t}[k-1]$} = x$.
+Therefore, $h^*$ is the same as the effective horizontal segment of $\bm{t}$, proving claim 1.
 
@@
 It remains to show that no two distinct segments $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$.
-Assume, on the contrary, that $h_1^*$ and $h_2^*$ are mapped to the same tuple $\bm{t}\in \J(R_1,...,R_{k-3}, H',V)$.
+Assume, on the contrary, that $h_1^*$ and $h_2^*$ are both mapped to $\bm{t}\in \J(R_1,...,R_{k-3}, H',V)$.
 Suppose that $h_1^*$ (resp., $h_2^*$) is the minimal segment of $h_1'$ (resp., $h_2'$).
-Under our mapping, it must be true that  $h_1' = h_2' = \bm{t}[k-2]$.
+Under our mapping, it must hold that  $h_1' = h_2' = \bm{t}[k-2]$.
 However, the distinctness of $h_1^*$ and $h_2^*$ requires $h_1' \ne h_2'$, which yields a contradiction.
@@
 The effective horizontal segment of $\bm{t}$ can then be represented as $[x_1,x] \times y$.
-
-%Then, $B_\bm{t}$, which is the intersection of $h'$ and $v$, is $(x,y)$.
 

### 2024-06-04 msj.tex
yufei

@@
 
-It remains to show that no two distinct segments $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$.
-Assume, on the contrary, that $h_1^*$ and $h_2^*$ are both mapped to $\bm{t}\in \J(R_1,...,R_{k-3}, H',V)$.
+It remains to show that no two distinct $h_1^*, h_2^* \in H^*$ can be mapped to the same tuple in $\J(R_1,...,R_{k-3}, H',V)$.
+Assume, on the contrary, that $h_1^*$ and $h_2^*$ are both mapped to $\bm{t}\in \J(R_1,...,R_{k-3},$ $H',V)$.
 Suppose that $h_1^*$ (resp., $h_2^*$) is the minimal segment of $h_1'$ (resp., $h_2'$).
@@
 
-\extraspacing \underline{\em Proof of Statement (2).} Let $h', h^*$, and $\bm{t}$ be as defined in statement (2).
-Let us represent $h'$ as $[x_1,x_2] \times y$; accordingly, $h^*$ can be written as $[x_1, \minleft(h')] \times y$.
+\extraspacing \underline{\em Proof of Statement (2).} Take any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-3}, H', V)$.
+Set $h' = \bm{t}[k-2]$, and define $h^*$ as the minimal segment of $h'$.
+Let us write $h'$ as $[x_1,x_2] \times y$; accordingly, $h^*$ can be written as $[x_1, \minleft(h')] \times y$.
 Additionally, let $x$ be the x-coordinate of the vertical segment $\bm{t}[k-1]$.
@@
 
-Firstly, we will prove $\contained_{R_{k-2}}(\bm{t})\subseteq \contained_{R_{k-2}}(h^*)$.
-For this purpose, we will show that any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ also covers $h^*$.
-It suffices to show that $h^*$ is contained in the effective horizontal segment of $\bm{t}$.
-Combining the definition of $\minleft(h')$ (see \eqref{eqn:minleft}), and the fact that $x$ is the x-coordinate of $\bm{t}[k-1]$, we have $\minleft(h') \le x$.
-This implies that $h^* = [x_1, \minleft(h')] \times y$ is contained in $[x_1, x] \times y$, i.e.
-the effective horizontal segment of $\bm{t}$.
+We will first prove $\contained_{R_{k-2}}(\bm{t})\subseteq \contained_{R_{k-2}}(h^*)$, or equivalently, every rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ covers $h^*$.
+It suffices to show that $h^*$ is contained in the effective horizontal segment of $\bm{t}$, or equivalently, $\minleft(h') \le x$.
+By the definition in \eqref{eqn:minleft}, $\minleft(h')$ is the minimum x-coordinate of $\bm{t'}[k-1]$ among all $\bm{t'} \in  \J(R'_1, ..., R'_{k-3}, H', V)$ with $\bm{t'}[k-2] = h'$, Thus, $\minleft(h') \le x$ holds because $\bm{t}$ is merely one such $\bm{t'}$.
 

### 2024-06-04 msj.tex
yufei

@@
 
-Secondly, assuming that the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ have been sorted in descending order of $\xright(r)$, we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
-It suffices to establish the following equivalent statement:
+Next, assuming that the rectangles $r \in \contained_{R_{k-2}}(h^*)$ have been sorted in descending order of $\xright(r)$, we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
+It suffices to establish the following equivalent claim:
 
@@
 \minipg{0.8\linewidth} {
-    If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies the condition that $\xright(r) \ge x$ (where as defined before $x$ is the x-coordinate of $\bm{t}[k-1]$), then $r$ must be in $\contained_{R_{k-2}}(\bm{t})$.
+    {\bf Claim 2:} If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies the condition that $\xright(r) \ge x$ (where $x$ is the x-coordinate of $\bm{t}[k-1]$), then $r \in \contained_{R_{k-2}}(\bm{t})$.
 }
@@
 Since $r \in \contained_{R_{k-2}}(h^*)$, we know that $r$ covers the segment $[x_1, \minleft(h')] \times y$.
-Therefore, $[x_1, \minleft(h')] \subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
-Using also the given condition $\xright(r) \ge x$, we can derive $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
-This means that $[x_1, x] \times y$, i.e.
+Hence, $[x_1, \minleft(h')]$ $\subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.
+Using also the condition $\xright(r) \ge x$ given in claim 2, we can derive $[x_1, x] \subseteq [\xleft(r), \xright(r)]$.
+Therefore, $[x_1, x] \times y$, i.e.
 the effective horizontal segment of $\bm{t}$, is contained in $r$.

### 2024-06-04 msj.tex

@@
 
-\noindent To prove the above, we must explain why $r$ contains the effective horizontal segment of $\bm{t}$.
+\noindent To prove the above, we must explain why the rectangle $r$ in the claim contains the effective horizontal segment of $\bm{t}$.
 Since $r \in \contained_{R_{k-2}}(h^*)$, we know that $r$ covers the segment $[x_1, \minleft(h')] \times y$.
@@
 Therefore, $[x_1, x] \times y$, i.e.
-the effective horizontal segment of $\bm{t}$, is contained in $r$.
+the effective horizontal segment of $\bm{t}$, is contained in $r$, establishing claim 2.
 

### 2024-06-06 msj.tex
yufei

@@
 In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extends the conventional equi-join in two ways.
-First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with a ``non-empty intersection'' predicate on the attributes involved.
+First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in an equi-join is replaced with a ``non-empty intersection'' predicate on the attributes involved.
 The $k$-SJ problem can be converted to a join under the framework of \cite{kcko22} as defined next.
@@
 The theorem implies a recursive nature of $k$-SJ.
-Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + k \cdot \out)$.
+Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional cost of $O(n \log n + k \cdot \out)$ per instance.
 For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.

### 2024-06-06 msj.tex
yufei

@@
 \def\dcross{\mathrm{d\text{-}cross}}
-\def\contained{\mathrm{contained}}
+\def\contain{\mathrm{contain}}
 \def\cat{\mathrm{cat1}}
@@
 
-Given a horizontal segment $h = [x_1, x_2] \times y$, we say that a rectangle $r$ is a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
-A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s \cap r \ne \emptyset$ but $r$ covers neither of the two endpoints of $s$.
-A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.
+Given a horizontal segment $h = [x_1, x_2] \times y$, we say that a rectangle $r$ is a {\em left-end covering rectangle} of $h$ if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$).
+A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$ if $s \cap r \ne \emptyset$ but $r$ covers neither of the two endpoints of $s$.
+A rectangle $r$ {\em contains} a horizontal/vertical segment $s$ if $r$ covers both endpoints of $s$.
 
@@
 \myeqn{
-    \contained_R(h) = \{r \in R \mid \text{$r$ contains $h$} \};
+    \contain_R(h) = \{r \in R \mid \text{$r$ contains $h$} \};
     \label{eqn:contained}
 }
-namely, $\contained_R(h)$ is the set of rectangles in $R$ containing $h$.
+namely, $\contain_R(h)$ is the set of rectangles in $R$ containing $h$.
 
@@
 \extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
-The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries.
-Formally, if $r_1, r_2, ..., r_z$ for some $z \ge 1$ are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
+The goal is to report, for each segment $h \in H$, the set $\contain_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries.
+Formally, if $r_1, r_2, ..., r_z$ for some $z \ge 1$ are all the rectangles in $\contain_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
 Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3), (h_3: \emptyset)}$.
@@
 \myeqn{
-    \contained_{R_{k-2}}(\bm{t}) = \{r \in R_{k-2} \mid \text{$r$ contains the effective horizontal segment of $\bm{t}$}   \}
+    \contain_{R_{k-2}}(\bm{t}) = \{r \in R_{k-2} \mid \text{$r$ contains the effective horizontal segment of $\bm{t}$}   \}
     \label{eqn:type2:contained-t}
 }
-The above should not be confused with \eqref{eqn:contained}, where the ``$\contained$'' function takes a segment as the parameter, rather than a tuple.
+The above should not be confused with \eqref{eqn:contained}, where the ``$\contain$'' function takes a segment as the parameter, rather than a tuple.
 
@@
 Accordingly, as can be seen from Figure~\ref{fig:hv:type2:ex}c,
-$\contained_{R_{k-2}}(\bm{t}_1) = \contained_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contained_{R_2}(\bm{t}_3) = \set{\beta_3}$.
+$\contain_{R_{k-2}}(\bm{t}_1) = \contain_{R_2}(\bm{t}_1) = \set{\beta_1}$, $\contain_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, and $\contain_{R_2}(\bm{t}_3) = \set{\beta_3}$.
 \done
@@
 Denote by $h$ the full segment of $\bm{t}[k-2]$.
-Then, for any $r \in \contained_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ must belong to $\J_2$.
+Then, for any $r \in \contain_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ must belong to $\J_2$.
 
@@
 Let $h'$ be the trimmed segment of $h$ and set $\bm{t} = (r_1, ..., r_{k-3}, h', v)$.
-Then, $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ and $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
-
-        \item $\sum_{\bm{t}} |\contained_{R_{k-2}}(\bm{t})| \le \out$, where the summation is over all $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$.
+Then, $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ and $r_{k-2} \in \contain_{R_{k-2}}(\bm{t})$.
+
+        \item $\sum_{\bm{t}} |\contain_{R_{k-2}}(\bm{t})| \le \out$, where the summation is over all $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$.
     }
@@
     \> 3.\>\> $h \leftarrow$ the full segment of $\bm{t}[k-2]$ \\
-    \> 4.\>\> {\bf for} each $r \in \contained_{R_{k-2}}(\bm{t})$ {\bf do} \\
+    \> 4.\>\> {\bf for} each $r \in \contain_{R_{k-2}}(\bm{t})$ {\bf do} \\
     \> 5.\>\>\> add $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ to $\J_2$
@@
 The correctness of the algorithm follows from statements (1) and (2) of Lemma~\ref{lmm:hv:type2:properties}.
-Furthermore, if we are given $\contained_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of Lemma~\ref{lmm:hv:type2:properties}
-assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+Furthermore, if we are given $\contain_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of Lemma~\ref{lmm:hv:type2:properties}
+assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contain_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 
 \extraspacing {\underline{\em Example 5.4.}} For $\bm{t}_1 = (\alpha, h_2', v_1)$, the full segment of $h_2'$ is $h_2$.
-As $\beta_1$ is the only rectangle in $\contained_{R_2}(\bm{t}_1)$, Line 5 of the algorithm adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$.
+As $\beta_1$ is the only rectangle in $\contain_{R_2}(\bm{t}_1)$, Line 5 of the algorithm adds tuple $(\alpha, \beta_1, h_2, v_1)$ to $\J_2$.
 For $\bm{t}_2 = (\alpha, h_3', v_1)$, the full segment of $h_3'$ is $h_3$.
-As $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, Line 5 adds $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_1)$ to $\J_2$.
+As $\contain_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$, Line 5 adds $(\alpha, \beta_1, h_3, v_1)$, $(\alpha, \beta_2, h_3, v_1)$, and $(\alpha, \beta_3, h_3, v_1)$ to $\J_2$.
 Finally, the processing of $\bm{t}_3 = (\alpha, h_3', v_2)$ adds $(\alpha, \beta_3, h_3, v_2)$ to $\J_2$.
@@
 
-We are interested in the sets $\contained_{R_{k-2}}(h^*)$ of the segments $h^*$ in $H^*$, where $\contained_{R_{k-2}}(h^*)$ --- defined in \eqref{eqn:contained} --- is the set of rectangles in $R_{k-2}$ containing $h^*$.
+We are interested in the sets $\contain_{R_{k-2}}(h^*)$ of the segments $h^*$ in $H^*$, where $\contain_{R_{k-2}}(h^*)$ --- defined in \eqref{eqn:contained} --- is the set of rectangles in $R_{k-2}$ containing $h^*$.
 These sets have some useful properties:
@@
     \myenums{
-        \item $\sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)| \le \out$.
+        \item $\sum_{h^* \in H^*} |\contain_{R_{k-2}}(h^*)| \le \out$.
 
@@
 Set $h' = \bm{t}[k-2]$ and let $h^*$ be the minimal segment of $h'$.
-Then, $$\contained_{R_{k-2}}(\bm{t}) \subseteq \contained_{R_{k-2}}(h^*).$$ Furthermore, if the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
+Then, $$\contain_{R_{k-2}}(\bm{t}) \subseteq \contain_{R_{k-2}}(h^*).$$ Furthermore, if the rectangles $r$ in $\contain_{R_{k-2}}(h^*)$ are sorted in descending order of $\xright(r)$, then $\contain_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
     }
@@
 
-\extraspacing {\underline{\em Example 5.6.}} It is clear from Figure~\ref{fig:hv:type2:ex}(d) that $\contained_{R_2}(h_3^*)$ has rectangles $\beta_3, \beta_2$, $\beta_1$, sorted in descending order of their right boundaries' x-coordinates.
+\extraspacing {\underline{\em Example 5.6.}} It is clear from Figure~\ref{fig:hv:type2:ex}(d) that $\contain_{R_2}(h_3^*)$ has rectangles $\beta_3, \beta_2$, $\beta_1$, sorted in descending order of their right boundaries' x-coordinates.
 Consider $\bm{t}_2 = (\alpha, h_3', v_1)$ and $\bm{t}_3 = (\alpha, h_3', v_2)$.
 Segment $h_3^*$ is the minimal segment of $h_3'$.
-As stated in Lemma~\ref{lmm:hv:type2:contained-h*}, both $\contained_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$ and $\contained_{R_2}(\bm{t}_3) = \set{\beta_1}$ are prefixes of the sorted order of $\contained_{R_2}(h_3^*)$.
-Regarding $h_2^*$, it is the minimal segment of $h_2'$, and $\contained_{R_2}(h_2^*)$ contains only  $\beta_1$.
-For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contained_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contained_{R_2}(h_2^*)$, as is also consistent with the lemma.
+As stated in Lemma~\ref{lmm:hv:type2:contained-h*}, both $\contain_{R_2}(\bm{t}_2) = \set{\beta_1, \beta_2, \beta_3}$ and $\contain_{R_2}(\bm{t}_3) = \set{\beta_1}$ are prefixes of the sorted order of $\contain_{R_2}(h_3^*)$.
+Regarding $h_2^*$, it is the minimal segment of $h_2'$, and $\contain_{R_2}(h_2^*)$ contains only  $\beta_1$.
+For $\bm{t}_1 = (\alpha, h_2', v_1)$, $\contain_{R_2}(\bm{t}_1) = \set{\beta_1}$ is a (trivial) prefix of $\contain_{R_2}(h_2^*)$, as is also consistent with the lemma.
 \done
 
-Finding the $\contained_{R_{k-2}}(h^*)$ sets of all $h^* \in H^*$ is an instance of the find-all-sorted version of Problem $\mathscr{E}$ (with $H^*$ and $R_{k-2}$ as the input).
+Finding the $\contain_{R_{k-2}}(h^*)$ sets of all $h^* \in H^*$ is an instance of the find-all-sorted version of Problem $\mathscr{E}$ (with $H^*$ and $R_{k-2}$ as the input).
 The cost is $O(n \log n + \out)$ according statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*} and the discussion in Section~\ref{sec:bricks}.
-Note that, for each $\contained_{R_{k-2}}(h^*)$ computed, the rectangles $r$ therein have been sorted in descending order of $\xright(r)$.
-
-\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*} allows us to produce $\contained_{R_{k-2}}(\bm{t})$ --- defined in \eqref{eqn:type2:contained-t} --- for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ as follows.
+Note that, for each $\contain_{R_{k-2}}(h^*)$ computed, the rectangles $r$ therein have been sorted in descending order of $\xright(r)$.
+
+\extraspacing {\bf Computing the ``$\bm{\contain_{R_{k-2}}(t)}$'' Sets.} Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*} allows us to produce $\contain_{R_{k-2}}(\bm{t})$ --- defined in \eqref{eqn:type2:contained-t} --- for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ as follows.
 First, fetch the (already computed) minimal segment $h^*$ of $\bm{t}[k-2]$ in $O(1)$ time.
-Then, scan the rectangles $r$ of $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$.
-For each $r$ scanned, check whether $r \in \contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers $B_\bm{t}$ (recall that $B_\bm{t}$ is a point); the cost of this inspection is $O(1)$.
+Then, scan the rectangles $r$ of $\contain_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$.
+For each $r$ scanned, check whether $r \in \contain_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers $B_\bm{t}$ (recall that $B_\bm{t}$ is a point); the cost of this inspection is $O(1)$.
 Abort the scan as soon as $r \notin
-\contained_{R_{k-2}}(\bm{t})$.
-This way, $\contained_{R_{k-2}}(t)$ can be decided in $O(k + |\contained_{R_{k-2}}(\bm{t})|)$ time.
-Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3},$ $H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(\bm{t})|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+\contain_{R_{k-2}}(\bm{t})$.
+This way, $\contain_{R_{k-2}}(t)$ can be decided in $O(k + |\contain_{R_{k-2}}(\bm{t})|)$ time.
+Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3},$ $H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contain_{R_{k-2}}(\bm{t})|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 
@@
 Next, for each $j \in [k-1]$, solve the reporting version of Problem $\mathscr{A}$ by feeding $P'$ and $R_j$ as the input.
-This produces the set $\contained_{R_j}(p)$ for each point $p \in P'$, where $\contained_{R_j}(p)$ is defined in \eqref{eqn:contained} (treating $p$ as a degenerated ``horizontal segment'') and includes all rectangles of $R_j$ covering $p$.
+This produces the set $\contain_{R_j}(p)$ for each point $p \in P'$, where $\contain_{R_j}(p)$ is defined in \eqref{eqn:contained} (treating $p$ as a degenerated ``horizontal segment'') and includes all rectangles of $R_j$ covering $p$.
 By the discussion in Section~\ref{sec:bricks}, the total cost of this step is bounded by
 \myeqn{
-    O\Big(kn \log n + \sum_{p \in P'} \sum_{j \in [k-1]} |\contained_{R_j}(p)| \Big). \label{eqn:ksj:cat1:help1}
+    O\Big(kn \log n + \sum_{p \in P'} \sum_{j \in [k-1]} |\contain_{R_j}(p)| \Big). \label{eqn:ksj:cat1:help1}
 }
@@
 \myeqn{
-    (r_1, ..., r_{k-1}) \in \contained_{R_1}(p) \times ... \times \contained_{R_{k-1}}(p) \nn
+    (r_1, ..., r_{k-1}) \in \contain_{R_1}(p) \times ... \times \contain_{R_{k-1}}(p) \nn
     %\label{eqn:ksj:cat1:help2}
@@
 Performing the above for all $p \in P'$ generates the whole $\J^\cat_k$ in $O(1 + k \cdot|\J^\cat_k|)$ time.
-The way $P'$ is computed ensures that $\contained_{R_j}(p) \ne \emptyset$ for each $j \in [k-1]$.
-Hence, $\sum_{j \in [k-1]} |\contained_{R_j}(p)| \le \prod_{j \in [k-1]} |\contained_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|) = O(kn \log n + k \cdot \out)$ .
+The way $P'$ is computed ensures that $\contain_{R_j}(p) \ne \emptyset$ for each $j \in [k-1]$.
+Hence, $\sum_{j \in [k-1]} |\contain_{R_j}(p)| \le \prod_{j \in [k-1]} |\contain_{R_j}(p)|$, which implies that \eqref{eqn:ksj:cat1:help1} is bounded by $O(kn \log n + k \cdot |\J^\cat_k|) = O(kn \log n + k \cdot \out)$ .
 
@@
 All these queries demand a total cost of $O(n \log n + \sum_h K_h)$.
-The sum $\sum_h K_h$ here is equal to the total size of $\contained_R(h)$ for all the $h \in H$.
+The sum $\sum_h K_h$ here is equal to the total size of $\contain_R(h)$ for all the $h \in H$.
 The total size is equivalent to $\out$.
@@
 
-Finally, we generate, for each $h$, its set $\contained_R(h)$ (i.e., the set of rectangles covering $h$), where the rectangles $r$ are sorted by $\xright(r)$.
+Finally, we generate, for each $h$, its set $\contain_R(h)$ (i.e., the set of rectangles covering $h$), where the rectangles $r$ are sorted by $\xright(r)$.
 To start with, initialize an empty linked list for every $h \in H$.
@@
 \begin{proposition} \label{prop:app:hv:type2:help2}
-    For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$ (where $H'$ is defined in \eqref{eqn:hv:type2:H'}), the set $\contained_{R_{k-2}}(\bm{t})$ is non-empty.
+    For each tuple $\bm{t} \in \J(R_1,...,R_{k-3},H',V)$ (where $H'$ is defined in \eqref{eqn:hv:type2:H'}), the set $\contain_{R_{k-2}}(\bm{t})$ is non-empty.
 \end{proposition}
@@
 Thus, the effective horizontal segment of $\bm{t}$ (defined in Section~\ref{sec:hv:type2}), which is contained in $h'$, must also be covered by $r_{k-2}$.
-Hence, $\contained_{R_{k-2}}(\bm{t})$ contains $r_{k-2}$ and thus cannot be empty.
+Hence, $\contain_{R_{k-2}}(\bm{t})$ contains $r_{k-2}$ and thus cannot be empty.
 \end{proof}
@@
 
-Consider any rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$.
-From the definition of $\contained_{R_{k-2}}(\bm{t})$, we know that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ whose right endpoint is $h' \cap v$.
+Consider any rectangle $r_{k-2} \in \contain_{R_{k-2}}(\bm{t})$.
+From the definition of $\contain_{R_{k-2}}(\bm{t})$, we know that $r_{k-2}$ covers the effective horizontal segment of $\bm{t}$ whose right endpoint is $h' \cap v$.
 This means that $h \cap v$ falls in $r_{k-2}$.
@@
 
-Next, we argue that $r_{k-2}\in \contained_{R_{k-2}}(\bm{t})$.
+Next, we argue that $r_{k-2}\in \contain_{R_{k-2}}(\bm{t})$.
 Let $p'$ be the left endpoint of $h$.
 Because$h' \cap v = h \cap v = p$, the effective horizontal segment of $\bm{t}$ is the segment $p'p$.
-To prove $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers this segment.
+To prove $r_{k-2} \in \contain_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers this segment.
 This is true because:
@@
 \myeqn{
-    \sum_{\bm{t} \in \J(R_1, ..., R_3, H', V)} |\contained_{R_{k-2}}(\bm{t})| \nn
+    \sum_{\bm{t} \in \J(R_1, ..., R_3, H', V)} |\contain_{R_{k-2}}(\bm{t})| \nn
 }
@@
 \extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type2:recur-output}.} Again, we resort to the algorithm generate-$\J_2$ .
-For each tuple $\bm{t} \in \J(R_1,...,$ $R_{k-3},$ $H',V)$, we know by Proposition~\ref{prop:app:hv:type2:help2} that $\contained_{R_{k-2}}(\bm{t}) \ne \emptyset$; thus, the algorithm adds at least one tuple to $\J_2$.
+For each tuple $\bm{t} \in \J(R_1,...,$ $R_{k-3},$ $H',V)$, we know by Proposition~\ref{prop:app:hv:type2:help2} that $\contain_{R_{k-2}}(\bm{t}) \ne \emptyset$; thus, the algorithm adds at least one tuple to $\J_2$.
 Therefore, $|\J(R_1,...,R_{k-3},H',V)| \le |\J_2| \le \out$, where the derivation used the definition of $\J_2$; see \eqref{eqn:hv:type2:J2}.
@@
 
-\extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\contained_{R_{k-2}}(h^*) = \contained_{R_{k-2}}(\bm{t})$.
+\extraspacing \underline{\em Proof of Statement (1).} We will map each segment $h^* \in H^*$ to a unique tuple $\bm{t}\in \J(R_1,...,R_{k-3},H',V)$ satisfying $\contain_{R_{k-2}}(h^*) = \contain_{R_{k-2}}(\bm{t})$.
 The mapping allows us to derive
 \myeqn{
-    \sum_{h^* \in H^*} |\contained_{R_{k-2}}(h^*)|
+    \sum_{h^* \in H^*} |\contain_{R_{k-2}}(h^*)|
     \le
-    \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\contained_{R_{k-2}}(\bm{t})|
+    \sum_{\bm{t} \in \J(R_1,...,R_{k-3},H',V)} |\contain_{R_{k-2}}(\bm{t})|
     \leq \out \nn
@@
 \begin{center}
-    {\bf Claim 1:} a rectangle $r \in \contained_{R_{k-2}}(h^*)$ if and only if $r \in \contained_{R_{k-2}}(\bm{t})$.
+    {\bf Claim 1:} a rectangle $r \in \contain_{R_{k-2}}(h^*)$ if and only if $r \in \contain_{R_{k-2}}(\bm{t})$.
 \end{center}
-According to the definitions of $\contained_{R_{k-2}}(h^*)$ and $\contained_{R_{k-2}}(\bm{t})$ --- see \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
+According to the definitions of $\contain_{R_{k-2}}(h^*)$ and $\contain_{R_{k-2}}(\bm{t})$ --- see \eqref{eqn:contained} and \eqref{eqn:type2:contained-t}, respectively --- it suffices to show that $h^*$ is the same as the effective horizontal segment of $\bm{t}$.
 
@@
 
-We will first prove $\contained_{R_{k-2}}(\bm{t})\subseteq \contained_{R_{k-2}}(h^*)$, or equivalently, every rectangle $r_{k-2} \in \contained_{R_{k-2}}(\bm{t})$ covers $h^*$.
+We will first prove $\contain_{R_{k-2}}(\bm{t})\subseteq \contain_{R_{k-2}}(h^*)$, or equivalently, every rectangle $r_{k-2} \in \contain_{R_{k-2}}(\bm{t})$ covers $h^*$.
 It suffices to show that $h^*$ is contained in the effective horizontal segment of $\bm{t}$, or equivalently, $\minleft(h') \le x$.
@@
 
-Next, assuming that the rectangles $r \in \contained_{R_{k-2}}(h^*)$ have been sorted in descending order of $\xright(r)$, we will prove that $\contained_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
+Next, assuming that the rectangles $r \in \contain_{R_{k-2}}(h^*)$ have been sorted in descending order of $\xright(r)$, we will prove that $\contain_{R_{k-2}}(\bm{t})$ includes a prefix of the sorted order.
 It suffices to establish the following equivalent claim:
@@
 \minipg{0.8\linewidth} {
-    {\bf Claim 2:} If a rectangle $r \in \contained_{R_{k-2}}(h^*)$ satisfies the condition that $\xright(r) \ge x$ (where $x$ is the x-coordinate of $\bm{t}[k-1]$), then $r \in \contained_{R_{k-2}}(\bm{t})$.
+    {\bf Claim 2:} If a rectangle $r \in \contain_{R_{k-2}}(h^*)$ satisfies the condition that $\xright(r) \ge x$ (where $x$ is the x-coordinate of $\bm{t}[k-1]$), then $r \in \contain_{R_{k-2}}(\bm{t})$.
 }
@@
 \noindent To prove the above, we must explain why the rectangle $r$ in the claim contains the effective horizontal segment of $\bm{t}$.
-Since $r \in \contained_{R_{k-2}}(h^*)$, we know that $r$ covers the segment $[x_1, \minleft(h')] \times y$.
+Since $r \in \contain_{R_{k-2}}(h^*)$, we know that $r$ covers the segment $[x_1, \minleft(h')] \times y$.
 Hence, $[x_1, \minleft(h')]$ $\subseteq [\xleft(r), \xright(r)]$ and $y \in [\ybot(r), \ytop(r)]$.

### 2024-06-06 msj.tex
yufei

@@
 
-In the {\em reporting version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$.
-In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_1, \emptyset), (p_2: r_1, r_2), (p_3: r_2)}$.
+In the {\em reporting version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$; if no such $r$ exists, report nothing for $p$.
+In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$.
 As shown in Appendix~\ref{app:bricks}, The problem can be solved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.

### 2024-06-06 msj.tex
yufei

@@
 \extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles.
-The goal is to report, for each segment $h \in H$, the set $\contain_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries.
+The goal is to report, for each segment $h \in H$, the set $\contain_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries; if $\contain_R(h)$ is empty, output nothing for $h$.
 Formally, if $r_1, r_2, ..., r_z$ for some $z \ge 1$ are all the rectangles in $\contain_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
-Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3), (h_3: \emptyset)}$.
+Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

### 2024-06-06 msj.tex
yufei

@@
 \extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$ (the solid rectangle), $R_2 = \set{\beta_2, \beta_2}$ (the dashed rectangles), $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
-The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_6)$.
+The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_5)$.
 \done

### 2024-06-06 msj.tex
yufei

@@
 The theorem implies a recursive nature of $k$-SJ.
-Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional cost of $O(n \log n + k \cdot \out)$ per instance.
+Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional cost of $O(k^3) \cdot (n \log n + k \cdot \out)$.
 For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$.

### 2024-06-06 msj.tex
yufei

@@
 Computing the ``segment $h_\bot$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input).
-By symmetry, so is the computation of $h_\top, v_\vdash$, and $v_\dashv$ segments for each $r \in R_i$.
+By symmetry, so is computing the $h_\top, v_\vdash$, and $v_\dashv$ segments for each $r \in R_i$.
 It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(kn \log n)$ total time.
@@
 By statements (1) and (2) of Lemma~\ref{lmm:hv:type1:properties}, the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}.
-Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, the above algorithm runs in $O(1 + |\J(R'_1, ..., R'_{k-2})| + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time, where the derivation used \eqref{eqn:hv:type1:J*} and Lemma~\ref{lmm:hv:type1:recur-output}.
+Furthermore, if we are given $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ for each $\bm{t}$, the above algorithm runs in $O(1 + k \cdot |\J(R'_1, ..., R'_{k-2})| + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time, where the derivation used \eqref{eqn:hv:type1:J*} and Lemma~\ref{lmm:hv:type1:recur-output}.
 

### 2024-06-06 msj.tex
yufei

@@
 For instance, $\alpha'$, which is trimmed from rectangle $\alpha$, is decided by $h_\bot$ $=$ $h_2$, $h_\top$ $=$ $h_6$, $v_\vdash = v_3$, and $v_\dashv = v_5$.
-The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2, ..., R'_{k-2}) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
+The $(k-2)$-SJ on $R'_1$ and $R'_2$ returns $\J(R_1', R'_2) = \set{(\alpha', \beta_1'), (\alpha', \beta_2')}$.
 \done
@@
 Note that $B_\bm{t}$ --- defined in \eqref{eqn:B_t} --- is the point $h' \cap v$ (the intersection of $h'$ and $v$).
-Suppose $h' = [x_1, x_2] \times y$ and $B_\bm{t} = (x, y)$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $[x_1, x] \times y$.
+Suppose that $h' = [x_1, x_2] \times y$ and $B_\bm{t} = (x, y)$; we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $[x_1, x] \times y$.
 This allows us to define
@@
 Denote by $h$ the full segment of $\bm{t}[k-2]$.
-Then, for any $r \in \contain_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ must belong to $\J_2$.
+Then, for any $r \in \contain_{R_{k-2}}(\bm{t})$, the $k$-tuple $(\bm{t}[1], ..., \bm{t}[k-3], r, h, \bm{t}[k-1])$ belongs to $\J_2$.
 

### 2024-06-06 msj.tex
yufei

@@
 
-Finding the $\contain_{R_{k-2}}(h^*)$ sets of all $h^* \in H^*$ is an instance of the find-all-sorted version of Problem $\mathscr{E}$ (with $H^*$ and $R_{k-2}$ as the input).
+Finding the $\contain_{R_{k-2}}(h^*)$ sets of all $h^* \in H^*$ is an instance of Problem $\mathscr{E}$ (with $H^*$ and $R_{k-2}$ as the input).
 The cost is $O(n \log n + \out)$ according statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*} and the discussion in Section~\ref{sec:bricks}.

### 2024-06-06 msj.tex
yufei

@@
 At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.
-The algorithm must produce the correct output on all $(n_1, n_2, ..., n_k)$-inputs.
+The algorithm is required to produce the correct output on all $(n_1, n_2, ..., n_k)$-inputs.
 Its {\em worst-case cost} is the height of $\T$.
@@
 As can be verified shortly, $h$ is the lowest segment in $H$ that crosses $r$; we therefore output $(r, h)$.
-After that, $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time.
+After that, $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $\T$ in $O(\log n)$ time.
 As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future.

### 2024-06-06 msj.tex
yufei

@@
 Performing the above for all $i, j \in [k]$ with $i \ne j$ leads to a total time complexity of $O(k^3) \cdot (F_{k-1}(n, \out) + n \log n + k \cdot \out)$.
-A category-2 result tuple $\bm{t}$ may be seen more than once (this can happen if, for example, more than one rectangle in $\bm{t}$ has the same top-edge or the same left-edge).
+A category-2 result tuple $\bm{t}$ may be seen more than once (this can happen if, for example, more than one rectangle in $\bm{t}$ has the same top-edge).
 Again, duplicate removal can be achieved at no extra cost asymptotically.
@@
 Fix any $h \in \dcross_H(\bm{t})$ and any $v \in \dcross_V(\bm{t})$.
-The segments $h$ and $v$ both cross $B_\bm{t}$, as can be verified from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
+The segments $h$ and $v$ both cross $B_\bm{t}$, as can be seen directly from the definitions of $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$.
 Thus, $h \cap v$ is a point in $B_\bm{t} = \bigcap_{i=1}^{k-2} \bm{t}[i]$.
@@
 \extraspacing \noindent {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.} Our proof again resorts to the algorithm generate-$\J^*$ in Section~\ref{sec:hv:type1}.
-As argued in Statement (3) of Lemma~\ref{lmm:hv:type1:properties}, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ are non-empty for any $\bm{t} \in \J(R_1',$ $...,R_{k-2}')$.
+By Proposition~\ref{prop:app:hv:type1:help2}, both $\dcross_H(\bm{t})$ and $\dcross_V(\bm{t})$ are non-empty for any $\bm{t} \in \J(R_1',$ $...,R_{k-2}')$.
 Therefore, in processing this $\bm{t}$, the algorithm adds at least one new tuple to $\J^*$.
@@
 Let us take the y-coordinate $y_h$ of $h$.
-We will prove that the point $p = (\xleft(B_{\bm{t}}), y_h)$ is on the segment $h \cap B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$.
+We will prove that the point $p = (\xleft(B_{\bm{t}}), y_h)$ is on the segment $h$ and is also in $B_{\bm{t}}$, suggesting $h\cap B_{\bm{t}} \neq \emptyset$.
 Once this is done, we can assert that $h$ crosses $B_{\bm{t}}$ (as no endpoint of $h$ falls in $B_\bm{t}$).

### 2024-06-06 msj.tex
yufei

@@
 \begin{proof}
-    Let $h$ be the full segment of $\bm{t}[k-2]$.
+    Set $h' = \bm{t}[k-2]$.
 By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ covering both endpoints of $h$.
-Thus, the effective horizontal segment of $\bm{t}$ (defined in Section~\ref{sec:hv:type2}), which is contained in $h'$, must also be covered by $r_{k-2}$.
+Thus, the effective horizontal segment of $\bm{t}$ (defined in Section~\ref{sec:hv:type2}) --- which must be contained in $h'$ --- must also be covered by $r_{k-2}$.
 Hence, $\contain_{R_{k-2}}(\bm{t})$ contains $r_{k-2}$ and thus cannot be empty.

### 2024-06-07 msj.tex
yufei

@@
     Set $h' = \bm{t}[k-2]$.
-By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ covering both endpoints of $h$.
+By the definition of a trimmed segment, there exists a rectangle $r_{k-2} \in R_{k-2}$ covering both endpoints of $h'$.
 Thus, the effective horizontal segment of $\bm{t}$ (defined in Section~\ref{sec:hv:type2}) --- which must be contained in $h'$ --- must also be covered by $r_{k-2}$.
@@
 Because$h' \cap v = h \cap v = p$, the effective horizontal segment of $\bm{t}$ is the segment $p'p$.
-To prove $r_{k-2} \in \contain_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers this segment.
+To prove $r_{k-2} \in \contain_{R_{k-2}}(\bm{t})$, we need to explain why $r_{k-2}$ covers segment $p'p$.
 This is true because:

### 2024-09-11 msj.tex
typo

@@
 The runtime  is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
-Previously, the state of the art was an algorithm with running time $O(n \log^{2k} n + \out)$.
+Previously, the state of the art was an algorithm with running time $O(n \log^{\red{2k-1}} n + \out)$.
 \end{abstract}
@@
 \extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$.
-Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ...
+Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the \red{Cartesian} product $S_1 \times S_2 \times ...
 \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$.
@@
 It is clear that there is one-one correspondence between the result of this join and that of k-SJ.
-Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time.
+Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{\red{2k-1}} n + \out)$ time.
 
@@
         $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & \\
-        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & \\
+        $\ge 3$ & \cite{kcko22} & $O(n \log^{\red{2k-1}} n + \out)$ & \\
         \hline
@@
 \myeqn{
-    B_\bm{t} &=& \bigcap_{i=1}^t r_i; \label{eqn:B_t}
+    B_\bm{t} &=& \bigcap_{i=1}^{\red{k}} r_i; \label{eqn:B_t}
 }
@@
 In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$.
-As shown in Appendix~\ref{app:bricks}, The problem can be solved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
+As shown in Appendix~\ref{app:bricks}, \red{the} problem can be solved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
 
@@
         \includegraphics[height=27mm]{./artwork/prob-b}  \\[2mm]
-        (a) Problems $\mathscr{A}$ &
+        (a) \red{Problem} $\mathscr{A}$ &
         \hspace{3mm}
@@
 Formally, if $r_1, r_2, ..., r_z$ for some $z \ge 1$ are all the rectangles in $\contain_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
-Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.
+Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: \red{r_3, r_2})}$.
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
@@
 
-\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$ (the solid rectangle), $R_2 = \set{\beta_2, \beta_2}$ (the dashed rectangles), $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
+\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$ (the solid rectangle), $R_2 = \set{\red{\beta_1}, \beta_2}$ (the dashed rectangles), $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
 The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_5)$.
@@
 Furthermore, if we are given $\contain_{R_{k-2}}(\bm{t})$ for each $\bm{t}$, statement (3) of Lemma~\ref{lmm:hv:type2:properties}
-assures us that the algorithm runs in $O(1 + |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contain_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
+assures us that the algorithm runs in $O(1 + k \cdot |\J(R_1, ..., R_{k-3}, H', V)| + k \sum_\bm{t} |\contain_{R_{k-2}}(\bm{t})|) = O(1 + k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
 

### 2024-09-19 msj.tex
yufei

@@
 }
-namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, define:
+namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, if $B_\bm{t}$ is not empty, define:
 \myitems{
@@
 
-Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ.
+Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ (note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ).
 To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time.
@@
 
-The rest of the section will focus on how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time.
+The rest of the section will focus on how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(k n \log n + k \cdot \out)$ time.
 An analogous method can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ within the same time complexity.
@@
 Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$.
-Specifically, when $\ell$ hits the bottom edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert $[x_1, x_2]$ into $\T$ and mark $r$ as active, which can be done in $O(\log n)$ time.
+Specifically, when $\ell$ hits the top edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert $[x_1, x_2]$ into $\T$ and mark $r$ as active, which can be done in $O(\log n)$ time.
 The rectangle $r$ will be referred to as the {\em host} of $[x_1, x_2]$ and is stored together with $[x_1, x_2]$ in $\T$.

### 2024-09-19 msj.tex
yufei

@@
 The runtime  is asymptotically optimal in the class of comparison-based algorithms, to which our solution belongs.
-Previously, the state of the art was an algorithm with running time $O(n \log^{\red{2k-1}} n + \out)$.
+Previously, the state of the art was an algorithm with running time $O(n \log^{2k-1} n + \out)$.
 \end{abstract}
@@
 \extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$.
-Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the \red{Cartesian} product $S_1 \times S_2 \times ...
+Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the Cartesian product $S_1 \times S_2 \times ...
 \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$.
@@
 It is clear that there is one-one correspondence between the result of this join and that of k-SJ.
-Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{\red{2k-1}} n + \out)$ time.
+Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k-1} n + \out)$ time.
 
@@
         $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & \\
-        $\ge 3$ & \cite{kcko22} & $O(n \log^{\red{2k-1}} n + \out)$ & \\
+        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k-1} n + \out)$ & \\
         \hline
@@
 \myeqn{
-    B_\bm{t} &=& \bigcap_{i=1}^{\red{k}} r_i; \label{eqn:B_t}
+    B_\bm{t} &=& \bigcap_{i=1}^{k} r_i; \label{eqn:B_t}
 }
@@
 In Figure~\ref{fig:guard}a, for instance, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$.
-As shown in Appendix~\ref{app:bricks}, \red{the} problem can be solved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
+As shown in Appendix~\ref{app:bricks}, the problem can be solved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.
 
@@
         \includegraphics[height=27mm]{./artwork/prob-b}  \\[2mm]
-        (a) \red{Problem} $\mathscr{A}$ &
+        (a) Problem $\mathscr{A}$ &
         \hspace{3mm}
@@
 Formally, if $r_1, r_2, ..., r_z$ for some $z \ge 1$ are all the rectangles in $\contain_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$.
-Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: \red{r_3, r_2})}$.
+Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_3, r_2)}$.
 In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
@@
 
-\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$ (the solid rectangle), $R_2 = \set{\red{\beta_1}, \beta_2}$ (the dashed rectangles), $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
+\extraspacing {\underline{\em Example 4.1.}} To illustrate our algorithm, we will utilize the running example in Figure~\ref{fig:hv:type1:ex}a, where $k = 4$, and $R_1 = \set{\alpha}$ (the solid rectangle), $R_2 = \set{\beta_1, \beta_2}$ (the dashed rectangles), $H$ $=$ $\set{h_1,$ $h_2,$ $..., h_6}$, and $V = \set{v_1, v_2, ..., v_5}$.
 The set $\J_1$ contains the following tuples: $(\alpha, \beta_2, h_2, v_3)$, $(\alpha, \beta_2, h_2, v_5)$, $(\alpha, \beta_2, h_5, v_3)$, and $(\alpha,$ $\beta_2, h_5, v_5)$.

### 2024-09-19 msj.tex
yufei

@@
 %\documentclass[sigconf]{acmart}
-\documentclass[acmsmall,review,anonymous]{acmart}
+%\documentclass[acmsmall,review,anonymous]{acmart}
+\documentclass[acmsmall]{acmart}
 % \documentclass{...}
@@
 
-\def\vgap{\vspace{1mm}}
-\def\extraspacing{\vspace{2mm} \noindent}
+\def\vgap{\vspace{0mm}}
+\def\extraspacing{\vspace{1mm} \noindent}
 \def\figcapup{\vspace{-2mm}}
@@
 %====== from acm ======
-\acmDOI{}
-\acmISBN{}
-
-\acmConference[]{...}{...}{...}
-\acmYear{...}
-\copyrightyear{}
-\acmArticle{}
-\acmPrice{}
+\setcopyright{acmlicensed}
+\acmJournal{PACMMOD}
+\acmYear{2024} \acmVolume{2} \acmNumber{5 (PODS)} \acmArticle{210} \acmMonth{11}\acmDOI{10.1145/3695828}
+
+\received{May 2024}
+\received[accepted]{August 2024}
+
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10003752.10003809</concept_id>
+       <concept_desc>Theory of computation~Design and analysis of algorithms</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Design and analysis of algorithms}
 %======================
@@
 \title{Optimal (Multiway) Spatial Joins}
-
-% \author{}
-% \affiliation{
-% 	\institution{Chinese University of Hong Kong}
-% 	\city{Hong Kong}
-% 	\country{China}}
-
-\author{}
+\titlenote{This work was supported in part by GRF projects 14207820, 14203421, and 14222822 from HKRGC.}
+
+\author{Ru Wang}
+\email{rwang21@cse.cuhk.edu.hk}
+\author{Yufei Tao}
+\email{taoyf@cse.cuhk.edu.hk}
+\affiliation{%
+  \institution{The Chinese University of Hong Kong}
+  \city{Hong Kong}
+  \state{Shatin}
+  \country{China}
+}
 
@@
 \maketitle
+
+% \begin{acks}
+%     This work was supported in part by GRF projects 14207820, 14203421, and 14222822 from HKRGC.
+% \end{acks}
 
@@
 
+% \begin{figure*}
+%     \begin{tabular}{cc}
+%         \includegraphics[height=30mm]{./artwork/prob-a} &
+%         \includegraphics[height=27mm]{./artwork/prob-b}  \\[2mm]
+%         (a) Problem $\mathscr{A}$ &
+%         \hspace{3mm}
+%         (b) Problem $\mathscr{B}$ \\[3mm]
+%     \end{tabular}
+%     %\vspace{3mm}
+%     \begin{tabular}{ccc}
+%         \includegraphics[height=30mm]{./artwork/prob-c} &
+%         \includegraphics[height=30mm]{./artwork/prob-d} &
+%         \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
+%         (c) Problem $\mathscr{C}$ &
+%         (d) Problem $\mathscr{D}$ &
+%         (e) Problem $\mathscr{E}$
+%     \end{tabular}
+%
+%     \figcapup
+%     \caption{Four geometric building brick problems}
+%     \label{fig:probs}
+%     \figcapdown
+% \end{figure*}
+
 \begin{figure*}
-    \begin{tabular}{cc}
-        \includegraphics[height=30mm]{./artwork/prob-a} &
-        \includegraphics[height=27mm]{./artwork/prob-b}  \\[2mm]
+    \begin{tabular}{ccccc}
+        \includegraphics[height=23mm]{./artwork/prob-a} &
+        \includegraphics[height=23mm]{./artwork/prob-b} &
+        \includegraphics[height=23mm]{./artwork/prob-c} &
+        \includegraphics[height=23mm]{./artwork/prob-d} &
+        \includegraphics[height=25mm]{./artwork/prob-e} \\[2mm]
         (a) Problem $\mathscr{A}$ &
-        \hspace{3mm}
-        (b) Problem $\mathscr{B}$ \\[3mm]
-    \end{tabular}
-    %\vspace{3mm}
-    \begin{tabular}{ccc}
-        \includegraphics[height=30mm]{./artwork/prob-c} &
-        \includegraphics[height=30mm]{./artwork/prob-d} &
-        \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
+        (b) Problem $\mathscr{B}$ &
         (c) Problem $\mathscr{C}$ &
@@
     \end{tabular}
-
     \figcapup
-    \caption{Four geometric building brick problems}
+    \caption{Five geometric building brick problems}
     \label{fig:probs}
@@
 \myeqn{
-    \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers the left endpoint of $h$}\}. \label{eqn:hv:type2:J2}
+    \J_2  &=& \{(r_1, ..., r_{k-2}, h, v) \in \J(R_1, ..., R_{k-2}, H, V) \mid \text{$r_{k-2}$ covers the left endpoint of $h$}\}. \label{eqn:hv:type2:J2}
 }
@@
 We now complete the proof of Theorem~\ref{thm:main-recur}.
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-\bibliography{ref}
-
-\balance
 
@@
 
+\bibliographystyle{plainurl}% the mandatory bibstyle
+\bibliography{ref}
+
 %\end{sloppy}

### 2024-09-19 msj.tex
yufei

@@
 \def\figcapup{\vspace{-2mm}}
-\def\figcapdown{\vspace{-4mm}}
+\def\figcapdown{\vspace{-5mm}}
 
@@
 
-It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based $k$-SJ algorithms (see Appendix~\ref{app:comp-class} for a formal definition of such algorithms), even for $k = 2$.
-This can be established via a reduction from the {\em element distinctness} problem; see \cite{dl79}.
-
-%where we are given $n$ real values $e_1, e_2, ..., e_n$ and need to decide whether there are distinct $i, j \in [n]$ satisfying $e_i = e_j$.
-The ED problem demands $\Omega(n \log n)$ comparisons to solve \cite{dl79}.
+It is worth noting that $\Omega(n \log n)$ is a lower bound on the runtime of any comparison-based $k$-SJ algorithms
+%(see Appendix~\ref{app:comp-class} for a formal definition of such algorithms),
+even for $k = 2$. This can be established via a reduction from the {\em element distinctness} problem; see \cite{dl79}.
 
@@
 \section*{Appendix}
-
-\section{Comparison-Based $k$-SJ Algorithms} \label{app:comp-class}
-
-Fix an integer $k \ge 3$ and $k$ integers $n_1, n_2, ..., n_k$ at least 1.
+%
+% \section{Comparison-Based $k$-SJ Algorithms} \label{app:comp-class}
+%
+% Fix an integer $k \ge 3$ and $k$ integers $n_1, n_2, ..., n_k$ at least 1.
 Define $n = \sum_{i=1}^k n_i$.
@@
 We assume that each rectangle in $R_i$ ($i \in [k]$) is given a distinct ID from $[n_i]$.
-
-\vgap
-
-A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a tree $\T$ defined as follows.
+%
+% \vgap
+%
+% A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a tree $\T$ defined as follows.
 Each internal node $u$ of $\T$ has three child nodes, referred as the {\em left}, {\em middle}, and {\em right} child, respectively.
 Each node is associated with an {\em instruction} of the form $(i, \lambda, \side; i', \lambda', \side')$ where
-\myitems{
-    \item $i, i' \in [1, k]$;
-    \item $\lambda \in [n_i], \lambda' \in [n_{i'}]$;
-    \item $\side$ and $\side'$ are functions chosen from the set $\{\xleft, \xright, \ybot$, $\ytop\}$.
-}
-Each leaf is associated with an {\em result}, which is defined to be a subset of $[n_1] \times [n_2] \times ... \times [n_k]$.
-
-\vgap
-
-Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$.
+% \myitems{
+%     \item $i, i' \in [1, k]$;
+%     \item $\lambda \in [n_i], \lambda' \in [n_{i'}]$;
+%     \item $\side$ and $\side'$ are functions chosen from the set $\{\xleft, \xright, \ybot$, $\ytop\}$.
+% }
+% Each leaf is associated with an {\em result}, which is defined to be a subset of $[n_1] \times [n_2] \times ... \times [n_k]$.
+%
+% \vgap
+%
+% Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$.
 In general, at an internal node with the instruction $(i, \lambda, \side; i', \lambda', \side')$, we first identify
-\myitems{
-    \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
-    \item $r'$ as the rectangle with ID $\lambda'$ in $R_{i'}$.
-}
-Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side' = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
-\myitems{
-    \item the left child if $\side(r) < \side'(r')$;
-    \item the middle child if $\side(r) = \side'(r')$;
-    \item the right child if $\side(r) > \side'(r')$.
-}
-At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.
+% \myitems{
+%     \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
+%     \item $r'$ as the rectangle with ID $\lambda'$ in $R_{i'}$.
+% }
+% Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side' = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
+% \myitems{
+%     \item the left child if $\side(r) < \side'(r')$;
+%     \item the middle child if $\side(r) = \side'(r')$;
+%     \item the right child if $\side(r) > \side'(r')$.
+% }
+% At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.
 The algorithm is required to produce the correct output on all $(n_1, n_2, ..., n_k)$-inputs.
 Its {\em worst-case cost} is the height of $\T$.
-
-\vgap
-
-Through a reduction from element distinctness \cite{dl79}, one can show that, for $k = 2$, any comparison-based algorithm correctly solving all $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.
+%
+%
+% \vgap
+%
+% Through a reduction from element distinctness \cite{dl79}, one can show that, for $k = 2$, any comparison-based algorithm correctly solving all $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.
 
@@
 \extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Sort the set of y-coordinates defined by all the segments of $H$ and $V$.
-Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+Next, sweep a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
 At all times, maintain the set $V_\ell$ of segments in $V$ that intersect with $\ell$.
@@
 \extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Sort the set of y-coordinates defined by all the segments in $H$ and all the rectangles in $R$.
-Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+Next, sweep a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
 At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
@@
 
-Sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+Sweep a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
 At all times, we maintain the set $R_\ell$ of {\em active} rectangles in $R$ that intersect with $\ell$.
@@
 Sort the set of y-coordinates defined by all the horizontal segments and rectangles.
-Sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
+Sweep a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$.
 At all times, we maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$.
@@
     (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3 \cdot (n\log n + k \cdot\out) \nn\\
-    &\leq& ... \nn \\
-    &\leq& c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3 \cdot (n\log n + k \cdot\out) \nn \\
+    &\leq ... \leq & %\nn \\
+    c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3 \cdot (n\log n + k \cdot\out) \nn \\
     &\leq & 2c^{k-1} \cdot (k!)^3 \cdot O(n\log n + k \cdot\out) \nn

### 2024-09-19 msj.tex
yufei

@@
 We now proceed to sort $L$ in ascending order of the $\lambda$-field (which is a rectangle ID), breaking ties arbitrarily.
-Because the IDs are integers created by the algorithm, we are permitted to sort $L$ using {\em counting sort} without violating the comparison-based requirements (see Appendix~\ref{app:comp-class}).
+Because the IDs are integers created by the algorithm, we are permitted to sort $L$ using {\em counting sort} without violating the comparison-based requirements.
+%(see Appendix~\ref{app:comp-class}).
 The counting sort finishes in $O(|R| + |L|) = O(n + \out)$ time, recalling that all the IDs are in $[|R|]$.
